### 2019-05-07 lagos_depth.Rmd
initial commit

---
title: "LAGOS lake depth"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(LAGOSNE)
library(ggplot2)
library(sf)
library(dplyr)

lg       <- lagosne_load("1.087.1")
lg_depth <- dplyr::select(lg$lakes_limno, 
                          lagoslakeid, maxdepth, maxdepthsource) %>% 
  left_join(lg$locus, by = "lagoslakeid") %>%
  dplyr::filter(!is.na(maxdepth)) %>%
  coordinatize()
```

```{r viz}
hist(lg_depth$maxdepth)
knitr::kable(t(quantile(lg_depth$maxdepth, c(0, 0.05, 0.5, 0.95, 1))))

```

### 2019-05-07 lagos_depth.Rmd
add map

@@
 
-```{r message=FALSE}
+```{r data, message=FALSE}
 library(LAGOSNE)
@@
 
+ggplot() + 
+  geom_sf(data = LAGOSNE::lg_extent) +
+  geom_sf(data = dplyr::sample_frac(lg_depth, 0.2), 
+          aes(color = cut(maxdepth, round(quantile(maxdepth)))), 
+          size = 1, alpha = 0.4, shape = 20) +
+  scale_color_viridis_d(direction = -1) +
+  coord_sf(datum = NA) + theme_void() +
+  labs(color = "depth (m)")
 ```

### 2019-06-03 lagos_depth.Rmd
compute some summary stats

--- 2019-05-07 lagos_depth.Rmd
@@
   left_join(lg$locus, by = "lagoslakeid") %>%
-  dplyr::filter(!is.na(maxdepth)) %>%
+  mutate(has_limno = lagoslakeid %in% lg$epi_nutr$lagoslakeid) %>%
+  # dplyr::filter(!is.na(maxdepth)) %>%
   coordinatize()
@@
 hist(lg_depth$maxdepth)
-knitr::kable(t(quantile(lg_depth$maxdepth, c(0, 0.05, 0.5, 0.95, 1))))
+knitr::kable(t(quantile(lg_depth$maxdepth, c(0, 0.05, 0.5, 0.95, 1), 
+                        na.rm = TRUE)))
 
+paste0(round(sum(!is.na(lg_depth$maxdepth)) / nrow(lg_depth) * 100, 2),
+       "% of lakes have maxdepth values.")
+
+paste0(round(sum(!is.na(dplyr::filter(lg_depth, has_limno)$maxdepth)) / 
+  length(unique(lg$epi_nutr$lagoslakeid)) * 100, 2), 
+  "% of lakes with limno data have maxdepth values.")
+
+paste0(round(sum(dplyr::filter(lg_depth, !is.na(maxdepth))$lagoslakeid %in% 
+            lg$epi_nutr$lagoslakeid) / 
+        nrow(lg_depth) * 100, 2), 
+       "% of lakes with maxdepth values have limno data.")
+
+ggplot() + 
+  geom_sf(data = LAGOSNE::lg_extent) +
+  geom_sf(data = dplyr::sample_frac(lg_depth, 0.2), 
+          aes(color = cut(maxdepth, round(quantile(maxdepth, c(0, 0.05, 0.5, 0.95, 1), 
+                                                   na.rm = TRUE)))), 
+          size = 1, alpha = 0.4, shape = 21) +
+  scale_color_viridis_d(direction = -1) +
+  coord_sf(datum = NA) + theme_void() +
+  labs(color = "depth (m)")
 ```

### 2019-06-03 lagos_depth.Rmd
reorganize

--- 2019-05-07 lagos_depth.Rmd
@@
 
-```{r message=FALSE}
+```{r data, message=FALSE, echo=FALSE}
 library(LAGOSNE)
@@
 library(dplyr)
+library(cowplot)
+library(ggforce)
 
@@
                           lagoslakeid, maxdepth, maxdepthsource) %>% 
+  mutate(color = cut(maxdepth, round(quantile(maxdepth, c(0, 0.05, 0.5, 0.95, 1), 
+                                                   na.rm = TRUE)))) %>% 
   left_join(lg$locus, by = "lagoslakeid") %>%
-  dplyr::filter(!is.na(maxdepth)) %>%
+  mutate(has_limno = lagoslakeid %in% lg$epi_nutr$lagoslakeid) %>%
+  # dplyr::filter(!is.na(maxdepth)) %>%
   coordinatize()
@@
 
-```{r viz}
-hist(lg_depth$maxdepth)
-knitr::kable(t(quantile(lg_depth$maxdepth, c(0, 0.05, 0.5, 0.95, 1))))
+```{r summary_stats, echo = FALSE}
+paste0(round(sum(!is.na(lg_depth$maxdepth)) / nrow(lg_depth) * 100, 2),
+       "% of lakes have maxdepth values.")
+
+paste0(round(sum(dplyr::filter(lg_depth, !is.na(maxdepth))$lagoslakeid %in% 
+            lg$epi_nutr$lagoslakeid) / 
+        nrow(lg_depth) * 100, 2), 
+       "% of lakes with maxdepth values have limno data.")
+
+paste0(round(sum(!is.na(dplyr::filter(lg_depth, has_limno)$maxdepth)) / 
+  length(unique(lg$epi_nutr$lagoslakeid)) * 100, 2), 
+  "% of lakes with limno data have maxdepth values.")
+```
+
+```{r viz, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4}
+knitr::kable(t(quantile(lg_depth$maxdepth, c(0, 0.05, 0.5, 0.95, 1), 
+                        na.rm = TRUE)), caption = "Lake depth quantiles")
+knitr::kable(round(t(quantile(lg_depth$lake_area_ha, c(0, 0.05, 0.5, 0.95, 1), 
+                        na.rm = TRUE)), 2), caption = "Lake area quantiles")
+
+cowplot::plot_grid(
+  ggplot(data = dplyr::filter(lg_depth, !is.na(lake_area_ha), !is.na(maxdepth))) + 
+    geom_mark_rect(aes(fill = "", x = lake_area_ha, y = maxdepth,
+                     filter = maxdepth <= 26 & lake_area_ha <= 53), inherit.aes = FALSE) +
+  geom_point(aes(x = lake_area_ha, y = maxdepth)) +
+  theme_minimal_grid() + theme(legend.position = 0, 
+                               axis.text.x = element_text(angle = 90)) +
+  scale_x_log10(expand = c(0, 0)) + 
+    scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
+  xlab("log(Lake Area)"), 
+  ggplot() + 
+    geom_point(data = dplyr::filter(lg_depth, lake_area_ha < 53, maxdepth < 26), 
+               aes(x = lake_area_ha, y = maxdepth)) +
+    theme_minimal_grid() + 
+      xlab("Lake Area")
+)
 
 ```
+
+```{r map, echo=FALSE, message=FALSE, error = FALSE}
+ggplot() + 
+  geom_sf(data = LAGOSNE::lg_extent) +
+  geom_sf(data = dplyr::sample_frac(
+    dplyr::filter(lg_depth, !is.na(maxdepth)), 0.3), 
+          aes(color = color), 
+    size = 1, alpha = 0.4, shape = 20) +
+  # scale_color_viridis_d(direction = -1) +
+  # scale_color_brewer(palette = "Set2") +
+  scale_fill_brewer(palette = "Set2") +
+  coord_sf(datum = NA) + theme_void() +
+  labs(color = "depth (m)")
+```

### 2019-06-03 lagos_depth.Rmd
add study table

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
 paste0(round(sum(!is.na(lg_depth$maxdepth)) / nrow(lg_depth) * 100, 2),
-       "% of lakes have maxdepth values.")
+       "% (n=", sum(!is.na(lg_depth$maxdepth)), ") of lakes have maxdepth values.")
 
@@
         nrow(lg_depth) * 100, 2), 
-       "% of lakes with maxdepth values have limno data.")
+       "% (n=", sum(dplyr::filter(lg_depth, !is.na(maxdepth))$lagoslakeid %in% 
+            lg$epi_nutr$lagoslakeid), ") of lakes with maxdepth values have limno data.")
 
@@
   length(unique(lg$epi_nutr$lagoslakeid)) * 100, 2), 
-  "% of lakes with limno data have maxdepth values.")
+  "% (n=", sum(!is.na(dplyr::filter(lg_depth, has_limno)$maxdepth)), ") of lakes with limno data have maxdepth values.")
 ```
 
-```{r viz, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4}
+```{r viz, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3}
 knitr::kable(t(quantile(lg_depth$maxdepth, c(0, 0.05, 0.5, 0.95, 1), 
@@
 ```{r map, echo=FALSE, message=FALSE, error = FALSE}
+states_base <- jsta::get_intersects(jsta::usa_sf(), lg_depth, 
+                                    threshold = 1500)
+
 ggplot() + 
-  geom_sf(data = LAGOSNE::lg_extent) +
+  geom_sf(data = states_base) +
   geom_sf(data = dplyr::sample_frac(
@@
 ```
+
+```{r }
+knitr::kable(
+  data.frame(stringsAsFactors=FALSE,
+  lead_author = c("Hollister", "Heathcote", "Oliver", "Sobek", "Emmerton",
+                    "Johansson", "Neumann"),
+  year = c(2011L, 2015L, 2016L, 2011L, 2007L, 2007L, 1959L),
+  location = c("NE/NY Huc", "Boreal Canada", "LAGOSNE", "Sweden",
+                 "Canadian Arctic", "Sweden", "Global"),
+  lake_numbers = c(133L, 433L, 7347L, 6618L, 81L, 105L, 107L)
+))
+
+```

### 2019-06-10 lagos_depth.Rmd
add chunk timings

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+# https://stackoverflow.com/a/24600461/3362993
+knitr::knit_hooks$set(timeit = local({
+  now = NULL
+  function(before, options) {
+    if (before) {
+      now <<- Sys.time()
+    } else {
+      res = difftime(Sys.time(), now)
+      now <<- NULL
+      # use options$label if you want the chunk label as well
+      paste('Time for this code chunk:', as.character(round(res, 2)))
+    }
+  }})
+)
+
+knitr::opts_chunk$set(echo = FALSE, timeit = TRUE)
 ```
@@
                           lagoslakeid, maxdepth, maxdepthsource) %>% 
-  mutate(color = cut(maxdepth, round(quantile(maxdepth, c(0, 0.05, 0.5, 0.95, 1), 
+  mutate(color = cut(maxdepth, 
+                     round(quantile(maxdepth, c(0, 0.05, 0.5, 0.95, 1), 
                                     na.rm = TRUE)))) %>% 
@@
 
-```{r }
+```{r papers}
 knitr::kable(
@@
 ))
-
 ```

### 2019-06-10 data/planning.R
add gdoc planning script

@@ -1,0 +1,8 @@
+# collapse log file by state
+#
+# 1. take files with a lagoslakeid field
+# 2. check if fields contain depth from the provider
+# 2. bind rows
+# 3. select lat, lon, name, and llid columns
+
+

### 2019-08-20 lagos_depth.Rmd
stash akima tests

@@
 
-```{r summary_stats, echo = FALSE}
+```{r summary_stats, echo = FALSE, eval=FALSE}
 paste0(round(sum(!is.na(lg_depth$maxdepth)) / nrow(lg_depth) * 100, 2),
@@
                  "Canadian Arctic", "Sweden", "Global"),
-  lake_numbers = c(133L, 433L, 7347L, 6618L, 81L, 105L, 107L)
+  lake_numbers = c(133L, 433L, 7347L, 6618L, 81L, 105L, 107L), 
+  rmse_m = c(6, 1.8, 7.1, NA, NA, NA, NA)
 ))
 ```
+
+```{r akima}
+library(gfplot)
+library(dplyr)
+library(PBSdata)
+library(akima)
+
+bath        <- gfplot:::load_bath(utm_zone = 9) %>% 
+  filter(depth > 0) %>%
+  filter(X < 600) %>% filter(X > 330) %>%
+  filter(Y < 5850) %>% filter(Y > 5630)
+bath_interp <- akima::interp(
+      x = bath$X,
+      y = bath$Y,
+      z = log(bath$depth),
+      extrap = TRUE, linear = TRUE
+    )
+
+z   <- reshape2::melt(bath_interp$z)
+z$x <- bath_interp$x[z$Var1]
+z$y <- bath_interp$y[z$Var2]
+z   <- rename(z, X = x, Y = y, akima_depth = value) %>%
+      dplyr::select(-Var1, -Var2) %>%
+  filter(!is.na(akima_depth)) %>%
+  mutate(akima_depth = exp(akima_depth)) %>%
+  dplyr::select(x = X, y = Y, akima_depth)
+z   <- raster::rasterFromXYZ(z)
+
+# rasterize bath object. See - https://gis.stackexchange.com/a/79074/32531
+bath_ras <- rasterize(bath[,1:2], setValues(z, NA), 
+                      bath$depth, fun = mean)
+
+par(mfrow = c(1, 2))
+plot(bath_ras)
+plot(z)
+```

### 2019-08-23 check_progress.Rmd
check progress via Rmd

---
title: "LAGOSUS depth progress"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(dplyr)
library(googledrive)
library(sf)
library(ggplot2)
library(jsta)
library(LAGOSNE)
library(assertr)
```

```{r data}
# drive_find(type = "spreadsheet")
if(!file.exists("data/backups/depth_log_all.csv")){
  drive_download("depth_log_all", type = "csv",
                 path = "data/backups/depth_log_all.csv", overwrite = TRUE)
}

dt_raw <- read.csv("data/backups/depth_log_all.csv", stringsAsFactors = FALSE)
```

```{r clean, warning=FALSE, message=FALSE}
# convert ft to m
dt <- dt_raw %>%
  mutate_at(vars(contains("depth")), as.numeric) %>%
  mutate(max_depth_m = case_when(
    is.na(max_depth_m) & !is.na(max_depth_ft) ~ max_depth_ft * 0.3048,
    TRUE ~ max_depth_m
  )) %>%
  mutate(mean_depth_m = case_when(
    is.na(mean_depth_m) & !is.na(mean_depth_ft) ~ mean_depth_ft * 0.3048,
    TRUE ~ mean_depth_m
  )) %>%
  dplyr::select(-mean_depth_ft, -max_depth_ft) %>%
  # assign GLNC to state(s)
  coordinatize("lat", "lon") %>%
  mutate(has_max = !is.na(max_depth_m) & nchar(max_depth_m) > 0) %>%
  st_join(jsta::usa_sf()) %>%
  mutate(state.x = case_when(
    state.x == "Western States GLNC" & !is.na(state.y) ~ state.y, 
    TRUE ~ state.x)) %>%
  dplyr::select(-state.y) %>% dplyr::rename(state = state.x) %>%
  dplyr::filter(state != "Western States GLNC")

# TODO: assertr the arg below
# which(dt$mean_depth_m == dt$max_depth_m)
```

```{r analysis}
# How many states have been touched?
# Of those states, how what is the response rate for max/mean depth?

frac <- dt %>%
  st_drop_geometry() %>%
  group_by(state) %>%
  # dplyr::filter(state == "ID") %>%
  summarize_at(vars(max_depth_m:mean_depth_m),
               function(x) round(sum(!is.na(x) & nchar(x > 0)) / length(x), 2)) %>%
  data.frame() %>% setNames(c("state", "max_frac", "mean_frac"))

n <- dt %>%
  st_drop_geometry() %>%
  group_by(state) %>%
  summarize_at(vars(max_depth_m:mean_depth_m),
               function(x) sum(!is.na(x))) %>%
  data.frame() %>% setNames(c("state", "max_n", "mean_n"))

res <- left_join(frac, n, by = "state") %>%
  arrange(desc(max_frac))
```

```{r tables}
knitr::kable(res)
knitr::kable(arrange(res, desc(max_n)))
```

```{r fun_facts, warning=FALSE, eval=TRUE}
# calculate total number of lakes we're working with here
lg <- lagosne_load()

paste0(sum(res$max_n), " lakes in not NE")

paste0((sum(res$max_n) + sum(!is.na(lg$lakes_limno$maxdepth))),
       " lakes in NE and not NE")

paste0(round(sum(res$max_n) / nrow(dt) * 100, 2),
       " percent of lakes with maxdepth not in NE")

paste0(round(sum(!is.na(lg$lakes_limno$maxdepth)) / nrow(lg$lakes_limno) * 100, 2),
       " percent of lakes with maxdepth in NE")
```

```{r map}
ggplot() +
  geom_sf(data = jsta::usa_sf()) +
  geom_sf(data = dt, aes(color = has_max), size = 0.6)

# mapview::mapview(dt_sf)
```

### 2019-08-23 scripts/99_check_progress.R

@@ -1,59 +1,0 @@
-library(readxl)
-library(dplyr)
-library(googledrive)
-library(assertr)
-# drive_find(type = "spreadsheet")
-if(!file.exists("data/backups/depth_log_all.csv")){
-  drive_download("depth_log_all", type = "csv",
-                 path = "data/backups/depth_log_all.csv", overwrite = TRUE)
-}
-
-dt_raw <- read.csv("data/backups/depth_log_all.csv", stringsAsFactors = FALSE)
-
-# convert ft to m
-dt <- dt_raw %>%
-  mutate_at(vars(contains("depth")), as.numeric) %>%
-  mutate(max_depth_m = case_when(
-    is.na(max_depth_m) & !is.na(max_depth_ft) ~ max_depth_ft * 0.3048,
-    TRUE ~ max_depth_m
-  )) %>%
-  mutate(mean_depth_m = case_when(
-    is.na(mean_depth_m) & !is.na(mean_depth_ft) ~ mean_depth_ft * 0.3048,
-    TRUE ~ mean_depth_m
-  )) %>%
-  dplyr::select(-mean_depth_ft, -max_depth_ft)
-
-# TODO: assertr the arg below
-# which(dt$mean_depth_m == dt$max_depth_m)
-
-# How many states have been touched?
-# Of those states, how what is the response rate for max/mean depth?
-
-frac <- dt %>%
-  group_by(state) %>%
-  summarize_at(vars(max_depth_m:mean_depth_m),
-               function(x) round(sum(!is.na(x)) / length(x), 2)) %>%
-  data.frame() %>% setNames(c("state", "max_frac", "mean_frac"))
-
-n <- dt %>%
-  group_by(state) %>%
-  summarize_at(vars(max_depth_m:mean_depth_m),
-               function(x) sum(!is.na(x))) %>%
-  data.frame() %>% setNames(c("state", "max_n", "mean_n"))
-
-res <- left_join(frac, n, by = "state") %>%
-  arrange(desc(max_frac))
-
-jsta::pdf_table(knitr::kable(res), "frac_sorted.pdf")
-jsta::pdf_table(knitr::kable(arrange(res, desc(max_n))), "n_sorted.pdf")
-
-
-# calculate total number of lakes we're working with here
-lg <- LAGOSNE::lagosne_load()
-sum(res$max_n) + sum(!is.na(lg$lakes_limno$maxdepth))
-
[…]

### 2019-08-30 scripts/rayshader.R
passable rayshader view

@@ -4,37 +4,57 @@ library(LAGOSNEgis)
 library(dplyr)
 library(mapview)
 library(raster)
-
-# arc_raster.exe from: https://github.com/r-barnes/ArcRasterRescue
-# curl::curl_download("ftp://ftp.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_dnr/water_lake_bathymetry/fgdb_water_lake_bathymetry.zip", "data/mn_lake_bathy_gdb.zip")
-# unzip("data/mn_lake_bathy_gdb.zip", exdir = "data/mn_bathy")
-# system(paste0("arc_raster.exe data/mn_bathy/water_lake_bathymetry.gdb/"))
-# system(paste0("arc_raster.exe data/mn_bathy/water_lake_bathymetry.gdb/ 0 lake_bathymetric_elevation_model"))
+library(rayshader)
+library(elevatr)
 
 # Lake Bathymetric Digital Elevation Model (DEM): A digital elevation
 # model (DEM) representing lake bathymetry. Cell size is most often 5m,
-# although 10m cells were used for some lakes to reduce grid file size. This grid
-# contains one attribute DEPTH that represents lake depth in (negative) feet. Use
-# in combination with other Lake Bathymetric GIS products. Reclassify DEM based on
-# various depth intervals. Calculate zonal and neighborhood statistics. Derive slope
-# surface. Model depth data with other cell-based parameters (e.g., slope, vegetation,
-# substrate, chemistry) to predict habitat suitability, functional niches, etc. (Note:
-# These raster analyses require Spatial Analyst or Arc Grid.)
+# although 10m cells were used for some lakes to reduce grid file size. This
+# grid # contains one attribute DEPTH that represents lake depth in (negative)
+# feet. Use in combination with other Lake Bathymetric GIS products. Reclassify
+# DEM based on various depth intervals. Calculate zonal and neighborhood
+# statistics. Derive slope surface. Model depth data with other cell-based
+# parameters (e.g., slope, vegetation, substrate, chemistry) to predict habitat
+# suitability, functional niches, etc. (Note: These raster analyses require
+# Spatial Analyst or Arc Grid.)
[…]

### 2019-09-23 lagos_depth.Rmd
use merged dataset over lagosne

@@
 
-knitr::opts_chunk$set(echo = FALSE, timeit = TRUE)
+knitr::opts_chunk$set(echo = FALSE, timeit = TRUE, eval = FALSE)
+source("scripts/99_utils.R")
 ```
 
-```{r data, message=FALSE, echo=FALSE}
-library(LAGOSNE)
-library(ggplot2)
-library(sf)
-library(dplyr)
-library(cowplot)
-library(ggforce)
+```{r data, message=FALSE, echo=FALSE, eval=TRUE}
+dt <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
+  mutate(maxdepth_class = cut(max_depth_m, 
+                     round(quantile(max_depth_m, c(0, 0.05, 0.5, 0.95, 1), 
+                                    na.rm = TRUE)))) %>% 
+  coordinatize(latname = "lat", longname = "long")
+```
 
-lg       <- lagosne_load("1.087.1")
-lg_depth <- dplyr::select(lg$lakes_limno, 
-                          lagoslakeid, maxdepth, maxdepthsource) %>% 
-  mutate(color = cut(maxdepth, 
-                     round(quantile(maxdepth, c(0, 0.05, 0.5, 0.95, 1), 
-                                    na.rm = TRUE)))) %>% 
-  left_join(lg$locus, by = "lagoslakeid") %>%
-  mutate(has_limno = lagoslakeid %in% lg$epi_nutr$lagoslakeid) %>%
-  # dplyr::filter(!is.na(maxdepth)) %>%
-  coordinatize()
-```
+ * Number of lakes w/ max depth by state
+
+ * Fraction of lakes w/ max depth by state
+
+ * Ridgeline plot of lake depth
 

### 2019-09-24 lagos_depth.Rmd
add area and conny class

@@
 
-```{r setup, include=FALSE}
+```{r timings, include=FALSE}
 # https://stackoverflow.com/a/24600461/3362993
@@
 )
+```
 
+```{r setup, include=FALSE}
+source("scripts/99_utils.R")
 knitr::opts_chunk$set(echo = FALSE, timeit = TRUE, eval = FALSE)
-source("scripts/99_utils.R")
 ```
@@
                                     na.rm = TRUE)))) %>% 
+  # TODO: join with locus preview
   coordinatize(latname = "lat", longname = "long")
@@
  * Ridgeline plot of lake depth
-
-```{r summary_stats, echo = FALSE, eval=FALSE}
-paste0(round(sum(!is.na(lg_depth$maxdepth)) / nrow(lg_depth) * 100, 2),
-       "% (n=", sum(!is.na(lg_depth$maxdepth)), ") of lakes have maxdepth values.")
-
-paste0(round(sum(dplyr::filter(lg_depth, !is.na(maxdepth))$lagoslakeid %in% 
-            lg$epi_nutr$lagoslakeid) / 
-        nrow(lg_depth) * 100, 2), 
-       "% (n=", sum(dplyr::filter(lg_depth, !is.na(maxdepth))$lagoslakeid %in% 
-            lg$epi_nutr$lagoslakeid), ") of lakes with maxdepth values have limno data.")
-
-paste0(round(sum(!is.na(dplyr::filter(lg_depth, has_limno)$maxdepth)) / 
-  length(unique(lg$epi_nutr$lagoslakeid)) * 100, 2), 
-  "% (n=", sum(!is.na(dplyr::filter(lg_depth, has_limno)$maxdepth)), ") of lakes with limno data have maxdepth values.")
-```
 
@@
 
+```
+
+```{r summary_stats, echo = FALSE, eval=FALSE}
+paste0(round(sum(!is.na(lg_depth$maxdepth)) / nrow(lg_depth) * 100, 2),
+       "% (n=", sum(!is.na(lg_depth$maxdepth)), ") of lakes have maxdepth values.")
+
+paste0(round(sum(dplyr::filter(lg_depth, !is.na(maxdepth))$lagoslakeid %in% 
+            lg$epi_nutr$lagoslakeid) / 
+        nrow(lg_depth) * 100, 2), 
+       "% (n=", sum(dplyr::filter(lg_depth, !is.na(maxdepth))$lagoslakeid %in% 
+            lg$epi_nutr$lagoslakeid), ") of lakes with maxdepth values have limno data.")
+
+paste0(round(sum(!is.na(dplyr::filter(lg_depth, has_limno)$maxdepth)) / 
+  length(unique(lg$epi_nutr$lagoslakeid)) * 100, 2), 
+  "% (n=", sum(!is.na(dplyr::filter(lg_depth, has_limno)$maxdepth)), ") of lakes with limno data have maxdepth values.")
 ```

### 2019-09-24 scripts/00_get_lagosne.R

@@ -6,11 +6,19 @@ lg <- lagosne_load("1.087.3")
 # llid, name, legacy_name, state, max_depth_m, mean_depth_m, source, lat, long
 res <- mutate(lg$lakes_limno,
        legacy_name = NA) %>%
-  left_join(dplyr::select(lg$locus, lagoslakeid, gnis_name, state_zoneid)) %>%
+  left_join(dplyr::select(lg$locus, lagoslakeid, gnis_name, state_zoneid,
+                          lake_area_ha)) %>%
   left_join(dplyr::select(lg$state, state, state_zoneid)) %>%
+  left_join(dplyr::select(lg$lakes.geo, lagoslakeid, lakeconnection)) %>%
+  # rename to lagosus conny codes
+  mutate(lakeconnection = case_when(lakeconnection == "DR_Stream" ~ "Drainage",
+                                    lakeconnection == "DR_LakeStream" ~ "DrainageLk",
+                                    TRUE ~ lakeconnection)) %>%
 dplyr::select(llid = lagoslakeid, name = gnis_name, state,
               max_depth_m = maxdepth, mean_depth_m = meandepth,
-              source = maxdepthsource, lat = nhd_lat, long = nhd_long) %>%
+              source = maxdepthsource, lake_waterarea_ha = lake_area_ha,
+              lake_connectivity_permanent = lakeconnection,
+              lat = nhd_lat, long = nhd_long) %>%
   # TODO: eventually cross-walk lagosne llids with lagosus llids
   dplyr::mutate(llid = NA) %>%
   dplyr::filter(!is.na(max_depth_m) | !is.na(mean_depth_m))

### 2019-09-24 scripts/00_get_manual.R

@@ -52,6 +52,13 @@ raw <- raw %>%
   data.frame(stringsAsFactors = FALSE) %>%
   janitor::clean_names("snake")
 
+# ---- join_lake_area ----
+locus <- read.csv("data/00_lagosus_locus/lake_characteristics.csv",
+                  stringsAsFactors = FALSE)
+raw <- left_join(raw, dplyr::select(locus, lagoslakeid, lake_waterarea_ha,
+                                    lake_connectivity_permanent),
+                 by = c("linked_lagoslakeid" = "lagoslakeid"))
+
 # ---- graph_checks ----
 if(interactive()){
 # histograms by state
@@ -176,6 +183,7 @@ res <- raw[,!duplicated(names(raw))] %>%
     mutate(legacy_name = NA) %>%
     rename(source = url) %>%
     select(llid, name, legacy_name, state, max_depth_m, mean_depth_m, source,
+           lake_waterarea_ha, lake_connectivity_permanent,
            lat, long = lon) %>%
     dplyr::filter(!is.na(max_depth_m) | !is.na(mean_depth_m))
 

### 2019-09-24 scripts/00_get_nla.R

@@ -64,7 +64,15 @@ dt_2007 <- dt_raw_2007 %>%
 # deal with 2007/2012 dups?
 dt_2012 <- select(dt_2012, -siteid_07)
 
-# anticipated columns:
-# llid, name, legacy_name, state, max_depth_m, mean_depth_m, source, lat, long
-res <- dplyr::bind_rows(dt_2007, dt_2012)
+
+
+  # ---- join_lake_area ----
+locus <- read.csv("data/00_lagosus_locus/lake_characteristics.csv",
+                  stringsAsFactors = FALSE)
+res     <- dplyr::bind_rows(dt_2007, dt_2012) %>%
+  left_join(dplyr::select(locus, llid = lagoslakeid, lake_waterarea_ha,
+                                    lake_connectivity_permanent),
+                 by = c("llid" = "llid"))
+
+# join lake area and connectivity from locus preview
 write.csv(res, "data/00_nla/00_nla.csv", row.names = FALSE)

### 2019-09-24 scripts/02_qa.R

@@ -1,10 +1,9 @@
 source("scripts/99_utils.R")
 
-# look at distributions, outliers, and area(etc) relationships -
-#     broken down by data source
+# TODO: make sure max depth is at least equal to the maximum sample depth from limno
 
-# make sure max depth is at least equal to the maximum sample depth from limno
-
-dt <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE)
-
-
+dt <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
+  # ## max_depth_m > mean_depth_m
+  dplyr::filter(max_depth_m != mean_depth_m) %>%
+  dplyr::filter(!is.na(max_depth_m) & !is.na(mean_depth_m)) %>%
+  dplyr::filter(max_depth_m > mean_depth_m)

### 2019-09-24 lagos_depth.Rmd
use merged data for analysis

@@
 
-```{r timings, include=FALSE}
+```{r timings, include=FALSE, eval=FALSE}
 # https://stackoverflow.com/a/24600461/3362993
@@
 dt <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
-  mutate(maxdepth_class = cut(max_depth_m, 
+  mutate(max_depth_m_class = cut(max_depth_m, 
                      round(quantile(max_depth_m, c(0, 0.05, 0.5, 0.95, 1), 
                                     na.rm = TRUE)))) %>% 
-  # TODO: join with locus preview
+    mutate(area_class =
+           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 400,
+                                                          1000, 20000, Inf),
+             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
+         ) %>%
   coordinatize(latname = "lat", longname = "long")
@@
 
- * Number of lakes w/ max depth by state
-
- * Fraction of lakes w/ max depth by state
-
- * Ridgeline plot of lake depth
-
-```{r viz, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3}
-knitr::kable(t(quantile(lg_depth$maxdepth, c(0, 0.05, 0.5, 0.95, 1), 
-                        na.rm = TRUE)), caption = "Lake depth quantiles")
-knitr::kable(round(t(quantile(lg_depth$lake_area_ha, c(0, 0.05, 0.5, 0.95, 1), 
-                        na.rm = TRUE)), 2), caption = "Lake area quantiles")
-
-cowplot::plot_grid(
-  ggplot(data = dplyr::filter(lg_depth, !is.na(lake_area_ha), !is.na(maxdepth))) + 
-    geom_mark_rect(aes(fill = "", x = lake_area_ha, y = maxdepth,
-                     filter = maxdepth <= 26 & lake_area_ha <= 53), inherit.aes = FALSE) +
-  geom_point(aes(x = lake_area_ha, y = maxdepth)) +
-  theme_minimal_grid() + theme(legend.position = 0, 
-                               axis.text.x = element_text(angle = 90)) +
-  scale_x_log10(expand = c(0, 0)) + 
-    scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
-  xlab("log(Lake Area)"), 
-  ggplot() + 
-    geom_point(data = dplyr::filter(lg_depth, lake_area_ha < 53, maxdepth < 26), 
-               aes(x = lake_area_ha, y = maxdepth)) +
-    theme_minimal_grid() + 
-      xlab("Lake Area")
-)
-
-```
-
-```{r summary_stats, echo = FALSE, eval=FALSE}
-paste0(round(sum(!is.na(lg_depth$maxdepth)) / nrow(lg_depth) * 100, 2),
-       "% (n=", sum(!is.na(lg_depth$maxdepth)), ") of lakes have maxdepth values.")
-
-paste0(round(sum(dplyr::filter(lg_depth, !is.na(maxdepth))$lagoslakeid %in% 
-            lg$epi_nutr$lagoslakeid) / 
-        nrow(lg_depth) * 100, 2), 
-       "% (n=", sum(dplyr::filter(lg_depth, !is.na(maxdepth))$lagoslakeid %in% 
-            lg$epi_nutr$lagoslakeid), ") of lakes with maxdepth values have limno data.")
-
-paste0(round(sum(!is.na(dplyr::filter(lg_depth, has_limno)$maxdepth)) / 
-  length(unique(lg$epi_nutr$lagoslakeid)) * 100, 2), 
-  "% (n=", sum(!is.na(dplyr::filter(lg_depth, has_limno)$maxdepth)), ") of lakes with limno data have maxdepth values.")
-```
-
-```{r map, echo=FALSE, message=FALSE, error = FALSE}
-states_base <- jsta::get_intersects(jsta::usa_sf(), lg_depth, 
-                                    threshold = 1500)
+```{r map, echo=FALSE, message=FALSE, error = FALSE, eval=TRUE}
+states_base <- jsta::get_intersects(jsta::usa_sf(), dt, 
+                                    threshold = 1)
 
@@
   geom_sf(data = dplyr::sample_frac(
-    dplyr::filter(lg_depth, !is.na(maxdepth)), 0.3), 
-          aes(color = color), 
+    dplyr::filter(dt, !is.na(max_depth_m)), 0.3), 
+          aes(color = max_depth_m_class), 
     size = 1, alpha = 0.4, shape = 20) +
@@
   coord_sf(datum = NA) + theme_void() +
-  labs(color = "depth (m)")
+  labs(color = "max depth (m)")
 ```
+
+```{r summary_stats, echo = FALSE, eval=TRUE}
+lg <- lagosne_load("1.087.3")
+
+dt_plot <- dplyr::bind_rows(
+   read.csv("data/00_manual/00_manual.csv", stringsAsFactors = FALSE))
+
+paste0(round(sum(!is.na(dt_plot$max_depth_m)) / nrow(dt_plot) * 100, 2),
+       "% (n=", sum(!is.na(dt_plot$max_depth_m)), ") of lakes with limno data in lagosus have max_depth_m values.")
+
+dt_plot <- dplyr::bind_rows(
+   read.csv("data/00_lagosne/00_lagosne.csv", stringsAsFactors = FALSE)) %>% 
+  mutate(has_limno = llid %in% lg$epi_nutr$lagoslakeid)
+
+paste0(round(sum(!is.na(dplyr::filter(dt_plot, has_limno)$max_depth_m)) / 
+  length(unique(lg$epi_nutr$lagoslakeid)) * 100, 2), 
+  "% (n=", sum(!is.na(dplyr::filter(dt_plot, has_limno)$max_depth_m)), ") of lakes with limno data in lagosne have max_depth_m values.")
+```
+
+```{r viz, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, eval=TRUE}
+knitr::kable(t(quantile(dt$max_depth_m, c(0, 0.05, 0.5, 0.95, 1), 
+                        na.rm = TRUE)), caption = "Lake depth quantiles")
+knitr::kable(round(t(quantile(dt$lake_waterarea_ha, c(0, 0.05, 0.5, 0.95, 1), 
+                        na.rm = TRUE)), 2), caption = "Lake area quantiles")
+```
+
+## Number of lakes with max/mean depth by state
+
+```{r num_lakes, eval = TRUE, echo=FALSE, message=FALSE}
+dt_plot <- dt %>%
+  st_drop_geometry() %>%
+  dplyr::filter(!is.na(max_depth_m)) %>%
+  group_by(state) %>%
+  tally() %>%
+  mutate(n_class = smart_cut(.$n, 5, "groups", 
+                        labels = ~paste(sep = "-", round(.y[1]), round(.y[2])))) %>%
+  right_join(usa_sf()) %>%
+  st_set_geometry("geometry")
+
+ggplot() + 
+  geom_sf(data = dt_plot, aes(fill = n_class)) +
+  labs(fill = "") +
+  ggtitle("Number of lakes with max depth")
+
+dt_plot <- dt %>%
+  st_drop_geometry() %>%
+  dplyr::filter(!is.na(mean_depth_m)) %>%
+  group_by(state) %>%
+  tally() %>%
+  mutate(n_class = smart_cut(.$n, 5, "groups", 
+                        labels = ~paste(sep = "-", round(.y[1]), round(.y[2])))) %>%
+  right_join(usa_sf()) %>%
+  st_set_geometry("geometry")
+
+ggplot() + 
+  geom_sf(data = dt_plot, aes(fill = n_class)) +
+  labs(fill = "") +
+  ggtitle("Number of lakes with mean depth")
+```
+
+```{r mean_vs_max, warning=FALSE}
+ggplot(data = dt) + 
+  geom_point(aes(x = mean_depth_m, y = max_depth_m))
+```
+
+\newpage
+
+## What are reasonable cutoffs for depth prediction efforts?
+
+> Oliver used 4 ha < max_depth < 1000 ha
+
+```{r cutoffs, eval = TRUE, echo=FALSE}
+
+ dt_plot <- dplyr::bind_rows(
+   read.csv("data/00_manual/00_manual.csv", stringsAsFactors = FALSE), 
+   read.csv("data/00_lagosne/00_lagosne.csv", stringsAsFactors = FALSE)) %>%
+  mutate(area_class =
+           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 400,
+                                                          1000, 20000, Inf),
+             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
+         ) %>%
+  drop_na(area_class) %>%
+  group_by(area_class) %>%
+  mutate(prop_maxdepth = round(mean(!is.na(max_depth_m)), 2)) %>%
+  add_tally() %>%
+  distinct(area_class, prop_maxdepth, n) %>%
+  arrange(area_class)
+
+dt_plot %>%
+  ggplot() +
+  geom_col(aes(y = n, x = area_class)) +
+  geom_text(aes(y = n, x = area_class, label = prop_maxdepth),
+            vjust = -0.5, size = 4) +
+  ylim(0, 100000) + xlab("Area (ha)") +
+  ggtitle("Proportion max depth availability by \n lake area class (LAGOSNE and Manual)")
+ ```
+
+ * Ridgeline plot of lake depth
 

### 2019-09-24 scripts/00_get_lagosne.R

@@ -20,7 +20,10 @@ dplyr::select(llid = lagoslakeid, name = gnis_name, state,
               lake_connectivity_permanent = lakeconnection,
               lat = nhd_lat, long = nhd_long) %>%
   # TODO: eventually cross-walk lagosne llids with lagosus llids
-  dplyr::mutate(llid = NA) %>%
-  dplyr::filter(!is.na(max_depth_m) | !is.na(mean_depth_m))
+  # for now: limit results to those with WQ data in epi_nutr
+  # dplyr::filter(llid %in% lg$epi_nutr$lagoslakeid) %>%
+  dplyr::mutate(# llid = NA,
+                effort = "LAGOSNE") # %>%
+  #dplyr::filter(!is.na(max_depth_m) | !is.na(mean_depth_m))
 
 write.csv(res, "data/00_lagosne/00_lagosne.csv", row.names = FALSE)

### 2019-09-24 scripts/00_get_manual.R

@@ -85,18 +85,6 @@ test <- left_join(raw, ll_locus, by = c("linked_lagoslakeid" = "lagoslakeid"))
 # labelled histogram of max depth availability by area class
 library(cutr) # devtools::install_github("moodymudskipper/cutr")
 
-thousand_k <- function(x){
-  res <- rep(NA, length(x))
-  for(i in 1:length(x)){
-      if(x[i] >= 1000 & x[i] < Inf){
-        res[i] <- paste0(substring(x[i], 1, 1), ".", substring(x[i], 2, 2), "k")
-      }else{
-        res[i] <- x[i]
-      }
-    }
-  res
-}
-
 test2 <- test %>%
   mutate(area_class =
            smart_cut(test$lake_waterarea_ha, c(1, 4, 40, 80, 400,
@@ -181,10 +169,11 @@ dplyr::filter(raw, stringr::str_detect(url, ".kdheks.gov"))
 res <- raw[,!duplicated(names(raw))] %>%
     rename(llid = linked_lagoslakeid) %>%
     mutate(legacy_name = NA) %>%
+    mutate(effort = "manual") %>%
     rename(source = url) %>%
     select(llid, name, legacy_name, state, max_depth_m, mean_depth_m, source,
            lake_waterarea_ha, lake_connectivity_permanent,
-           lat, long = lon) %>%
-    dplyr::filter(!is.na(max_depth_m) | !is.na(mean_depth_m))
+           lat, long = lon)# %>%
+    # dplyr::filter(!is.na(max_depth_m) | !is.na(mean_depth_m))
 
 write.csv(res, "data/00_manual/00_manual.csv", row.names = FALSE)

### 2019-09-24 scripts/00_get_nla.R

@@ -64,15 +64,15 @@ dt_2007 <- dt_raw_2007 %>%
 # deal with 2007/2012 dups?
 dt_2012 <- select(dt_2012, -siteid_07)
 
-
-
-  # ---- join_lake_area ----
+# ---- join_lake_area ----
 locus <- read.csv("data/00_lagosus_locus/lake_characteristics.csv",
                   stringsAsFactors = FALSE)
 res     <- dplyr::bind_rows(dt_2007, dt_2012) %>%
   left_join(dplyr::select(locus, llid = lagoslakeid, lake_waterarea_ha,
                                     lake_connectivity_permanent),
                  by = c("llid" = "llid"))
+
+res <- mutate(res, effort = "NLA")
 
 # join lake area and connectivity from locus preview
 write.csv(res, "data/00_nla/00_nla.csv", row.names = FALSE)

### 2019-09-24 scripts/01_merge.R

@@ -11,6 +11,7 @@ lagosne_raw <- read.csv("data/00_lagosne/00_lagosne.csv", stringsAsFactors = FAL
 
 res <- dplyr::bind_rows(manual_raw,
                         nla_raw,
-                        lagosne_raw)
+                        lagosne_raw) %>%
+  dplyr::filter(!is.na(max_depth_m) | !is.na(mean_depth_m))
 
 write.csv(res, "data/lagosus_depth.csv", row.names = FALSE)

### 2019-09-24 scripts/99_utils.R

@@ -11,6 +11,9 @@ library(ggplot2)
 suppressMessages(library(janitor))
 library(stringr)
 library(ggforce)
+suppressMessages(library(cowplot))
+library(cutr) # devtools::install_github("moodymudskipper/cutr")
+library(broom)
 
 # ---- misc fxn ----
 # jsta::get_if_not_exists
@@ -53,3 +56,26 @@ key_state <- function(x){
   dplyr::left_join(x, key,
                    by = c("state.name"))
 }
+
+thousand_k <- function(x){
+  res <- rep(NA, length(x))
+  for(i in 1:length(x)){
+    if(x[i] >= 1000 & x[i] < Inf){
+      res[i] <- paste0(substring(x[i], 1, 1), ".", substring(x[i], 2, 2), "k")
+    }else{
+      res[i] <- x[i]
+    }
+  }
+  res
+}
+
+# jsta::usa_sf()
+usa_sf <- function(crs){
+  res <- sf::st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))
+  state_key <- data.frame(state = datasets::state.abb,
+                          ID = tolower(datasets::state.name),
+                          stringsAsFactors = FALSE)
+  res <- dplyr::left_join(res, state_key, by = "ID")
+  dplyr::filter(res, !is.na(.data$state))
+}
+

### 2019-09-26 lagos_depth.Rmd
towards ridgeline plot of lake depth

@@
 
-```{r mean_vs_max, warning=FALSE}
-ggplot(data = dt) + 
-  geom_point(aes(x = mean_depth_m, y = max_depth_m))
-```
-
 \newpage
@@
 
- * Ridgeline plot of lake depth
+```{r ridgeline} 
+# Ridgeline plot of lake depth
+library(raster)
+r <- raster(ncol = 1000, nrow = 1000)
+test <- as_Spatial(
+  dplyr::sample_frac(dplyr::select(dt, max_depth_m), 0.2))
+test <- raster::rasterize(test, r, field = "max_depth_m")
+
+n <- 1000
+vals <- 1:n
+r <- raster(ncols=36, nrows=18)
+n <- 1000
+set.seed(123)
+x <- runif(n) * 360 - 180
+y <- runif(n) * 180 - 90
+xy <- cbind(x, y)
+p <- data.frame(xy, name=vals)
+coordinates(p) <- ~x+y
+
+  raster::rasterToPoints() %>%
+  as_data_frame() %>%
+    dim()
+  rename(z = glp00g30) %>%
+  mutate(z = ifelse(z < 5000, NA, z)) %>%
+  complete(x, y) %>%
+  mutate(z = y + z/0.5e6)
+
+test %>%
+  ggplot() +
+  geom_path(data = map_data("world"), aes(long, lat, group = group), colour = "grey", alpha = 0.1) +
+  geom_polygon(data = map_data("world"), aes(long, lat, group = group), fill = "grey", alpha = 0.1) +
+  geom_line(aes(x, z, group = y), lwd = 0.2) +
+  # st_bbox(jsta::usa_sf())
+  coord_quickmap(ylim = c(25, 50),
+                 xlim = c(-125, -66)) +
+  scale_x_continuous(NULL, NULL) +
+  scale_y_continuous(NULL, NULL)
+
+```
+
+```{r mean_vs_max, warning=FALSE}
+ggplot(data = dt) + 
+  geom_point(aes(x = mean_depth_m, y = max_depth_m))
+```
 

### 2019-10-04 lagos_depth.Rmd
viz hypso collection

@@
 
-```{r timings, include=FALSE, eval=FALSE}
-# https://stackoverflow.com/a/24600461/3362993
-knitr::knit_hooks$set(timeit = local({
-  now = NULL
-  function(before, options) {
-    if (before) {
-      now <<- Sys.time()
-    } else {
-      res = difftime(Sys.time(), now)
-      now <<- NULL
-      # use options$label if you want the chunk label as well
-      paste('Time for this code chunk:', as.character(round(res, 2)))
-    }
-  }})
-)
-```
-
 ```{r setup, include=FALSE}
 source("scripts/99_utils.R")
-knitr::opts_chunk$set(echo = FALSE, timeit = TRUE, eval = FALSE)
+knitr::opts_chunk$set(echo = FALSE, eval = FALSE, message = FALSE,
+                      fig.path = "figures/", fig.width = 5, fig.height = 6,
+                      dev = c("pdf"))
 ```
@@
   coordinatize(latname = "lat", longname = "long")
+
+hypso_ct <- read.csv("data/ct_hypso.csv", stringsAsFactors = FALSE)
+hypso_mn <- read.csv("data/mn_hypso.csv", stringsAsFactors = FALSE)
 ```
@@
 
-```{r viz, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, eval=TRUE}
+```{r viz, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, eval=FALSE}
 knitr::kable(t(quantile(dt$max_depth_m, c(0, 0.05, 0.5, 0.95, 1), 
@@
   ggtitle("Number of lakes with max depth")
+```
 
+```{r num-lakes-mean, eval=TRUE}
 dt_plot <- dt %>%
@@
 
-```{r cutoffs, eval = TRUE, echo=FALSE}
+```{r cutoffs, eval = TRUE, echo=FALSE, fig.height = 4}
 
@@
   ggtitle("Proportion max depth availability by \n lake area class (LAGOSNE and Manual)")
+```
+
+```{r hypsography, eval=TRUE, fig.height = 4}
+ids <- factor(c("concave", "convex"))
+polys <- data.frame(
+  id = rep(ids, each = 3),
+  x = c(0, 75, 0, 
+        25, 100, 100), 
+  y = c(75, 0, 0, 
+        100, 100, 25))
+
+plot_grid(
+ggplot(data = hypso_ct) +
+  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
+  geom_line(aes(x = area_percent, y = depth_percent, group = id)) +
+  ylab("Depth (percent)") + xlab("Area (percent)") + 
+  ggtitle(paste0("CT", " (", length(unique(hypso_ct$id)), ") lakes")) + 
+  theme(legend.position = "bottom"), 
+ggplot(data = hypso_mn) +
+    geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
+  ylab("Depth (percent)") + xlab("Area (percent)") +
+  ggtitle(paste0("MN", " (", length(unique(hypso_mn$llid)), ") lakes")) +
+  theme(axis.title.y = element_blank()), 
+rel_widths = c(1, 0.7))
+
 ```

### 2019-10-04 scripts/00_get_hypso_ct.R

@@ -64,9 +64,10 @@ get_hypso <- function(rsub, id){
   rc <- rsub %>%
     as.data.frame() %>%
     group_by(layer) %>% tally() %>%
+    tidyr::drop_na(layer) %>%
+    arrange(desc(layer)) %>%
     mutate(n = cumsum(n)) %>%
     rename(depth = layer) %>%
-    tidyr::drop_na(depth) %>%
     mutate(area_m2 = n * res(rsub)[1] * res(rsub)[2]) %>%
     mutate(area_percent = scales::rescale(area_m2, to = c(0, 100))) %>%
     mutate(depth_percent = scales::rescale(depth, to = c(0, 100)))
@@ -94,4 +95,4 @@ if(interactive()){
 }
 
 write.csv(hypso, "data/ct_hypso.csv", row.names = FALSE)
-# test <- read.csv("data/ct_hypso.csv", stringsAsFactors = FALSE)
+# hypso <- read.csv("data/ct_hypso.csv", stringsAsFactors = FALSE)

### 2019-10-04 scripts/00_get_hypso_mn.R

@@ -40,7 +40,7 @@ get_hypso <- function(rsub){
   rc <- cut(rsub, breaks = depth_int) %>%
     as.data.frame() %>% tidyr::drop_na(layer) %>%
     group_by(layer) %>% tally() %>% cumsum() %>%
-    mutate(area_m2 = n * 5 *5) %>%
+    mutate(area_m2 = n * 5 * 5) %>%
     mutate(depth_int = rev(# add interval midpoints
       as.numeric(na.omit((depth_int + lag(depth_int))/2)) * -1)) %>%
     mutate(area_percent = scales::rescale(area_m2, to = c(0, 100))) %>%
@@ -58,10 +58,10 @@ hypso <- lapply(seq_len(length(rsubs)),
 hypso <- dplyr::bind_rows(hypso)
 
 if(interactive()){
-  ggplot(data = hypso) +
+  ggplot(data = hypso_mn) +
     geom_line(aes(x = area_percent, y = depth_percent, group = llid))
   # TODO: plot the line of an ideal cone shape
 }
 
 write.csv(hypso, "data/mn_hypso.csv", row.names = FALSE)
-# test <- read.csv("data/mn_hypso.csv", stringsAsFactors = FALSE)
+# hypso_mn <- read.csv("data/mn_hypso.csv", stringsAsFactors = FALSE)

### 2019-10-04 lagos_depth.Rmd
add abline

@@
   geom_line(aes(x = area_percent, y = depth_percent, group = id)) +
+  geom_abline(intercept = 100, slope = -1, color = "red") +
   ylab("Depth (percent)") + xlab("Area (percent)") + 

### 2019-10-07 scripts/99_utils.R
avoid raster and dplyr conflict

@@ -1,6 +1,5 @@
 # ---- load packages ----
 library(googledrive)
-suppressMessages(library(dplyr))
 suppressMessages(library(LAGOSNE))
 suppressMessages(library(sf))
 library(assertr)
@@ -18,6 +17,7 @@ suppressMessages(library(raster))
 library(LAGOSNEgis)
 suppressMessages(library(scales))
 library(elevatr)
+suppressMessages(library(dplyr))
 
 # ---- misc fxn ----
 # jsta::get_if_not_exists

### 2019-10-07 lagos_depth.Rmd
towards raster hypso panel

@@
 
-hypso_ct <- read.csv("data/ct_hypso.csv", stringsAsFactors = FALSE)
+hypso_ct <- read.csv("data/ct_hypso.csv", stringsAsFactors = FALSE) %>%
+  select(-n) %>% group_by(id) %>%  add_tally() %>% ungroup() %>%
+  dplyr::filter(n > 3) %>%
+  filter(!str_detect(id, "Cove"))
 hypso_mn <- read.csv("data/mn_hypso.csv", stringsAsFactors = FALSE)
@@
 
+```{r raster-to-profile-to-hypso}
+# 3 panel figure of a single starting with rayshader, then depth profile, then hypsography
+
+# use plotly::ggplotly to select specific lakes
+dplyr::filter(hypso_ct, id == "lake_saltonstall") # super deep
+"pickerel_lake" # convex
+"long_pond" # concave
+
+rsub1 <- raster::raster("data/ct_bathy/pickerel_lake.tif")
+rsub2 <- raster::raster("data/ct_bathy/long_pond.tif")
+
+# load hypso function(s)
+get_hypso <- function(rsub, id){
+  # rsub <- rsubs[[46]]
+
+  maxdepth <- abs(cellStats(rsub, "max"))
+
+  # calculate area of each class
+  rc <- rsub %>%
+    as.data.frame() %>%
+    group_by(layer) %>% tally() %>%
+    tidyr::drop_na(layer) %>%
+    arrange(desc(layer)) %>%
+    mutate(n_cs = cumsum(n)) %>%
+    rename(depth = layer) %>%
+    mutate(area_m2 = n_cs * res(rsub)[1] * res(rsub)[2]) %>%
+    mutate(area_percent = scales::rescale(area_m2, to = c(0, 100))) %>%
+    mutate(depth_percent = scales::rescale(depth, to = c(0, 100)))
+
+  rc
+}
+
+names(rsub1) <- "layer"
+hypso1 <- get_hypso(rsub1)
+hypso1$test <- scales::rescale(c(diff(hypso1$n_cs), 0), to = c(0, 100))
+
+ggplot(data = as.data.frame(rsub1, xy = TRUE)) + 
+  geom_raster(aes(x = x, y = y, fill = layer)) +
+  coord_quickmap()
+
+cowplot::plot_grid(
+ggplot(data = hypso1, aes(x = test, y = depth_percent)) +
+  geom_point(aes(y = rev(depth_percent))) + xlab("Area (percent)") + ylab("Depth (percent)"), 
+ggplot(data = hypso1, aes(x = area_percent, y = depth_percent)) +
+  geom_point() + xlab("Area > depth (percent)") + ylab("Depth (percent)")
+)
+
+names(rsub2) <- "layer"
+hypso2 <- get_hypso(rsub2)
+
+```
+
 ```{r ridgeline} 

### 2019-10-07 scripts/02_qa.R
show bad lagosne depths

@@ -7,3 +7,12 @@ dt <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
   dplyr::filter(max_depth_m != mean_depth_m) %>%
   dplyr::filter(!is.na(max_depth_m) & !is.na(mean_depth_m)) %>%
   dplyr::filter(max_depth_m > mean_depth_m)
+
+dt_bad <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
+  dplyr::filter(max_depth_m == mean_depth_m) %>%
+  mutate(test = urltools::domain(source))
+
+table(dt_bad$test)
+jsta::pdf_table(knitr::kable(table(dt_bad$test)))
+jsta::pdf_table(knitr::kable(t(table(dt_bad$state))))
+

### 2019-10-07 scripts/00_get_hypso_ct.R
write rsub rasters

@@ -66,9 +66,9 @@ get_hypso <- function(rsub, id){
     group_by(layer) %>% tally() %>%
     tidyr::drop_na(layer) %>%
     arrange(desc(layer)) %>%
-    mutate(n = cumsum(n)) %>%
+    mutate(n_cs = cumsum(n)) %>%
     rename(depth = layer) %>%
-    mutate(area_m2 = n * res(rsub)[1] * res(rsub)[2]) %>%
+    mutate(area_m2 = n_cs * res(rsub)[1] * res(rsub)[2]) %>%
     mutate(area_percent = scales::rescale(area_m2, to = c(0, 100))) %>%
     mutate(depth_percent = scales::rescale(depth, to = c(0, 100)))
 
@@ -81,6 +81,12 @@ rsubs <- lapply(unique(ps$WBNAME), function(x){
   get_rsub(dt)
 })
 names(rsubs) <- unique(ps$WBNAME)
+
+for (i in 1:length(rsubs)) {
+  writeRaster(rsubs[[i]],
+              paste0("data/ct_bathy/", snakecase::to_snake_case(names(rsubs[i]))),
+              format='GTiff')
+}
 
 hypso <- lapply(seq_len(length(rsubs)), function(x){
   dplyr::mutate(get_hypso(rsubs[[x]]),

### 2019-10-07 lagos_depth.Rmd
more work on raster to hypso panel

--- 2019-10-04 lagos_depth.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE, eval = FALSE, message = FALSE, 
+                      warning = FALSE,
                       fig.path = "figures/", fig.width = 5, fig.height = 6,
@@
   geom_line(aes(x = area_percent, y = depth_percent, group = id)) +
+  geom_abline(intercept = -100, slope = 1, color = "red") +
+  ylim(c(100, 0)) +
   ylab("Depth (percent)") + xlab("Area (percent)") + 
@@
     geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
+  ylim(c(100, 0)) +
   ylab("Depth (percent)") + xlab("Area (percent)") +
@@
 rel_widths = c(1, 0.7))
+
+```
+
+```{r raster-to-profile-to-hypso, eval=TRUE, fig.height = 4}
+# 3 panel figure of a single starting with rayshader, then depth profile, then hypsography
+
+# use plotly::ggplotly to select specific lakes
+# dplyr::filter(hypso_ct, id == "lake_saltonstall") # super deep
+# "pickerel_lake" # convex
+# "long_pond" # concave
+
+rsub1 <- raster::raster("data/ct_bathy/pickerel_lake.tif")
+rsub2 <- raster::raster("data/ct_bathy/long_pond.tif")
+names(rsub1) <- "layer"
+names(rsub2) <- "layer"
+
+cowplot::plot_grid( 
+ggplot(data = as.data.frame(rsub1, xy = TRUE)) +
+  geom_raster(aes(x = x, y = y, fill = layer)) +
+  coord_quickmap() +
+  scale_fill_continuous(tran = "reverse") +
+  labs(fill = "") + ggtitle("convex") +
+  theme_void(),
+ggplot(data = as.data.frame(rsub2, xy = TRUE)) +
+  geom_raster(aes(x = x, y = y, fill = layer)) +
+  coord_quickmap() +
+  scale_fill_continuous(tran = "reverse") +
+  labs(fill = "") + ggtitle("concave") +
+  theme_void()
+)
+```
+
+```{r get-hypso, eval=TRUE, fig.height = 4}
+get_hypso <- function(rsub, id){
+  # rsub <- rsubs[[46]]
+
+  maxdepth <- abs(cellStats(rsub, "max"))
+
+  # calculate area of each class
+  rc <- rsub %>%
+    as.data.frame() %>%
+    group_by(layer) %>% tally() %>%
+    tidyr::drop_na(layer) %>%
+    arrange(desc(layer)) %>%
+    mutate(n_cs = cumsum(n)) %>%
+    rename(depth = layer) %>%
+    mutate(area_m2 = n_cs * res(rsub)[1] * res(rsub)[2]) %>%
+    mutate(area_percent = scales::rescale(area_m2, to = c(0, 100))) %>%
+    mutate(depth_percent = scales::rescale(depth, to = c(0, 100))) %>%
+    mutate(n_diff = scales::rescale(c(0, diff(n_cs)), 
+                                    to = c(0, 100))) %>%
+    mutate(radius_percent = rev(scales::rescale(sqrt(area_m2 / pi)) * 100))
+
+  rc
+}
+
+hypso1 <- get_hypso(rsub1)
+hypso1$type <- "convex"
+hypso2 <- get_hypso(rsub2)
+hypso2$type <- "concave"
+hypso <- dplyr::rbind_all(list(hypso1, hypso2))
+
+# par(mfrow = c(1, 2))
+# plot(hypso1$radius_percent, hypso1$depth, 
+#      ylim = rev(range(hypso1$depth)))
+# plot(hypso1$area_percent, hypso1$depth, 
+#      ylim = rev(range(hypso1$depth)))
+
+cowplot::plot_grid(
+  ggplot(data = hypso, aes(x = radius_percent, y = depth)) +
+    geom_point(aes(color = type)) + geom_line(aes(color = type)) +
+    xlab("Radius distance (percent)") + ylab("Depth") +
+    ylim(c(max(hypso$depth), 0)) +
+    theme(legend.position = "none"), 
+  ggplot(data = hypso, aes(x = area_percent, y = depth)) +
+    geom_point(aes(color = type)) + geom_line(aes(color = type)) +
+    xlab("Area > depth (percent)") + ylab("Depth") +
+    ylim(c(max(hypso$depth), 0)),
+rel_widths = c(0.6, 1)
+)
+```
+
+```{r eval=TRUE, fig.height = 4}
+cowplot::plot_grid(
+  ggplot(data = hypso, aes(x = radius_percent, y = depth_percent)) +
+    geom_point(aes(color = type)) + geom_line(aes(color = type)) +
+    xlab("Radius distance (percent)") + ylab("Depth (percent)") + ylim(c(100, 0)), 
+ggplot(data = hypso, aes(x = area_percent, y = depth_percent)) +
+  geom_point(aes(color = type)) + geom_line(aes(color = type)) +
+  xlab("Area > depth (percent)") + ylab("Depth (percent)") +
+    ylim(c(100, 0))
+)
 

### 2019-10-07 lagos_depth.Rmd
finish out panel calcs

@@
 ```{r hypsography, eval=TRUE, fig.height = 4}
-ids <- factor(c("concave", "convex"))
+ids <- factor(c("convex", "concave"))
 polys <- data.frame(
@@
   scale_fill_continuous(tran = "reverse") +
-  labs(fill = "") + ggtitle("convex") +
+  labs(fill = "") + ggtitle("concave") +
   theme_void(),
@@
   scale_fill_continuous(tran = "reverse") +
-  labs(fill = "") + ggtitle("concave") +
+  labs(fill = "") + ggtitle("convex") +
   theme_void()
@@
 get_hypso <- function(rsub, id){
-  # rsub <- rsubs[[46]]
-
   maxdepth <- abs(cellStats(rsub, "max"))
@@
                                     to = c(0, 100))) %>%
-    mutate(radius_percent = rev(scales::rescale(sqrt(area_m2 / pi)) * 100))
+    mutate(radius_percent = 
+             rev(scales::rescale(cumsum(sqrt((n * res(rsub)[1] * res(rsub)[2])/ pi)) * 100)))
 
@@
 hypso1 <- get_hypso(rsub1)
-hypso1$type <- "convex"
+hypso1$type <- "concave"
 hypso2 <- get_hypso(rsub2)
-hypso2$type <- "concave"
+hypso2$type <- "convex"
 hypso <- dplyr::rbind_all(list(hypso1, hypso2))
@@
     geom_point(aes(color = type)) + geom_line(aes(color = type)) +
-    xlab("Radius distance (percent)") + ylab("Depth") +
+    xlab("Radius distance (proportion)") + ylab("Depth") +
     ylim(c(max(hypso$depth), 0)) +

### 2019-10-07 lagos_depth.Rmd
figure arrangement

@@
     geom_point(aes(color = type)) + geom_line(aes(color = type)) +
-    xlab("Radius distance (percent)") + ylab("Depth (percent)") + ylim(c(100, 0)), 
+    xlab("Radius distance (percent)") + ylab("Depth (percent)") + 
+    ylim(c(100, 0)) +  theme(legend.position = "none"),
 ggplot(data = hypso, aes(x = area_percent, y = depth_percent)) +
@@
   xlab("Area > depth (percent)") + ylab("Depth (percent)") +
-    ylim(c(100, 0))
+    ylim(c(100, 0)),
+rel_widths = c(0.6, 1)
 )
 
+```
+
+```{r compare-slopes, eval=TRUE, fig.height = 4}
+dt <- dplyr::filter(lg$lakes_limno,
+                    !is.na(maxdepth) & !is.na(meandepth)) %>%
+  dplyr::select(lagoslakeid, nhd_lat, nhd_long, maxdepth, meandepth) %>%
+  dplyr::left_join(dplyr::select(lg$locus, lagoslakeid, lake_area_ha, gnis_name, state_zoneid)) %>%
+  left_join(dplyr::select(lg$buffer100m.lulc,
+                          buffer100m_slope_mean, lagoslakeid)) %>%
+  dplyr::left_join(dplyr::select(lg$state, state_zoneid, state)) %>%
+  dplyr::filter(lake_area_ha >= 4 & lake_area_ha <= 400) %>%
+  dplyr::filter(state == "MN") %>%
+  dplyr::filter(meandepth != maxdepth) %>%
+  dplyr::filter(meandepth > 1)
+
+# using lake areas:
+# calculate a vector of lake radii
+dt <- mutate(dt, radius = sqrt(lake_area_ha / pi))
+# TODO: instead of straight-up radii use half the fetch or half the minimum bounding width
+# Kath did geoprocessing using ArcGIS v 10.5 with tool Minimum Bounding Geometry in Features in Data Management Tools (System Toolbox)
+
+# calculate "slopes" as max_depth / r
+dt <- mutate(dt, slope = maxdepth / radius)
+
+# calculate "half-slopes" as mean_depth / (r/2)
+# how do they compare with full "slopes"?
+dt <- mutate(dt, half_slope = meandepth / (radius / 2))
+
+plot(dt$buffer100m_slope_mean, dt$slope)
+abline(0, 1)
+```
+
+```{r eval=TRUE, fig.height = 4}
+plot(dt$half_slope, dt$slope)
+abline(0, 1)
 ```

### 2019-10-07 scripts/00_get_lagosne.R
pull buffer slope

@@ -10,6 +10,8 @@ res <- mutate(lg$lakes_limno,
                           lake_area_ha)) %>%
   left_join(dplyr::select(lg$state, state, state_zoneid)) %>%
   left_join(dplyr::select(lg$lakes.geo, lagoslakeid, lakeconnection)) %>%
+  left_join(dplyr::select(lg$buffer100m.lulc,
+                          buffer100m_slope_mean, lagoslakeid)) %>%
   # rename to lagosus conny codes
   mutate(lakeconnection = case_when(lakeconnection == "DR_Stream" ~ "Drainage",
                                     lakeconnection == "DR_LakeStream" ~ "DrainageLk",
@@ -17,6 +19,7 @@ res <- mutate(lg$lakes_limno,
 dplyr::select(llid = lagoslakeid, name = gnis_name, state,
               max_depth_m = maxdepth, mean_depth_m = meandepth,
               source = maxdepthsource, lake_waterarea_ha = lake_area_ha,
+              buffer100m_slope_mean,
               lake_connectivity_permanent = lakeconnection,
               lat = nhd_lat, long = nhd_long) %>%
   # TODO: eventually cross-walk lagosne llids with lagosus llids

### 2019-10-09 lagos_depth.Rmd
produce rayshader panels

@@
 ggplot(data = hypso_mn) +
+  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
   geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
+  geom_abline(intercept = -100, slope = 1, color = "red") +
   ylim(c(100, 0)) +
@@
   ggtitle(paste0("MN", " (", length(unique(hypso_mn$llid)), ") lakes")) +
-  theme(axis.title.y = element_blank()), 
+  theme(axis.title.y = element_blank(), 
+        legend.position = "none"), 
 rel_widths = c(1, 0.7))
@@
 
-cowplot::plot_grid( 
-ggplot(data = as.data.frame(rsub1, xy = TRUE)) +
+# gg_raster <- cowplot::plot_grid( 
+gg_convave_raster <- ggplot(data = as.data.frame(rsub1, xy = TRUE)) +
   geom_raster(aes(x = x, y = y, fill = layer)) +
@@
   labs(fill = "") + ggtitle("concave") +
-  theme_void(),
-ggplot(data = as.data.frame(rsub2, xy = TRUE)) +
+  theme_void()
+
+gg_convex_raster <- ggplot(data = as.data.frame(rsub2, xy = TRUE)) +
   geom_raster(aes(x = x, y = y, fill = layer)) +
@@
   theme_void()
+```
+
+```{r gg_rayshader, eval=TRUE}
+library(rayshader)
+library(elevatr)
+
+tex <- create_texture("#AFEEEE", "#B0C4DE", "#B0E2FF", "#B2DFEE", "#BCD2EE")
+
+rsub1 <- raster::raster("data/ct_bathy/pickerel_lake.tif")
+rsub2 <- raster::raster("data/ct_bathy/long_pond.tif")
+
+if(!(file.exists("data/elevatr/pickerel_lake.gri") & 
+     file.exists("data/elevatr/long_pond.gri"))){
+  elev <- get_elev_raster(rsub1, 13)
+  raster::writeRaster(elev, paste0("data/elevatr/", names(rsub1)))
+
+  elev <- get_elev_raster(rsub2, 13)
+  raster::writeRaster(elev, paste0("data/elevatr/", names(rsub2)))
+}
+elev1 <- raster::raster("data/elevatr/pickerel_lake.gri")
+elev2 <- raster::raster("data/elevatr/long_pond.gri")
+
+prep_rayshader <- function(rsub, elev){
+  rsub <- rsub * -1
+  maxdepth <- abs(cellStats(rsub, "min"))
+
+  test <- crop(elev, rsub)
+  test <- test - abs(cellStats(test, "min")) # test should have a min of 0
+  rsub <- rsub + abs(cellStats(rsub, "max")) # rsub should have a max of 0
+
+  test  <- mask(test, test <= (cellStats(test, "min") + 1), maskvalue = TRUE)
+  test  <- resample(test, rsub)
+  test2 <- merge(rsub, test)
+  # test2[is.na(test2)] <- 0
+
+  rmat                <- matrix(test2 + maxdepth, 
+                                nrow = ncol(rsub), ncol = nrow(rsub))
+  rmat[rmat > maxdepth] <- NA
+  list(rmat = rmat, maxdepth = maxdepth, r = test2)
+}
+```
+
+```{r gg-rayshader-concave, eval=TRUE}
+
+res_rsub1 <- prep_rayshader(rsub1, elev1)
+# plot(res_rsub1$r)
+
+sphere_shade(res_rsub1$rmat, texture = tex) %>%
+  plot_3d(res_rsub1$rmat, water = TRUE, waterdepth = res_rsub1$maxdepth,
+          zscale = 0.4, solidcolor = "white", solidalpha = 0.4,
+          shadow = FALSE, solidlinecolor = "white", phi = 10, theta = -20)
+render_snapshot("figures/concave-rayshader.png", clear = TRUE)
+
+cowplot::plot_grid(
+  gg_concave_raster, 
+  ggdraw() +
+  draw_image("figures/concave-rayshader.png")
+)
+```
+
+```{r gg-rayshader-convex, eval=TRUE}
+res_rsub2 <- prep_rayshader(rsub2, elev2)
+
+sphere_shade(res_rsub2$rmat, texture = tex) %>%
+  plot_3d(res_rsub2$rmat, water = TRUE, waterdepth = res_rsub2$maxdepth,
+          zscale = 0.4, solidcolor = "white", solidalpha = 0.4,
+          shadow = FALSE, solidlinecolor = "white", phi = 10, theta = -20)
+render_snapshot("figures/convex-rayshader.png", clear = TRUE)
+
+cowplot::plot_grid(
+  gg_convex_raster, 
+  ggdraw() +
+  draw_image("figures/convex-rayshader.png")
 )
@@
 ```{r get-hypso, eval=TRUE, fig.height = 4}
+rsub1 <- raster::raster("data/ct_bathy/pickerel_lake.tif")
+rsub2 <- raster::raster("data/ct_bathy/long_pond.tif")
+names(rsub1) <- "layer"
+names(rsub2) <- "layer"
+
 get_hypso <- function(rsub, id){
@@
 
-```{r eval=TRUE, fig.height = 4}
+```{r get-hypso-normalized, eval=TRUE, fig.height = 4}
 cowplot::plot_grid(
@@
 abline(0, 1)
+```
+
+```{r }
+
 ```

### 2019-10-09 lagos_depth.Rmd
typo fix

@@
 # gg_raster <- cowplot::plot_grid( 
-gg_convave_raster <- ggplot(data = as.data.frame(rsub1, xy = TRUE)) +
+gg_concave_raster <- ggplot(data = as.data.frame(rsub1, xy = TRUE)) +
   geom_raster(aes(x = x, y = y, fill = layer)) +

### 2019-10-14 scripts/00_get_hypso_ct.R
return names and llids

@@ -114,10 +114,12 @@ rsubs <- lapply(unique(ps$lagoslakeid), function(x){
 })
 names(rsubs) <- unique(ps$lagoslakeid)
 
+if(!interactive()){
 for (i in 1:length(rsubs)) {
   writeRaster(rsubs[[i]],
               paste0("data/ct_bathy/", snakecase::to_snake_case(names(rsubs[i]))),
               format='GTiff')
+}
 }
 
 hypso <- lapply(seq_len(length(rsubs)), function(x){
@@ -125,6 +127,13 @@ hypso <- lapply(seq_len(length(rsubs)), function(x){
                 llid = names(rsubs)[x])
 })
 hypso <- dplyr::bind_rows(hypso)
+
+name_key <- distinct(st_drop_geometry(
+  dplyr::mutate(
+    dplyr::select(ps, lagoslakeid, WBNAME),
+    lagoslakeid = as.character(lagoslakeid))))
+hypso <- left_join(hypso, name_key,
+                  by = c("llid" = "lagoslakeid"))
 
 if(interactive()){
   ggplot(data = left_join(hypso,

### 2019-10-14 lagos_depth.Rmd
towards workflow fig

@@
 hypso_ct <- read.csv("data/ct_hypso.csv", stringsAsFactors = FALSE) %>%
-  select(-n) %>% group_by(id) %>%  add_tally() %>% ungroup() %>%
+  select(-n) %>% group_by(llid, WBNAME) %>%  
+  add_tally() %>% ungroup() %>%
   dplyr::filter(n > 3) %>%
-  filter(!str_detect(id, "Cove"))
+  filter(!str_detect(WBNAME, "Cove"))
 hypso_mn <- read.csv("data/mn_hypso.csv", stringsAsFactors = FALSE)
@@
 # dplyr::filter(hypso_ct, id == "lake_saltonstall") # super deep
-# "pickerel_lake" # convex
-# "long_pond" # concave
+# "pickerel_lake" 7096 # convex
+# "long_pond" 99642 # concave
 
-rsub1 <- raster::raster("data/ct_bathy/pickerel_lake.tif")
-rsub2 <- raster::raster("data/ct_bathy/long_pond.tif")
+# dplyr::filter(hypso_ct, 
+#               str_detect(WBNAME, c("Pickerel", "Long"))) %>%
+#   distinct(llid)
+
+rsub1 <- raster::raster("data/ct_bathy/7096.tif")
+rsub2 <- raster::raster("data/ct_bathy/99642.tif")
 names(rsub1) <- "layer"
@@
 ```{r get-hypso-normalized, eval=TRUE, fig.height = 4}
-cowplot::plot_grid(
-  ggplot(data = hypso, aes(x = radius_percent, y = depth_percent)) +
+gg_radius_normal <- ggplot(data = hypso, 
+                           aes(x = radius_percent, y = depth_percent)) +
     geom_point(aes(color = type)) + geom_line(aes(color = type)) +
     xlab("Radius distance (percent)") + ylab("Depth (percent)") + 
-    ylim(c(100, 0)) +  theme(legend.position = "none"),
-ggplot(data = hypso, aes(x = area_percent, y = depth_percent)) +
+    ylim(c(100, 0)) +  theme(legend.position = "none")
+
+gg_hypso_normal <- ggplot(data = hypso, aes(x = area_percent, y = depth_percent)) +
   geom_point(aes(color = type)) + geom_line(aes(color = type)) +
   xlab("Area > depth (percent)") + ylab("Depth (percent)") +
-    ylim(c(100, 0)),
+    ylim(c(100, 0))
+
+cowplot::plot_grid(
+  gg_radius_normal, gg_hypso_normal,
 rel_widths = c(0.6, 1)
@@
 
+```
+
+```{r workflow}
+gg_radius_convex <- 
+  ggplot(data = dplyr::filter(hypso, type == "convex"), 
+         aes(x = radius_percent, y = depth)) +
+    geom_point() + geom_line() +
+    xlab("Radius distance (proportion)") + ylab("Depth") +
+    scale_y_reverse()
+
+gg_hypso_convex <- ggplot(data = dplyr::filter(hypso, type == "convex"), 
+                           aes(x = area_percent, y = depth)) +
+    geom_point() + geom_line() +
+    xlab("Area > depth (percent)") + ylab("Depth") +
+  scale_y_reverse()
+
+cowplot::plot_grid(
+  gg_convex_raster, 
+  ggdraw() +
+  draw_image("figures/convex-rayshader.png"),
+  gg_radius_convex, 
+  gg_hypso_convex, 
+  ncol = 1
+)
 ```

### 2019-10-14 lagos_depth.Rmd
fig cleanup

@@
 
-plot_grid(
-ggplot(data = hypso_ct) +
+gg_hypso_ct <- ggplot(data = hypso_ct) +
   geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
-  geom_line(aes(x = area_percent, y = depth_percent, group = id)) +
+  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
   geom_abline(intercept = -100, slope = 1, color = "red") +
@@
   ylab("Depth (percent)") + xlab("Area (percent)") + 
-  ggtitle(paste0("CT", " (", length(unique(hypso_ct$id)), ") lakes")) + 
-  theme(legend.position = "bottom"), 
-ggplot(data = hypso_mn) +
+  ggtitle(paste0("CT", " (", length(unique(hypso_ct$llid)), ") lakes")) + 
+  theme(legend.position = "bottom")
+
+gg_hypso_mn <- ggplot(data = hypso_mn) +
   geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
@@
   theme(axis.title.y = element_blank(), 
-        legend.position = "none"), 
+        legend.position = "none")
+
+plot_grid(gg_hypso_ct, gg_hypso_mn, 
 rel_widths = c(1, 0.7))
+
+gg_hypso_ct + theme(plot.title = element_blank())
 
@@
 
-rsub1 <- raster::raster("data/ct_bathy/pickerel_lake.tif")
-rsub2 <- raster::raster("data/ct_bathy/long_pond.tif")
+# "pickerel_lake" 7096 # convex
+# "long_pond" 99642 # concave
+rsub1 <- raster::raster("data/ct_bathy/7096.tif")
+rsub2 <- raster::raster("data/ct_bathy/99642.tif")
 
-if(!(file.exists("data/elevatr/pickerel_lake.gri") & 
-     file.exists("data/elevatr/long_pond.gri"))){
+if(!(file.exists("data/elevatr/7096.gri") & 
+     file.exists("data/elevatr/99642.gri"))){
   elev <- get_elev_raster(rsub1, 13)
-  raster::writeRaster(elev, paste0("data/elevatr/", names(rsub1)))
+  raster::writeRaster(elev, paste0("data/elevatr/", 
+                                   gsub("X", "", names(rsub1))
+                                   ))
 
   elev <- get_elev_raster(rsub2, 13)
-  raster::writeRaster(elev, paste0("data/elevatr/", names(rsub2)))
+  raster::writeRaster(elev, paste0("data/elevatr/", 
+                                   gsub("X", "", names(rsub2))
+                                   ))
 }
-elev1 <- raster::raster("data/elevatr/pickerel_lake.gri")
-elev2 <- raster::raster("data/elevatr/long_pond.gri")
+# "pickerel_lake" 7096 # convex
+# "long_pond" 99642 # concave
+elev1 <- raster::raster("data/elevatr/7096.gri")
+elev2 <- raster::raster("data/elevatr/99642.gri")
 
@@
 render_snapshot("figures/concave-rayshader.png", clear = TRUE)
+magick::image_read("figures/concave-rayshader.png") %>%
+  magick::image_trim() %>%
+  magick::image_write("figures/concave-rayshader.png")
 
@@
 render_snapshot("figures/convex-rayshader.png", clear = TRUE)
+magick::image_read("figures/convex-rayshader.png") %>%
+  magick::image_trim() %>%
+  magick::image_write("figures/convex-rayshader.png")
 
@@
 ```{r get-hypso, eval=TRUE, fig.height = 4}
-rsub1 <- raster::raster("data/ct_bathy/pickerel_lake.tif")
-rsub2 <- raster::raster("data/ct_bathy/long_pond.tif")
+# "pickerel_lake" 7096 # convex
+# "long_pond" 99642 # concave
+rsub1 <- raster::raster("data/ct_bathy/7096.tif")
+rsub2 <- raster::raster("data/ct_bathy/99642.tif")
 names(rsub1) <- "layer"
@@
 
-get_hypso <- function(rsub, id){
+get_hypso <- function(rsub, llid){
   maxdepth <- abs(cellStats(rsub, "max"))
@@
 
-```{r workflow}
+```{r workflow, eval=TRUE}
 gg_radius_convex <- 
@@
 cowplot::plot_grid(
-  gg_convex_raster, 
+  gg_convex_raster + theme(plot.title = element_blank()), 
   ggdraw() +
@@
   gg_hypso_convex, 
-  ncol = 1
+  ncol = 2, nrow = 2
 )

### 2019-10-14 scripts/00_get_hypso_mi.R
typo fix

@@ -8,8 +8,8 @@ download.file("https://opendata.arcgis.com/datasets/d49160d2e5af4123b15d48c2e9c7
 
 lg_poly <- query_gis_(query = paste0("SELECT * FROM LAGOS_NE_All_Lakes_4ha WHERE ",
                                    paste0("State_Name LIKE '", "Michigan'", collapse = " OR ")))
-ct      <- st_read("data/mi_bathy/contours.geojson")
-ct      <- st_transform(ct, st_crs(lg_poly))
+mi      <- st_read("data/mi_bathy/contours.geojson")
+mi      <- st_transform(mi, st_crs(lg_poly))
 
 ct_large <- st_join(ct, lg_poly) %>%
   dplyr::filter(!is.na(lagoslakeid)) %>%

### 2019-10-30 lagos_depth.Rmd
classify shape classes

@@
 
-```{r hypsography, eval=TRUE, fig.height = 4}
+```{r hypsography, eval=TRUE, fig.height = 3.2}
 ids <- factor(c("convex", "concave"))
@@
   id = rep(ids, each = 3),
-  x = c(0, 75, 0, 
-        25, 100, 100), 
-  y = c(75, 0, 0, 
-        100, 100, 25))
+  x = c(0, 100, 0, 
+        0, 100, 100), 
+  y = c(100, 0, 0, 
+        100, 100, 0))
 
@@
   ggtitle(paste0("CT", " (", length(unique(hypso_ct$llid)), ") lakes")) + 
-  theme(legend.position = "bottom")
+  theme(legend.position = c(0.84, 0.2), 
+        legend.title = element_blank())
 
@@
 
-gg_hypso_ct + theme(plot.title = element_blank())
+gg_hypso_ct <- gg_hypso_ct + theme(plot.title = element_blank())
+# for poster
+ggdraw(gg_hypso_ct, ylim = c(0, 1), clip = "off") + 
+  draw_label("200 MN and CT lakes with bathymetry [1,4]", 
+             y = 0.06, x = 0.22, size = 8) +
+  draw_label("33 percent concave, 66 percent convex", 
+             y = 0.01, x = 0.2, size = 8)
+```
+
+```{r classify, eval = TRUE}
+# round the x value and test if y is greater or less than a straight-line curve
+hypso_ct_classes <- hypso_ct %>%
+  group_by(WBNAME) %>%
+  summarize_all(median) %>%
+  mutate(area_percent_rounded = round(area_percent)) %>%
+  left_join(data.frame(y = rev(seq(0, 100, by = 1)), x = seq(0, 100, by = 1)), 
+            by = c("area_percent_rounded" = "x")) %>%
+  data.frame() %>%
+  mutate(shape_class = case_when(depth_percent < y ~ "convex", 
+                                 depth_percent > y ~ "concave", 
+                                 TRUE ~ "neither")) %>%
+  ungroup() %>%   dplyr::filter(!is.na(shape_class)) %>% 
+    dplyr::select(WBNAME, llid, shape_class)
+
+# table(hypso_ct_classes$shape_class)
+
+hypso_ct_w_class <- left_join(hypso_ct_classes, hypso_ct)
+
+ggplot(data = hypso_ct_w_class) +
+  geom_line(aes(x = area_percent, y = depth_percent, group = llid, 
+                color = shape_class)) +
+  geom_abline(intercept = -100, slope = 1) +
+  ylim(c(100, 0)) +
+  ylab("Depth (percent)") + xlab("Area (percent)") + 
+  ggtitle(paste0("CT", " (", length(unique(hypso_ct$llid)), ") lakes")) + 
+  theme(legend.position = "bottom") +
+  labs(fill = "")
+```
+
+```{r class-depths, eval = TRUE}
+
+distinct(hypso_ct_w_class, llid, .keep_all = TRUE) %>%
+  dplyr::filter(shape_class != "neither") %>%
+  group_by(shape_class) %>%
+  summarize(depth = list(tibble::enframe(
+    quantile(depth, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = shape_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Depth (m)") + xlab("")
 
@@
 
-```{r raster-to-profile-to-hypso, eval=TRUE, fig.height = 4}
+```{r raster-to-profile-to-hypso, eval=FALSE, fig.height = 4}
 # 3 panel figure of a single starting with rayshader, then depth profile, then hypsography
@@
 
-```{r gg_rayshader, eval=TRUE}
+```{r gg_rayshader, eval=FALSE}
 library(rayshader)
@@
 
-```{r gg-rayshader-concave, eval=TRUE}
+```{r gg-rayshader-concave, eval=FALSE}
 
@@
 
-```{r gg-rayshader-convex, eval=TRUE}
+```{r gg-rayshader-convex, eval=FALSE}
 res_rsub2 <- prep_rayshader(rsub2, elev2)
@@
 
-```{r get-hypso, eval=TRUE, fig.height = 4}
+```{r get-hypso, eval=FALSE, fig.height = 4}
 # "pickerel_lake" 7096 # convex
@@
 
-```{r get-hypso-normalized, eval=TRUE, fig.height = 4}
+```{r get-hypso-normalized, eval=FALSE, fig.height = 4}
 gg_radius_normal <- ggplot(data = hypso, 
@@
 
-```{r workflow, eval=TRUE}
+```{r workflow, eval=FALSE}
 gg_radius_convex <- 
@@
 
-```{r compare-slopes, eval=TRUE, fig.height = 4}
+```{r compare-slopes, eval=FALSE, fig.height = 4}
 dt <- dplyr::filter(lg$lakes_limno,
@@
 
-```{r eval=TRUE, fig.height = 4}
+```{r eval=FALSE, fig.height = 4}
 plot(dt$half_slope, dt$slope)
 abline(0, 1)
-```
-
-```{r }
-
 ```

### 2019-10-31 lagos_depth.Rmd
calculate depth classes

@@
 
-```{r num_lakes, eval = TRUE, echo=FALSE, message=FALSE}
+```{r num_lakes, eval=TRUE, echo=FALSE, message=FALSE}
 dt_plot <- dt %>%
@@
 
-```{r hypsography, eval=TRUE, fig.height = 3.2}
+```{r hypsography, eval=TRUE, fig.height = 2.8}
 ids <- factor(c("convex", "concave"))
@@
   ylab("Depth (percent)") + xlab("Area (percent)") + 
-  ggtitle(paste0("CT", " (", length(unique(hypso_ct$llid)), ") lakes")) + 
+  ggtitle(paste0("CT", " (n=", length(unique(hypso_ct$llid)), ") lakes [1]")) + 
   theme(legend.position = c(0.84, 0.2), 
@@
   ylab("Depth (percent)") + xlab("Area (percent)") +
-  ggtitle(paste0("MN", " (", length(unique(hypso_mn$llid)), ") lakes")) +
+  ggtitle(paste0("MN", " (n=", length(unique(hypso_mn$llid)), ") lakes [4]")) +
   theme(axis.title.y = element_blank(), 
-        legend.position = "none")
+        axis.ticks.y = element_blank(), 
+        axis.text.y = element_blank(),
+        legend.position = c(0.72, 0.22), 
+        legend.margin = margin(0, 4, 1, 1),
+        legend.title = element_blank())
 
-plot_grid(gg_hypso_ct, gg_hypso_mn, 
-rel_widths = c(1, 0.7))
+plot_grid(gg_hypso_ct + theme(legend.position = "none"), gg_hypso_mn, 
+rel_widths = c(1, 0.88))
 
@@
 
-```{r classify, eval = TRUE}
+```{r classify, eval=TRUE}
 # round the x value and test if y is greater or less than a straight-line curve
@@
 
-```{r class-depths, eval = TRUE}
+```{r class-depths, eval=TRUE}
 
@@
 
+```
+
+```{r depth-histogram, eval=TRUE}
+# using Lauren's rounded (pretty) depth classes
+dt2 <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
+  dplyr::filter(!is.na(max_depth_m)) %>%
+  mutate(max_depth_m_class = smart_cut(max_depth_m, c(10, 30, 70, 200, Inf)))
+
+table(dt2$max_depth_m_class)
+round((table(dt2$max_depth_m_class) / nrow(dt2)) * 100, 2)
+dplyr::filter(dt2, max_depth_m > 200)
 ```

### 2019-12-24 manuscript/figures.Rmd
start figures doc

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{colortbl}
---

\arrayrulecolor{white}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
```

\pagebreak

# map

### 2019-12-26 README.Rmd
add taxonomy/schema

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

## LAGOS-US-Depth: Lake depth at continental scales

[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)]() [![Code DOI](https://zenodo.org/badge/123951266.svg)](https://zenodo.org/badge/latestdoi/123951266)

### Data (see [data/lagosus_depth.csv]())

```{r echo=FALSE}
# see scripts/99_make_taxonomy.R

dt <- read.csv("data/lagosus_depth_taxonomy.csv", stringsAsFactors = FALSE)

knitr::kable(setNames(dt, c("Variable name", "Description")))
```

### Products

Figures: [manuscript/figures.pdf](manuscript/figures.pdf)

Appendix: [manuscript/appendix.pdf](manuscript/figures.pdf)

### Abstract

Lake depth is widely regarded as a critical predictor of nutrient cycling in lakes. Unfortunately, existing databases lack widespread information on this basic measure. Lack of depth data is a key impediment to effective lake nutrient prediction in thousands of lakes at regional to continental scales. Although a number of investigators have developed predictive models of lake depth based on topography surrounding a lake, predictions from such models have proved to be fairly imprecise. To improve prediction of lake depth, we compiled a database of maximum depth for more than 15,000 lakes throughout the conterminous United States and used it to develop predictive models of lake depth based on factors such as lake elevation, topography, geometry, impoundment status, and landform. We used fitted models to predict lake depth for nearly all lakes in the conterminous United States and show how our statistical models, which are stratified on lake and watershed characteristics, represent an improvement over existing models. Ultimately, our lake depth predictions could be used not only as an aid in predicting lake nutrients but also as a rough check on estimates of total lake volume used in Earth System Models.

### 2019-12-26 scripts/99_make_taxonomy.R

@@ -1,0 +1,27 @@
+# Make taxonomy table by referencing `limno_taxonomy` file**
+
+# names(read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE))
+
+taxonomy_key <- data.frame(
+  col_name = c("lagoslakeid", "lake_namegnis", "lake_states", "lake_state",
+               "lake_lat_decdeg", "lake_lon_decdeg", "lake_maxdepth_m",
+               "lake_meandepth_m", "lake_waterarea_ha", "programtype_depth",
+               "programlink_depth", "lagos_effort", "predicted_maxdepth_m"),
+                           description = c(
+                  "unique lake identifier developed for LAGOS-US",
+                  "lake name from the gnis database",
+                  "abbreviation(s) of state(s) intersecting the lake polygon",
+                  "abbrevaiation of the state used to search for a lake's depth",
+                  "the latitude of the lake center point (NAD83).",
+                  "the longitude of the lake center point (NAD83).",
+                  "lake maximum depth in meters",
+                  "lake mean depth in meters",
+                  "surface area of lake waterbody polygon from NHD (excludes islands)",
+                  "type of program used as the source of depth data",
+                  "link or name of the program used as the source of depth data",
+                  "name of depth searching effort used for internal tracking",
+                  "lake maximum depth in meters predicted from Stachelek et. al"),
+  stringsAsFactors = FALSE)
+
+write.csv(taxonomy_key, "data/lagosus_depth_taxonomy.csv", row.names = FALSE)
+

### 2019-12-27 figures/00_maps.Rmd
start maps figure

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 11_map, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
study_bbox    <- st_read("../data/gis.gpkg", "study_bbox", quiet = TRUE)
dt            <- st_read("../data/gis.gpkg", "dt", quiet = TRUE) %>%
  mutate(maxdepth_class = smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, Inf)))
states_all    <- st_read("../data/gis.gpkg", "states_all", quiet = TRUE)
states_focal  <- st_read("../data/gis.gpkg", "states_focal", quiet = TRUE)
hu4_focal     <- st_read("../data/gis.gpkg", "hu4s_focal_simple", quiet = TRUE)
hu4_focal     <- st_crop(st_transform(
  lwgeom::st_make_valid(hu4_focal), st_crs(study_bbox)), 
  study_bbox)

# outline <- ne_countries(country = c("canada", "mexico", "united states of america"), 
#                      returnclass = "sf")
# inset_bbox <- mapedit::drawFeatures()
# sf::st_write(inset_bbox, "../data/inset_bbox.gpkg", 
#              layer_options = "OVERWRITE=YES")
# inset_bbox <- st_read("../data/inset_bbox.gpkg")
# outline <- st_crop(outline, inset_bbox)
# outline <- outline %>%
#   st_transform(LAGOSNEgis:::albers_conic()) %>%
#   st_snap(x = ., y = ., tolerance = 0.000001) %>%
#   st_union()

gg_states <- 
  ggplot() + 
    geom_sf(data = study_bbox, alpha = 1, size = 0.6, 
          color = "black", fill = "#1D9DCF") +
    geom_sf(data = states_all, fill = "gray", color = "black", size = 0.3) +
    # geom_sf(data = outline, color = "black", alpha = 0, size = 0.8) +
    # geom_sf(data = hu4, fill = "grey") +
    geom_sf(data = states_focal, fill = "white", color = "black", size = 0.4) + 
    geom_sf(data = dt, aes(color = maxdepth_class), size = 0.9) +
    # scale_fill_brewer(palette = "Set2") +
    coord_sf(datum = NA) +
    theme_void()

gg_states

# gg_hu4 <- 
#   ggplot() + 
#       geom_sf(data = study_bbox, alpha = 1, size = 0.6, 
#               color = "black", fill = "#1D9DCF") +
#       geom_sf(data = states_all, fill = "gray", color = "black", size = 0.3) +
#       geom_sf(data = hu4_focal, fill = "white", color = "black", size = 0.4) +
#       # geom_sf(data = ep, size = 1) +
#       coord_sf(datum = NA) +
#       theme_void() + 
#       theme(legend.position = 0)
# 
# cowplot::plot_grid(gg_states, gg_hu4, ncol = 2, labels = "AUTO", 
#                    vjust = 8, hjust = -1)
```

### 2019-12-27 scripts/00_get_gis.R

@@ -1,0 +1,66 @@
+source("scripts/99_utils.R")
+
+lg <- LAGOSUS::lagosus_load("locus")
+
+# ---- size_comparison ----
+gpkg_path  <- "data/gis.gpkg"
+dt         <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
+  dplyr::filter(!is.na(lake_maxdepth_m)) %>%
+  st_as_sf(coords = c("lake_lon_decdeg", "lake_lat_decdeg"), crs = 4326)
+
+states_all <- ne_states(country = c("united states of america", "canada", "mexico"),
+                        returnclass = "sf")
+
+state_codes <- unique(dt$lake_state)
+states_focal <- states_all %>%
+  dplyr::filter(., postal %in% state_codes & iso_a2 == "US")
+study_bbox   <- st_as_sfc(st_bbox(states_focal))
+states_all   <- states_all %>%
+  st_crop(study_bbox)
+
+# add zoneids
+
+# use LAGOSUS to pull hu ids that correspond to states
+# lg <- lagosne_load()
+# pad states with NA to 10 characters
+hu4_zones <- distinct(lg$locus$locus_information,
+                      lake_centroidstate, hu4_zoneid, lagoslakeid) %>%
+  dplyr::filter(lake_centroidstate %in% unique(dt$lake_state)) %>%
+  distinct(hu4_zoneid, .keep_all = TRUE) %>% arrange(hu4_zoneid)
+hu4_focal <- left_join(st_drop_geometry(dt), hu4_zones) %>%
+  distinct(hu4_zoneid) %>% dplyr::filter(!is.na(hu4_zoneid))
+
+# hu4s       <- LAGOSUSgis::query_gis("hu4", "hu4_zoneid", hu4_zones$hu4_zoneid)
+hu4s_focal <- LAGOSUSgis::query_gis("hu4", "hu4_zoneid", hu4_focal$hu4_zoneid)
+
+study_bbox   <- st_transform(st_as_sfc(st_bbox(hu4s_focal)), st_crs(study_bbox))
+
+hu4s_focal_simple <- rmapshaper::ms_simplify(
+  st_crop(lwgeom::st_make_valid(hu4s_focal),
+          st_transform(study_bbox, st_crs(hu4s_focal)))
+)
+
+# study_bbox   <- concaveman::concaveman(st_cast(hu4s_focal, "POINT"))
+
+# unlink("data/gis.gpkg")
+# st_layers("data/gis.gpkg")
[…]

### 2019-12-27 scripts/99_utils.R

@@ -22,6 +22,7 @@ library(progress)
 suppressMessages(library(LAGOSUS))
 library(polylabelr)
 library(esri2sf) # install_github("yonghah/esri2sf")
+library(rnaturalearth)
 
 # ---- misc fxn ----
 # jsta::get_if_not_exists

### 2019-12-28 figures/00_maps.Rmd
create figures pipeline

@@
 
-```{r 11_map, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r 00_map, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
 study_bbox    <- st_read("../data/gis.gpkg", "study_bbox", quiet = TRUE)

### 2019-12-28 scripts/00_get_gis.R
non geographic bbox

@@ -33,7 +33,7 @@ hu4_focal <- left_join(st_drop_geometry(dt), hu4_zones) %>%
 # hu4s       <- LAGOSUSgis::query_gis("hu4", "hu4_zoneid", hu4_zones$hu4_zoneid)
 hu4s_focal <- LAGOSUSgis::query_gis("hu4", "hu4_zoneid", hu4_focal$hu4_zoneid)
 
-study_bbox   <- st_transform(st_as_sfc(st_bbox(hu4s_focal)), st_crs(study_bbox))
+study_bbox   <- st_as_sfc(st_bbox(states_all))
 
 hu4s_focal_simple <- rmapshaper::ms_simplify(
   st_crop(lwgeom::st_make_valid(hu4s_focal),

### 2019-12-28 figures/00_maps.Rmd
cleanup

@@
 dt            <- st_read("../data/gis.gpkg", "dt", quiet = TRUE) %>%
-  mutate(maxdepth_class = smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, Inf)))
+  mutate(maxdepth_class = smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, Inf))) %>%
+  dplyr::filter(!is.na(maxdepth_class))
 states_all    <- st_read("../data/gis.gpkg", "states_all", quiet = TRUE)
@@
   ggplot() + 
-    geom_sf(data = study_bbox, alpha = 1, size = 0.6, 
+    geom_sf(data = study_bbox[2,], alpha = 1, size = 0.6, 
           color = "black", fill = "#1D9DCF") +
@@
     geom_sf(data = states_focal, fill = "white", color = "black", size = 0.4) + 
-    geom_sf(data = dt, aes(color = maxdepth_class), size = 0.9) +
+    geom_sf(data = dt, aes(color = maxdepth_class), size = 0.6, alpha = 0.8) +
+    labs(color = "Max depth (m)") +
     # scale_fill_brewer(palette = "Set2") +

### 2019-12-28 figures/00_maps.Rmd
better legend text

@@
 study_bbox    <- st_read("../data/gis.gpkg", "study_bbox", quiet = TRUE)
-dt            <- st_read("../data/gis.gpkg", "dt", quiet = TRUE) %>%
-  mutate(maxdepth_class = smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, Inf))) %>%
+dt            <- st_read("../data/gis.gpkg", "dt", quiet = TRUE)
+max_depth <- dt$lake_maxdepth_m
+dt <- dt %>%
+  mutate(lake_maxdepth_m = round(lake_maxdepth_m, 2)) %>%
+  mutate(maxdepth_class = 
+           smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, max(max_depth)), 
+                     brackets = NULL, sep = " - ")
+         ) %>% # unique(dt$maxdepth_class)
   dplyr::filter(!is.na(maxdepth_class))

### 2019-12-29 figures/00_cutoffs.Rmd
add cutoffs figure

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 00_cutoffs, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
lg <- lagosus_load("locus")
dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  dplyr::select(lagoslakeid, lake_maxdepth_m)

dt_plot <- lg$locus$locus_characteristics %>%
  left_join(dt) %>%
  mutate(area_class =
           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 300,
                                          400, 1000, 20000, Inf),
             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
         ) %>% drop_na(area_class) %>% group_by(area_class) %>%
  mutate(prop_maxdepth = round(mean(!is.na(lake_maxdepth_m)), 2)) %>%
  add_tally() %>%
  distinct(area_class, prop_maxdepth, n) %>%
  arrange(area_class)

# dt_plot %>%
#   ggplot() +
#   geom_col(aes(y = n, x = area_class)) +
#   geom_text(aes(y = n, x = area_class, label = prop_maxdepth),
#             vjust = -0.5, size = 4) +
#   ylim(0, max(dt_plot$n) + 10000) + 
#   xlab("Area (ha)") +
#   ggtitle("Proportion max depth availability by \n lake area class")

dt_plot %>%
  ggplot() +
  geom_point(aes(y = prop_maxdepth, x = area_class)) +
  # ylim(0, max(dt_plot$n) + 10000) + 
  ylim(1, 0) +
  xlab("Area (ha)") +
  ylab("Max depth availability (prop)") + 
  theme(axis.text.x = element_text(angle = 90))
```

### 2019-12-29 manuscript/figures.Rmd
create figures pipeline
fill in caption for map
add cutoffs figure

@@
 
-# map
+```{r maps, fig.cap = "Map of lake maximum depth in the continental United States.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/00_map-1.pdf")
+```
+
+```{r cutoffs, fig.cap = "Proportion max depth availability by lake area class.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/00_cutoffs-1.pdf")
+```
+
+# exploratory analyses
+
+# hypsography classification figure (panelized by state)

### 2019-12-29 figures/01_heatmap.Rmd
move heatmap to separate file

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 01_heatmap, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
library(corrr)
library(pheatmap)

dt_raw_ne <- read.csv("../data/lagosne_depth_predictors.csv", 
                      stringsAsFactors = FALSE)

pretty_names <- gsub("mbgconhull_", "", names(dt_raw_ne))
pretty_names <- gsub("lake_", "", pretty_names)
pretty_names <- gsub("_nounits", "", pretty_names)
pretty_names <- gsub("focallake", "", pretty_names)

test <- dt_raw_ne %>%
  setNames(pretty_names) %>%
  dplyr::select_if(is.numeric) %>%
  dplyr::select(-lagoslakeid, -contains("decdeg"),
                -contains("nhd"),
                -contains("lagoslakeid"),
                -huc12) %>%
  correlate() %>%
  data.frame()
test_rowname    <- test$rowname
test            <- dplyr::select(test, -rowname)
row.names(test) <- test_rowname

pheatmap(t(test),
        color = colorRampPalette(
          RColorBrewer::brewer.pal(n = 7, name = "Reds"))(100))
```

### 2019-12-29 scripts/04_depth_predict.R

@@ -49,33 +49,6 @@ write.csv(dt_raw_ne, "data/lagosne_depth_predictors.csv", row.names = FALSE)
 
 if(interactive()){
 
-# ---- correlation-heatmap ----
-
-  library(corrr)
-  library(pheatmap)
-
-  pretty_names <- gsub("mbgconhull_", "", names(dt_raw_ne))
-  pretty_names <- gsub("lake_", "", pretty_names)
-  pretty_names <- gsub("_nounits", "", pretty_names)
-  pretty_names <- gsub("focallake", "", pretty_names)
-
-  test <- dt_raw_ne %>%
-    setNames(pretty_names) %>%
-    dplyr::select_if(is.numeric) %>%
-    dplyr::select(-llid, -contains("decdeg"),
-                  -contains("nhd"),
-                  -contains("lagoslakeid"),
-                  -huc12) %>%
-    correlate() %>%
-    data.frame()
-  test_rowname    <- test$rowname
-  test            <- dplyr::select(test, -rowname)
-  row.names(test) <- test_rowname
-
-  pheatmap(t(test),
-          color = colorRampPalette(
-            RColorBrewer::brewer.pal(n = 7, name = "Reds"))(100))
-
 # ---- oliver-random-huc-slopes ----
   names(dt_raw_ne)
 

### 2019-12-29 lagos_depth.Rmd
move hypsography panel to separate file

@@
   ggtitle("Proportion max depth availability by \n lake area class (LAGOSNE and Manual)")
-```
-
-```{r hypsography, eval=TRUE, fig.height = 2.8}
-ids <- factor(c("convex", "concave"))
-polys <- data.frame(
-  id = rep(ids, each = 3),
-  x = c(0, 100, 0, 
-        0, 100, 100), 
-  y = c(100, 0, 0, 
-        100, 100, 0))
-
-gg_hypso_ct <- ggplot(data = hypso_ct) +
-  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
-  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
-  geom_abline(intercept = -100, slope = 1, color = "red") +
-  ylim(c(100, 0)) +
-  ylab("Depth (percent)") + xlab("Area (percent)") + 
-  ggtitle(paste0("CT", " (n=", length(unique(hypso_ct$llid)), ") lakes [1]")) + 
-  theme(legend.position = c(0.84, 0.2), 
-        legend.title = element_blank())
-
-gg_hypso_mn <- ggplot(data = hypso_mn) +
-  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
-  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
-  geom_abline(intercept = -100, slope = 1, color = "red") +
-  ylim(c(100, 0)) +
-  ylab("Depth (percent)") + xlab("Area (percent)") +
-  ggtitle(paste0("MN", " (n=", length(unique(hypso_mn$llid)), ") lakes [4]")) +
-  theme(axis.title.y = element_blank(), 
-        axis.ticks.y = element_blank(), 
-        axis.text.y = element_blank(),
-        legend.position = c(0.72, 0.22), 
-        legend.margin = margin(0, 4, 1, 1),
-        legend.title = element_blank())
-
-plot_grid(gg_hypso_ct + theme(legend.position = "none"), gg_hypso_mn, 
-rel_widths = c(1, 0.88))
-
-gg_hypso_ct <- gg_hypso_ct + theme(plot.title = element_blank())
-# for poster
-ggdraw(gg_hypso_ct, ylim = c(0, 1), clip = "off") + 
-  draw_label("200 MN and CT lakes with bathymetry [1,4]", 
-             y = 0.06, x = 0.22, size = 8) +
-  draw_label("33 percent concave, 66 percent convex", 
-             y = 0.01, x = 0.2, size = 8)
 ```

### 2019-12-29 figures/00_cutoffs.Rmd
move cutoffs figure to separate file

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
+knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 4, fig.height = 4)
 # setwd("figures")

### 2019-12-29 lagos_depth.Rmd
move cutoffs figure to separate file

@@
 \newpage
-
-## What are reasonable cutoffs for depth prediction efforts?
-
-> Oliver used 4 ha < max_depth < 1000 ha
-
-```{r cutoffs, eval = TRUE, echo=FALSE, fig.height = 4}
-
- dt_plot <- dplyr::bind_rows(
-   read.csv("data/00_manual/00_manual.csv", stringsAsFactors = FALSE), 
-   read.csv("data/00_lagosne/00_lagosne.csv", stringsAsFactors = FALSE)) %>%
-  mutate(area_class =
-           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 400,
-                                                          1000, 20000, Inf),
-             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
-         ) %>%
-  drop_na(area_class) %>%
-  group_by(area_class) %>%
-  mutate(prop_maxdepth = round(mean(!is.na(max_depth_m)), 2)) %>%
-  add_tally() %>%
-  distinct(area_class, prop_maxdepth, n) %>%
-  arrange(area_class)
-
-dt_plot %>%
-  ggplot() +
-  geom_col(aes(y = n, x = area_class)) +
-  geom_text(aes(y = n, x = area_class, label = prop_maxdepth),
-            vjust = -0.5, size = 4) +
-  ylim(0, 100000) + xlab("Area (ha)") +
-  ggtitle("Proportion max depth availability by \n lake area class (LAGOSNE and Manual)")
-```
 

### 2019-12-29 manuscript/figures.Rmd
fill out figure pipeline

@@
 
-# exploratory analyses
+\pagebreak
 
-# hypsography classification figure (panelized by state)
+<!-- # exploratory analyses -->
+
+```{r heatmap, fig.cap = "Heatmap caption.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_heatmap-1.pdf")
+```
+
+<!-- # hypsography classification figure (panelized by state) -->
+```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_hypsography-1.pdf")
+```

### 2020-01-01 figures/00_cutoffs.Rmd
add figure description

@@
 ```{r 00_cutoffs, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+# proportion max depth availability by lake area class
 lg <- lagosus_load("locus")

### 2020-01-01 figures/00_maps.Rmd
add proportion of lakes with max depth by state

@@
 
-# outline <- ne_countries(country = c("canada", "mexico", "united states of america"), 
-#                      returnclass = "sf")
-# inset_bbox <- mapedit::drawFeatures()
-# sf::st_write(inset_bbox, "../data/inset_bbox.gpkg", 
-#              layer_options = "OVERWRITE=YES")
-# inset_bbox <- st_read("../data/inset_bbox.gpkg")
-# outline <- st_crop(outline, inset_bbox)
-# outline <- outline %>%
-#   st_transform(LAGOSNEgis:::albers_conic()) %>%
-#   st_snap(x = ., y = ., tolerance = 0.000001) %>%
-#   st_union()
-
 gg_states <- 
@@
     geom_sf(data = states_focal, fill = "white", color = "black", size = 0.4) + 
-    geom_sf(data = dt, aes(color = maxdepth_class), size = 0.6, alpha = 0.8) +
+    geom_sf(data = dplyr::filter(dt, lake_maxdepth_m < 10), 
+            aes(color = maxdepth_class), size = 0.3, alpha = 0.8) +
+    geom_sf(data = dplyr::filter(dt, lake_maxdepth_m > 10), 
+            aes(color = maxdepth_class), size = 0.45, alpha = 0.8) +
+    # scale_color_brewer(palette = "Set1") +
     labs(color = "Max depth (m)") +
@@
 
-# gg_hu4 <- 
-#   ggplot() + 
-#       geom_sf(data = study_bbox, alpha = 1, size = 0.6, 
-#               color = "black", fill = "#1D9DCF") +
-#       geom_sf(data = states_all, fill = "gray", color = "black", size = 0.3) +
-#       geom_sf(data = hu4_focal, fill = "white", color = "black", size = 0.4) +
-#       # geom_sf(data = ep, size = 1) +
-#       coord_sf(datum = NA) +
-#       theme_void() + 
-#       theme(legend.position = 0)
-# 
-# cowplot::plot_grid(gg_states, gg_hu4, ncol = 2, labels = "AUTO", 
-#                    vjust = 8, hjust = -1)
 ```
+
+```{r 00_map_state, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+# proportion of lakes > 4ha with max depth by state
+lg <- lagosus_load("locus")
+dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)
+
+dt_plot <- dplyr::select(lg$locus$locus_information, 
+                         -contains("decdeg"), 
+                         -contains("states"), 
+                         -contains("gnis")) %>%
+  distinct(lagoslakeid, .keep_all = TRUE) %>%
+  left_join(dt, by = "lagoslakeid") %>%
+  dplyr::filter(lake_waterarea_ha >= 4) %>%
+  group_by(lake_centroidstate) %>%
+  summarize(prop_maxdepth = mean(!is.na(lake_maxdepth_m))) %>%
+  arrange(desc(prop_maxdepth)) %>%
+  left_join(dplyr::filter(
+    dplyr::select(states_all, postal, iso_a2), iso_a2 == "US"), 
+            by = c("lake_centroidstate" = "postal")) %>%
+  dplyr::filter(!is.na(prop_maxdepth))
+st_geometry(dt_plot) <- dt_plot$geom
+
+ggplot() + 
+  geom_sf(data = dt_plot, aes(fill = prop_maxdepth)) + 
+  coord_sf(datum = NA) + theme_void()
+
+```

### 2020-01-05 scripts/04_hypso_predictors.R
deal with multimatching llids

@@ -1,12 +1,14 @@
 source("scripts/99_utils.R")
 
 fname <- "data/00_hypso/hypso_predictors.csv"
+# res <- read.csv(fname, stringsAsFactors = FALSE)
 # unlink("data/00_hypso/hypso_predictors.csv")
 
 hypso_classes <- read.csv("data/00_hypso/hypso_classes.csv",
                           stringsAsFactors = FALSE)
 dt_raw <- read.csv("data/lagosus_depth_predictors.csv",
-                   stringsAsFactors = FALSE)
+                   stringsAsFactors = FALSE) %>%
+  distinct(lagoslakeid, .keep_all = TRUE)
 
 res <- left_join(hypso_classes, dt_raw,
                  by = c("llid" = "lagoslakeid"))

### 2020-01-05 figures/00_maps.Rmd
reorder legend

@@
     geom_sf(data = states_focal, fill = "white", color = "black", size = 0.4) + 
+    geom_sf(data = dt, aes(color = maxdepth_class), size = 0.3, alpha = 0) +
+    scale_color_manual(values = viridisLite::viridis(5)) +
     geom_sf(data = dplyr::filter(dt, lake_maxdepth_m < 10),
@@
     geom_sf(data = dplyr::filter(dt, lake_maxdepth_m > 10), 
-            aes(color = maxdepth_class), size = 0.45, alpha = 0.8) +
-    # scale_color_brewer(palette = "Set1") +
+            aes(color = maxdepth_class), size = 0.3, alpha = 0.8) +
     labs(color = "Max depth (m)") +
-    # scale_fill_brewer(palette = "Set2") +
     coord_sf(datum =  NA) +
@@
   geom_sf(data = dt_plot, aes(fill = prop_maxdepth)) + 
+  labs(fill = "") +
   coord_sf(datum = NA) + theme_void()

### 2020-01-06 figures/00_maps.Rmd
add proportion labels

@@
 
+# add text labels of proportions
+
 ggplot() + 
   geom_sf(data = dt_plot, aes(fill = prop_maxdepth)) + 
+  geom_sf_text(data = dt_plot, 
+               aes(label = round(prop_maxdepth, 2)), 
+               size = 2.5) +
   labs(fill = "") +

### 2020-01-06 scripts/00_get_manual.R
pull data from early collection efforts

@@ -8,14 +8,52 @@ source("scripts/99_utils.R")
 dt_raw <- suppressWarnings(suppressMessages(
   read_csv("data/00_manual/depth_log_all.csv", col_types = cols())))
 
-dt_lw <- invisible(suppressWarnings(suppressMessages(
-  get_if_not_exists(x = drive_download,
-                            destfile = "data/00_manual/lw.csv",
-                            read_function = read_csv,
-                           file = "FL_LAKEWATCH_sites_linked_R",
-                            type = "csv", path = "data/00_manual/lw.csv",
-                            overwrite = FALSE, col_types = cols())
-  )))
+# collected in very early efforts and lack a lagoslakeid
+one_off_files <- c("data/00_manual/ar.csv" = "AR_ADEQ_SWQM_sites_lakedepth",
+                   "data/00_manual/ca.csv" = "CA_USGS_WSC_sites_lakedepth",
+                   "data/00_manual/fl.csv" = "FL_LAKEWATCH_station_lakedepth")
+one_off_skips <- c(0, 0, 1)
+# lapply(seq_len(length(misc_one_offs)),
+#        function(x) drive_download(
+#          file = as.character(misc_one_offs[x]),
+#          path = names(misc_one_offs)[x],
+#          overwrite = TRUE))
+
+one_offs <- lapply(seq_len(length(one_off_files)), function(i){
+  # i <- 3
+  state <- toupper(gsub(".csv", "", basename(names(one_off_files)[i])))
+  message(paste0("dealing with one-off file from ", state))
+  res <- read.csv(names(one_off_files)[i], stringsAsFactors = FALSE,
+                    header = TRUE, skip = one_off_skips[i]) %>%
+    janitor::clean_names() %>%
+    mutate(file_name = as.character(one_off_files[i]),
+           state = state) %>%
+    select(lat = latitude_measure, lon = longitude_measure,
+           name = lake_name_google_earth, file_name, state,
+           max_depth_ft, mean_depth_ft,
+           max_depth_m, mean_depth_m, url = url_with_depth_info,
+           comments) %>%
+    mutate(lon = as.numeric(lon), lat = as.numeric(lat)) %>%
+    dplyr::filter(!is.na(lat) & !is.na(lon))
+
[…]

### 2020-01-07 manuscript/figures.Rmd
add proportion maxdepth by state fig

@@
 
-```{r maps, fig.cap = "Map of lake maximum depth in the continental United States.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r maps_dot, fig.cap = "Map of lake maximum depth in the continental United States.", warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/00_map-1.pdf")
+```
+
+```{r maps_state, fig.cap = "Proportion of lakes greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/00_map_state-1.pdf")
 ```

### 2020-01-07 figures/00_maps.Rmd
linting

--- 2020-01-05 figures/00_maps.Rmd
@@
 
+# add text labels of proportions
+
 ggplot() + 
   geom_sf(data = dt_plot, aes(fill = prop_maxdepth)) + 
+  geom_sf_text(data = dt_plot, 
+               aes(label = round(prop_maxdepth, 2)), 
+               size = 2.5) +
   labs(fill = "") +
   coord_sf(datum = NA) + theme_void()
-
 ```

### 2020-01-07 tables/01_predictors.Rmd
stash predictors table code

---
output:
  pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
# setwd("figures")
```

```{r summary_table, eval=TRUE}
library(kableExtra)
suppressMessages(library(dplyr))
library(magrittr)

# pred_key <- read.csv("../data/predictor_key.csv", 
#                      stringsAsFactors = FALSE)
# 
# # scripts/02_aggregate_predictors.R
# dt_units <- readRDS("../data/dt_units.rds")

# cdl_vars <- readRDS("../data/dt.rds") %>%
#   dplyr::select(ag:wheat) %>%
#   summarize_all(median, na.rm = TRUE) %>%
#   tidyr::gather() %>%
#   dplyr::filter(value >= 1.4) %>%
#   arrange(desc(value)) %>%
#   pull(key)

# setwd("tables")
dt <- read.csv("../data/lagosus_depth_predictors.csv", 
               stringsAsFactors = FALSE) %>%
  distinct(lagoslakeid, .keep_all = TRUE) %>%
  dplyr::select(-contains("lagoslakeid"), 
                -contains("lake_nhd"), 
                -contains("lake_huc")) %>%
  tidyr::gather(key = "variable") %>%
  group_by(variable) %>%
  mutate(value = as.numeric(value)) %>%
  mutate(q10 = quantile(value, probs = c(0.1), na.rm = TRUE),
         q25 = quantile(value, probs = c(0.25), na.rm = TRUE),
         q50 = quantile(value, probs = c(0.5), na.rm = TRUE),
         q75 = quantile(value, probs = c(0.75), na.rm = TRUE),
         q90 = quantile(value, probs = c(0.9), na.rm = TRUE), 
         ratio = q90 / q10, 
         n = sum(!is.na(value), na.rm = TRUE)) %>%
  dplyr::filter(!is.na(q50)) %>%
  dplyr::distinct(signif(q10, 2), signif(q25, 2), signif(q50, 2), signif(q75, 2), signif(q90, 2), ratio, n) %>%
  dplyr::mutate_if(is.numeric, funs(as.character(signif(., 3)))) %>%
  dplyr::select(variable, 
         q10 = `signif(q10, 2)`,
         q25 = `signif(q25, 2)`,
         q50 = `signif(q50, 2)`,
         q75 = `signif(q75, 2)`,
         q90 = `signif(q90, 2)`,
         q10_q90_ratio = ratio,
         n)#  %>%
  # left_join(dt_units, by = "variable") %>%
  # dplyr::select(variable, units, everything())

dt <- ungroup(dt)
# dt <- left_join(dt, pred_key, by = c("variable" = "var"))
# dt <- dplyr::filter(dt, !is.na(pretty))

# dt <- mutate(dt, pretty = case_when(
#   nchar(units) > 0 ~ paste0(pretty, " (", units, ")"),
#   TRUE ~ pretty), 
#   IQR = paste0(q25, " - ", q75))
# dt <- dplyr::select(dt, Variable = pretty, 
#                     Median = q50, Q25 = q25, Q75 = q75)

knitr::kable(dt, digits = 0, format = "markdown") 
#%>%
 # add_header_above(c(" ", " ", Ranges = 3, " ", " "))
```

### 2020-01-07 scripts/04_depth_predictors.R
include hypso classes in depth predictors

@@ -8,8 +8,11 @@ lg_x_walk <- lagosus_load(modules = "locus")$locus$locus_link %>%
   dplyr::select(lagosne_lagoslakeid, lagoslakeid, lagosus_centroidstate) %>%
   dplyr::rename(lagosus_lagoslakeid = lagoslakeid) %>%
   distinct(lagosus_lagoslakeid, .keep_all = TRUE)
+hypso_classes <- read.csv(
+  "data/00_hypso/hypso_classes.csv", stringsAsFactors = FALSE)
 
 dt_raw <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
+  dplyr::left_join(hypso_classes, by = c("lagoslakeid" = "llid")) %>%
   dplyr::select(-contains("waterarea")) %>%
   dplyr::filter(!is.na(lake_maxdepth_m)) %>%
   left_join(lg$locus$locus_characteristics,

### 2020-01-07 scripts/04_hypso_predictors.R

@@ -4,14 +4,9 @@ fname <- "data/00_hypso/hypso_predictors.csv"
 # res <- read.csv(fname, stringsAsFactors = FALSE)
 # unlink("data/00_hypso/hypso_predictors.csv")
 
-hypso_classes <- read.csv("data/00_hypso/hypso_classes.csv",
-                          stringsAsFactors = FALSE)
-dt_raw <- read.csv("data/lagosus_depth_predictors.csv",
+res <- read.csv("data/lagosus_depth_predictors.csv",
                    stringsAsFactors = FALSE) %>%
   distinct(lagoslakeid, .keep_all = TRUE)
-
-res <- left_join(hypso_classes, dt_raw,
-                 by = c("llid" = "lagoslakeid"))
 res$shape_class <- factor(res$shape_class)
 
 write.csv(res, fname, row.names = FALSE)

### 2020-01-07 figures/01_hypsography.Rmd
improved title

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 01_hypsography, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

ids <- factor(c("convex", "concave"))
polys <- data.frame(
  id = rep(ids, each = 3),
  x = c(0, 100, 0, 
        0, 100, 100), 
  y = c(100, 0, 0, 
        100, 100, 0))

dt_raw <- read.csv("../data/00_hypso/hypso.csv", stringsAsFactors = FALSE)

ggplot(data = dt_raw) +
  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
  geom_abline(intercept = -100, slope = 1, color = "red") +
  ylim(c(100, 0)) +
  ylab("Depth (percent)") + xlab("Area (percent)") + 
  ggtitle(paste0("Normalized hypsography for ", length(unique(dt_raw$llid)), " lakes")) + 
  theme(legend.position = c(0.84, 0.2), 
        legend.title = element_blank()) +
  facet_wrap(~state)

```

### 2020-01-07 figures/01_contrasts.Rmd
add contrasts figures

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r gg_depth_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

dt <- read.csv("../data/00_hypso/hypso_predictors.csv", 
               stringsAsFactors = FALSE)

gg_depth_shape_class <- dt %>%
  dplyr::filter(shape_class != "neither") %>%
  group_by(shape_class) %>%
  summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = shape_class, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab("")
```

```{r gg_depth_connectivity, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

gg_depth_conny_class <- dt %>%
  dplyr::filter(!is.na(lake_connectivity_class)) %>%
  group_by(lake_connectivity_class) %>%
  summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = lake_connectivity_class, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab("")
```

```{r 01_contrasts_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 4}
cowplot::plot_grid(gg_depth_shape_class + 
                     theme(axis.text.x = element_text(angle = 90)), 
                   gg_depth_conny_class + 
                     theme(# axis.text.y = element_blank(), 
                           axis.title.y = element_blank(), 
                           axis.text.x = element_text(angle = 90)))
```

```{r gg_tally_profileshapeclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# panel histograms that tally number of lakes with max depth by profileshape_class
# include sum across all types

gg_tally_profileshapeclass <- read.csv(
  "../data/00_hypso/hypso_predictors.csv", stringsAsFactors = FALSE) %>% 
  group_by(shape_class) %>%
  tally() %>%
  ggplot() + 
  geom_col(aes(x = shape_class, y = n))
```

```{r gg_tally_programtype, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# panel histograms that tally number of lakes with max depth by programtype_depth
# include sum across all types

gg_tally_programtype <- read.csv(
  "../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>% 
   dplyr::filter(!is.na(lake_maxdepth_m)) %>%
   group_by(programtype_depth) %>% 
   tally() %>% 
   ggplot(aes(x = programtype_depth, y = n)) + 
   geom_col() + 
   geom_text(aes(label = n), vjust = -0.5) + 
   ylim(0, 11000) +
  xlab("Program type")
```

```{r gg_tally_maxdepthclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# panel histograms that tally number of lakes with max depth by maxdepth_class
# include sum across all types

dt <- read.csv(
  "../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE)
max_depth <- dt$lake_maxdepth_m
gg_tally_maxdepthclass <- dt %>%
  mutate(maxdepth_class = 
           smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, max(max_depth)), 
                     brackets = NULL, sep = " - ")
         ) %>%
  group_by(maxdepth_class) %>%
  tally() %>%
  ggplot(aes(x = maxdepth_class, y = n)) +
  geom_col() + 
  geom_text(aes(label = n), vjust = -0.5) +
  ylim(0, 15000) + 
  xlab("Max depth class (m)")
```

```{r gg_tally_waterareaclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# panel histograms that tally number of lakes with max depth by waterarea_class
# include sum across all types
gg_tally_waterareaclass <- read.csv("../data/lagosus_depth_predictors.csv", 
               stringsAsFactors = FALSE) %>%
  mutate(area_class =
           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 300, 400,
                                               1000, 20000, Inf),
             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
         ) %>%
  group_by(area_class) %>%
  tally() %>%
  ggplot(aes(x = area_class, y = n)) +
  geom_col() + 
  geom_text(aes(label = n), vjust = -0.5) +
  ylim(0, 10000) + 
  xlab("Lake area class (ha)")
```

```{r gg_tally_hypsoclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_tally_hypsoclass <- read.csv(
  "../data/lagosus_depth_predictors.csv", 
               stringsAsFactors = FALSE) %>%
  dplyr::filter(!is.na(shape_class)) %>%
  distinct(lagoslakeid, shape_class, .keep_all = TRUE) %>%
  group_by(shape_class) %>%
  tally() %>%
  ggplot(aes(x = shape_class, y = n)) +
  geom_col() + 
  geom_text(aes(label = n), vjust = -0.5) +
  ylim(0, 3000) + 
  xlab("Shape class")
```

```{r 01_contrasts_tally, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 7.6}
cowplot::plot_grid(gg_tally_programtype, gg_tally_maxdepthclass, 
                   gg_tally_waterareaclass, gg_tally_hypsoclass, 
                   ncol = 1)
```

### 2020-01-07 manuscript/figures.Rmd
add contrasts figures

@@
 
-<!-- # exploratory analyses -->
-
-```{r heatmap, fig.cap = "Heatmap caption.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r heatmap, fig.cap = "Correlation matrix heatmap.", warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/01_heatmap-1.pdf")
@@
 
-<!-- # hypsography classification figure (panelized by state) -->
 ```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE}
@@
 ```
+
+```{r contrasts_depth, fig.cap = "Lake maximum depth by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_contrasts_depth-1.pdf")
+```
+
+```{r contrasts_tally, fig.cap = "Number of lakes by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_contrasts_tally-1.pdf")
+```

### 2020-01-07 figures/01_geometry.Rmd
stash hypso panel source

--- 2020-01-07 figures/01_hypsography.Rmd
@@
   ylab("Depth (percent)") + xlab("Area (percent)") + 
-  ggtitle(paste0("Normalized hypsography for ", length(unique(dt_raw$llid)), " lakes")) + 
+  ggtitle(paste0("Normalized hypsography", " (n=", length(unique(dt_raw$llid)), ") lakes")) + 
   theme(legend.position = c(0.84, 0.2), 

### 2020-01-07 manuscript/figures.Rmd
include predictors table in figures doc
add diagrams to figures doc

@@
 ```
+
+```{r 01_predictors, results='asis', warning=FALSE, echo = FALSE}
+knitr::kable('\\includegraphics[width=6.5in]{"../tables/01_predictors"}', col.names = NULL, caption = "Summary of lake depth predictor variables.")
+```
+
+![Slope diagram caption](../figures/slope_diagram.pdf){width=100%}
+
+![Lake shape caption](../figures/lake_shape.pdf){width=100%}

### 2020-01-07 tables/01_predictors.Rmd
cleanup

@@
 #   IQR = paste0(q25, " - ", q75))
-# dt <- dplyr::select(dt, Variable = pretty, 
-#                     Median = q50, Q25 = q25, Q75 = q75)
+dt <- dplyr::select(dt, variable, # Variable = pretty,
+                    Median = q50, Q25 = q25, Q75 = q75, n)
 

### 2020-01-08 figures/01_contrasts.Rmd
add glaciation contrast

@@
 
-```{r 01_contrasts_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 4}
+```{r gg_depth_glaciated, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+
+gg_depth_glaciated_class <- dt %>%
+  dplyr::filter(!is.na(lake_glaciatedlatewisc)) %>%
+  group_by(lake_glaciatedlatewisc) %>%
+  summarize(lake_maxdepth_m = list(tibble::enframe(
+    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = lake_glaciatedlatewisc, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Depth (m)") + xlab("")
+```
+
+```{r 01_contrasts_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.5}
+theme_opts <- theme(axis.text.x = element_text(angle = 90))
+
 cowplot::plot_grid(gg_depth_shape_class + 
-                     theme(axis.text.x = element_text(angle = 90)), 
+                     theme_opts, 
                    gg_depth_conny_class + 
-                     theme(# axis.text.y = element_blank(), 
-                           axis.title.y = element_blank(), 
-                           axis.text.x = element_text(angle = 90)))
+                     theme_opts +
+                     theme(axis.title.y = element_blank()), 
+                   gg_depth_glaciated_class + 
+                     theme_opts +
+                     theme(axis.title.y = element_blank()), 
+                   nrow = 1)
 ```

### 2020-01-08 figures/01_contrasts.Rmd
add programs tally and glaciated figs

@@
 ```{r gg_tally_programtype, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
-# panel histograms that tally number of lakes with max depth by programtype_depth
-# include sum across all types
-
+# tally number of lakes with max depth by programtype_depth
 gg_tally_programtype <- read.csv(
@@
 ```
+
+```{r 01_programs_tally, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+# number of programs in each type
+read.csv(
+  "../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>% 
+   dplyr::filter(!is.na(lake_maxdepth_m)) %>%
+   group_by(programtype_depth, programlink_depth) %>% 
+   tally() %>% 
+   ungroup() %>% group_by(programtype_depth) %>%
+   tally() %>%
+   ggplot(aes(x = programtype_depth, y = n)) + 
+   geom_col() + 
+   geom_text(aes(label = n), vjust = -0.5) + 
+   ylim(0, 250) +
+  xlab("Program type")
+```
+
+```{r glaciated_vs_shapeclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+dt_sub <- dplyr::filter(dt, !is.na(shape_class), shape_class != "neither")
+as.data.frame(
+  table(paste(dt_sub$lake_glaciatedlatewisc, dt_sub$shape_class))) %>%
+  tidyr::separate(Var1, c("glaciation", "shape_class"), sep = " ") %>%
+  ggplot() + 
+  geom_point(aes(glaciation, shape_class, size = Freq)) +
+  scale_size_continuous(range = c(5, 20)) +
+  theme(legend.position = "none")
+```

### 2020-01-08 figures/02_depth_model.Rmd
add figures showing initial modelling results

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 02_depth_model, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

dt <- readRDS("../data/01_depth_model/depth_model.rds")
```

```{r 02_depth_model_fitted, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

gg_noshape <- ggplot() +
  geom_point(data = dplyr::filter(dt$no_shape, shape_class == "convex"),
             aes(y = lake_maxdepth_m, x = fitted_values, color = shape_class),
             alpha = 0.3) +
  geom_point(data = dplyr::filter(dt$no_shape, shape_class == "concave"),
             aes(y = lake_maxdepth_m, x = fitted_values, color = shape_class),
             alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0) +
  ylab("Measured max depth") + xlab("Predicted max depth") +
  ylim(0, 80) + xlim(0, 80) +
  labs(color = "") +
  ggtitle("Max depth ~ Lake Area + Buffer slope")
  # theme(legend.position = "none")

gg_shape <- ggplot() +
  geom_point(data = dplyr::filter(dt$shape, shape_class == "convex"),
             aes(y = lake_maxdepth_m, x = fitted_values, color = shape_class),
             alpha = 0.3) +
  geom_point(data = dplyr::filter(dt$shape, shape_class == "concave"),
             aes(y = lake_maxdepth_m, x = fitted_values, color = shape_class),
             alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0) +
  ylab("Measured max depth") + xlab("Predicted max depth") +
  ylim(0, 80) + xlim(0, 80) +
  labs(color = "") +
  ggtitle("Max depth ~ Lake Area + Buffer slope + Shape class")

legend <- get_legend(
  # create some space to the left of the legend
  gg_noshape + theme(#legend.box.margin = margin(0, 0, 0, 12), 
                     legend.direction = "horizontal")
)

theme_opts <- theme(legend.position = "none", 
                    plot.title = element_text(size = 8))
cowplot::plot_grid(gg_noshape + theme_opts, 
                   gg_shape + theme_opts,
                   legend, 
                   rel_heights = c(1, 0.1)
                   )

```

```{r 02_depth_model_resid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# residual = observed - predicted
# positive value = underprediction
# negative value = overprediction

ggplot() +
  geom_boxplot(data = dt$no_shape, aes(y = resid, x = shape_class),
               outlier.shape = NA) +
  ylim(-20, 20)

```

### 2020-01-08 figures/02_hypso_model.Rmd
add figures showing initial modelling results

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 02_hypso_model, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

dt <- readRDS("../data/01_hypso_model/hypso_model.rds")
```

```{r 02_hypso_model_fitted, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=4}

# names(dt$ranger)
# class(dt$ranger$confusion)

dt_sub <- data.frame(dt$ranger$confusion) %>%
  mutate(goodbad = ifelse(data.frame(dt$ranger$confusion)$response == 
                      data.frame(dt$ranger$confusion)$truth, "good", "bad")) %>%
  group_by(truth) %>%
  mutate(prop = Freq/sum(Freq))

gg_ranger <- ggplot(data = dt_sub, 
                    aes(x = truth, y = response, 
                        fill = goodbad, alpha = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_manual(values = c(good = "green", bad = "red")) +
  theme_bw() + 
  ggtitle(paste0("Random Forest (acc=", 
                 round(dt$ranger$score, 2), ")"))

dt_sub <- data.frame(dt$rpart$confusion) %>%
  mutate(goodbad = ifelse(data.frame(dt$rpart$confusion)$response == 
                      data.frame(dt$rpart$confusion)$truth, "good", "bad")) %>%
  group_by(truth) %>%
  mutate(prop = Freq/sum(Freq))

gg_rpart <- ggplot(data = dt_sub, 
                    aes(x = truth, y = response, 
                        fill = goodbad, alpha = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_manual(values = c(good = "green", bad = "red")) +
  theme_bw() + 
  ggtitle(paste0("Recursive Partitioning (acc=", 
                 round(dt$rpart$score, 2), ")"))

theme_opts <- theme(legend.position = "none", 
                    plot.title = element_text(size = 8.5))
cowplot::plot_grid(gg_rpart + theme_opts, 
                   gg_ranger + theme_opts,
                   rel_heights = c(1, 0.1)
                   )

```

```{r 02_hypso_model_importance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=4}

# names(dt$rpart)
data.frame(importance = dt$rpart$importance, stringsAsFactors = FALSE) %>% 
  mutate(variable = row.names(.)) %>%
  mutate(variable = forcats::fct_reorder(variable, importance)) %>%
  arrange(desc(importance)) %>%
  ggplot() + 
  geom_col(aes(x = variable, y = importance)) +
  coord_flip() +
  ggtitle("Recursive Partitioning Importance")

```

### 2020-01-08 manuscript/figures.Rmd
cleanup
add figures showing initial modelling results

@@
 
+![Lake shape caption](../figures/lake_shape.pdf){width=100%}
+
 ```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE}
@@
 
-```{r contrasts_tally, fig.cap = "Number of lakes by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r glaciated_vs_shapeclass, fig.cap = "Cross tabulation between lake glaciation and shape class.
+Circles are proportional to the number of lakes in each category.
+If a lake has been glaciated it is more likely to be convex.
+However, if a lake has not been glaciated there is no tendencey towards a particular shape class.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/glaciated_vs_shapeclass-1.pdf")
+```
+
+```{r contrasts_tally, fig.cap = "Number of lake depth measurements by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/01_contrasts_tally-1.pdf")
+```
+
+```{r 01_programs_tally, fig.cap = "Number of programs by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_programs_tally-1.pdf")
 ```
@@
 
+```{r 02_depth_model_fitted, fig.cap = "A model that does not include shape class over predicts depth in concave lakes but does not systematically under predict depth in convex lakes.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/02_depth_model_fitted-1.pdf")
+```
+
 ![Lake shape caption](../figures/lake_shape.pdf){width=100%}
+
+```{r 02_hypso_model_fitted, fig.cap = "Confusion matrix comparing two classification methods", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/02_hypso_model_fitted-1.pdf")
+```
+
+```{r 02_hypso_model_importance, fig.cap = "Recusive Partitioning variable importance.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/02_hypso_model_importance-1.pdf")
+```

### 2020-01-08 scripts/05_depth_model.R

@@ -5,21 +5,25 @@ dt <- read.csv("data/lagosne_depth_predictors.csv",
                stringsAsFactors = FALSE) %>%
   dplyr::filter(!is.na(shape_class))
 
+res <- list()
+
 #---- Model without shape_class ----
 fit <- lm(lake_maxdepth_m ~
             buffer100m_slope_max + lake_waterarea_ha,# + shape_class,
           data = dt)
-fit_dt <- data.frame(resid = fit$residuals,
+fit_dt_noshape <- data.frame(resid = fit$residuals,
                      shape_class = dt$shape_class,
                      fitted_values = fit$fitted.values,
                      lake_maxdepth_m = dt$lake_maxdepth_m,
                      stringsAsFactors = FALSE)
 
+res[["no_shape"]] <- fit_dt_noshape
+
 ggplot() +
-  geom_point(data = dplyr::filter(fit_dt, shape_class == "convex"),
+  geom_point(data = dplyr::filter(fit_dt_noshape, shape_class == "convex"),
              aes(x = lake_maxdepth_m, y = fitted_values, color = shape_class),
              alpha = 0.3) +
-  geom_point(data = dplyr::filter(fit_dt, shape_class == "concave"),
+  geom_point(data = dplyr::filter(fit_dt_noshape, shape_class == "concave"),
              aes(x = lake_maxdepth_m, y = fitted_values, color = shape_class),
              alpha = 0.3) +
   geom_abline(slope = 1, intercept = 0) +
@@ -28,8 +32,11 @@ ggplot() +
   ggtitle("Max depth ~ Lake Area + Buffer slope")
   # theme(legend.position = "none")
 
+# residual = observed - predicted
+# positive value = underprediction
+# negative value = overprediction
 ggplot() +
-  geom_boxplot(data = fit_dt, aes(y = resid, x = shape_class),
+  geom_boxplot(data = fit_dt_noshape, aes(y = resid, x = shape_class),
                outlier.shape = NA) +
   ylim(-20, 20)
 
@@ -37,17 +44,19 @@ ggplot() +
 fit <- lm(lake_maxdepth_m ~
             buffer100m_slope_max + lake_waterarea_ha + shape_class,
           data = dt)
-fit_dt <- data.frame(resid = fit$residuals,
+fit_dt_shape <- data.frame(resid = fit$residuals,
                      shape_class = dt$shape_class,
                      fitted_values = fit$fitted.values,
[…]

### 2020-01-08 scripts/05_hypso_model.R

@@ -1,25 +1,23 @@
 source("scripts/99_utils.R")
 # setwd("../")
 
-dt <- read.csv("data/00_hypso/hypso_predictors.csv", stringsAsFactors = FALSE)
-dt$shape_class <- factor(dt$shape_class)
-
-ggplot(data = dplyr::filter(dt,
-                            !is.na(shape_class) &
-                              shape_class != "neither")) +
-  geom_boxplot(aes(x = shape_class, y = lake_waterarea_ha), outlier.shape = NA) +
-  ylim(c(0, 300)) +
-  facet_wrap(~lake_centroidstate)
-
-# logistic model yes/no bowl-shaped lake
-test <- glm(shape_class ~ lake_waterarea_ha + ws_mgbconhull_length_m,
-            data = dplyr::filter(dt, shape_class != "neither"), family = "binomial")
-hist(test$fitted.values)
-
 library(mlr3)
 library(mlr3learners)
 library(ranger)
 
+dt <- read.csv("data/00_hypso/hypso_predictors.csv", stringsAsFactors = FALSE)
+dt$shape_class <- factor(dt$shape_class)
+
+res <- list()
+
+# ggplot(data = dplyr::filter(dt,
+#                             !is.na(shape_class) &
+#                               shape_class != "neither")) +
+#   geom_boxplot(aes(x = shape_class, y = lake_waterarea_ha), outlier.shape = NA) +
+#   ylim(c(0, 300)) +
+#   facet_wrap(~lake_centroidstate)
+
+# ---- model setup ----
 dt_sub       <- dt %>%
   # dplyr::filter(lake_centroidstate %in% c("MN", "MI", "CT")) %>%
   dplyr::filter(!is.na(shape_class) & shape_class != "neither") %>%
@@ -41,6 +39,7 @@ train_set  <- sample(class_task$nrow, 0.9 * class_task$nrow)
 test_set   <- setdiff(seq_len(class_task$nrow), train_set)
 measure    <- msr("classif.acc")
 
+# ---- fit rpart model ----
 learner    <- lrn("classif.rpart", cp = 0)
 learner$train(class_task, row_ids = train_set)
 learner$importance()
@@ -52,7 +51,14 @@ pred = rr$prediction()
 pred$confusion
 pred$score(measure)
 
-# ----
+fit_rpart <- list(
+  importance = learner$importance(),
+  confusion = pred$confusion,
+  score = pred$score(measure))
+
[…]

### 2020-01-09 figures/01_geometry.Rmd
add geometry figures

--- 2020-01-07 figures/01_hypsography.Rmd
@@
 
-```{r 01_hypsography, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r 01_geometry, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+# example map of true deepest point versus visual center
 
-ids <- factor(c("convex", "concave"))
-polys <- data.frame(
-  id = rep(ids, each = 3),
-  x = c(0, 100, 0, 
-        0, 100, 100), 
-  y = c(100, 0, 0, 
-        100, 100, 0))
+dt <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
 
-dt_raw <- read.csv("../data/00_hypso/hypso.csv", stringsAsFactors = FALSE)
+# plot of true deepest versus visual centers
 
-ggplot(data = dt_raw) +
-  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
-  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
-  geom_abline(intercept = -100, slope = 1, color = "red") +
-  ylim(c(100, 0)) +
-  ylab("Depth (percent)") + xlab("Area (percent)") + 
-  ggtitle(paste0("Normalized hypsography for ", length(unique(dt_raw$llid)), " lakes")) + 
-  theme(legend.position = c(0.84, 0.2), 
-        legend.title = element_blank()) +
-  facet_wrap(~state)
+(gg_test <- ggplot(data = st_drop_geometry(dt)) + 
+  geom_point(aes(x = dist_deepest, y = dist_viscenter)) + # , color = llid)) +
+  geom_abline(aes(intercept = 0, slope = 1)) +
+  xlab("Distance to shore from deepest point") +
+  ylab("Distance to shore from visual center"))
+
+# plotly::ggplotly(gg_test)
+# plot(res$dist_deepest, res$dist_viscenter)
+
+get_pnt_map <- function(x){
+  # x <- 3
+  lg_example <- LAGOSUSgis::query_gis(
+    "LAGOS_US_All_Lakes_1ha", "lagoslakeid", dt$llid[x])
+  crs_example <- st_crs(
+    raster(list.files(paste0("../data/", tolower(dt$state[x]), "_bathy"), 
+               include.dirs = TRUE, full.names = TRUE)[1])
+    )
+  st_crs(dt$pnt_deepest)   <- crs_example
+  st_crs(dt$pnt_viscenter) <- crs_example
+
+  ggplot() + 
+    geom_sf(data = lg_example) +
+    geom_sf(data = dt$pnt_deepest[x], color = "red") + 
+    geom_sf(data = dt$pnt_viscenter[x], color = "green") +
+    theme_void()
+}
+
+# which(dt$llid == 5636)
+# which(dt$llid == 257862)
+# get_pnt_map(2352)
+cowplot::plot_grid(get_pnt_map(which(dt$llid == 3791)) +
+                     ggtitle(3791), 
+                   get_pnt_map(which(dt$llid == 2772)) +
+                     ggtitle(2772), 
+                   get_pnt_map(which(dt$llid == 2655)) +
+                     ggtitle(2655), 
+                   get_pnt_map(which(dt$llid == 1822)) +
+                     ggtitle(1822)
+                   )
 

### 2020-01-09 manuscript/figures.Rmd
add messager equation table
add geometry figures

@@
 
-![Slope diagram caption](../figures/slope_diagram.pdf){width=100%}
-
 ```{r 02_depth_model_fitted, fig.cap = "A model that does not include shape class over predicts depth in concave lakes but does not systematically under predict depth in convex lakes.", warning=FALSE, echo = FALSE, eval=TRUE}
@@
 ```
+
+![Messager et al. 2016 equation table](../figures/messager_eq_table.png){width=100%}
+
+```{r 01_geometry_points, fig.cap = "Caption.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_geometry-1.pdf")
+```
+
+```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_geometry-2.pdf")
+```
+
+![Slope diagram caption](../figures/slope_diagram.pdf){width=100%}

### 2020-01-10 figures/01_contrasts.Rmd
add reservoir classifications

@@
 
-```{r 01_contrasts_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.5}
+```{r gg_depth_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+
+gg_depth_reservoir <- dt %>%
+  dplyr::filter(!is.na(reservoir_class)) %>%
+  group_by(reservoir_class) %>%
+  summarize(lake_maxdepth_m = list(tibble::enframe(
+    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = reservoir_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Depth (m)") + xlab("")
+```
+
+```{r gg_depth_xtabs, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+dt_sub <- dplyr::filter(dt, 
+                        !is.na(shape_class), shape_class != "neither", 
+                        !is.na(reservoir_class))
+# group_by(dt_sub, lake_glaciatedlatewisc, reservoir_class) %>%
+#   tally() %>%
+#   arrange(lake_glaciatedlatewisc) %>%
+#   gridExtra::grid.table()
+
+gg_depth_xtabs <- group_by(dt_sub, 
+                           lake_glaciatedlatewisc, reservoir_class) %>%
+    summarize(lake_maxdepth_m = list(tibble::enframe(
+    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  tidyr::unite(cat, lake_glaciatedlatewisc:reservoir_class) %>%
+  ggplot() +
+geom_pointrange(aes(x = cat, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Depth (m)") + xlab("")
+
+```
+
+```{r 01_contrasts_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 6}
 theme_opts <- theme(axis.text.x = element_text(angle = 90))
 
+cowplot::plot_grid(  
 cowplot::plot_grid(gg_depth_shape_class + 
@@
                      theme(axis.title.y = element_blank()), 
-                   nrow = 1)
+                   gg_depth_reservoir + 
+                     theme_opts +
+                     theme(axis.title.y = element_blank()), 
+                   nrow = 1), 
+cowplot::plot_grid(gg_depth_xtabs + theme_opts, 
+                   ggplot() + geom_blank() + 
+                     theme(panel.background = element_blank()), nrow = 1), 
+nrow = 2)
+
 ```
@@
 
-```{r glaciated_vs_shapeclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r gg_tally_glaciated_vs_shapeclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
 dt_sub <- dplyr::filter(dt, !is.na(shape_class), shape_class != "neither")

### 2020-01-10 scripts/00_get_reservoir_class.R

@@ -1,0 +1,34 @@
+# setwd("../")
+# From Dropbox: contlimno2 > manuscripts > RSVRs (or Reservoirs)
+
+library(readxl)
+library(dplyr)
+
+lg <- LAGOSUS::lagosus_load("locus")
+
+manual_class_raw <- readxl::read_excel(
+  "data/00_reservoir_classification/LAGOSNE_MANUAL_CLASSIFICATIONS_V1.0.xlsx") %>%
+  janitor::clean_names() %>%
+  rename(reservoir_class = lake_type) %>%
+  mutate(reservoir_class = case_when(
+    reservoir_class == "RSVR" ~ "Res",
+    TRUE ~ reservoir_class))
+
+pred_class_raw <- readxl::read_excel(
+  "data/00_reservoir_classification/LAGOSNE_PREDICTION_V1.0.xlsx") %>%
+  janitor::clean_names() %>%
+  dplyr::filter(confidence >= 0.75) %>%
+  rename(reservoir_class = prediction)
+
+dt <- dplyr::bind_rows(manual_class_raw, pred_class_raw) %>%
+  arrange(lagoslakeid) %>%
+  rename(lagosne_lagoslakeid = lagoslakeid) %>%
+  left_join(dplyr::select(lg$locus$locus_link, lagoslakeid, lagosne_lagoslakeid),
+                          by = "lagosne_lagoslakeid") %>%
+  dplyr::filter(!is.na(lagoslakeid)) %>%
+  distinct(lagoslakeid, .keep_all = TRUE) %>%
+  dplyr::select(lagoslakeid, reservoir_class)
+
+write.csv(dt,
+          "data/00_reservoir_classification/reservoir_classes_clean.csv",
+          row.names = FALSE)

### 2020-01-10 scripts/04_depth_predictors.R

@@ -46,10 +46,16 @@ dt_raw <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
                 -contains("predicted"))
 # fill missing nws data with ws data?
 
-dt_raw_ne <- dplyr::filter(dt_raw, !is.na(buffer100m_slope_max))
+# add reservoir class labels
+dt_raw <- dt_raw %>%
+  left_join(dplyr::select(read.csv(
+  "data/00_reservoir_classification/reservoir_classes_clean.csv",
+  stringsAsFactors = FALSE), lagoslakeid, reservoir_class), by = "lagoslakeid")
 
 write.csv(dt_raw, "data/lagosus_depth_predictors.csv", row.names = FALSE)
 # dt_raw <- read.csv("data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE)
+
+dt_raw_ne <- dplyr::filter(dt_raw, !is.na(buffer100m_slope_max))
 write.csv(dt_raw_ne, "data/lagosne_depth_predictors.csv", row.names = FALSE)
 
 if(interactive()){

### 2020-01-15 figures/00_maps.Rmd
rm duplicate measures based on program reliability

@@
 
-dt_plot <- dplyr::select(lg$locus$locus_information, 
-                         -contains("decdeg"), 
-                         -contains("states"), 
-                         -contains("gnis")) %>%
+dt_plot <- dt %>%
   distinct(lagoslakeid, .keep_all = TRUE) %>%
-  left_join(dt, by = "lagoslakeid") %>%
   dplyr::filter(lake_waterarea_ha >= 4) %>%
-  group_by(lake_centroidstate) %>%
+  group_by(lake_state) %>%
   summarize(prop_maxdepth = mean(!is.na(lake_maxdepth_m))) %>%
@@
     dplyr::select(states_all, postal, iso_a2), iso_a2 == "US"), 
-            by = c("lake_centroidstate" = "postal")) %>%
+            by = c("lake_state" = "postal")) %>%
   dplyr::filter(!is.na(prop_maxdepth))

### 2020-01-15 scripts/02_qa.R

@@ -1,31 +1,104 @@
+# setwd("../")
 source("scripts/99_utils.R")
+
+lg        <- lagosus_load(module = "locus")
+
+# ---- remove obs that seem wrong in lakes with repeated measures ----
+threshold <- 0.4
+
+dt_raw <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
+  dplyr::filter(!is.na(lake_maxdepth_m))
+dt     <- dt_raw %>%
+  group_by(lagoslakeid) %>%
+  add_tally() %>%
+  mutate(diff = (max(lake_maxdepth_m) - min(lake_maxdepth_m)) /
+           max(lake_maxdepth_m)) %>%
+  mutate(diff = case_when(n < 2 ~ 0,
+                          TRUE ~ diff)) %>%
+  arrange(desc(diff))
+
+# remove NLA first because we're calling it the most unreliable
+dt     <- dt %>%
+  ungroup(dt) %>%
+  dplyr::filter(!(lagos_effort == "NLA" & diff > threshold)) %>%
+  dplyr::select(-n, -diff) %>%
+  group_by(lagoslakeid) %>%
+  add_tally() %>%
+  mutate(diff = (max(lake_maxdepth_m) - min(lake_maxdepth_m)) /
+           max(lake_maxdepth_m)) %>%
+  mutate(diff = case_when(n < 2 ~ 0,
+                          TRUE ~ diff)) %>%
+  arrange(desc(diff))
+
+# remove LAGOSNE second because we're calling it the second most unreliable
+dt     <- dt %>%
+  ungroup(dt) %>%
+  dplyr::filter(!(lagos_effort == "LAGOSNE" & diff > threshold)) %>%
+  dplyr::select(-n, -diff) %>%
+  group_by(lagoslakeid) %>%
+  add_tally() %>%
+  mutate(diff = (max(lake_maxdepth_m) - min(lake_maxdepth_m)) /
+           max(lake_maxdepth_m)) %>%
+  mutate(diff = case_when(n < 2 ~ 0,
+                          TRUE ~ diff)) %>%
+  arrange(desc(diff))
+
+# remove bathymetry third because this is likely due to NHD hi-res polygons
+# being mismatched to gnis names?
+# dt <- dt %>%
+#   ungroup(dt) %>%
[…]

### 2020-01-15 figures/01_contrasts.Rmd
print intermediate figs

@@
 
-gg_depth_reservoir <- dt %>%
+(gg_depth_reservoir <- dt %>%
   dplyr::filter(!is.na(reservoir_class)) %>%
@@
                       ymin = `5%`, ymax = `95%`)) +
-  ylab("Depth (m)") + xlab("")
+  ylab("Depth (m)") + xlab(""))
 ```
@@
 
-gg_depth_xtabs <- group_by(dt_sub, 
+(gg_depth_xtabs <- group_by(dt_sub, 
                            lake_glaciatedlatewisc, reservoir_class) %>%
@@
                       ymin = `5%`, ymax = `95%`)) +
-  ylab("Depth (m)") + xlab("")
+  ylab("Depth (m)") + xlab(""))
 

### 2020-01-15 scripts/00_get_manual.R
bring in PA aftermarket data

@@ -8,6 +8,22 @@ source("scripts/99_utils.R")
 dt_raw <- suppressWarnings(suppressMessages(
   read_csv("data/00_manual/depth_log_all.csv", col_types = cols())))
 
+# collected in later efforts but not merged in overall log
+# drive_download(file = "PA_USGS_sites_linked_R",
+#                path = "data/00_manual/pa.csv", overwrite = TRUE)
+dt_pa <- suppressMessages(read_csv("data/00_manual/pa.csv")) %>%
+  janitor::clean_names() %>%
+  mutate(file_name = "PA_USGS_sites_linked_R",
+         state = "PA") %>%
+  dplyr::select(Linked_lagoslakeid = linked_lagoslakeid,
+                lat = latitude, lon = longitude,
+                name = lakename_google, file_name,
+                state, max_depth_ft, mean_depth_ft,
+                lakename_google, max_depth_m, mean_depth_m,
+                url, comments)
+
+dt_raw <- dplyr::bind_rows(dt_raw, dt_pa)
+
 # collected in very early efforts and lack a lagoslakeid
 one_off_files <- c("data/00_manual/ar.csv" = "AR_ADEQ_SWQM_sites_lakedepth",
                    "data/00_manual/ca.csv" = "CA_USGS_WSC_sites_lakedepth",

### 2020-01-15 manuscript/figures.Rmd
more specific caption

@@
 
-```{r maps_state, fig.cap = "Proportion of lakes greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r maps_state, fig.cap = "Proportion of ALL lakes greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/00_map_state-1.pdf")

### 2020-01-17 scripts/00_get_gis.R
deal with adding new fields to dt layer

@@ -1,3 +1,4 @@
+# setwd("../")
 source("scripts/99_utils.R")
 
 lg <- LAGOSUS::lagosus_load("locus")
@@ -64,3 +65,7 @@ st_write(study_bbox, gpkg_path, layer = "study_bbox", update = TRUE,
          layer_options = c("OVERWRITE=yes"))
 st_write(dt, gpkg_path, layer = "dt", update = TRUE,
          layer_options = c("OVERWRITE=yes"))
+# library(RSQLite)
+# db = SQLite()
+# con = dbConnect(db, "data/gis.gpkg")
+# dbRemoveTable(con, "dt")

### 2020-01-17 figures/00_maps.Rmd
add map of depth availability in limno lakes

@@
 
-# add text labels of proportions
+ggplot() + 
+  geom_sf(data = dt_plot, aes(fill = prop_maxdepth)) + 
+  geom_sf_text(data = dt_plot, 
+               aes(label = round(prop_maxdepth, 2)), 
+               size = 2.5) +
+  labs(fill = "") +
+  coord_sf(datum = NA) + theme_void()
+
+# limit to lakes with limno data
+dt_plot <- dt %>%
+  distinct(lagoslakeid, .keep_all = TRUE) %>%
+  dplyr::filter(lake_waterarea_ha >= 4) %>%
+  dplyr::filter(has_limno == 1) %>%
+  group_by(lake_state) %>%
+  summarize(prop_maxdepth = mean(!is.na(lake_maxdepth_m))) %>%
+  arrange(desc(prop_maxdepth)) %>%
+  left_join(dplyr::filter(
+    dplyr::select(states_all, postal, iso_a2), iso_a2 == "US"), 
+            by = c("lake_state" = "postal")) %>%
+  dplyr::filter(!is.na(prop_maxdepth))
+st_geometry(dt_plot) <- dt_plot$geom
 

### 2020-01-17 manuscript/figures.Rmd
add map of depth availability in limno lakes

--- 2020-01-09 manuscript/figures.Rmd
@@
 
-```{r maps_state, fig.cap = "Proportion of lakes greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r maps_state_all, fig.cap = "Proportion of ALL lakes greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/00_map_state-1.pdf")
+```
+
+```{r maps_state_limno, fig.cap = "Proportion of lakes with limno data greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/00_map_state-2.pdf")
 ```

### 2020-01-24 figures/01_geometry.Rmd
plot geometry distances as grids

--- 2020-01-07 figures/01_hypsography.Rmd
@@
 
-```{r 01_hypsography, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r 01_geometry, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}
+# example map of true deepest point versus visual center
 
-ids <- factor(c("convex", "concave"))
-polys <- data.frame(
-  id = rep(ids, each = 3),
-  x = c(0, 100, 0, 
-        0, 100, 100), 
-  y = c(100, 0, 0, 
-        100, 100, 0))
+# setwd("figures")
+dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
+hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
+                       stringsAsFactors = FALSE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid))
+dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class), 
+                                  by = c("llid" = "lagoslakeid"))
 
-dt_raw <- read.csv("../data/00_hypso/hypso.csv", stringsAsFactors = FALSE)
+# plot of true deepest versus visual centers
 
-ggplot(data = dt_raw) +
-  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
-  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
-  geom_abline(intercept = -100, slope = 1, color = "red") +
-  ylim(c(100, 0)) +
-  ylab("Depth (percent)") + xlab("Area (percent)") + 
-  ggtitle(paste0("Normalized hypsography for ", length(unique(dt_raw$llid)), " lakes")) + 
-  theme(legend.position = c(0.84, 0.2), 
-        legend.title = element_blank()) +
-  facet_wrap(~state)
+gg_scatter_state <- ggplot(data = st_drop_geometry(dt)) + 
+  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = state)) +
+  geom_abline(aes(intercept = 0, slope = 1)) +
+  xlab("Distance to shore from deepest point") +
+  ylab("Distance to shore from visual center")
+
+# plotly::ggplotly(gg_test)
+# plot(res$dist_deepest, res$dist_viscenter)
+
+gg_scatter_reservoir <- ggplot(data = st_drop_geometry(dt)) + 
+  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = reservoir_class)) +
+  geom_abline(aes(intercept = 0, slope = 1)) +
+  xlab("Distance to shore from deepest point") +
+  ylab("Distance to shore from visual center")
+
+cowplot::plot_grid(gg_scatter_state + 
+                     theme(axis.title.x = element_text(color = "white")), 
+                   gg_scatter_reservoir + 
+                     labs(color = "") +
+                     theme(axis.title.y = element_blank(), 
+                           axis.text.y = element_blank()), 
+                   rel_widths = c(1, 0.8))
+
+gg_pntrng_deepest <- st_drop_geometry(dt) %>%
+    dplyr::filter(!is.na(reservoir_class)) %>%
+  group_by(reservoir_class) %>%
+  summarize(dist_deepest = list(tibble::enframe(
+    quantile(dist_deepest, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = reservoir_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Distance to shore (m)") + xlab("")
+
+gg_pntrng_viscenter <- st_drop_geometry(dt) %>%
+    dplyr::filter(!is.na(reservoir_class)) %>%
+  group_by(reservoir_class) %>%
+  summarize(dist_viscenter = list(tibble::enframe(
+    quantile(dist_viscenter, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = reservoir_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Distance to lake center (m)") + xlab("")
+
+cowplot::plot_grid(gg_pntrng_deepest, gg_pntrng_viscenter)
+
+get_pnt_map <- function(x){
+  # x <- 3
+  # x <- which(dt$llid == 257862)
+  lg_example <- LAGOSUSgis::query_gis(
+    "LAGOS_US_All_Lakes_1ha", "lagoslakeid", dt$llid[x])
+  crs_example <- st_crs(
+    raster(list.files(paste0("../data/", tolower(dt$state[x]), "_bathy"), 
+               include.dirs = TRUE, full.names = TRUE)[1])
+    )
+  st_crs(dt$pnt_deepest)   <- crs_example
+  st_crs(dt$pnt_viscenter) <- crs_example
+
+  ggplot() + 
+    geom_sf(data = lg_example) +
+    geom_sf(data = dt$pnt_deepest[x], color = "red") + 
+    geom_sf(data = dt$pnt_viscenter[x], color = "green") +
+    theme_void()
+}
+
+# get_pnt_map(which(dt$llid == 5636))
+# get_pnt_map(which(dt$llid == 257862))
+cowplot::plot_grid(get_pnt_map(which(dt$llid == 3791)) +
+                     ggtitle(3791), 
+                   get_pnt_map(which(dt$llid == 2772)) +
+                     ggtitle(2772), 
+                   get_pnt_map(which(dt$llid == 2655)) +
+                     ggtitle(2655), 
+                   get_pnt_map(which(dt$llid == 1822)) +
+                     ggtitle(1822)
+                   )
 
 ```
+
+```{r }
+# compare inlake slope against buffer slope
+
+dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
+hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
+                       stringsAsFactors = FALSE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid))
+dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, 
+                                  buffer100m_slope_max), 
+                                  by = c("llid" = "lagoslakeid"))
+
+plot(dt$inlake_slope, dt$buffer100m_slope_max/10)
+
+```

### 2020-01-24 figures/02_depth_model.Rmd
add oliver 2016 figures

@@
   ylim(-20, 20)
+```
 
+```{r 02_depth_oliver, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.4}
+# setwd("figures")
+oliver2015 <- LAGOSNE::lagos_load_oliver_2015() %>%
+  mutate(lake_maxdepth_m_oliver = zmaxpredict)
+dt <- readRDS("../data/01_depth_model/oliver_model.rds")
+
+rmse_group <- group_by(dt$test, hu4_zoneid) %>%
+  summarize(r_rmse = yardstick::rmse_vec(
+    exp(lake_maxdepth_m), 
+    exp(lake_maxdepth_m_predicted)) / mean(lake_maxdepth_m))
+hu4_gis <- st_read("../data/gis.gpkg", layer = "hu4s_focal_simple") %>%
+# ggplot() + 
+#   geom_sf(data = hu4_gis) +
+#   geom_sf_text(data = hu4_gis, aes(label = hu4_zoneid))
+  left_join(rmse_group) %>%
+  dplyr::filter(!is.na(r_rmse) & r_rmse != Inf & r_rmse < 12)
+ggplot() +
+  geom_sf(data = hu4_gis, aes(fill = r_rmse))
+
+## predicted vs obs
+ggplot(data = dt$test) + 
+  geom_point(aes(x = lake_maxdepth_m_predicted, 
+                 y = lake_maxdepth_m)) +
+  geom_abline(aes(slope = 1, intercept = 0))
+
+## calculate metrics
+gridExtra::grid.arrange(gridExtra::tableGrob(
+bind_rows(yardstick::rmse(dt$test, 
+                          exp(lake_maxdepth_m), exp(lake_maxdepth_m_predicted)), 
+          yardstick::rsq(dt$test,
+                         exp(lake_maxdepth_m),
+                         exp(lake_maxdepth_m_predicted)))))
+
+## replication vs oliver
+# dt$test <- dt$test %>%
+#   left_join(dplyr::select(dplyr::mutate(oliver2015, lagoslakeid = as.character(lagoslakeid)), lagoslakeid,
+#                           lake_maxdepth_m_oliver))
+# ggplot(data = dt$test) + 
+#   geom_point(aes(x = exp(lake_maxdepth_m_predicted), 
+#                  y = lake_maxdepth_m_oliver)) +
+#   geom_abline(aes(slope = 1, intercept = 0))
+
+cowplot::plot_grid(
+  ggplot(data = oliver2015) +
+    geom_point(aes(x = zmaxpredict, y = zmaxobs)) +
+    geom_abline(aes(slope = 1, intercept = 0)), 
+  ggplot(data = oliver2015) +
+    geom_point(aes(x = log(zmaxpredict), y = log(zmaxobs))) +
+    geom_abline(aes(slope = 1, intercept = 0))
+)
 ```

### 2020-01-24 scripts/05_depth_model.R

@@ -109,23 +109,27 @@ test  <- dplyr::anti_join(dt, train, by = 'lagoslakeid')
 
 ## fit model
 library(lme4)
-eq <- as.formula("lake_maxdepth_m ~ buffer100m_slope_max + lake_waterarea_ha + ws_lake_arearatio + lake_shorelinedevfactor_nounits + (lake_waterarea_ha | hu4_zoneid)")
+eq <- as.formula("lake_maxdepth_m ~ buffer100m_slope_max + lake_waterarea_ha + (lake_waterarea_ha | hu4_zoneid)")
 fit <- lme4::lmer(eq, data = train, REML = FALSE)
-
-## predict on holdouts
 test$lake_maxdepth_m_predicted <- predict(fit, test)
-plot(test$lake_maxdepth_m_predicted, test$lake_maxdepth_m, xlim = c(0, 5), ylim = c(0, 5))
-abline(0, 1)
 
-## calculate rmse
-yardstick::rmse(test, exp(lake_maxdepth_m), exp(lake_maxdepth_m_predicted))
-yardstick::rsq(test, exp(lake_maxdepth_m), exp(lake_maxdepth_m_predicted))
+saveRDS(list(fit = fit, test = test, train = train),
+        "data/01_depth_model/oliver_model.rds")
 
-test <- test %>%
-  left_join(dplyr::select(dplyr::mutate(oliver2015, lagoslakeid = as.character(lagoslakeid)), lagoslakeid,
-                          lake_maxdepth_m_oliver))
-plot(exp(test$lake_maxdepth_m_predicted), test$lake_maxdepth_m_oliver)
-abline(0, 1)
+if(interactive()){
+  plot(test$lake_maxdepth_m_predicted, test$lake_maxdepth_m, xlim = c(0, 5), ylim = c(0, 5))
+  abline(0, 1)
 
-plot(oliver2015$zmaxpredict, oliver2015$zmaxobs)
-abline(0, 1)
+  ## calculate rmse
+  yardstick::rmse(test, exp(lake_maxdepth_m), exp(lake_maxdepth_m_predicted))
+  yardstick::rsq(test, exp(lake_maxdepth_m), exp(lake_maxdepth_m_predicted))
+
+  test <- test %>%
+    left_join(dplyr::select(dplyr::mutate(oliver2015, lagoslakeid = as.character(lagoslakeid)), lagoslakeid,
[…]

### 2020-01-24 figures/01_geometry.Rmd
debugging distance outliers

@@
 dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, 
-                                  buffer100m_slope_max), 
+                                  buffer100m_slope_mean), 
                                   by = c("llid" = "lagoslakeid"))
 
-plot(dt$inlake_slope, dt$buffer100m_slope_max/10)
+plot(dt$inlake_slope, dt$buffer100m_slope_mean/10, 
+     xlim = c(0, 3), ylim = c(0, 3))
+abline(0, 1)
+
+# View(arrange(dt, buffer100m_slope_mean))
 

### 2020-01-24 manuscript/figures.Rmd
rebuild

@@
 
+```{r constrasts_distance, fig.cap = "Contrasting distance from shore to the deepest point in the lake and distance to the lake visual center for natural lakes and reservoirs.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_geometry-2.pdf")
+```
+
 ```{r glaciated_vs_shapeclass, fig.cap = "Cross tabulation between lake glaciation and shape class.
@@
 
+```{r 02_depth_oliver-2, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/02_depth_oliver-2.pdf")
+```
+
+```{r 02_depth_oliver-1, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/02_depth_oliver-1.pdf")
+```
+
+```{r 02_depth_oliver-4, fig.cap = "Oliver 2016 observed versus predicted.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/02_depth_oliver-4.pdf")
+```
+
 ![Lake shape caption](../figures/lake_shape.pdf){width=100%}
@@
 
-```{r 01_geometry_points, fig.cap = "Caption.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r 01_geometry_points, fig.cap = "Distance to deepest point versus distance to lake visual center.", warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/01_geometry-1.pdf")
@@
 
+\pagebreak
+
 ```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_geometry-2.pdf")
+knitr::include_graphics("../figures/01_geometry-3.pdf")
 ```

### 2020-01-28 figures/00_maps.Rmd
compress map figure, closes #19

@@
 states_focal  <- st_read("../data/gis.gpkg", "states_focal", quiet = TRUE)
-hu4_focal     <- st_read("../data/gis.gpkg", "hu4s_focal_simple", quiet = TRUE)
-hu4_focal     <- st_crop(st_transform(
-  lwgeom::st_make_valid(hu4_focal), st_crs(study_bbox)), 
-  study_bbox)
+# hu4_focal     <- st_read("../data/gis.gpkg", "hu4s_focal_simple", quiet = TRUE)
+# hu4_focal     <- st_crop(st_transform(
+#   lwgeom::st_make_valid(hu4_focal), st_crs(study_bbox)), 
+#   study_bbox)
+
+dt_notsf <- cbind(st_drop_geometry(dt), st_coordinates(dt))
 
@@
     geom_sf(data = states_focal, fill = "white", color = "black", size = 0.4) + 
-    geom_sf(data = dt, aes(color = maxdepth_class), size = 0.3, alpha = 0) +
+    geom_point_rast(data = dt_notsf, aes(x = X, y = Y, color = maxdepth_class), 
+                    size = 0.3, alpha = 0) +
     scale_color_manual(values = viridisLite::viridis(5)) +
-    geom_sf(data = dplyr::filter(dt, lake_maxdepth_m < 10),
-            aes(color = maxdepth_class), size = 0.3, alpha = 0.8) +
-    geom_sf(data = dplyr::filter(dt, lake_maxdepth_m > 10), 
-            aes(color = maxdepth_class), size = 0.3, alpha = 0.8) +
+    geom_point_rast(data = dplyr::filter(dt_notsf, lake_maxdepth_m < 10), 
+            aes(x = X, y = Y, color = maxdepth_class), size = 0.3, alpha = 0.8) +
+    geom_point_rast(data = dplyr::filter(dt_notsf, lake_maxdepth_m > 10), 
+            aes(x = X, y = Y, color = maxdepth_class), size = 0.3, alpha = 0.8) +
     labs(color = "Max depth (m)") +

### 2020-01-28 scripts/00_get_gis.R

@@ -43,6 +43,8 @@ hu4s_focal_simple <- rmapshaper::ms_simplify(
           st_transform(study_bbox, st_crs(hu4s_focal)))
 )
 
+states_all   <- rmapshaper::ms_simplify(states_all)
+
 # study_bbox   <- concaveman::concaveman(st_cast(hu4s_focal, "POINT"))
 
 # unlink("data/gis.gpkg")

### 2020-01-28 scripts/99_utils.R

@@ -25,6 +25,7 @@ library(polylabelr)
 library(esri2sf) # install_github("yonghah/esri2sf")
 library(rnaturalearth)
 library(mapview)
+library(ggrastr)
 
 # ---- misc fxn ----
 # jsta::get_if_not_exists

### 2020-01-28 figures/01_geometry.Rmd
fix axis label

@@
                       ymin = `5%`, ymax = `95%`)) +
-  ylab("Distance to shore (m)") + xlab("")
+  ylab("Distance from shore to deepest point (m)") + xlab("")
 

### 2020-01-31 figures/01_geometry.Rmd
rm ggplotly tests

@@
 # plot of true deepest versus visual centers
-
 gg_scatter_state <- ggplot(data = st_drop_geometry(dt)) + 
@@
   ylab("Distance to shore from visual center")
-
-# plotly::ggplotly(gg_test)
-# plot(res$dist_deepest, res$dist_viscenter)
 

### 2020-01-31 figures/00_maps.Rmd
add map of bathymetry locations, closes #21

@@
 ```
+
+```{r 00_map_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+bathy_pnts <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
+dt            <- st_read("../data/gis.gpkg", "dt", quiet = TRUE)
+max_depth <- dt$lake_maxdepth_m
+dt <- dt %>%
+  mutate(lake_maxdepth_m = round(lake_maxdepth_m, 2)) %>%
+  mutate(maxdepth_class = 
+           smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, max(max_depth)), 
+                     brackets = NULL, sep = " - ")
+         ) %>% # unique(dt$maxdepth_class)
+  dplyr::filter(!is.na(maxdepth_class))
+
+dt_bathy <- st_drop_geometry(bathy_pnts) %>% 
+  left_join(
+    dplyr::distinct(dplyr::mutate(dplyr::select(dt, lagoslakeid, maxdepth_class, 
+                                                has_limno, lake_waterarea_ha),
+                  lagoslakeid = as.character(lagoslakeid)), lagoslakeid, .keep_all = TRUE),
+            by = c("llid" = "lagoslakeid"))
+
+dt_bathy <- st_sf(dt_bathy, geometry = dt_bathy$geom)
+
+gg_bathy <- 
+  ggplot() + 
+    geom_sf(data = study_bbox[2,], alpha = 1, size = 0.6, 
+          color = "black", fill = "#1D9DCF") +
+    geom_sf(data = states_all, fill = "gray", color = "black", size = 0.3) +
+    geom_sf(data = states_focal, fill = "white", color = "black", size = 0.4) + 
+    geom_sf(data = dt_bathy, aes(color = maxdepth_class), 
+                    size = 0.3, alpha = 1) +
+    scale_color_manual(values = viridisLite::viridis(5)) +
+    labs(color = "Max depth (m)") +
+    coord_sf(datum =  NA) +
+    theme_void()
+
+gg_bathy
+```

### 2020-01-31 manuscript/figures.Rmd
add map of bathymetry locations, closes #21

@@
 knitr::include_graphics("../figures/00_map-1.pdf")
+```
+
+![Summary of lake depth predictor variables.](../tables/01_predictors.pdf){width=65%}
+
+```{r maps_dot_bathy, fig.cap = "Map of lake maximum depth measurements derived from lake bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/00_map_bathy-1.pdf")
 ```
@@
 
-```{r 01_predictors, results='asis', warning=FALSE, echo = FALSE}
-knitr::kable('\\includegraphics[width=6.5in]{"../tables/01_predictors"}', col.names = NULL, caption = "Summary of lake depth predictor variables.")
-```
-
 ```{r 02_depth_model_fitted, fig.cap = "A model that does not include shape class over predicts depth in concave lakes but does not systematically under predict depth in convex lakes.", warning=FALSE, echo = FALSE, eval=TRUE}

### 2020-01-31 scripts/00_get_hypso_me.R
remove lakes with less than 3 soundings

@@ -17,11 +17,10 @@ if(!file.exists("data/me_bathy/lakedpths.zip")){
   unzip("data/me_bathy/lakedpths.zip", exdir = "data/me_bathy/")
 }
 
-pnts <- sf::st_read("data/me_bathy/lakedpth.shp")
-
 # pull lagosus points
 lg_poly <- LAGOSUSgis::query_gis_(query =
                                     "SELECT * FROM LAGOS_US_All_Lakes_1ha WHERE lake_centroidstate LIKE 'ME' AND lake_totalarea_ha > 4")
+pnts <- sf::st_read("data/me_bathy/lakedpth.shp")
 pnts <- st_transform(pnts, st_crs(lg_poly))
 pnts <- st_join(pnts, lg_poly) %>%
   dplyr::filter(!is.na(lagoslakeid))
@@ -32,6 +31,12 @@ lg_xwalk <- read.csv("data/00_lagosne/00_lagosne_xwalk.csv",
                      stringsAsFactors = FALSE)
 pnts <- dplyr::filter(pnts,
                            lagoslakeid %in% unique(lg_xwalk$lagoslakeid))
+few_soundings <- pnts %>%
+  group_by(lagoslakeid) %>%
+  add_tally() %>%
+  dplyr::filter(n < 3) %>%
+  pull(lagoslakeid)
+pnts <- dplyr::filter(pnts, !(lagoslakeid %in% few_soundings))
 
 lg_ps <- dplyr::filter(lg_poly, lagoslakeid %in% pnts$lagoslakeid)
 
@@ -44,7 +49,7 @@ rsubs <- lapply(seq_along(unique(pnts$lagoslakeid)),
                 function(i){
                   pb$tick(tokens = list(llid = unique(pnts$lagoslakeid)[i]))
 
-                  # i <- 3
+                  # i <- which(unique(pnts$lagoslakeid) == 56028)
                   fname <- paste0("data/me_bathy/", unique(pnts$lagoslakeid)[i], ".tif")
                   if(!file.exists(fname)){
                     dt <- dplyr::filter(pnts, lagoslakeid == unique(pnts$lagoslakeid)[i])
@@ -75,7 +80,7 @@ rsubs <- lapply(seq_along(unique(pnts$lagoslakeid)),
                   list(r = res$r, width = res$wh)
                 })
 # whs          <- unlist(lapply(rsubs, function(x) x$width))
-# flist        <- list.files("data/ma_bathy/", pattern = "\\d.tif",
[…]

### 2020-02-06 manuscript/data.Rmd
split data viz and ms viz, closes #22

--- 2020-01-17 manuscript/figures.Rmd
@@
 ```
-
-\pagebreak
 
@@
 
-\pagebreak
-
-```{r heatmap, fig.cap = "Correlation matrix heatmap.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_heatmap-1.pdf")
-```
-
-![Lake shape caption](../figures/lake_shape.pdf){width=100%}
-
-```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_hypsography-1.pdf")
-```
-
 ```{r contrasts_depth, fig.cap = "Lake maximum depth by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/01_contrasts_depth-1.pdf")
-```
-
-```{r glaciated_vs_shapeclass, fig.cap = "Cross tabulation between lake glaciation and shape class.
-Circles are proportional to the number of lakes in each category.
-If a lake has been glaciated it is more likely to be convex.
-However, if a lake has not been glaciated there is no tendencey towards a particular shape class.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/glaciated_vs_shapeclass-1.pdf")
 ```
@@
 ```
-
-```{r 01_predictors, results='asis', warning=FALSE, echo = FALSE}
-knitr::kable('\\includegraphics[width=6.5in]{"../tables/01_predictors"}', col.names = NULL, caption = "Summary of lake depth predictor variables.")
-```
-
-```{r 02_depth_model_fitted, fig.cap = "A model that does not include shape class over predicts depth in concave lakes but does not systematically under predict depth in convex lakes.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/02_depth_model_fitted-1.pdf")
-```
-
-![Lake shape caption](../figures/lake_shape.pdf){width=100%}
-
-```{r 02_hypso_model_fitted, fig.cap = "Confusion matrix comparing two classification methods", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/02_hypso_model_fitted-1.pdf")
-```
-
-```{r 02_hypso_model_importance, fig.cap = "Recusive Partitioning variable importance.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/02_hypso_model_importance-1.pdf")
-```
-
-![Messager et al. 2016 equation table](../figures/messager_eq_table.png){width=100%}
-
-```{r 01_geometry_points, fig.cap = "Caption.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_geometry-1.pdf")
-```
-
-```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_geometry-2.pdf")
-```
-
-![Slope diagram caption](../figures/slope_diagram.pdf){width=100%}

### 2020-02-06 manuscript/figures.Rmd
split data viz and ms viz, closes #22

@@
 
-\pagebreak
-
-```{r maps_dot, fig.cap = "Map of lake maximum depth in the continental United States.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/00_map-1.pdf")
-```
-
 ![Summary of lake depth predictor variables.](../tables/01_predictors.pdf){width=65%}
@@
 ```
-
-```{r maps_state_all, fig.cap = "Proportion of ALL lakes greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/00_map_state-1.pdf")
-```
-
-```{r maps_state_limno, fig.cap = "Proportion of lakes with limno data greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/00_map_state-2.pdf")
-```
-
-```{r cutoffs, fig.cap = "Proportion max depth availability by lake area class.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/00_cutoffs-1.pdf")
-```
-
-\pagebreak
 

### 2020-02-06 figures/01_contrasts.Rmd
expand y axes

@@
    geom_text(aes(label = n), vjust = -0.5) + 
-   ylim(0, 11000) +
+   ylim(0, 13000) +
   xlab("Program type")
@@
   geom_text(aes(label = n), vjust = -0.5) +
-  ylim(0, 10000) + 
+  ylim(0, 11000) + 
   xlab("Lake area class (ha)")
@@
   geom_text(aes(label = n), vjust = -0.5) +
-  ylim(0, 3000) + 
+  ylim(0, 4500) + 
   xlab("Shape class")

### 2020-02-06 figures/00_maps.Rmd
rm duplicate labels

@@
 gg_states
-
 ```
@@
             by = c("lake_state" = "postal")) %>%
-  dplyr::filter(!is.na(prop_maxdepth))
+  dplyr::filter(!is.na(prop_maxdepth)) %>%
+  distinct(lake_state, .keep_all = TRUE)
 st_geometry(dt_plot) <- dt_plot$geom
@@
   group_by(lake_state) %>%
-  summarize(prop_maxdepth = mean(!is.na(lake_maxdepth_m))) %>%
+  summarize(prop_maxdepth = round(mean(!is.na(lake_maxdepth_m)), 2)) %>%
   arrange(desc(prop_maxdepth)) %>%
@@
   geom_sf_text(data = dt_plot,
-               aes(label = round(prop_maxdepth, 2)), 
+               aes(label = prop_maxdepth),
                size = 2.5) +

### 2020-02-06 figures/00_qa.Rmd
add figure showing variation among multiple data sources, closes #17

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 4, fig.height = 4)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 00_qa, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=2.5}
# variation among measurements for the same lake

dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  group_by(lagoslakeid) %>%
  add_tally() %>%
  dplyr::filter(n > 1) %>%
  # dplyr::filter(lagoslakeid == 5426)
  summarize(# lake_maxdepth_std = sd(lake_maxdepth_m), 
         lake_maxdepth_range = range(lake_maxdepth_m)[2] - 
           range(lake_maxdepth_m)[1], 
         lake_maxdepth_prop = lake_maxdepth_range / range(lake_maxdepth_m)[2]) %>%
  setNames(c("lagoslakeid", 
             "Range", "Range / largest"))
# View(dplyr::arrange(dt, desc(Range)))
# plot(dt$Range, dt$`Range / largest`)

ggplot(data = tidyr::pivot_longer(dt, -lagoslakeid)) +
  geom_histogram(aes(x = value)) +
  facet_wrap(~name, scales = "free") +
  xlab("meters")
```

```{r lgne-area_vs_lgus-area, eval=FALSE}
lg    <- lagosus_load("locus")
lg_ne <- lagosne_load()
dt    <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)

names(lg$locus$locus_link$l)
res <- dt %>% 
  left_join(
    dplyr::select(lg$locus$locus_link, lagoslakeid, lagosne_lagoslakeid)) %>%
  left_join(
    dplyr::select(lg$locus$locus_characteristics, 
                  lagoslakeid, lake_totalarea_ha)) %>%
  dplyr::filter(!is.na(lagosne_lagoslakeid)) %>%
  dplyr::filter(lagoslakeid == lagosne_lagoslakeid) %>%
  dplyr::filter(!is.na(lake_maxdepth_m)) %>%
  distinct(lagoslakeid, .keep_all = TRUE) %>%
  dplyr::select(lagoslakeid, lagosne_lagoslakeid, lake_waterarea_ha, 
                lake_totalarea_ha, lake_maxdepth_m) %>%
  left_join(dplyr::select(lg_ne$locus, lagosne_lagoslakeid = lagoslakeid, 
                          lake_area_ha)) %>%
  mutate(diff = lake_area_ha - lake_waterarea_ha) %>%
  arrange(desc(diff))

res2 <- res %>%
  dplyr::filter(abs(diff) < 0.1)
hist(res2$diff)

res3 <- res %>%
  dplyr::filter(abs(diff) > 1)
i <- 2
mapview(LAGOSUSgis::query_lake_poly(1222))
dplyr::filter(lg$locus$locus_characteristics, lagoslakeid == 14794)
dplyr::filter(dt, lagoslakeid == 14794)
dplyr::filter(lg_ne$locus, lagoslakeid == 14794)

+
  mapview(LAGOSNEgis::query_gis(
    "LAGOS_NE_All_Lakes_4ha", "lagoslakeid", res3$lagoslakeid[i]))

mapview(LAGOSNEgis::query_gis(
    "LAGOS_NE_All_Lakes_4ha", "lagoslakeid", 61727))

```

```{r misc, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
threshold <- 5

dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  dplyr::filter(!is.na(lake_maxdepth_m)) %>%
  group_by(lagoslakeid) %>%
  add_tally() %>%
  dplyr::filter(n >= 2) %>%
  mutate(diff = (max(lake_maxdepth_m) - min(lake_maxdepth_m)) / 
           max(lake_maxdepth_m)) %>%
  arrange(desc(diff))

%>%
  dplyr::filter(!(lagos_effort == "NLA" & diff > threshold)) %>%
  dplyr::filter(!(lagos_effort == "LAGOSNE" & diff > threshold))

dt %>% View()
distinct(dt, lagoslakeid, diff) %>%
  pull(diff) %>% hist(n = 100)

mapview(LAGOSUSgis::query_lake_poly(365518))

r <- raster("../data/nh_bathy/5787.tif")

ggplot() + geom_blank() + 
  theme(panel.background = element_blank())
LAGOSNE::lake_info(c(5592, 5787)) %>%
  dplyr::select(lagoslakeid, lagosname1, maxdepth, gnis_name, 
                lake_area_ha) %>%
  t() %>%
  gridExtra::grid.table()

hist(dt$diff, xlim = c(0, 10), n = 100)
median(dt$diff, na.rm = TRUE)

res_all <- read.csv("../data/00_bathy_depth/bathy_geometry.csv",
                    stringsAsFactors = FALSE)
dt_raw_ne <- read.csv("../data/lagosne_depth_predictors.csv",
                      stringsAsFactors = FALSE) %>%
  mutate(llid = as.character(lagoslakeid))
test      <- left_join(res_all, dt_raw_ne) %>%
  data.frame()

test$calc_depth_lgne <- calc_depth(test$buffer100m_slope_max, test$dist_deepest,
                                  grain = 10)
test$calc_depth_lgus <- calc_depth(test$inlake_slope, test$dist_deepest)
plot(test$maxdepth, test$calc_depth_lgus)
# plot(test$lake_maxdepth_m, test$calc_depth_lgus)
# abline(0, 1)
#
# ggplot(data = test) +
#   geom_point(aes(x = lake_maxdepth_m, y = calc_depth_lgus)) +
#   geom_abline(aes(slope = 1, intercept = 0)) +
#   facet_wrap(~state)

# plotly::ggplotly())

#
# hist(test$inlake_slope)
# hist(test$dist_deepest)
# hist(test$calc_depth_lgus)
#
# plot(test$calc_depth_lgne, test$calc_depth_lgus)
# plot(test$lake_maxdepth_m, test$calc_depth_lgne)

#
# # 2654
#
#
#
# arrange(res, desc(dist_between)) %>%
#   View()
#
# hist(res$dist_between)
# plot(res$dist_deepest, res$dist_viscenter)
# abline(0, 1)

# pnt_surface  <- st_point_on_surface(dt_poly)
## point furthest from shore
# test <- as_Spatial(dt_poly)
# r <- rasterize(SpatialPolygons(test@polygons), dt)
# r[is.na(r)] <- 0
# r[r == 1] <- NA
# test2 <- distance(r)
# which.max(test2)
# xy           <- xyFromCell(dt, which.max(test2[]))
# pnt_furthest  <- st_sfc(st_multipoint(xy), crs = st_crs(dt))
# pnt_centroid <- st_centroid(dt_poly)

# dt           <- raster("data/mn_bathy/2419.tif")
# compare_deepest_center(dt)
# mapview::mapview(dt_poly) +
#   mapview::mapview(pnt_viscenter) +
# mapview::mapview(r) +
#   mapview::mapview(pnt_deepest)

```

### 2020-02-06 figures/00_maps.Rmd
clean up bathy map

@@
 dt_bathy <- st_sf(dt_bathy, geometry = dt_bathy$geom)
+dt_bathy <- mutate(dt_bathy, maxdepth_class = droplevels(dt_bathy$maxdepth_class)) 
+dt_bathy <- dplyr::filter(dt_bathy, !is.na(maxdepth_class))
 
@@
   ggplot() + 
-    geom_sf(data = study_bbox[2,], alpha = 1, size = 0.6, 
+    geom_sf(data = st_as_sfc(st_bbox(dt_bathy)), alpha = 1, size = 0.6, 
           color = "black", fill = "#1D9DCF") +
-    geom_sf(data = states_all, fill = "gray", color = "black", size = 0.3) +
-    geom_sf(data = states_focal, fill = "white", color = "black", size = 0.4) + 
+    geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))), fill = "gray", color = "black", size = 0.3) +
+    geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))), fill = "white", color = "black", size = 0.4) + 
     geom_sf(data = dt_bathy, aes(color = maxdepth_class), 
-                    size = 0.3, alpha = 1) +
-    scale_color_manual(values = viridisLite::viridis(5)) +
+                    size = 0.2, alpha = 0.8) +
+    scale_color_manual(values = viridisLite::viridis(4)) +
     labs(color = "Max depth (m)") +
-    coord_sf(datum =  NA) +
+    coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
     theme_void()

### 2020-02-10 tables/00_data.Rmd
add summary table to data doc

---
output:
  pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
# setwd("figures")
```

```{r summary_table, eval=TRUE}
library(kableExtra)
suppressMessages(library(dplyr))
library(magrittr)

# setwd("tables")
dt <- read.csv("../data/lagosus_depth.csv", 
               stringsAsFactors = FALSE) %>%
  distinct(lagoslakeid, .keep_all = TRUE) %>%
  dplyr::select(-contains("lagoslakeid"), 
                -contains("decdeg"), 
                -contains("waterarea"), 
                -has_limno) %>%
  tidyr::gather(key = "variable") %>%
  group_by(variable) %>%
  mutate(value = as.numeric(value)) %>%
  mutate(q10 = quantile(value, probs = c(0.1), na.rm = TRUE),
         q25 = quantile(value, probs = c(0.25), na.rm = TRUE),
         q50 = quantile(value, probs = c(0.5), na.rm = TRUE),
         q75 = quantile(value, probs = c(0.75), na.rm = TRUE),
         q90 = quantile(value, probs = c(0.9), na.rm = TRUE), 
         ratio = q90 / q10, 
         n = sum(!is.na(value), na.rm = TRUE)) %>%
  dplyr::filter(!is.na(q50)) %>%
  dplyr::distinct(signif(q10, 2), signif(q25, 2), signif(q50, 2), signif(q75, 2), signif(q90, 2), ratio, n) %>%
  dplyr::mutate_if(is.numeric, funs(as.character(signif(., 3)))) %>%
  dplyr::select(variable, 
         q10 = `signif(q10, 2)`,
         q25 = `signif(q25, 2)`,
         q50 = `signif(q50, 2)`,
         q75 = `signif(q75, 2)`,
         q90 = `signif(q90, 2)`,
         q10_q90_ratio = ratio,
         n)

dt <- ungroup(dt)
dt <- dplyr::select(dt, variable, # Variable = pretty,
                    Median = q50, Q25 = q25, Q75 = q75, n)

knitr::kable(dt, digits = 0, format = "markdown") 
#%>%
 # add_header_above(c(" ", " ", Ranges = 3, " ", " "))
```

### 2020-02-11 figures/00_qa.Rmd
create a qa plot for mean depth

@@
 
+```{r lgnemanual-bathy-meandepth}
+dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)
+dt_sub <- dt %>%
+  dplyr::filter(
+    lagos_effort %in% c("LAGOSNE", "bathymetry", "manual") &
+    !is.na(lake_meandepth_m)) %>%
+  dplyr::select(lagoslakeid, lake_state, lagos_effort, lake_meandepth_m) %>%
+  tidyr::pivot_wider(names_from = lagos_effort, 
+                     values_from = c(lake_meandepth_m)) %>%
+  dplyr::filter(!is.na(bathymetry)) %>%
+  mutate(combined = coalesce(manual, LAGOSNE)) %>%
+  dplyr::filter(!is.na(combined))
+
+(gg <- ggplot(data = dplyr::filter(dt_sub, !(lake_state %in% c("ME", "NH")))) +
+  geom_text(aes(x = bathymetry, y = combined, color = lake_state, 
+                 label = lagoslakeid)) +
+  geom_abline(aes(intercept = 0, slope = 1)))
+
+# plotly::ggplotly(gg)
+```
+
 ```{r lgne-area_vs_lgus-area, eval=FALSE}
@@
 
-names(lg$locus$locus_link$l)
 res <- dt %>% 

### 2020-02-11 scripts/00_get_hypso_me.R

@@ -49,7 +49,7 @@ rsubs <- lapply(seq_along(unique(pnts$lagoslakeid)),
                 function(i){
                   pb$tick(tokens = list(llid = unique(pnts$lagoslakeid)[i]))
 
-                  # i <- which(unique(pnts$lagoslakeid) == 8026)
+                  # i <- which(unique(pnts$lagoslakeid) == 8511)
                   fname <- paste0("data/me_bathy/", unique(pnts$lagoslakeid)[i], ".tif")
                   if(!file.exists(fname)){
                     dt <- dplyr::filter(pnts, lagoslakeid == unique(pnts$lagoslakeid)[i])

### 2020-02-11 manuscript/data.Rmd
show qa fig
add summary table to data doc
test fig size caption, closes #25

@@
 knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
+knitr::opts_knit$set(eval.after = "fig.cap")
 ```
 
-```{r maps_dot, fig.cap = "Map of lake maximum depth in the continental United States.", warning=FALSE, echo = FALSE, eval=TRUE}
+\pagebreak
+
+![Summary of lake depth measurements](../tables/00_data.pdf){width=65%}
+
+```{r maps_dot, fig.cap = paste0("Map of lake maximum depth in the continental United States (", paste0(system("fsizemb ../figures/00_map-1.pdf", intern = TRUE), " MB)")), warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/00_map-1.pdf")
+# 
 ```
@@
 ```
+
+```{r 00_qa, fig.cap = "Range of depth measurements for lakes with data from multiple sources.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/00_qa-1.pdf")
+```

### 2020-02-11 figures/00_qa.Rmd
add manual vs bathy panels to makefile

@@
 
-```{r lgnemanual-bathy-meandepth}
+```{r lgnemanual-vs-bathy-depth, fig.height=3}
 dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)
+
 dt_sub <- dt %>%
@@
 
-(gg <- ggplot(data = dplyr::filter(dt_sub, !(lake_state %in% c("ME", "NH")))) +
-  geom_text(aes(x = bathymetry, y = combined, color = lake_state, 
-                 label = lagoslakeid)) +
-  geom_abline(aes(intercept = 0, slope = 1)))
+gg_mean <- ggplot() +
+  geom_point(data = dt_sub, 
+             aes(x = bathymetry, y = combined, color = lake_state), 
+             alpha = 0.6) +
+  geom_point(data = dplyr::filter(dt_sub, !(lake_state %in% c("ME"))), 
+             aes(x = bathymetry, y = combined, color = lake_state)) +
+  geom_abline(aes(intercept = 0, slope = 1)) +
+  xlab("Calculated (m)") + ylab("Reported (m)") +
+  ggtitle("Mean Depth")
 
 # plotly::ggplotly(gg)
+
+dt_sub <- dt %>%
+  dplyr::filter(
+    lagos_effort %in% c("LAGOSNE", "bathymetry", "manual") &
+    !is.na(lake_maxdepth_m) &
+    !(lagoslakeid %in% c(456169))) %>% # remove llids with duplicate manual depth
+  dplyr::select(lagoslakeid, lake_state, lagos_effort, lake_maxdepth_m) %>%
+  tidyr::pivot_wider(names_from = lagos_effort, 
+                     values_from = c(lake_maxdepth_m)) %>%
+  dplyr::filter(!is.na(bathymetry)) %>%
+  mutate(combined = coalesce(manual, LAGOSNE)) %>%
+  dplyr::filter(!is.na(combined))
+
+gg_max <- ggplot(data = dt_sub) +
+  # geom_text(aes(x = bathymetry, y = combined, color = lake_state, 
+  #                label = lagoslakeid)) +
+  geom_point(aes(x = bathymetry, y = combined, color = lake_state)) +
+  geom_abline(aes(intercept = 0, slope = 1)) +
+  xlab("Calculated (m)") + ylab("Reported (m)") +
+  ggtitle("Max Depth")
+
+cowplot::plot_grid(gg_max + theme(legend.position = "none"), 
+                   gg_mean + theme(axis.title.y = element_blank()), 
+                   rel_widths = c(0.7, 1))
 ```

### 2020-02-11 manuscript/figures.Rmd
rm programs tally
add manual vs bathy panels to makefile

@@
 
-![Summary of lake depth predictor variables.](../tables/01_predictors.pdf){width=65%}
+\pagebreak
+
+![Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right)](../figures/lake_shape.pdf){width=100%}
 
@@
 
+```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported max depth and max depth calculated from bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/lgnemanual-vs-bathy-depth-1.pdf")
+```
+
+![Summary of lake depth predictor variables.](../tables/01_predictors.pdf){width=65%}
+
 ```{r heatmap, fig.cap = "Correlation matrix heatmap.", warning=FALSE, echo = FALSE, eval=TRUE}
@@
 ```
-
-![Lake shape caption](../figures/lake_shape.pdf){width=100%}
 
@@
 
-```{r 01_programs_tally, fig.cap = "Number of programs by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_programs_tally-1.pdf")
-```
-
 ```{r 02_depth_model_fitted, fig.cap = "A model that does not include shape class over predicts depth in concave lakes but does not systematically under predict depth in convex lakes.", warning=FALSE, echo = FALSE, eval=TRUE}

### 2020-02-11 figures/00_maps.Rmd
create a colocated map

@@
 
-```{r 00_map_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r gg_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
 bathy_pnts <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
@@
 max_depth <- dt$lake_maxdepth_m
+states_all    <- st_read("../data/gis.gpkg", "states_all", quiet = TRUE)
+states_focal  <- st_read("../data/gis.gpkg", "states_focal", quiet = TRUE)
 dt <- dt %>%
@@
 ```
+
+```{r gg_colocated, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+is_colocated <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
+  dplyr::filter(
+    lagos_effort %in% c("LAGOSNE", "bathymetry", "manual") &
+    !is.na(lake_maxdepth_m) &
+    !(lagoslakeid %in% c(456169))) %>% # remove llids with duplicate manual depth)
+  dplyr::select(lagoslakeid, lake_state, lagos_effort, lake_maxdepth_m) %>%
+  tidyr::pivot_wider(names_from = lagos_effort, 
+                     values_from = c(lake_maxdepth_m)) %>%
+  dplyr::filter(!is.na(bathymetry)) %>%
+  mutate(combined = coalesce(manual, LAGOSNE)) %>%
+  dplyr::filter(!is.na(combined)) %>%
+  dplyr::filter(!is.na(bathymetry) & !is.na(combined)) %>%
+  pull(lagoslakeid)
+
+dt_bathy <- mutate(dt_bathy, 
+               colocated = case_when(llid %in% is_colocated ~ "colocated", 
+                                     TRUE ~ "bathymetry only")) %>%
+  arrange(desc(colocated))
+
+gg_colocated <- ggplot() + 
+    geom_sf(data = st_as_sfc(st_bbox(dt_bathy)), alpha = 1, size = 0.6, 
+          color = "black", fill = "#1D9DCF") +
+    geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))), fill = "gray", color = "black", size = 0.3) +
+    geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))), fill = "white", color = "black", size = 0.4) + 
+    geom_sf(data = dt_bathy, aes(color = colocated),
+                    size = 0.2, alpha = 0.8) +
+    coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
+    theme_void()
+
+gg_colocated
+
+# gg_shapeclass <-
+
+```

### 2020-02-11 manuscript/figures.Rmd
adjust fig dimensions

@@
 
-![Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right)](../figures/lake_shape.pdf){width=100%}
+![Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right)](../figures/lake_shape.pdf){width=85%}
 

### 2020-02-13 figures/00_maps.Rmd
add shape class panel to map

@@
 gg_colocated
+```
 
-# gg_shapeclass <-
+```{r gg_shapeclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+dt           <- st_read("../data/gis.gpkg", "dt", quiet = TRUE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  distinct(lagoslakeid, .keep_all = TRUE)
+dt_withshape <- read.csv("../data/lagosus_depth_predictors.csv", 
+               stringsAsFactors = FALSE) %>%
+  dplyr::distinct(lagoslakeid, shape_class) %>%
+  dplyr::filter(!is.na(shape_class)) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid))
+dt           <- inner_join(dt, dt_withshape) %>%
+  arrange(desc(shape_class))
 
+gg_shapeclass <- ggplot() + 
+    geom_sf(data = st_as_sfc(st_bbox(dt_bathy)), alpha = 1, size = 0.6, 
+          color = "black", fill = "#1D9DCF") +
+    geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))), fill = "gray", color = "black", size = 0.3) +
+    geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))), fill = "white", color = "black", size = 0.4) + 
+    geom_sf(data = dplyr::filter(dt, shape_class == "convex"), aes(color = shape_class),
+                    size = 0.03, alpha = 0.3) +
+    geom_sf(data = dplyr::filter(dt, shape_class != "convex"), aes(color = shape_class),
+                    size = 0.07, alpha = 0.8) +
+    labs(color = "Shape class") +
+    coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
+    theme_void()
+
+gg_shapeclass
+
+# table of shape_class by state
+# st_drop_geometry(dt) %>%
+#   group_by(lake_state, shape_class) %>%
+#   count() %>%
+#   tidyr::pivot_wider(names_from = shape_class, values_from = n)
 ```
+
+```{r 00_map_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+cowplot::plot_grid(gg_bathy, 
+                   gg_shapeclass, ncol = 1)
+```

### 2020-02-13 manuscript/figures.Rmd
add shape class panel to map

@@
 
-```{r maps_dot_bathy, fig.cap = "Map of lake maximum depth measurements derived from lake bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r maps_dot_bathy, fig.cap = "Map of lake maximum depth measurements calculated from lake bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/00_map_bathy-1.pdf")

### 2020-02-13 figures/00_maps.Rmd
adjust fig dimensions

--- 2020-02-11 figures/00_maps.Rmd
@@
 gg_colocated
+```
 
-# gg_shapeclass <-
+```{r gg_shapeclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+dt           <- st_read("../data/gis.gpkg", "dt", quiet = TRUE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  distinct(lagoslakeid, .keep_all = TRUE)
+dt_withshape <- read.csv("../data/lagosus_depth_predictors.csv", 
+               stringsAsFactors = FALSE) %>%
+  dplyr::distinct(lagoslakeid, shape_class) %>%
+  dplyr::filter(!is.na(shape_class)) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid))
+dt           <- inner_join(dt, dt_withshape) %>%
+  arrange(desc(shape_class))
 
+gg_shapeclass <- ggplot() + 
+    geom_sf(data = st_as_sfc(st_bbox(dt_bathy)), alpha = 1, size = 0.6, 
+          color = "black", fill = "#1D9DCF") +
+    geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))), fill = "gray", color = "black", size = 0.3) +
+    geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))), fill = "white", color = "black", size = 0.4) + 
+    geom_sf(data = dplyr::filter(dt, shape_class == "convex"), aes(color = shape_class),
+                    size = 0.03, alpha = 0.3) +
+    geom_sf(data = dplyr::filter(dt, shape_class != "convex"), aes(color = shape_class),
+                    size = 0.07, alpha = 0.8) +
+    labs(color = "Shape class   ") +
+    coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
+    theme_void()
+
+gg_shapeclass
+
+# table of shape_class by state
+# st_drop_geometry(dt) %>%
+#   group_by(lake_state, shape_class) %>%
+#   count() %>%
+#   tidyr::pivot_wider(names_from = shape_class, values_from = n)
 ```
+
+```{r 00_map_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6}
+cowplot::plot_grid(gg_bathy, 
+                   gg_shapeclass, 
+                   ncol = 1)
+```

### 2020-02-13 figures/00_qa.Rmd
adjust fig dimensions

@@
 
-```{r lgnemanual-vs-bathy-depth, fig.height=3}
+```{r lgnemanual-vs-bathy-depth, fig.height=2.7, fig.width=6}
 dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)
@@
                    gg_mean + theme(axis.title.y = element_blank()), 
-                   rel_widths = c(0.7, 1))
+                   rel_widths = c(0.8, 1))
 ```

### 2020-02-13 manuscript/figures.Rmd
reorder figs

@@
 
+![Diagram showing the assumption of most lake depth models that the nearshore slope is extends into the lake and that all lakes have a common relationship between nearshore and in-lake slope.
+](../figures/slope_diagram.pdf){width=90%}
+
 ![Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right)](../figures/lake_shape.pdf){width=85%}
+
+```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported max depth and max depth calculated from bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE}
+# see 00_qa.Rmd
+knitr::include_graphics("../figures/lgnemanual-vs-bathy-depth-1.pdf")
+```
 
@@
 
-```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported max depth and max depth calculated from bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/lgnemanual-vs-bathy-depth-1.pdf")
+```{r 01_geometry_points, fig.cap = "Distance to deepest point versus distance to lake visual center.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_geometry-1.pdf")
 ```
@@
 
-```{r 01_geometry_points, fig.cap = "Distance to deepest point versus distance to lake visual center.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_geometry-1.pdf")
-```
-
 \pagebreak
@@
 ```
-
-![Slope diagram caption](../figures/slope_diagram.pdf){width=100%}

### 2020-02-14 scripts/00_get_nearshore.R
extract online slopes in a buffer

@@ -11,12 +11,11 @@ dt         <- read.csv("data/lagosus_depth.csv",
 bathy_pnts <- readRDS("data/00_bathy_depth/bathy_pnts.rds")
 dt_pred    <- read.csv("data/lagosne_depth_predictors.csv",
                     stringsAsFactors = FALSE) %>%
-  dplyr::filter(lagoslakeid %in% bathy_pnts$llid &
-                  !is.na(shape_class))
+  dplyr::filter(lagoslakeid %in% bathy_pnts$llid)
 
 # attempt at manual calculation of max buffer slope
 
-ll_ids            <- sample(dt_pred$lagoslakeid, 60)
+ll_ids            <- sample(dt_pred$lagoslakeid, 80)
 flist             <- list.files("data/elevatr", pattern = "\\d*.tif",
                     full.names = TRUE, include.dirs = TRUE)
 existing_surfaces <- gsub(".tif", "",
@@ -73,8 +72,9 @@ get_slope <- function(ll_id){
                                      st_transform(ll_poly_hull, st_crs(r_crs)))
   buffer_line  <- sf::st_nearest_points(shore_pnt,
                                         st_transform(ll_buff_hull, st_crs(r_crs)))
+  buffer_line <- st_buffer(buffer_line, 20)
 
-  # mapview(elev) + mapview(ll_poly_hull) + mapview(ll_buff_hull) + mapview(buffer_line)
+  # mapview(elev) + mapview(ll_poly_hull) + mapview(ll_buff_hull) + mapview(buffer_line) + mapview(st_buffer(buffer_line, 20))
 
   slope_online        <- raster::extract(slope, as_Spatial(buffer_line))[[1]]
   slope_online_mean   <- mean(slope_online, na.rm = TRUE) * res(elev)[1]

### 2020-02-14 figures/01_geometry.Rmd
compare slopes with a heatmap

@@
 
-```{r }
-# compare inlake slope against buffer slope
+```{r slope-compare, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
+# compare inlake slope, buffer slope, and online slope
+# setwd("figures") 
+bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
+dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
+  mutate(llid = as.character(llid))
+dt <- left_join(dt, st_drop_geometry(bp), 
+                by = "llid") %>%
+  dplyr::filter(!is.na(inlake_slope_mean))
+
+jsta::jheatmap(dt)
+jsta::jheatmap(
+  dplyr::select(dt, 
+                slope_online_mean,
+                inlake_slope_mean,
+                slope_mean, 
+                slope_mean_ne,
+                dist_deepest, 
+                dist_viscenter, 
+                maxdepth,
+                meandepth))
+
+plot(dt$inlake_slope_mean, dt$slope_online_mean)
+plot(dt$inlake_slope_mean, dt$slope_mean)
+plot(dt$inlake_slope_mean, dt$slope_mean_ne)
+plot(dt$slope_mean, dt$slope_mean_ne)
+dt$res_elev_1
+plot(dt$inlake_slope_median, dt$slope_online_median)
+
+plot(dt$inlake_slope, dt$slope_mean_ne)
+plot(dt$inlake_slope, dt$slope_online)
+plot(dt$slope_mean, dt$slope_mean_ne)
+
+plot(dt$inlake_slope, dt$buffer100m_slope_mean/10, 
+     xlim = c(0, 3), ylim = c(0, 3))
+abline(0, 1)
 

### 2020-02-17 figures/01_geometry.Rmd
add slope comparison figure

--- 2020-01-31 figures/01_geometry.Rmd
@@
 
-```{r }
-# compare inlake slope against buffer slope
+```{r slope-compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+# compare inlake slope, buffer slope, and online slope
+# setwd("figures") 
+bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
+dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
+  mutate(llid = as.character(llid))
+dt <- left_join(dt, st_drop_geometry(bp), 
+                by = "llid") %>%
+  dplyr::filter(!is.na(inlake_slope_mean))
 
-dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
-hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
-                       stringsAsFactors = FALSE) %>%
+# jsta::jheatmap(dt)
+jsta::jheatmap(
+  dplyr::select(dt, 
+                slope_online_mean,
+                inlake_slope_mean,
+                slope_mean, 
+                slope_mean_ne,
+                dist_deepest, 
+                dist_viscenter, 
+                maxdepth,
+                meandepth))
+
+GGally::ggpairs(dplyr::select(dt, 
+                inlake_slope_mean, inlake_slope.y,
+                slope_online_mean,
+                slope_mean, 
+                # dist_deepest, 
+                # dist_viscenter, 
+                maxdepth), 
+                progress = FALSE, 
+                ) +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1), 
+          strip.text.y = element_text(angle = 0))
+
+gg_scatter_slope <- function(x, y, dt){
+  ggplot(data = dt) + 
+    geom_point(aes(UQ(rlang::sym(x)), UQ(rlang::sym(y))))
+}
+
+# dt[22,]
+# which(dt$inlake_slope.y > 0.4)
+# which(dt$inlake_slope.x > 0.4)
+dt <- dplyr::filter(dt, inlake_slope.y < 0.4)
+
+cowplot::plot_grid(
+  gg_scatter_slope("slope_mean", "maxdepth", dt),
+  gg_scatter_slope("inlake_slope.y", "maxdepth", dt), 
+  gg_scatter_slope("slope_mean", "inlake_slope.y", dt)
+  )
+```
+
+```{r buffer-vs-inlake_slope, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4}
+lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
   mutate(lagoslakeid = as.character(lagoslakeid))
-dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, 
-                                  buffer100m_slope_mean), 
-                                  by = c("llid" = "lagoslakeid"))
+test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>%
+  dplyr::filter(shape_class != "neither")
 
-plot(dt$inlake_slope, dt$buffer100m_slope_mean/10, 
-     xlim = c(0, 3), ylim = c(0, 3))
-abline(0, 1)
-
-# View(arrange(dt, buffer100m_slope_mean))
-
+cowplot::plot_grid(
+ggplot(data = test) + 
+    geom_point(aes(slope_mean, inlake_slope.y, color = reservoir_class)) +
+  xlab("Buffer slope (mean)") +
+  ylab("In-lake slope") +
+  labs(color = "") + 
+  theme(axis.title.x = element_blank()),
+ggplot(data = test) + 
+    geom_point(aes(slope_mean, inlake_slope.y, color = shape_class)) +
+  xlab("Buffer slope (mean)") +
+  ylab("In-lake slope") +
+  labs(color = ""), 
+nrow = 2 
+)
 ```

### 2020-02-18 figures/00_maps.Rmd
embiggen legend points, closes #30

@@
 ```{r 00_map_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6}
-cowplot::plot_grid(gg_bathy, 
-                   gg_shapeclass, 
+cowplot::plot_grid(gg_bathy + 
+                     guides(
+                       color = guide_legend(
+                         override.aes = list(size = 2))), 
+                   gg_shapeclass + 
+                     guides(
+                       color = guide_legend(
+                         override.aes = list(size = 2))), 
                    ncol = 1)

### 2020-02-18 figures/01_geometry.Rmd
create grid of slope and distance comparisons, closes #29

--- 2020-01-31 figures/01_geometry.Rmd
@@
   mutate(lagoslakeid = as.character(lagoslakeid))
-dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class), 
+dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, shape_class), 
                                   by = c("llid" = "lagoslakeid"))
@@
   geom_abline(aes(intercept = 0, slope = 1)) +
-  xlab("Distance to shore from deepest point") +
-  ylab("Distance to shore from visual center")
+  xlab("Deepest point distance (m)") +
+  ylab("Visual center distance (m)")
+
+gg_scatter_shape <- st_drop_geometry(dt) %>%
+  dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::arrange(desc(factor(shape_class, 
+                        levels = c("neither", "concave", "convex")))) %>%
+  ggplot() + 
+  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = shape_class)) +
+  geom_abline(aes(intercept = 0, slope = 1)) +
+  labs(color = "") +
+  xlab("Deepest point distance (m)") +
+  ylab("Visual center distance (m)")
 
@@
   geom_abline(aes(intercept = 0, slope = 1)) +
-  xlab("Distance to shore from deepest point") +
-  ylab("Distance to shore from visual center")
+  xlab("Deepest point distance (m)") +
+  ylab("Visual center distance (m)")
+```
 
-cowplot::plot_grid(gg_scatter_state + 
+```{r gg_distance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3}
+gg_distance <- cowplot::plot_grid(gg_scatter_shape + 
                      theme(axis.title.x = element_text(color = "white")), 
@@
                    rel_widths = c(1, 0.8))
+gg_distance
 
@@
                       ymin = `5%`, ymax = `95%`)) +
-  ylab("Distance from shore to deepest point (m)") + xlab("")
+  ylab("Deepest point distance (m)") + xlab("")
 
@@
 
-```{r }
-# compare inlake slope against buffer slope
+```{r slope-compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}
+# compare inlake slope, buffer slope, and online slope
+# setwd("figures") 
+bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
+dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
+  mutate(llid = as.character(llid))
+dt <- left_join(dt, st_drop_geometry(bp), 
+                by = "llid") %>%
+  dplyr::filter(!is.na(inlake_slope_mean))
 
-dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
-hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
-                       stringsAsFactors = FALSE) %>%
+# jsta::jheatmap(dt)
+# jsta::jheatmap(dt, "inlake_slope.y")
+jsta::jheatmap(
+  dplyr::select(dt, 
+                slope_online_mean,
+                inlake_slope_mean,
+                slope_mean, 
+                slope_mean_ne,
+                dist_deepest, 
+                dist_viscenter, 
+                maxdepth,
+                meandepth))
+
+GGally::ggpairs(dplyr::select(dt, 
+                inlake_slope_mean, inlake_slope.y,
+                slope_online_mean,
+                slope_mean, 
+                # dist_deepest, 
+                # dist_viscenter, 
+                maxdepth), 
+                progress = FALSE, 
+                ) +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1), 
+          strip.text.y = element_text(angle = 0))
+
+gg_scatter_slope <- function(x, y, dt){
+  ggplot(data = dt) + 
+    geom_point(aes(UQ(rlang::sym(x)), UQ(rlang::sym(y))))
+}
+
+# dt[22,]
+# which(dt$inlake_slope.y > 0.4)
+# which(dt$inlake_slope.x > 0.4)
+dt <- dplyr::filter(dt, inlake_slope.y < 0.4)
+
+cowplot::plot_grid(
+  gg_scatter_slope("slope_mean", "maxdepth", dt),
+  gg_scatter_slope("inlake_slope.y", "maxdepth", dt), 
+  gg_scatter_slope("slope_mean", "inlake_slope.y", dt)
+  )
+```
+
+```{r buffer-vs-inlake_slope, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4}
+lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
   mutate(lagoslakeid = as.character(lagoslakeid))
-dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, 
-                                  buffer100m_slope_mean), 
-                                  by = c("llid" = "lagoslakeid"))
+test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
+  dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::arrange(desc(factor(shape_class, 
+                        levels = c("neither", "concave", "convex"))))
 
-plot(dt$inlake_slope, dt$buffer100m_slope_mean/10, 
-     xlim = c(0, 3), ylim = c(0, 3))
-abline(0, 1)
+# GGally::ggpairs(dplyr::select(test, matches("class")))
 
-# View(arrange(dt, buffer100m_slope_mean))
+gg_slope_shape <- ggplot(data = test, 
+                         aes(slope_mean, inlake_slope.y, 
+                             color = shape_class)) + 
+    geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  xlab("Buffer slope (mean)") +
+  ylab("In-lake slope") +
+  labs(color = "")
 
+gg_slope_reservoir <- ggplot(data = test, 
+                             aes(slope_mean, inlake_slope.y, 
+                                 color = reservoir_class)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  xlab("Buffer slope (mean)") +
+  ylab("In-lake slope") +
+  labs(color = "")
 ```
+
+```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5}
+theme_opts <- theme(legend.position = "none")
+cowplot::plot_grid(gg_scatter_shape + 
+                     labs(color = "") +
+                     # theme(axis.title.x = element_blank()) +
+                     theme_opts,
+                   gg_scatter_reservoir + 
+                     labs(color = "") +
+                     theme(#axis.title.x = element_blank(), 
+                           axis.title.y = element_blank()) +
+                     theme_opts,
+                   gg_slope_shape +
+                     theme_opts,
+                   gg_slope_reservoir + 
+                     theme(axis.title.y = element_blank()) +
+                     theme_opts, 
+                   rel_widths = c(1, 0.93, 1, 1))
+```

### 2020-02-18 manuscript/figures.Rmd
size adjustment
add slope comparison figure
create grid of slope and distance comparisons, closes #29

@@
 ![Diagram showing the assumption of most lake depth models that the nearshore slope is extends into the lake and that all lakes have a common relationship between nearshore and in-lake slope.
-](../figures/slope_diagram.pdf){width=90%}
+](../figures/slope_diagram.pdf){width=85%}
 
@@
 
-```{r 01_geometry_points, fig.cap = "Distance to deepest point versus distance to lake visual center.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_geometry-1.pdf")
+```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) buffer slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
 ```
@@
 ```{r constrasts_distance, fig.cap = "Contrasting distance from shore to the deepest point in the lake and distance to the lake visual center for natural lakes and reservoirs.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_geometry-2.pdf")
+knitr::include_graphics("../figures/gg_distance-2.pdf")
 ```
@@
 ```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_geometry-3.pdf")
+knitr::include_graphics("../figures/gg_distance-3.pdf")
 ```

### 2020-02-18 README.Rmd
link to tables not appendix
link fix
link to data summary pdf
rm badges

@@
 
-[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)]() [![Code DOI](https://zenodo.org/badge/123951266.svg)](https://zenodo.org/badge/latestdoi/123951266)
-
-### Data (see [data/lagosus_depth.csv]())
+### Data (see [data/lagosus_depth.csv](data/lagosus_depth.csv))
 
@@
 
+Data Summary: [manuscript/data.pdf](manuscript/data.pdf)
+
 Figures: [manuscript/figures.pdf](manuscript/figures.pdf)
 
-Appendix: [manuscript/appendix.pdf](manuscript/figures.pdf)
+Tables: [manuscript/tables.pdf](manuscript/tables.pdf)
 

### 2020-02-18 README.Rmd
rm abstract

@@
 Tables: [manuscript/tables.pdf](manuscript/tables.pdf)
-
-### Abstract
-
-Lake depth is widely regarded as a critical predictor of nutrient cycling in lakes.
-Unfortunately, existing databases lack widespread information on this basic measure.
-Lack of depth data is a key impediment to effective lake nutrient prediction in thousands of lakes at regional to continental scales.
-Although a number of investigators have developed predictive models of lake depth based on topography surrounding a lake, predictions from such models have proved to be fairly imprecise.
-To improve prediction of lake depth, we compiled a database of maximum depth for more than 15,000 lakes throughout the conterminous United States and used it to develop predictive models of lake depth based on factors such as lake elevation, topography, geometry, impoundment status, and landform.
-We used fitted models to predict lake depth for nearly all lakes in the conterminous United States and show how our statistical models, which are stratified on lake and watershed characteristics, represent an improvement over existing models.
-Ultimately, our lake depth predictions could be used not only as an aid in predicting lake nutrients but also as a rough check on estimates of total lake volume used in Earth System Models.

### 2020-02-19 figures/01_contrasts.Rmd
align source fields with data dictionary

@@
    dplyr::filter(!is.na(lake_maxdepth_m)) %>%
-   group_by(programtype_depth) %>% 
+   group_by(sourcetype_depth) %>% 
    tally() %>% 
-   ggplot(aes(x = programtype_depth, y = n)) + 
+   ggplot(aes(x = sourcetype_depth, y = n)) + 
    geom_col() + 
@@
    ylim(0, 13000) +
-  xlab("Program type")
+  xlab("Source type")
 ```
@@
    dplyr::filter(!is.na(lake_maxdepth_m)) %>%
-   group_by(programtype_depth, programlink_depth) %>% 
+   group_by(sourcetype_depth, sourceurl_depth) %>% 
    tally() %>% 
-   ungroup() %>% group_by(programtype_depth) %>%
+   ungroup() %>% group_by(sourcetype_depth) %>%
    tally() %>%
-   ggplot(aes(x = programtype_depth, y = n)) + 
+   ggplot(aes(x = sourcetype_depth, y = n)) + 
    geom_col() + 

### 2020-02-19 scripts/01_merge.R

@@ -13,14 +13,24 @@ lagosne_raw      <- read.csv(
 bathy_raw        <- read.csv(
   "data/00_bathy_depth/00_bathy_depth.csv", stringsAsFactors = FALSE)
 
-# quality assurance
+# merge and limit to 1ft min depth
 res_raw <- dplyr::bind_rows(manual_raw, manual_extra_raw,
                             nla_raw, lagosne_raw, bathy_raw) %>%
   dplyr::filter(is.na(max_depth_m) | is.na(mean_depth_m) |
                   max_depth_m != mean_depth_m) %>%
   dplyr::filter(is.na(max_depth_m) | is.na(mean_depth_m) |
                   max_depth_m > mean_depth_m) %>%
-  dplyr::filter(max_depth_m >= 0.3048 | is.na(max_depth_m)) %>%
+  dplyr::filter(max_depth_m >= 0.3048 | is.na(max_depth_m))
+
+# clean up source columns
+res_raw$source_is_url <- sapply(res_raw$source, is_url)
+res_raw <- res_raw %>%
+  mutate(source_name = case_when(
+    !source_is_url ~ source,
+    TRUE ~ NA_character_)) %>%
+  mutate(source = case_when(
+    source_is_url ~ source,
+    TRUE ~ NA_character_)) %>%
   mutate(source = urltools::domain(source))
 
 # source classification
@@ -32,14 +42,14 @@ res_raw <- res_raw %>%
     !is.na(source_type) ~ source_type,
     TRUE ~ NA_character_)) %>%
   mutate(source_type = case_when(
-    grepl("Mix", source_type) | is.na(source_type) ~ "Misc",
-    grepl("Agency", source_type) | is.na(source_type) ~ "Government",
+    # grepl("Mix", source_type) | is.na(source_type) ~ "Misc",
+    grepl("Agency", source_type)~ "Government",
     !is.na(source_type) ~ source_type))
 
 # column naming and arrangement
 res <- res_raw %>%
   dplyr::select(llid, name, legacy_name, state, max_depth_m,
-                mean_depth_m, source, source_type, effort, lat, long, lake_waterarea_ha, has_limno) %>%
+                mean_depth_m, source, source_name, source_type, effort, lat, long, lake_waterarea_ha, has_limno) %>%
   left_join(lg$locus$locus_information, by = c("llid" = "lagoslakeid"))
 
 res <- res %>%
@@ -48,10 +58,10 @@ res <- res %>%
[…]

### 2020-02-19 scripts/99_utils.R

@@ -263,6 +263,12 @@ calc_depth <- function(slope, distance, grain = 1){
   depth
 }
 
+is_url <- function(x){
+  # https://stackoverflow.com/a/3809435/3362993
+  length(grep("[-a-zA-Z0-9@:%._\\+~#=]\\.[a-zA-Z0-9()]{2,10}\\b([-a-zA-Z0-9()@:%_\\+.~#?&\\/\\/=]*)",
+              x)) > 0
+}
+
 # library(maptools)
 # library(spatstat)
 # test4 <- as(as_Spatial(dt), "ppp")

### 2020-02-20 README.Rmd
formatting
rm a lot of the non bathy files and info

@@
 ```{r, echo = FALSE}
+library(details)
+
 knitr::opts_chunk$set(
@@
 
-## LAGOS-US-Depth: Lake depth at continental scales
-
-### Data (see [data/lagosus_depth.csv](data/lagosus_depth.csv))
-
-```{r echo=FALSE}
-# see scripts/99_make_taxonomy.R
-
-dt <- read.csv("data/lagosus_depth_taxonomy.csv", stringsAsFactors = FALSE)
-
-knitr::kable(setNames(dt, c("Variable name", "Description")))
-```
+## Bathymetry manuscript
 
@@
 Tables: [manuscript/tables.pdf](manuscript/tables.pdf)
+
+Data: [bathymetry derived lake depth](data/00_bathy_depth/00_bathy_depth.csv)
+
+```{details, echo=FALSE, details.summary = "Data dictionary"}
+# see scripts/99_make_taxonomy.R
+
+dt <- read.csv("data/00_bathy_depth/00_bathy_depth.csv", stringsAsFactors = FALSE)
+knitr::kable(setNames(data.frame(names(dt)), "field"))
+```
+
+Data: [bathymetry derived lake geometry](data/00_bathy_depth/bathy_geometry.csv)
+
+```{details, echo=FALSE, details.summary = "Data dictionary"}
+# see scripts/99_make_taxonomy.R
+
+dt <- read.csv("data/00_bathy_depth/bathy_geometry.csv", stringsAsFactors = FALSE)
+knitr::kable(setNames(data.frame(names(dt)), "field"))
+```

### 2020-02-20 scripts/00_get_lagosus.R

@@ -1,0 +1,6 @@
+source("scripts/99_utils.R")
+
+lg <- LAGOSUS::lagosus_load("depth")
+lg <- lg$depth$depth
+
+write.csv(lg, "data/lagosus_depth.csv", row.names = FALSE)

### 2020-02-20 scripts/00_get_manual.R

@@ -1,224 +1,0 @@
-source("scripts/99_utils.R")
-# merge depth_log_all with lakewatch and rm lines with missing lakes
-
-# ---- download_from_drive ----
-# unlink("data/00_manual/depth_log_all.csv")
-if(!file.exists("data/00_manual/depth_log_all.csv")){
-  drive_download(file = "depth_log_all",
-                 path = "data/00_manual/depth_log_all.csv", overwrite = TRUE)
-}
-dt_raw <- suppressWarnings(suppressMessages(
-  read_csv("data/00_manual/depth_log_all.csv", col_types = cols())))
-
-# collected in later efforts but not merged in overall log
-# drive_download(file = "PA_USGS_sites_linked_R",
-#                path = "data/00_manual/pa.csv", overwrite = TRUE)
-dt_pa <- suppressMessages(read_csv("data/00_manual/pa.csv")) %>%
-  janitor::clean_names() %>%
-  mutate(file_name = "PA_USGS_sites_linked_R",
-         state = "PA") %>%
-  dplyr::select(Linked_lagoslakeid = linked_lagoslakeid,
-                lat = latitude, lon = longitude,
-                name = lakename_google, file_name,
-                state, max_depth_ft, mean_depth_ft,
-                lakename_google, max_depth_m, mean_depth_m,
-                url, comments)
-
-dt_raw <- dplyr::bind_rows(dt_raw, dt_pa)
-
-# collected in very early efforts and lack a lagoslakeid
-one_off_files <- c("data/00_manual/ar.csv" = "AR_ADEQ_SWQM_sites_lakedepth",
-                   "data/00_manual/ca.csv" = "CA_USGS_WSC_sites_lakedepth",
-                   "data/00_manual/fl.csv" = "FL_LAKEWATCH_station_lakedepth")
-one_off_skips <- c(0, 0, 1)
-# lapply(seq_len(length(misc_one_offs)),
-#        function(x) drive_download(
-#          file = as.character(misc_one_offs[x]),
-#          path = names(misc_one_offs)[x],
-#          overwrite = TRUE))
-
-one_offs <- lapply(seq_len(length(one_off_files)), function(i){
-  # i <- 3
-  state <- toupper(gsub(".csv", "", basename(names(one_off_files)[i])))
[…]

### 2020-02-20 scripts/00_get_manual_extra.R

@@ -1,39 +1,0 @@
-source("scripts/99_utils.R")
-# merge data in Depth_Collection_Extra
-
-locus <- lagosus_load("locus")$locus$locus_characteristics
-
-# files <- c("co_cpw", "ct_healy", "ks_befs", "mt_fwp", "nc_deq", "nd_gfapps",
-#            "nm_env", "ut_deq")
-# lapply(files, function(x) drive_download(file = x,
-#                path = paste0("data/00_manual_extra/", x, ".csv"),
-#                overwrite = TRUE))
-# drive_download(file = "mt_fwp",
-#                path = paste0("data/00_manual_extra/", "mt_fwp", ".csv"),
-#                overwrite = TRUE)
-
-flist <- list.files("data/00_manual_extra/", pattern = ".csv",
-                    full.names = TRUE, include.dirs = TRUE)
-flist <- flist[!grepl("00_manual_extra.csv", flist)]
-
-dt_raw <- lapply(flist, function(x) read.csv(x, stringsAsFactors = FALSE))
-
-dt_raw  <- dplyr::bind_rows(dt_raw) %>%
-  dplyr::filter(!is.na(max_depth_m) | !is.na(mean_depth_m) |
-                !is.na(max_depth_ft) | !is.na(mean_depth_ft))
-
-dt  <- convert_ft_m(dt_raw)
-dt  <- left_join(dt, dplyr::select(locus, lagoslakeid, lake_waterarea_ha,
-                                    lake_connectivity_permanent),
-                 by = c("llid" = "lagoslakeid"))
-dt <- dt[,!duplicated(names(dt))] %>%
-  mutate(legacy_name = NA) %>%
-  mutate(effort = "manual") %>%
-  rename(source = url) %>%
-  select(llid, name, legacy_name, state,
-         max_depth_m, mean_depth_m, source,
-         lake_waterarea_ha, lake_connectivity_permanent, effort,
-         lat, long = lon)
-dt  <- rm_dups(dt)
-
-write.csv(dt, "data/00_manual_extra/00_manual_extra.csv", row.names = FALSE)

### 2020-02-20 scripts/00_get_nla.R

@@ -1,79 +1,0 @@
-source("scripts/99_utils.R")
-
-# get nla files from the cont-limno depth folder and merge with data from nlaR
-# drive_download(file = "US_EPA_NLA2012_siteinfo_v1_linked_R",
-#                path = "data/00_nla/2012_nla.csv")
-# drive_download(file = "US_EPA_NLA2007_siteinfo_v1_linked_R",
-#                path = "data/00_nla/2007_nla.csv")
-
-dt_raw_2012 <- read_csv("data/00_nla/2012_nla.csv")
-dt_raw_2007 <- read_csv("data/00_nla/2007_nla.csv") #nla 2007 file already has depth
-
-nl2012 <- nla_load(2012)$wide_siteinfo %>%
-  dplyr::select(SITE_ID, SITEID_07, EVAL_NAME, LAT_DD83, LON_DD83) %>%
-  distinct(SITE_ID, .keep_all = TRUE) %>%
-  left_join(dplyr::select(nla_load(2012)$wide_profile,
-                          SITE_ID, INDEX_SITE_DEPTH)) %>%
-  tidyr::drop_na(INDEX_SITE_DEPTH) %>%
-  group_by(SITE_ID) %>%
-  slice(which.max(INDEX_SITE_DEPTH)) %>%
-  LAGOSNE::coordinatize("LAT_DD83", "LON_DD83")
-
-dt_2012 <- dt_raw_2012 %>%
-  distinct(SITE_ID, .keep_all = TRUE) %>%
-  left_join(nl2012, by = c("SITE_ID", "LON_DD83", "LAT_DD83")) %>%
-  dplyr::filter(!is.na(INDEX_SITE_DEPTH)) %>%
-  data.frame(stringsAsFactors = FALSE) %>%
-  janitor::clean_names("snake") %>%
-  rename(llid = linked_lagoslakeid,
-         legacy_name = site_id,
-         name = eval_name,
-         lat = lat_dd83,
-         long = lon_dd83) %>%
-  mutate(source = "NLA2012",
-         source_type = "Government",
-         state = single_state(state_nla),
-         max_depth_m = index_site_depth) %>%
-  select(llid, name, legacy_name, state, max_depth_m, mean_depth_m, source, source_type,
-         lat, long, siteid_07) %>%
-  dplyr::filter(!is.na(max_depth_m) | !is.na(mean_depth_m))
-
[…]

### 2020-02-20 scripts/01_merge.R

@@ -1,84 +1,0 @@
-source("scripts/99_utils.R")
-# merge manual, nla, and lagosne data
-
-lg               <- lagosus_load(module = "locus")
-manual_raw       <- read.csv(
-  "data/00_manual/00_manual.csv", stringsAsFactors = FALSE)
-manual_extra_raw <- read.csv(
-  "data/00_manual_extra/00_manual_extra.csv", stringsAsFactors = FALSE)
-nla_raw          <- read.csv(
-  "data/00_nla/00_nla.csv", stringsAsFactors = FALSE)
-lagosne_raw      <- read.csv(
-  "data/00_lagosne/00_lagosne.csv", stringsAsFactors = FALSE)
-bathy_raw        <- read.csv(
-  "data/00_bathy_depth/00_bathy_depth.csv", stringsAsFactors = FALSE)
-
-# merge and limit to 1ft min depth
-res_raw <- dplyr::bind_rows(manual_raw, manual_extra_raw,
-                            nla_raw, lagosne_raw, bathy_raw) %>%
-  dplyr::filter(is.na(max_depth_m) | is.na(mean_depth_m) |
-                  max_depth_m != mean_depth_m) %>%
-  dplyr::filter(is.na(max_depth_m) | is.na(mean_depth_m) |
-                  max_depth_m > mean_depth_m) %>%
-  dplyr::filter(max_depth_m >= 0.3048 | is.na(max_depth_m))
-
-# clean up source columns
-res_raw$source_is_url <- sapply(res_raw$source, is_url)
-res_raw <- res_raw %>%
-  mutate(source_name = case_when(
-    !source_is_url ~ source,
-    TRUE ~ NA_character_)) %>%
-  mutate(source = case_when(
-    source_is_url ~ source,
-    TRUE ~ NA_character_)) %>%
-  mutate(source = urltools::domain(source))
-
-# source classification
-res_raw <- res_raw %>%
-  mutate(source_type = case_when(
-    grepl("\\.gov|\\.us$|_dnr_|dwq|_dep_|_deq_|_dfw_|^epa_", source) ~ "Government",
-    is.na(source_type) & grepl("\\.edu", source) ~ "University",
-    grepl("mn_mpca", source) ~ "Citizen Monitoring",
-    !is.na(source_type) ~ source_type,
-    TRUE ~ NA_character_)) %>%
-  mutate(source_type = case_when(
-    # grepl("Mix", source_type) | is.na(source_type) ~ "Misc",
[…]

### 2020-02-20 scripts/02_qa.R

@@ -1,112 +1,0 @@
-# setwd("../")
-source("scripts/99_utils.R")
-
-lg        <- lagosus_load(module = "locus")
-
-# ---- remove obs that seem wrong in lakes with repeated measures ----
-diff_threshold    <- 8
-percent_threshold <- 0.3
-
-dt_raw        <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE)
-has_limno_ids <- dt_raw %>%
-  dplyr::filter(has_limno == 1) %>%
-  distinct(lagoslakeid) %>%
-  pull(lagoslakeid)
-dt_raw        <- dt_raw %>%
-  dplyr::filter(!is.na(lake_maxdepth_m))
-
-calc_diff_metrics <- function(dt, field = "lake_maxdepth_m"){
-  dt %>%
-  group_by(lagoslakeid) %>%
-    add_tally() %>%
-    mutate(diff = (max(UQ(rlang::sym(field))) - min(UQ(rlang::sym(field)))),
-           percent = diff / max(UQ(rlang::sym(field)))) %>%
-    mutate(diff = case_when(n < 2 ~ 0,
-                            TRUE ~ diff),
-           percent = case_when(n < 2 ~ 0,
-                               TRUE ~ percent)) %>%
-    arrange(desc(diff))
-}
-
-dt     <- dt_raw %>%
-  calc_diff_metrics()
-
-# remove NLA first because we're calling it the most unreliable
-dt     <- dt %>%
-  ungroup(dt) %>%
-  dplyr::filter(!(lagos_effort == "NLA" & percent > percent_threshold)) %>%
-  dplyr::filter(!(lagos_effort == "NLA" & diff > diff_threshold)) %>%
-  dplyr::select(-n, -diff, -percent) %>%
-  calc_diff_metrics()
-
-# remove LAGOSNE second because we're calling it the second most unreliable
-dt     <- dt %>%
-  ungroup(dt) %>%
-  dplyr::filter(!(lagos_effort == "LAGOSNE" & percent > percent_threshold)) %>%
-  dplyr::filter(!(lagos_effort == "LAGOSNE" & diff > diff_threshold)) %>%
-  dplyr::select(-n, -diff, -percent) %>%
-  calc_diff_metrics()
-
-# remove manual third because we assume that NHD hi-res polygons are correct
-dt     <- dt %>%
-  ungroup(dt) %>%
[…]

### 2020-02-20 scripts/99_apportion_lakes.R

@@ -1,53 +1,0 @@
-library(dplyr)
-library(googledrive)
-# drive_find(type = "spreadsheet")
-if(!file.exists("data/backups/depth_log_all.csv")){
-  drive_download("depth_log_all", type = "csv",
-                 path = "data/backups/depth_log_all.csv", overwrite = TRUE)
-}
-
-dt <- read.csv("data/backups/depth_log_all.csv", stringsAsFactors = FALSE) %>%
-  rename(people = Assigned.to.)
-
-# unique(dt$people)
-# dplyr::filter(dt, people == "")
-
-# set apportionment goal
-people <- data.frame(stringsAsFactors = FALSE,
-                    people = c("Lauren", "Jessica", "Joe",
-                               "Ian", "Katelyn", "Jake",
-                               "Arika", "Allie", "Lindsie",
-                               "Sam"),
-                    num_states = c(1L, 1L, 1L,
-                                   2L, 2L, 2L,
-                                   1L, 3L, 10L,
-                                   9L)) %>%
-  mutate(target_frac = num_states / sum(num_states))
-
-res <- dt %>%
-  group_by(people) %>%
-  tally() %>% ungroup() %>%
-  mutate(current_frac = n / sum(n)) %>%
-  left_join(people) %>%
-  mutate(n_target = target_frac * sum(n)) %>%
-  mutate(n_diff = n_target - n) %>%
-  arrange(desc(n_diff)) # %>%
-  # dplyr::filter(n_diff > 0)
-
-# Which people have the most untouched lakes?
-touched <- dt %>%
-  group_by(state) %>%
-  # dplyr::filter(state == "AL") %>%
-  summarize_at(vars(max_depth_ft:comments),
-               function(x) sum(is.na(x)) / length(x)) %>%
-  data.frame() %>% # dt$state == unique(dt_raw$state)
-  dplyr::select(-state) %>%
-  apply(1, function(x) min(x[x > 0])) %>%
-  as.numeric() != 1
-data.frame(touched = !touched, state = unique(dt$state),
-           stringsAsFactors = FALSE) %>%
-  arrange(touched) %>%
-  left_join(distinct(dplyr::select(dt, people, state), people, state, .keep_all = TRUE)) %>%
-  arrange(people) %>%
-  View()
-

### 2020-02-20 scripts/99_log_create.R

@@ -1,143 +1,0 @@
-# collapse log file by state
-#
-# 1. take files with a lagoslakeid field
-# 2. check if fields contain depth from the provider
-# 2. bind rows
-# 3. select lat, lon, name, and llid columns
-
-library(assertr)
-library(dplyr)
-library(tidyr)
-library(janitor)
-library(googledrive)
-# drive_find(type = "spreadsheet")
-if(!file.exists("data/depth_log.csv")){
-  drive_download("LakeDepthLog_for_Reference", type = "csv",
-                 path = "data/depth_log.csv")
-}
-
-# jsta::get_if_not_exists
-get_if_not_exists <- function(x, destfile, read_function = readRDS,
-         overwrite = FALSE, ...){
-
-  if(is.function(x)){
-    if(!file.exists(destfile) | overwrite){
-      res <- x(destfile, ...)
-      return(res)
-    }else{
-      message(paste0("A local evaulation of x already exists on disk"))
-      return(read_function(destfile))
-    }
-  }
-
-  if(!is.function(x)){
-    if(!file.exists(destfile) | overwrite){
-      download.file(x, destfile)
-    }else{
-      message(paste0("A local copy of ", x, " already exists on disk"))
-    }
-    invisible(x)
-  }
-}
-
-depth_log <- read.csv("data/depth_log.csv", stringsAsFactors = FALSE, skip = 3) %>%
-  janitor::clean_names() %>%
-  dplyr::filter(nchar(dataset_assigned_to) <= 0) %>%
-  dplyr::filter(notes != "DEPTHS IN LAGOS-NE") %>%
-  dplyr::filter(state != "Tribal") %>%
-  dplyr::mutate(number_of_linked_lake_sites =
-                  as.numeric(number_of_linked_lake_sites)) %>%
-  dplyr::filter(!is.na(number_of_linked_lake_sites)) %>%
-  rowwise() %>%
-  mutate(file_name = strsplit(dataset_path, "\\\\|/")[[1]][2]) %>%
-  dplyr::select(state, program_name,
-                number_of_linked_lake_sites, file_name)
-
-out_names <- paste0("data/", depth_log$file_name, ".csv")
-get_csv   <- function(destfile, drive_name){
-  tryCatch(drive_download(paste0(drive_name, "_R"), type = "csv",
-                 path = destfile),
-           error = function(e){
-             print(drive_name)
-             write.csv(NA, file = destfile, row.names = FALSE)
-             return(NA)
[…]

### 2020-02-20 scripts/99_log_create_extra.R

@@ -1,43 +1,0 @@
-## create logs for individual state for lakes without wq data
-library(dplyr)
-
-target_state <- "NC"
-file_name    <- "nc_deq.csv"
-
-# pull locus
-ls_raw <- read.csv("data/00_lagosus_locus/lake_information_20190913.csv",
-               stringsAsFactors = FALSE)
-
-# pull manual log
-log <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE)
-
-# remove dups
-ls <- dplyr::filter(ls_raw, !(lagoslakeid %in% log$llid)) %>%
-  dplyr::filter(lake_centroidstate == target_state) %>%
-  # add fields
-  mutate(rowid = seq_len(nrow(.)),
-         max_depth_ft = NA, mean_depth_ft = NA,
-         lakename_google = NA, max_depth_m = NA, mean_depth_m = NA,
-         url = NA) %>%
-  dplyr::select(llid = lagoslakeid, rowid, lat = lake_lat_decdeg,
-                lon = lake_lon_decdeg, name = lake_namegnis,
-                state = lake_centroidstate, max_depth_ft, mean_depth_ft,
-                lakename_google, max_depth_m, mean_depth_m,
-                url)
-
-# distinct !is.na(name) but keep all blanks
-ls_noname    <- dplyr::filter(ls, nchar(name) <= 1)
-ls_name <- dplyr::filter(ls, nchar(name) > 1) %>%
-  distinct(name, .keep_all = TRUE) %>%
-  arrange(name)
-
-ls <- rbind(ls_name, ls_noname)
-
-# export to csv
-write.csv(ls, paste0("data/99_depth_collection_extra/", file_name),
-          row.names = FALSE)
-
-# write to GDrive
-googledrive::drive_upload(media = paste0("data/99_depth_collection_extra/", file_name),
-    path = paste0("Continental-limnology/HUBS/LIMNO-Hub/LAGOS-lake-depth/",
-    "LAGOSUS-LakeDepth/Depth_Collection_Extra/", file_name))

### 2020-02-20 scripts/99_log_update.R

@@ -1,81 +1,0 @@
-source("scripts/99_utils.R")
-
-# ---- read-in-todo-files ----
-dt         <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE)
-flist_todo <- drive_ls("Depth_TODO")
-
-out_names <- paste0("data/99_depth_collection/", flist_todo$name)
-
-depth_files <- lapply(seq_len(length(out_names)), function(x){
-  get_if_not_exists(get_csv, out_names[x], read.csv,
-                    drive_name = flist_todo$name[x])
-}) %>%
-  setNames(flist_todo$name)
-
-# ---- id-unique-llids ----
-
-has_llid    <- unlist(lapply(depth_files,
-                             function(x) "Linked_lagoslakeid" %in% names(x)))
-# View(cbind(names(depth_files), depth_log$file_name, has_llid))
-depth_files <- depth_files[has_llid]
-
-# remove duplicate llids among files
-llids <- lapply(depth_files, function(x){
-  dplyr::select(x, Linked_lagoslakeid)
-})
-llids <- llids  %>%
-  unlist(recursive = FALSE) %>%
-  tibble::enframe() %>%
-  unnest()
-llids <- llids[!duplicated(llids$value),]
-llids <- llids %>%
-  mutate(file_name = gsub(".Linked_lagoslakeid", "", name)) %>%
-  rename(llid = value) %>%
-  dplyr::filter(!is.na(llid))
-llids <- llids[!(llids$llid %in% dt$llid),]
-
-# ---- clean-input-files-for-log ----
-
-depth_log_all <- lapply(depth_files,
-                        function(x) dplyr::select(x,
-          matches("^lat|_lat_|y_coord|latitude$|latdd$|^latitude|ylat"),
-          matches("^lon|_lon_|x_coord|longitude$|longdd$|^longitude|xlon"),
-          matches("waterbodyname|commonname|county_name|waterbody_name|station_name"),
-          "Linked_lagoslakeid")) %>%
-  lapply(function(x) mutate_all(x, as.character)) %>%
-  lapply(function(x) setNames(x, tolower(names(x))))
-
-lat_col  <- lapply(depth_files, function(x)
-  names(dplyr::select(x,
-  matches("^lat|_lat_|y_coord|latitude$|latdd$|^latitude|ylat|lat_ddmmss"))))
-lat_col <- unlist(lapply(lat_col, function(x) x[length(x)]))
-lon_col  <- lapply(depth_files,
-              function(x) names(dplyr::select(x,
[…]

### 2020-02-20 scripts/99_make_taxonomy.R

@@ -1,28 +1,0 @@
-# Make taxonomy table by referencing `limno_taxonomy` file**
-
-# names(read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE))
-
-taxonomy_key <- data.frame(
-  col_name = c("lagoslakeid", "lake_namegnis", "lake_states", "lake_state",
-               "lake_lat_decdeg", "lake_lon_decdeg", "lake_maxdepth_m",
-               "lake_meandepth_m", "lake_waterarea_ha",
-               "sourcename_depth", "sourceurl_depth", "sourcetype_depth",
-               "lagos_effort", "lagos_effort_reliablility"),
-                           description = c(
-                  "unique lake identifier developed for LAGOS-US",
-                  "lake name from the gnis database by way of the LAGOS-US Locus module",
-                  "abbreviation(s) of state(s) intersecting the lake polygon from the LAGOS-US Locus module",
-                  "abbreviation of the state used to search for a lake's depth",
-                  "the latitude of the lake center point (NAD83) from the LAGOS-US Locus module",
-                  "the longitude of the lake center point (NAD83) from the LAGOS-US Locus module",
-                  "lake maximum depth in meters",
-                  "lake mean depth in meters",
-                  "surface area of lake waterbody polygon from NHD (excludes islands) by way of the LAGOS-US Locus module",
-                  "name of the source of depth data",
-                  "url link of the source of depth data",
-                  "type of the source of depth data; one of Citizen Monitoring, Government, University, Commercial",
-                  "name of depth searching effort used for internal tracking; one of LAGOSNE, LAGOSUS, bathymetry, NLA",
-                  "reliability of depth searching effort used for quality assurance and de-duplication; values in ascending order with 1 being the most reliable"),
-  stringsAsFactors = FALSE)
-
-write.csv(taxonomy_key, "data/lagosus_depth_taxonomy.csv", row.names = FALSE)

### 2020-02-20 scripts/nla_vs_lagosus.R

@@ -1,20 +1,0 @@
-library(nlaR)
-library(dplyr)
-library(tidyr)
-library(LAGOSNE)
-library(ggplot2)
-
-nl2012 <- nla_load(2012)
-dt     <- dplyr::select(nl2012$wide_siteinfo,
-                    SITE_ID, LAT_DD83, LON_DD83) %>%
-  distinct(SITE_ID, .keep_all = TRUE) %>%
-  left_join(dplyr::select(nl2012$wide_profile,
-                          SITE_ID, INDEX_SITE_DEPTH)) %>%
-  tidyr::drop_na(INDEX_SITE_DEPTH) %>%
-  group_by(SITE_ID) %>%
-  slice(which.max(INDEX_SITE_DEPTH)) %>%
-  LAGOSNE::coordinatize("LAT_DD83", "LON_DD83")
-
-ggplot() +
-  geom_sf(data = dt, aes(color = INDEX_SITE_DEPTH))
-

### 2020-02-20 README.Rmd
formatting

@@
 ```
-
 Data: [bathymetry derived lake geometry](data/00_bathy_depth/bathy_geometry.csv)

### 2020-02-20 README.Rmd
display markdown table data dictionaries

--- 2020-02-18 README.Rmd
@@
 Tables: [manuscript/tables.pdf](manuscript/tables.pdf)
+
+Data: [bathymetry derived lake depth](data/00_bathy_depth/00_bathy_depth.csv)
+
+<details>
+<summary>Data dictionary</summary>
+```{r, echo=FALSE}
+# see scripts/99_make_taxonomy.R
+
+dt <- read.csv("data/00_bathy_depth/00_bathy_depth.csv", stringsAsFactors = FALSE)
+knitr::kable(setNames(data.frame(names(dt)), "field")) 
+```
+</details> &nbsp;
+
+Data: [bathymetry derived lake geometry](data/00_bathy_depth/bathy_geometry.csv)
+
+<details>
+<summary>Data dictionary</summary>
+```{r, echo=FALSE}
+# see scripts/99_make_taxonomy.R
+
+dt <- read.csv("data/00_bathy_depth/bathy_geometry.csv", stringsAsFactors = FALSE)
+knitr::kable(setNames(data.frame(names(dt)), "field"))
+```
+</details>

### 2020-02-20 figures/01_geometry.Rmd
remove junk from geometry panel, add legends

@@
   geom_point(aes(x = dist_deepest, y = dist_viscenter, color = shape_class)) +
-  geom_abline(aes(intercept = 0, slope = 1)) +
+  # geom_abline(aes(intercept = 0, slope = 1)) +
   labs(color = "") +
@@
 
-gg_scatter_reservoir <- ggplot(data = st_drop_geometry(dt)) + 
+gg_scatter_reservoir <- ggplot(data = dplyr::filter(
+  st_drop_geometry(dt), !is.na(reservoir_class))) + 
   geom_point(aes(x = dist_deepest, y = dist_viscenter, color = reservoir_class)) +
-  geom_abline(aes(intercept = 0, slope = 1)) +
+  # geom_abline(aes(intercept = 0, slope = 1)) +
   xlab("Deepest point distance (m)") +
@@
     geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
+  geom_smooth(data = dplyr::filter(test, shape_class != "neither"), 
+              method = "lm", se = FALSE) +
   xlab("Buffer slope (mean)") +
@@
 
-gg_slope_reservoir <- ggplot(data = test, 
+gg_slope_reservoir <- ggplot(data = dplyr::filter(test, !is.na(reservoir_class)), 
                              aes(slope_mean, inlake_slope.y, 
@@
 theme_opts <- theme(legend.position = "none")
+
+legend_shape     <- cowplot::get_legend(gg_scatter_shape)
+legend_reservoir <- cowplot::get_legend(gg_scatter_reservoir + labs(color = ""))
+legend_panel     <- cowplot::plot_grid(legend_shape, legend_reservoir, 
+                                       ncol = 1, 
+                                       align = "hv")
+
+cowplot::plot_grid(
   cowplot::plot_grid(gg_scatter_shape + 
@@
                        theme_opts, 
-                   rel_widths = c(1, 0.93, 1, 1))
+                     rel_widths = c(1, 0.93, 1, 1)), 
+  legend_panel, rel_widths = c(1, 0.2)
+)
 ```

### 2020-02-21 figures/00_maps.Rmd
add reservoir map panel

--- 2020-02-13 figures/00_maps.Rmd
@@
     scale_color_manual(values = viridisLite::viridis(4)) +
-    labs(color = "Max depth (m)") +
+    labs(color = "Max depth (m)  ") +
     coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
@@
 
+```{r gg_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+dt           <- st_read("../data/gis.gpkg", "dt", quiet = TRUE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  distinct(lagoslakeid, .keep_all = TRUE)
+dt_withres <- read.csv("../data/lagosus_depth_predictors.csv", 
+               stringsAsFactors = FALSE) %>%
+  dplyr::distinct(lagoslakeid, reservoir_class) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid))
+dt           <- inner_join(dt, dt_withres) %>%
+  arrange(desc(reservoir_class))
+
+gg_reservoir <- 
+  ggplot() + 
+    geom_sf(data = st_as_sfc(st_bbox(dt_bathy)), alpha = 1, size = 0.6, 
+          color = "black", fill = "#1D9DCF") +
+    geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))), fill = "gray", color = "black", size = 0.3) +
+    geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))), fill = "white", color = "black", size = 0.4) + 
+    geom_sf(data = dplyr::filter(dt, !is.na(reservoir_class)), 
+            aes(color = reservoir_class), size = 0.03, alpha = 0.3) +
+    labs(color = "Reservoir class") +
+    coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
+    theme_void()
+
+gg_reservoir
+```
+
 ```{r 00_map_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6}
-cowplot::plot_grid(gg_bathy, 
-                   gg_shapeclass, 
+theme_opts <- guides(color = guide_legend(
+  override.aes = list(size = 2)))
+cowplot::plot_grid(gg_bathy + theme_opts, 
+                   gg_shapeclass + theme_opts,
+                   gg_reservoir + theme_opts,
                    ncol = 1)

### 2020-02-21 manuscript/figures.Rmd
add reservoir map panel

@@
 
-```{r maps_dot_bathy, fig.cap = "Map of lake maximum depth measurements calculated from lake bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE}
 knitr::include_graphics("../figures/00_map_bathy-1.pdf")

### 2020-02-21 scripts/00_get_gis.R
limit focal lakes to bathy, closes #7

@@ -17,7 +17,12 @@ states_focal <- states_all %>%
 study_bbox   <- st_as_sfc(st_bbox(states_focal))
 states_all   <- states_all %>%
   st_crop(study_bbox)
-states_focal <- dplyr::filter(states_all, iso_a2 == "US")
+states_focal <- dplyr::filter(states_all, iso_a2 == "US") %>%
+  dplyr::filter(postal %in%
+                  pull(distinct(
+                    dplyr::filter(st_drop_geometry(dt),
+                                  lagos_effort == "bathymetry"),
+                    lake_state), lake_state))
 
 # add zoneids
 

### 2020-02-21 figures/00_maps.Rmd
adjust figure dimensions

@@
     geom_point_rast(data = dt_notsf, aes(x = X, y = Y, color = maxdepth_class), 
-                    size = 0.3, alpha = 0) +
+                    alpha = 0) +
     scale_color_manual(values = viridisLite::viridis(5)) +
     geom_point_rast(data = dplyr::filter(dt_notsf, lake_maxdepth_m < 10), 
-            aes(x = X, y = Y, color = maxdepth_class), size = 0.3, alpha = 0.8) +
+            aes(x = X, y = Y, color = maxdepth_class), size = 0.2, alpha = 0.8) +
     geom_point_rast(data = dplyr::filter(dt_notsf, lake_maxdepth_m > 10), 
-            aes(x = X, y = Y, color = maxdepth_class), size = 0.3, alpha = 0.8) +
+            aes(x = X, y = Y, color = maxdepth_class), size = 0.2, alpha = 0.8) +
     labs(color = "Max depth (m)") +
@@
                stringsAsFactors = FALSE) %>%
+  dplyr::filter(!is.na(shape_class)) %>%
   dplyr::distinct(lagoslakeid, reservoir_class) %>%

### 2020-02-21 figures/01_geometry.Rmd
adjust figure dimensions

@@
                         levels = c("neither", "concave", "convex")))) %>%
-  ggplot() + 
-  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = shape_class)) +
+  ggplot(aes(x = dist_deepest, y = dist_viscenter, color = shape_class)) + 
+  geom_point() +
   # geom_abline(aes(intercept = 0, slope = 1)) +
@@
 
-```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5}
+```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5, fig.height=4.2}
 theme_opts <- theme(legend.position = "none")

### 2020-02-21 manuscript/figures.Rmd
adjust figure dimensions

@@
 ![Diagram showing the assumption of most lake depth models that the nearshore slope is extends into the lake and that all lakes have a common relationship between nearshore and in-lake slope.
-](../figures/slope_diagram.pdf){width=85%}
+](../figures/slope_diagram.pdf){width=80%}
 
-![Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right)](../figures/lake_shape.pdf){width=85%}
+![Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right)](../figures/lake_shape.pdf){width=80%}
 

### 2020-02-24 figures/02_depth_model.Rmd
create residual density plots

@@
 
-```{r 02_depth_model, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r 02_depth_model_grid_resid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
 
-dt <- readRDS("../data/01_depth_model/depth_model.rds")
+lgus <- read.csv("../data/lagosus_depth_predictors.csv", 
+                 stringsAsFactors = FALSE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  dplyr::select(-lake_maxdepth_m) %>%
+  distinct(lagoslakeid, .keep_all = TRUE)
+dt   <- readRDS("../data/01_depth_model/depth_grid.rds") %>%
+  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  left_join(lgus, by = "lagoslakeid")
+
+ggplot(data = dt) + 
+  geom_density(aes(x = resid, fill = shape_class)) +
+  geom_vline(aes(xintercept = 0)) +
+  facet_wrap(~ model)
+
+ggplot(data = dt) + 
+  geom_density(aes(x = resid, fill = reservoir_class)) +
+  geom_vline(aes(xintercept = 0)) +
+  facet_wrap(~ model)
 ```
@@
 ```{r 02_depth_model_fitted, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+dt <- readRDS("../data/01_depth_model/depth_model.rds")
 

### 2020-02-26 scripts/00_get_geometry.R
handle reprojected bathy_pnts

@@ -11,15 +11,16 @@ lg <- lagosus_load("locus")
 #   the distance between these points
 #   the true in-lake "slope"
 get_geometry <- function(r, llid, deep_positive = TRUE, ft = 1){
-  # llid <- 5636
-  # r <- raster(paste0("data/nh_bathy/", llid, ".tif"))
+  # llid <- 2119
+  # r <- raster(paste0("data/mi_bathy/", llid, ".tif"))
   # deep_positive = TRUE
   # ft <- 3.281
   # ft = 1
 
   dt_poly <- LAGOSUSgis::query_gis("LAGOS_US_All_Lakes_1ha",
                                    "lagoslakeid", llid)
-  proj_str <- st_crs(r)$proj4string
+  proj_str_init <- st_crs(r)$proj4string
+  proj_str <- proj_str_init
   if(nrow(st_coordinates(st_transform(dt_poly, proj_str))) >= 1){
     dt_poly <- st_transform(dt_poly, proj_str)
   }else{
@@ -112,6 +113,11 @@ get_geometry <- function(r, llid, deep_positive = TRUE, ft = 1){
   inlake_slope_mean   <- mean(r_slope@data@values, na.rm = TRUE) * res(r)[1]
   inlake_slope_median <- median(r_slope@data@values, na.rm = TRUE) * res(r)[1]
 
+  # make sure pnt deepest and pnt_viscenter match
+  # the projection of the on disk raster
+  # attempt to project pnts to `proj_str_init`
+  # if empty project raster to `proj_str`
+
   list(pnt_deepest = pnt_deepest, pnt_viscenter = pnt_viscenter,
        dist_deepest = dist_deepest, dist_viscenter = dist_viscenter,
        dist_between = dist_between, inlake_slope = inlake_slope,
@@ -128,10 +134,10 @@ rm_bad_rasters <- function(rsubs){
 }
 
 loop_state <- function(fpath, outname, deep_positive, ft = 1){
-  # fpath <- "data/me_bathy/"
-  # outname <- "data/00_bathy_depth/00_bathy_depth_me.rds"
+  # fpath <- "data/mi_bathy/"
+  # outname <- "data/00_bathy_depth/00_bathy_depth_mi.rds"
   # deep_positive = TRUE
-  # ft = 1
+  # ft = 3.281
   flist <- list.files(fpath, pattern = "\\d.tif",
                          full.names = TRUE, include.dirs = TRUE)
   # print(flist)

### 2020-02-26 scripts/00_get_hypso_mi.R

@@ -54,7 +54,7 @@ rsubs <- lapply(seq_along(unique(mi$lagoslakeid)),
   pb$tick(tokens = list(llid = unique(mi$lagoslakeid)[i]))
 
   # i <- 1
-  # i <- which(unique(mi$lagoslakeid) == 107971)
+  # i <- which(unique(mi$lagoslakeid) == 2645)
   fname <- paste0("data/mi_bathy/", unique(mi$lagoslakeid)[i], ".tif")
   if(!file.exists(fname)){
     dt <- dplyr::filter(mi, lagoslakeid == unique(mi$lagoslakeid)[i])
@@ -64,7 +64,14 @@ rsubs <- lapply(seq_along(unique(mi$lagoslakeid)),
     #     data = dplyr::filter(lg_poly, lagoslakeid == unique(mi$lagoslakeid)[i]))
 
     res <- poly_to_filled_raster(dt, "DEPTH", 27, proj = 32616)
+
     if(cellStats(res$r, max) != 0){
+      dt_poly <- dplyr::filter(lg_poly, lagoslakeid == unique(mi$lagoslakeid)[i])
+      if(nrow(st_coordinates(st_transform(dt_poly,
+        st_crs(res$r)$proj4string))) < 1){
+        proj_str <- st_crs(dt_poly)
+        res$r <- projectRaster(res$r, crs = proj_str$proj4string)
+      }
       writeRaster(res$r, fname)
     }
   }else{

### 2020-02-26 scripts/00_get_nearshore.R

@@ -5,7 +5,7 @@ source("scripts/99_utils.R")
 
 set.seed(55)
 max_buffer_dist <- 100 # to match Hollister (2011)
-n_lakes         <- 170
+n_lakes         <- 510
 
 dt         <- read.csv("data/lagosus_depth.csv",
                     stringsAsFactors = FALSE)
@@ -27,9 +27,11 @@ ll_ids <- ll_ids[!duplicated(ll_ids)]
 
 get_slope <- function(ll_id){
   # ll_id <- ll_ids[1]
+  # ll_id <- 2645
   ll_poly <- query_gis("LAGOS_NE_All_Lakes_4ha", "lagoslakeid", ll_id)
+  ll_poly <- lwgeom::st_make_valid(ll_poly)
   ll_iws  <- query_gis("IWS", "lagoslakeid", ll_id)
-  # TODO: make this a one-sided buffer [sf doesn't offer this :(]
+  # hack together a one-sided buffer [sf doesn't offer this :(]
   ll_buff <- st_buffer(ll_poly, max_buffer_dist)
   ll_buff <- st_difference(ll_buff, ll_poly)
 
@@ -62,6 +64,10 @@ get_slope <- function(ll_id){
     paste0("data/", tolower(deepest_pnt$state), "_bathy/", ll_id, ".tif")
     )
   st_crs(deepest_pnt) <- st_crs(r_crs)
+  if(is.na(st_distance(st_transform(ll_poly, st_crs(r_crs)), deepest_pnt))){
+    st_crs(deepest_pnt) <- st_crs(ll_poly)
+    r_crs <- projectRaster(r_crs, crs = st_crs(ll_poly)$proj4string)
+  }
 
   ll_poly_hull <- st_zm(st_cast(
     concaveman::concaveman(st_cast(ll_poly, "MULTILINESTRING")),
@@ -70,12 +76,13 @@ get_slope <- function(ll_id){
     concaveman::concaveman(st_cast(ll_buff, "MULTILINESTRING")),
     "MULTILINESTRING"))
   shore_pnt <- sf::st_nearest_points(deepest_pnt,
-                                     st_transform(ll_poly_hull, st_crs(r_crs)))
+                                  st_transform(ll_poly_hull, st_crs(deepest_pnt)))
   buffer_line  <- sf::st_nearest_points(shore_pnt,
-                                        st_transform(ll_buff_hull, st_crs(r_crs)))
[…]

### 2020-02-26 manuscript/figures.Rmd
outline figures todo

--- 2020-02-18 manuscript/figures.Rmd
@@
 - \usepackage{colortbl}
+- \usepackage{float}
 ---
@@
 knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
+knitr::opts_chunk$set(fig.pos = 'H')
 ```
@@
 ![Diagram showing the assumption of most lake depth models that the nearshore slope is extends into the lake and that all lakes have a common relationship between nearshore and in-lake slope.
-](../figures/slope_diagram.pdf){width=85%}
+](../figures/slope_diagram.pdf){width=75%}
 
-![Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right)](../figures/lake_shape.pdf){width=85%}
+```{r lake_shape, echo=FALSE, out.extra='', fig.align='center', fig.cap="Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right)"}
+knitr::include_graphics("../figures/lake_shape.pdf")
+```
 
-```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported max depth and max depth calculated from bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+knitr::include_graphics("../figures/00_map_bathy-1.pdf")
+```
+
+```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported max depth and max depth calculated from bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 # see 00_qa.Rmd
@@
 
-```{r maps_dot_bathy, fig.cap = "Map of lake maximum depth measurements calculated from lake bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/00_map_bathy-1.pdf")
-```
+### predictor summary table
 
-```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) buffer slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE}
+### variable importance for true slopes and distances (5)
+
+### model metrics grid (table)
+
+### 2 x 4 grid of residual box plots for shape and reservoir by true and proxy dist/slope (6)
+
+### 4 x 2 grid of dotplots of depth, slope, area, dist by shape and reservoir class (7)
+
+# Appendix
+
+```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) buffer slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 knitr::include_graphics("../figures/01_geometry_grid-1.pdf")

### 2020-02-26 tables/01_predictors.Rmd
create bathy specific predictor table

@@
 
-# pred_key <- read.csv("../data/predictor_key.csv", 
-#                      stringsAsFactors = FALSE)
-# 
-# # scripts/02_aggregate_predictors.R
-# dt_units <- readRDS("../data/dt_units.rds")
-
-# cdl_vars <- readRDS("../data/dt.rds") %>%
-#   dplyr::select(ag:wheat) %>%
-#   summarize_all(median, na.rm = TRUE) %>%
-#   tidyr::gather() %>%
-#   dplyr::filter(value >= 1.4) %>%
-#   arrange(desc(value)) %>%
-#   pull(key)
-
 # setwd("tables")
+nearshore <- read.csv("../data/00_geometry/nearshore.csv", 
+                      stringsAsFactors = FALSE) %>%
+  dplyr::select(llid, buffer_slope = slope_mean)
 dt <- read.csv("../data/lagosus_depth_predictors.csv", 
@@
   distinct(lagoslakeid, .keep_all = TRUE) %>%
+  left_join(nearshore, by = c("lagoslakeid" = "llid")) %>%
   dplyr::select(-contains("lagoslakeid"), 
@@
 
-knitr::kable(dt, digits = 0, format = "markdown") 
+# knitr::kable(dt, digits = 0, format = "markdown") 
 #%>%
@@
 ```
+
+```{r predictors_table}
+# setwd("figures")
+dt <- dt %>%
+  dplyr::filter(variable %in% 
+              c("lake_meandepth_m", 
+              names(readRDS("../data/01_depth_model/depth_training.rds")), 
+              "buffer_slope")) %>%
+  mutate(n = dt$n[which(dt$variable == "inlake_slope")])
+
+knitr::kable(dt, digits = 0, format = "markdown") 
+```

### 2020-02-26 figures/01_heatmap.Rmd
make a normalized heatmap

@@
 library(pheatmap)
+library(recipes)
 
@@
 
-pretty_names <- gsub("mbgconhull_", "", names(dt_raw_ne))
+dt_raw_ne_normal_rec <- recipe(. ~ ., data = dt_raw_ne) %>%
+  step_normalize(all_numeric()) %>%
+  prep()
+dt_raw_ne_normal <- juice(dt_raw_ne_normal_rec)
+
+create_heatmap <- function(x){
+  print(range(x$lake_maxdepth_m))
+  pretty_names <- gsub("mbgconhull_", "", names(x))
   pretty_names <- gsub("lake_", "", pretty_names)
@@
 
-test <- dt_raw_ne %>%
+  test <- x %>%
     setNames(pretty_names) %>%
@@
             RColorBrewer::brewer.pal(n = 7, name = "Reds"))(100))
+}
+
+create_heatmap(dt_raw_ne)
+create_heatmap(dt_raw_ne_normal)
 ```

### 2020-02-26 figures/02_depth_model.Rmd
add model metrics table

@@
 source("../scripts/99_utils.R")
+```
+
+```{r depth_model_metrics}
+dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds")
+jsta::pdf_table(dt_metrics, "../tables/depth_model_metrics-1.pdf")
 ```
@@
   geom_vline(aes(xintercept = 0)) +
-  facet_wrap(~ model)
+  facet_wrap(~ model) + 
+    theme(legend.position = "none",
+        plot.margin = unit(c(1.5,1.5,0.2,0.2), "cm"))
+grid::grid.text(unit(0.93, "npc"), 0.32, 
+                label = 'proxy slope', rot = 270) # bottom right
+grid::grid.text(unit(0.93, "npc"), 0.65, 
+                label = 'true slope', rot = 270) # top right
+grid::grid.text(unit(0.7, "npc"), 0.89, 
+                label = 'proxy distance') # right
+grid::grid.text(unit(0.3, "npc"), 0.89, 
+                label = 'true distance') # left
 
@@
   geom_vline(aes(xintercept = 0)) +
-  facet_wrap(~ model)
+  facet_wrap(~ model) +
+  theme(legend.position = "none",
+        plot.margin = unit(c(1.5,1.5,0.2,0.2), "cm"))
+grid::grid.text(unit(0.93, "npc"), 0.32, 
+                label = 'proxy slope', rot = 270) # bottom right
+grid::grid.text(unit(0.93, "npc"), 0.65, 
+                label = 'true slope', rot = 270) # top right
+grid::grid.text(unit(0.7, "npc"), 0.89, 
+                label = 'proxy distance') # right
+grid::grid.text(unit(0.3, "npc"), 0.89, 
+                label = 'true distance') # left
 ```

### 2020-02-26 manuscript/figures.Rmd
create bathy specific predictor table
add model metrics table

@@
 
-### predictor summary table
+![Summary of lake depth predictor variables.](../tables/01_predictors.pdf){width=65%}
 
-### variable importance for true slopes and distances (5)
+```{r metric_correlations, fig.cap = "metric correlations", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+knitr::include_graphics("../figures/metric_correlations-1.pdf")
+```
 
-### model metrics grid (table)
+```{r depth_model_metrics, fig.cap = "depth model metrics", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+knitr::include_graphics("../tables/depth_model_metrics-1.pdf")
+```
 
@@
 ```
-
-![Summary of lake depth predictor variables.](../tables/01_predictors.pdf){width=65%}
 

### 2020-02-26 figures/01_heatmap.Rmd
add true metrics correlation figure

@@
 ```
+
+```{r metric_correlations_setup, echo=FALSE, warning=FALSE}
+
+dt <- dplyr::select(dt_raw_ne, names(dt_raw_ne)[names(dt_raw_ne) %in%
+                    c("lake_meandepth_m", 
+              names(readRDS("../data/01_depth_model/depth_training.rds")))])
+names(dt) <- gsub("^lake_", "", names(dt))
+
+gg_slope <- jsta::jheatmap(dt, focal_columns = "inlake_slope")$gtable
+gg_dist <- jsta::jheatmap(dt, focal_columns = "dist_deepest")$gtable
+```
+
+```{r metric_correlations, echo=FALSE, warning=FALSE}
+cowplot::plot_grid(
+  gg_slope, 
+  gg_dist, 
+  rel_widths = c(1, 0.95))
+```

### 2020-02-26 figures/02_depth_model.Rmd
add residual density plots facetting by class

--- 2020-02-24 figures/02_depth_model.Rmd
@@
 
-ggplot(data = dt) + 
-  geom_density(aes(x = resid, fill = shape_class)) +
-  geom_vline(aes(xintercept = 0)) +
-  facet_wrap(~ model)
+gg_shape <- ggplot(data = dt) + 
+  geom_density(aes(x = resid, fill = shape_class), alpha = 0.75, adjust = 1.3) +
+  geom_vline(aes(xintercept = 0), linetype = "dashed") +
+  facet_wrap(~ model) + 
+    theme(legend.position = "bottom",
+        strip.background = element_blank(), 
+        strip.text = element_blank(),
+        plot.margin = unit(c(1.2,0.2,0.2,0.2), "cm"))  +
+  guides(fill = guide_legend(nrow = 2))
 
-ggplot(data = dt) + 
-  geom_density(aes(x = resid, fill = reservoir_class)) +
-  geom_vline(aes(xintercept = 0)) +
-  facet_wrap(~ model)
+gg_shape <- ggdraw(gg_shape)  +
+  # draw_label("proxy", x = 0.83, y = 0.32, angle = 270, size = 10) +
+  # draw_label("true", x = 0.83, y = 0.65, angle = 270, size = 10) +
+  # draw_label("slope", x = 0.93, y = 0.5, size = 10, angle = 270) +
+  draw_label("proxy", x = 0.7, y = 0.89, size = 10) +
+  draw_label("true", x = 0.3, y = 0.89, size = 10) +
+  draw_label("distance", x = 0.5, y = 0.94, size = 10)# +
+  #draw_label("Shape class", x = 0.5, y = 0.98, size = 10, fontface = "bold")
+
+gg_reservoir <- ggplot(data = dt) + 
+  geom_density(aes(x = resid, fill = reservoir_class), alpha = 0.75, adjust = 1.3) +
+  geom_vline(aes(xintercept = 0), linetype = "dashed") +
+  facet_wrap(~ model) +
+  theme(legend.position = "bottom",
+        strip.background = element_blank(), 
+        strip.text = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = unit(c(1.2,1.2,0.2,0.2), "cm")) +
+  guides(fill = guide_legend(nrow = 2))
+
+gg_reservoir <- ggdraw(gg_reservoir) +
+  draw_label("proxy", x = 0.83, y = 0.44, angle = 270, size = 10) +
+  draw_label("true", x = 0.83, y = 0.77, angle = 270, size = 10) +
+  draw_label("slope", x = 0.93, y = 0.62, size = 10, angle = 270) +
+  draw_label("proxy", x = 0.7, y = 0.89, size = 10) +
+  draw_label("true", x = 0.3, y = 0.89, size = 10) +
+  draw_label("distance", x = 0.5, y = 0.94, size = 10)# +
+  # draw_label("Reservoir class", x = 0.5, y = 0.98, size = 10, fontface = "bold")
+
+cowplot::plot_grid(gg_shape, gg_reservoir, 
+                   nrow = 1, rel_widths = c(1, 0.95))
 ```

### 2020-02-26 manuscript/figures.Rmd
add residual density plots facetting by class

@@
 
-### 2 x 4 grid of residual box plots for shape and reservoir by true and proxy dist/slope (6)
+```{r depth_model_grid_resid, fig.cap = "Depth model residuals by shape and reservoir class.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+knitr::include_graphics("../figures/02_depth_model_grid_resid-1.pdf")
+```
 

### 2020-02-28 manuscript/figures.Rmd
cleanup, fight with figure numbering

@@
 - \usepackage{float}
+- \usepackage{caption}
 ---
@@
 
-![Summary of lake depth predictor variables.](../tables/01_predictors.pdf){width=65%}
+\captionsetup[figure]{name=Table}
+\captionsetup[figure]{labelformat=empty}
+```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth predictor variables", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../tables/01_predictors.pdf")
+```
+\captionsetup[figure]{name=Figure}
+\captionsetup[figure]{labelformat=default}
+\setcounter{figure}{4}
 
@@
 
-```{r depth_model_metrics, fig.cap = "depth model metrics", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+\captionsetup[figure]{name=Table}
+\captionsetup[figure]{labelformat=empty}
+```{r depth_model_metrics, fig.cap = "Table 2: depth model metrics", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 knitr::include_graphics("../tables/depth_model_metrics-1.pdf")
 ```
+\captionsetup[figure]{name=Figure}
+\captionsetup[figure]{labelformat=default}
+\setcounter{figure}{5}
 
@@
 # Appendix
+\setcounter{figure}{0}
 
@@
 ```{r heatmap, fig.cap = "Correlation matrix heatmap.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_heatmap-1.pdf")
+knitr::include_graphics("../figures/slope-compare-1.pdf")
 ```
@@
 
-```{r contrasts_tally, fig.cap = "Number of lake depth measurements by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r contrasts_tally, fig.cap = "Number of lake depth measurements by categories.", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/01_contrasts_tally-1.pdf")
@@
 
-```{r 02_depth_model_fitted, fig.cap = "A model that does not include shape class over predicts depth in concave lakes but does not systematically under predict depth in convex lakes.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r 02_depth_model_fitted, fig.cap = "A model that does not include shape class over predicts depth in concave lakes but does not systematically under predict depth in convex lakes.", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/02_depth_model_fitted-1.pdf")
@@
 
-```{r 02_depth_oliver-2, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r 02_depth_oliver-2, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/02_depth_oliver-2.pdf")
@@
 
-```{r 02_depth_oliver-1, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r 02_depth_oliver-1, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/02_depth_oliver-1.pdf")
@@
 
-```{r 02_depth_oliver-4, fig.cap = "Oliver 2016 observed versus predicted.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r 02_depth_oliver-4, fig.cap = "Oliver 2016 observed versus predicted.", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/02_depth_oliver-4.pdf")
@@
 
-![Messager et al. 2016 equation table](../figures/messager_eq_table.png){width=100%}
-
-\pagebreak
+<!-- ![Messager et al. 2016 equation table](../figures/messager_eq_table.png){width=100%} -->
 

### 2020-02-28 figures/01_contrasts.Rmd
add more detailed contrasts figure

@@
   ylab("Depth (m)") + xlab("")
+```
+
+```{r gg_area_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+gg_area_shape_class <- dt %>%
+  dplyr::filter(shape_class != "neither") %>%
+  group_by(shape_class) %>%
+  summarize(lake_waterarea_ha = list(tibble::enframe(
+    quantile(lake_waterarea_ha, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = shape_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Area (ha)") + xlab("")
+```
+
+```{r gg_slope_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+gg_slope_shape_class <- dt %>%
+  dplyr::filter(shape_class != "neither") %>%
+  group_by(shape_class) %>%
+  summarize(inlake_slope = list(tibble::enframe(
+    quantile(inlake_slope, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = shape_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Inlake slope (m/m)") + xlab("")
+```
+
+```{r gg_dist_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+gg_dist_shape_class <- dt %>%
+  dplyr::filter(shape_class != "neither") %>%
+  group_by(shape_class) %>%
+  summarize(dist_deepest = list(tibble::enframe(
+    quantile(dist_deepest, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = shape_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Deepest point distance (m)") + xlab("")
 ```
@@
 
+```{r gg_area_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+
+(gg_area_reservoir <- dt %>%
+  dplyr::filter(!is.na(reservoir_class)) %>%
+  group_by(reservoir_class) %>%
+  summarize(lake_waterarea_ha = list(tibble::enframe(
+    quantile(lake_waterarea_ha, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = reservoir_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Area (ha)") + xlab(""))
+```
+
+```{r gg_slope_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+
+(gg_slope_reservoir <- dt %>%
+  dplyr::filter(!is.na(reservoir_class)) %>%
+  group_by(reservoir_class) %>%
+  summarize(inlake_slope = list(tibble::enframe(
+    quantile(inlake_slope, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = reservoir_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Inlake slope (m/m)") + xlab(""))
+```
+
+```{r gg_dist_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+
+(gg_dist_reservoir <- dt %>%
+  dplyr::filter(!is.na(reservoir_class)) %>%
+  group_by(reservoir_class) %>%
+  summarize(dist_deepest = list(tibble::enframe(
+    quantile(dist_deepest, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+  unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = reservoir_class, y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab("Deepest point distance (m)") + xlab(""))
+```
+
 ```{r gg_depth_xtabs, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
@@
 
-```{r 01_contrasts_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 6}
-theme_opts <- theme(axis.text.x = element_text(angle = 90))
+```{r 01_contrasts, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 4}
+theme_opts <- theme(axis.text.x = element_text(angle = 90), 
+                    axis.title = element_blank(),
+                    plot.title = element_text(size = 8))
 
-cowplot::plot_grid(  
-cowplot::plot_grid(gg_depth_shape_class + 
-                     theme_opts, 
-                   gg_depth_conny_class + 
-                     theme_opts +
-                     theme(axis.title.y = element_blank()), 
-                   gg_depth_glaciated_class + 
-                     theme_opts +
-                     theme(axis.title.y = element_blank()), 
-                   gg_depth_reservoir + 
-                     theme_opts +
-                     theme(axis.title.y = element_blank()), 
-                   nrow = 1), 
+gg_shape <- cowplot::plot_grid(
+  gg_depth_shape_class + ggtitle("Depth (m)") + theme_opts,
+  gg_area_shape_class + ggtitle("Area (ha)") + theme_opts, 
+  gg_area_shape_class + ggtitle("Inlake slope (m/m)") + theme_opts, 
+  gg_area_shape_class + ggtitle("Deepest point distance (m)") + theme_opts, 
+  nrow = 1)
+
+gg_reservoir <- cowplot::plot_grid(gg_depth_reservoir + theme_opts, 
+                                   gg_area_reservoir + theme_opts, 
+                                   gg_slope_reservoir + theme_opts, 
+                                   gg_dist_reservoir + theme_opts, 
+                                   nrow = 1)
+
+cowplot::plot_grid(gg_shape, gg_reservoir, 
+                   nrow = 2, rel_heights = c(1, 0.85))
+
 cowplot::plot_grid(gg_depth_xtabs + theme_opts, 
                    ggplot() + geom_blank() + 
-                     theme(panel.background = element_blank()), nrow = 1), 
-nrow = 2)
-
+                     theme(panel.background = element_blank()), nrow = 1)
 ```

### 2020-02-28 manuscript/figures.Rmd
add more detailed contrasts figure

@@
 
-```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../figures/00_map_bathy-1.pdf")
@@
 
-```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported max depth and max depth calculated from bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth calculated from bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 # see 00_qa.Rmd
@@
 \captionsetup[figure]{labelformat=empty}
-```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth predictor variables", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth predictor variables.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../tables/01_predictors.pdf")
@@
 
-```{r metric_correlations, fig.cap = "metric correlations", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 knitr::include_graphics("../figures/metric_correlations-1.pdf")
@@
 \captionsetup[figure]{labelformat=empty}
-```{r depth_model_metrics, fig.cap = "Table 2: depth model metrics", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuarcy metrics for all combinations of true and proxy slopes and distances.
+", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 knitr::include_graphics("../tables/depth_model_metrics-1.pdf")
@@
 
-### 4 x 2 grid of dotplots of depth, slope, area, dist by shape and reservoir class (7)
+```{r contrasts, fig.cap = "Lake characteristics by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_contrasts-1.pdf")
+```
 
+\pagebreak
 # Appendix
@@
 knitr::include_graphics("../figures/01_hypsography-1.pdf")
-```
-
-```{r contrasts_depth, fig.cap = "Lake maximum depth by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_contrasts_depth-1.pdf")
-```
-
-```{r constrasts_distance, fig.cap = "Contrasting distance from shore to the deepest point in the lake and distance to the lake visual center for natural lakes and reservoirs.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/gg_distance-2.pdf")
 ```

### 2020-02-28 figures/01_contrasts.Rmd
figure resizing

@@
 
-```{r 01_contrasts, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 4}
+```{r 01_contrasts, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.5}
 theme_opts <- theme(axis.text.x = element_text(angle = 90), 

### 2020-02-28 manuscript/figures.Rmd
figure resizing

@@
 
-```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', out.width="11cm", fig.align='center'}
 knitr::include_graphics("../figures/metric_correlations-1.pdf")
@@
 \captionsetup[figure]{labelformat=empty}
-```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuarcy metrics for all combinations of true and proxy slopes and distances.
-", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuarcy metrics for all combinations of true (inlakeslope, distdeepest) and proxy (slopemean, distviscenter) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../tables/depth_model_metrics-1.pdf")
@@
 
-```{r contrasts, fig.cap = "Lake characteristics by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r contrasts, fig.cap = "Lake characteristics by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 knitr::include_graphics("../figures/01_contrasts-1.pdf")
@@
 If a lake has been glaciated it is more likely to be convex.
-However, if a lake has not been glaciated there is no tendencey towards a particular shape class.", warning=FALSE, echo = FALSE, eval=TRUE}
+However, if a lake has not been glaciated there is no tendencey towards a particular shape class.", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/glaciated_vs_shapeclass-1.pdf")
@@
 
-```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/gg_distance-3.pdf")

### 2020-02-28 figures/02_depth_model.Rmd
rounding in metrics table

@@
 ```{r depth_model_metrics}
-dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds")
+dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
+  mutate_if(is.numeric, function(x) round(x, 2)) %>%
+  dplyr::select(slope, dist, everything(), -.estimator)
+
 jsta::pdf_table(dt_metrics, "../tables/depth_model_metrics-1.pdf")

### 2020-02-28 figures/01_contrasts.Rmd
typo fix

@@
   gg_area_shape_class + ggtitle("Area (ha)") + theme_opts, 
-  gg_area_shape_class + ggtitle("Inlake slope (m/m)") + theme_opts, 
-  gg_area_shape_class + ggtitle("Deepest point distance (m)") + theme_opts, 
+  gg_slope_shape_class + ggtitle("Inlake slope (m/m)") + theme_opts, 
+  gg_dist_shape_class + ggtitle("Deepest point distance (m)") + theme_opts, 
   nrow = 1)

### 2020-03-13 scripts/00_get_hypso_me.R
exclude bad me lakes

@@ -25,6 +25,10 @@ pnts <- st_transform(pnts, st_crs(lg_poly))
 pnts <- st_join(pnts, lg_poly) %>%
   dplyr::filter(!is.na(lagoslakeid))
 
+# remove problematic lakes
+bad_llids <- c(7922)
+pnts <- dplyr::filter(pnts, !(lagoslakeid %in% bad_llids))
+
 # remove lakes not in lagosne xwalk table as they likely have a
 # non-functional basin split issue
 lg_xwalk <- read.csv("data/00_lagosne/00_lagosne_xwalk.csv",

### 2020-03-13 figures/00_qa.Rmd
adjust figure dimensions
figure cleanup

--- 2020-02-11 figures/00_qa.Rmd
@@
 
-```{r lgnemanual-vs-bathy-depth, fig.height=3}
+```{r lgnemanual-vs-bathy-depth, fig.height=2.5, fig.width=5.5}
 dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)
@@
   geom_abline(aes(intercept = 0, slope = 1)) +
+  labs(color = "") +
   xlab("Calculated (m)") + ylab("Reported (m)") +
@@
 
-cowplot::plot_grid(gg_max + theme(legend.position = "none"), 
-                   gg_mean + theme(axis.title.y = element_blank()), 
-                   rel_widths = c(0.7, 1))
+cowplot::plot_grid(gg_max + cowplot::theme_cowplot() +
+                     theme(legend.position = "none"), 
+                   gg_mean + cowplot::theme_cowplot() + 
+                     theme(axis.title.y = element_blank()), 
+                   rel_widths = c(0.8, 1))
 ```

### 2020-03-13 manuscript/figures.Rmd
caption adjustments

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
+knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", dev = c("pdf"))
 knitr::opts_chunk$set(fig.pos = 'H')
@@
 
-![Diagram showing the assumption of most lake depth models that the nearshore slope is extends into the lake and that all lakes have a common relationship between nearshore and in-lake slope.
+![Diagram showing the assumption of most lake depth models that the nearshore slope extends into the lake and that all lakes have a common relationship between nearshore and in-lake slope.
 ](../figures/slope_diagram.pdf){width=75%}
@@
 
-```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth calculated from bathymetry surfaces.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth calculated from bathymetry surfaces by US State.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 # see 00_qa.Rmd

### 2020-03-13 scripts/00_get_nearshore.R
increase sample size, use lagosus polygons

@@ -5,7 +5,7 @@ source("scripts/99_utils.R")
 
 set.seed(55)
 max_buffer_dist <- 100 # to match Hollister (2011)
-n_lakes         <- 510
+n_lakes         <- 1000
 
 dt         <- read.csv("data/lagosus_depth.csv",
                     stringsAsFactors = FALSE)
@@ -27,10 +27,10 @@ ll_ids <- ll_ids[!duplicated(ll_ids)]
 
 get_slope <- function(ll_id){
   # ll_id <- ll_ids[1]
-  # ll_id <- 2645
-  ll_poly <- query_gis("LAGOS_NE_All_Lakes_4ha", "lagoslakeid", ll_id)
+  # ll_id <- 7922
+  ll_poly <- LAGOSUSgis::query_gis("LAGOS_US_All_Lakes_1ha", "lagoslakeid", ll_id)
   ll_poly <- lwgeom::st_make_valid(ll_poly)
-  ll_iws  <- query_gis("IWS", "lagoslakeid", ll_id)
+  ll_iws  <- LAGOSUSgis::query_gis("ws", "lagoslakeid", ll_id)
   # hack together a one-sided buffer [sf doesn't offer this :(]
   ll_buff <- st_buffer(ll_poly, max_buffer_dist)
   ll_buff <- st_difference(ll_buff, ll_poly)

### 2020-03-20 figures/02_depth_model.Rmd
cleanup model metrics grid figure

--- 2020-02-26 figures/02_depth_model.Rmd
@@
 
-```{r depth_model_metrics}
-dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds")
-jsta::pdf_table(dt_metrics, "../tables/depth_model_metrics-1.pdf")
+```{r depth_model_metrics, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3, fig.width=4.5}
+dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
+  mutate_if(is.numeric, function(x) as.character(round(x, 2))) %>%
+  mutate(rsq = as.numeric(rsq)) %>%
+  mutate(rsq = case_when(
+    as.character(model) == "theta_true_true" ~ NA_real_,
+    TRUE ~ rsq)) %>%
+  mutate(rmse = case_when(
+    as.character(model) == "theta_true_true" ~ "-",
+    TRUE ~ rmse)) %>%
+  mutate(model = gsub("theta_", "", model)) %>%
+  mutate(model = gsub("false", "proxy", model)) %>%
+  tidyr::separate(model, c("slope", "distance")) %>%
+  dplyr::select(slope, distance, everything(), -.estimator)
+
+ggplot(data = dt_metrics, 
+                    aes(x = reorder(slope, desc(slope)), y = distance, 
+                        fill = rsq)) +
+  geom_tile() +
+  scale_fill_viridis_c(begin = 0.4, end = 0.7, na.value = "transparent") +
+  scale_x_discrete(limits = rev(levels(dt_metrics$slope)), 
+                   position = "top") +
+  geom_text(aes(label = rmse), vjust = .5, fontface  = "bold", alpha = 1) +
+  xlab("slope") +
+  cowplot::theme_cowplot() +
+  theme(axis.line = element_blank())
 ```
@@
   mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  mutate(resid = case_when(theta == "theta_true_true" ~ NA_real_, 
+                           TRUE ~ resid)) %>%
+  mutate(theta = factor(theta, levels = c("theta_true_true", "theta_true_false", 
+                                  "theta_false_true", "theta_false_false"))) %>%
   left_join(lgus, by = "lagoslakeid")
@@
   geom_vline(aes(xintercept = 0), linetype = "dashed") +
-  facet_wrap(~ model) + 
+  facet_wrap(~ theta) + 
     theme(legend.position = "bottom",
@@
   geom_vline(aes(xintercept = 0), linetype = "dashed") +
-  facet_wrap(~ model) +
+  facet_wrap(~ theta) +
   theme(legend.position = "bottom",

### 2020-03-20 manuscript/figures.Rmd
cleanup model metrics grid figure

@@
 \captionsetup[figure]{labelformat=empty}
-```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuarcy metrics for all combinations of true (inlakeslope, distdeepest) and proxy (slopemean, distviscenter) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../tables/depth_model_metrics-1.pdf")
+```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuracy metrics for all combinations of true (inlakeslope, distdeepest) and proxy (slopemean, distviscenter) metrics.
+Tiles are colored by their R2 values and labeled with their RMSE values.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/depth_model_metrics-1.pdf")
 ```

### 2020-03-20 scripts/05_depth_model.R

@@ -134,8 +134,8 @@ abline(0, 1)
   bind_cols(data.frame(model = theta_vec)))
 
 # saveRDS(dt_train, "data/01_depth_model/depth_training.rds")
-# saveRDS(bind_rows(dt_fits),
-#         "data/01_depth_model/depth_grid.rds")
+saveRDS(bind_rows(lapply(dt_fits, function(x) x$res)),
+        "data/01_depth_model/depth_grid.rds")
 saveRDS(dt_metrics,
         "data/01_depth_model/depth_grid_metrics.rds")
 

### 2020-03-20 manuscript/figures.Rmd
move heatmap to appendix

@@
 
-```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', out.width="11cm", fig.align='center'}
-knitr::include_graphics("../figures/metric_correlations-1.pdf")
-```
-
-\captionsetup[figure]{name=Table}
-\captionsetup[figure]{labelformat=empty}
-```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuracy metrics for all combinations of true (inlakeslope, distdeepest) and proxy (slopemean, distviscenter) metrics.
+```{r depth_model_metrics, fig.cap = "Model fit and predictive accuracy metrics for all combinations of true (inlakeslope, distdeepest) and proxy (slopemean, distviscenter) metrics.
 Tiles are colored by their R2 values and labeled with their RMSE values.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
@@
 ```
-\captionsetup[figure]{name=Figure}
-\captionsetup[figure]{labelformat=default}
-\setcounter{figure}{5}
 
@@
 knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
+```
+
+```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.width="11cm", fig.align='center', out.extra=''}
+knitr::include_graphics("../figures/metric_correlations-1.pdf")
 ```

### 2020-03-20 manuscript/figures.Rmd
separate reporting of bathy predictors and lgus stats

@@
 \captionsetup[figure]{labelformat=empty}
-```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth predictor variables.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth predictor variables for modelling efforts (and for lakes continent-wide).", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../tables/01_predictors.pdf")

### 2020-03-20 tables/01_predictors.Rmd
separate reporting of bathy predictors and lgus stats

@@
 library(magrittr)
+library(stringr)
 
@@
   dplyr::select(llid, buffer_slope = slope_mean)
-dt <- read.csv("../data/lagosus_depth_predictors.csv", 
+dt_raw <- read.csv("../data/lagosus_depth_predictors.csv", 
                stringsAsFactors = FALSE) %>%
@@
                 -contains("lake_nhd"), 
-                -contains("lake_huc")) %>%
+                -contains("lake_huc"))
+
+compute_stats <- function(dt){
+  # dt <- dt_raw
+  dt <- dt %>%
   tidyr::gather(key = "variable") %>%
@@
                       Median = q50, Q25 = q25, Q75 = q75, n)
+  dt
+}
+
+# compute stats for bathymetry lakes
+dt_bathy <- dplyr::filter(dt_raw, lagos_effort == "bathymetry")
+dt_bathy <- compute_stats(dt_bathy)
+dt_bathy <- data.frame(dt_bathy)
+# compute stats for all of lagosus
+dt_all <- data.frame(compute_stats(dt_raw))
+# interleave
+dt_bathy <- bind_cols(dplyr::select(dt_bathy, variable), 
+          setNames(
+            bind_cols(lapply(names(dt_bathy)[2:5], 
+                           function(i) paste0(dt_bathy[,i], 
+                                              " (", dt_all[,i], ")"))), 
+            names(dt_bathy)[2:5]))
 
@@
 # setwd("figures")
-dt <- dt %>%
-  dplyr::filter(variable %in% 
-              c("lake_meandepth_m", 
-              names(readRDS("../data/01_depth_model/depth_training.rds")), 
-              "buffer_slope")) %>%
-  mutate(n = dt$n[which(dt$variable == "inlake_slope")])
+dt <- dt_bathy %>%
+  dplyr::filter(str_detect(variable, 
+                           "^lake_|^ws_lake_arearatio|^dist_|_slope")) %>%
+  dplyr::filter(!str_detect(variable,
+                      "totalarea|islandperimeter|between|buffer100m")) %>%
+  mutate(n = dt_bathy$n[which(dt_bathy$variable == "inlake_slope")])
 

### 2020-03-20 figures/02_depth_model.Rmd
outline tiles

@@
                         fill = rsq)) +
-  geom_tile() +
+  geom_tile(color = "black") +
   scale_fill_viridis_c(begin = 0.4, end = 0.7, na.value = "transparent") +

### 2020-03-23 figures/01_geometry.Rmd
theme cleanup

--- 2020-02-20 figures/01_geometry.Rmd
@@
   dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::filter(dist_viscenter >= dist_deepest) %>%
   dplyr::arrange(desc(factor(shape_class, 
                         levels = c("neither", "concave", "convex")))) %>%
-  ggplot() + 
-  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = shape_class)) +
+  ggplot(aes(x = dist_deepest, y = dist_viscenter, color = shape_class)) + 
+  geom_point() +
   # geom_abline(aes(intercept = 0, slope = 1)) +
@@
 
-gg_scatter_reservoir <- ggplot(data = dplyr::filter(
-  st_drop_geometry(dt), !is.na(reservoir_class))) + 
+gg_scatter_reservoir <- st_drop_geometry(dt) %>%
+  dplyr::filter(!is.na(reservoir_class)) %>%
+  dplyr::filter(dist_viscenter >= dist_deepest) %>%
+  ggplot() + 
   geom_point(aes(x = dist_deepest, y = dist_viscenter, color = reservoir_class)) +
+  scale_color_discrete(labels = c("NL     ", "Res")) + 
   # geom_abline(aes(intercept = 0, slope = 1)) +
@@
 
-```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5}
-theme_opts <- theme(legend.position = "none")
+```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5.5, fig.height=4.2}
+theme_opts <- cowplot::theme_cowplot() + 
+  theme(legend.position = "none", 
+        axis.text = element_text(size = 10), 
+        axis.title = element_text(size = 11))
 
@@
 
+cowplot::plot_grid(
   cowplot::plot_grid(
@@
                        theme_opts, 
+                     ncol = 2),
+  cowplot::plot_grid(ggplot() + geom_blank() + 
+                       theme_opts + theme(axis.line = element_blank()), 
                      gg_slope_shape +
@@
                        theme_opts,
-                     rel_widths = c(1, 0.93, 1, 1)), 
-  legend_panel, rel_widths = c(1, 0.2)
+                     ncol = 3, rel_widths = c(0.08, 1, 1)), 
+                     rel_widths = c(1, 1), nrow = 2), 
+  legend_panel, rel_widths = c(1, 0.12)
 )

### 2020-03-23 scripts/00_get_nearshore.R

@@ -5,7 +5,7 @@ source("scripts/99_utils.R")
 
 set.seed(55)
 max_buffer_dist <- 100 # to match Hollister (2011)
-n_lakes         <- 1000
+n_lakes         <- 1500
 
 dt         <- read.csv("data/lagosus_depth.csv",
                     stringsAsFactors = FALSE)

### 2020-03-23 scripts/00_get_nearshore.R
turn off data gathering, fix slope units

@@ -3,9 +3,7 @@ source("scripts/99_utils.R")
 # lagos geo module should have terrain metrics for mean slope
 # Oliver (year) found that max slope was most informative
 
-set.seed(55)
 max_buffer_dist <- 100 # to match Hollister (2011)
-n_lakes         <- 1500
 
 dt         <- read.csv("data/lagosus_depth.csv",
                     stringsAsFactors = FALSE)
@@ -16,12 +14,13 @@ dt_pred    <- read.csv("data/lagosne_depth_predictors.csv",
 
 # attempt at manual calculation of max buffer slope
 
-ll_ids            <- sample(dt_pred$lagoslakeid, n_lakes)
+ll_ids            <- unique(dt_pred$lagoslakeid)
 flist             <- list.files("data/elevatr", pattern = "\\d*.tif",
                     full.names = TRUE, include.dirs = TRUE)
 existing_surfaces <- gsub(".tif", "",
                 stringr::str_extract(flist, "\\d*(!?.tif)"))
-ll_ids <- c(ll_ids, existing_surfaces)
+ll_ids <- c(existing_surfaces)
+# ll_ids <- c(ll_ids, existing_surfaces)
 ll_ids <- ll_ids[!duplicated(ll_ids)]
 # sapply(flist[!(ll_ids %in% dt_pred$lagoslakeid)], "unlink")
 
@@ -50,9 +49,9 @@ get_slope <- function(ll_id){
   elev      <- mask(elev, as_Spatial(ll_buff))
   slope     <- terrain(elev, "slope")
 
-  slope_mean   <- mean(slope@data@values, na.rm = TRUE) * res(elev)[1]
-  slope_median <- median(slope@data@values, na.rm = TRUE) * res(elev)[1]
-  slope_max    <- max(slope@data@values, na.rm = TRUE) * res(elev)[1]
+  slope_mean   <- mean(slope@data@values, na.rm = TRUE) # * res(elev)[1]
+  slope_median <- median(slope@data@values, na.rm = TRUE) # * res(elev)[1]
+  slope_max    <- max(slope@data@values, na.rm = TRUE) # * res(elev)[1]
   slope_ne     <- dplyr::filter(dt_pred, lagoslakeid == ll_id) %>%
     dplyr::select(buffer100m_slope_mean, buffer100m_slope_max) %>%
     mutate_all(function(x) x / 10) # lagosne slopes are per 10m

### 2020-03-23 manuscript/figures.Rmd
move slope and distance scatter to main figure area

@@
 
+```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) buffer slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
+```
+
 \pagebreak
@@
 \setcounter{figure}{0}
-
-```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) buffer slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
-knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
-```
 

### 2020-03-23 figures/02_depth_model.Rmd
fight with label placement

@@
 
-```{r 02_depth_model_grid_resid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r 02_depth_model_grid_resid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5, fig.width=8}
 
@@
 
-gg_shape <- ggplot(data = dt) + 
+gg_shape <- lapply(unique(dt$theta), function(x){ 
+  ggplot(data = dplyr::filter(dt, theta == x)) + 
   geom_density(aes(x = resid, fill = shape_class), alpha = 0.75, adjust = 1.3) +
   geom_vline(aes(xintercept = 0), linetype = "dashed") +
-  facet_wrap(~ theta) + 
+    xlim(-25, 25) +
+    xlab("Residual") +
+    theme_cowplot() +
     theme(legend.position = "bottom",
@@
   guides(fill = guide_legend(nrow = 2))
+  })
+gg_shape[[1]] <- ggplot() + geom_blank() + cowplot::theme_cowplot() +
+  theme(axis.line = element_blank())
+leg            <- cowplot::get_legend(gg_shape[[2]] + labs(fill = "Shape class") +
+                                        theme(legend.direction = "horizontal") +
+                                        guides(fill = guide_legend(ncol = 2)))
+theme_opts     <- function(x) x + theme(legend.position = "none", 
+                                        axis.title = element_blank())
+rm_legend     <- function(x) x + theme(legend.position = "none")
+gg_shape      <- lapply(gg_shape, rm_legend)
+gg_shape[[2]] <- gg_shape[[2]] + theme(axis.text = element_blank(), 
+                                       axis.title = element_blank())
+gg_shape[[3]] <- gg_shape[[3]] + theme(axis.title = element_blank())
+gg_shape[[4]] <- gg_shape[[4]] + theme(axis.text.y = element_blank(), 
+                                       axis.title = element_blank())
 
-gg_shape <- ggdraw(gg_shape)  +
-  # draw_label("proxy", x = 0.83, y = 0.32, angle = 270, size = 10) +
-  # draw_label("true", x = 0.83, y = 0.65, angle = 270, size = 10) +
-  # draw_label("slope", x = 0.93, y = 0.5, size = 10, angle = 270) +
-  draw_label("proxy", x = 0.7, y = 0.89, size = 10) +
-  draw_label("true", x = 0.3, y = 0.89, size = 10) +
-  draw_label("distance", x = 0.5, y = 0.94, size = 10)# +
-  #draw_label("Shape class", x = 0.5, y = 0.98, size = 10, fontface = "bold")
+gg_shape <- cowplot::plot_grid(gg_shape[[1]], 
+          cowplot::plot_grid(
+            cowplot::plot_grid(gg_shape[[1]], gg_shape[[2]], 
+                               rel_widths = c(1, 0.85)),
+            cowplot::plot_grid(gg_shape[[3]], gg_shape[[4]], 
+                               rel_widths = c(1, 0.85)), 
+            leg, 
+            nrow = 3, rel_heights = c(0.85, 1, 0.15)), 
+          ncol = 2, rel_widths = c(0.1, 1)) %>%
+  ggdraw() +
+  draw_label("proxy", x = 0.06, y = 0.32, angle = 90, size = 10) +
+  draw_label("true", x = 0.06, y = 0.75, angle = 90, size = 10) +
+  draw_label("slope", x = 0.01, y = 0.55, size = 10, angle = 90) +
+  draw_label("proxy", x = 0.8, y = 0.9, size = 10) +
+  draw_label("true", x = 0.35, y = 0.9, size = 10) +
+  draw_label("distance", x = 0.55, y = 0.96, size = 10) +
+  draw_label("Residual distribution", 
+             x = 0.55, y = 0.99, size = 10, fontface = "bold")
 
-gg_reservoir <- ggplot(data = dt) + 
+gg_reservoir <- lapply(unique(dt$theta), function(x){ 
+  ggplot(data = dplyr::filter(dt, theta == x)) + 
   geom_density(aes(x = resid, fill = reservoir_class), alpha = 0.75, adjust = 1.3) +
   geom_vline(aes(xintercept = 0), linetype = "dashed") +
-  facet_wrap(~ theta) +
+    xlim(-25, 25) +
+    xlab("Residual") +
+    theme_cowplot() +
     theme(legend.position = "bottom",
@@
         strip.text = element_blank(),
-        axis.title.y = element_blank(),
-        axis.text.y = element_blank(),
-        plot.margin = unit(c(1.2,1.2,0.2,0.2), "cm")) +
+        plot.margin = unit(c(1.2,0.2,0.2,0.2), "cm"))  +
   guides(fill = guide_legend(nrow = 2))
+  })
+gg_reservoir[[1]] <- ggplot() + geom_blank() + cowplot::theme_cowplot() +
+  theme(axis.line = element_blank())
+leg            <- cowplot::get_legend(gg_reservoir[[2]] + labs(fill = "Reservoir class") +
+                                        theme(legend.direction = "horizontal") +
+                                        guides(fill = guide_legend(ncol = 2)))
+theme_opts     <- function(x) x + theme(legend.position = "none", 
+                                        axis.title = element_blank())
+rm_legend     <- function(x) x + theme(legend.position = "none")
+gg_reservoir      <- lapply(gg_reservoir, rm_legend)
+gg_reservoir[[2]] <- gg_reservoir[[2]] + theme(axis.text = element_blank(), 
+                                       axis.title = element_blank())
+gg_reservoir[[3]] <- gg_reservoir[[3]] + theme(axis.title = element_blank())
+gg_reservoir[[4]] <- gg_reservoir[[4]] + theme(axis.text.y = element_blank(), 
+                                       axis.title = element_blank())
 
-gg_reservoir <- ggdraw(gg_reservoir) +
-  draw_label("proxy", x = 0.83, y = 0.44, angle = 270, size = 10) +
-  draw_label("true", x = 0.83, y = 0.77, angle = 270, size = 10) +
-  draw_label("slope", x = 0.93, y = 0.62, size = 10, angle = 270) +
-  draw_label("proxy", x = 0.7, y = 0.89, size = 10) +
-  draw_label("true", x = 0.3, y = 0.89, size = 10) +
-  draw_label("distance", x = 0.5, y = 0.94, size = 10)# +
-  # draw_label("Reservoir class", x = 0.5, y = 0.98, size = 10, fontface = "bold")
+gg_reservoir <- cowplot::plot_grid(gg_reservoir[[1]], 
+          cowplot::plot_grid(
+            cowplot::plot_grid(gg_reservoir[[1]], gg_reservoir[[2]], 
+                               rel_widths = c(1, 0.85)),
+            cowplot::plot_grid(gg_reservoir[[3]], gg_reservoir[[4]], 
+                               rel_widths = c(1, 0.85)), 
+            leg, 
+            nrow = 3, rel_heights = c(0.85, 1, 0.15)), 
+          ncol = 2, rel_widths = c(0.1, 1)) %>%
+  ggdraw() +
+  draw_label("proxy", x = 0.06, y = 0.32, angle = 90, size = 10) +
+  draw_label("true", x = 0.06, y = 0.75, angle = 90, size = 10) +
+  draw_label("slope", x = 0.01, y = 0.55, size = 10, angle = 90) +
+  draw_label("proxy", x = 0.8, y = 0.9, size = 10) +
+  draw_label("true", x = 0.35, y = 0.9, size = 10) +
+  draw_label("distance", x = 0.55, y = 0.96, size = 10) +
+  draw_label("Residual distribution", 
+             x = 0.55, y = 0.99, size = 10, fontface = "bold")
 
-cowplot::plot_grid(gg_shape, gg_reservoir, 
-                   nrow = 1, rel_widths = c(1, 0.95))
+cowplot::plot_grid(gg_shape, gg_reservoir, ncol = 2)
 ```

### 2020-03-24 README.Rmd
add stand-in title
towards a complete data dictionary

@@
 
-## Bathymetry manuscript
+## Bathymetry data from thousands of lakes show that difficulty modeling in-lake slope limits lake depth prediction
 
@@
 <summary>Data dictionary</summary>
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 # see scripts/99_make_taxonomy.R
-
-dt <- read.csv("data/00_bathy_depth/00_bathy_depth.csv", stringsAsFactors = FALSE)
-knitr::kable(setNames(data.frame(names(dt)), "field")) 
+library(dplyr)
+taxonomy <- read.csv("data/taxonomy.csv", 
+                     stringsAsFactors = FALSE)
+dt       <- read.csv("data/00_bathy_depth/00_bathy_depth.csv", 
+               stringsAsFactors = FALSE) %>%
+  names() %>%  as.data.frame() %>%  setNames("col_name") %>%
+  left_join(taxonomy)
+knitr::kable(dt) 
 ```
@@
 
-Data: [bathymetry derived lake geometry](data/00_bathy_depth/bathy_geometry.csv)
+Data: [bathymetry derived lake geometry](data/depth_predictors.csv)
 
@@
 
-dt <- read.csv("data/00_bathy_depth/bathy_geometry.csv", stringsAsFactors = FALSE)
+dt <- read.csv("data/depth_predictors.csv", stringsAsFactors = FALSE)
 knitr::kable(setNames(data.frame(names(dt)), "field"))

### 2020-03-24 scripts/99_make_taxonomy.R

@@ -1,0 +1,65 @@
+# Make taxonomy table by referencing `locus_taxonomy` file on Dropbox**
+
+# dt <- read.csv("data/00_bathy_depth/00_bathy_depth.csv",
+#                stringsAsFactors = FALSE)
+# dt2 <- read.csv("data/depth_predictors.csv",
+#                stringsAsFactors = FALSE)
+# cat(unique(c(names(dt), names(dt2))))
+
+taxonomy_key <- bind_rows(
+  c(col_name = "llid",
+    description = "unique lake identifier developed for LAGOS-US"),
+  c(col_name = "state",
+    description = "abbreviation of the state used to search for a lake's depth"),
+  c(col_name = "max_depth_m",
+    description = "lake maximum depth in meters"),
+  c(col_name = "mean_depth_m",
+    description = "lake mean depth in meters"),
+  c(col_name = "source",
+    description = "url link of the source of depth data"),
+  c(col_name = "lake_waterarea_ha",
+    description = "surface area of lake waterbody polygon from NHD (excludes islands) by way of the LAGOS-US Locus module"),
+  c(col_name = "lake_connectivity_permanent",
+    description = "connectivity of focal lake to upstream features (DrainageLK = drainage lake with an upstream lake, Drainage = drainage lake with upstream stream, Headwater = lake with outlet but no inlet, Isolated = lake with no inlets or outlets) by way of the LAGOS-US Locus module"),
+  c(col_name = "lake_lat_decdeg", description = "the latitude of the lake center point (NAD83) from the LAGOS-US Locus module"),
+  c(col_name = "lake_lon_decdeg",
+    description = "the longitude of the lake center point (NAD83) from the LAGOS-US Locus module"),
+  c(col_name = "lat",
+    description = "the latitude of the lake center point"),
+  c(col_name = "long",
+    description = "the longitude of the lake center point"),
+  c(col_name = "lagoslakeid",
+    description = "unique lake identifier developed for LAGOS-US"),
+  c(col_name = "lake_maxdepth_m",
+    description = "lake maximum depth in meters"),
+  c(col_name = "maxdepth_true_true",
+    description = "lake maximum depth calculated from inlake_slope and dist_deepest in meters"),
+  c(col_name = "maxdepth_true_false",
+    description = "lake maximum depth calculated from inlake_slope and dist_viscenter"),
+  c(col_name = "maxdepth_false_true",
[…]

### 2020-03-24 README.Rmd
fight with kable rendering

@@
   names() %>%  as.data.frame() %>%  setNames("col_name") %>%
-  left_join(taxonomy)
+  left_join(taxonomy) %>% data.frame()
+
 knitr::kable(dt) 
@@
 
-Data: [bathymetry derived lake geometry](data/depth_predictors.csv)
+Datad: [bathymetry derived lake geometry](data/depth_predictors.csv)
 

### 2020-03-24 README.Rmd
complete data dictionary

--- 2020-03-24 README.Rmd
@@
 
-<details>
-<summary>Data dictionary</summary>
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
   names() %>%  as.data.frame() %>%  setNames("col_name") %>%
-  left_join(taxonomy)
+  left_join(taxonomy) %>% data.frame() %>%
+  dplyr::filter(!is.na(description)) %>%
+  setNames(c("column name", "description"))
+
 knitr::kable(dt)
 ```
-</details> &nbsp;
 
@@
 
-<details>
-<summary>Data dictionary</summary>
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 # see scripts/99_make_taxonomy.R
+dt <- read.csv("data/depth_predictors.csv", 
+               stringsAsFactors = FALSE) %>% 
+  names() %>%  as.data.frame() %>%  setNames("col_name") %>%
+  left_join(taxonomy) %>% data.frame() %>%
+  dplyr::filter(!is.na(description)) %>%
+  setNames(c("column name", "description"))
 
-dt <- read.csv("data/depth_predictors.csv", stringsAsFactors = FALSE)
-knitr::kable(setNames(data.frame(names(dt)), "field"))
+knitr::kable(dt)
 ```
-</details>

### 2020-03-24 scripts/99_make_taxonomy.R

@@ -6,6 +6,8 @@
 #                stringsAsFactors = FALSE)
 # cat(unique(c(names(dt), names(dt2))))
 
+library(dplyr)
+
 taxonomy_key <- bind_rows(
   c(col_name = "llid",
     description = "unique lake identifier developed for LAGOS-US"),
@@ -56,8 +58,7 @@ taxonomy_key <- bind_rows(
     description = "surface area of islands within outer boundary of lake waterbody polygon from NHD by way of the LAGOS-US Locus module"),
   c(col_name = "lake_elevation_m",
     description = "the elevation of the lake polygon centroid, in meters (referenced to the North American Vertical Datum of 1988 (NAVD88)) and obtained from the National Elevation Dataset by way of the LAGOS-US Locus module"),
-  c(col_name = "lake_shorelinedevfactor_nounits",
-    description = "shoreline development factor calculated as the lake_perimeter_m / (2*√(π*lake_waterarea_ha*10000)) by way of the LAGOS-US Locus module"),
+  c(col_name = "lake_shorelinedevfactor_nounits", description = "shoreline development factor calculated as the lake_perimeter_m / (2*√(π*lake_waterarea_ha*10000)) by way of the LAGOS-US Locus module"),
   c(col_name = "ws_lake_arearatio", description = "ratio between interlake watershed area and lake water area by way of the LAGOS-US Locus module"),
   c(col_name = "hu4_zoneid", description = "unique identifier assigned by LAGOS-US for zones in the spatial division HU4")
   )

### 2020-03-24 README.Rmd
better link text
title adjustment

@@
 
-## Bathymetry data from thousands of lakes show that difficulty modeling in-lake slope limits lake depth prediction
+## Bathymetry data from thousands of lakes show that difficulty modeling in-lake slope confounds lake depth prediction
 
@@
 
-Data: [bathymetry derived lake geometry](data/depth_predictors.csv)
+Data: [bathymetry derived depth predictors](data/depth_predictors.csv)
 

### 2020-03-24 README.Rmd
cleanup

@@
 
-Data Summary: [manuscript/data.pdf](manuscript/data.pdf)
-
 Figures: [manuscript/figures.pdf](manuscript/figures.pdf)
-
-Tables: [manuscript/tables.pdf](manuscript/tables.pdf)
-
-Data: [bathymetry derived lake depth](data/00_bathy_depth/00_bathy_depth.csv)
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# see scripts/99_make_taxonomy.R
-library(dplyr)
-taxonomy <- read.csv("data/taxonomy.csv", 
-                     stringsAsFactors = FALSE)
-dt       <- read.csv("data/00_bathy_depth/00_bathy_depth.csv", 
-               stringsAsFactors = FALSE) %>%
-  names() %>%  as.data.frame() %>%  setNames("col_name") %>%
-  left_join(taxonomy) %>% data.frame() %>%
-  dplyr::filter(!is.na(description)) %>%
-  setNames(c("column name", "description"))
-
-knitr::kable(dt)
-```
 
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
+library(dplyr)
+taxonomy <- read.csv("data/taxonomy.csv", 
+                     stringsAsFactors = FALSE)
+
 # see scripts/99_make_taxonomy.R
@@
 ```
+
+Data: [bathymetry derived lake depth](data/00_bathy_depth/00_bathy_depth.csv)
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# see scripts/99_make_taxonomy.R
+dt       <- read.csv("data/00_bathy_depth/00_bathy_depth.csv", 
+               stringsAsFactors = FALSE) %>%
+  names() %>%  as.data.frame() %>%  setNames("col_name") %>%
+  left_join(taxonomy) %>% data.frame() %>%
+  dplyr::filter(!is.na(description)) %>%
+  setNames(c("column name", "description"))
+
+knitr::kable(dt)
+```

### 2020-03-25 figures/02_depth_model.Rmd
carry over theta renaming to plot generation

@@
   mutate(rsq = case_when(
-    as.character(model) == "theta_true_true" ~ NA_real_,
+    as.character(model) == "maxdepth_true_true" ~ NA_real_,
     TRUE ~ rsq)) %>%
   mutate(rmse = case_when(
-    as.character(model) == "theta_true_true" ~ "-",
+    as.character(model) == "maxdepth_true_true" ~ "-",
     TRUE ~ rmse)) %>%
-  mutate(model = gsub("theta_", "", model)) %>%
+  mutate(model = gsub("maxdepth_", "", model)) %>%
   mutate(model = gsub("false", "proxy", model)) %>%
@@
   mutate(lagoslakeid = as.character(lagoslakeid)) %>%
-  mutate(resid = case_when(theta == "theta_true_true" ~ NA_real_, 
+  mutate(resid = case_when(maxdepth == "maxdepth_true_true" ~ NA_real_, 
                            TRUE ~ resid)) %>%
-  mutate(theta = factor(theta, levels = c("theta_true_true", "theta_true_false", 
-                                  "theta_false_true", "theta_false_false"))) %>%
+  mutate(maxdepth = factor(maxdepth, levels = c("maxdepth_true_true", "maxdepth_true_false", 
+                                  "maxdepth_false_true", "maxdepth_false_false"))) %>%
   left_join(lgus, by = "lagoslakeid")
 
-gg_shape <- lapply(unique(dt$theta), function(x){ 
-  ggplot(data = dplyr::filter(dt, theta == x)) + 
+gg_shape <- lapply(unique(dt$maxdepth), function(x){ 
+  ggplot(data = dplyr::filter(dt, maxdepth == x)) + 
   geom_density(aes(x = resid, fill = shape_class), alpha = 0.75, adjust = 1.3) +
@@
 
-gg_reservoir <- lapply(unique(dt$theta), function(x){ 
-  ggplot(data = dplyr::filter(dt, theta == x)) + 
+gg_reservoir <- lapply(unique(dt$maxdepth), function(x){ 
+  ggplot(data = dplyr::filter(dt, maxdepth == x)) + 
   geom_density(aes(x = resid, fill = reservoir_class), alpha = 0.75, adjust = 1.3) +
@@
 
-```{r 02_depth_oliver, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.4}
+```{r 02_depth_oliver, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide', fig.height=3.4}
 # setwd("figures")

### 2020-03-25 scripts/05_depth_model.R
exclude class predictors, use random forest model

@@ -40,15 +40,15 @@ dt_raw$dist_viscenter_norm <- scales::rescale(dt_raw$dist_viscenter,
 # plot(dt_raw$dist_viscenter_norm, dt_raw$dist_deepest)
 # abline(0, 1)
 
-dt_raw$theta_true_true   <- calc_depth(dt_raw$inlake_slope, dt_raw$dist_deepest)
-dt_raw$theta_true_false  <- calc_depth(dt_raw$inlake_slope, dt_raw$dist_viscenter_norm)
-dt_raw$theta_false_true  <- calc_depth(dt_raw$slope_mean_norm, dt_raw$dist_deepest)
-dt_raw$theta_false_false <- calc_depth(dt_raw$slope_mean_norm, dt_raw$dist_viscenter_norm)
+dt_raw$maxdepth_true_true   <- calc_depth(dt_raw$inlake_slope, dt_raw$dist_deepest)
+dt_raw$maxdepth_true_false  <- calc_depth(dt_raw$inlake_slope, dt_raw$dist_viscenter_norm)
+dt_raw$maxdepth_false_true  <- calc_depth(dt_raw$slope_mean_norm, dt_raw$dist_deepest)
+dt_raw$maxdepth_false_false <- calc_depth(dt_raw$slope_mean_norm, dt_raw$dist_viscenter_norm)
 
-# plot(dt_raw$theta_true_true, dt_raw$lake_maxdepth_m)
-# plot(dt_raw$theta_true_false, dt_raw$lake_maxdepth_m)
-# plot(dt_raw$theta_false_true, dt_raw$lake_maxdepth_m)
-# plot(dt_raw$theta_false_false, dt_raw$lake_maxdepth_m)
+# plot(dt_raw$maxdepth_true_true, dt_raw$lake_maxdepth_m)
+# plot(dt_raw$maxdepth_true_false, dt_raw$lake_maxdepth_m)
+# plot(dt_raw$maxdepth_false_true, dt_raw$lake_maxdepth_m)
+# plot(dt_raw$maxdepth_false_false, dt_raw$lake_maxdepth_m)
 # abline(0, 1)
 
 # data prep
@@ -60,8 +60,8 @@ set.seed(1234)
 
 dt <- dt_raw %>%
   dplyr::select(lagoslakeid, lake_maxdepth_m,
-                theta_true_true, theta_true_false,
-                theta_false_true, theta_false_false,
+                maxdepth_true_true, maxdepth_true_false,
[…]

### 2020-03-25 figures/02_depth_model.Rmd
create variable importance plot

@@
 knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
-# setwd("figures")
+if(interactive()){
+  setwd("figures")
+}
 source("../scripts/99_utils.R")
@@
   theme(axis.line = element_blank())
+```
+
+```{r 02_depth_model_importance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5, fig.width=8}
+dt_fits <- readRDS("../data/01_depth_model/depth_fits.rds")
+
+tidy_importance <- function(x){
+  # x <- dt_fits[[2]]$fit$fit
+  res <- ranger::importance(x) %>%
+  data.frame() %>% cbind(row.names(.), .[,1]) %>%
+  setNames(c("percent_inc_mse", "variable"))
+  row.names(res) <- NULL
+  res <- res[,c(2, 1)]
+  res <- arrange(res, percent_inc_mse)
+  # res$variable <- factor(res$variable, levels = res$variable)
+  res <- mutate(res, variable = as.character(variable))
+  res <- mutate(res, variable = case_when(
+    stringr::str_detect(variable, "maxdepth") ~ "trig_depth", 
+    TRUE ~ variable)
+    )
+  res
+}
+
+dt_imp <- lapply(2:4, function(i) tidy_importance(dt_fits[[i]]$fit$fit))
+model_names <- c("true_proxy", 
+                 "proxy_true", 
+                 "proxy_proxy")
+dt_imp <- bind_rows(lapply(1:3, function(i){
+  mutate(dt_imp[[i]], model = model_names[i])
+  }))
+dt_imp$model <- factor(dt_imp$model, levels = model_names)
+
+var_order <- 
+  group_by(dt_imp, variable) %>% 
+  summarize(max = max(percent_inc_mse)) %>%
+  arrange(max) %>%
+  pull(variable)
+dt_imp$variable <- factor(dt_imp$variable, levels = var_order)
+
+ggplot() + 
+  geom_point(data = dt_imp, aes(percent_inc_mse, variable, color = model)) +   
+  # geom_hline(linetype = 3, size = 1, colour = "gray", 
+  #                yintercept = 1:nlevels(dt$variable)) + 
+  theme(text = element_text(family="Times"),
+        panel.background = element_blank(), panel.grid = element_blank(),
+        panel.border = element_rect(fill = NA), # legend.position = "none", 
+        axis.title.x = element_text(family="sans", face = "bold", 
+        vjust = -0.5, size = 12), 
+        axis.text = element_text(family="sans",size = 11, 
+                                 color = "black")) + 
+        ylab("") + xlab("Percent Increase in Mean Square Error")
 ```

### 2020-03-25 manuscript/figures.Rmd
create variable importance plot

@@
 
+```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) buffer slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
+```
+
 ```{r depth_model_metrics, fig.cap = "Model fit and predictive accuracy metrics for all combinations of true (inlakeslope, distdeepest) and proxy (slopemean, distviscenter) metrics.
@@
 
+\pagebreak
+# Appendix
+\setcounter{figure}{0}
+
+```{r 02_depth_model_importance, fig.cap = "Importance plot for random forest variables showing increase in mean square error.
+Higher values indicate greater importance to model predictions.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")
+```
+
 ```{r contrasts, fig.cap = "Lake characteristics by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
@@
 ```
-
-```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) buffer slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
-```
-
-\pagebreak
-# Appendix
-\setcounter{figure}{0}
 

### 2020-03-25 scripts/05_depth_model.R

@@ -151,6 +151,7 @@ abline(0, 1)
   bind_cols(data.frame(model = maxdepth_vec)))
 
 # saveRDS(dt_train, "data/01_depth_model/depth_training.rds")
+saveRDS(dt_fits, "data/01_depth_model/depth_fits.rds")
 saveRDS(bind_rows(lapply(dt_fits, function(x) x$res)),
         "data/01_depth_model/depth_grid.rds")
 saveRDS(dt_metrics,

### 2020-03-25 manuscript/figures.Rmd
organization

@@
 
-```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.width="11cm", fig.align='center', out.extra=''}
+```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=FALSE, out.width="11cm", fig.align='center', out.extra=''}
 knitr::include_graphics("../figures/metric_correlations-1.pdf")
@@
 
-```{r heatmap, fig.cap = "Correlation matrix heatmap.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r heatmap, fig.cap = "Correlation matrix heatmap.", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/slope-compare-1.pdf")
@@
 
-```{r 02_hypso_model_fitted, fig.cap = "Confusion matrix comparing two classification methods", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r 02_hypso_model_fitted, fig.cap = "Confusion matrix comparing two classification methods", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/02_hypso_model_fitted-1.pdf")
@@
 
-```{r 02_hypso_model_importance, fig.cap = "Recusive Partitioning variable importance.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r 02_hypso_model_importance, fig.cap = "Recusive Partitioning variable importance.", warning=FALSE, echo = FALSE, eval=FALSE}
 knitr::include_graphics("../figures/02_hypso_model_importance-1.pdf")

### 2020-03-25 figures/02_depth_model.Rmd
align axes among plots

@@
 
-```{r depth_model_metrics, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3, fig.width=4.5}
+```{r depth_model_metrics, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=2.8, fig.width=4.2}
 dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
@@
 ggplot(data = dt_metrics, 
-                    aes(x = reorder(slope, desc(slope)), y = distance, 
+                    aes(y = slope, 
+                        x = reorder(distance, desc(distance)),
                         fill = rsq)) +
@@
   geom_text(aes(label = rmse), vjust = .5, fontface  = "bold", alpha = 1) +
-  xlab("slope") +
+  xlab("distance") + ylab("slope") +
   cowplot::theme_cowplot() +
@@
   geom_density(aes(x = resid, fill = shape_class), alpha = 0.75, adjust = 1.3) +
+  scale_fill_manual(values = hue_pal()(3)[1:2]) +
   geom_vline(aes(xintercept = 0), linetype = "dashed") +
@@
   draw_label("true", x = 0.35, y = 0.9, size = 10) +
-  draw_label("distance", x = 0.55, y = 0.96, size = 10) +
-  draw_label("Residual distribution", 
-             x = 0.55, y = 0.99, size = 10, fontface = "bold")
+  draw_label("distance", x = 0.55, y = 0.96, size = 10)#  +
+  # draw_label("Residual distribution", 
+  #            x = 0.55, y = 0.99, size = 10, fontface = "bold")
 
@@
   ggdraw() +
-  draw_label("proxy", x = 0.06, y = 0.32, angle = 90, size = 10) +
-  draw_label("true", x = 0.06, y = 0.75, angle = 90, size = 10) +
-  draw_label("slope", x = 0.01, y = 0.55, size = 10, angle = 90) +
+  # draw_label("proxy", x = 0.06, y = 0.32, angle = 90, size = 10) +
+  # draw_label("true", x = 0.06, y = 0.75, angle = 90, size = 10) +
+  # draw_label("slope", x = 0.01, y = 0.55, size = 10, angle = 90) +
   draw_label("proxy", x = 0.8, y = 0.9, size = 10) +
   draw_label("true", x = 0.35, y = 0.9, size = 10) +
-  draw_label("distance", x = 0.55, y = 0.96, size = 10) +
-  draw_label("Residual distribution", 
-             x = 0.55, y = 0.99, size = 10, fontface = "bold")
+  draw_label("distance", x = 0.55, y = 0.96, size = 10) # +
+  # draw_label("Residual distribution", 
+  #            x = 0.55, y = 0.99, size = 10, fontface = "bold")
 
-cowplot::plot_grid(gg_shape, gg_reservoir, ncol = 2)
+cowplot::plot_grid(gg_shape, gg_reservoir, ncol = 2, rel_widths = c(1, 0.9))
 ```

### 2020-03-25 tables/01_predictors.Rmd
towards sample size fixes

@@
   dplyr::select(llid, buffer_slope = slope_mean)
-dt_raw <- read.csv("../data/lagosus_depth_predictors.csv", 
+dt_all <- read.csv("../data/lagosus_depth_predictors.csv", 
                stringsAsFactors = FALSE) %>%
-  distinct(lagoslakeid, .keep_all = TRUE) %>%
-  left_join(nearshore, by = c("lagoslakeid" = "llid")) %>%
+  distinct(lagoslakeid, .keep_all = TRUE)
+dt_bathy <- dt_all %>%
+  right_join(nearshore, by = c("lagoslakeid" = "llid")) %>%
   dplyr::select(-contains("lagoslakeid"), 
@@
 # compute stats for bathymetry lakes
-dt_bathy <- dplyr::filter(dt_raw, lagos_effort == "bathymetry")
 dt_bathy <- compute_stats(dt_bathy)
@@
 # compute stats for all of lagosus
-dt_all <- data.frame(compute_stats(dt_raw))
+dt_all <- data.frame(compute_stats(dt_all))
+dt_all <- dplyr::filter(dt_all, variable %in% dt_bathy$variable) 
 # interleave
-dt_bathy <- bind_cols(dplyr::select(dt_bathy, variable), 
-          setNames(
-            bind_cols(lapply(names(dt_bathy)[2:5], 
+dt_combined <- bind_cols(lapply(names(dt_bathy)[2:5], 
                            function(i) paste0(dt_bathy[,i], 
-                                              " (", dt_all[,i], ")"))), 
-            names(dt_bathy)[2:5]))
+                                              " (", dt_all[,i], ")"))) %>%
+  setNames(names(dt_bathy)[2:5])
+
+dt_combined <- bind_cols(
+  dplyr::select(dt_bathy, variable), 
+  dt_combined)
 
@@
 # setwd("figures")
-dt <- dt_bathy %>%
+dt <- dt_combined %>%
   dplyr::filter(str_detect(variable, 
@@
   dplyr::filter(!str_detect(variable,
-                      "totalarea|islandperimeter|between|buffer100m")) %>%
-  mutate(n = dt_bathy$n[which(dt_bathy$variable == "inlake_slope")])
+                      "totalarea|islandperimeter|between|buffer100m")) # %>%
+  # mutate(n = dt_bathy$n[which(dt_bathy$variable == "inlake_slope")])
+
+# remove sample size reporting for bathy specific variables
+dt[
+  dt$variable %in% c("dist_deepest", "dist_viscenter", "inlake_slope"), 2:5] <- 
+  dt[
+  dt$variable %in% c("dist_deepest", "dist_viscenter", "inlake_slope"), 2:5] %>%
+apply(2, function(x) stringr::str_replace(x, "\\(([^\\)]+)\\)", "(-)"))
 

### 2020-03-25 figures/02_depth_model.Rmd
importance legend adjustments

@@
 
-```{r 02_depth_model_importance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5, fig.width=8}
+```{r 02_depth_model_importance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5, fig.width=6}
 dt_fits <- readRDS("../data/01_depth_model/depth_fits.rds")
@@
   geom_point(data = dt_imp, aes(percent_inc_mse, variable, color = model)) +   
-  # geom_hline(linetype = 3, size = 1, colour = "gray", 
-  #                yintercept = 1:nlevels(dt$variable)) + 
+  labs(color = "slope-distance\nmeasure") +
+  geom_hline(linetype = 3, size = 1, colour = "gray",
+                 yintercept = 1:nlevels(dt_imp$variable)) +
   theme(text = element_text(family="Times"),
@@
         axis.text = element_text(family="sans",size = 11, 
-                                 color = "black")) + 
+                                 color = "black"), 
+         legend.key = element_blank()) + 
         ylab("") + xlab("Percent Increase in Mean Square Error")

### 2020-03-26 README.Rmd
use pretty variable names

@@
   dplyr::filter(!is.na(description)) %>%
-  setNames(c("column name", "description"))
+  setNames(c("column name", "description", "variable")) %>%
+  select(`column name`, variable, description)
 
+options(knitr.kable.NA = '-')
 knitr::kable(dt)
@@
   dplyr::filter(!is.na(description)) %>%
-  setNames(c("column name", "description"))
+  setNames(c("column name", "description", "variable")) %>%
+  select(`column name`, variable, description)
 

### 2020-03-26 figures/02_depth_model.Rmd
use pretty variable names

--- 2020-03-25 figures/02_depth_model.Rmd
@@
 
-```{r depth_model_metrics, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=2.8, fig.width=4.2}
+```{r depth_model_metrics, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=2.6, fig.width=4}
 dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
@@
 
-```{r 02_depth_model_importance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5, fig.width=8}
+```{r 02_depth_model_importance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.2, fig.width=5}
 dt_fits      <- readRDS("../data/01_depth_model/depth_fits.rds")
+taxonomy_key <- read.csv("../data/taxonomy.csv", stringsAsFactors = FALSE)
 
@@
 var_order <- 
-  group_by(dt_imp, variable) %>% 
+  left_join(dt_imp, taxonomy_key, by = c("variable" = "col_name")) %>%
+  group_by(pretty) %>% 
   summarize(max = max(percent_inc_mse)) %>%
   arrange(max) %>%
-  pull(variable)
-dt_imp$variable <- factor(dt_imp$variable, levels = var_order)
+  pull(pretty)
+dt_imp <- left_join(dt_imp, taxonomy_key, by = c("variable" = "col_name")) %>%
+  mutate(pretty = factor(pretty, levels = var_order)) %>%
+  mutate(variable = pretty) %>%  dplyr::select(-description) %>%
+  dplyr::select(-pretty)
 
@@
         panel.border = element_rect(fill = NA), # legend.position = "none", 
-        axis.title.x = element_text(family="sans", face = "bold", 
-        vjust = -0.5, size = 12), 
-        axis.text = element_text(family="sans",size = 11, 
-                                 color = "black")) + 
+        axis.title.x = element_text(family = "sans", face = "bold", 
+        vjust = -0.5, size = 10), 
+        axis.text = element_text(family="sans",size = 9, 
+                                 color = "black"), 
+         legend.key = element_blank()) + 
         ylab("") + xlab("Percent Increase in Mean Square Error")

### 2020-03-26 manuscript/figures.Rmd
use pretty variable names

@@
 
-```{r 02_hypso_model_importance, fig.cap = "Recusive Partitioning variable importance.", warning=FALSE, echo = FALSE, eval=FALSE}
+```{r 02_hypso_model_importance, fig.cap = "Recusive Partitioning variable importance.", warning=FALSE, echo = FALSE, eval=FALSE, out.extra=''}
 knitr::include_graphics("../figures/02_hypso_model_importance-1.pdf")

### 2020-03-26 tables/01_predictors.Rmd
use pretty variable names

@@
 
+taxonomy_key <- read.csv("../data/taxonomy.csv", stringsAsFactors = FALSE)
+
+dt <- left_join(dt, taxonomy_key, by = c("variable" = "col_name")) %>%
+  dplyr::mutate(variable = pretty) %>% select(-pretty, -description)
+
+options(knitr.kable.NA = '-')
 knitr::kable(dt, digits = 0, format = "markdown") 

### 2020-03-26 scripts/99_make_taxonomy.R

@@ -12,15 +12,19 @@ taxonomy_key <- bind_rows(
   c(col_name = "llid",
     description = "unique lake identifier developed for LAGOS-US"),
   c(col_name = "state",
-    description = "abbreviation of the state used to search for a lake's depth"),
+    description = "abbreviation of the state used to search for a lake's depth",
+    pretty = "State"),
   c(col_name = "max_depth_m",
-    description = "lake maximum depth in meters"),
+    description = "lake maximum depth in meters",
+    pretty = "Max depth (m)"),
   c(col_name = "mean_depth_m",
-    description = "lake mean depth in meters"),
+    description = "lake mean depth in meters",
+    pretty = "Mean depth (m)"),
   c(col_name = "source",
     description = "url link of the source of depth data"),
   c(col_name = "lake_waterarea_ha",
-    description = "surface area of lake waterbody polygon from NHD (excludes islands) by way of the LAGOS-US Locus module"),
+    description = "surface area of lake waterbody polygon from NHD (excludes islands) by way of the LAGOS-US Locus module",
+    pretty = "Area (ha)"),
   c(col_name = "lake_connectivity_permanent",
     description = "connectivity of focal lake to upstream features (DrainageLK = drainage lake with an upstream lake, Drainage = drainage lake with upstream stream, Headwater = lake with outlet but no inlet, Isolated = lake with no inlets or outlets) by way of the LAGOS-US Locus module"),
   c(col_name = "lake_lat_decdeg", description = "the latitude of the lake center point (NAD83) from the LAGOS-US Locus module"),
@@ -33,7 +37,11 @@ taxonomy_key <- bind_rows(
   c(col_name = "lagoslakeid",
     description = "unique lake identifier developed for LAGOS-US"),
   c(col_name = "lake_maxdepth_m",
-    description = "lake maximum depth in meters"),
+    description = "lake maximum depth in meters",
+    pretty = "Max depth (m)"),
+  c(col_name = "lake_meandepth_m",
+    description = "lake mean depth in meters",
+    pretty = "Mean depth (m)"),
   c(col_name = "maxdepth_true_true",
     description = "lake maximum depth calculated from inlake_slope and dist_deepest in meters"),
   c(col_name = "maxdepth_true_false",
@@ -42,25 +50,46 @@ taxonomy_key <- bind_rows(
     description = "lake maximum depth calculated from slope_mean and dist_deepest in meters"),
   c(col_name = "maxdepth_false_false",
[…]

### 2020-03-26 scripts/04_depth_predictors.R
add wd placeholder

@@ -1,3 +1,4 @@
+# setwd("../")
 source("scripts/99_utils.R")
 
 # unlink("data/lagosus_depth_predictors.csv")

### 2020-03-26 figures/01_contrasts.Rmd
towards switching between violin and pointrange plots

@@
 
+```{r define-plots}
+contrast_plot <- function(..., plot_type){
+  if(plot_type == "violin"){
+    violin_plot(...)
+  }else{
+    pointrange_plot(...)
+  }
+}
+
+violin_plot <- function(x, grp, varb, label){
+  ggplot(data = x) +
+  geom_violin(aes(x = UQ(rlang::sym(grp)), y = UQ(rlang::sym(varb))), 
+              adjust = 0.8, scale = "count", 
+              draw_quantiles = c(0.25, 0.5, 0.75)) +
+    ylab(label) + xlab("")
+}
+
+quantile_collapse <- function(x){
+  list(tibble::enframe(
+    quantile(x, probs = c(0.05,0.5,0.95), na.rm = TRUE)))
+}
+
+pointrange_plot <- function(x, grp, varb, label){
+  x %>%
+  group_by(UQ(rlang::sym(grp))) %>%
+    dplyr::select(UQ(rlang::sym(grp)), UQ(rlang::sym(varb))) %>%
+    summarize_all(quantile_collapse) %>%
+    unnest() %>%
+  tidyr::spread(name, value) %>%
+  ggplot() +
+geom_pointrange(aes(x = UQ(rlang::sym(grp)), y = `50%`, 
+                      ymin = `5%`, ymax = `95%`)) +
+  ylab(label) + xlab("")
+}
+```
+
 ```{r gg_depth_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
@@
 
+# gg_depth_shape_class <- dt %>%
+#   dplyr::filter(shape_class != "neither") %>%  
+#   contrast_plot(grp = "shape_class", varb = "lake_maxdepth_m", 
+#                 label = "Depth (m)", plot_type = "violin")
+
 gg_depth_shape_class <- dt %>%
   dplyr::filter(shape_class != "neither") %>%  
-  group_by(shape_class) %>%
-  summarize(lake_maxdepth_m = list(tibble::enframe(
-    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
-  unnest() %>%
-  tidyr::spread(name, value) %>%
-  ggplot() +
-geom_pointrange(aes(x = shape_class, y = `50%`, 
-                      ymin = `5%`, ymax = `95%`)) +
-  ylab("Depth (m)") + xlab("")
+  contrast_plot(grp = "shape_class", varb = "lake_maxdepth_m", 
+                label = "Depth (m)", plot_type = "pointrange")
 ```
@@
   dplyr::filter(shape_class != "neither") %>%  
-  group_by(shape_class) %>%
-  summarize(lake_waterarea_ha = list(tibble::enframe(
-    quantile(lake_waterarea_ha, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
-  unnest() %>%
-  tidyr::spread(name, value) %>%
-  ggplot() +
-geom_pointrange(aes(x = shape_class, y = `50%`, 
-                      ymin = `5%`, ymax = `95%`)) +
-  ylab("Area (ha)") + xlab("")
+  contrast_plot(grp = "shape_class", varb = "lake_waterarea_ha", 
+                label = "Area (ha)", plot_type = "pointrange")
 ```

### 2020-03-26 scripts/99_make_taxonomy.R
change buffer to nearshore

@@ -58,7 +58,7 @@ taxonomy_key <- bind_rows(
     pretty = "Mean slope (m/m)"),
   c(col_name = "buffer_slope",
     description = "average slope of the land surrounding a lake in a 100m buffer",
-    pretty = "Buffer slope (m/m)"),
+    pretty = "Nearshore slope (m/m)"),
   c(col_name = "dist_deepest",
     description = "distance from the lake shoreline to the deepest point of the lake in meters",
     pretty = "Deepest point distance (m)"),

### 2020-03-26 manuscript/figures.Rmd
rm hypenation

@@
 
-![Diagram showing the assumption of most lake depth models that the nearshore slope extends into the lake and that all lakes have a common relationship between nearshore and in-lake slope.
+![Diagram showing the assumption of most lake depth models that the nearshore slope extends into the lake and that all lakes have a common relationship between nearshore and inlake slope.
 ](../figures/slope_diagram.pdf){width=75%}
@@
 
+```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE}
+knitr::include_graphics("../figures/01_hypsography-1.pdf")
+```
+
 ```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=FALSE, out.width="11cm", fig.align='center', out.extra=''}
@@
 knitr::include_graphics("../figures/slope-compare-1.pdf")
-```
-
-```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE}
-knitr::include_graphics("../figures/01_hypsography-1.pdf")
 ```

### 2020-03-26 figures/01_contrasts.Rmd
fix multi tally situation with dplyr 1.0

@@
    dplyr::filter(!is.na(lake_maxdepth_m)) %>%
-   group_by(sourcetype_depth, sourceurl_depth) %>% 
-   tally() %>% 
+   distinct(sourcetype_depth, sourceurl_depth) %>% 
    ungroup() %>% group_by(sourcetype_depth) %>%

### 2020-03-26 figures/00_maps.Rmd
size alignment

@@
 
+us_states <-dplyr::filter(
+    dplyr::select(states_all, postal, iso_a2), iso_a2 == "US")
 dt_plot <- dt %>%
@@
   summarize(prop_maxdepth = mean(!is.na(lake_maxdepth_m))) %>%
-  arrange(desc(prop_maxdepth)) %>%
-  left_join(dplyr::filter(
-    dplyr::select(states_all, postal, iso_a2), iso_a2 == "US"), 
-            by = c("lake_state" = "postal")) %>%
+  arrange(desc(prop_maxdepth))
+dt_plot <- dt_plot %>%
+  left_join(us_states, by = c("lake_state" = "postal")) %>%
   dplyr::filter(!is.na(prop_maxdepth)) %>%
@@
 
-```{r 00_map_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6}
+```{r 00_map_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6.5}
 theme_opts <- guides(color = guide_legend(

### 2020-03-26 manuscript/figures.Rmd
more caption detail

@@
 
-```{r lake_shape, echo=FALSE, out.extra='', fig.align='center', fig.cap="Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right)"}
+```{r lake_shape, echo=FALSE, out.extra='', fig.align='center', fig.cap="Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
+Dashed lines represent extrapolated nearshore slope while solid lines represent the lake bottom."}
 knitr::include_graphics("../figures/lake_shape.pdf")
@@
 
-```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth calculated from bathymetry surfaces by US State.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth calculated from bathymetry surfaces by US State.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 # see 00_qa.Rmd

### 2020-03-26 scripts/00_get_nearshore.R
turn off incremental pulling, closes #9

@@ -19,8 +19,10 @@ flist             <- list.files("data/elevatr", pattern = "\\d*.tif",
                     full.names = TRUE, include.dirs = TRUE)
 existing_surfaces <- gsub(".tif", "",
                 stringr::str_extract(flist, "\\d*(!?.tif)"))
-ll_ids <- c(existing_surfaces)
-# ll_ids <- c(ll_ids, existing_surfaces)
+ll_ids_new <- ll_ids[!(ll_ids %in% existing_surfaces)]
+# ll_ids <- c(existing_surfaces)
+# ll_ids <- c(ll_ids_new)
+ll_ids <- c(ll_ids_new, existing_surfaces)
 ll_ids <- ll_ids[!duplicated(ll_ids)]
 # sapply(flist[!(ll_ids %in% dt_pred$lagoslakeid)], "unlink")
 

### 2020-03-26 figures/01_contrasts.Rmd
enable violin plot switching

@@
 
-```{r define-plots}
+```{r define-plots, echo=FALSE}
 contrast_plot <- function(..., plot_type){
+  theme_opts <- cowplot::theme_cowplot()
+
   if(plot_type == "violin"){
-    violin_plot(...)
+    violin_plot(...) + theme_opts
   }else{
-    pointrange_plot(...)
+    pointrange_plot(...) + theme_opts
   }
@@
   dplyr::filter(shape_class != "neither") %>%  
-  group_by(shape_class) %>%
-  summarize(inlake_slope = list(tibble::enframe(
-    quantile(inlake_slope, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
-  unnest() %>%
-  tidyr::spread(name, value) %>%
-  ggplot() +
-geom_pointrange(aes(x = shape_class, y = `50%`, 
-                      ymin = `5%`, ymax = `95%`)) +
-  ylab("Inlake slope (m/m)") + xlab("")
+  contrast_plot(grp = "shape_class", varb = "inlake_slope", 
+                label = "Inlake slope (m/m)", plot_type = "pointrange")
 ```
@@
   dplyr::filter(shape_class != "neither") %>%  
-  group_by(shape_class) %>%
-  summarize(dist_deepest = list(tibble::enframe(
-    quantile(dist_deepest, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
-  unnest() %>%
-  tidyr::spread(name, value) %>%
-  ggplot() +
-geom_pointrange(aes(x = shape_class, y = `50%`, 
-                      ymin = `5%`, ymax = `95%`)) +
-  ylab("Deepest point distance (m)") + xlab("")
+  contrast_plot(grp = "shape_class", varb = "dist_deepest", 
+                label = "Deepest point distance (m)", plot_type = "pointrange")
 ```
@@
 ```{r gg_depth_connectivity, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
-
 gg_depth_conny_class <- dt %>%
@@
 ```{r gg_depth_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
-
-(gg_depth_reservoir <- dt %>%
+gg_depth_reservoir <- dt %>%
   dplyr::filter(!is.na(reservoir_class)) %>%  
-  group_by(reservoir_class) %>%
-  summarize(lake_maxdepth_m = list(tibble::enframe(
-    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
-  unnest() %>%
-  tidyr::spread(name, value) %>%
-  ggplot() +
-geom_pointrange(aes(x = reservoir_class, y = `50%`, 
-                      ymin = `5%`, ymax = `95%`)) +
-  ylab("Depth (m)") + xlab(""))
+  contrast_plot(grp = "reservoir_class", varb = "lake_maxdepth_m", 
+                label = "Depth (m)", plot_type = "pointrange")
 ```
@@
 ```{r gg_area_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
-
-(gg_area_reservoir <- dt %>%
+gg_area_reservoir <- dt %>%
   dplyr::filter(!is.na(reservoir_class)) %>%  
-  group_by(reservoir_class) %>%
-  summarize(lake_waterarea_ha = list(tibble::enframe(
-    quantile(lake_waterarea_ha, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
-  unnest() %>%
-  tidyr::spread(name, value) %>%
-  ggplot() +
-geom_pointrange(aes(x = reservoir_class, y = `50%`, 
-                      ymin = `5%`, ymax = `95%`)) +
-  ylab("Area (ha)") + xlab(""))
+  contrast_plot(grp = "reservoir_class", varb = "lake_waterarea_ha", 
+                label = "Area (ha)", plot_type = "pointrange")
 ```
@@
 ```{r gg_slope_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
-
-(gg_slope_reservoir <- dt %>%
+gg_slope_reservoir <- dt %>%
   dplyr::filter(!is.na(reservoir_class)) %>%  
-  group_by(reservoir_class) %>%
-  summarize(inlake_slope = list(tibble::enframe(
-    quantile(inlake_slope, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
-  unnest() %>%
-  tidyr::spread(name, value) %>%
-  ggplot() +
-geom_pointrange(aes(x = reservoir_class, y = `50%`, 
-                      ymin = `5%`, ymax = `95%`)) +
-  ylab("Inlake slope (m/m)") + xlab(""))
+  contrast_plot(grp = "reservoir_class", varb = "inlake_slope", 
+                label = "Inlake slope (m/m)", plot_type = "pointrange")
 ```
@@
 ```{r gg_dist_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
-
-(gg_dist_reservoir <- dt %>%
+gg_dist_reservoir <- dt %>%
   dplyr::filter(!is.na(reservoir_class)) %>%  
-  group_by(reservoir_class) %>%
-  summarize(dist_deepest = list(tibble::enframe(
-    quantile(dist_deepest, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
-  unnest() %>%
-  tidyr::spread(name, value) %>%
-  ggplot() +
-geom_pointrange(aes(x = reservoir_class, y = `50%`, 
-                      ymin = `5%`, ymax = `95%`)) +
-  ylab("Deepest point distance (m)") + xlab(""))
+  contrast_plot(grp = "reservoir_class", varb = "dist_deepest", 
+                label = "Deepest point distance (m)", plot_type = "pointrange")
 ```
@@
 cowplot::plot_grid(gg_shape, gg_reservoir, 
-                   nrow = 2, rel_heights = c(1, 0.85))
+                   nrow = 2, rel_heights = c(1, 0.76))
 

### 2020-03-26 figures/01_geometry.Rmd
standardize slope names

@@
               method = "lm", se = FALSE) +
-  xlab("Buffer slope (mean)") +
-  ylab("In-lake slope") +
+  xlab("Nearshore slope (mean)") +
+  ylab("Inlake slope") +
   labs(color = "")
@@
   geom_smooth(method = "lm", se = FALSE) +
-  xlab("Buffer slope (mean)") +
-  ylab("In-lake slope") +
+  xlab("Nearshore slope (mean)") +
+  ylab("Inlake slope") +
   labs(color = "")

### 2020-03-30 manuscript/figures.Rmd
fix figure caption terms

@@
 
-```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../figures/00_map_bathy-1.pdf")
-```
-
 ```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth calculated from bathymetry surfaces by US State.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
@@
 \captionsetup[figure]{labelformat=default}
-\setcounter{figure}{4}
+\setcounter{figure}{3}
 
-```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) buffer slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
@@
 
-```{r depth_model_metrics, fig.cap = "Model fit and predictive accuracy metrics for all combinations of true (inlakeslope, distdeepest) and proxy (slopemean, distviscenter) metrics.
-Tiles are colored by their R2 values and labeled with their RMSE values.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+```{r depth_model_metrics, fig.cap = "Model fit and predictive accuracy metrics for all combinations of true (inlake slope, deepest point distance) and proxy (nearshore slope, visual center distance) metrics.
+Tiles are colored by their $R^2$ values and labeled with their RMSE values.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../figures/depth_model_metrics-1.pdf")
+```
+
+```{r 02_depth_model_importance, fig.cap = "Importance plot for random forest variables showing increase in mean square error.
+Higher values indicate greater importance to model predictions.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")
 ```
@@
 
-```{r 02_depth_model_importance, fig.cap = "Importance plot for random forest variables showing increase in mean square error.
-Higher values indicate greater importance to model predictions.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
-knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")
+```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/00_map_bathy-1.pdf")
 ```
@@
 
-```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE}
+```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 knitr::include_graphics("../figures/01_hypsography-1.pdf")

### 2020-04-03 figures/01_geometry.Rmd
add one to one line and better legend placement

--- 2020-02-21 figures/01_geometry.Rmd
@@
   dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::filter(dist_viscenter >= dist_deepest) %>%
+  dplyr::filter(dist_deepest < 3200) %>%
   dplyr::arrange(desc(factor(shape_class, 
@@
   ggplot(aes(x = dist_deepest, y = dist_viscenter, color = shape_class)) + 
-  geom_point() +
+  geom_point(size = 1) +
   # geom_abline(aes(intercept = 0, slope = 1)) +
@@
 
-gg_scatter_reservoir <- ggplot(data = dplyr::filter(
-  st_drop_geometry(dt), !is.na(reservoir_class))) + 
-  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = reservoir_class)) +
+gg_scatter_reservoir <- st_drop_geometry(dt) %>%
+  dplyr::filter(!is.na(reservoir_class)) %>%
+  dplyr::filter(dist_viscenter >= dist_deepest) %>%
+  ggplot() + 
+  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = reservoir_class), 
+             size = 1) +
+  scale_color_discrete(labels = c("NL     ", "Res")) + 
   # geom_abline(aes(intercept = 0, slope = 1)) +
@@
 
-gg_slope_shape <- ggplot(data = test, 
-                         aes(slope_mean, inlake_slope.y, 
-                             color = shape_class)) + 
-    geom_point() +
-  geom_smooth(data = dplyr::filter(test, shape_class != "neither"), 
+gg_slope_shape <- ggplot(data = test) + 
+  geom_point(aes(slope_mean, inlake_slope.y, 
+                             color = shape_class), size = 1) +
+  geom_smooth(aes(slope_mean, inlake_slope.y, 
+                             color = shape_class), data = dplyr::filter(test, shape_class != "neither"), 
               method = "lm", se = FALSE) +
-  xlab("Buffer slope (mean)") +
-  ylab("In-lake slope") +
+  geom_segment(aes(x = 0, xend = 0.35, y = 0, yend = 0.35), 
+               linetype = "dashed", size = 0.4) +
+  # geom_abline(slope = 1, intercept = 0) +
+  xlab("Nearshore slope (mean)") +
+  ylab("Inlake slope") +
   labs(color = "")
 
-gg_slope_reservoir <- ggplot(data = dplyr::filter(test, !is.na(reservoir_class)), 
-                             aes(slope_mean, inlake_slope.y, 
-                                 color = reservoir_class)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  xlab("Buffer slope (mean)") +
-  ylab("In-lake slope") +
+gg_slope_reservoir <- ggplot(data = dplyr::filter(test, !is.na(reservoir_class))) +
+  geom_point(aes(slope_mean, inlake_slope.y, 
+                                 color = reservoir_class), size = 1) +
+  geom_smooth(aes(slope_mean, inlake_slope.y, 
+                                 color = reservoir_class), 
+              method = "lm", se = FALSE) +
+  geom_segment(aes(x = 0, xend = 0.35, y = 0, yend = 0.35), 
+               linetype = "dashed", size = 0.4) +
+  xlab("Nearshore slope (mean)") +
+  ylab("Inlake slope") +
   labs(color = "")
@@
 
-```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5, fig.height=4.2}
-theme_opts <- theme(legend.position = "none")
+```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5.5, fig.height=4.2}
+theme_opts <- cowplot::theme_cowplot() + 
+  theme(legend.position = "none", 
+        axis.text = element_text(size = 10), 
+        axis.title = element_text(size = 11))
+gg_spacer <- ggplot() + geom_blank() + theme(panel.background = element_blank())
 
-legend_shape     <- cowplot::get_legend(gg_scatter_shape)
-legend_reservoir <- cowplot::get_legend(gg_scatter_reservoir + labs(color = ""))
-legend_panel     <- cowplot::plot_grid(legend_shape, legend_reservoir, 
-                                       ncol = 1, 
+legend_shape     <- cowplot::get_legend(gg_scatter_shape +
+                                          theme(legend.direction = "horizontal"))
+legend_reservoir <- cowplot::get_legend(gg_scatter_reservoir + labs(color = "") +
+                                          theme(legend.direction = "horizontal"))
+legend_panel     <- cowplot::plot_grid(gg_spacer, legend_shape, 
+                                       gg_spacer, legend_reservoir, 
+                                       nrow = 1, rel_widths = c(0.1, 1, 0.08, 1),
                                        align = "hv")
 
+cowplot::plot_grid(
   cowplot::plot_grid(
@@
                        theme_opts,
-                     rel_widths = c(1, 0.93, 1, 1)), 
-  legend_panel, rel_widths = c(1, 0.2)
+                     ncol = 3, rel_widths = c(0.08, 1, 1)), 
+                     rel_widths = c(1, 1), nrow = 2), 
+ legend_panel, rel_heights = c(1, 0.12), ncol = 1
 )

### 2020-04-03 manuscript/figures.Rmd
add one to one line and better legend placement

@@
 
-```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.
+Also shown is a 1:1 line for comparison.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../figures/01_geometry_grid-1.pdf")

### 2020-04-08 manuscript/appendix.Rmd
format appendix for WRR

--- 2020-03-30 manuscript/figures.Rmd
@@
 
+\def\figurename{Figure S\hspace{-0.3em}}
+\def\tablename{Table S\hspace{-0.3em}}
+
 \arrayrulecolor{white}
@@
 
-![Diagram showing the assumption of most lake depth models that the nearshore slope extends into the lake and that all lakes have a common relationship between nearshore and inlake slope.
-](../figures/slope_diagram.pdf){width=75%}
+\centering
+\includegraphics{../../../../JournalSubmissions/logos/agu_publications.png}
 
-```{r lake_shape, echo=FALSE, out.extra='', fig.align='center', fig.cap="Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
-Dashed lines represent extrapolated nearshore slope while solid lines represent the lake bottom."}
-knitr::include_graphics("../figures/lake_shape.pdf")
-```
+\emph{Water Resources Research}
 
-```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth calculated from bathymetry surfaces by US State.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-# see 00_qa.Rmd
-knitr::include_graphics("../figures/lgnemanual-vs-bathy-depth-1.pdf")
-```
+\vspace{1em}
 
-\captionsetup[figure]{name=Table}
-\captionsetup[figure]{labelformat=empty}
-```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth predictor variables for modelling efforts (and for lakes continent-wide).", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../tables/01_predictors.pdf")
-```
-\captionsetup[figure]{name=Figure}
-\captionsetup[figure]{labelformat=default}
-\setcounter{figure}{3}
+Supporting information for
 
-```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
-```
+\vspace{1em}
 
-```{r depth_model_metrics, fig.cap = "Model fit and predictive accuracy metrics for all combinations of true (inlake slope, deepest point distance) and proxy (nearshore slope, visual center distance) metrics.
-Tiles are colored by their $R^2$ values and labeled with their RMSE values.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../figures/depth_model_metrics-1.pdf")
-```
+\textbf{Bathymetry data from thousands of lakes show that lake depth prediction is confounded by difficulty modeling inlake slope}
 
-```{r 02_depth_model_importance, fig.cap = "Importance plot for random forest variables showing increase in mean square error.
-Higher values indicate greater importance to model predictions.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
-knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")
-```
+\vspace{1em}
 
-```{r depth_model_grid_resid, fig.cap = "Depth model residuals by shape and reservoir class.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
-knitr::include_graphics("../figures/02_depth_model_grid_resid-1.pdf")
-```
+J.Stachelek\textsuperscript{1},
+
+\vspace{1em}
+
+\textsuperscript{1}Department of Fisheries and Wildlife, Michigan State University, 480 Wilson Rd., East Lansing, Michigan 48824 USA
+
+\vspace{8em}
+
+\textbf{Contents of this file}
+
+\vspace{4em}
+
+\textbf{Introduction}
 
 \pagebreak
-# Appendix
 \setcounter{figure}{0}

### 2020-04-13 manuscript/appendix.Rmd
use GIS-derived intead of map-derived

@@
 
-```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=FALSE, out.width="11cm", fig.align='center', out.extra=''}
+```{r metric_correlations, fig.cap = "Pearson correlation coefficients of GIS-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=FALSE, out.width="11cm", fig.align='center', out.extra=''}
 knitr::include_graphics("../figures/metric_correlations-1.pdf")

### 2020-04-13 manuscript/appendix.Rmd
move figs to separate appendix file

@@
 
+How do reported values of lake depth compare to depth calculated from detailed bathymetry?
+
 \pagebreak
@@
 knitr::include_graphics("../figures/00_map_bathy-1.pdf")
+```
+
+```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth calculated from bathymetry surfaces by US State.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+# see 00_qa.Rmd
+knitr::include_graphics("../figures/lgnemanual-vs-bathy-depth-1.pdf")
 ```

### 2020-04-13 manuscript/figures.Rmd
move figs to separate appendix file

@@
 knitr::include_graphics("../figures/lake_shape.pdf")
-```
-
-```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth calculated from bathymetry surfaces by US State.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-# see 00_qa.Rmd
-knitr::include_graphics("../figures/lgnemanual-vs-bathy-depth-1.pdf")
 ```
@@
 ```
-
-\pagebreak
-# Appendix
-\setcounter{figure}{0}
-
-```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../figures/00_map_bathy-1.pdf")
-```
-
-```{r contrasts, fig.cap = "Lake characteristics by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
-knitr::include_graphics("../figures/01_contrasts-1.pdf")
-```
-
-```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
-knitr::include_graphics("../figures/01_hypsography-1.pdf")
-```
-
-```{r metric_correlations, fig.cap = "Pearson correlation coefficients of GIS-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=FALSE, out.width="11cm", fig.align='center', out.extra=''}
-knitr::include_graphics("../figures/metric_correlations-1.pdf")
-```
-
-```{r heatmap, fig.cap = "Correlation matrix heatmap.", warning=FALSE, echo = FALSE, eval=FALSE}
-knitr::include_graphics("../figures/slope-compare-1.pdf")
-```
-
-```{r glaciated_vs_shapeclass, fig.cap = "Cross tabulation between lake glaciation and shape class.
-Circles are proportional to the number of lakes in each category.
-If a lake has been glaciated it is more likely to be convex.
-However, if a lake has not been glaciated there is no tendencey towards a particular shape class.", warning=FALSE, echo = FALSE, eval=FALSE}
-knitr::include_graphics("../figures/glaciated_vs_shapeclass-1.pdf")
-```
-
-```{r contrasts_tally, fig.cap = "Number of lake depth measurements by categories.", warning=FALSE, echo = FALSE, eval=FALSE}
-knitr::include_graphics("../figures/01_contrasts_tally-1.pdf")
-```
-
-```{r 02_depth_model_fitted, fig.cap = "A model that does not include shape class over predicts depth in concave lakes but does not systematically under predict depth in convex lakes.", warning=FALSE, echo = FALSE, eval=FALSE}
-knitr::include_graphics("../figures/02_depth_model_fitted-1.pdf")
-```
-
-```{r 02_depth_oliver-2, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=FALSE}
-knitr::include_graphics("../figures/02_depth_oliver-2.pdf")
-```
-
-```{r 02_depth_oliver-1, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=FALSE}
-knitr::include_graphics("../figures/02_depth_oliver-1.pdf")
-```
-
-```{r 02_depth_oliver-4, fig.cap = "Oliver 2016 observed versus predicted.", warning=FALSE, echo = FALSE, eval=FALSE}
-knitr::include_graphics("../figures/02_depth_oliver-4.pdf")
-```
-
-```{r 02_hypso_model_fitted, fig.cap = "Confusion matrix comparing two classification methods", warning=FALSE, echo = FALSE, eval=FALSE}
-knitr::include_graphics("../figures/02_hypso_model_fitted-1.pdf")
-```
-
-```{r 02_hypso_model_importance, fig.cap = "Recusive Partitioning variable importance.", warning=FALSE, echo = FALSE, eval=FALSE, out.extra=''}
-knitr::include_graphics("../figures/02_hypso_model_importance-1.pdf")
-```
-
-<!-- ![Messager et al. 2016 equation table](../figures/messager_eq_table.png){width=100%} -->
-
-```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=FALSE}
-knitr::include_graphics("../figures/gg_distance-3.pdf")
-```

### 2020-04-13 figures/01_hypsography.Rmd
include class breakdown in hypso fig

@@
 
-ids <- factor(c("convex", "concave"))
+ids <- factor(c("convex", "concave"), levels = c("convex", "concave"))
 polys <- data.frame(
@@
 
-ggplot(data = dt_raw) +
-  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
-  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
+# generate labels with percent in each class
+dt_classes <- read.csv("../data/00_hypso/hypso_classes.csv", 
+                       stringsAsFactors = FALSE) %>%
+  left_join(dt_raw)
+class_tally <- dt_classes %>%
+  dplyr::filter(shape_class != "neither") %>%
+  group_by (state, shape_class) %>%
+  summarise (n=n()) %>%
+  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%")) %>%
+  dplyr::select(-n) %>%
+  tidyr::pivot_wider(names_from = state, values_from = rel.freq) %>%
+  t()
+class_tally <- setNames(
+  tibble::as_tibble(class_tally[-1,]), c("concave", "convex")) %>%
+  mutate(state = row.names(class_tally)[-1]) %>%
+  mutate(state_label = paste0(state, " (", convex, ",", concave, ")"))
+dt_raw <- left_join(dt_raw, class_tally, by = "state")
+
+res <- ggplot(data = dt_raw) +
+  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.5) +
+  scale_fill_manual(values = c(concave = "#F8766D", convex = "#00BA38")) +
+  geom_line(aes(x = area_percent, y = depth_percent, group = llid), size = 0.2) +
   geom_abline(intercept = -100, slope = 1, color = "red") +
@@
         legend.title = element_blank()) +
-  facet_wrap(~state)
+  facet_wrap(~state_label)
+
+leg <- cowplot::get_legend(res + theme(legend.direction = "horizontal"))
+
+cowplot::plot_grid(res + theme(legend.position = "none"), 
+                   leg, 
+                   nrow = 2, rel_heights = c(1, 0.09))
 

### 2020-04-14 manuscript/appendix.Rmd
formatting, stash text

--- 2020-04-13 manuscript/appendix.Rmd
@@
 
-J.Stachelek\textsuperscript{1},
+J. Stachelek\textsuperscript{1}, P. Hanly\textsuperscript{1}, and P.A. Soranno\textsuperscript{1}
 
@@
 
+\raggedright
 \textbf{Contents of this file}
 
-\vspace{4em}
+\textbf{Figure S1} Map of study lakes
 
-\textbf{Introduction}
+\textbf{Figure S2} Comparison between reported depth and depth estimated from bathymetry surfaces
+
+\textbf{Figure S3} Lake characteristics by categorical variables
+
+\textbf{Figure S4} Hypsography classification by state
+
+<!-- \vspace{4em} -->
+
+<!-- \textbf{Introduction} -->
+
+<!-- How do reported values of lake depth compare to depth calculated from detailed bathymetry? -->
+
+<!-- we estimated calculated the maximum depth of approximately 5,000 lakes in the Northeastern and Midwestern US from bathymetry data and compared the results against a database of lake depths developed by (LAGOSUS-Depth citation).
+-->
+
+<!-- One of our objectives was to see whether uncertainty or bias in reported lake depths is likely responsible for the poor performance of prior depth prediction studies.  -->
+
+<!-- We refer to these values as “calculated depth” to distinguish them from “reported depth” values one might find listed in a report or other document.  -->
+
+<!-- To evaluate the accuracy of reported lake depth values, we compared depths calculated from bathymetry surfaces against reported depths in the LAGOSUS-Depth product (citation).
+No reported depth values originated from the same source as its corresponding calculated value.
+-->
+
+<!-- How do reported values of lake depth compare to depth calculated from detailed bathymetry? -->
+<!-- Given the importance of depth-dependent processes in determining lake function, it is encouraging that reported maximum depth values are closely aligned with “true” depth values calculated from bathymetry.
+This suggests that uncertainty or bias in reported lake depths is not likely responsible for the poor performance of prior depth prediction studies.
+-->
+
+<!-- We propose that the median spread among reported and calculated values of maximum be used as a benchmark for future studies.
+Future studies should not expect to have prediction errors smaller than our observed spread (0.3 m) among reported and calculated maximum depth values.
+-->
+
+<!-- Although it might seem that the mismatch and relatively poor correspondence between reported and calculated values of mean depth calls into question the basis of prior studies depending on the mean depth metric, there are several possible explanations for our finding.
+-->
+
+<!-- First, the mismatch between reported and calculated mean depth may be a result of using different definitions for lake boundaries (or a result of variation in lake water levels).
+In our study, we defined a lake boundary based on the waterbody polygons in the NHD but on-the-ground surveys could very well be using different boundaries.
+This could lead to discrepancies in mean depth values because calculated values would be based on different lake basin extents or entirely different numbers of basins.
+-->
+
+<!-- An alternative explanation is that reported values of mean depth are not based on the same definition of mean depth as we have used in our calculations.
+Whereas we have calculated mean depth mathematically, reported mean depth values may be based on expert judgement or simply a qualitative estimate of the general depth one might experience out on a lake.
+-->
+
+<!-- Finally, our calculated mean depth values may be sensitive to the resolution and type of the source bathymetry data.
+For example, in the case of the Maine data the source bathymetry data was point measurements made along depth sounding transects rather than depth contours, polygons, or interpolated raster products.
+As a result, the density of data used for smoothing interpolations in Maine lakes was much lower than in other states.
+This could lead to overestimation of mean depth due to limited sampling in shallow areas as they are relatively inaccessible by boat (Furnans and Austin 2008).
+-->
 
@@
 knitr::include_graphics("../figures/00_map_bathy-1.pdf")
+```
+
+```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth estimated from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product (citation).
+For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+# see 00_qa.Rmd
+knitr::include_graphics("../figures/lgnemanual-vs-bathy-depth-1.pdf")
 ```

### 2020-04-14 figures/01_geometry.Rmd
show r2 on scatter

@@
                        # theme(axis.title.x = element_blank()) +
+                       annotate("text", x = 2000, y = 300, label = "R^2 == 0.80",
+                                parse = TRUE) +
                        theme_opts,
@@
                        labs(color = "") +
+                       # annotate("text", x = 1800, y = 500, label = "R^2 == 0.80",
+                       #          parse = TRUE) +
                        theme(#axis.title.x = element_blank(), 
@@
                      gg_slope_shape +
+                       annotate("text", x = 0.33, y = 0.03, label = "R^2 == 0.17",
+                                parse = TRUE) +
                        theme_opts,
                      gg_slope_reservoir + 
+                         # annotate("text", x = 0.3, y = 0.03, label = "R^2 == 0.17",
+                         #        parse = TRUE) +
                        theme(axis.title.y = element_blank()) +

### 2020-04-14 manuscript/appendix.Rmd
stash text

@@
 <!-- How do reported values of lake depth compare to depth calculated from detailed bathymetry? -->
+
+<!-- Reported maximum depth values were very similar to calculated maximum depth values as evidenced by their alignment along a 1:1 line (Figure 3).
+Conversely, the alignment between reported and calculated mean depth values was more messy.
+In particular, the calculated mean depth of Maine lakes appeared to be consistently underestimated compared to reported values.
+-->
+
 <!-- Given the importance of depth-dependent processes in determining lake function, it is encouraging that reported maximum depth values are closely aligned with “true” depth values calculated from bathymetry.

### 2020-04-14 manuscript/figures.Rmd
fix figure numbering
show proxy measures on diagram

@@
 
-![Diagram showing the assumption of most lake depth models that the nearshore slope extends into the lake and that all lakes have a common relationship between nearshore and inlake slope.
-](../figures/slope_diagram.pdf){width=75%}
+![Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
+Calculated depth from Equation 1 requires a distance and a slope metric.](../figures/slope_diagram_new.pdf){width=75%}
 
@@
 \captionsetup[figure]{labelformat=default}
-\setcounter{figure}{3}
+\setcounter{figure}{2}
 

### 2020-04-14 scripts/06_stats.R
stash stats script

@@ -1,0 +1,43 @@
+
+# What percentage of lakes are in each shape class?
+dt <- read.csv("data/00_hypso/hypso_classes.csv", stringsAsFactors = FALSE)
+table(dt$shape_class)[1] / sum(table(dt$shape_class)) * 100
+table(dt$shape_class)[2] / sum(table(dt$shape_class)) * 100
+
+# What percentage of lakes are in each reservoir class?
+dt <- read.csv("data/00_reservoir_classification/reservoir_classes_clean.csv",
+               stringsAsFactors = FALSE)
+table(dt$reservoir_class)[1] / sum(table(dt$reservoir_class))
+table(dt$reservoir_class)[2] / sum(table(dt$reservoir_class))
+
+# What are the summary stats of the papers looked at by Hakanson (1977)?
+hak_areas <- as.numeric(c("821000", "598000", "578000", "25700", "19000", "5650",
+                          "3580", "2070", "1910", "1140", "484", "128",
+                          "94.7", "92.0", "63.8", "54.2", "23.8", "18.6",
+                          "16.0", "14.9", "10.0", "7.10", "5.15", "4.96",
+                          "3.88", "3.28", "2.80", "2.66", "2.00", "1.90",
+                          "1.80", "1.60", "1.59", "1.47", "1.13", "1.1",
+                          "1.02", "0.91", "0.87", "0.84", "0.57", "0.56", "
+             0.36", "0.32", "0.31", "0.17", "0.097", "0.029")) *
+  100 # convert from km2 to ha
+quantile(hak_areas,
+         probs = c(0.25, 0.5, 0.75))
+hist(hak_areas, n = 80)
+
+# How related are true and proxy geometry metrics?
+bp <- readRDS("data/00_bathy_depth/bathy_pnts.rds")
+dt <- read.csv("data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
+  mutate(llid = as.character(llid))
+dt <- left_join(dt, st_drop_geometry(bp),
+                by = "llid") %>%
+  dplyr::filter(!is.na(inlake_slope_mean))
+lgus_pred <- read.csv("data/lagosus_depth_predictors.csv",
+                      stringsAsFactors = FALSE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  dplyr::select(-contains("dist"))
[…]

### 2020-04-15 scripts/99_make_taxonomy.R
change derived depth to geometric depth

@@ -89,7 +89,7 @@ taxonomy_key <- bind_rows(
     pretty = "HUC4 ID"),
   c(col_name = "trig_depth",
     description = "Trigonometrically derived max depth (m)",
-    pretty = "Derived max depth (m)")
+    pretty = "Geometric max depth (m)")
   )
 
 write.csv(taxonomy_key, "data/taxonomy.csv", row.names = FALSE)

### 2020-04-15 manuscript/figures.Rmd
point to defn of geometric depth

@@
 ```{r 02_depth_model_importance, fig.cap = "Importance plot for random forest variables showing increase in mean square error.
-Higher values indicate greater importance to model predictions.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+Higher values indicate greater importance to model predictions.
+See Equation 1 for a definition of geometric max depth.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")

### 2020-04-15 scripts/06_stats.R
calculate percent lakes with bathy maps

@@ -1,16 +1,17 @@
+source("scripts/99_utils.R")
 
-# What percentage of lakes are in each shape class?
+# ---- What percentage of lakes are in each shape class? ----
 dt <- read.csv("data/00_hypso/hypso_classes.csv", stringsAsFactors = FALSE)
 table(dt$shape_class)[1] / sum(table(dt$shape_class)) * 100
 table(dt$shape_class)[2] / sum(table(dt$shape_class)) * 100
 
-# What percentage of lakes are in each reservoir class?
+# ---- What percentage of lakes are in each reservoir class? ----
 dt <- read.csv("data/00_reservoir_classification/reservoir_classes_clean.csv",
                stringsAsFactors = FALSE)
 table(dt$reservoir_class)[1] / sum(table(dt$reservoir_class))
 table(dt$reservoir_class)[2] / sum(table(dt$reservoir_class))
 
-# What are the summary stats of the papers looked at by Hakanson (1977)?
+# ---- What are the summary stats of the papers looked at by Hakanson (1977)? ----
 hak_areas <- as.numeric(c("821000", "598000", "578000", "25700", "19000", "5650",
                           "3580", "2070", "1910", "1140", "484", "128",
                           "94.7", "92.0", "63.8", "54.2", "23.8", "18.6",
@@ -24,7 +25,7 @@ quantile(hak_areas,
          probs = c(0.25, 0.5, 0.75))
 hist(hak_areas, n = 80)
 
-# How related are true and proxy geometry metrics?
+# ---- How related are true and proxy geometry metrics? ----
 bp <- readRDS("data/00_bathy_depth/bathy_pnts.rds")
 dt <- read.csv("data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
   mutate(llid = as.character(llid))
@@ -41,3 +42,22 @@ dt <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>%
 
 broom::glance(lm(dist_viscenter~dist_deepest, data = dt))$r.squared # 0.80
 broom::glance(lm(inlake_slope ~ slope_mean, data = dt))$r.squared # 0.17
+
+# ---- the percentage of lakes in the study footprint with bathy maps ----
+
+# pull bathy llids + states
+bd <- read.csv("data/00_bathy_depth/00_bathy_depth.csv",
[…]

### 2020-04-15 figures/01_geometry.Rmd
avoid cutoff axis label

--- 2020-04-03 figures/01_geometry.Rmd
@@
   xlab("Deepest point distance (m)") +
-  ylab("Visual center distance (m)")
+  ylab("Visual center\ndistance (m)")
 
@@
   xlab("Deepest point distance (m)") +
-  ylab("Visual center distance (m)")
+  ylab("Visual center\ndistance (m)")
 ```
@@
                                        gg_spacer, legend_reservoir, 
-                                       nrow = 1, rel_widths = c(0.1, 1, 0.08, 1),
+                                       nrow = 1, rel_widths = c(0.2, 1, 0.08, 1),
                                        align = "hv")
 
+text_size <- 2.5
 cowplot::plot_grid(
@@
                        # theme(axis.title.x = element_blank()) +
+                       annotate("text", x = 2250, y = 300, label = "R^2 == 0.80",
+                                parse = TRUE, size = text_size) +
                        theme_opts,
+                     gg_spacer,
                      gg_scatter_reservoir + 
                        labs(color = "") +
+                       # annotate("text", x = 1800, y = 500, label = "R^2 == 0.80",
+                       #          parse = TRUE) +
                        theme(#axis.title.x = element_blank(), 
@@
                        theme_opts, 
-                     ncol = 2),
+                     ncol = 3, rel_widths = c(1, 0, 1)),
   cowplot::plot_grid(ggplot() + geom_blank() + 
@@
                      gg_slope_shape +
+                       annotate("text", x = 0.36, y = 0.03, label = "R^2 == 0.17",
+                                parse = TRUE, size = text_size) +
                        theme_opts,
+                     gg_spacer,
                      gg_slope_reservoir + 
+                         # annotate("text", x = 0.3, y = 0.03, label = "R^2 == 0.17",
+                         #        parse = TRUE) +
                        theme(axis.title.y = element_blank()) +
                        theme_opts,
-                     ncol = 3, rel_widths = c(0.08, 1, 1)), 
-                     rel_widths = c(1, 1), nrow = 2), 
+                     ncol = 4, rel_widths = c(0.11, 1, 0.11, 1)), 
+                     rel_widths = c(0.7, 1), nrow = 2), 
  legend_panel, rel_heights = c(1, 0.12), ncol = 1

### 2020-04-15 figures/02_depth_model.Rmd
rm dashes, avoid axis text overlap

@@
   mutate(rmse = case_when(
-    as.character(model) == "maxdepth_true_true" ~ "-",
+    as.character(model) == "maxdepth_true_true" ~ " ",
     TRUE ~ rmse)) %>%
@@
   cowplot::theme_cowplot() +
-  theme(axis.line = element_blank())
+  theme(axis.line = element_blank(), 
+        axis.ticks = element_blank())
 ```
@@
         strip.background = element_blank(), 
+        axis.text = element_text(size = 9.5),
         strip.text = element_blank(),
@@
         strip.text = element_blank(),
+        axis.text = element_text(size = 9.5),
         plot.margin = unit(c(1.2,0.2,0.2,0.2), "cm"))  +

### 2020-04-15 manuscript/appendix.Rmd
explain percentages

@@
 
-```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r hypsography, fig.cap = "Hypsography classification by state.
+Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 knitr::include_graphics("../figures/01_hypsography-1.pdf")

### 2020-04-15 manuscript/figures.Rmd
explain panel discrepancies

@@
 ```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.
-Also shown is a 1:1 line for comparison.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+A dashed 1:1 line is shown for comparison.
+Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
@@
 Tiles are colored by their $R^2$ values and labeled with their RMSE values.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+# see 02_depth_model.Rmd
 knitr::include_graphics("../figures/depth_model_metrics-1.pdf")
@@
 
-```{r depth_model_grid_resid, fig.cap = "Depth model residuals by shape and reservoir class.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r depth_model_grid_resid, fig.cap = "Depth model residuals by cross-section shape and reservoir class.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+# see 02_depth_model.Rmd
 knitr::include_graphics("../figures/02_depth_model_grid_resid-1.pdf")

### 2020-04-29 figures/02_depth_model.Rmd
specify units of meters

@@
     as.character(model) == "maxdepth_true_true" ~ " ",
-    TRUE ~ rmse)) %>%
+    TRUE ~ paste0(rmse, " m"))) %>%
   mutate(model = gsub("maxdepth_", "", model)) %>%
@@
   distinct(lagoslakeid, .keep_all = TRUE)
-dt   <- readRDS("../data/01_depth_model/depth_grid.rds") %>%
+dt   <- readRDS("../data/01_depth_model/depth_grid.rds") %>% # 05_depth_model.R
   mutate(lagoslakeid = as.character(lagoslakeid)) %>%

### 2020-04-29 manuscript/figures.Rmd
specify units of meters

@@
 
-```{r depth_model_grid_resid, fig.cap = "Depth model residuals by cross-section shape and reservoir class.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r depth_model_grid_resid, fig.cap = "Depth model residuals in meters by cross-section shape and reservoir class.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 # see 02_depth_model.Rmd

### 2020-04-29 README.Rmd
title grammar
point to combined figs and appendix

@@
 
-## Bathymetry data from thousands of lakes show that difficulty modeling in-lake slope confounds lake depth prediction
+## Bathymetry data from thousands of lakes show that lake depth prediction is confounded by difficulty modeling inlake slope
 
@@
 
-Figures: [manuscript/figures.pdf](manuscript/figures.pdf)
+Figures: [manuscript/figures.pdf](manuscript/combined.pdf)
 

### 2020-04-29 manuscript/figures.Rmd
cleanup predictor table

@@
 \captionsetup[figure]{labelformat=empty}
-```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth predictor variables for modelling efforts (and for lakes continent-wide).", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth and predictor variables for modelling efforts (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
+Predictor variables used in Eq 2 are printed in bold face.
+Dashes (-) indicate identical sample size among this study and that of the contiguous United States.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../tables/01_predictors.pdf")

### 2020-04-29 tables/01_predictors.Rmd
cleanup predictor table

@@
   dplyr::filter(!str_detect(variable,
-                      "totalarea|islandperimeter|between|buffer100m")) # %>%
+                      "totalarea|islandperimeter|between|buffer100m|lake_meandepth_m")) # %>%
   # mutate(n = dt_bathy$n[which(dt_bathy$variable == "inlake_slope")])
@@
 
+dt <- dplyr::rename(dt, Variable = variable)
+
+# bold Eq 2 predictor variables
+dt$Variable[!stringr::str_detect(dt$Variable, "slope|distance|depth")] <- 
+  paste0("**", 
+         dt$Variable[!stringr::str_detect(dt$Variable, "slope|distance|depth")],
+         "**")
+
 options(knitr.kable.NA = '-')

### 2020-04-29 tables/01_predictors.Rmd
rm continental comparison for nearshore slope

@@
 dt[
-  dt$variable %in% c("dist_deepest", "dist_viscenter", "inlake_slope"), 2:5] <- 
+  dt$variable %in% c("dist_deepest", "dist_viscenter", 
+                     "inlake_slope", "buffer_slope"), 2:5] <- 
   dt[
-  dt$variable %in% c("dist_deepest", "dist_viscenter", "inlake_slope"), 2:5] %>%
+  dt$variable %in% c("dist_deepest", "dist_viscenter", 
+                     "inlake_slope", "buffer_slope"), 2:5] %>%
 apply(2, function(x) stringr::str_replace(x, "\\(([^\\)]+)\\)", "(-)"))

### 2020-05-01 manuscript/figures.Rmd
eq 1 produces geometric depth not calculated depth
describe hu4 in figure caption

@@
 ![Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
-Calculated depth from Equation 1 requires a distance and a slope metric.](../figures/slope_diagram_new.pdf){width=75%}
+Geometric depth calculated via Equation 1 requires a single distance and slope metric.](../figures/slope_diagram_new.pdf){width=75%}
 
@@
 Higher values indicate greater importance to model predictions.
-See Equation 1 for a definition of geometric max depth.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+See Equation 1 for a definition of geometric max depth.
+HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")

### 2020-05-04 figures/01_contrasts.Rmd
look at shoreline dev by shape class

@@
 
+```{r gg_shorelinedev_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+# compare shoreline dev for each shapeclass
+gg_shorelinedev_shape_class <- dt %>%
+  dplyr::filter(shape_class != "neither") %>%  
+  contrast_plot(grp = "shape_class", varb = "lake_shorelinedevfactor_nounits", 
+                label = "Shoreline Dev.", plot_type = "pointrange")
+```
+
 ```{r 01_contrasts, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.5}

### 2020-05-04 README.Rmd
point to combined figs and appendix
formatting

--- 2020-03-26 README.Rmd
@@
 
-## Bathymetry data from thousands of lakes show that difficulty modeling in-lake slope confounds lake depth prediction
+[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Docker Build](https://img.shields.io/badge/Docker%20Image-jsta/bathymetry-green.svg)](https://cloud.docker.com/repository/docker/jsta/bathymetry)
+
+Code and data
+for:
+
+**Stachelek et al. In prep**. Bathymetry data from thousands of lakes show that lake depth prediction is confounded by difficulty modeling inlake slope.
 
@@
 
-Figures: [manuscript/figures.pdf](manuscript/figures.pdf)
+Figures: [manuscript/combined.pdf](manuscript/combined.pdf)
 

### 2020-05-11 figures/01_geometry.Rmd
list more package dependencies, remove jsta package dependence

@@
 
-# jsta::jheatmap(dt)
-# jsta::jheatmap(dt, "inlake_slope.y")
-jsta::jheatmap(
+# jsta::jheatmap
+jheatmap(
   dplyr::select(dt, 

### 2020-05-11 scripts/99_utils.R

@@ -15,17 +15,25 @@ suppressMessages(library(cowplot))
 library(cutr) # devtools::install_github("moodymudskipper/cutr")
 library(broom)
 suppressMessages(library(raster))
-library(LAGOSNEgis)
+library(LAGOSNEgis) # install_github("cont-limno/LAGOSNEgis")
 suppressMessages(library(scales))
 library(elevatr)
 suppressMessages(library(dplyr))
 library(progress)
-suppressMessages(library(LAGOSUS))
+suppressMessages(library(LAGOSUS)) # install_github("cont-limno/LAGOSUS")
 library(polylabelr)
 library(esri2sf) # install_github("yonghah/esri2sf")
 library(rnaturalearth)
 library(mapview)
-library(ggrastr)
+library(ggrastr) # install_github('VPetukhov/ggrastr')
+library(lwgeom)
+library(rmapshaper)
+library(kableExtra)
+library(LAGOSUSgis) # install_github("cont-limno/LAGOSUSgis")
+library(rgdal)
+library(rnaturalearthhires) # install.packages("rnaturalearthhires", repos = "http://packages.ropensci.org", type = "source")
+library(GGally)
+library(ranger)
 
 # ---- misc fxn ----
 # jsta::get_if_not_exists
@@ -279,3 +287,49 @@ is_url <- function(x){
 # area_ratio <- st_area(
 #   lwgeom::st_minimum_bounding_circle(concaveman::concaveman(dt))) /
 #   st_area(concaveman::concaveman(dt))
+
+# jsta::jheatmap
+jheatmap <- function(dt, focal_columns = NULL) {
+  test <- data.frame(dt) %>% dplyr::select_if(is.numeric) %>%
+    corrr::correlate(quiet = TRUE) %>% data.frame()
+  test_rowname <- test$rowname
+  test <- dplyr::select(test, -"rowname")
+  row.names(test) <- test_rowname
+  if (length(focal_columns) > 0) {
+    test <- dplyr::select(test, focal_columns)
+  }
+  hmap.palette_red <- colorRampPalette(RColorBrewer::brewer.pal(n = 7,
+                                                                name = "Reds"))
+  hmap.palette_blue <- colorRampPalette(RColorBrewer::brewer.pal(n = 7,
+                                                                 name = "Blues"))
+  frac_color <- ceiling(range(as.numeric(unlist(test)), na.rm = TRUE) *
[…]

### 2020-05-11 scripts/00_get_gis.R
st_make_valid lives in the sf package now

@@ -43,7 +43,7 @@ hu4s_focal <- LAGOSUSgis::query_gis("hu4", "hu4_zoneid", hu4_focal$hu4_zoneid)
 study_bbox   <- st_as_sfc(st_bbox(states_all))
 
 hu4s_focal_simple <- rmapshaper::ms_simplify(
-  st_crop(lwgeom::st_make_valid(hu4s_focal),
+  st_crop(sf::st_make_valid(hu4s_focal),
           st_transform(study_bbox, st_crs(hu4s_focal)))
 )
 

### 2020-05-11 scripts/00_get_hypso_mi.R

@@ -11,7 +11,7 @@ download.file("https://opendata.arcgis.com/datasets/d49160d2e5af4123b15d48c2e9c7
 
 lg_poly <- LAGOSUSgis::query_gis_(query =
                                     "SELECT * FROM LAGOS_US_All_Lakes_1ha WHERE lake_centroidstate LIKE 'MI' AND lake_totalarea_ha > 4")
-lg_poly <- lwgeom::st_make_valid(lg_poly)
+lg_poly <- sf::st_make_valid(lg_poly)
 
 # remove problematic llids
 bad_llids <- c(107971, 4101, 1141, 2723)

### 2020-05-11 scripts/00_get_nearshore.R

@@ -30,7 +30,7 @@ get_slope <- function(ll_id){
   # ll_id <- ll_ids[1]
   # ll_id <- 7922
   ll_poly <- LAGOSUSgis::query_gis("LAGOS_US_All_Lakes_1ha", "lagoslakeid", ll_id)
-  ll_poly <- lwgeom::st_make_valid(ll_poly)
+  ll_poly <- sf::st_make_valid(ll_poly)
   ll_iws  <- LAGOSUSgis::query_gis("ws", "lagoslakeid", ll_id)
   # hack together a one-sided buffer [sf doesn't offer this :(]
   ll_buff <- st_buffer(ll_poly, max_buffer_dist)

### 2020-05-11 scripts/00_get_lagosus.R
use all depths lagosus product

@@ -1,6 +1,9 @@
 source("scripts/99_utils.R")
 
-lg <- LAGOSUS::lagosus_load("depth")
-lg <- lg$depth$depth
+# lg <- LAGOSUS::lagosus_load("depth")
+# lg <- lg$depth$depth
+
+# use all depths product
+lg <- read.csv("../lagos_depth/data/lagosus_depth_all.csv", stringsAsFactors = FALSE)
 
 write.csv(lg, "data/lagosus_depth.csv", row.names = FALSE)

### 2020-05-11 figures/00_maps.Rmd
use correct study bbox layer

@@
   ggplot() + 
-    geom_sf(data = study_bbox[2,], alpha = 1, size = 0.6, 
+    geom_sf(data = study_bbox[1,], alpha = 1, size = 0.6, 
           color = "black", fill = "#1D9DCF") +

### 2020-05-12 figures/01_contrasts.Rmd
turn off programtype graphs

@@
 
-```{r gg_tally_programtype, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r gg_tally_programtype, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
 # tally number of lakes with max depth by programtype_depth
@@
 ```{r 01_contrasts_tally, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 7.6}
-cowplot::plot_grid(gg_tally_programtype, gg_tally_maxdepthclass, 
+cowplot::plot_grid(gg_tally_maxdepthclass, 
                    gg_tally_waterareaclass, gg_tally_hypsoclass, 
@@
 
-```{r 01_programs_tally, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r 01_programs_tally, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
 # number of programs in each type

### 2020-05-15 figures/01_geometry.Rmd
drop class comparison from proxy versus true fig in main text

@@
 
+```{r 01_geometry_base, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
+theme_opts <- cowplot::theme_cowplot() + 
+  theme(legend.position = "none", 
+        axis.text = element_text(size = 10), 
+        axis.title = element_text(size = 11))
+text_size <- 3
+
+# distance data
+dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
+hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
+                       stringsAsFactors = FALSE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid))
+dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, shape_class), 
+                                  by = c("llid" = "lagoslakeid"))
+
+gg_dist_base <- st_drop_geometry(dt) %>%
+  dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::filter(dist_viscenter >= dist_deepest) %>%
+  dplyr::filter(dist_deepest < 3200) %>%
+  ggplot(aes(x = dist_deepest, y = dist_viscenter)) + 
+  geom_point(size = 1) +
+  geom_smooth(method = "lm", se = FALSE) +
+  annotate("text", x = 2550, y = 300, label = "R^2 == 0.80",
+                                parse = TRUE, size = text_size) +
+  labs(color = "") +
+  xlab("Deepest point distance (m)") +
+  ylab("Visual center\ndistance (m)") +
+  xlim(0, 3000) +
+  theme_opts
+
+# slope data
+bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
+dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
+  mutate(llid = as.character(llid))
+dt <- left_join(dt, st_drop_geometry(bp), 
+                by = "llid") %>%
+  dplyr::filter(!is.na(inlake_slope_mean)) %>%
+  dplyr::filter(inlake_slope.y < 0.4)
+lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid))
+test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
+  dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::arrange(desc(factor(shape_class, 
+                        levels = c("neither", "concave", "convex"))))
+
+gg_slope_base <- ggplot(data = test) + 
+  geom_point(aes(slope_mean, inlake_slope.y), size = 1) +
+  geom_smooth(aes(slope_mean, inlake_slope.y), method = "lm", se = FALSE) +
+  geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
+  alpha = 1, fill = "white") +
+  annotate("text", x = 0.34, y = 0.03, label = "R^2 == 0.17",
+                                parse = TRUE, size = text_size) +
+  # geom_abline(slope = 1, intercept = 0) +
+  xlab("Nearshore slope (mean)") +
+  ylab("Inlake slope") +
+  labs(color = "") +
+  theme_opts
+
+cowplot::plot_grid(
+  gg_dist_base, 
+  gg_slope_base, 
+  rel_widths = c(1, 0.9)
+)
+
+```
+
 ```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5.5, fig.height=4.2}

### 2020-05-15 manuscript/appendix.Rmd
drop class comparison from proxy versus true fig in main text

@@
 
+```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.
+A dashed 1:1 line is shown for comparison.
+Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
+```
+
 ```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}

### 2020-05-15 manuscript/figures.Rmd
drop class comparison from proxy versus true fig in main text

@@
 
-```{r lake_shape, echo=FALSE, out.extra='', fig.align='center', fig.cap="Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
+```{r lake_shape, echo=FALSE, out.extra='', fig.align='center', out.width="4in",fig.cap="Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
 Dashed lines represent extrapolated nearshore slope while solid lines represent the lake bottom."}
@@
 
-```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.
-A dashed 1:1 line is shown for comparison.
-Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
+```{r 01_geometry_base, fig.cap = "Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance to lake visual center and B) nearshore slope versus inlake slope.
+A best-fit line and coefficient of determination is shown to illustrate representativeness.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/01_geometry_base-1.pdf")
 ```

### 2020-05-15 manuscript/figures.Rmd
replace model metrics fig with a table

@@
 
-```{r depth_model_metrics, fig.cap = "Model fit and predictive accuracy metrics for all combinations of true (inlake slope, deepest point distance) and proxy (nearshore slope, visual center distance) metrics.
-Tiles are colored by their $R^2$ values and labeled with their RMSE values.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-# see 02_depth_model.Rmd
-knitr::include_graphics("../figures/depth_model_metrics-1.pdf")
-```
-
-```{r 02_depth_model_importance, fig.cap = "Importance plot for random forest variables showing increase in mean square error.
-Higher values indicate greater importance to model predictions.
-See Equation 1 for a definition of geometric max depth.
-HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
-knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")
+```{r depth_model_metrics, fig.cap = "Model fit and predictive accuracy metrics for all combinations of true (inlake slope, deepest point distance) and proxy (nearshore slope, visual center distance) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+# see tables/02_model_metrics.Rmd
+knitr::include_graphics("../tables/02_model_metrics.pdf")
 ```

### 2020-05-15 tables/02_model_metrics.Rmd
replace model metrics fig with a table

---
output:
  pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
# setwd("figures")
```

```{r 02_model_metrics, eval=TRUE}
library(kableExtra)
suppressMessages(library(dplyr))
library(magrittr)
library(stringr)

dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
  mutate_if(is.numeric, function(x) as.character(round(x, 2))) %>%
  mutate(rsq = as.numeric(rsq)) %>%
  mutate(rsq = case_when(
    as.character(model) == "maxdepth_true_true" ~ NA_real_,
    TRUE ~ rsq)) %>%
  mutate(rmse = case_when(
    as.character(model) == "maxdepth_true_true" ~ NA_character_,
    TRUE ~ paste0(rmse, " m"))) %>%
  mutate(model = gsub("maxdepth_", "", model)) %>%
  mutate(model = gsub("false", "proxy", model)) %>%
  tidyr::separate(model, c("slope", "distance")) %>%
  dplyr::select(slope, distance, everything(), -.estimator) %>%
  dplyr::select(-mae)

options(knitr.kable.NA = '-')
knitr::kable(dt_metrics, digits = 2, format = "markdown") 
```

### 2020-05-15 manuscript/appendix.Rmd
move random forest plot to appendix

@@
 
+```{r 02_depth_model_importance, fig.cap = "Importance plot for random forest variables showing increase in mean square error.
+Higher values indicate greater importance to model predictions.
+See Equation 1 for a definition of geometric max depth.
+HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")
+```
+
 ```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}

### 2020-05-18 manuscript/figures.Rmd
cleanup table captions

@@
 \captionsetup[figure]{labelformat=empty}
-```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth and predictor variables for modelling efforts (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
+```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth and predictor variables for computing random forest offsets (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
 Predictor variables used in Eq 2 are printed in bold face.
-Dashes (-) indicate identical sample size among this study and that of the contiguous United States.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../tables/01_predictors.pdf")
@@
 
-```{r depth_model_metrics, fig.cap = "Model fit and predictive accuracy metrics for all combinations of true (inlake slope, deepest point distance) and proxy (nearshore slope, visual center distance) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+\captionsetup[figure]{name=Table}
+\captionsetup[figure]{labelformat=empty}
+```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuracy metrics for all combinations of true (inlake slope, deepest point distance) and proxy (nearshore slope, visual center distance) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 # see tables/02_model_metrics.Rmd
@@
 ```
+\captionsetup[figure]{name=Figure}
+\captionsetup[figure]{labelformat=default}
+\setcounter{figure}{3}
 

### 2020-05-26 tables/02_model_metrics.Rmd
drop extra sig figs in rmse, closes #11

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
-# setwd("figures")
+# setwd("tables")
 ```
@@
     TRUE ~ rsq)) %>%
+  mutate(rmse = round(as.numeric(rmse), 1)) %>%
   mutate(rmse = case_when(

### 2020-05-29 manuscript/figures.Rmd
move residual explanation to figure caption

@@
 
-```{r depth_model_grid_resid, fig.cap = "Depth model residuals in meters by cross-section shape and reservoir class.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r depth_model_grid_resid, fig.cap = "Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 # see 02_depth_model.Rmd

### 2020-06-01 figures/01_contrasts.Rmd
sync terminology with ms text

@@
   contrast_plot(grp = "shape_class", varb = "inlake_slope", 
-                label = "Inlake slope (m/m)", plot_type = "pointrange")
+                label = "In-lake slope (m/m)", plot_type = "pointrange")
 ```
@@
   contrast_plot(grp = "shape_class", varb = "dist_deepest", 
-                label = "Deepest point distance (m)", plot_type = "pointrange")
+                label = "Distance to deepest point (m)", plot_type = "pointrange")
 ```
@@
   contrast_plot(grp = "reservoir_class", varb = "inlake_slope", 
-                label = "Inlake slope (m/m)", plot_type = "pointrange")
+                label = "In-lake slope (m/m)", plot_type = "pointrange")
 ```
@@
   contrast_plot(grp = "reservoir_class", varb = "dist_deepest", 
-                label = "Deepest point distance (m)", plot_type = "pointrange")
+                label = "Distance to deepest point (m)", plot_type = "pointrange")
 ```
@@
   gg_area_shape_class + ggtitle("Area (ha)") + theme_opts, 
-  gg_slope_shape_class + ggtitle("Inlake slope (m/m)") + theme_opts, 
-  gg_dist_shape_class + ggtitle("Deepest point distance (m)") + theme_opts, 
+  gg_slope_shape_class + ggtitle("In-lake slope (m/m)") + theme_opts, 
+  gg_dist_shape_class + ggtitle("Dist. deepest pnt (m)") + theme_opts, 
   nrow = 1)

### 2020-06-01 manuscript/appendix.Rmd
organize appendix
sync terminology with ms text

@@
 
-\textbf{Bathymetry data from thousands of lakes show that lake depth prediction is confounded by difficulty modeling inlake slope}
+\textbf{Geometric models to predict lake depth are sensitive to mis-measurement of in-lake slope}
 
@@
 
+\vspace{1em}
+
 \textbf{Figure S1} Map of study lakes
 
-\textbf{Figure S2} Comparison between reported depth and depth estimated from bathymetry surfaces
+\textbf{Figure S2} Comparison between reported depth and depth extracted from bathymetry surfaces
 
@@
 \textbf{Figure S4} Hypsography classification by state
+
+\textbf{Figure S5} Comparison among lake shape and reservoir classes for true and proxy geometry measures
+
+\textbf{Figure S6} Importance plot for random forest variables showing increase in mean square error
 
@@
 
-```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.
-A dashed 1:1 line is shown for comparison.
-Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
-```
-
-```{r 02_depth_model_importance, fig.cap = "Importance plot for random forest variables showing increase in mean square error.
-Higher values indicate greater importance to model predictions.
-See Equation 1 for a definition of geometric max depth.
-HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
-knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")
-```
-
 ```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
@@
 
-```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth estimated from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product (citation).
+```{r lgnemanual-vs-bathy-depth, fig.cap="Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product (citation).
 For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
@@
 knitr::include_graphics("../figures/01_hypsography-1.pdf")
+```
+
+```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.
+A dashed 1:1 line is shown for comparison.
+Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/01_geometry_grid-1.pdf")
+```
+
+```{r 02_depth_model_importance, fig.cap = "Importance plot for random forest variables showing increase in mean square error.
+Higher values indicate greater importance to model predictions.
+See Equation 1 for a definition of geometric max depth.
+HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")
 ```

### 2020-06-01 manuscript/figures.Rmd
sync terminology with ms text

--- 2020-05-18 manuscript/figures.Rmd
@@
 ```{r lake_shape, echo=FALSE, out.extra='', fig.align='center', out.width="4in",fig.cap="Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
-Dashed lines represent extrapolated nearshore slope while solid lines represent the lake bottom."}
+Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom."}
 knitr::include_graphics("../figures/lake_shape.pdf")
@@
 \captionsetup[figure]{labelformat=empty}
-```{r 01_predictors, fig.cap = "Table 1: Summary of lake depth and predictor variables for computing random forest offsets (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
-Predictor variables used in Eq 2 are printed in bold face.
+```{r 01_predictors, fig.cap = "Table 1: Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
+Predictor variables for computing random forest offsets (Eq 2) are printed in bold face.
 Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
@@
 
-```{r 01_geometry_base, fig.cap = "Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance to lake visual center and B) nearshore slope versus inlake slope.
+```{r 01_geometry_base, fig.cap = "Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
 A best-fit line and coefficient of determination is shown to illustrate representativeness.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
@@
 \captionsetup[figure]{labelformat=empty}
-```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuracy metrics for all combinations of true (inlake slope, deepest point distance) and proxy (nearshore slope, visual center distance) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuracy metrics (RMSE = root mean square error, R2 = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 # see tables/02_model_metrics.Rmd
@@
 
-```{r depth_model_grid_resid, fig.cap = "Depth model residuals in meters by cross-section shape and reservoir class.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+```{r depth_model_grid_resid, fig.cap = "Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
 # see 02_depth_model.Rmd

### 2020-06-01 tables/01_predictors.Rmd
sync terminology with ms text

@@
 # bold Eq 2 predictor variables
-dt$Variable[!stringr::str_detect(dt$Variable, "slope|distance|depth")] <- 
+dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")] <- 
   paste0("**", 
-         dt$Variable[!stringr::str_detect(dt$Variable, "slope|distance|depth")],
+         dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")],
          "**")

### 2020-06-08 manuscript/manuscript.Rmd
start converting gdoc ms

---
title: Geometric models overestimate lake depth due to imperfect slope measurement

# Use letters for affiliations
author:
  - name: Joseph Stachelek
    affiliation: a
  - name: Patrick Hanly
    affiliation: a
  - name: Patricia A. Soranno
    affiliation: a

address:
  - code: a
    address: Department of Fisheries and Wildlife, Michigan State University, 480 Wilson Rd., East Lansing, MI 48824, USA

# Optional: line of arbitrary text with additional information.
# Could be used, for example, to mention the bibliographic info in a post-print.
# If not specified, defaults to "This version was compiled on \today"
#date_subtitle: Published in *Journal of Statistical Software*, 2018
# date_subtitle: > 
# This is the accepted version of the following article: J. Stachelek et al., Ecological Applications. In Press, which has been published in final form at DOI:  [DOI](https://doi.org).  This article may be used for non-commercial purposes in accordance with Wiley Terms and Conditions for Use of Self-Archived Versions.

# For footer text  TODO(fold into template, allow free form two-authors)
lead_author_surname: Stachelek et al.

# Place eg a DOI URL or CRAN Package URL here
# doi_footer: "https://doi.org/10.5281/zenodo.3754916"

# Abstract
abstract: |
  Lake depth is a critical characteristic that influences many important ecological processes in lakes. Unfortunately, lake depth measurements are labor-intensive to gather and are only available for a small fraction of lakes globally. Therefore, scientists have tried to predict lake depth from characteristics that are easily obtained for all lakes such as lake surface area or the slope of the land surrounding a lake. One approach for predicting lake depth simulates lake basins using a geometric model where nearshore land slope is assumed to be a representative proxy for in-lake slope and the distance to the center of the lake is assumed to be a representative proxy for the distance to the deepest point of the lake. However, these assumptions have rarely been tested in a broad range of lakes. We used bathymetry data from 5,000 lakes and reservoirs to test these assumptions and to examine whether differences in lake type or shape influences depth prediction error. We found that nearshore land slope was not a representative proxy of in-lake slope and using it for prediction increases prediction error substantially relative to models using true in-lake slope. We also found that models using nearshore land slope as a proxy systematically overpredicts lake depth in concave lakes (i.e. bowl-shaped; up to 18% of lakes in the study population) and reservoir lakes (up to 30% of lakes in the study population), suggesting caution in using geometric models for lake prediction in unsampled lakes.

# Optional: Acknowledgements
acknowledgements: |
  Funding was provided by the US NSF Macrosystems Biology Program grants, DEB-1638679; DEB-1638550, DEB-1638539, DEB-1638554. PAS was also supported by USDA National Institute of Food and Agriculture Hatch Project, Grant Number: 176820. Author contributions: JS conceived of the study, built models, analyzed data, and wrote the paper. PH and PAS provided interpretation of results and edited the paper.

# Optional: One or more keywords
keywords:
  - lake depth
  - bathymetry
  - hypsography
  - geometry
  - slope

# Paper size for the document, values of letter and a4
papersize: letter

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 9pt

# Optional: Force one-column layout, default is two-column
#one_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
#numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography 
bibliography: lagosdepth
biblio-style: jsta

# Optional: Enable a 'Draft' watermark on the document
#watermark: true

# Customize footer, eg by referencing the vignette
# footer_contents: "Agriculture-lake nutrient relationships"

# Produce a pinp document
output: pinp::pinp

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{YourPackage-vignetteentry}
  %\VignetteKeywords{YourPackage, r, anotherkeyword}
  %\VignettePackage{YourPackage}
  %\VignetteEngine{knitr::rmarkdown}
---

## Introduction

Lake depth is a major factor controlling lake physics, chemistry, and biota (Li et al. 2020, Fee 1996, Read et al. 2014). However, because measured depth data is only available for a small fraction of all lakes, our ability to understand and predict depth-dependent processes is limited. The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake (Heathcote 2015, Oliver 2016, Sobek 2011). The approaches taken in these studies have been varied. In addition, they have explored a variety of strategies to maximize prediction accuracy such as employing more diverse covariates (Oliver 2016), varying lake buffer sizes (Heathcote 2015), or estimating hidden groupings among lakes (Sobek 2011). 
One approach for predicting lake depth involves a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid (Hollister 2011, Neumann 1959, Yigsaw 2018). All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae. In the simplest case, where lakes basins are treated as cones (Eq 1, Figure 1), two assumptions are required to make depth predictions for all lakes: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the lake is a representative proxy for the distance to the deepest point of the lake (Figure 1). This cone model imposes the following fixed (i.e. geometric) relationship between slope and horizontal distance:

Geometric formula: depth_geometric = tan(slope) * distance             (Equation 1)

where the product of slope and horizontal distance yields an exact geometric depth estimate (depth_geometric). 
The assumptions of this model as well as other geometric models can be tested by comparing proxy measures of lake geometry against corresponding “true” (i.e. in-lake) values derived from bathymetric maps. However, due to the scarcity of bathymetric map data, few studies have tested these assumptions except in individual studies of very large lakes or for small numbers of lakes (< 100) (Johansson et al. 2007). As a result, it is unclear what effect proxies have on lake depth predictions using geometric models and whether depth predictions are more sensitive to errors in the horizontal dimension (i.e. distance to the deepest point of the lake) or errors in the vertical dimension (i.e. in-lake slope). Additionally it is unclear whether model prediction error is related to differences in lake type such those with different cross-section shapes or those classified as reservoirs versus natural lakes. This information could help direct future research efforts to focus on particular dimensions of lake geometry (i.e. horizontal versus vertical) or to stratify model predictions based on specific lake types and cross-section shapes. 
Given the knowledge gaps identified above, we asked three research questions: 1) Is nearshore land slope representative of in-lake slope; and is the distance to the center of a lake representative of the distance to the deepest point of a lake? 2) How does the use of proxies for lake geometry affect lake depth prediction error? 3) How does lake shape (i.e. concave versus convex) and lake type (i.e. natural lake vs reservoir) affect depth prediction error? 
To answer these questions, we extracted the maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 lakes in the Northeastern and Midwestern US from bathymetric map data. First, we examined whether measures of lake geometry (in-lake slope and distance to the deepest point of lakes) were related to geometry proxies (nearshore land slope and distance to lake centers). Next, we computed geometric depth estimates (Equation 1) and offsets to these estimates using the random forest algorithm (Equation 2). Covariates used in offset modeling included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table 1). We examined differences in overall prediction error corresponding to different inputs to geometric depth as well as differences in the relative distribution of prediction error in lakes with different characteristics.
Given that, by definition, the distance proxy (distance to the center of the lake) must always be greater or equal to the true distance value (distance to the deepest point of the lake), we expect that the use of this proxy will lead to overestimation of lake depth (Figure 1). Furthermore we expect to see greater overestimation error in reservoirs as compared to natural lakes because many reservoirs are known to be drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e. next to the dam) rather than in the center of the reservoir (Lanza and Silvey 1985). Similar to distance proxies, we expect that the use of a nearshore land slope proxy will lead to prediction error according to differences in cross-section shape (Figure 2) such that the depth of U-shaped (i.e. concave) lakes will be overpredicted whereas the depth of V-shaped (i.e. convex) lakes will be underpredicted. Finally, we expect that offsets to geometric depth estimates will be most strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies (Oliver et al. 2016).

### 2020-06-09 manuscript/manuscript.Rmd
towards linked citations
copy over text up to results section

@@
 
-## Introduction
+## INTRODUCTION
 
-Lake depth is a major factor controlling lake physics, chemistry, and biota (Li et al.
-2020, Fee 1996, Read et al.
-2014).
+Lake depth is a major factor controlling lake physics, chemistry, and biota
+\citep{liSignificantContributionLake2020, feeEffectsLakeSize1996, readSimulating2368Temperate2014}.
 However, because measured depth data is only available for a small fraction of all lakes, our ability to understand and predict depth-dependent processes is limited.
-The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake (Heathcote 2015, Oliver 2016, Sobek 2011).
+The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \citep{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
 The approaches taken in these studies have been varied.
-In addition, they have explored a variety of strategies to maximize prediction accuracy such as employing more diverse covariates (Oliver 2016), varying lake buffer sizes (Heathcote 2015), or estimating hidden groupings among lakes (Sobek 2011).
-One approach for predicting lake depth involves a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid (Hollister 2011, Neumann 1959, Yigsaw 2018).
+In addition, they have explored a variety of strategies to maximize prediction accuracy such as employing more diverse covariates \citep{oliver2016prediction}, varying lake buffer sizes \citep{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings among lakes \citep{sobekPredictingDepthVolume2011}.
+
+One approach for predicting lake depth involves a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \citep{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
 All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae.
-In the simplest case, where lakes basins are treated as cones (Eq 1, Figure 1), two assumptions are required to make depth predictions for all lakes: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the lake is a representative proxy for the distance to the deepest point of the lake (Figure 1).
+In the simplest case, where lakes basins are treated as cones (Eq \ref{eq1}, Figure \ref{fig1}), two assumptions are required to make depth predictions for all lakes: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the lake is a representative proxy for the distance to the deepest point of the lake (Figure 1).
 This cone model imposes the following fixed (i.e.
 geometric) relationship between slope and horizontal distance:
-
-Geometric formula: depth_geometric = tan(slope) * distance             (Equation 1)
+\begin{align}
+  depth_{geometric} = tan(slope) * distance \label{eq1}
+\end{align}
 
 where the product of slope and horizontal distance yields an exact geometric depth estimate (depth_geometric).
+
 The assumptions of this model as well as other geometric models can be tested by comparing proxy measures of lake geometry against corresponding “true” (i.e.
 in-lake) values derived from bathymetric maps.
-However, due to the scarcity of bathymetric map data, few studies have tested these assumptions except in individual studies of very large lakes or for small numbers of lakes (< 100) (Johansson et al.
-2007).
+However, due to the scarcity of bathymetric map data, few studies have tested these assumptions except in individual studies of very large lakes or for small numbers of lakes (< 100) \citep{johanssonNewApproachesModelling2007}.
 As a result, it is unclear what effect proxies have on lake depth predictions using geometric models and whether depth predictions are more sensitive to errors in the horizontal dimension (i.e.
@@
 horizontal versus vertical) or to stratify model predictions based on specific lake types and cross-section shapes.
+
 Given the knowledge gaps identified above, we asked three research questions: 1) Is nearshore land slope representative of in-lake slope; and is the distance to the center of a lake representative of the distance to the deepest point of a lake?
@@
 natural lake vs reservoir) affect depth prediction error?
+
 To answer these questions, we extracted the maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 lakes in the Northeastern and Midwestern US from bathymetric map data.
@@
 We examined differences in overall prediction error corresponding to different inputs to geometric depth as well as differences in the relative distribution of prediction error in lakes with different characteristics.
+
 Given that, by definition, the distance proxy (distance to the center of the lake) must always be greater or equal to the true distance value (distance to the deepest point of the lake), we expect that the use of this proxy will lead to overestimation of lake depth (Figure 1).
 Furthermore we expect to see greater overestimation error in reservoirs as compared to natural lakes because many reservoirs are known to be drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
-next to the dam) rather than in the center of the reservoir (Lanza and Silvey 1985).
+next to the dam) rather than in the center of the reservoir \citep{lanza1985interactions}.
 Similar to distance proxies, we expect that the use of a nearshore land slope proxy will lead to prediction error according to differences in cross-section shape (Figure 2) such that the depth of U-shaped (i.e.
@@
 convex) lakes will be underpredicted.
-Finally, we expect that offsets to geometric depth estimates will be most strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies (Oliver et al.
-2016).
+Finally, we expect that offsets to geometric depth estimates will be most strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies \citep{oliver2016prediction}.
+
+## METHODS
+### Data description
+We compiled bathymetry data on approximately 5,000 lakes in the Northeastern and Midwestern US from nine official state databases (Figure S1).
+The original data came in a variety of formats including pre-interpolated rasters (Minnesota), contour lines (Nebraska, Michigan, Massachusetts, Kansas, Iowa), contour polygons (New Hampshire, Connecticut), or point depth soundings (Maine).
+For the Minnesota data, we simply clipped the raster for each lake to its outline.
+For data from the remaining states, we processed each lake by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the raster R package \citep{hijmansRasterGeographicData2019}.
+The size of the moving window was adjusted iteratively to ensure that each bathymetry raster contained no missing data.
+
+All lake bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) NHD \citep{usgsNationalHydrographyDataset2019} waterbodies such that source data and bathymetry surface outputs were clipped to the area of each lake polygon.
+We restricted the lakes in our study to those with an area of at least 4 ha and a maximum depth of at least 0.3 m (1 ft).
+The purpose of these restrictions was to ensure that lakes had enough contours (or points, or polygons) to generate adequately smooth interpolations with which to calculate in-lake geometry metrics.
+
+We used our generated bathymetry surfaces to find the location of the deepest point in the lake and we resolved ties by choosing the deepest point that was closest to the center of the lake.
+We calculated the center of the lake not as its centroid but rather by finding the point farthest from the lake shoreline (i.e.
+its “visual distance to lake center”).
+For these calculations, we used the polylabelr R package \citep{larssonPolylabelrFindPole2019d}, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
+We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model (~15x15m grain) accessed using the elevatr R package \citep{hollisterElevatrAccessElevation2017} and the terrain function in the raster R package \citep{hijmansRasterGeographicData2019}.
+
+We categorized lakes based on their cross-section shape and reservoir class.
+For cross-section shape, we categorized lakes as either convex or concave following the method of \citet{hakansonLakeFormLake1977} by computing normalized lake depth-area relationships (i.e.
+hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S4).
+We further classified lakes using data from \citet{polus2020}, which uses the output of a machine learning algorithm to assign a probability to each lake regarding whether it is a reservoir or a natural lake.
+For our purposes, we determined a lake to be a reservoir if the classification probability was 0.75 or greater.
+The \citet{polus2020} data product defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water, beyond simply controlling water level.
+It makes no distinction between different dam types or dam heights.
+
+Covariates used in random forest modeling (Table 1, Equation 2) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
+HUC4s), were obtained from the LAGOS-US LOCUS database (Smith et al.
+In Prep).
+Shoreline development is a measure of lake perimeter shape defined as:
+
+\begin{align}
+shoreline_{devel} = perimeter / (2 * \sqrt{(\pi * waterarea * 10000)})        
+\end{align}
+
+where sinuous lakes have larger values of shoreline development and circular lakes have smaller values of shoreline development.
+Watershed to lake area ratio is an approximation of water residence time and is defined as watershed area divided by lake area.
+All of our data and data processing code are available at [Zenodo DOI].
+
+2.2. Proxy evaluation
+We conducted a qualitative assessment of whether or not proxy measures of lake geometry are representative of their true values by visual inspection (i.e.
+plotting each proxy measure against its corresponding true value) and by computing coefficients of determination ($R^2$).
+We further tested proxy measures by examining their effect on lake depth prediction error.
+Our approach involved several steps.
+In the first step, we computed a geometric estimate of lake depth using only geometry information (depth_geometric, Equation 1).
+In the second step, we fit a random forest model to predict observed (i.e.
+true) depth as a function of geometric depth along with several covariates available for all lakes (Table 1).
+The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of depth_geometric and among different lake types.
+Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table 2).
+
+## Model description
+### Geometric model 
+We used a geometric model of lakes where basins are treated as cones with a fixed relationship between slope and distance (Equation 1).
+Note that Equation 1 is a geometric formula and has no intercept or “coefficients” and it produces a perfect depth estimate given true values of slope and distance.
+To use this model to predict the depth of all lakes, there is a necessary assumption that proxy slope and distance measures, which are available for all lakes, are representative of true slope and distance (Figure 1).
+
+### Random forest models
+Nearly all prior studies predicting lake depth using geometric models include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy.
+For our purposes, this offset modeling enabled us to test our expectations that prediction error would be different among different formulations of depth_geometric and among different lake types.
+It also facilitated direct comparison against prior models of lake depth including those that are non-geometric.
+We generated an “offset” to geometric depth (sensu Hollister 2011) using the random forest algorithm and the ranger R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation 1) along with the lake elevation, area, perimeter, and ratio/index measures listed in Table 1:
+
+\begin{align}
+  depth_{observed} ~ depth_{geometric} + covariates         
+\end{align}
+
+Neither cross-section shape nor reservoir class was used as a covariate in any random forest models.
+We used the random forest algorithm because it makes no assumptions about the distribution of model residuals, allows for non-linearity, and is insensitive to interactions (i.e.
+multicollinearity) among covariates \citep{prasadNewerClassificationRegression2006}.
+
+### Model comparisons
+We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation 1 (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
+Prior to entry into Equation 1, we standardized proxy distances to have the same numeric range as their true counterpart.
+The purpose of this standardization was to prevent lakes with extremely long proxy distances from having an outsized impact on model evaluation metrics.
+
+### Model evaluations
+We evaluated model fit and prediction error using root-mean-square error (RMSE) and coefficient of determination ($R^2$) metrics on a holdout set containing 25\% of all lakes.
+We evaluated the residuals (residual = observed - predicted) of each model relative to lake cross-section shape and reservoir classes to determine whether depth is consistently over or under predicted for some lake types relative to others.
+All of our code for model fitting and evaluation is available at [Zenodo DOI].
+
+## RESULTS
+Lakes belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Figure S1).
+For example, concave lakes were nearly absent from Michigan whereas Maine lakes had an overabundance of lakes categorized as neither concave nor convex.
+Lakes in the southern portions of our study area tended to be classified as reservoirs whereas lakes in the northern portions of our study area were a more even mix between reservoirs and natural lakes (Figure S1).
+Approximately 18\%, 80\%, and 2\% of lakes were classified as having a concave, convex, or neither shape respectively whereas approximately 30\% and 70\% of lakes were classified as being a reservoir or a natural lake.
+
+Although proxy distance to lake center was often much larger in magnitude compared to the true distance to the deepest point of lakes’, they were strongly related ($R^2$ = 0.8).
+In contrast, proxy nearshore land slope and true in-lake slope were more weakly related ($R^2$ = 0.17).
+For slope measures, most lakes had lower magnitude (i.e.
+shallower) nearshore land slope compared to true in-lake slope (Figure 3).
+Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes while proxy nearshore land slope is not representative of true in-lake slope.
+In addition to overall differences between slope and distance measures, we found differences in these relationships among lake shape classes.
+For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S5).
+However,  there were not similar differences among slope and distance metrics for natural lakes versus reservoirs (Figure S5).
+
+The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table 2).
+More specifically, the true slope - proxy distance model had a better fit (R2 = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model (R2 = 0.26, RMSE = 6.87m).
+Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure 4).
+We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure 4).
+
+The most important covariates in offset models were those relating to spatial location, lake area, and perimeter (Figure S6).
+Conversely, watershed metrics and lake elevation had little contribution to random forest model fit (Figure S6).
+The spatial location (i.e.
+HUC4) covariate was notably less importance in the true slope model compared to the two proxy slope models.
+Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Eq 1 (Figure S6).

### 2020-06-09 manuscript/manuscript.Rmd
add inline figures and citations

@@
 
+\begin{figure}[H]
+  \begin{center}
+    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/slope_diagram_new}
+  \end{center}
+  \caption{Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
+Geometric depth calculated via Equation 1 requires a single distance and slope metric.}\label{fig1}
+\end{figure}
+
 The assumptions of this model as well as other geometric models can be tested by comparing proxy measures of lake geometry against corresponding “true” (i.e.
@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/lake_shape}
+  \end{center}
+  \caption{Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
+Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom.}\label{fig2}
+\end{figure}
+
 ## METHODS
@@
 For the Minnesota data, we simply clipped the raster for each lake to its outline.
-For data from the remaining states, we processed each lake by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the raster R package \citep{hijmansRasterGeographicData2019}.
+For data from the remaining states, we processed each lake by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the \texttt{raster} R package \citep{hijmansRasterGeographicData2019}.
 The size of the moving window was adjusted iteratively to ensure that each bathymetry raster contained no missing data.
@@
 its “visual distance to lake center”).
-For these calculations, we used the polylabelr R package \citep{larssonPolylabelrFindPole2019d}, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
-We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model (~15x15m grain) accessed using the elevatr R package \citep{hollisterElevatrAccessElevation2017} and the terrain function in the raster R package \citep{hijmansRasterGeographicData2019}.
+For these calculations, we used the \texttt{polylabelr} R package \citep{larssonPolylabelrFindPole2019d}, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
+We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model (~15x15m grain) accessed using the \texttt{elevatr} R package \citep{hollisterElevatrAccessElevation2017} and the terrain function in the \texttt{raster} R package \citep{hijmansRasterGeographicData2019}.
 
@@
 It also facilitated direct comparison against prior models of lake depth including those that are non-geometric.
-We generated an “offset” to geometric depth (sensu Hollister 2011) using the random forest algorithm and the ranger R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation 1) along with the lake elevation, area, perimeter, and ratio/index measures listed in Table 1:
+We generated an “offset” to geometric depth \citep[sensu][]{hollisterPredictingMaximumLake2011} using the random forest algorithm and the \texttt{ranger} R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation 1) along with the lake elevation, area, perimeter, and ratio/index measures listed in Table 1:
 
 \begin{align}
-  depth_{observed} ~ depth_{geometric} + covariates         
+  depth_{observed} \sim depth_{geometric} + covariates         
 \end{align}
@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/01_geometry_base-1}
+  \end{center}
+  \caption{Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
+A best-fit line and coefficient of determination is shown to illustrate representativeness.}\label{fig3}
+\end{figure}
+
 The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table 2).
-More specifically, the true slope - proxy distance model had a better fit (R2 = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model (R2 = 0.26, RMSE = 6.87m).
+More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model ($R^2$ = 0.26, RMSE = 6.87m).
 Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure 4).
 We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure 4).
+
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/02_depth_model_grid_resid-1}
+  \end{center}
+  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig4}
+\end{figure*}
 
@@
 Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Eq 1 (Figure S6).
+
+## DISCUSSION
+
+Our tests of geometric lake depth models show that specific proxy measures of lake geometry are not representative of true measures of lake geometry across a broad array of lakes.
+Using a cone model example, we show that nearshore land slope is not representative of in-lake slope.
+Furthermore, our results indicate that use of this metric for prediction results in increased prediction error and systematic overestimation of depth in concave and reservoir lakes.
+
+### Representativeness of proxy measures of lake geometry
+In comparing among lake geometry measures, our analysis suggests that proxy distance to lake center is representative of true distance to the deepest point of the lakes but that proxy nearshore land slope is not representative of true in-lake slope.
+A simple indication of this non representativeness is that proxy nearshore land slope was often (in > 74\% of cases) steeper than true in-lake slope.
+This finding is consistent with the suggestion by \citet{heathcotePredictingBathymetricFeatures2015} that in-lake slope matches nearshore land slope at the time of lake formation, but as lakes age, infilling via sedimentation causes them to develop shallower in-lake slopes compared to the surrounding land.
+If true, then these results also lend support to prior work \citep{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}, conducted on a more limited number of lakes, showing that in-lake slopes are controlled by erosion and sediment transport processes.
+
+One surprising finding with respect to the relationship between true and proxy geometry measures when examined by lake class was the fact that there was no greater difference between proxy and true distances in reservoirs compared to natural lakes.
+This is contrary to the idea that most reservoirs are drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
+next to the dam) rather than in the center of the reservoir \citep{lanza1985interactions}.
+One possible explanation is that the reservoir classification data from \citet{polus2020} uses a more general definition of a reservoir (i.e.
+any permanent waterbody that has a water control structure likely to significantly impact flow or pool water) compared to that of conventional classifications that are tied to specific dam types or dam heights.
+Another possible explanation is that conventional reservoir classifications are conceptually biased towards more southern areas with few natural lakes (Figure S1).
+
+We found other differences between convex and concave lakes.
+One finding was that convex lakes, when compared to concave lakes, had longer distances to lake centers relative to corresponding distances to the deepest point of lakes.
+In addition, convex lakes often had steeper in-lake slopes relative to nearshore land slopes as compared to concave lakes.
+These differences are reflected in the class-wise differences in shoreline development (e.g.
+shorelines were more sinuous in convex lakes compared to concave lakes, Figure S3).
+It was notable that convex lakes were deeper than concave lakes despite having similar distributions of lake surface area (Figure S3).
+Given the similarity in lake surface area, the underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of lake formation whereby concave lakes were formed as a result of more intense glacial scouring compared to convex lakes \citep{gorhamPhysicalLimnologyNorthern1958}.
+While there is some evidence in support of this idea, namely that there is a geographic hotspot of concave lakes associated with the glaciated “prairie pothole region” \citep[see][]{hayashiSimpleEquationsRepresent2000}, the overall geographic distribution of lake cross-section shapes does not support this idea.
+Instead of a concentrated area of concave lakes in formerly glaciated regions, there appears to be a fairly even mix of concave and convex lakes distributed amongst the northern (i.e.
+glaciated) and southern (non-glaciated) portions of our study area (Figure S1).
+
+### Effects of proxy measures of lake geometry depth prediction error
+Models using only proxy variables had prediction error rates (RMSE = 6.6m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting lake depth at broad geographic extents \citep{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
+When only a single proxy measure was used there was a difference in model sensitivity depending on if it was a horizontal distance measure or a vertical slope measure.
+In the case of a true slope and proxy distance combination, models were more accurate ($\pm$ 4.2m) than even the most accurate of prior studies \citep{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
+Conversely, models using a proxy slope and true distance combination had prediction error rates ($\pm$ 6.9m) of a similar magnitude as that of the baseline proxy-proxy model ($\pm$ 6.6m).
+The greater sensitivity of depth predictions to proxy slope measures relative to proxy distance measures may be explained by the fact that proxy slope measures were a more imperfect representation of true in-lake slopes relative to proxy versus true distances.
+In addition, these results help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \citep{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S3).
+
+### Effects of lake shape and lake type on depth prediction error
+As expected, we found that the maximum depth of concave lakes was systematically overpredicted by a simple geometric model using proxy nearshore land slope.
+However, contrary to our expectation, we did not observe underprediction of depth in convex lakes.
+The reason we did not observe underprediction of the depth of convex lakes is likely because geometric depth itself was always greater than observed maximum depth owing to the fact that proxy distance is constrained to be greater than true distance.
+Although the magnitude of overestimation is likely related to class imbalance in our dataset (i.e.
+there was a greater number of convex lakes), these results suggest that broad scale estimates of lake depth are overestimated particularly when those estimates encompass large numbers of lakes with diverse cross-section shapes.
+
+### Future research
+One of our models (true slope, proxy distance) was more accurate than even the most accurate of prior studies.
+However, parameterization of this model requires data on bathymetry which is not available for all lakes.
+We propose that the error rate of this model ($\pm$ 4.2m) be used as an out-of-sample prediction benchmark for future studies such that they should attempt to match it but not expect to exceed it.
+Because this model requires bathymetry data, this suggests that it may not be possible with current data and models to produce depth predictions for all lakes with error rates below about 6m.
+To approach our benchmark using data available for all lakes, future studies could explore alternative modeling approaches such as ordinal modeling, which would capture whether or not a lake crosses some important depth threshold, but would not seek to predict a specific depth value.
+These studies could also explore emerging data types to boost prediction accuracy such as “topobathymetric” products that integrate both topographic and bathymetric data in a seamless fashion rather than treating them as separate entities.
+This would allow for more robust tests of the representativeness of geometric model inputs.
+Unfortunately, topobathymetric products are rare, have mostly been limited nearshore marine environments, and as such are not widely available for inland waters \citep{danielsonTopobathymetricElevationModel2016}.
+
+Finally, our findings indicate that geometry measures differ according to lake cross-section shape.
+This makes it an attractive target for inclusion in depth prediction models.
+Unfortunately, identifying a lake’s cross-section shape requires bathymetry data which is unavailable for most lakes.
+However, given the conceptual links between cross-section shape, glaciation, and sedimentation \citep{johanssonNewApproachesModelling2007} it may be advantageous for future studies to compile data on sedimentation to determine if this data can be used to predict cross-section shape and boost depth prediction accuracy.

### 2020-06-09 manuscript/manuscript.Rmd
add inline tables

@@
 
+\begin{table}
+\centering
+\setlength\tabcolsep{1.5pt} % default value: 6pt
+\begin{tabular}{lllll}
+  \hline
+  Variable & Median & Q25 & Q75 & n\\
+  \hline
+  Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4850 (17700)\\
+  \textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4850 (17700)\\
+  \textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4850 (17700)\\
+  \textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4850 (17700)\\
+  \textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4850 (17700)\\
+  \textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4850 (17700)\\
+  \textbf{Watershed-lake ratio} & 7.8 (10) & 3.9 (4.4) & 17 (28) & 4850 (17700)\\
+  Distance to deepest point (m) & 180 (-) & 110 (-) & 290 (-) & 4850 (-)\\
+  Distance to lake center (m) & 240 (-) & 160 (-) & 380 (-) & 4850 (-)\\
+  In-lake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4850 (-)\\
+  Nearshore land slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4850 (-)\\  
+  \bottomrule
+  \end{tabular}
+  \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
+Predictor variables for computing random forest offsets (Eq 2) are printed in bold face.
+Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
+\end{table}
+
 Covariates used in random forest modeling (Table 1, Equation 2) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
@@
 
+\begin{table}
+  \centering
+  % \setlength\tabcolsep{1.5pt} % default value: 6pt
+\begin{tabular}{llll}
+  \hline
+  slope & distance & rmse & rsq\\
+  \hline
+  true & true & - & -\\
+  true & proxy & 4.2 m & 0.73\\
+  proxy & true & 6.9 m & 0.26\\
+  proxy & proxy & 6.6 m & 0.31\\
+  \bottomrule
+\end{tabular}
+\caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table2}
+\end{table}
+
 \begin{figure*}

### 2020-06-09 tables/01_predictors.Rmd
add inline tables

@@
   - \pagenumbering{gobble}
+editor_options: 
+  chunk_output_type: console
 ---
@@
 # bold Eq 2 predictor variables
-dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")] <- 
+dt_md <- dt
+dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth")] <- 
   paste0("**", 
-         dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")],
+         dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth")],
          "**")
@@
 options(knitr.kable.NA = '-')
-knitr::kable(dt, digits = 0, format = "markdown") 
+knitr::kable(dt_md, digits = 0, format = "markdown")
+
+# ---- latex output ----
+
+# bold Eq 2 predictor variables
+dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")] <- 
+  paste0("\\textbf{", 
+         dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")],
+         "}")
+
+# clipr::write_clip(
+#   knitr::kable(dt, format = 'latex', align = c("lllll"), escape = FALSE))
 ```

### 2020-06-09 tables/02_model_metrics.Rmd
add inline tables

@@
 knitr::kable(dt_metrics, digits = 2, format = "markdown")
+
+# clipr::write_clip(
+#   knitr::kable(dt_metrics, format = 'latex', align = c("llll"),
+# escape = TRUE))
 ```

### 2020-06-09 manuscript/manuscript.Rmd
cross-link figures,tables, and equations

@@
 # Optional: Enable a 'Draft' watermark on the document
-#watermark: true
+watermark: true
 
 # Customize footer, eg by referencing the vignette
-# footer_contents: "Agriculture-lake nutrient relationships"
+footer_contents: "Geometric lake depth models and bathymetry"
 
@@
 All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae.
-In the simplest case, where lakes basins are treated as cones (Eq \ref{eq1}, Figure \ref{fig1}), two assumptions are required to make depth predictions for all lakes: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the lake is a representative proxy for the distance to the deepest point of the lake (Figure 1).
+In the simplest case, where lakes basins are treated as cones (Eq \ref{eq1}, Figure \ref{fig1}), two assumptions are required to make depth predictions for all lakes: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the lake is a representative proxy for the distance to the deepest point of the lake (Figure \ref{fig1}).
 This cone model imposes the following fixed (i.e.
@@
   \caption{Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
-Geometric depth calculated via Equation 1 requires a single distance and slope metric.}\label{fig1}
+Geometric depth calculated via Equation \ref{eq1} requires a single distance and slope metric.}\label{fig1}
 \end{figure}
@@
 First, we examined whether measures of lake geometry (in-lake slope and distance to the deepest point of lakes) were related to geometry proxies (nearshore land slope and distance to lake centers).
-Next, we computed geometric depth estimates (Equation 1) and offsets to these estimates using the random forest algorithm (Equation 2).
-Covariates used in offset modeling included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table 1).
+Next, we computed geometric depth estimates (Equation \ref{eq1}) and offsets to these estimates using the random forest algorithm (Equation \ref{eq2}).
+Covariates used in offset modeling included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table \ref{table1}).
 We examined differences in overall prediction error corresponding to different inputs to geometric depth as well as differences in the relative distribution of prediction error in lakes with different characteristics.
 
-Given that, by definition, the distance proxy (distance to the center of the lake) must always be greater or equal to the true distance value (distance to the deepest point of the lake), we expect that the use of this proxy will lead to overestimation of lake depth (Figure 1).
+Given that, by definition, the distance proxy (distance to the center of the lake) must always be greater or equal to the true distance value (distance to the deepest point of the lake), we expect that the use of this proxy will lead to overestimation of lake depth (Figure \ref{fig1}).
 Furthermore we expect to see greater overestimation error in reservoirs as compared to natural lakes because many reservoirs are known to be drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
 next to the dam) rather than in the center of the reservoir \citep{lanza1985interactions}.
-Similar to distance proxies, we expect that the use of a nearshore land slope proxy will lead to prediction error according to differences in cross-section shape (Figure 2) such that the depth of U-shaped (i.e.
+Similar to distance proxies, we expect that the use of a nearshore land slope proxy will lead to prediction error according to differences in cross-section shape (Figure \ref{fig2}) such that the depth of U-shaped (i.e.
 concave) lakes will be overpredicted whereas the depth of V-shaped (i.e.
@@
   \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
-Predictor variables for computing random forest offsets (Eq 2) are printed in bold face.
+Predictor variables for computing random forest offsets (Eq \ref{eq2}) are printed in bold face.
 Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
@@
 
-Covariates used in random forest modeling (Table 1, Equation 2) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
+Covariates used in random forest modeling (Table \ref{table1}, Equation \ref{eq2}) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
 HUC4s), were obtained from the LAGOS-US LOCUS database (Smith et al.
@@
 
-2.2. Proxy evaluation
+### Proxy evaluation
 We conducted a qualitative assessment of whether or not proxy measures of lake geometry are representative of their true values by visual inspection (i.e.
@@
 Our approach involved several steps.
-In the first step, we computed a geometric estimate of lake depth using only geometry information (depth_geometric, Equation 1).
+In the first step, we computed a geometric estimate of lake depth using only geometry information (depth_geometric, Equation \ref{eq1}).
 In the second step, we fit a random forest model to predict observed (i.e.
-true) depth as a function of geometric depth along with several covariates available for all lakes (Table 1).
+true) depth as a function of geometric depth along with several covariates available for all lakes (Table \ref{table1}).
 The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of depth_geometric and among different lake types.
-Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table 2).
+Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table \ref{table2}).
 
@@
 ### Geometric model 
-We used a geometric model of lakes where basins are treated as cones with a fixed relationship between slope and distance (Equation 1).
-Note that Equation 1 is a geometric formula and has no intercept or “coefficients” and it produces a perfect depth estimate given true values of slope and distance.
-To use this model to predict the depth of all lakes, there is a necessary assumption that proxy slope and distance measures, which are available for all lakes, are representative of true slope and distance (Figure 1).
+We used a geometric model of lakes where basins are treated as cones with a fixed relationship between slope and distance (Equation \ref{eq1}).
+Note that Equation \ref{eq1} is a geometric formula and has no intercept or “coefficients” and it produces a perfect depth estimate given true values of slope and distance.
+To use this model to predict the depth of all lakes, there is a necessary assumption that proxy slope and distance measures, which are available for all lakes, are representative of true slope and distance (Figure \ref{fig1}).
 
@@
 It also facilitated direct comparison against prior models of lake depth including those that are non-geometric.
-We generated an “offset” to geometric depth \citep[sensu][]{hollisterPredictingMaximumLake2011} using the random forest algorithm and the \texttt{ranger} R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation 1) along with the lake elevation, area, perimeter, and ratio/index measures listed in Table 1:
+We generated an “offset” to geometric depth \citep[sensu][]{hollisterPredictingMaximumLake2011} using the random forest algorithm and the \texttt{ranger} R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation \ref{eq1}) along with the lake elevation, area, perimeter, and ratio/index measures listed in Table 1:
 
@@
 ### Model comparisons
-We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation 1 (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
-Prior to entry into Equation 1, we standardized proxy distances to have the same numeric range as their true counterpart.
+We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
+Prior to entry into Equation \ref{eq1}, we standardized proxy distances to have the same numeric range as their true counterpart.
 The purpose of this standardization was to prevent lakes with extremely long proxy distances from having an outsized impact on model evaluation metrics.
@@
 For slope measures, most lakes had lower magnitude (i.e.
-shallower) nearshore land slope compared to true in-lake slope (Figure 3).
+shallower) nearshore land slope compared to true in-lake slope (Figure \ref{fig3}).
 Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes while proxy nearshore land slope is not representative of true in-lake slope.
@@
 
-The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table 2).
+The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table \ref{table2}).
 More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model ($R^2$ = 0.26, RMSE = 6.87m).
-Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure 4).
-We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure 4).
+Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure \ref{fig4}).
+We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure \ref{fig4}).
 
@@
   \hline
-  slope & distance & rmse & rsq\\
+  slope & distance & RMSE & $R^2$\\
   \hline
@@
 HUC4) covariate was notably less importance in the true slope model compared to the two proxy slope models.
-Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Eq 1 (Figure S6).
+Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Eq \ref{eq1} (Figure S6).
 

### 2020-06-09 manuscript/agujournaltemplate.tex
copy over to wrr template

--- 2020-06-09 manuscript/manuscript.Rmd
@@
----
-title: Geometric models overestimate lake depth due to imperfect slope measurement
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
+%
+% This file includes commands and instructions
+% given in the order necessary to produce a final output that will
+% satisfy AGU requirements, including customized APA reference formatting.
+%
+% You may copy this file and give it your
+% article name, and enter your text.
+%
+%
+% Step 1: Set the \documentclass
+%
+%
 
-# Use letters for affiliations
-author:
-  - name: Joseph Stachelek
-    affiliation: a
-  - name: Patrick Hanly
-    affiliation: a
-  - name: Patricia A. Soranno
-    affiliation: a
+%% To submit your paper:
+\documentclass[draft]{agujournal2019}
+\usepackage{url} %this package should fix any errors with URLs in refs.
+\usepackage{lineno}
+\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
+\usepackage{soul}
+\linenumbers
+%%%%%%%
+% As of 2018 we recommend use of the TrackChanges package to mark revisions.
+% The trackchanges package adds five new LaTeX commands:
+%
+%  \note[editor]{The note}
+%  \annote[editor]{Text to annotate}{The note}
+%  \add[editor]{Text to add}
+%  \remove[editor]{Text to remove}
+%  \change[editor]{Text to remove}{Text to add}
+%
+% complete documentation is here: http://trackchanges.sourceforge.net/
+%%%%%%%
 
-address:
-  - code: a
-    address: Department of Fisheries and Wildlife, Michigan State University, 480 Wilson Rd., East Lansing, MI 48824, USA
+\draftfalse
 
-# Optional: line of arbitrary text with additional information.
-# Could be used, for example, to mention the bibliographic info in a post-print.
-# If not specified, defaults to "This version was compiled on \today"
-#date_subtitle: Published in *Journal of Statistical Software*, 2018
-# date_subtitle: > 
-# This is the accepted version of the following article: J.
-Stachelek et al., Ecological Applications.
-In Press, which has been published in final form at DOI:  [DOI](https://doi.org).
-This article may be used for non-commercial purposes in accordance with Wiley Terms and Conditions for Use of Self-Archived Versions.
+%% Enter journal name below.
+%% Choose from this list of Journals:
+%
+% JGR: Atmospheres
+% JGR: Biogeosciences
+% JGR: Earth Surface
+% JGR: Oceans
+% JGR: Planets
+% JGR: Solid Earth
+% JGR: Space Physics
+% Global Biogeochemical Cycles
+% Geophysical Research Letters
+% Paleoceanography and Paleoclimatology
+% Radio Science
+% Reviews of Geophysics
+% Tectonics
+% Space Weather
+% Water Resources Research
+% Geochemistry, Geophysics, Geosystems
+% Journal of Advances in Modeling Earth Systems (JAMES)
+% Earth's Future
+% Earth and Space Science
+% Geohealth
+%
+% ie, \journalname{Water Resources Research}
 
-# For footer text  TODO(fold into template, allow free form two-authors)
-lead_author_surname: Stachelek et al.
+\journalname{Water Resources Research}
 
-# Place eg a DOI URL or CRAN Package URL here
-# doi_footer: "https://doi.org/10.5281/zenodo.3754916"
+\begin{document}
 
-# Abstract
-abstract: |
-  Lake depth is a critical characteristic that influences many important ecological processes in lakes.
-Unfortunately, lake depth measurements are labor-intensive to gather and are only available for a small fraction of lakes globally.
-Therefore, scientists have tried to predict lake depth from characteristics that are easily obtained for all lakes such as lake surface area or the slope of the land surrounding a lake.
-One approach for predicting lake depth simulates lake basins using a geometric model where nearshore land slope is assumed to be a representative proxy for in-lake slope and the distance to the center of the lake is assumed to be a representative proxy for the distance to the deepest point of the lake.
-However, these assumptions have rarely been tested in a broad range of lakes.
-We used bathymetry data from 5,000 lakes and reservoirs to test these assumptions and to examine whether differences in lake type or shape influences depth prediction error.
-We found that nearshore land slope was not a representative proxy of in-lake slope and using it for prediction increases prediction error substantially relative to models using true in-lake slope.
-We also found that models using nearshore land slope as a proxy systematically overpredicts lake depth in concave lakes (i.e.
-bowl-shaped; up to 18% of lakes in the study population) and reservoir lakes (up to 30% of lakes in the study population), suggesting caution in using geometric models for lake prediction in unsampled lakes.
+%% ------------------------------------------------------------------------ %%
+%  Title
+%
+% (A title should be specific, informative, and brief. Use
+% abbreviations only if they are defined in the abstract. Titles that
+% start with general keywords then specific terms are optimized in
+% searches)
+%
+%% ------------------------------------------------------------------------ %%
 
-# Optional: Acknowledgements
-acknowledgements: |
-  Funding was provided by the US NSF Macrosystems Biology Program grants, DEB-1638679; DEB-1638550, DEB-1638539, DEB-1638554.
-PAS was also supported by USDA National Institute of Food and Agriculture Hatch Project, Grant Number: 176820.
-Author contributions: JS conceived of the study, built models, analyzed data, and wrote the paper.
-PH and PAS provided interpretation of results and edited the paper.
+% Example: \title{This is a test title}
 
-# Optional: One or more keywords
-keywords:
-  - lake depth
-  - bathymetry
-  - hypsography
-  - geometry
-  - slope
+\title{Geometric models overestimate lake depth due to imperfect slope
+measurement}
 
-# Paper size for the document, values of letter and a4
-papersize: letter
+%% ------------------------------------------------------------------------ %%
+%
+%  AUTHORS AND AFFILIATIONS
+%
+%% ------------------------------------------------------------------------ %%
 
-# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
-fontsize: 9pt
+% Authors are individuals who have significantly contributed to the
+% research and preparation of the article. Group authors are allowed, if
+% each author in the group is separately identified in an appendix.)
 
-# Optional: Force one-column layout, default is two-column
-#one_column: true
+% List authors by first name or initial followed by last name and
+% separated by commas. Use \affil{} to number affiliations, and
+% \thanks{} for author notes.
+% Additional author notes should be indicated with \thanks{} (for
+% example, for current addresses).
 
-# Optional: Enables lineno mode, but only if one_column mode is also true
-#lineno: true
+% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
+% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
 
-# Optional: Enable one-sided layout, default is two-sided
-#one_sided: true
+\authors{J. Stachelek\affil{1}, P. Hanly\affil{1}, and P. A. Soranno\affil{1}}
 
-# Optional: Enable section numbering, default is unnumbered
-#numbersections: true
+\affiliation{1}{Department of Fisheries and Wildlife, Michigan State University, 480
+Wilson Rd., East Lansing, MI 48824, USA}
+% \affiliation{2}{Second Affiliation}
+% \affiliation{3}{Third Affiliation}
+% \affiliation{4}{Fourth Affiliation}
 
-# Optional: Specify the depth of section number, default is 5
-#secnumdepth: 5
+% \affiliation{=number=}{=Affiliation Address=}
+%(repeat as many times as is necessary)
 
-# Optional: Skip inserting final break between acknowledgements, default is false
-skip_final_break: true
+%% Corresponding Author:
+% Corresponding author mailing address and e-mail address:
 
-# Optional: Bibliography 
-bibliography: lagosdepth
-biblio-style: jsta
+% (include name and email addresses of the corresponding author.  More
+% than one corresponding author is allowed in this LaTeX file and for
+% publication; but only one corresponding author is allowed in our
+% editorial system.)
 
-# Optional: Enable a 'Draft' watermark on the document
-watermark: true
+% Example: \correspondingauthor{First and Last Name}{email@address.edu}
 
-# Customize footer, eg by referencing the vignette
-footer_contents: "Geometric lake depth models and bathymetry"
+\correspondingauthor{J. Stachelek}{stachel2@msu.edu}
 
-# Produce a pinp document
-output: pinp::pinp
+%% Keypoints, final entry on title page.
 
-# Required: Vignette metadata for inclusion in a package.
-vignette: >
-  %\VignetteIndexEntry{YourPackage-vignetteentry}
-  %\VignetteKeywords{YourPackage, r, anotherkeyword}
-  %\VignettePackage{YourPackage}
-  %\VignetteEngine{knitr::rmarkdown}
----
+%  List up to three key points (at least one is required)
+%  Key Points summarize the main points and conclusions of the article
+%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
 
-## INTRODUCTION
+% Example:
+% \begin{keypoints}
+% \item	List up to three key points (at least one is required)
+% \item	Key Points summarize the main points and conclusions of the article
+% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
+% \end{keypoints}
 
+\begin{keypoints}
+\item Geometric models to predict lake depth, which require in-lake slope, assume that nearshore land slope is a good proxy for in-lake slope.
+\item Using data from thousands of lakes, we show that nearshore land slope is a poor proxy for in-lake slope and increases prediction error.
+\item Increases in prediction error were systematic such that depth was overpredicted in concave and reservoir lakes.
+\end{keypoints}
+
+%% ------------------------------------------------------------------------ %%
+%
+%  ABSTRACT and PLAIN LANGUAGE SUMMARY
+%
+% A good Abstract will begin with a short description of the problem
+% being addressed, briefly describe the new data or analyses, then
+% briefly states the main conclusion(s) and how they are supported and
+% uncertainties.
+
+% The Plain Language Summary should be written for a broad audience,
+% including journalists and the science-interested public, that will not have 
+% a background in your field.
+%
+% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
+% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
+% see http://sharingscience.agu.org/creating-plain-language-summary/)
+%
+%% ------------------------------------------------------------------------ %%
+
+%% \begin{abstract} starts the second page
+
+\begin{abstract}
+      Lake depth is a critical characteristic that influences many important
+      ecological processes in lakes. Unfortunately, lake depth measurements
+      are labor-intensive to gather and are only available for a small
+      fraction of lakes globally. Therefore, scientists have tried to predict
+      lake depth from characteristics that are easily obtained for all lakes
+      such as lake surface area or the slope of the land surrounding a lake.
+      One approach for predicting lake depth simulates lake basins using a
+      geometric model where nearshore land slope is assumed to be a
+      representative proxy for in-lake slope and the distance to the center of
+      the lake is assumed to be a representative proxy for the distance to the
+      deepest point of the lake. However, these assumptions have rarely been
+      tested in a broad range of lakes. We used bathymetry data from 5,000
+      lakes and reservoirs to test these assumptions and to examine whether
+      differences in lake type or shape influences depth prediction error. We
+      found that nearshore land slope was not a representative proxy of
+      in-lake slope and using it for prediction increases prediction error
+      substantially relative to models using true in-lake slope. We also found
+      that models using nearshore land slope as a proxy systematically
+      overpredicts lake depth in concave lakes (i.e.~bowl-shaped; up to 18\%
+      of lakes in the study population) and reservoir lakes (up to 30\% of
+      lakes in the study population), suggesting caution in using geometric
+      models for lake prediction in unsampled lakes.
+\end{abstract}
+
+%% ------------------------------------------------------------------------ %%
+%
+%  TEXT
+%
+%% ------------------------------------------------------------------------ %%
+
+%%% Suggested section heads:
+% \section{Introduction}
+%
+% The main text should start with an introduction. Except for short
+% manuscripts (such as comments and replies), the text should be divided
+% into sections, each with its own heading.
+
+% Headings should be sentence fragments and do not begin with a
+% lowercase letter or number. Examples of good headings are:
+
+% \section{Materials and Methods}
+% Here is text on Materials and Methods.
+%
+% \subsection{A descriptive heading about methods}
+% More about Methods.
+%
+% \section{Data} (Or section title might be a descriptive heading about data)
+%
+% \section{Results} (Or section title might be a descriptive heading about the
+% results)
+%
+% \section{Conclusions}
+
+\section{Introduction}
+\noindent
 Lake depth is a major factor controlling lake physics, chemistry, and biota
-\citep{liSignificantContributionLake2020, feeEffectsLakeSize1996, readSimulating2368Temperate2014}.
+\cite{liSignificantContributionLake2020, feeEffectsLakeSize1996, readSimulating2368Temperate2014}.
 However, because measured depth data is only available for a small fraction of all lakes, our ability to understand and predict depth-dependent processes is limited.
-The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \citep{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
+The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \cite{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
 The approaches taken in these studies have been varied.
-In addition, they have explored a variety of strategies to maximize prediction accuracy such as employing more diverse covariates \citep{oliver2016prediction}, varying lake buffer sizes \citep{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings among lakes \citep{sobekPredictingDepthVolume2011}.
+In addition, they have explored a variety of strategies to maximize prediction accuracy such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings among lakes \cite{sobekPredictingDepthVolume2011}.
 
-One approach for predicting lake depth involves a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \citep{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
+One approach for predicting lake depth involves a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \cite{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
 All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae.
@@
 geometric) relationship between slope and horizontal distance:
-\begin{align}
+
+\begin{linenomath*}
+\begin{equation}
       depth_{geometric} = tan(slope) * distance \label{eq1}
-\end{align}
+\end{equation}
+\label{eq1}
+\end{linenomath*}
 
-where the product of slope and horizontal distance yields an exact geometric depth estimate (depth_geometric).
+\noindent
+where the product of slope and horizontal distance yields an exact geometric depth estimate ($depth_{geometric}$).
 
-\begin{figure}[H]
-  \begin{center}
-    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/slope_diagram_new}
-  \end{center}
+\begin{figure}
+\noindent\includegraphics[width=0.85\textwidth]{../figures/slope_diagram_new}
 \caption{Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
@@
 in-lake) values derived from bathymetric maps.
-However, due to the scarcity of bathymetric map data, few studies have tested these assumptions except in individual studies of very large lakes or for small numbers of lakes (< 100) \citep{johanssonNewApproachesModelling2007}.
+However, due to the scarcity of bathymetric map data, few studies have tested these assumptions except in individual studies of very large lakes or for small numbers of lakes (< 100) \cite{johanssonNewApproachesModelling2007}.
 As a result, it is unclear what effect proxies have on lake depth predictions using geometric models and whether depth predictions are more sensitive to errors in the horizontal dimension (i.e.
@@
 Furthermore we expect to see greater overestimation error in reservoirs as compared to natural lakes because many reservoirs are known to be drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
-next to the dam) rather than in the center of the reservoir \citep{lanza1985interactions}.
+next to the dam) rather than in the center of the reservoir \cite{lanza1985interactions}.
 Similar to distance proxies, we expect that the use of a nearshore land slope proxy will lead to prediction error according to differences in cross-section shape (Figure \ref{fig2}) such that the depth of U-shaped (i.e.
@@
 convex) lakes will be underpredicted.
-Finally, we expect that offsets to geometric depth estimates will be most strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies \citep{oliver2016prediction}.
+Finally, we expect that offsets to geometric depth estimates will be most strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies \cite{oliver2016prediction}.
 
@@
       \begin{center}
-    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/lake_shape}
+        \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/lake_shape}
       \end{center}
@@
 
-## METHODS
-### Data description
+\section{Methods}
+\subsection{Data description}
+\noindent
 We compiled bathymetry data on approximately 5,000 lakes in the Northeastern and Midwestern US from nine official state databases (Figure S1).
@@
 For the Minnesota data, we simply clipped the raster for each lake to its outline.
-For data from the remaining states, we processed each lake by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the \texttt{raster} R package \citep{hijmansRasterGeographicData2019}.
+For data from the remaining states, we processed each lake by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the \texttt{raster} R package \cite{hijmansRasterGeographicData2019}.
 The size of the moving window was adjusted iteratively to ensure that each bathymetry raster contained no missing data.
 
-All lake bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) NHD \citep{usgsNationalHydrographyDataset2019} waterbodies such that source data and bathymetry surface outputs were clipped to the area of each lake polygon.
+All lake bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) NHD \cite{usgsNationalHydrographyDataset2019} waterbodies such that source data and bathymetry surface outputs were clipped to the area of each lake polygon.
 We restricted the lakes in our study to those with an area of at least 4 ha and a maximum depth of at least 0.3 m (1 ft).
@@
 its “visual distance to lake center”).
-For these calculations, we used the \texttt{polylabelr} R package \citep{larssonPolylabelrFindPole2019d}, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
-We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model (~15x15m grain) accessed using the \texttt{elevatr} R package \citep{hollisterElevatrAccessElevation2017} and the terrain function in the \texttt{raster} R package \citep{hijmansRasterGeographicData2019}.
+For these calculations, we used the \texttt{polylabelr} R package \cite{larssonPolylabelrFindPole2019d}, which interfaces with the Mapbox pole of inaccessibility algorithm \cite{agafonkinJSLibraryFinding2019}.
+We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model (~15x15m grain) accessed using the \texttt{elevatr} R package \cite{hollisterElevatrAccessElevation2017} and the terrain function in the \texttt{raster} R package \cite{hijmansRasterGeographicData2019}.
 
 We categorized lakes based on their cross-section shape and reservoir class.
-For cross-section shape, we categorized lakes as either convex or concave following the method of \citet{hakansonLakeFormLake1977} by computing normalized lake depth-area relationships (i.e.
+For cross-section shape, we categorized lakes as either convex or concave following the method of \citeA{hakansonLakeFormLake1977} by computing normalized lake depth-area relationships (i.e.
 hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S4).
-We further classified lakes using data from \citet{polus2020}, which uses the output of a machine learning algorithm to assign a probability to each lake regarding whether it is a reservoir or a natural lake.
+We further classified lakes using data from \citeA{polus2020}, which uses the output of a machine learning algorithm to assign a probability to each lake regarding whether it is a reservoir or a natural lake.
 For our purposes, we determined a lake to be a reservoir if the classification probability was 0.75 or greater.
-The \citet{polus2020} data product defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water, beyond simply controlling water level.
+The \citeA{polus2020} data product defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water, beyond simply controlling water level.
 It makes no distinction between different dam types or dam heights.
@@
 \begin{table}
+\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
+Predictor variables for computing random forest offsets (Eq \ref{eq2}) are printed in bold face.
+Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
 \centering
@@
   Nearshore land slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4850 (-)\\  
-  \bottomrule
+  \hline    
   \end{tabular}  
-  \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
-Predictor variables for computing random forest offsets (Eq \ref{eq2}) are printed in bold face.
-Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
 \end{table}
@@
 
-\begin{align}
-shoreline_{devel} = perimeter / (2 * \sqrt{(\pi * waterarea * 10000)})       
-\end{align}
+\begin{linenomath*}
+      \begin{equation}
+            shoreline_{devel} = perimeter / (2 * \sqrt{(\pi * waterarea * 10000)}) \label{eq2}        
+      \end{equation}
+      \end{linenomath*}
 
+\noindent
 where sinuous lakes have larger values of shoreline development and circular lakes have smaller values of shoreline development.
@@
 
-### Proxy evaluation
+\subsubsection{Proxy evaluation}
+\noindent
 We conducted a qualitative assessment of whether or not proxy measures of lake geometry are representative of their true values by visual inspection (i.e.
@@
 Our approach involved several steps.
-In the first step, we computed a geometric estimate of lake depth using only geometry information (depth_geometric, Equation \ref{eq1}).
+In the first step, we computed a geometric estimate of lake depth using only geometry information ($depth_{geometric}$, Equation \ref{eq1}).
 In the second step, we fit a random forest model to predict observed (i.e.
 true) depth as a function of geometric depth along with several covariates available for all lakes (Table \ref{table1}).
-The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of depth_geometric and among different lake types.
+The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of $depth_{geometric}$ and among different lake types.
 Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table \ref{table2}).
 
-## Model description
-### Geometric model 
+\subsection{Model description}
+\subsubsection{Geometric model}
+\noindent
 We used a geometric model of lakes where basins are treated as cones with a fixed relationship between slope and distance (Equation \ref{eq1}).
@@
 
-### Random forest models
+\subsubsection{Random forest models}
+\noindent
 Nearly all prior studies predicting lake depth using geometric models include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy.
-For our purposes, this offset modeling enabled us to test our expectations that prediction error would be different among different formulations of depth_geometric and among different lake types.
+For our purposes, this offset modeling enabled us to test our expectations that prediction error would be different among different formulations of $depth_{geometric}$ and among different lake types.
 It also facilitated direct comparison against prior models of lake depth including those that are non-geometric.
-We generated an “offset” to geometric depth \citep[sensu][]{hollisterPredictingMaximumLake2011} using the random forest algorithm and the \texttt{ranger} R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation \ref{eq1}) along with the lake elevation, area, perimeter, and ratio/index measures listed in Table 1:
+We generated an “offset” to geometric depth \cite<sensu>{hollisterPredictingMaximumLake2011} using the random forest algorithm and the \texttt{ranger} R package \cite{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation \ref{eq1}) along with the lake elevation, area, perimeter, and ratio/index measures listed in Table 1:
 
-\begin{align}
-  depth_{observed} \sim depth_{geometric} + covariates         
-\end{align}
+\begin{linenomath*}
+      \begin{equation}
+            depth_{observed} \sim depth_{geometric} + covariates \label{eq3}
+      \end{equation}
+      \end{linenomath*}
 
+\noindent      
 Neither cross-section shape nor reservoir class was used as a covariate in any random forest models.
 We used the random forest algorithm because it makes no assumptions about the distribution of model residuals, allows for non-linearity, and is insensitive to interactions (i.e.
-multicollinearity) among covariates \citep{prasadNewerClassificationRegression2006}.
+multicollinearity) among covariates \cite{prasadNewerClassificationRegression2006}.
 
-### Model comparisons
+\subsubsection{Model comparisons}
+\noindent
 We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
@@
 
-### Model evaluations
+\subsubsection{Model evaluations}
+\noindent
 We evaluated model fit and prediction error using root-mean-square error (RMSE) and coefficient of determination ($R^2$) metrics on a holdout set containing 25\% of all lakes.
@@
 
-## RESULTS
+\section{Results}
+\noindent
 Lakes belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Figure S1).
@@
   \begin{center}
-    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/01_geometry_base-1}
+    \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/01_geometry_base-1}
   \end{center}
@@
 \begin{table}
+  \caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table2}
   \centering
@@
   proxy & proxy & 6.6 m & 0.31\\
-  \bottomrule
+  \hline
 \end{tabular}
-\caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table2}
 \end{table}
 
-\begin{figure*}
+\begin{figure}
   \begin{center}
-    \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/02_depth_model_grid_resid-1}
+    \includegraphics[width=\textwidth,keepaspectratio]{../figures/02_depth_model_grid_resid-1}
   \end{center}
   \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig4}
-\end{figure*}
+\end{figure}
 
@@
 
-## DISCUSSION
-
+\section{Discussion}
+\noindent
 Our tests of geometric lake depth models show that specific proxy measures of lake geometry are not representative of true measures of lake geometry across a broad array of lakes.
@@
 
-### Representativeness of proxy measures of lake geometry
+\subsection{Representativeness of proxy measures of lake geometry}
+\noindent
 In comparing among lake geometry measures, our analysis suggests that proxy distance to lake center is representative of true distance to the deepest point of the lakes but that proxy nearshore land slope is not representative of true in-lake slope.
 A simple indication of this non representativeness is that proxy nearshore land slope was often (in > 74\% of cases) steeper than true in-lake slope.
-This finding is consistent with the suggestion by \citet{heathcotePredictingBathymetricFeatures2015} that in-lake slope matches nearshore land slope at the time of lake formation, but as lakes age, infilling via sedimentation causes them to develop shallower in-lake slopes compared to the surrounding land.
-If true, then these results also lend support to prior work \citep{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}, conducted on a more limited number of lakes, showing that in-lake slopes are controlled by erosion and sediment transport processes.
+This finding is consistent with the suggestion by \citeA{heathcotePredictingBathymetricFeatures2015} that in-lake slope matches nearshore land slope at the time of lake formation, but as lakes age, infilling via sedimentation causes them to develop shallower in-lake slopes compared to the surrounding land.
+If true, then these results also lend support to prior work \cite{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}, conducted on a more limited number of lakes, showing that in-lake slopes are controlled by erosion and sediment transport processes.
 
@@
 This is contrary to the idea that most reservoirs are drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
-next to the dam) rather than in the center of the reservoir \citep{lanza1985interactions}.
-One possible explanation is that the reservoir classification data from \citet{polus2020} uses a more general definition of a reservoir (i.e.
+next to the dam) rather than in the center of the reservoir \cite{lanza1985interactions}.
+One possible explanation is that the reservoir classification data from \citeA{polus2020} uses a more general definition of a reservoir (i.e.
 any permanent waterbody that has a water control structure likely to significantly impact flow or pool water) compared to that of conventional classifications that are tied to specific dam types or dam heights.
@@
 It was notable that convex lakes were deeper than concave lakes despite having similar distributions of lake surface area (Figure S3).
-Given the similarity in lake surface area, the underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of lake formation whereby concave lakes were formed as a result of more intense glacial scouring compared to convex lakes \citep{gorhamPhysicalLimnologyNorthern1958}.
-While there is some evidence in support of this idea, namely that there is a geographic hotspot of concave lakes associated with the glaciated “prairie pothole region” \citep[see][]{hayashiSimpleEquationsRepresent2000}, the overall geographic distribution of lake cross-section shapes does not support this idea.
+Given the similarity in lake surface area, the underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of lake formation whereby concave lakes were formed as a result of more intense glacial scouring compared to convex lakes \cite{gorhamPhysicalLimnologyNorthern1958}.
+While there is some evidence in support of this idea, namely that there is a geographic hotspot of concave lakes associated with the glaciated “prairie pothole region” \cite<see>{hayashiSimpleEquationsRepresent2000}, the overall geographic distribution of lake cross-section shapes does not support this idea.
 Instead of a concentrated area of concave lakes in formerly glaciated regions, there appears to be a fairly even mix of concave and convex lakes distributed amongst the northern (i.e.
@@
 
-### Effects of proxy measures of lake geometry depth prediction error
-Models using only proxy variables had prediction error rates (RMSE = 6.6m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting lake depth at broad geographic extents \citep{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
+\subsection{Effects of proxy measures of lake geometry depth prediction error}
+\noindent
+Models using only proxy variables had prediction error rates (RMSE = 6.6m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting lake depth at broad geographic extents \cite{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
 When only a single proxy measure was used there was a difference in model sensitivity depending on if it was a horizontal distance measure or a vertical slope measure.
-In the case of a true slope and proxy distance combination, models were more accurate ($\pm$ 4.2m) than even the most accurate of prior studies \citep{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
+In the case of a true slope and proxy distance combination, models were more accurate ($\pm$ 4.2m) than even the most accurate of prior studies \cite{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
 Conversely, models using a proxy slope and true distance combination had prediction error rates ($\pm$ 6.9m) of a similar magnitude as that of the baseline proxy-proxy model ($\pm$ 6.6m).
 The greater sensitivity of depth predictions to proxy slope measures relative to proxy distance measures may be explained by the fact that proxy slope measures were a more imperfect representation of true in-lake slopes relative to proxy versus true distances.
-In addition, these results help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \citep{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S3).
+In addition, these results help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \cite{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S3).
 
-### Effects of lake shape and lake type on depth prediction error
+\subsection{Effects of lake shape and lake type on depth prediction error}
+\noindent
 As expected, we found that the maximum depth of concave lakes was systematically overpredicted by a simple geometric model using proxy nearshore land slope.
@@
 
-### Future research
+\subsection{Future research}
+\noindent
 One of our models (true slope, proxy distance) was more accurate than even the most accurate of prior studies.
@@
 This would allow for more robust tests of the representativeness of geometric model inputs.
-Unfortunately, topobathymetric products are rare, have mostly been limited nearshore marine environments, and as such are not widely available for inland waters \citep{danielsonTopobathymetricElevationModel2016}.
+Unfortunately, topobathymetric products are rare, have mostly been limited nearshore marine environments, and as such are not widely available for inland waters \cite{danielsonTopobathymetricElevationModel2016}.
 
@@
 Unfortunately, identifying a lake’s cross-section shape requires bathymetry data which is unavailable for most lakes.
-However, given the conceptual links between cross-section shape, glaciation, and sedimentation \citep{johanssonNewApproachesModelling2007} it may be advantageous for future studies to compile data on sedimentation to determine if this data can be used to predict cross-section shape and boost depth prediction accuracy.
+However, given the conceptual links between cross-section shape, glaciation, and sedimentation \cite{johanssonNewApproachesModelling2007} it may be advantageous for future studies to compile data on sedimentation to determine if this data can be used to predict cross-section shape and boost depth prediction accuracy.
+
+%%
+
+%% Enter Figures and Tables near as possible to where they are first mentioned:
+%
+% DO NOT USE \psfrag or \subfigure commands.
+%
+% Figure captions go below the figure.
+% Table titles go above tables;  other caption information
+%  should be placed in last line of the table, using
+% \multicolumn2l{$^a$ This is a table note.}
+%
+%----------------
+% EXAMPLE FIGURES
+%
+% \begin{figure}
+% \includegraphics{example.png}
+% \caption{caption}
+% \end{figure}
+%
+% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
+% \begin{figure}
+% \noindent\includegraphics[width=\textwidth]{anothersample.png}
+%\caption{caption}
+%\label{pngfiguresample}
+%\end{figure}
+%
+%
+% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
+This is common when trying to add a PDF figure without pdflatex.
+% \begin{figure}
+% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
+%\caption{caption}
+%\label{pdffiguresample}
+%\end{figure}
+%
+%
+% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
+%
+
+%
+% ---------------
+% EXAMPLE TABLE
+%
+% \begin{table}
+% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
+% \centering
+% \begin{tabular}{l c}
+% \hline
+%  Run  & Time (min)  \\
+% \hline
+%   $l1$  & 260   \\
+%   $l2$  & 300   \\
+%   $l3$  & 340   \\
+%   $h1$  & 270   \\
+%   $h2$  & 250   \\
+%   $h3$  & 380   \\
+%   $r1$  & 370   \\
+%   $r2$  & 390   \\
+% \hline
+% \multicolumn{2}{l}{$^{a}$Footnote text here.}
+% \end{tabular}
+% \end{table}
+
+%% SIDEWAYS FIGURE and TABLE
+% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
+%
+% \begin{sidewaysfigure}
+% \includegraphics[width=20pc]{figsamp}
+% \caption{caption here}
+% \label{newfig}
+% \end{sidewaysfigure}
+%
+%  \begin{sidewaystable}
+%  \caption{Caption here}
+% \label{tab:signif_gap_clos}
+%  \begin{tabular}{ccc}
+% one&two&three\\
+% four&five&six
+%  \end{tabular}
+%  \end{sidewaystable}
+
+%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
+%\begin{linenomath*}
+%\begin{equation}
+%y|{f} \sim g(m, \sigma),
+%\end{equation}
+%\end{linenomath*}
+
+%%% End of body of article
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional Appendix goes here
+%
+% The \appendix command resets counters and redefines section heads
+%
+% After typing \appendix
+%
+%\section{Here Is Appendix Title}
+% will show
+% A: Here Is Appendix Title
+%
+%\appendix
+%\section{Here is a sample appendix}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Optional Glossary, Notation or Acronym section goes here:
+%
+%%%%%%%%%%%%%%
+% Glossary is only allowed in Reviews of Geophysics
+%  \begin{glossary}
+%  \term{Term}
+%   Term Definition here
+%  \term{Term}
+%   Term Definition here
+%  \term{Term}
+%   Term Definition here
+%  \end{glossary}
+
+%
+%%%%%%%%%%%%%%
+% Acronyms
+%   \begin{acronyms}
+%   \acro{Acronym}
+%   Definition here
+%   \acro{EMOS}
+%   Ensemble model output statistics
+%   \acro{ECMWF}
+%   Centre for Medium-Range Weather Forecasts
+%   \end{acronyms}
+
+%
+%%%%%%%%%%%%%%
+% Notation
+%   \begin{notation}
+%   \notation{$a+b$} Notation Definition here
+%   \notation{$e=mc^2$}
+%   Equation in German-born physicist Albert Einstein's theory of special
+%  relativity that showed that the increased relativistic mass ($m$) of a
+%  body comes from the energy of motion of the body—that is, its kinetic
+%  energy ($E$)—divided by the speed of light squared ($c^2$).
+%   \end{notation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%  ACKNOWLEDGMENTS
+%
+% The acknowledgments must list:
+%
+% >>>>	A statement that indicates to the reader where the data
+% 	supporting the conclusions can be obtained (for example, in the
+% 	references, tables, supporting information, and other databases).
+%
+% 	All funding sources related to this work from all authors
+%
+% 	Any real or perceived financial conflicts of interests for any
+%	author
+%
+% 	Other affiliations for any author that may be perceived as
+% 	having a conflict of interest with respect to the results of this
+% 	paper.
+%
+%
+% It is also the appropriate place to thank colleagues and other contributors.
+% AGU does not normally allow dedications.
+
+\acknowledgments
+Funding was provided by the US NSF Macrosystems Biology Program grants, DEB-1638679; DEB-1638550, DEB-1638539, DEB-1638554.
+PAS was also supported by USDA National Institute of Food and Agriculture Hatch Project, Grant Number: 176820.
+Author contributions: JS conceived of the study, built models, analyzed data, and wrote the paper.
+PH and PAS provided interpretation of results and edited the paper.
+
+%% ------------------------------------------------------------------------ %%
+%% References and Citations
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% \bibliography{<name of your .bib file>} don't specify the file extension
+%
+% don't specify bibliographystyle
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\bibliography{lagosdepth}
+
+%Reference citation instructions and examples:
+%
+% Please use ONLY \cite and \citeA for reference citations.
+% \cite for parenthetical references
+% ...as shown in recent studies (Simpson et al., 2019)
+% \citeA for in-text citations
+% ...Simpson et al. (2019) have shown...
+%
+%
+%...as shown by \citeA{jskilby}.
+%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
+%...has been shown \cite{jskilbye}.
+%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
+%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
+%...has been shown by \cite <e.g.,>[and others]{lewin76}.
+%
+% apacite uses < > for prenotes and [ ] for postnotes
+% DO NOT use other cite commands (e.g., \citet, \cite, \citeyear, \nocite, \citealp, etc.).
+%
+
+\end{document}
+
+More Information and Advice:
+
+%% ------------------------------------------------------------------------ %%
+%
+%  SECTION HEADS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Capitalize the first letter of each word (except for
+% prepositions, conjunctions, and articles that are
+% three or fewer letters).
+
+% AGU follows standard outline style; therefore, there cannot be a section 1 without
+% a section 2, or a section 2.3.1 without a section 2.3.2.
+% Please make sure your section numbers are balanced.
+% ---------------
+% Level 1 head
+%
+% Use the \section{} command to identify level 1 heads;
+% type the appropriate head wording between the curly
+% brackets, as shown below.
+%
+%An example:
+%\section{Level 1 Head: Introduction}
+%
+% ---------------
+% Level 2 head
+%
+% Use the \subsection{} command to identify level 2 heads.
+%An example:
+%\subsection{Level 2 Head}
+%
+% ---------------
+% Level 3 head
+%
+% Use the \subsubsection{} command to identify level 3 heads
+%An example:
+%\subsubsection{Level 3 Head}
+%
+%---------------
+% Level 4 head
+%
+% Use the \subsubsubsection{} command to identify level 3 heads
+% An example:
+%\subsubsubsection{Level 4 Head} An example.
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  IN-TEXT LISTS
+%
+%% ------------------------------------------------------------------------ %%
+%
+% Do not use bulleted lists; enumerated lists are okay.
+% \begin{enumerate}
+% \item
+% \item
+% \item
+% \end{enumerate}
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  EQUATIONS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Single-line equations are centered.
+% Equation arrays will appear left-aligned.
+
+Math coded inside display math mode \[ ...\]
+ will not be numbered, e.g.,:
+ \[ x^2=y^2 + z^2\]
+
+ Math coded inside \begin{equation} and \end{equation} will
+ be automatically numbered, e.g.,:
+ \begin{equation}
+ x^2=y^2 + z^2
+ \end{equation}
+
+% To create multiline equations, use the
+% \begin{eqnarray} and \end{eqnarray} environment
+% as demonstrated below.
+\begin{eqnarray}
+  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
+        && + (y - y_{0}) \sin \Theta  \nonumber \\
+  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
+        && + (y - y_{0}) \cos \Theta.
+\end{eqnarray}
+
+%If you don't want an equation number, use the star form:
+%\begin{eqnarray*}...\end{eqnarray*}
+
+% Break each line at a sign of operation
+% (+, -, etc.) if possible, with the sign of operation
+% on the new line.
+
+% Indent second and subsequent lines to align with
+% the first character following the equal sign on the
+% first line.
+
+% Use an \hspace{} command to insert horizontal space
+% into your equation if necessary. Place an appropriate
+% unit of measure between the curly braces, e.g.
+% \hspace{1in}; you may have to experiment to achieve
+% the correct amount of space.
+
+%% ------------------------------------------------------------------------ %%
+%
+%  EQUATION NUMBERING: COUNTER
+%
+%% ------------------------------------------------------------------------ %%
+
+% You may change equation numbering by resetting
+% the equation counter or by explicitly numbering
+% an equation.
+
+% To explicitly number an equation, type \eqnum{}
+% (with the desired number between the brackets)
+% after the \begin{equation} or \begin{eqnarray}
+% command.  The \eqnum{} command will affect only
+% the equation it appears with; LaTeX will number
+% any equations appearing later in the manuscript
+% according to the equation counter.
+%
+
+% If you have a multiline equation that needs only
+% one equation number, use a \nonumber command in
+% front of the double backslashes (\\) as shown in
+% the multiline equation above.
+
+% If you are using line numbers, remember to surround
+% equations with \begin{linenomath*}...\end{linenomath*}
+
+%  To add line numbers to lines in equations:
+%  \begin{linenomath*}
+%  \begin{equation}
+%  \end{equation}
+%  \end{linenomath*}

### 2020-06-09 manuscript/manuscript.Rmd
copy over to wrr template

@@
 
-where the product of slope and horizontal distance yields an exact geometric depth estimate (depth_geometric).
+where the product of slope and horizontal distance yields an exact geometric depth estimate ($depth_{geometric}$).
 
@@
   \end{tabular}  
-  \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
-Predictor variables for computing random forest offsets (Eq \ref{eq2}) are printed in bold face.
-Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
 \end{table}
@@
 Our approach involved several steps.
-In the first step, we computed a geometric estimate of lake depth using only geometry information (depth_geometric, Equation \ref{eq1}).
+In the first step, we computed a geometric estimate of lake depth using only geometry information ($depth_{geometric}$, Equation \ref{eq1}).
 In the second step, we fit a random forest model to predict observed (i.e.
 true) depth as a function of geometric depth along with several covariates available for all lakes (Table \ref{table1}).
-The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of depth_geometric and among different lake types.
+The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of $depth_{geometric}$ and among different lake types.
 Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table \ref{table2}).
@@
 Nearly all prior studies predicting lake depth using geometric models include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy.
-For our purposes, this offset modeling enabled us to test our expectations that prediction error would be different among different formulations of depth_geometric and among different lake types.
+For our purposes, this offset modeling enabled us to test our expectations that prediction error would be different among different formulations of $depth_{geometric}$ and among different lake types.
 It also facilitated direct comparison against prior models of lake depth including those that are non-geometric.
@@
 \end{tabular}
-\caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table2}
 \end{table}

### 2020-06-18 figures/01_geometry.Rmd
fix visual center distance labels

@@
   xlab("Deepest point distance (m)") +
-  ylab("Visual center\ndistance (m)")
+  ylab("Distance to lake center (m)")
 
@@
   xlab("Deepest point distance (m)") +
-  ylab("Visual center\ndistance (m)")
+  ylab("Distance to lake center (m)")
 ```
@@
 # get_pnt_map(which(dt$llid == 257862))
-cowplot::plot_grid(get_pnt_map(which(dt$llid == 3791)) +
-                     ggtitle(3791), 
-                   get_pnt_map(which(dt$llid == 2772)) +
-                     ggtitle(2772), 
-                   get_pnt_map(which(dt$llid == 2655)) +
-                     ggtitle(2655), 
-                   get_pnt_map(which(dt$llid == 1822)) +
-                     ggtitle(1822)
-                   )
+# cowplot::plot_grid(get_pnt_map(which(dt$llid == 3791)) +
+#                      ggtitle(3791), 
+#                    get_pnt_map(which(dt$llid == 2772)) +
+#                      ggtitle(2772), 
+#                    get_pnt_map(which(dt$llid == 2655)) +
+#                      ggtitle(2655), 
+#                    get_pnt_map(which(dt$llid == 1822)) +
+#                      ggtitle(1822)
+#                    )
 
@@
   xlab("Deepest point distance (m)") +
-  ylab("Visual center\ndistance (m)") +
+  ylab("Distance to lake center (m)") +
   xlim(0, 3000) +

### 2020-06-18 figures/01_geometry.Rmd
align geometry grid

@@
   xlab("Deepest point distance (m)") +
-  ylab("Distance to lake center (m)")
+  ylab("Distance to \n lake center (m)")
 
@@
   xlab("Deepest point distance (m)") +
-  ylab("Distance to lake center (m)")
+  ylab("Distance to \n lake center (m)")
 ```
@@
   xlab("Deepest point distance (m)") +
-  ylab("Distance to lake center (m)") +
+  ylab("Distance to \n lake center (m)") +
   xlim(0, 3000) +

### 2020-06-22 figures/01_geometry.Rmd
add dashes to some remaining inlake tags

@@
   xlab("Nearshore slope (mean)") +
-  ylab("Inlake slope") +
+  ylab("In-lake slope") +
   labs(color = "")
@@
   xlab("Nearshore slope (mean)") +
-  ylab("Inlake slope") +
+  ylab("In-lake slope") +
   labs(color = "") +

### 2020-06-22 manuscript/appendix.Rmd
add dashes to some remaining inlake tags

@@
 
-```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.
+```{r 01_geometry_grid, fig.cap = "Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake center and C-D) nearshore slope versus inlake slope.
 A dashed 1:1 line is shown for comparison.
@@
 
-```{r metric_correlations, fig.cap = "Pearson correlation coefficients of GIS-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=FALSE, out.width="11cm", fig.align='center', out.extra=''}
+```{r metric_correlations, fig.cap = "Pearson correlation coefficients of proxy geometry metrics against in-lake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=FALSE, out.width="11cm", fig.align='center', out.extra=''}
 knitr::include_graphics("../figures/metric_correlations-1.pdf")

### 2020-06-23 figures/01_geometry.Rmd
rm white background in spacers and legend

@@
         axis.title = element_text(size = 11))
-gg_spacer <- ggplot() + geom_blank() + theme(panel.background = element_blank())
+gg_spacer <- ggplot() + geom_blank() + theme(panel.background = element_rect(fill = NA), plot.background = element_rect(fill = NA))
 
 legend_shape     <- cowplot::get_legend(gg_scatter_shape +
-                                          theme(legend.direction = "horizontal"))
+                                          theme(legend.direction = "horizontal", legend.background = element_rect(fill = NA)))
 legend_reservoir <- cowplot::get_legend(gg_scatter_reservoir + labs(color = "") +
-                                          theme(legend.direction = "horizontal"))
+                                          theme(legend.direction = "horizontal", legend.background = element_rect(fill = NA)))
 legend_panel     <- cowplot::plot_grid(gg_spacer, legend_shape, 

### 2020-07-07 scripts/06_stats.R
calculate johansson lake area stats

@@ -25,6 +25,14 @@ quantile(hak_areas,
          probs = c(0.25, 0.5, 0.75))
 hist(hak_areas, n = 80)
 
+# ---- What are the summary stats of the lakes used in Johansson et al. (2007)? ----
+# library(digitize)
+# digitize("~/Pictures/Screenshot from 2020-07-07 16-00-54.png")
+jhs <- data.frame(area = c(0.055, 0.55, 5.5, 55, 5500) * 10e6,
+                  count = c(12, 51, 38, 3, 2), stringsAsFactors = FALSE)
+jhs$area * 0.0001
+# most are > 500 ha
+
 # ---- How related are true and proxy geometry metrics? ----
 bp <- readRDS("data/00_bathy_depth/bathy_pnts.rds")
 dt <- read.csv("data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%

### 2020-07-07 figures/01_contrasts.Rmd
start bathy contrast panel

@@
 ```
+
+```{r gg_effort, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.5}
+# compare characteristics of bathy lakes versus overall lakes
+
+# load LAGOSUS DEPTH
+lg_raw <- lagosus_load("depth")$depth$depth
+lg <- lg_raw %>%
+  mutate(lagos_effort = 
+           forcats::fct_collapse(lg_raw$lagos_effort, 
+                                 bathymetry = "bathymetry", 
+                                 other = c("LAGOSNE", "manual", "NLA")))
+
+# jsta::gg_quantdot
+gg_quantdot <- function(dt, grp, var){
+  # dt <- mtcars; grp <- "cyl"; var <- "mpg"
+
+  # https://tbradley1013.github.io/2018/10/01/calculating-quantiles-for-groups-with-dplyr-summarize-and-purrr-partial/
+  p       <- c(0.05, 0.5, 0.95)
+  p_names <- paste0(as.character(p*100), "%")
+  p_funs  <- lapply(seq_along(p), function(x){
+    function(...){quantile(probs = p[x], na.rm = TRUE, ...)}
+    })
+  p_funs <- setNames(p_funs, p_names)
+
+  dt %>%
+    group_by({{grp}}) %>%
+    summarise_at({{var}}, p_funs) %>% 
+    ggplot() + 
+    geom_pointrange(aes(x = {{grp}}, y = .data$`50%`,
+                      ymin = .data$`5%`, ymax = .data$`95%`)) +
+    ylab(var)
+}
+
+cowplot::plot_grid(
+  gg_quantdot(lg, lagos_effort, "lake_maxdepth_m"), 
+  gg_quantdot(lg, lagos_effort, "lake_waterarea_ha") +
+    scale_y_log10(), 
+  ncol = 1)
+
+# library(tigris)
+# urbanareas <- tigris::urban_areas()
+# urbanareas <- st_as_sf(urbanareas)
+# urbanareas <- st_transform(urbanareas, LAGOSUSgis:::albers_conic())
+# urbanareas <- tidyr::separate(urbanareas, NAME10,
+#                               c("urban_name", "urban_state"), sep = ", ")
+# urbanareas <- dplyr::filter(urbanareas, !(urban_state %in% c("AK", "HI")))
+# saveRDS(urbanareas, "../data/urbanareas.rds")
+urbanareas <- readRDS("../data/urbanareas.rds")
+
+test <- apply(st_distance(LAGOSUS::coordinatize(lg, crs = LAGOSUSgis:::albers_conic()),
+            urbanareas), 1, function(x) x[which.min(x)])
+```

### 2020-07-08 figures/01_contrasts.Rmd
finish a working bathy representativness fig

--- 2020-06-01 figures/01_contrasts.Rmd
@@
 ```
+
+```{r gg_effort, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.5}
+# compare characteristics of bathy lakes versus overall lakes
+
+# load LAGOSUS DEPTH
+lg_raw <- lagosus_load("depth")$depth$depth
+lg <- lg_raw %>%
+  mutate(lagos_effort = 
+           forcats::fct_collapse(lg_raw$lagos_effort, 
+                                 bathymetry = "bathymetry", 
+                                 other = c("LAGOSNE", "manual", "NLA")))
+
+# library(tigris)
+# urbanareas <- tigris::urban_areas()
+# urbanareas <- st_as_sf(urbanareas)
+# urbanareas <- st_transform(urbanareas, LAGOSUSgis:::albers_conic())
+# urbanareas <- tidyr::separate(urbanareas, NAME10,
+#                               c("urban_name", "urban_state"), sep = ", ")
+# urbanareas <- dplyr::filter(urbanareas, !(urban_state %in% c("AK", "HI")))
+# urbanareas <- urbanareas[st_area(urbanareas) > 
+#                      units::as_units(9500000, "m2"),]
+# for(i in seq_len(nrow(urbanareas))){
+#   print(i)
+#   urbanareas[i,] <- rmapshaper::ms_simplify(urbanareas[1,])
+# }
+# saveRDS(urbanareas, "../data/urbanareas.rds")
+urbanareas <- readRDS("../data/urbanareas.rds")
+
+lg$urban_distance <- apply(
+  st_distance(
+    st_transform(
+      LAGOSUS::coordinatize(lg, crs = 4326), 
+      LAGOSUSgis:::albers_conic()),
+    urbanareas), 
+  1, function(x) x[which.min(x)])
+
+cowplot::plot_grid(
+  gg_quantdot(lg, lagos_effort, "lake_maxdepth_m"), 
+  gg_quantdot(lg, lagos_effort, "lake_waterarea_ha") +
+    scale_y_log10(), 
+  gg_quantdot(lg, lagos_effort, "urban_distance"),
+  ncol = 1)
+```

### 2020-07-08 scripts/99_utils.R

@@ -333,3 +333,24 @@ jheatmap <- function(dt, focal_columns = NULL) {
     pheatmap::pheatmap(t(test), color = hmap_cols)
   }
 }
+
+# jsta::gg_quantdot
+gg_quantdot <- function(dt, grp, var){
+  # dt <- mtcars; grp <- "cyl"; var <- "mpg"
+
+  # https://tbradley1013.github.io/2018/10/01/calculating-quantiles-for-groups-with-dplyr-summarize-and-purrr-partial/
+  p       <- c(0.05, 0.5, 0.95)
+  p_names <- paste0(as.character(p*100), "%")
+  p_funs  <- lapply(seq_along(p), function(x){
+    function(...){quantile(probs = p[x], na.rm = TRUE, ...)}
+  })
+  p_funs <- setNames(p_funs, p_names)
+
+  dt %>%
+    group_by({{grp}}) %>%
+    summarise_at({{var}}, p_funs) %>%
+    ggplot() +
+    geom_pointrange(aes(x = {{grp}}, y = .data$`50%`,
+                        ymin = .data$`5%`, ymax = .data$`95%`)) +
+    ylab(var)
+}

### 2020-07-08 figures/01_contrasts.Rmd
finish out bathy comparison fig, closes #12

@@
 
-```{r gg_effort, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.5}
+```{r gg_effort, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.7, fig.width=2.7}
 # compare characteristics of bathy lakes versus overall lakes
@@
   1, function(x) x[which.min(x)])
+lg$urban_distance <- lg$urban_distance * 0.001 # m to km
 
+theme_opts <- theme(axis.title.x = element_blank())
 cowplot::plot_grid(
-  gg_quantdot(lg, lagos_effort, "lake_maxdepth_m"), 
+  gg_quantdot(lg, lagos_effort, "lake_maxdepth_m") +
+    ylab("Max depth (m) \n") +
+    theme_opts +
+    theme(axis.text.x = element_blank()), 
   gg_quantdot(lg, lagos_effort, "lake_waterarea_ha") +
-    scale_y_log10(), 
-  gg_quantdot(lg, lagos_effort, "urban_distance"),
-  ncol = 1)
+    scale_y_log10() +
+    ylab("Area (ha)") +
+    theme_opts +
+    theme(axis.text.x = element_blank()), 
+  gg_quantdot(lg, lagos_effort, "urban_distance") +
+    scale_y_log10() +
+    ylab("Distance to urban (km)") +
+    theme_opts,
+  rel_heights = c(0.85, 0.85, 1),  ncol = 1)
 ```

### 2020-07-08 manuscript/appendix.Rmd
finish out bathy comparison fig, closes #12

@@
 
+```{r gg_effort, fig.cap = "Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources.
+Distance to urban areas uses data from the 2018 US Census Urban and Rural Classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/gg_effort-1.pdf")
+```
+
 ```{r metric_correlations, fig.cap = "Pearson correlation coefficients of proxy geometry metrics against in-lake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=FALSE, out.width="11cm", fig.align='center', out.extra=''}

### 2020-07-16 manuscript/agujournaltemplate.tex
stash pre dissertation edit text
sync ms text with dissertation edits

@@
 
-\authors{J. Stachelek\affil{1}, P. Hanly\affil{1}, and P. A. Soranno\affil{1}}
+\authors{J. Stachelek\affil{1}, P. J. Hanly\affil{1}, and P. A. Soranno\affil{1}}
 
@@
       that models using nearshore land slope as a proxy systematically
-      overpredicts lake depth in concave lakes (i.e.~bowl-shaped; up to 18\%
+      overpredicts lake depth in concave lakes (i.e. bowl-shaped; up to 18\%
       of lakes in the study population) and reservoir lakes (up to 30\% of
       lakes in the study population), suggesting caution in using geometric
-      models for lake prediction in unsampled lakes.
+      models for depth prediction in unsampled lakes.
 \end{abstract}
@@
 \noindent
-Lake depth is a major factor controlling lake physics, chemistry, and biota
-\cite{liSignificantContributionLake2020, feeEffectsLakeSize1996, readSimulating2368Temperate2014}.
-However, because measured depth data is only available for a small fraction of all lakes, our ability to understand and predict depth-dependent processes is limited.
+\noindent
+Lake depth is an important factor controlling lake physics, chemistry, and biota.
+Deeper lakes generally have higher water clarity and less complete mixing compared to shallow lakes \cite{feeEffectsLakeSize1996, readSimulating2368Temperate2014}.
+These differences are reflected in variation among lakes in terms of biological productivity \cite{qinWaterDepthUnderpins2020} and rates of greenhouse gas production \cite{liSignificantContributionLake2020}.
+However, because measured depth data is only available for a small fraction ($\sim$25\%) of all lakes, our ability to understand and predict depth-dependent processes is limited.
 The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \cite{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
-The approaches taken in these studies have been varied.
-In addition, they have explored a variety of strategies to maximize prediction accuracy such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings among lakes \cite{sobekPredictingDepthVolume2011}.
+Given the limited prediction accuracy of prior attempts ($\pm$ 6-7 m), studies have explored strategies such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings (e.g.
+fitting different models for distinct size classes) among lakes \cite{sobekPredictingDepthVolume2011}.
 
-One approach for predicting lake depth involves a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \cite{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
+One approach for predicting lake depth involves using a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \cite{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
 All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae.
@@
 \end{equation}
-\label{eq1}
 \end{linenomath*}
@@
 
-The assumptions of this model as well as other geometric models can be tested by comparing proxy measures of lake geometry against corresponding “true” (i.e.
-in-lake) values derived from bathymetric maps.
-However, due to the scarcity of bathymetric map data, few studies have tested these assumptions except in individual studies of very large lakes or for small numbers of lakes (< 100) \cite{johanssonNewApproachesModelling2007}.
-As a result, it is unclear what effect proxies have on lake depth predictions using geometric models and whether depth predictions are more sensitive to errors in the horizontal dimension (i.e.
-distance to the deepest point of the lake) or errors in the vertical dimension (i.e.
+The assumptions of the cone model (as well as other geometric models) can be tested by comparing proxy measures of lake geometry against corresponding “true” (i.e.
+in-lake) values derived from bathymetric maps and by evaluating how lake cross-section shapes differ from that of an idealized cone \cite{johanssonNewApproachesModelling2007}.
+For instance, lake cross-section shapes have been shown to vary from narrow "convex" forms to outstretched "concave" forms \cite{hakansonLakeFormLake1977}.
+Because tests of geometric model assumptions require bathymetric map data, which is only available for about 15\% of all lakes, existing evidence may not be applicable to all lakes.
+The few studies that have tested these assumptions have mostly been limited to individual studies of very large ($>$ 500 ha) lakes or studies on small numbers ($<$ 100) of lakes \cite{johanssonNewApproachesModelling2007}.
+Studies focused specifically on reservoirs (as opposed to the more typical case where reservoirs and natural lakes are combined), have been even more restricted to that of extremely large lakes $>$ 1000 ha \cite{lehnerHighresolutionMappingWorld2011, messagerEstimatingVolumeAge2016}.
+
+As a result of this limited testing, we lack knowledge on both the predictive performance of geometric models, the effect of proxies on depth prediction, and whether depth predictions are more sensitive to measurement errors in the horizontal dimension (i.e.
+distance to the deepest point of the lake) or measurement errors in the vertical dimension (i.e.
 in-lake slope).
 Additionally it is unclear whether model prediction error is related to differences in lake type such those with different cross-section shapes or those classified as reservoirs versus natural lakes.
-This information could help direct future research efforts to focus on particular dimensions of lake geometry (i.e.
-horizontal versus vertical) or to stratify model predictions based on specific lake types and cross-section shapes.
 
-Given the knowledge gaps identified above, we asked three research questions: 1) Is nearshore land slope representative of in-lake slope; and is the distance to the center of a lake representative of the distance to the deepest point of a lake?
+Given the knowledge gaps identified above, we asked three research questions: 1) How representative is nearshore land slope of in-lake slope; and how representative is the distance to the center of a lake compared to the distance to the deepest point of a lake?
 2) How does the use of proxies for lake geometry affect lake depth prediction error?
-3) How does lake shape (i.e.
+3) How does lake cross-section shape (i.e.
 concave versus convex) and lake type (i.e.
 natural lake vs reservoir) affect depth prediction error?
-
-To answer these questions, we extracted the maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 lakes in the Northeastern and Midwestern US from bathymetric map data.
+To answer these questions, we extracted maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, cross-section shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 lakes in the Northeastern and Midwestern US from bathymetric map data and supplemented this data with classification estimates of whether lakes are reservoirs or natural lakes.
 First, we examined whether measures of lake geometry (in-lake slope and distance to the deepest point of lakes) were related to geometry proxies (nearshore land slope and distance to lake centers).
-Next, we computed geometric depth estimates (Equation \ref{eq1}) and offsets to these estimates using the random forest algorithm (Equation \ref{eq2}).
+Next, we computed geometric depth estimates (Equation \ref{eq1}) and prediction "offsets" to these estimates using the random forest algorithm (Equation \ref{eq2}).
 Covariates used in offset modeling included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table \ref{table1}).
@@
 
-Given that, by definition, the distance proxy (distance to the center of the lake) must always be greater or equal to the true distance value (distance to the deepest point of the lake), we expect that the use of this proxy will lead to overestimation of lake depth (Figure \ref{fig1}).
+Given that, by definition, the distance proxy (distance to the center of the lake) must always be greater or equal to the true distance value (distance to the deepest point of the lake), we expect that the use of this proxy will lead to overestimation of lake depth (Figure \ref{fig3}).
 Furthermore we expect to see greater overestimation error in reservoirs as compared to natural lakes because many reservoirs are known to be drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
 next to the dam) rather than in the center of the reservoir \cite{lanza1985interactions}.
-Similar to distance proxies, we expect that the use of a nearshore land slope proxy will lead to prediction error according to differences in cross-section shape (Figure \ref{fig2}) such that the depth of U-shaped (i.e.
+In a similar fashion, we expect to see overestimation error associated with using a nearshore land slope proxy in lakes with differing cross-section shape (Figure \ref{fig2}) such that the depth of U-shaped (i.e.
 concave) lakes will be overpredicted whereas the depth of V-shaped (i.e.
 convex) lakes will be underpredicted.
-Finally, we expect that offsets to geometric depth estimates will be most strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies \cite{oliver2016prediction}.
+Finally, we expect that overall depth prediction will be strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies \cite{oliver2016prediction}.
+
+By testing these expectations, we can establish whether barriers to increased depth prediction accuracy lie in lack of correspondence between true and proxy measures of lake geometry or in hidden grouping (such as lake cross-section shape or reservoir status).
+This information could help direct future research efforts to focus on particular dimensions of lake geometry (i.e.
+horizontal versus vertical) or to stratify model predictions based on specific lake types and cross-section shapes.
+Ultimately, achieving increased depth prediction accuracy would allow for more precise estimates of depth-dependent biotic and chemical processes across broad spatial extents.
 
@@
 its “visual distance to lake center”).
-For these calculations, we used the \texttt{polylabelr} R package \cite{larssonPolylabelrFindPole2019d}, which interfaces with the Mapbox pole of inaccessibility algorithm \cite{agafonkinJSLibraryFinding2019}.
-We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model (~15x15m grain) accessed using the \texttt{elevatr} R package \cite{hollisterElevatrAccessElevation2017} and the terrain function in the \texttt{raster} R package \cite{hijmansRasterGeographicData2019}.
+For these calculations, we used the \texttt{polylabelr} R package \cite{larssonPolylabelrFindPole2019}, which interfaces with the Mapbox pole of inaccessibility algorithm \cite{agafonkinJSLibraryFinding2019}.
+We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model ($\sim$15x15m grain) accessed using the \texttt{elevatr} R package \cite{hollisterElevatrAccessElevation2017} and computed using the terrain function in the \texttt{raster} R package \cite{hijmansRasterGeographicData2019}.
 
@@
 hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S4).
-We further classified lakes using data from \citeA{polus2020}, which uses the output of a machine learning algorithm to assign a probability to each lake regarding whether it is a reservoir or a natural lake.
+We further classified lakes using data from \citeA{polus2020}, which uses the output of a machine learning algorithm to assign a probability to each lake as to whether it is a reservoir or a natural lake.
 For our purposes, we determined a lake to be a reservoir if the classification probability was 0.75 or greater.
@@
 \begin{table}
-\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
+\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from \citeA{stachelek2020}).
 Predictor variables for computing random forest offsets (Eq \ref{eq2}) are printed in bold face.
@@
 Covariates used in random forest modeling (Table \ref{table1}, Equation \ref{eq2}) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
-HUC4s), were obtained from the LAGOS-US LOCUS database (Smith et al.
-In Prep).
-Shoreline development is a measure of lake perimeter shape defined as:
+HUC4s), were obtained from the LAGOS-US LOCUS database \cite{smith2020}.
+One such measure, that of shoreline development, is a measure of lake perimeter shape defined as:
 
@@
 \noindent
-We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
+We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation \ref{eq1} (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
 Prior to entry into Equation \ref{eq1}, we standardized proxy distances to have the same numeric range as their true counterpart.
@@
 Using a cone model example, we show that nearshore land slope is not representative of in-lake slope.
-Furthermore, our results indicate that use of this metric for prediction results in increased prediction error and systematic overestimation of depth in concave and reservoir lakes.
+Furthermore, our results indicate that the use of nearshore land slope for prediction results in increased error and systematic overestimation of depth in concave and reservoir lakes.
+Although our analysis was limited to lakes with available bathymetry data, these lakes did not have characteristics that differed from that of the overall lake population (Figure S7, S2).
+This lack of difference suggests that our results are likely to be broadly applicable to all lakes although there is a possibility that there is some hidden bias not explored for in our analyses.
 
@@
 A simple indication of this non representativeness is that proxy nearshore land slope was often (in > 74\% of cases) steeper than true in-lake slope.
-This finding is consistent with the suggestion by \citeA{heathcotePredictingBathymetricFeatures2015} that in-lake slope matches nearshore land slope at the time of lake formation, but as lakes age, infilling via sedimentation causes them to develop shallower in-lake slopes compared to the surrounding land.
-If true, then these results also lend support to prior work \cite{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}, conducted on a more limited number of lakes, showing that in-lake slopes are controlled by erosion and sediment transport processes.
+This finding is consistent with \citeA{heathcotePredictingBathymetricFeatures2015} whos results suggest that in-lake slopes are shallower compared to the surrounding land.
+The shallow nature of in-lake slopes is likley a function of erosion and sediment transport processes \cite{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}.
 
@@
 
-We found other differences between convex and concave lakes.
+We found other differences among lake geometry measures according to lake cross-section shape.
 One finding was that convex lakes, when compared to concave lakes, had longer distances to lake centers relative to corresponding distances to the deepest point of lakes.
@@
 However, given the conceptual links between cross-section shape, glaciation, and sedimentation \cite{johanssonNewApproachesModelling2007} it may be advantageous for future studies to compile data on sedimentation to determine if this data can be used to predict cross-section shape and boost depth prediction accuracy.
+
+\section{Conclusion}
+\noindent
+To our knowledge, the present study is the largest and most comprehensive test to date of geometric models of lake depth.
+Using bathymetry data on approximately 5,000 lakes, we show that proxy slope measures are not representative of true in-lake slope and this leads to inaccuracies in predicting the depth of concave and reservoir lakes.
+These innaccuracies suggest that caution is warranted in using geometric models for depth prediction in unsampled lakes.
+Despite these apparent biases, overall prediction accuracy was equivalent to that of prior depth prediction studies ($\pm$ 6-7m).
+Only our models using a true measure of in-lake slope, which is limited in availability to lakes with bathymetry data and where we already know lake depth, had greater accuracy than that of prior studies ($\pm$ 4.2m).
+Lack of improved prediction accuracy (short of including data that is unavailable for most lakes) suggests that improved prediction may require new types of data or novel analysis techniques.
 

### 2020-07-16 manuscript/si_template_2019.tex
sync ms text with dissertation edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified December 2018
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% FOR FIGURES, DO NOT USE \psfrag
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% IMPORTANT NOTE:
%
% SUPPORTING INFORMATION DOCUMENTATION IS NOT COPYEDITED BEFORE PUBLICATION.
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%
% Choose the journal abbreviation for the journal you are
% submitting to:

% jgrga JOURNAL OF GEOPHYSICAL RESEARCH (use for all of them)
% gbc   GLOBAL BIOCHEMICAL CYCLES
% grl   GEOPHYSICAL RESEARCH LETTERS
% pal   PALEOCEANOGRAPHY
% ras   RADIO SCIENCE
% rog   REVIEWS OF GEOPHYSICS
% tec   TECTONICS
% wrr   WATER RESOURCES RESEARCH
% gc    GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw    SPACE WEATHER
% ms    JAMES
% ef    EARTH'S FUTURE
%
%
%
% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

\documentclass[draft,wrr]{agutexSI2019}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  SUPPORTING INFORMATION TEMPLATE
%
%% ------------------------------------------------------------------------ %%
%
%
%Please use this template when formatting and submitting your Supporting Information.

%This template serves as both a “table of contents” for the supporting information for your article and as a summary of files.
%
%
%OVERVIEW
%
%Please note that all supporting information will be peer reviewed with your manuscript. It will not be copyedited if the paper is accepted.
%In general, the purpose of the supporting information is to enable authors to provide and archive auxiliary information such as data tables, method information, figures, video, or computer software, in digital formats so that other scientists can use it.
%The key criteria are that the data:
% 1. supplement the main scientific conclusions of the paper but are not essential to the conclusions (with the exception of
%    including %data so the experiment can be reproducible);
% 2. are likely to be usable or used by other scientists working in the field;
% 3. are described with sufficient precision that other scientists can understand them, and
% 4. are not exe files.
%
%USING THIS TEMPLATE
%
%***All references should be included in the reference list of the main paper so that they can be indexed, linked, and counted as citations.  The reference section does not count toward length limits.
%
%All Supporting text and figures should be included in this document. Insert supporting information content into each appropriate section of the template. To add additional captions, simply copy and paste each sample as needed.

%Tables may be included, but can also be uploaded separately, especially if they are larger than 1 page, or if necessary for retaining table formatting. Data sets, large tables, movie files, and audio files should be uploaded separately. Include their captions in this document and list the file name with the caption. You will be prompted to upload these files on the Upload Files tab during the submission process, using file type “Supporting Information (SI)”

%IMPORTANT NOTE ON FIGURES AND TABLES
% Placeholders for figures and tables appear after the \end{article} command, after references.
% DO NOT USE \psfrag or \subfigure commands.
%
 \usepackage{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
 \setkeys{Gin}{draft=false}
%
% You may need to use one of these options for graphicx depending on the driver program you are using.
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

% Author names in capital letters:
%\authorrunninghead{BALES ET AL.}

% Shorter version of title entered in capital letters:
%\titlerunninghead{SHORT TITLE}

%Corresponding author mailing address and e-mail address:
%\authoraddr{Corresponding author: A. B. Smith,
%Department of Hydrology and Water Resources, University of
%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
%(a.b.smith@hwr.arizona.edu)}

\begin{document}

%% ------------------------------------------------------------------------ %%
%
%  TITLE
%
%% ------------------------------------------------------------------------ %%

%\includegraphics{agu_pubart-white_reduced.eps}

\title{Supporting Information for "Geometric models overestimate lake depth due to imperfect slope
measurement"}
%
% e.g., \title{Supporting Information for "Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$"}
%
%DOI: 10.1002/%insert paper number here%

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{J. Stachelek\affil{1}, P. J. Hanly\affil{1}, and P. A. Soranno\affil{1}}

\affiliation{1}{Department of Fisheries and Wildlife, Michigan State University, 480
Wilson Rd., East Lansing, MI 48824, USA}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

% \affiliation{=number=}{=Affiliation Address=}
%(repeat as many times as is necessary)

%% ------------------------------------------------------------------------ %%
%
%  BEGIN ARTICLE
%
%% ------------------------------------------------------------------------ %%

% The body of the article must start with a \begin{article} command
%
% \end{article} must follow the references section, before the figures
%  and tables.

\begin{article}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

\noindent\textbf{Contents of this file}
%%%Remove or add items as needed%%%
\begin{enumerate}
\item \textbf{Figure S1:} Map of study lakes

\item \textbf{Figure S2:} Comparison between reported depth and depth extracted from bathymetry surfaces

\item \textbf{Figure S3:} Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status

\item \textbf{Figure S4:} Hypsography classification by state

\item \textbf{Figure S5:} Comparison among lake shape and reservoir classes for true and proxy geometry measures

\item \textbf{Figure S6:} Importance plot for random forest variables showing increase in mean square error

%if Tables are larger than 1 page, upload as separate excel file
\end{enumerate}

% \noindent\textbf{Introduction}
%Type or paste your text here. The introduction gives a brief overview of the supporting information. You should include information %about as many of the following as possible (when appropriate):
% 1. a general overview of the kind of data files;
% 2. information about when and how the data were collected or created;
% 3. a general description of processing steps used;
% 4. any known imperfections or anomalies in the data.

%\clearpage

%Delete all unused file types below. Copy/paste for multiples of each file type as needed.
%\noindent\textbf{Text S1.}
%Type or paste text here. This should be additional explanatory text, such as: extended descriptions of results, full details of models, extended lists of acknowledgements etc.  It should not be additional discussion, analysis, interpretation or critique. It should not be an additional scientific experiment or paper.
%
%Repeat for any additional Supporting Text

%%Enter Data Set, Movie, and Audio captions here
%%EXAMPLE CAPTIONS

% \noindent\textbf{Data Set S1.} %Type or paste caption here.
%upload your dataset(s) to AGU's journal submission site and select "Supporting Information (SI)" as the file type. Following naming %convention: ds01.

%Repeat for any additional Supporting data sets

% \noindent\textbf{Movie S1.} %Type or paste caption here.
%upload your movie(s) to AGU's journal submission site and select, "Supporting Information %(SI)" as the file type. Following naming convention: ms01.

%Repeat any additional Supporting movies

% \noindent\textbf{Audio S1.} %Type or paste caption here.
%upload your audio file(s) to AGU's journal submission site and select "Supporting Information %(SI)" as the file type. Following naming convention: auds01.

%Repeat for any additional Supporting audio files

%%% End of body of article:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Notation section goes here
%
% Notation -- End each entry with a period.
% \begin{notation}
% Term & definition.\\
% Second term & second definition.\\
% \end{notation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% ------------------------------------------------------------------------ %%
%%  REFERENCE LIST AND TEXT CITATIONS

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
\bibliography{lagosdepth}
%
% no need to specify bibliographystyle
%
% Note that ALL references in this supporting information file must also be referenced in the primary manuscript
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% if you get an error about newblock being undefined, uncomment this line:
%\newcommand{\newblock}{}

% \bibliography{ uncomment this line and enter the name of your bibtex file here }

%Reference citation instructions and examples:
%
% Please use ONLY \cite and \citeA for reference citations.
% \cite for parenthetical references
% ...as shown in recent studies (Simpson et al., 2019)
% \citeA for in-text citations
% ...Simpson et al (2019) have shown...
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
%
%
%...as shown by \citeA{jskilby}.
%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
%...has been shown \cite<e.g.,>{jskilbye}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%
% apacite uses < > for prenotes, not [ ]
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
%

%% ------------------------------------------------------------------------ %%
%
%  END ARTICLE
%
%% ------------------------------------------------------------------------ %%
\end{article}
\clearpage

\begin{figure}
      \begin{center}\includegraphics[width=0.65\textwidth]{../figures/00_map_bathy-1}\end{center}
      \caption{Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.}\label{figs1}
      \end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_hypsography-1}
      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figs4}
      \end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_geometry_grid-1}
      \caption{Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope. A dashed 1:1 line is shown for comparison. Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figs5}
      \end{figure}

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
      \caption{Importance plot for random forest variables showing increase in mean square error. Higher values indicate greater importance to model predictions. See Equation 1 for a definition of geometric max depth. HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figs6}
      \end{figure}

\begin{figure}
      \begin{center}\includegraphics[width=0.45\textwidth]{../figures/gg_effort-1}\end{center}
            \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product \cite{stachelek2020}. The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figs7}
      \end{figure}

\clearpage
\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}
      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figs3}
      \end{figure}

\clearpage

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}
      \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product \cite{stachelek2020}. For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figs2}
      \end{figure}

% Copy/paste for multiples of each file type as needed.

% enter figures and tables below here: %%%%%%%
%
%
%
%
% EXAMPLE FIGURES
% ---------------
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
% \begin{figure}
%\setfigurenum{S1} %%You can change number for each figure if you want, not required. "S" prepended automatically.
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.eps}
%\caption{caption}
%\label{epsfiguresample}
%\end{figure}
%
%
% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
% \begin{figure}
% \noindent\includegraphics[width=\textwidth]{anothersample.png}
%\caption{caption}
%\label{pngfiguresample}
%\end{figure}
%
%
%\begin{figure}
%\noindent\includegraphics[width=\textwidth]{athirdsample.pdf}
%\caption{A pdf test figure}
%\label{pdffiguresample}
%\end{figure}
%
% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
%
%
% ---------------
% EXAMPLE TABLE
%
%\begin{table}
%\settablenum{S1} %%Change number for each table
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\centering
%\begin{tabular}{l c}
%\hline
% Run  & Time (min)  \\
%\hline
%  $l1$  & 260   \\
%  $l2$  & 300   \\
%  $l3$  & 340   \\
%  $h1$  & 270   \\
%  $h2$  & 250   \\
%  $h3$  & 380   \\
%  $r1$  & 370   \\
%  $r2$  & 390   \\
%\hline
%\end{tabular}
%\tablenotetext{a}{Footnote text here.}
%\end{table}
% ---------------
%
% EXAMPLE LARGE TABLE (UPLOADED SEPARATELY)
%\begin{table}
%\settablenum{S1} %%Change number for each table
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\end{table}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
% BREAK THE EQUATIONS INTO TWO OR MORE LINES
% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
% using double backslashes (\\).

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

%% ------------------------------------------------------------------------ %%
%
%  SIDEWAYS FIGURE AND TABLE EXAMPLES
%
%% ------------------------------------------------------------------------ %%
%
% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{samplefigure.eps}
% \caption{caption here}
% \label{label_here}
% \end{sidewaysfigure}
%
%
%
% \begin{sidewaystable}
% \caption{}
% \begin{tabular}
% Table layout here.
% \end{tabular}
% \end{sidewaystable}
%
%

### 2020-07-16 manuscript/si_template_2019.tex
fix supp fig listing

@@
 
-\item \textbf{Figure S2:} Comparison between reported depth and depth extracted from bathymetry surfaces
+\item \textbf{Figure S2:} Hypsography classification by state
 
-\item \textbf{Figure S3:} Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status
+\item \textbf{Figure S3:} Comparison among lake shape and reservoir classes for true and proxy geometry measures
 
-\item \textbf{Figure S4:} Hypsography classification by state
+\item \textbf{Figure S4:} Importance plot for random forest variables showing increase in mean square error
 
-\item \textbf{Figure S5:} Comparison among lake shape and reservoir classes for true and proxy geometry measures
+\item \textbf{Figure S5:} Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources
 
-\item \textbf{Figure S6:} Importance plot for random forest variables showing increase in mean square error
+\item \textbf{Figure S6:} Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status
+
+\item \textbf{Figure S7:} Comparison between reported depth and depth extracted from bathymetry surfaces
 
@@
 \begin{figure}
-      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}
+      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
       \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figs3}
@@
 \begin{figure}
-      \noindent\includegraphics[width=0.85\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}
+      \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
       \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product \cite{stachelek2020}.

### 2020-07-28 manuscript/agujournaltemplate.tex
use pjh for initials
fill out acknowledgements, closes #13
prep for submission

@@
 \begin{document}
-
+%TC:ignore
 %% ------------------------------------------------------------------------ %%
@@
 \item Using data from thousands of lakes, we show that nearshore land slope is a poor proxy for in-lake slope and increases prediction error.
-\item Increases in prediction error were systematic such that depth was overpredicted in concave and reservoir lakes.
+\item Prediction errors were systematic such that depth was overpredicted in concave and reservoir lakes.
 \end{keypoints}
@@
 %% ------------------------------------------------------------------------ %%
-
+%TC:endignore
 %% \begin{abstract} starts the second page
-
 \begin{abstract}
-      Lake depth is a critical characteristic that influences many important
-      ecological processes in lakes. Unfortunately, lake depth measurements
-      are labor-intensive to gather and are only available for a small
-      fraction of lakes globally. Therefore, scientists have tried to predict
-      lake depth from characteristics that are easily obtained for all lakes
-      such as lake surface area or the slope of the land surrounding a lake.
-      One approach for predicting lake depth simulates lake basins using a
-      geometric model where nearshore land slope is assumed to be a
-      representative proxy for in-lake slope and the distance to the center of
-      the lake is assumed to be a representative proxy for the distance to the
-      deepest point of the lake. However, these assumptions have rarely been
-      tested in a broad range of lakes. We used bathymetry data from 5,000
-      lakes and reservoirs to test these assumptions and to examine whether
-      differences in lake type or shape influences depth prediction error. We
-      found that nearshore land slope was not a representative proxy of
-      in-lake slope and using it for prediction increases prediction error
-      substantially relative to models using true in-lake slope. We also found
-      that models using nearshore land slope as a proxy systematically
-      overpredicts lake depth in concave lakes (i.e. bowl-shaped; up to 18\%
-      of lakes in the study population) and reservoir lakes (up to 30\% of
-      lakes in the study population), suggesting caution in using geometric
-      models for depth prediction in unsampled lakes.
+Lake depth is a critical characteristic that influences many important ecological processes in lakes.
+Unfortunately, lake depth measurements are labor-intensive to gather and are only available for a small fraction of lakes globally.
+Therefore, scientists have tried to predict lake depth from characteristics that are easily obtained for all lakes such as lake surface area or the slope of the land surrounding a lake.
+One approach for predicting lake depth simulates lake basins using a geometric model where nearshore land slope is assumed to be a representative proxy for in-lake slope and the distance to the center of the lake is assumed to be a representative proxy for the distance to the deepest point of the lake.
+However, these assumptions have rarely been tested in a broad range of lakes.
+We used bathymetry data from approximately 5,000 lakes and reservoirs to test these assumptions and to examine whether differences in lake type or shape influences depth prediction error.
+We found that nearshore land slope was not representative of in-lake slope and using it for prediction increases error substantially relative to models using true in-lake slope.
+We also found that models using nearshore land slope as a proxy systematically overpredict lake depth in concave lakes (i.e.
+bowl-shaped; up to 18\% of lakes in the study population) and reservoir lakes (up to 30\% of lakes in the study population), suggesting caution in using geometric models for depth prediction in unsampled lakes.
 \end{abstract}
@@
 The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \cite{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
-Given the limited prediction accuracy of prior attempts ($\pm$ 6-7 m), studies have explored strategies such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings (e.g.
+Given the limited prediction accuracy of prior depth prediction efforts ($\pm$ 6-7 m), a major focus has been on  improving prediction accuracy using strategies such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings (e.g.
 fitting different models for distinct size classes) among lakes \cite{sobekPredictingDepthVolume2011}.
@@
 All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae.
-In the simplest case, where lakes basins are treated as cones (Eq \ref{eq1}, Figure \ref{fig1}), two assumptions are required to make depth predictions for all lakes: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the lake is a representative proxy for the distance to the deepest point of the lake (Figure \ref{fig1}).
+In the simplest case, where lakes basins are treated as cones (Equation \ref{eq1}, Figure \ref{fig1}), two assumptions are required to make depth predictions for all lakes: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the lake is a representative proxy for the distance to the deepest point of the lake (Figure \ref{fig1}).
 This cone model imposes the following fixed (i.e.
@@
 Because tests of geometric model assumptions require bathymetric map data, which is only available for about 15\% of all lakes, existing evidence may not be applicable to all lakes.
-The few studies that have tested these assumptions have mostly been limited to individual studies of very large ($>$ 500 ha) lakes or studies on small numbers ($<$ 100) of lakes \cite{johanssonNewApproachesModelling2007}.
+The few studies that have tested these assumptions have been limited to individual studies of very large ($>$ 500 ha) lakes or studies on small numbers ($<$ 100) of lakes \cite{johanssonNewApproachesModelling2007}.
 Studies focused specifically on reservoirs (as opposed to the more typical case where reservoirs and natural lakes are combined), have been even more restricted to that of extremely large lakes $>$ 1000 ha \cite{lehnerHighresolutionMappingWorld2011, messagerEstimatingVolumeAge2016}.
@@
 Additionally it is unclear whether model prediction error is related to differences in lake type such those with different cross-section shapes or those classified as reservoirs versus natural lakes.
-
-Given the knowledge gaps identified above, we asked three research questions: 1) How representative is nearshore land slope of in-lake slope; and how representative is the distance to the center of a lake compared to the distance to the deepest point of a lake?
+Given these knowledge gaps, we asked three research questions: 1) How representative is nearshore land slope of in-lake slope; and how representative is the distance to the center of a lake compared to the distance to the deepest point of a lake?
 2) How does the use of proxies for lake geometry affect lake depth prediction error?
@@
 natural lake vs reservoir) affect depth prediction error?
-To answer these questions, we extracted maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, cross-section shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 lakes in the Northeastern and Midwestern US from bathymetric map data and supplemented this data with classification estimates of whether lakes are reservoirs or natural lakes.
-First, we examined whether measures of lake geometry (in-lake slope and distance to the deepest point of lakes) were related to geometry proxies (nearshore land slope and distance to lake centers).
-Next, we computed geometric depth estimates (Equation \ref{eq1}) and prediction "offsets" to these estimates using the random forest algorithm (Equation \ref{eq2}).
-Covariates used in offset modeling included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table \ref{table1}).
-We examined differences in overall prediction error corresponding to different inputs to geometric depth as well as differences in the relative distribution of prediction error in lakes with different characteristics.
+To answer these questions, we extracted maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, cross-section shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 lakes from bathymetric map data and supplemented this data with classification estimates of whether lakes are reservoirs or natural lakes.
+We used this data to compute geometric depth estimates (Equation \ref{eq1}) and prediction "offsets" to these estimates using the random forest algorithm (Equation \ref{eq3}).
+Covariates used in offset modeling included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table S1).
 
-Given that, by definition, the distance proxy (distance to the center of the lake) must always be greater or equal to the true distance value (distance to the deepest point of the lake), we expect that the use of this proxy will lead to overestimation of lake depth (Figure \ref{fig3}).
+By definition, the distance proxy (distance to the center of the lake) must always be greater or equal to the true distance value (distance to the deepest point of the lake).
+Therefore, we expect that the use of this proxy will lead to overestimation of lake depth (Figure \ref{fig1}).
 Furthermore we expect to see greater overestimation error in reservoirs as compared to natural lakes because many reservoirs are known to be drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
 next to the dam) rather than in the center of the reservoir \cite{lanza1985interactions}.
-In a similar fashion, we expect to see overestimation error associated with using a nearshore land slope proxy in lakes with differing cross-section shape (Figure \ref{fig2}) such that the depth of U-shaped (i.e.
+In a similar fashion, we expect to see overestimation error associated with using a nearshore land slope proxy in lakes with differing cross-section shape such that the depth of U-shaped (i.e.
 concave) lakes will be overpredicted whereas the depth of V-shaped (i.e.
-convex) lakes will be underpredicted.
-Finally, we expect that overall depth prediction will be strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies \cite{oliver2016prediction}.
+convex) lakes will be underpredicted (Figure S1).
+Finally, we expect that depth predictions themselves will be strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies \cite{oliver2016prediction}.
 
-By testing these expectations, we can establish whether barriers to increased depth prediction accuracy lie in lack of correspondence between true and proxy measures of lake geometry or in hidden grouping (such as lake cross-section shape or reservoir status).
+By testing these expectations, we can establish whether barriers to increased depth prediction accuracy lie in lack of correspondence between true and proxy measures of lake geometry or in hidden groupings among lakes (such as lake cross-section shape or reservoir status).
 This information could help direct future research efforts to focus on particular dimensions of lake geometry (i.e.
@@
 
-\begin{figure}
-      \begin{center}
-        \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/lake_shape}
-      \end{center}
-      \caption{Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
-Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom.}\label{fig2}
-    \end{figure}
-
 \section{Methods}
@@
 \noindent
-We compiled bathymetry data on approximately 5,000 lakes in the Northeastern and Midwestern US from nine official state databases (Figure S1).
+We compiled bathymetry data on approximately 5,000 lakes in the Northeastern and Midwestern US from nine official state databases (Figure S2).
 The original data came in a variety of formats including pre-interpolated rasters (Minnesota), contour lines (Nebraska, Michigan, Massachusetts, Kansas, Iowa), contour polygons (New Hampshire, Connecticut), or point depth soundings (Maine).
@@
 For cross-section shape, we categorized lakes as either convex or concave following the method of \citeA{hakansonLakeFormLake1977} by computing normalized lake depth-area relationships (i.e.
-hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S4).
+hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S3).
 We further classified lakes using data from \citeA{polus2020}, which uses the output of a machine learning algorithm to assign a probability to each lake as to whether it is a reservoir or a natural lake.
@@
 
-\begin{table}
-\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from \citeA{stachelek2020}).
-Predictor variables for computing random forest offsets (Eq \ref{eq2}) are printed in bold face.
-Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
-\centering
-\setlength\tabcolsep{1.5pt} % default value: 6pt
-\begin{tabular}{lllll}
-  \hline
-  Variable & Median & Q25 & Q75 & n\\
-  \hline
-  Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4850 (17700)\\
-  \textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4850 (17700)\\
-  \textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4850 (17700)\\
-  \textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4850 (17700)\\
-  \textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4850 (17700)\\
-  \textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4850 (17700)\\
-  \textbf{Watershed-lake ratio} & 7.8 (10) & 3.9 (4.4) & 17 (28) & 4850 (17700)\\
-  Distance to deepest point (m) & 180 (-) & 110 (-) & 290 (-) & 4850 (-)\\
-  Distance to lake center (m) & 240 (-) & 160 (-) & 380 (-) & 4850 (-)\\
-  In-lake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4850 (-)\\
-  Nearshore land slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4850 (-)\\
-  \hline
-  \end{tabular}
-\end{table}
-
-Covariates used in random forest modeling (Table \ref{table1}, Equation \ref{eq2}) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
+Covariates used in random forest modeling (Table S1, Equation \ref{eq3}) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
 HUC4s), were obtained from the LAGOS-US LOCUS database \cite{smith2020}.
@@
 Watershed to lake area ratio is an approximation of water residence time and is defined as watershed area divided by lake area.
-All of our data and data processing code are available at [Zenodo DOI].
 
@@
 In the second step, we fit a random forest model to predict observed (i.e.
-true) depth as a function of geometric depth along with several covariates available for all lakes (Table \ref{table1}).
+true) depth as a function of geometric depth along with several covariates available for all lakes (Table S1).
 The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of $depth_{geometric}$ and among different lake types.
@@
 \noindent
-Nearly all prior studies predicting lake depth using geometric models include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy.
+Prior studies using geometric models to predict lake depth include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy.
 For our purposes, this offset modeling enabled us to test our expectations that prediction error would be different among different formulations of $depth_{geometric}$ and among different lake types.
@@
 We evaluated model fit and prediction error using root-mean-square error (RMSE) and coefficient of determination ($R^2$) metrics on a holdout set containing 25\% of all lakes.
-We evaluated the residuals (residual = observed - predicted) of each model relative to lake cross-section shape and reservoir classes to determine whether depth is consistently over or under predicted for some lake types relative to others.
-All of our code for model fitting and evaluation is available at [Zenodo DOI].
+We evaluated the residuals of each model relative to lake cross-section shape and reservoir classes to determine whether depth is consistently over or under predicted for some lake types relative to others.
 
@@
 \noindent
-Lakes belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Figure S1).
+Lakes belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Figure S2, S3).
 For example, concave lakes were nearly absent from Michigan whereas Maine lakes had an overabundance of lakes categorized as neither concave nor convex.
-Lakes in the southern portions of our study area tended to be classified as reservoirs whereas lakes in the northern portions of our study area were a more even mix between reservoirs and natural lakes (Figure S1).
+Lakes in the southern portions of our study area tended to be classified as reservoirs whereas lakes in the northern portions of our study area were a more even mix between reservoirs and natural lakes (Figure S2).
 Approximately 18\%, 80\%, and 2\% of lakes were classified as having a concave, convex, or neither shape respectively whereas approximately 30\% and 70\% of lakes were classified as being a reservoir or a natural lake.
@@
 shallower) nearshore land slope compared to true in-lake slope (Figure \ref{fig3}).
-Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes while proxy nearshore land slope is not representative of true in-lake slope.
+Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes whereas proxy nearshore land slope is not representative of true in-lake slope.
 In addition to overall differences between slope and distance measures, we found differences in these relationships among lake shape classes.
-For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S5).
-However,  there were not similar differences among slope and distance metrics for natural lakes versus reservoirs (Figure S5).
+For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S4).
+However,  there were not similar differences among slope and distance metrics for natural lakes versus reservoirs (Figure S4).
 
@@
 
-\begin{table}
+\begin{table}[h]
   \caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table2}
@@
 
-The most important covariates in offset models were those relating to spatial location, lake area, and perimeter (Figure S6).
-Conversely, watershed metrics and lake elevation had little contribution to random forest model fit (Figure S6).
+The most important covariates in offset models were those relating to spatial location, lake area, and perimeter (Figure S5).
+Conversely, watershed metrics and lake elevation had little contribution to random forest model fit (Figure S5).
 The spatial location (i.e.
 HUC4) covariate was notably less importance in the true slope model compared to the two proxy slope models.
-Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Eq \ref{eq1} (Figure S6).
+Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Equation \ref{eq1} (Figure S5).
 
@@
 \noindent
-Our tests of geometric lake depth models show that specific proxy measures of lake geometry are not representative of true measures of lake geometry across a broad array of lakes.
-Using a cone model example, we show that nearshore land slope is not representative of in-lake slope.
-Furthermore, our results indicate that the use of nearshore land slope for prediction results in increased error and systematic overestimation of depth in concave and reservoir lakes.
-Although our analysis was limited to lakes with available bathymetry data, these lakes did not have characteristics that differed from that of the overall lake population (Figure S7, S2).
-This lack of difference suggests that our results are likely to be broadly applicable to all lakes although there is a possibility that there is some hidden bias not explored for in our analyses.
+Our tests of geometric lake depth models show that specific proxy measures of lake geometry are not representative of true geometry measures across a broad array of lakes.
+Models using non-representative proxies showed increased error and systematic overestimation of depth in concave and reservoir lakes.
+Although our analysis was limited to lakes with available bathymetry data, these lakes did not have characteristics that differed from that of the overall lake population (Figure S6, S8).
+Although there is a possibility that there is some hidden bias not explored for in our analyses, this lack of difference suggests that our results are likely to be broadly applicable to all lakes.
 
@@
 In comparing among lake geometry measures, our analysis suggests that proxy distance to lake center is representative of true distance to the deepest point of the lakes but that proxy nearshore land slope is not representative of true in-lake slope.
-A simple indication of this non representativeness is that proxy nearshore land slope was often (in > 74\% of cases) steeper than true in-lake slope.
+A simple indication of this non representativeness is that proxy nearshore land slope was often (in $>$ 74\% of cases) steeper than true in-lake slope.
 This finding is consistent with \citeA{heathcotePredictingBathymetricFeatures2015} whos results suggest that in-lake slopes are shallower compared to the surrounding land.
@@
 any permanent waterbody that has a water control structure likely to significantly impact flow or pool water) compared to that of conventional classifications that are tied to specific dam types or dam heights.
-Another possible explanation is that conventional reservoir classifications are conceptually biased towards more southern areas with few natural lakes (Figure S1).
+Another possible explanation is that conventional reservoir classifications are conceptually biased towards more southern areas with few natural lakes (Figure S2).
 
@@
 In addition, convex lakes often had steeper in-lake slopes relative to nearshore land slopes as compared to concave lakes.
-These differences are reflected in the class-wise differences in shoreline development (e.g.
-shorelines were more sinuous in convex lakes compared to concave lakes, Figure S3).
-It was notable that convex lakes were deeper than concave lakes despite having similar distributions of lake surface area (Figure S3).
-Given the similarity in lake surface area, the underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of lake formation whereby concave lakes were formed as a result of more intense glacial scouring compared to convex lakes \cite{gorhamPhysicalLimnologyNorthern1958}.
-While there is some evidence in support of this idea, namely that there is a geographic hotspot of concave lakes associated with the glaciated “prairie pothole region” \cite<see>{hayashiSimpleEquationsRepresent2000}, the overall geographic distribution of lake cross-section shapes does not support this idea.
+Finally, it was notable that convex lakes were deeper than concave lakes despite having similar distributions of lake surface area (Figure S7).
+The underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of lake formation whereby the formation of concave lakes were a result of more intense glacial scouring compared to that of convex lakes \cite{gorhamPhysicalLimnologyNorthern1958}.
+While our findings provide some evidence in support of this idea, namely that there is a geographic hotspot of concave lakes associated with the glaciated “prairie pothole region” \cite<see>{hayashiSimpleEquationsRepresent2000}, the overall geographic distribution of lake cross-section shapes does not support this idea.
 Instead of a concentrated area of concave lakes in formerly glaciated regions, there appears to be a fairly even mix of concave and convex lakes distributed amongst the northern (i.e.
-glaciated) and southern (non-glaciated) portions of our study area (Figure S1).
+glaciated) and southern (non-glaciated) portions of our study area (Figure S2).
 
@@
 Models using only proxy variables had prediction error rates (RMSE = 6.6m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting lake depth at broad geographic extents \cite{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
-When only a single proxy measure was used there was a difference in model sensitivity depending on if it was a horizontal distance measure or a vertical slope measure.
+When only a single proxy measure was used, there was a difference in model sensitivity depending on if it was a horizontal distance measure or a vertical slope measure.
 In the case of a true slope and proxy distance combination, models were more accurate ($\pm$ 4.2m) than even the most accurate of prior studies \cite{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
@@
 The greater sensitivity of depth predictions to proxy slope measures relative to proxy distance measures may be explained by the fact that proxy slope measures were a more imperfect representation of true in-lake slopes relative to proxy versus true distances.
-In addition, these results help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \cite{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S3).
+In addition, these results help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \cite{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S7).
 
@@
 \noindent
-As expected, we found that the maximum depth of concave lakes was systematically overpredicted by a simple geometric model using proxy nearshore land slope.
+As expected, we found that the maximum depth of concave lakes was systematically overpredicted by a simple geometric model using proxy nearshore land slope (Figure S1).
 However, contrary to our expectation, we did not observe underprediction of depth in convex lakes.
 The reason we did not observe underprediction of the depth of convex lakes is likely because geometric depth itself was always greater than observed maximum depth owing to the fact that proxy distance is constrained to be greater than true distance.
-Although the magnitude of overestimation is likely related to class imbalance in our dataset (i.e.
-there was a greater number of convex lakes), these results suggest that broad scale estimates of lake depth are overestimated particularly when those estimates encompass large numbers of lakes with diverse cross-section shapes.
+This suggests that depth estimates in prior studies may be overestimated when they encompass large numbers of lakes with diverse cross-section shapes.
 
@@
 We propose that the error rate of this model ($\pm$ 4.2m) be used as an out-of-sample prediction benchmark for future studies such that they should attempt to match it but not expect to exceed it.
-Because this model requires bathymetry data, this suggests that it may not be possible with current data and models to produce depth predictions for all lakes with error rates below about 6m.
-To approach our benchmark using data available for all lakes, future studies could explore alternative modeling approaches such as ordinal modeling, which would capture whether or not a lake crosses some important depth threshold, but would not seek to predict a specific depth value.
-These studies could also explore emerging data types to boost prediction accuracy such as “topobathymetric” products that integrate both topographic and bathymetric data in a seamless fashion rather than treating them as separate entities.
-This would allow for more robust tests of the representativeness of geometric model inputs.
-Unfortunately, topobathymetric products are rare, have mostly been limited nearshore marine environments, and as such are not widely available for inland waters \cite{danielsonTopobathymetricElevationModel2016}.
+
+Because this most accurate model requires bathymetry data, this suggests that it may not be possible with current data and models to produce depth predictions for all lakes with error rates below 6m.
+To achieve high prediction accuracy using data available for all lakes, future studies could explore alternative modeling approaches such as ordinal modeling, which would capture whether or not a lake crosses some important depth threshold but would not seek to predict a specific depth value, or emerging data types such as “topobathymetric” products that integrate both topographic and bathymetric data in a seamless fashion rather than treating them as separate entities.
+Topobathymetry would allow for more robust tests of the representativeness of geometric model inputs.
+Unfortunately, topobathymetric products are rare, have mostly been limited nearshore marine environments, and as such are not yet widely available for inland waters \cite{danielsonTopobathymetricElevationModel2016}.
 
@@
 
+%TC:ignore
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\acknowledgments
+\acknowledgments All data as well as code for data processing, model fitting, and model evaluation is available at [Zenodo DOI].
 Funding was provided by the US NSF Macrosystems Biology Program grants, DEB-1638679; DEB-1638550, DEB-1638539, DEB-1638554.
@@
 Author contributions: JS conceived of the study, built models, analyzed data, and wrote the paper.
-PH and PAS provided interpretation of results and edited the paper.
+PJH and PAS provided interpretation of results and edited the paper.
+This work benefited from participation in the Global Lake Ecological Observatory Network (GLEON).
+We thank K.S.
+Cheruvelil for a friendly review of an earlier draft.
+%TC:endignore
 

### 2020-07-28 manuscript/si_template_2019.tex
prep for submission

@@
 \begin{enumerate}
-\item \textbf{Figure S1:} Map of study lakes
 
-\item \textbf{Figure S2:} Hypsography classification by state
+\item \textbf{Figure S1:} Diagram showing expectations regarding depth prediction bias
 
-\item \textbf{Figure S3:} Comparison among lake shape and reservoir classes for true and proxy geometry measures
+\item \textbf{Figure S2:} Map of study lakes
 
-\item \textbf{Figure S4:} Importance plot for random forest variables showing increase in mean square error
+\item \textbf{Figure S3:} Hypsography classification by state
 
-\item \textbf{Figure S5:} Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources
+\item \textbf{Figure S4:} Lake geometry comparison among shape and reservoir classes
 
-\item \textbf{Figure S6:} Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status
+\item \textbf{Figure S5:} Importance plot for random forest variables
 
-\item \textbf{Figure S7:} Comparison between reported depth and depth extracted from bathymetry surfaces
+\item \textbf{Figure S6:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
+
+\item \textbf{Figure S7:} Lake characteristics comparison among shape and reservoir classes
+
+\item \textbf{Figure S8:} Comparison between reported depth and bathymetry depth
 
@@
 \end{enumerate}
+\vspace{1em}
 
@@
 \begin{figure}
+      \begin{center}
+        \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/lake_shape}
+      \end{center}
+      \caption{Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
+Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom.}\label{fig2}
+    \end{figure}
+
+\clearpage
+
+\begin{figure}
       \begin{center}\includegraphics[width=0.65\textwidth]{../figures/00_map_bathy-1}\end{center}
@@
       \end{figure}
+
+\clearpage
+
+\begin{table}
+\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from \citeA{stachelek2020}).
+Predictor variables for computing random forest offsets (Equation 2) are printed in bold face.
+Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
+\centering
+\setlength\tabcolsep{1.5pt} % default value: 6pt
+\begin{tabular}{lllll}
+  \hline
+  Variable & Median & Q25 & Q75 & n\\
+  \hline
+  Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4850 (17700)\\
+  \textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4850 (17700)\\
+  \textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4850 (17700)\\
+  \textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4850 (17700)\\
+  \textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4850 (17700)\\
+  \textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4850 (17700)\\
+  \textbf{Watershed-lake ratio} & 7.8 (10) & 3.9 (4.4) & 17 (28) & 4850 (17700)\\
+  Distance to deepest point (m) & 180 (-) & 110 (-) & 290 (-) & 4850 (-)\\
+  Distance to lake center (m) & 240 (-) & 160 (-) & 380 (-) & 4850 (-)\\
+  In-lake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4850 (-)\\
+  Nearshore land slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4850 (-)\\
+  \hline
+  \end{tabular}
+\end{table}
 

### 2020-07-29 manuscript/agujournaltemplate.tex
rm in prep refs

@@
 hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S3).
-We further classified lakes using data from \citeA{polus2020}, which uses the output of a machine learning algorithm to assign a probability to each lake as to whether it is a reservoir or a natural lake.
+We further classified lakes using the output of a machine learning algorithm to assign a probability to each lake as to whether it is a reservoir or a natural lake.
 For our purposes, we determined a lake to be a reservoir if the classification probability was 0.75 or greater.
-The \citeA{polus2020} data product defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water, beyond simply controlling water level.
+Our reservoir classification data defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water, beyond simply controlling water level.
 It makes no distinction between different dam types or dam heights.
@@
 Covariates used in random forest modeling (Table S1, Equation \ref{eq3}) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
-HUC4s), were obtained from the LAGOS-US LOCUS database \cite{smith2020}.
+HUC4s), were obtained from the LAGOS-US LOCUS database.
 One such measure, that of shoreline development, is a measure of lake perimeter shape defined as:
@@
 next to the dam) rather than in the center of the reservoir \cite{lanza1985interactions}.
-One possible explanation is that the reservoir classification data from \citeA{polus2020} uses a more general definition of a reservoir (i.e.
+One possible explanation is that our reservoir classification data uses a more general definition of a reservoir (i.e.
 any permanent waterbody that has a water control structure likely to significantly impact flow or pool water) compared to that of conventional classifications that are tied to specific dam types or dam heights.

### 2020-07-29 manuscript/si_template_2019.tex
rm in prep refs

@@
 %
-\bibliography{lagosdepth}
+% \bibliography{lagosdepth}
 %
@@
       \begin{center}\includegraphics[width=0.45\textwidth]{../figures/gg_effort-1}\end{center}
-            \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product \cite{stachelek2020}.
+            \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product.
 The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figs7}
@@
       \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
-      \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product \cite{stachelek2020}.
+      \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product.
 For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figs2}
@@
 \begin{table}
-\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from \citeA{stachelek2020}).
+\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States).
 Predictor variables for computing random forest offsets (Equation 2) are printed in bold face.

### 2020-11-02 manuscript/agujournaltemplate.tex
initial minor text revisions

@@
 Lake depth is a critical characteristic that influences many important ecological processes in lakes.
-Unfortunately, lake depth measurements are labor-intensive to gather and are only available for a small fraction of lakes globally.
+Unfortunately, lake depth measurements are labor-intensive to gather and are only available for a tiny fraction of lakes globally.
 Therefore, scientists have tried to predict lake depth from characteristics that are easily obtained for all lakes such as lake surface area or the slope of the land surrounding a lake.
@@
 These differences are reflected in variation among lakes in terms of biological productivity \cite{qinWaterDepthUnderpins2020} and rates of greenhouse gas production \cite{liSignificantContributionLake2020}.
-However, because measured depth data is only available for a small fraction ($\sim$25\%) of all lakes, our ability to understand and predict depth-dependent processes is limited.
+However, because measured depth data is only available for a tiny fraction of lakes ($\sim$25\% of all lakes in our study footprint), our ability to understand and predict depth-dependent processes is limited.
 The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \cite{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.

### 2020-11-03 manuscript/agujournaltemplate.tex
more minor edits

@@
 These differences are reflected in variation among lakes in terms of biological productivity \cite{qinWaterDepthUnderpins2020} and rates of greenhouse gas production \cite{liSignificantContributionLake2020}.
-However, because measured depth data is only available for a tiny fraction of lakes ($\sim$25\% of all lakes in our study footprint), our ability to understand and predict depth-dependent processes is limited.
+However, because measured depth data is only available for a tiny fraction of lakes and $<$ 25\% of all lakes in our study footprint, our ability to understand and predict depth-dependent processes is limited.
 The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \cite{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
-Given the limited prediction accuracy of prior depth prediction efforts ($\pm$ 6-7 m), a major focus has been on  improving prediction accuracy using strategies such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings (e.g.
+Given the limited prediction accuracy of prior depth prediction efforts ($\pm$ 6-7 m), a major focus has been on improving accuracy using a variety of strategies such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings (e.g.
 fitting different models for distinct size classes) among lakes \cite{sobekPredictingDepthVolume2011}.
+These efforts have met with limited improvements in prediction accuracy.
 
-One approach for predicting lake depth involves using a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \cite{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
+One inuitive approach for predicting lake depth involves using a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \cite{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
 All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae.
@@
 For instance, lake cross-section shapes have been shown to vary from narrow "convex" forms to outstretched "concave" forms \cite{hakansonLakeFormLake1977}.
-Because tests of geometric model assumptions require bathymetric map data, which is only available for about 15\% of all lakes, existing evidence may not be applicable to all lakes.
+Because tests of geometric model assumptions require bathymetric map data, which is available for a tiny fraction of lakes (about 15\% of all lakes in our study footprint), existing evidence may not be applicable to all lakes.
 The few studies that have tested these assumptions have been limited to individual studies of very large ($>$ 500 ha) lakes or studies on small numbers ($<$ 100) of lakes \cite{johanssonNewApproachesModelling2007}.
@@
 its “visual distance to lake center”).
+The reason for this is to account for lakes where the centroid is located on land such as an island or peninsual and doesn't intersect lake bathymetry.
 For these calculations, we used the \texttt{polylabelr} R package \cite{larssonPolylabelrFindPole2019}, which interfaces with the Mapbox pole of inaccessibility algorithm \cite{agafonkinJSLibraryFinding2019}.
-We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model ($\sim$15x15m grain) accessed using the \texttt{elevatr} R package \cite{hollisterElevatrAccessElevation2017} and computed using the terrain function in the \texttt{raster} R package \cite{hijmansRasterGeographicData2019}.
+We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point:
+
+$<$Insert equation$>$
+
+We calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model ($\sim$15x15m grain) accessed using the \texttt{elevatr} R package \cite{hollisterElevatrAccessElevation2017}.
+More specifically, we calculated nearshore land slope by first passing a 3x3 moving window over the lake buffer to calculate slope using Horn's algorithm via the \texttt{terrain} function in the \texttt{raster} R package \cite{hijmansRasterGeographicData2019}.
+Reported slope values are the mean of all points in the buffer.
 

### 2020-11-04 manuscript/agujournaltemplate.tex
minor text edits

@@
 These differences are reflected in variation among lakes in terms of biological productivity \cite{qinWaterDepthUnderpins2020} and rates of greenhouse gas production \cite{liSignificantContributionLake2020}.
-However, because measured depth data is only available for a tiny fraction of lakes and $<$ 25\% of all lakes in our study footprint, our ability to understand and predict depth-dependent processes is limited.
+However, because measured depth data is only available for a tiny fraction of lakes (including $<$ 25\% of all lakes in our study footprint), our ability to understand and predict depth-dependent processes is limited.
 The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \cite{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
-Given the limited prediction accuracy of prior depth prediction efforts ($\pm$ 6-7 m), a major focus has been on improving accuracy using a variety of strategies such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings (e.g.
-fitting different models for distinct size classes) among lakes \cite{sobekPredictingDepthVolume2011}.
-These efforts have met with limited improvements in prediction accuracy.
+Given the limited prediction accuracy of prior depth prediction efforts ($\pm$ 6-7 m), a major focus has been on improving accuracy using strategies such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings (e.g.
+fitting different models for distinct size classes) among lakes \cite{caelVolumeMeanDepth2017, sobekPredictingDepthVolume2011}.
+Unfortunately, the results of these efforts have been mixed from a pure prediction standpoint.
 
-One inuitive approach for predicting lake depth involves using a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \cite{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
+One intuitive approach for predicting lake depth involves using a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \cite{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
 All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae.
@@
 For instance, lake cross-section shapes have been shown to vary from narrow "convex" forms to outstretched "concave" forms \cite{hakansonLakeFormLake1977}.
-Because tests of geometric model assumptions require bathymetric map data, which is available for a tiny fraction of lakes (about 15\% of all lakes in our study footprint), existing evidence may not be applicable to all lakes.
+Because tests of geometric model assumptions require bathymetric map data, which is only available for a tiny fraction of lakes (including about 15\% of all lakes in our study footprint), existing evidence may not be applicable to all lakes.
 The few studies that have tested these assumptions have been limited to individual studies of very large ($>$ 500 ha) lakes or studies on small numbers ($<$ 100) of lakes \cite{johanssonNewApproachesModelling2007}.
@@
 its “visual distance to lake center”).
-The reason for this is to account for lakes where the centroid is located on land such as an island or peninsual and doesn't intersect lake bathymetry.
+We did this to account for lakes where the centroid does not intersect lake bathymetry because it is located within an embedded island or peninsula.
 For these calculations, we used the \texttt{polylabelr} R package \cite{larssonPolylabelrFindPole2019}, which interfaces with the Mapbox pole of inaccessibility algorithm \cite{agafonkinJSLibraryFinding2019}.
-We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point:
-
-$<$Insert equation$>$
+We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point.
+Where maximum lake depth is a point measurement derived from a bathymetric surface and distance to the deepest point is the smallest straight-line distance from the waterbody perimeter.
 
 We calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model ($\sim$15x15m grain) accessed using the \texttt{elevatr} R package \cite{hollisterElevatrAccessElevation2017}.
-More specifically, we calculated nearshore land slope by first passing a 3x3 moving window over the lake buffer to calculate slope using Horn's algorithm via the \texttt{terrain} function in the \texttt{raster} R package \cite{hijmansRasterGeographicData2019}.
-Reported slope values are the mean of all points in the buffer.
+Slope computations proceeded by passing a 3x3 moving window over the 100-m buffer to calculate the slope at each point using Horn's algorithm via the \texttt{terrain} function in the \texttt{raster} R package \cite{hijmansRasterGeographicData2019}.
+Reported nearshore land slope values are the mean of all points in the buffer.
 
@@
 hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S3).
-We further classified lakes using the output of a machine learning algorithm to assign a probability to each lake as to whether it is a reservoir or a natural lake.
-For our purposes, we determined a lake to be a reservoir if the classification probability was 0.75 or greater.
-Our reservoir classification data defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water, beyond simply controlling water level.
-It makes no distinction between different dam types or dam heights.
+
+We further classified lakes using the output of a deep convolutional neural network model trained on sattelite images labeled according to whether there was visual evidence of a water control structure significantly impacting flow.
+This model had an overall validation accuracy of 81\% and produced a probability for each waterbody as to whether it is a reservoir or a natural lake.
+For our purposes, we set a conservative classification probability threshold of 0.75 to determine whether a lake would be considered a  reservoir.
+We tested the sensitivity of our model output relative to variations of this threshold and found little substantive difference in results.
+Note that our reservoir classification defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water.
+It makes no distinction between different dam types, heights, or uses/purposes.
 

### 2020-11-05 manuscript/agujournaltemplate.tex
more minor text edits

@@
 
-All lake bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) NHD \cite{usgsNationalHydrographyDataset2019} waterbodies such that source data and bathymetry surface outputs were clipped to the area of each lake polygon.
+All lake bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) National Hydrography Dataset \cite{usgsNationalHydrographyDataset2019} waterbodies such that source data and bathymetry surface outputs were clipped to the area of each lake polygon.
 We restricted the lakes in our study to those with an area of at least 4 ha and a maximum depth of at least 0.3 m (1 ft).
@@
 For these calculations, we used the \texttt{polylabelr} R package \cite{larssonPolylabelrFindPole2019}, which interfaces with the Mapbox pole of inaccessibility algorithm \cite{agafonkinJSLibraryFinding2019}.
-We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point.
-Where maximum lake depth is a point measurement derived from a bathymetric surface and distance to the deepest point is the smallest straight-line distance from the waterbody perimeter.
+We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point where maximum lake depth is a point measurement derived from a bathymetric surface and distance to the deepest point is the smallest straight-line distance from that point to the waterbody perimeter.
 
@@
 
-We further classified lakes using the output of a deep convolutional neural network model trained on sattelite images labeled according to whether there was visual evidence of a water control structure significantly impacting flow.
+We further classified lakes using the output of a deep convolutional neural network model trained on satellite images labeled according to whether there was visual evidence of a water control structure significantly impacting flow.
 This model had an overall validation accuracy of 81\% and produced a probability for each waterbody as to whether it is a reservoir or a natural lake.
@@
 where sinuous lakes have larger values of shoreline development and circular lakes have smaller values of shoreline development.
-Watershed to lake area ratio is an approximation of water residence time and is defined as watershed area divided by lake area.
+Watershed to lake area ratio is an approximation of water residence time and is defined as watershed area divided by lake area \cite{timmsGeomorphologyLakeBasins2009}.
 
@@
 \noindent
-Prior studies using geometric models to predict lake depth include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy.
+Prior studies using geometric models to predict lake depth include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy  \cite{hollisterPredictingMaximumLake2011, yigzawNewGlobalStorage2018}.
 For our purposes, this offset modeling enabled us to test our expectations that prediction error would be different among different formulations of $depth_{geometric}$ and among different lake types.
@@
 This finding is consistent with \citeA{heathcotePredictingBathymetricFeatures2015} whos results suggest that in-lake slopes are shallower compared to the surrounding land.
-The shallow nature of in-lake slopes is likley a function of erosion and sediment transport processes \cite{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}.
+The reason for these shallow in-lake slopes may be related to slope-induced turbidity currents which distribute sediment from shallow high-energy areas of lakes to deep low-energy areas \cite{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}.
+The strength of such sediment focusing is likely greater in "younger" lakes with steeper slopes leading to a smoothing of their bathymetry over time \cite{blaisInfluenceLakeMorphometry1995}.
 
@@
 Using bathymetry data on approximately 5,000 lakes, we show that proxy slope measures are not representative of true in-lake slope and this leads to inaccuracies in predicting the depth of concave and reservoir lakes.
-These innaccuracies suggest that caution is warranted in using geometric models for depth prediction in unsampled lakes.
+These inaccuracies suggest that caution is warranted in using geometric models for depth prediction in unsampled lakes.
 Despite these apparent biases, overall prediction accuracy was equivalent to that of prior depth prediction studies ($\pm$ 6-7m).

### 2020-11-05 manuscript/agujournaltemplate.tex
fix internal figure numbering

@@
 For slope measures, most lakes had lower magnitude (i.e.
-shallower) nearshore land slope compared to true in-lake slope (Figure \ref{fig3}).
+shallower) nearshore land slope compared to true in-lake slope (Figure \ref{fig2}).
 Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes whereas proxy nearshore land slope is not representative of true in-lake slope.
@@
   \caption{Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
-A best-fit line and coefficient of determination is shown to illustrate representativeness.}\label{fig3}
+A best-fit line and coefficient of determination is shown to illustrate representativeness.}\label{fig2}
 \end{figure}
@@
 More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model ($R^2$ = 0.26, RMSE = 6.87m).
-Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure \ref{fig4}).
-We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure \ref{fig4}).
+Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure \ref{fig3}).
+We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure \ref{fig3}).
 
@@
   \end{center}
-  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig4}
+  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig3}
 \end{figure}

### 2020-11-09 figures/01_nearshore.Rmd
stash nearshore exploration figures

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 01_nearshore, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide', fig.height=3.5}
# set llid
flist <- list.files("../data/elevatr", pattern = "\\d*.tif",
                    full.names = TRUE, include.dirs = TRUE)
ll_id           <- 2294
max_buffer_dist <- 100 # to match Hollister (2011)

bathy_pnts <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
ll_poly    <- query_gis("LAGOS_NE_All_Lakes_4ha", "lagoslakeid", ll_id)
ll_iws     <- query_gis("IWS", "lagoslakeid", ll_id)
elev       <- raster(flist[grep(ll_id, flist)])

deepest_pnt         <- dplyr::filter(bathy_pnts, llid == ll_id)
r                   <- raster(paste0("../data/", 
                                     tolower(deepest_pnt$state), "_bathy/", ll_id, ".tif"))
st_crs(deepest_pnt) <- st_crs(r)

elev    <- projectRaster(elev, crs = st_crs(r)$proj4string)
ll_poly <- st_transform(ll_poly, st_crs(r))
ll_iws  <- st_transform(ll_iws, st_crs(r))

# approximates functions in lakemorpho package
elev      <- mask(elev, as_Spatial(ll_iws))
elev      <- mask(elev, as_Spatial(st_buffer(ll_poly, max_buffer_dist)))
elev <- crop(elev, as_Spatial(st_as_sfc(st_bbox(st_buffer(ll_poly, 100)))))
slope     <- terrain(elev, "slope")

slope_mean <- mean(slope@data@values, na.rm = TRUE) * res(elev)[1]
slope_max  <- max(slope@data@values, na.rm = TRUE) * res(elev)[1]

# pull slope on-line from deepest point
shore_pnt <- sf::st_nearest_points(deepest_pnt, st_transform(
  st_cast(ll_poly, "MULTILINESTRING"), st_crs(r)))
ll_buf <- st_buffer(ll_poly, 100)

buffer_line <- sf::st_nearest_points(shore_pnt,
                               st_transform(st_cast(ll_buf, "MULTILINESTRING"), st_crs(r)))
slope_online <- raster::extract(slope, as_Spatial(buffer_line))
slope_online <- mean(slope_online[[1]], na.rm = TRUE) * res(slope)[2]

res <- list(slope_mean = slope_mean, slope_max = slope_max,
     slope_online = slope_online) %>%
  bind_rows()

ggplot() + 
  geom_sf(data = st_transform(ll_buf, st_crs(r)), size = 1.5) +
  geom_sf(data = ll_poly, size = 1.5) +
  geom_raster(data = as.data.frame(terrain(r, "slope"), xy = TRUE),
              aes(x = x, y = y, fill = slope)) +
  geom_raster(data = as.data.frame(slope, xy = TRUE),
              aes(x = x, y = y, fill = slope)) +
  scale_fill_gradient(na.value = NA) +
  geom_sf(data = deepest_pnt) +
  geom_sf(data = shore_pnt) +
  geom_sf(data = buffer_line, color = "red") +
  theme_void()
```

```{r slope_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}

# load csv file from 00_get_nearshore
dt <- read.csv("data/00_geometry/nearshore.csv", 
               stringsAsFactors = FALSE) %>%
    dplyr::filter(inlake_slope < 0.25) %>%
  dplyr::filter(slope_online_mean < 0.6)

cowplot::plot_grid(
  ggplot(data = dt) + 
    geom_point(aes(x = slope_mean, y = slope_online_mean)) +
    geom_abline(aes(slope = 1, intercept = 0)), 
   ggplot(data = dt) + 
    geom_point(aes(x = slope_mean, y = inlake_slope)) +
    geom_abline(aes(slope = 1, intercept = 0)), 
   ggplot(data = dt) + 
    geom_point(aes(x = slope_online_mean, y = inlake_slope)) +
    geom_abline(aes(slope = 1, intercept = 0))
)

plot(dt$inlake_slope, dt$slope_mean,
     xlim = c(0, 2), ylim = c(0, 2))
plot(dt$inlake_slope, dt$slope_online,
     xlim = c(0, 2), ylim = c(0, 2))

# plot(dt$slope_mean, dt$slope_online)
abline(0, 1)

cor.test(dt$slope_mean, dt$inlake_slope)
cor.test(dt$slope_online, dt$inlake_slope)

hist(dt$inlake_slope / dt$slope_online)

```

### 2020-11-10 manuscript/si_template_2019.tex
relabel appendix figs

@@
       \caption{Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
-Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom.}\label{fig2}
+Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom.}\label{figS1}
     \end{figure}
@@
       \begin{center}\includegraphics[width=0.65\textwidth]{../figures/00_map_bathy-1}\end{center}
-      \caption{Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.}\label{figs1}
+      \caption{Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.}\label{figS2}
       \end{figure}
@@
       \noindent\includegraphics[width=0.85\textwidth]{../figures/01_hypsography-1}
-      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figs4}
+      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figS3}
       \end{figure}
@@
 A dashed 1:1 line is shown for comparison.
-Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figs5}
+Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figS4}
       \end{figure}
@@
 See Equation 1 for a definition of geometric max depth.
-HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figs6}
+HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figS5}
       \end{figure}
@@
             \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product.
-The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figs7}
+The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS6}
       \end{figure}
@@
       \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
-      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figs3}
+      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS7}
       \end{figure}
@@
       \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product.
-For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figs2}
+For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figS8}
       \end{figure}
@@
 Predictor variables for computing random forest offsets (Equation 2) are printed in bold face.
-Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
+Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{tableS1}
 \centering
@@
 
-More Information and Advice:
+% More Information and Advice:
 
@@
 
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
+% Math coded inside display math mode \[ ...\]
+%  will not be numbered, e.g.,:
+%  \[ x^2=y^2 + z^2\]
+%
+%  Math coded inside \begin{equation} and \end{equation} will
+%  be automatically numbered, e.g.,:
+%  \begin{equation}
+%  x^2=y^2 + z^2
+%  \end{equation}
 
@@
 % as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
+% \begin{eqnarray}
+%   x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
+%         && + (y - y_{0}) \sin \Theta  \nonumber \\
+%   y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
+%         && + (y - y_{0}) \cos \Theta.
+% \end{eqnarray}
 

### 2020-11-10 figures/01_geometry.Rmd
prototype slope x shape x depth x area comparison figure #21

@@
 ```
+
+```{r slope_by_area_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5.5, fig.height=4.2}
+
+nearshore     <- read.csv("../data/00_geometry/nearshore.csv",
+                      stringsAsFactors = FALSE) %>%
+                  dplyr::select(-inlake_slope)
+dt_raw        <- read.csv("../data/lagosus_depth_predictors.csv", 
+                      stringsAsFactors = FALSE) %>%
+                  left_join(nearshore, b = c("lagoslakeid" = "llid")) %>% 
+                  dplyr::filter(shape_class %in% c("concave", "convex")) %>% 
+                  dplyr::filter(inlake_slope < 0.6)
+
+breaks_dt <- classInt::classIntervals(dt_raw$lake_waterarea_ha, n = 4)$brks
+dt        <- dt_raw %>%
+  dplyr::mutate(x_bins = cut(dt_raw$lake_waterarea_ha,
+      breaks = breaks_dt)) %>%
+  dplyr::filter(!is.na(x_bins))
+dt <- dt %>%
+  dplyr::mutate(x_bins = forcats::fct_recode(x_bins, 
+        !!!setNames(levels(dt$x_bins), 
+                    paste0("< ", as.character(round(breaks_dt[-1], 1))))
+      )
+    )
+
+gg_area <- ggplot(data = dt) +
+   geom_point(aes(x = slope_mean, y = inlake_slope, color = x_bins)) +
+   geom_smooth(aes(slope_mean, inlake_slope, color = x_bins),
+               method = "lm", se = FALSE) + 
+  facet_wrap(~shape_class) +
+    ggtitle("Area (ha)")
+
+breaks_dt <- classInt::classIntervals(dt_raw$lake_maxdepth_m, n = 4)$brks
+dt        <- dt_raw %>%
+  dplyr::mutate(x_bins = cut(dt_raw$lake_maxdepth_m,
+      breaks = breaks_dt)) %>%
+  dplyr::filter(!is.na(x_bins))
+dt <- dt %>%
+  dplyr::mutate(x_bins = forcats::fct_recode(x_bins, 
+        !!!setNames(levels(dt$x_bins), 
+                    paste0("< ", as.character(round(breaks_dt[-1], 1))))
+      )
+    )
+
+gg_depth <- ggplot(data = dt) +
+   geom_point(aes(x = slope_mean, y = inlake_slope, color = x_bins)) +
+   geom_smooth(aes(slope_mean, inlake_slope, color = x_bins),
+               method = "lm", se = FALSE) +
+    facet_wrap(~shape_class) +
+    ggtitle("Max Depth (m)")
+
+theme_opts <- cowplot::theme_cowplot() +
+  theme(legend.position = "bottom", 
+        legend.title = element_blank(), 
+        panel.spacing = unit(1, "lines")) 
+cowplot::plot_grid(
+  gg_area + theme_opts +
+    guides(color = guide_legend(nrow = 2, byrow = TRUE)),
+  gg_depth + theme_opts + 
+    theme(axis.title.y = element_blank(), 
+          axis.text.y = element_blank(), 
+          axis.line.y = element_blank(),
+          axis.ticks.y = element_blank()) +
+    guides(color = guide_legend(nrow = 2, byrow = TRUE)),
+  rel_widths = c(1, 0.87))
+
+```

### 2020-11-10 figures/01_geometry.Rmd
prototype Figure 2 best-fit lines, #20

@@
                                 parse = TRUE, size = text_size) +
+  annotate("text", x = 2100, y = 150, 
+           label = get_formula(
+             lm(dist_viscenter ~ dist_deepest, data = st_drop_geometry(dt))), 
+           size = text_size) +
   labs(color = "") +
@@
                                 parse = TRUE, size = text_size) +
+  annotate("text", x = 0.34, y = 0.01, 
+           label = get_formula(
+             lm(inlake_slope.y ~ slope_mean, data = test)), 
+           size = text_size) +
   # geom_abline(slope = 1, intercept = 0) +

### 2020-11-10 scripts/99_utils.R

@@ -359,3 +359,14 @@ gg_quantdot <- function(dt, grp, var){
                         ymin = .data$`5%`, ymax = .data$`95%`)) +
     ylab(var)
 }
+
+# https://stackoverflow.com/a/64376355/3362993
+# get_formula(lm(dist ~ speed, data = cars))
+get_formula <- function(model) {
+  broom::tidy(model)[, 1:2] %>%
+    mutate(sign = ifelse(sign(estimate) == 1, ' + ', ' - ')) %>% #coeff signs
+    mutate_if(is.numeric, ~ abs(round(., 2))) %>% #for improving formatting
+    mutate(a = ifelse(term == '(Intercept)', paste0('y ~ ', estimate), paste0(sign, estimate, ' * ', term))) %>%
+    summarise(formula = paste(a, collapse = '')) %>%
+    as.character
+}

### 2020-11-22 manuscript/agujournaltemplate.tex
cite getirana paper, note table source files

@@
 
-One intuitive approach for predicting lake depth involves using a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \cite{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
+One intuitive approach for predicting lake depth involves using a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \cite{getiranaDerivingThreeDimensional2018, hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
 All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae.
@@
 \begin{figure}
-\noindent\includegraphics[width=0.85\textwidth]{../figures/slope_diagram_new}
+\noindent\includegraphics[width=0.79\textwidth]{../figures/slope_diagram_new}
 \caption{Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
@@
 The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of $depth_{geometric}$ and among different lake types.
-Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table \ref{table2}).
+Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table \ref{table1}).
 
@@
 
-The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table \ref{table2}).
+The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table \ref{table1}).
 More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model ($R^2$ = 0.26, RMSE = 6.87m).
@@
 \begin{table}[h]
-  \caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table2}
+  \caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table1}
   \centering
-  % \setlength\tabcolsep{1.5pt} % default value: 6pt
+  % scripts/05_depth_model.R
+  % data/01_depth_model/depth_grid_metrics.rds
+  % tables/02_model_metrics.Rmd
 \begin{tabular}{llll}

### 2020-11-23 figures/01_geometry.Rmd
propagate bathy metric name changes, calculate mape model metric, closes #17

@@
 dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
-  mutate(llid = as.character(llid))
+  mutate(llid = as.character(llid)) %>%
+  dplyr::select(-contains("inlake_slope"))
 dt <- left_join(dt, st_drop_geometry(bp), 
@@
 
+slope_dist_str <- "online_mean|slope_pnt|inlake_slope|dist_v|dist_d|dists_|slope_mean"
 # jsta::jheatmap
@@
   dplyr::select(dt, 
-                slope_online_mean,
-                inlake_slope_mean,
-                slope_mean, 
-                slope_mean_ne,
-                dist_deepest, 
-                dist_viscenter, 
-                maxdepth,
-                meandepth))
+   matches(slope_dist_str))
+   # maxdepth,
+   # meandepth)
+  )
 
 GGally::ggpairs(dplyr::select(dt, 
-                inlake_slope_mean, inlake_slope.y,
-                slope_online_mean,
-                slope_mean, 
+                inlake_slope_mean, inlake_slope_pnt,
+                nearshore_slope_online_mean,
+                nearshore_slope_mean, 
                 # dist_deepest, 
@@
 # which(dt$inlake_slope.x > 0.4)
-dt <- dplyr::filter(dt, inlake_slope.y < 0.4)
+dt <- dplyr::filter(dt, inlake_slope_pnt < 0.4)
 
 cowplot::plot_grid(
-  gg_scatter_slope("slope_mean", "maxdepth", dt),
-  gg_scatter_slope("inlake_slope.y", "maxdepth", dt), 
-  gg_scatter_slope("slope_mean", "inlake_slope.y", dt)
+  gg_scatter_slope("nearshore_slope_mean", "maxdepth", dt),
+  gg_scatter_slope("inlake_slope_pnt", "maxdepth", dt), 
+  gg_scatter_slope("nearshore_slope_mean", "inlake_slope_pnt", dt)
   )
@@
 lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
-  mutate(lagoslakeid = as.character(lagoslakeid))
+  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  dplyr::select(-matches("inlake_slope"), 
+                -matches("dist_|dists_"))
 test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
@@
 gg_slope_shape <- ggplot(data = test) + 
-  geom_point(aes(slope_mean, inlake_slope.y, 
+  geom_point(aes(nearshore_slope_mean, inlake_slope_pnt, 
                              color = shape_class), size = 1) +
-  geom_smooth(aes(slope_mean, inlake_slope.y, 
+  geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt, 
                              color = shape_class), data = dplyr::filter(test, shape_class != "neither"), 
@@
 gg_slope_reservoir <- ggplot(data = dplyr::filter(test, !is.na(reservoir_class))) +
-  geom_point(aes(slope_mean, inlake_slope.y, 
+  geom_point(aes(nearshore_slope_mean, inlake_slope_pnt, 
                                  color = reservoir_class), size = 1) +
-  geom_smooth(aes(slope_mean, inlake_slope.y, 
+  geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt, 
                                  color = reservoir_class), 
@@
 dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
-  mutate(llid = as.character(llid))
+  mutate(llid = as.character(llid)) %>%
+  dplyr::select(-matches("inlake_slope")) # prevent name conflict with bp
 dt <- left_join(dt, st_drop_geometry(bp), 
@@
   dplyr::filter(!is.na(inlake_slope_mean)) %>%
-  dplyr::filter(inlake_slope.y < 0.4)
+  dplyr::filter(inlake_slope_pnt < 0.4)
 lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
-  mutate(lagoslakeid = as.character(lagoslakeid))
+  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  dplyr::select(-matches("inlake_slope"))
 test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
@@
 gg_slope_base <- ggplot(data = test) + 
-  geom_point(aes(slope_mean, inlake_slope.y), size = 1) +
-  geom_smooth(aes(slope_mean, inlake_slope.y), method = "lm", se = FALSE) +
+  geom_point(aes(nearshore_slope_mean, inlake_slope_pnt), size = 1) +
+  geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt), method = "lm", se = FALSE) +
   geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
@@
            label = get_formula(
-             lm(inlake_slope.y ~ slope_mean, data = test)), 
+             lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)), 
            size = text_size) +
@@
                       stringsAsFactors = FALSE) %>%
-                  dplyr::select(-inlake_slope)
+                  dplyr::select(-matches("inlake_slope")) 
 dt_raw        <- read.csv("../data/lagosus_depth_predictors.csv", 
@@
                   dplyr::filter(shape_class %in% c("concave", "convex")) %>% 
-                  dplyr::filter(inlake_slope < 0.6)
+                  dplyr::filter(inlake_slope_pnt < 0.6)
 
@@
 gg_area <- ggplot(data = dt) +
-   geom_point(aes(x = slope_mean, y = inlake_slope, color = x_bins)) +
-   geom_smooth(aes(slope_mean, inlake_slope, color = x_bins),
+   geom_point(aes(x = nearshore_slope_mean, y = inlake_slope_pnt, color = x_bins)) +
+   geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt, color = x_bins),
                method = "lm", se = FALSE) + 
@@
 gg_depth <- ggplot(data = dt) +
-   geom_point(aes(x = slope_mean, y = inlake_slope, color = x_bins)) +
-   geom_smooth(aes(slope_mean, inlake_slope, color = x_bins),
+   geom_point(aes(x = nearshore_slope_mean, y = inlake_slope_pnt, color = x_bins)) +
+   geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt, color = x_bins),
                method = "lm", se = FALSE) +

### 2020-11-23 tables/01_predictors.Rmd
propagate bathy metric name changes, calculate mape model metric, closes #17

@@
                       stringsAsFactors = FALSE) %>%
-  dplyr::select(llid, buffer_slope = slope_mean)
+  dplyr::select(llid, contains(c("slope_mean", "slope_online_mean", 
+                                 "slopes_mean", "slopes_online_mean")))
 dt_all <- read.csv("../data/lagosus_depth_predictors.csv", 
                stringsAsFactors = FALSE) %>%
-  distinct(lagoslakeid, .keep_all = TRUE)
+  distinct(lagoslakeid, .keep_all = TRUE) %>%
+  dplyr::select(-inlake_slope_mean)
 dt_bathy <- dt_all %>%
@@
 # interleave
-dt_combined <- bind_cols(lapply(names(dt_bathy)[2:5], 
+dt_combined <- suppressMessages(
+  bind_cols(lapply(names(dt_bathy)[2:5], 
                            function(i) paste0(dt_bathy[,i], 
-                                              " (", dt_all[,i], ")"))) %>%
+                                              " (", dt_all[,i], ")")))
+  ) %>%
   setNames(names(dt_bathy)[2:5])
@@
 # setwd("figures")
+(slope_distance_vars <- unique(dt_combined$variable)[
+  stringr::str_detect(unique(dt_combined$variable), 
+                     "dist_|e_slope_|dists_|e_slopes")])
+(exclude_vars <- unique(dt_combined$variable)[
+  stringr::str_detect(unique(dt_combined$variable), 
+                     "totalarea|islandperimeter|between|buffer100m|lake_meandepth_m")])
+
 dt <- dt_combined %>%
   dplyr::filter(str_detect(variable, 
-                           "^lake_|^ws_lake_arearatio|^dist_|_slope")) %>%
+                c(paste0(slope_distance_vars, collapse = "|"),
+                  "^lake_|^ws_lake_arearatio"))) %>%
   dplyr::filter(!str_detect(variable,
-                      "totalarea|islandperimeter|between|buffer100m|lake_meandepth_m")) # %>%
+                            paste0(exclude_vars, collapse = "|")))
+                       # %>%
   # mutate(n = dt_bathy$n[which(dt_bathy$variable == "inlake_slope")])
@@
 dt[
-  dt$variable %in% c("dist_deepest", "dist_viscenter", 
-                     "inlake_slope", "buffer_slope"), 2:5] <- 
+  dt$variable %in% slope_distance_vars, 2:5] <- 
   dt[
-  dt$variable %in% c("dist_deepest", "dist_viscenter", 
-                     "inlake_slope", "buffer_slope"), 2:5] %>%
+  dt$variable %in% slope_distance_vars, 2:5] %>%
 apply(2, function(x) stringr::str_replace(x, "\\(([^\\)]+)\\)", "(-)"))
@@
 dt_md <- dt
-dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth")] <- 
+dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)] <- 
   paste0("**", 
-         dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth")],
+         dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)],
          "**")
@@
 # bold Eq 2 predictor variables
-dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")] <- 
+dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)] <- 
   paste0("\\textbf{", 
-         dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")],
+         dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)],
          "}")

### 2020-11-23 tables/02_model_metrics.Rmd
propagate bathy metric name changes, calculate mape model metric, closes #17

@@
   tidyr::separate(model, c("slope", "distance")) %>%
-  dplyr::select(slope, distance, everything(), -.estimator) %>%
-  dplyr::select(-mae)
+  dplyr::select(slope, distance, everything(), -.estimator)
 

### 2020-11-23 scripts/00_get_geometry.R

@@ -220,8 +220,6 @@ res_all <- data.frame()
 fpath_stem   <- "data"
 outname_stem <- "data"
 
-flatten_multipoint <- function(x) list(as.data.frame(st_coordinates(x)))
-
 # MN
 message("Calculating MN geometries...")
 res_mn <- mutate(
@@ -363,9 +361,10 @@ res_all <- dplyr::filter(res_all, dist_deepest > 0.1)
 saveRDS(st_as_sf(res_all),
         "data/00_bathy_depth/bathy_pnts.rds")
 
-# write geometry stats without pnt geometry
-write.csv(select(res_all, -contains("pnt")),
+# write geometry stats without multipoint geometry
+write.csv(dplyr::select(res_all, -matches("pnts_|pnt_")),
           "data/00_bathy_depth/bathy_geometry.csv", row.names = FALSE)
+# test <- read.csv("data/00_bathy_depth/bathy_geometry.csv", stringsAsFactors = FALSE)
 
 # write max depth formatted like other max depth sources
 res_final <- res_all %>%

### 2020-11-23 scripts/00_get_nearshore.R

@@ -183,9 +183,10 @@ res <- lapply(ll_ids, function(x) {
 res <- res %>%
   bind_rows() %>%
   dplyr::mutate(llid = as.character(llid)) %>%
-  left_join(dplyr::select(sf::st_drop_geometry(bathy_pnts),
-                          llid, starts_with("inlake_slope")))
-
+  left_join(mutate(
+    dplyr::select(sf::st_drop_geometry(bathy_pnts),
+                          llid, starts_with("inlake_slope")),
+    llid = as.character(llid)))
 
 # write to csv
 write.csv(res, "data/00_geometry/nearshore.csv", row.names = FALSE)

### 2020-11-23 scripts/04_depth_predictors.R

@@ -2,11 +2,11 @@
 source("scripts/99_utils.R")
 
 # unlink("data/lagosus_depth_predictors.csv")
-lg        <- lagosus_load(modules = "locus")
-lg_ne     <- lagosne_load()
+lg        <- lagosus_load(modules = "locus", versions = 1.1)
+lg_ne     <- lagosne_load("1.087.3")
 # need x_walk to join lagosne slope data
-lg_x_walk <- lagosus_load(modules = "locus")$locus$locus_link %>%
-  dplyr::select(lagosne_lagoslakeid, lagoslakeid, lagosus_centroidstate) %>%
+lg_x_walk <- lg$locus$lake_link %>%
+  dplyr::select(lagosne_lagoslakeid, lagoslakeid, lake_centroidstate) %>%
   dplyr::rename(lagosus_lagoslakeid = lagoslakeid) %>%
   distinct(lagosus_lagoslakeid, .keep_all = TRUE)
 hypso_classes <- read.csv(
@@ -16,25 +16,20 @@ dt_raw <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
   dplyr::left_join(hypso_classes, by = c("lagoslakeid" = "llid")) %>%
   dplyr::select(-contains("waterarea")) %>%
   dplyr::filter(!is.na(lake_maxdepth_m)) %>%
-  left_join(lg$locus$locus_characteristics,
+  left_join(lg$locus$lake_characteristics,
             by = "lagoslakeid") %>%
-  left_join(dplyr::select(lg$locus$locus_information,
+  left_join(dplyr::select(lg$locus$lake_information,
                           -matches("\\b(?!hu4)\\w*zoneid\\b", perl = TRUE),
                           # -contains("name"),
                           -contains("reachcode"),
                           -lake_states, -contains("decdeg"),
                           -contains("namegnis")),
                           by = "lagoslakeid") %>%
-  left_join(dplyr::select(lg$locus$locus_nws, contains("area"),
+  left_join(dplyr::select(lg$locus$lake_watersheds, contains("area"),
                           contains("perimeter"), contains("width"),
                           contains("length"), contains("orientation"),
                           "lagoslakeid"),
                           by = c("lagoslakeid")) %>%
-  left_join(dplyr::select(lg$locus$locus_ws, contains("area"),
[…]

### 2020-11-23 scripts/05_depth_model.R

@@ -1,16 +1,27 @@
 # setwd("../")
 source("scripts/99_utils.R")
 
+dt_raw <- read.csv("data/lagosne_depth_predictors.csv",
+                   stringsAsFactors = FALSE) %>%
+  # test3 <- test2 %>%
+  dplyr::filter(lagos_effort == "bathymetry") %>%
+  dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::filter(!(shape_class %in% c("neither"))) %>%
+  dplyr::filter(inlake_slope_pnt < 0.25) %>%
+  dplyr::filter(!is.na(reservoir_class))
+# TODO: assert - any(is.na(dt_raw$ws_lake_arearatio))
+
 data_prep <- function(dt_raw,
-                      inlake_slope_var = "inlake_slope",
-                      nearshore_slope_var = "slope_mean",
+                      inlake_slope_var = "inlake_slope_pnt",
+                      nearshore_slope_var = "nearshore_slope_mean",
                       inlake_dist_var = "dist_deepest",
                       out_path = "data/depth_predictors.csv"){
 
   nearshore <- read.csv("data/00_geometry/nearshore.csv",
                         stringsAsFactors = FALSE) %>%
-    dplyr::select(lagoslakeid = llid, slope_mean = all_of(nearshore_slope_var))
-  dt_raw <- left_join(dt_raw, nearshore) %>%
+    dplyr::select(lagoslakeid = llid, slope_mean = all_of(nearshore_slope_var),
+                  inlake_slope = all_of(inlake_slope_var))
+  dt_raw <- left_join(dt_raw, nearshore, by = "lagoslakeid") %>%
     dplyr::filter(!is.na(slope_mean))
 
   # calculate the trig products as covariates
@@ -114,15 +125,13 @@ fit_model <- function(maxdepth, dt_train, dt_test){
   list(res = res, fit = fit1)
 }
 
-dt_raw <- read.csv("data/lagosne_depth_predictors.csv",
-               stringsAsFactors = FALSE) %>%
-  # test3 <- test2 %>%
-  dplyr::filter(lagos_effort == "bathymetry") %>%
-  dplyr::filter(!is.na(shape_class)) %>%
-  dplyr::filter(!(shape_class %in% c("neither"))) %>%
-  dplyr::filter(inlake_slope < 0.25) %>%
[…]

### 2020-11-23 scripts/99_utils.R

@@ -370,3 +370,5 @@ get_formula <- function(model) {
     summarise(formula = paste(a, collapse = '')) %>%
     as.character
 }
+
+flatten_multipoint <- function(x) list(as.data.frame(st_coordinates(x)))

### 2020-11-23 figures/01_contrasts.Rmd
more name change propagation

@@
   dplyr::filter(shape_class != "neither") %>%  
-  contrast_plot(grp = "shape_class", varb = "inlake_slope", 
+  contrast_plot(grp = "shape_class", varb = "inlake_slope_pnt", 
                 label = "In-lake slope (m/m)", plot_type = "pointrange")
@@
   dplyr::filter(!is.na(reservoir_class)) %>%  
-  contrast_plot(grp = "reservoir_class", varb = "inlake_slope", 
+  contrast_plot(grp = "reservoir_class", varb = "inlake_slope_pnt", 
                 label = "In-lake slope (m/m)", plot_type = "pointrange")
@@
   dplyr::filter(shape_class != "neither") %>%  
-  contrast_plot(grp = "shape_class", varb = "lake_shorelinedevfactor_nounits", 
+  contrast_plot(grp = "shape_class", varb = "lake_shorelinedevfactor", 
                 label = "Shoreline Dev.", plot_type = "pointrange")
@@
     st_transform(
-      LAGOSUS::coordinatize(lg, crs = 4326), 
+      LAGOSUS::coordinatize(dt = lg, crs_in = 4326), 
       LAGOSUSgis:::albers_conic()),

### 2020-11-23 manuscript/figures.Rmd
rm predictor table to sync with ms

@@
 
+<!-- \captionsetup[figure]{name=Table} -->
+<!-- \captionsetup[figure]{labelformat=empty} -->
+<!-- ```{r 01_predictors, fig.cap = "Table 1: Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
+Predictor variables for computing random forest offsets (Eq 2) are printed in bold face.
+Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'} -->
+<!-- knitr::include_graphics("../tables/01_predictors.pdf") -->
+<!-- ``` -->
+
+<!-- \captionsetup[figure]{name=Figure} -->
+<!-- \captionsetup[figure]{labelformat=default} -->
+<!-- \setcounter{figure}{0} -->
+
 ![Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
 Geometric depth calculated via Equation 1 requires a single distance and slope metric.](../figures/slope_diagram_new.pdf){width=75%}
-
-```{r lake_shape, echo=FALSE, out.extra='', fig.align='center', out.width="4in",fig.cap="Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
-Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom."}
-knitr::include_graphics("../figures/lake_shape.pdf")
-```
-
-\captionsetup[figure]{name=Table}
-\captionsetup[figure]{labelformat=empty}
-```{r 01_predictors, fig.cap = "Table 1: Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
-Predictor variables for computing random forest offsets (Eq 2) are printed in bold face.
-Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../tables/01_predictors.pdf")
-```
-\captionsetup[figure]{name=Figure}
-\captionsetup[figure]{labelformat=default}
-\setcounter{figure}{2}
 

### 2020-11-23 figures/01_geometry.Rmd
format best-fit equations, closes #20

@@
   geom_smooth(method = "lm", se = FALSE) +
-  annotate("text", x = 2550, y = 300, label = "R^2 == 0.80",
-                                parse = TRUE, size = text_size) +
-  annotate("text", x = 2100, y = 150, 
+  annotate("text", x = 2550, y = 400, label = "R^2 == 0.80",
+                                parse = TRUE, size = text_size-0.7) +
+  annotate("label", x = 1800, y = 150, 
            label = get_formula(
              lm(dist_viscenter ~ dist_deepest, data = st_drop_geometry(dt))), 
-           size = text_size) +
+           size = text_size-0.85, fill = "white", label.size = NA) +
   labs(color = "") +
@@
   alpha = 1, fill = "white") +
-  annotate("text", x = 0.34, y = 0.03, label = "R^2 == 0.17",
-                                parse = TRUE, size = text_size) +
-  annotate("text", x = 0.34, y = 0.01, 
+  annotate("text", x = 0.34, y = 0.05, label = "R^2 == 0.17",
+                                parse = TRUE, size = text_size-0.7) +
+  annotate("label", x = 0.27, y = 0.01, 
            label = get_formula(
              lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)), 
-           size = text_size) +
+           size = text_size-0.85, fill = "white", label.size = NA) +
   # geom_abline(slope = 1, intercept = 0) +

### 2020-11-23 figures/01_geometry.Rmd
prototype slope alternative scatter plots

--- 2020-11-23 figures/01_geometry.Rmd
@@
 ```{r 01_geometry_base, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
+# setwd("figures")
 theme_opts <- cowplot::theme_cowplot() + 
@@
 
+rsq <- round(broom::glance(lm(dist_viscenter ~ dist_deepest, data = dt))$r.squared, 2)
 gg_dist_base <- st_drop_geometry(dt) %>%
@@
   geom_smooth(method = "lm", se = FALSE) +
-  annotate("text", x = 2550, y = 300, label = "R^2 == 0.80",
-                                parse = TRUE, size = text_size) +
-  annotate("text", x = 2100, y = 150, 
+  annotate("text", x = 2550, y = 400, label = paste0("R^2 == ", rsq),
+                                parse = TRUE, size = text_size-0.7) +
+  annotate("label", x = 1800, y = 150, 
            label = get_formula(
              lm(dist_viscenter ~ dist_deepest, data = st_drop_geometry(dt))), 
-           size = text_size) +
+           size = text_size-0.85, fill = "white", label.size = NA) +
   labs(color = "") +
@@
 
+rsq <- round(broom::glance(lm(nearshore_slope_mean ~ inlake_slope_pnt, data = test))$r.squared, 2)
 gg_slope_base <- ggplot(data = test) + 
@@
   alpha = 1, fill = "white") +
-  annotate("text", x = 0.34, y = 0.03, label = "R^2 == 0.17",
-                                parse = TRUE, size = text_size) +
-  annotate("text", x = 0.34, y = 0.01, 
+  annotate("text", x = 0.34, y = 0.05, label = paste0("R^2 == ", rsq),
+                                parse = TRUE, size = text_size-0.7) +
+  annotate("label", x = 0.27, y = 0.01, 
            label = get_formula(
              lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)), 
-           size = text_size) +
+           size = text_size-0.85, fill = "white", label.size = NA) +
   # geom_abline(slope = 1, intercept = 0) +
@@
 
+```
+
+```{r 01_dist_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
+# fig 2A (dist) alternatives
+#   gg_dist_base <- dist_viscenter ~ dist_deepest
+#   gg_dists_base <- dist_viscenter ~ dists_deepest
+dt$dists_deepest <- sapply(dt$dists_deepest, as.numeric)
+rsq <- round(broom::glance(lm(dist_viscenter ~ dists_deepest, data = dt))$r.squared, 2)
+
+gg_dists_base <- dt %>%
+  # dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::filter(dist_viscenter >= dist_deepest) %>%
+  dplyr::filter(dist_deepest < 3200) %>%
+  ggplot(aes(x = dists_deepest, y = dist_viscenter)) + 
+  geom_point(size = 1) +
+  geom_smooth(method = "lm", se = FALSE) +
+  annotate("text", x = 1800, y = 400, label = paste0("R^2 == ", rsq),
+                                parse = TRUE, size = text_size-0.7) +
+  annotate("label", x = 1200, y = 150, 
+           label = get_formula(
+             lm(dist_viscenter ~ dists_deepest, data = dt)), 
+           size = text_size-0.85, fill = "white", label.size = NA) +
+  labs(color = "") +
+  xlab("Deepest point distance (m)") +
+  ylab("Distance to \n lake center (m)") +
+  xlim(0, 2000) +
+  theme_opts
+
+cowplot::plot_grid(gg_dist_base, gg_dists_base)
+```
+
+```{r 01_slope_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
+# fig 2B (slope) alternatives
+#   gg_slope_base <- nearshore_slope_mean ~ inlake_slope_pnt
+#   gg_slope_* <- nearshore_slope* ~ inlake_slope*
+
+inlake_slope_alternatives     <- c("inlake_slope_pnt", "inlake_slope_pnts",
+                                   "inlake_slope_mean",
+                                   "inlake_slope_online_mean",
+                                   "inlake_slopes_online_mean")
+nearshore_slope_alternatives  <- c("nearshore_slope_mean",
+                                   "nearshore_slope_online_mean",
+                                   "nearshore_slopes_online_mean")
+slope_alternatives   <- setNames(data.frame(expand.grid(
+  inlake_slope_alternatives,
+  nearshore_slope_alternatives,
+  stringsAsFactors = FALSE
+  )), c("inlake_slope", "nearshore_slope"))
+
+gg_slope_alternative <- function(nearshore_var, inlake_var, dt){
+  # dt <- test
+  # nearshore_var <- "nearshore_slope_mean"
+  # inlake_var <- "inlake_slope_pnt"
+  rsq <- round(broom::glance(
+    lm(as.formula(paste0(nearshore_var, " ~ ", inlake_var)), data = dt)
+    )$r.squared, 2)
+  res <-  ggplot(data = dt) + 
+    geom_point(aes(.data[[nearshore_var]], .data[[inlake_var]]), size = 1) +
+    geom_smooth(aes(.data[[nearshore_var]], .data[[inlake_var]]), 
+                method = "lm", se = FALSE) +
+    geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
+    alpha = 1, fill = "white") +
+    annotate("text", x = 0.34, y = 0.05, label = paste0("R^2 == ", rsq),
+                                  parse = TRUE, size = text_size-0.7) +
+    # annotate("label", x = 0.27, y = 0.01, 
+    #          label = get_formula(
+    #            lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)), 
+    #          size = text_size-0.85, fill = "white", label.size = NA) +
+    # geom_abline(slope = 1, intercept = 0) +
+    xlab(nearshore_var) +
+    ylab(inlake_var) +
+    labs(color = "") +
+    theme_opts
+
+  res
+}
+
+gg_slope_list <- lapply(seq_len(nrow(slope_alternatives)), function(i)
+  gg_slope_alternative(slope_alternatives$nearshore_slope[i],
+                     slope_alternatives$inlake_slope[i], test)
+)
+
+gg_slope_list[c(1:4, 6:9, 11:14)] <- lapply(
+  gg_slope_list[c(1:4, 6:9, 11:14)], function(x) x +
+     theme(axis.title.x = element_blank(), 
+           axis.text.x = element_blank()))
+
+gg_slope_list[c(6:15)] <- lapply(
+  gg_slope_list[c(6:15)], function(x) x +
+     theme(axis.title.y = element_blank(), 
+           axis.text.y = element_blank()))
+
+cowplot::plot_grid(plotlist = gg_slope_list, 
+                   nrow = 5, byrow = FALSE, 
+                   rel_heights = c(rep(0.6, 4), 1,
+                                   rep(0.6, 4), 1,
+                                   rep(0.6, 4), 1))
 ```

### 2020-11-24 figures/01_geometry.Rmd
cleanup fig2b alternatives, #22

@@
                 method = "lm", se = FALSE) +
-    geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
-    alpha = 1, fill = "white") +
-    annotate("text", x = 0.34, y = 0.05, label = paste0("R^2 == ", rsq),
+    # geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
+    # alpha = 1, fill = "white") +
+    annotate("label", x = 0.4, y = 0.52, label = paste0("R^2 == ", rsq),
                                   parse = TRUE, size = text_size-0.7) +
@@
     # geom_abline(slope = 1, intercept = 0) +
-    xlab(nearshore_var) +
-    ylab(inlake_var) +
+    xlab(stringr::str_replace_all(nearshore_var, c("nearshore_" = "", "_mean" = ""))) +
+    ylab(pad_str(stringr::str_replace_all(inlake_var, c("inlake_" = "")), 
+                 18)) +
+    xlim(c(0, 0.6)) + ylim(0, 0.6) +
     labs(color = "") +
-    theme_opts
+    theme_opts +
+    theme(axis.title = element_text(family = "mono"))
 
@@
 }
+
+# pad_str(
+#   stringr::str_replace_all(
+#     slope_alternatives$inlake_slope, c("inlake_" = "")),
+#   18)
 
@@
 )
+
+gg_slope_list[c(1:5)] <- lapply(
+  gg_slope_list[c(1:5)], function(x) x +
+     theme(
+       # axis.title.y = element_blank(), 
+       axis.title.y = element_text(angle = 0)))
 
@@
 
+xlab <- ggdraw() + draw_label("nearshore")
+ylab <- ggdraw() + draw_label("inlake", angle = 90)
+cowplot::plot_grid(
+  ylab,
+cowplot::plot_grid(
 cowplot::plot_grid(plotlist = gg_slope_list, 
                    nrow = 5, byrow = FALSE, 
-                   rel_heights = c(rep(0.6, 4), 1,
-                                   rep(0.6, 4), 1,
-                                   rep(0.6, 4), 1))
+                   rel_heights = c(rep(0.6, 4), 1),
+                   rel_widths = c(rep(1, 1), 
+                                  rep(0.37, 2))), 
+xlab, ncol = 1, rel_heights = c(1, 0.1)), 
+ncol = 2, rel_widths = c(0.1, 1))
 ```

### 2020-11-24 scripts/99_utils.R

@@ -371,4 +371,21 @@ get_formula <- function(model) {
     as.character
 }
 
-flatten_multipoint <- function(x) list(as.data.frame(st_coordinates(x)))
+flatten_multipoint <- function(x){
+  list(as.data.frame(st_coordinates(x)))
+}
+
+# pad_str("a", 3)
+# pad_str(c("a", "b"), 3)[1]
+pad_str <- function(x, n){
+
+  .pad_str <- function(x, n){
+    if(nchar(x) < n){
+      paste0(x, paste0(rep(" ", abs(nchar(x) - n)), collapse = ""))
+    }else{
+      x
+    }
+  }
+  as.character(sapply(x, function(x) .pad_str(x, n)))
+}
+

### 2020-11-24 tables/02_model_metrics.Rmd
add mape to table 1

@@
     TRUE ~ paste0(rmse, " m"))) %>%
+  mutate(mape = case_when(
+    as.character(model) == "maxdepth_true_true" ~ NA_character_,
+    TRUE ~ paste0(mape, " %"))) %>%
   mutate(model = gsub("maxdepth_", "", model)) %>%

### 2020-11-24 tables/02_model_metrics.Rmd
format mape

--- 2020-11-23 tables/02_model_metrics.Rmd
@@
     TRUE ~ paste0(rmse, " m"))) %>%
+  mutate(mape = as.numeric(mape)) %>%
+  mutate(mape = case_when(
+    as.character(model) == "maxdepth_true_true" ~ NA_character_,
+    TRUE ~ paste0(round(mape, 0), " %"))) %>%
   mutate(model = gsub("maxdepth_", "", model)) %>%

### 2020-11-24 tables/01_predictors.Rmd
linting

@@
 # setwd("figures")
-(slope_distance_vars <- unique(dt_combined$variable)[
+slope_distance_vars <- unique(dt_combined$variable)[
   stringr::str_detect(unique(dt_combined$variable), 
-                     "dist_|e_slope_|dists_|e_slopes")])
-(exclude_vars <- unique(dt_combined$variable)[
+                     "dist_|e_slope_|dists_|e_slopes")]
+exclude_vars <- unique(dt_combined$variable)[
   stringr::str_detect(unique(dt_combined$variable), 
-                     "totalarea|islandperimeter|between|buffer100m|lake_meandepth_m")])
+                     "totalarea|islandperimeter|between|buffer100m|lake_meandepth_m")]
 

### 2020-11-24 figures/01_geometry.Rmd
adjust 01_slope_compare figure dimensions

@@
 
-```{r 01_slope_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
+```{r 01_slope_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6.2, fig.height=5.3}
 # fig 2B (slope) alternatives

### 2020-11-24 manuscript/si_template_2019.tex
add response figure and table placeholders

@@
       \begin{center}
+        \includegraphics[width=\textwidth,keepaspectratio]{../figures/01_slope_compare-1}
+      \end{center}
+      \caption{Sensitivity analysis regarding different slope calculation methods.}\label{figS1}
+    \end{figure}
+\clearpage
+
+\begin{figure}
+      \begin{center}
         \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/lake_shape}
@@
 \begin{figure}
+      \begin{center}
+        \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/slope_by_area_depth-1}
+      \end{center}
+      \caption{Compare the mean lake slopes in concave and convex lakes of the same size and max depth.}\label{figS1}
+    \end{figure}
+\clearpage
+
+\begin{figure}
       \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
@@
 
+% source at tables/01_predictors.Rmd
 \begin{table}
@@
 \setlength\tabcolsep{1.5pt} % default value: 6pt
+
 \begin{tabular}{lllll}
@@
 \hline
-  Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4850 (17700)\\
-  \textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4850 (17700)\\
-  \textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4850 (17700)\\
-  \textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4850 (17700)\\
-  \textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4850 (17700)\\
-  \textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4850 (17700)\\
-  \textbf{Watershed-lake ratio} & 7.8 (10) & 3.9 (4.4) & 17 (28) & 4850 (17700)\\
-  Distance to deepest point (m) & 180 (-) & 110 (-) & 290 (-) & 4850 (-)\\
-  Distance to lake center (m) & 240 (-) & 160 (-) & 380 (-) & 4850 (-)\\
-  In-lake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4850 (-)\\
-  Nearshore land slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4850 (-)\\
+\textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4820 (17700)\\
+\textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4820 (17700)\\
+- & 1500 (1100) & 880 (640) & 2500 (2200) & 4820 (17700)\\
+- & 91 (88) & 41 (40) & 140 (140) & 4820 (17700)\\
+- & 710 (560) & 430 (320) & 1200 (1000) & 4820 (17700)\\
+- & 380 (290) & 230 (170) & 610 (500) & 4820 (17700)\\
+- & 0 (0) & 0 (0) & 58 (25) & 4820 (17700)\\
+- & 0 (0) & 0 (0) & 3 (3) & 4820 (17700)\\
+- & 0 (0) & 0 (0) & 1 (1) & 4820 (17700)\\
+- & 26000 (27000) & 23000 (23000) & 27000 (38000) & 4820 (17700)\\
+\textbf{Deepest point distance (m)} & 180 (-) & 110 (-) & 290 (-) & 4820 (-)\\
+\textbf{Visual center distance (m)} & 240 (-) & 160 (-) & 390 (-) & 4820 (-)\\
+- & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4820 (-)\\
+
+- & 0.062 (-) & 0.029 (-) & 0.14 (-) & 4800 (-)\\
+
+- & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4820 (-)\\
+
+- & 0.077 (-) & 0.038 (-) & 0.13 (-) & 4590 (-)\\
+
+- & 0.077 (-) & 0.037 (-) & 0.13 (-) & 4540 (-)\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{Show model metrics for sensitivity analysis. Describe how each different metric was calculated.} \label{tableS2}
+\centering
+\setlength\tabcolsep{1.5pt} % default value: 6pt
+
+\begin{tabular}{lllll}
+\hline
+Variable & Median & Q25 & Q75 & n\\
+\hline
+- & 0.077 (-) & 0.037 (-) & 0.13 (-) & 4540 (-)\\
 \hline

### 2020-11-24 manuscript/figures.Rmd
describe figure 2 best fit lines

@@
 ```{r 01_geometry_base, fig.cap = "Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
-A best-fit line and coefficient of determination is shown to illustrate representativeness.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+A best-fit line and equation is shown to shown to facilitate computation of correction factors for proxy values of lake geometry.
+Coefficients of determination are shown to illustrate representativeness.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 knitr::include_graphics("../figures/01_geometry_base-1.pdf")

### 2020-11-24 manuscript/agujournaltemplate.tex
text edits

@@
 \begin{figure}
-\noindent\includegraphics[width=0.79\textwidth]{../figures/slope_diagram_new}
+\noindent\includegraphics[width=0.74\textwidth]{../figures/slope_diagram_new}
 \caption{Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
@@
 We used our generated bathymetry surfaces to find the location of the deepest point in the lake and we resolved ties by choosing the deepest point that was closest to the center of the lake.
-We calculated the center of the lake not as its centroid but rather by finding the point farthest from the lake shoreline (i.e.
+To account for lakes where the centroid does not intersect lake bathymetry because it is located within an embedded island or peninsula, we calculated the center of the lake not as its centroid but rather by finding the point farthest from the lake shoreline (i.e.
 its “visual distance to lake center”).
-We did this to account for lakes where the centroid does not intersect lake bathymetry because it is located within an embedded island or peninsula.
 For these calculations, we used the \texttt{polylabelr} R package \cite{larssonPolylabelrFindPole2019}, which interfaces with the Mapbox pole of inaccessibility algorithm \cite{agafonkinJSLibraryFinding2019}.
 We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point where maximum lake depth is a point measurement derived from a bathymetric surface and distance to the deepest point is the smallest straight-line distance from that point to the waterbody perimeter.
-
 We calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model ($\sim$15x15m grain) accessed using the \texttt{elevatr} R package \cite{hollisterElevatrAccessElevation2017}.
@@
 Reported nearshore land slope values are the mean of all points in the buffer.
+In addition to the aforementioned methods of calculating in-lake (and nearshore) slopes and distances, we tried 7 alternative methods which are described in Table S2.
 
@@
 We used a geometric model of lakes where basins are treated as cones with a fixed relationship between slope and distance (Equation \ref{eq1}).
+One reason that we used the cone model is that, unlike other idealized shapes, it is well suited for modeling maximum depth and does not require any knowledge of lake volume or mean depth.
 Note that Equation \ref{eq1} is a geometric formula and has no intercept or “coefficients” and it produces a perfect depth estimate given true values of slope and distance.
@@
 The purpose of this standardization was to prevent lakes with extremely long proxy distances from having an outsized impact on model evaluation metrics.
+In addition to comparing among model runs using different metric combinations, we compared among sets of model runs where slope and distance measures were calculated using different methods (Table S2).
 
@@
 Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes whereas proxy nearshore land slope is not representative of true in-lake slope.
+The strong relationship between distance to the center of lakes and distance to the deepest point means that it is possible to compute a "correction factor" to convert between the two measures (Figure \ref{fig2}).
+
 In addition to overall differences between slope and distance measures, we found differences in these relationships among lake shape classes.
@@
 The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table \ref{table1}).
-More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model ($R^2$ = 0.26, RMSE = 6.87m).
+More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.7) and lower prediction error (RMSE = 4.4m, MAPE = 29\%) compared to the proxy slope - true distance model ($R^2$ = 0.32, RMSE = 6.6m, MAPE = 60\%).
 Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure \ref{fig3}).
@@
 \begin{table}[h]
-  \caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table1}
+  \caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination, MAPE = mean absolute percent error) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table1}
   \centering
@@
   % tables/02_model_metrics.Rmd
-\begin{tabular}{llll}
+\begin{tabular}{lllll}
   \hline
-  slope & distance & RMSE & $R^2$\\
+  slope & distance & RMSE & $R^2$ & MAPE\\
   \hline
-  true & true & - & -\\
-  true & proxy & 4.2 m & 0.73\\
-  proxy & true & 6.9 m & 0.26\\
-  proxy & proxy & 6.6 m & 0.31\\
+  true & true & - & - & -\\
+  true & proxy & 4.4 m & 0.70 & 29 \%\\
+  proxy & true & 6.6 m & 0.32 & 60 \%\\
+  proxy & proxy & 6.4 m & 0.35 & 59 \%\\
   \hline
@@
 \noindent
-Models using only proxy variables had prediction error rates (RMSE = 6.6m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting lake depth at broad geographic extents \cite{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
+Models using only proxy variables had prediction error rates (RMSE = 6.4m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting lake depth at broad geographic extents \cite{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
 When only a single proxy measure was used, there was a difference in model sensitivity depending on if it was a horizontal distance measure or a vertical slope measure.
-In the case of a true slope and proxy distance combination, models were more accurate ($\pm$ 4.2m) than even the most accurate of prior studies \cite{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
-Conversely, models using a proxy slope and true distance combination had prediction error rates ($\pm$ 6.9m) of a similar magnitude as that of the baseline proxy-proxy model ($\pm$ 6.6m).
+In the case of a true slope and proxy distance combination, models were more accurate ($\pm$ 4.4m, 29\%) than even the most accurate of prior studies \cite{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
+Conversely, models using a proxy slope and true distance combination had prediction error rates ($\pm$ 6.6m, 60\%) of a similar magnitude as that of the baseline proxy-proxy model ($\pm$ 6.4m, 59\%).
 The greater sensitivity of depth predictions to proxy slope measures relative to proxy distance measures may be explained by the fact that proxy slope measures were a more imperfect representation of true in-lake slopes relative to proxy versus true distances.
@@
 However, parameterization of this model requires data on bathymetry which is not available for all lakes.
-We propose that the error rate of this model ($\pm$ 4.2m) be used as an out-of-sample prediction benchmark for future studies such that they should attempt to match it but not expect to exceed it.
+We propose that the error rate of this model ($\pm$ 4.4m, 29\%) be used as an out-of-sample prediction benchmark for future studies such that they should attempt to match it but not expect to exceed it.
 
@@
 Despite these apparent biases, overall prediction accuracy was equivalent to that of prior depth prediction studies ($\pm$ 6-7m).
-Only our models using a true measure of in-lake slope, which is limited in availability to lakes with bathymetry data and where we already know lake depth, had greater accuracy than that of prior studies ($\pm$ 4.2m).
+Only our models using a true measure of in-lake slope, which is limited in availability to lakes with bathymetry data and where we already know lake depth, had greater accuracy than that of prior studies ($\pm$ 4.4m, 29\%).
 Lack of improved prediction accuracy (short of including data that is unavailable for most lakes) suggests that improved prediction may require new types of data or novel analysis techniques.

### 2020-11-25 tables/02_model_metrics.Rmd
generate alt metrics model table

@@
 ```
+
+```{r 02_alternate_metrics, eval=TRUE}
+
+dt_metrics <- read.csv("../data/01_depth_model/alternatives_metrics.csv",
+                       stringsAsFactors = FALSE) %>%
+  mutate_if(is.numeric, function(x) as.character(round(x, 2))) %>%
+  mutate(proxy_proxy_rmse = paste0(
+    round(as.numeric(proxy_proxy_rmse), 1), " m")) %>%
+  mutate(proxy_proxy_rsq = round(as.numeric(proxy_proxy_rsq), 2)) %>%
+  mutate(proxy_proxy_mape = paste0(
+    round(as.numeric(proxy_proxy_mape), 0), " %")) %>%
+  mutate(inlake_slope = gsub("inlake_", "", inlake_slope)) %>%
+  mutate(nearshore_slope = gsub("nearshore_", "", nearshore_slope)) %>%
+  arrange(proxy_proxy_rmse) %>%
+  setNames(c("Inlake slope", "Nearshore slope", 
+             "Inlake distance", "RMSE", "$R^2$", "MAPE"))
+
+options(knitr.kable.NA = '-')
+knitr::kable(dt_metrics, digits = 2, format = "markdown")
+
+dt_metrics[2,] <- paste0("**", dt_metrics[2,], "**") # highlight row from main text
+clipr::write_clip(
+  gsub("%\\*\\*", "\\\\%\\*\\*",
+  gsub("_", "\\\\_",
+  gsub("%\\\\", "\\\\%\\\\",
+  sub("\\\\midrule", "\\\\hline",
+  sub("\\\\bottomrule", "\\\\hline",
+  sub("\\\\toprule", "\\\\hline",
+      knitr::kable(dt_metrics, digits = 2, format = "latex",
+     booktabs = TRUE, linesep = "", align = c("lllllc"), escape = FALSE)))))))
+   )
+```

### 2020-11-25 manuscript/si_template_2019.tex
format table S2 and fill out caption

@@
 \begin{table}
-\caption{Show model metrics for sensitivity analysis. Describe how each different metric was calculated.} \label{tableS2}
+\caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination, MAPE = mean absolute percent error) for the proxy - proxy combination of geometry metrics (see main text Table 1).
+Each row shows model metrics when proxy and "true" measures are calculated with slight differences from the default (bolded) used in the main text.
+\texttt{slope\_mean} is the mean slope of all inlake or nearshore buffer points.
+\texttt{slope\_pnts} is the average \texttt{slope\_pnt} of all points at maximum depth.
+\texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope).
+\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except for all inlake points at maximum depth.
+\texttt{dists\_deepest} is the same as \texttt{dist\_deepest} except distance is calculated for all points at maximum depth.}  \label{tableS2}
 \centering
 \setlength\tabcolsep{1.5pt} % default value: 6pt
-
-\begin{tabular}{lllll}
+\begin{tabular}{lllllc}
 \hline
-Variable & Median & Q25 & Q75 & n\\
+Inlake slope & Nearshore slope & Inlake distance & RMSE & $R^2$ & MAPE\\
 \hline
-- & 0.077 (-) & 0.037 (-) & 0.13 (-) & 4540 (-)\\
+slope\_pnts & slope\_mean & dists\_deepest & 6.2 m & 0.38 & 58 \%\\
+\textbf{slope\_pnt} & \textbf{slope\_mean} & \textbf{dist\_deepest} & \textbf{6.4 m} & \textbf{0.35} & \textbf{59 \%}\\
+slope\_pnts & slopes\_online\_mean & dist\_deepest & 6.4 m & 0.32 & 61 \%\\
+slope\_online\_mean & slope\_mean & dists\_deepest & 6.5 m & 0.41 & 63 \%\\
+slope\_pnts & slope\_mean & dist\_deepest & 6.7 m & 0.44 & 58 \%\\
+slope\_online\_mean & slope\_mean & dist\_deepest & 6.7 m & 0.36 & 59 \%\\
+slope\_online\_mean & slopes\_online\_mean & dist\_deepest & 6.7 m & 0.32 & 66 \%\\
+slope\_mean & slope\_mean & dists\_deepest & 6.8 m & 0.36 & 59 \%\\
+slope\_pnt & slopes\_online\_mean & dists\_deepest & 6.8 m & 0.25 & 73 \%\\
+slope\_pnts & slope\_online\_mean & dist\_deepest & 6.9 m & 0.3 & 71 \%\\
+slope\_online\_mean & slope\_online\_mean & dist\_deepest & 6.9 m & 0.32 & 68 \%\\
+slope\_online\_mean & slope\_online\_mean & dists\_deepest & 6.9 m & 0.33 & 65 \%\\
+slope\_mean & slopes\_online\_mean & dists\_deepest & 7 m & 0.24 & 65 \%\\
+slope\_mean & slope\_mean & dist\_deepest & 7.1 m & 0.4 & 64 \%\\
+slopes\_online\_mean & slope\_mean & dist\_deepest & 7.1 m & 0.37 & 56 \%\\
+slope\_mean & slope\_online\_mean & dist\_deepest & 7.1 m & 0.3 & 69 \%\\
+slopes\_online\_mean & slopes\_online\_mean & dists\_deepest & 7.2 m & 0.32 & 63 \%\\
+slopes\_online\_mean & slopes\_online\_mean & dist\_deepest & 7.3 m & 0.25 & 64 \%\\
+slope\_pnt & slope\_mean & dists\_deepest & 7.3 m & 0.35 & 61 \%\\
+slopes\_online\_mean & slope\_mean & dists\_deepest & 7.3 m & 0.36 & 60 \%\\
+slope\_online\_mean & slopes\_online\_mean & dists\_deepest & 7.3 m & 0.29 & 58 \%\\
+slope\_pnts & slope\_online\_mean & dists\_deepest & 7.4 m & 0.27 & 64 \%\\
+slopes\_online\_mean & slope\_online\_mean & dists\_deepest & 7.4 m & 0.33 & 67 \%\\
+slopes\_online\_mean & slope\_online\_mean & dist\_deepest & 7.5 m & 0.26 & 61 \%\\
+slope\_pnt & slopes\_online\_mean & dist\_deepest & 7.5 m & 0.33 & 69 \%\\
+slope\_mean & slopes\_online\_mean & dist\_deepest & 7.6 m & 0.26 & 64 \%\\
+slope\_pnt & slope\_online\_mean & dist\_deepest & 7.7 m & 0.27 & 68 \%\\
+slope\_pnts & slopes\_online\_mean & dists\_deepest & 7.8 m & 0.3 & 65 \%\\
+slope\_pnt & slope\_online\_mean & dists\_deepest & 7.9 m & 0.27 & 67 \%\\
+slope\_mean & slope\_online\_mean & dists\_deepest & 7.9 m & 0.31 & 60 \%\\
 \hline

### 2020-11-25 tables/02_model_metrics.Rmd
format table S2 and fill out caption

@@
 dt_metrics[2,] <- paste0("**", dt_metrics[2,], "**") # highlight row from main text
-clipr::write_clip(
-  gsub("%\\*\\*", "\\\\%\\*\\*",
-  gsub("_", "\\\\_",
-  gsub("%\\\\", "\\\\%\\\\",
-  sub("\\\\midrule", "\\\\hline",
-  sub("\\\\bottomrule", "\\\\hline",
-  sub("\\\\toprule", "\\\\hline",
-      knitr::kable(dt_metrics, digits = 2, format = "latex",
-     booktabs = TRUE, linesep = "", align = c("lllllc"), escape = FALSE)))))))
-   )
+# clipr::write_clip(
+#   gsub("%\\*\\*", "\\\\%\\*\\*",
+#   gsub("_", "\\\\_",
+#   gsub("%\\\\", "\\\\%\\\\",
+#   sub("\\\\midrule", "\\\\hline",
+#   sub("\\\\bottomrule", "\\\\hline",
+#   sub("\\\\toprule", "\\\\hline",
+#       knitr::kable(dt_metrics, digits = 2, format = "latex",
+#      booktabs = TRUE, linesep = "", align = c("lllllc"), escape = FALSE)))))))
+#    )
 ```

### 2020-11-25 manuscript/si_template_2019.tex
fix table S1 variable names

@@
 \setlength\tabcolsep{1.5pt} % default value: 6pt
-
 \begin{tabular}{lllll}
@@
 \hline
-\textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4820 (17700)\\
-\textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4820 (17700)\\
-- & 1500 (1100) & 880 (640) & 2500 (2200) & 4820 (17700)\\
-- & 91 (88) & 41 (40) & 140 (140) & 4820 (17700)\\
-- & 710 (560) & 430 (320) & 1200 (1000) & 4820 (17700)\\
-- & 380 (290) & 230 (170) & 610 (500) & 4820 (17700)\\
-- & 0 (0) & 0 (0) & 58 (25) & 4820 (17700)\\
-- & 0 (0) & 0 (0) & 3 (3) & 4820 (17700)\\
-- & 0 (0) & 0 (0) & 1 (1) & 4820 (17700)\\
-- & 26000 (27000) & 23000 (23000) & 27000 (38000) & 4820 (17700)\\
-\textbf{Deepest point distance (m)} & 180 (-) & 110 (-) & 290 (-) & 4820 (-)\\
-\textbf{Visual center distance (m)} & 240 (-) & 160 (-) & 390 (-) & 4820 (-)\\
-- & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4820 (-)\\
-
-- & 0.062 (-) & 0.029 (-) & 0.14 (-) & 4800 (-)\\
-
-- & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4820 (-)\\
-
-- & 0.077 (-) & 0.038 (-) & 0.13 (-) & 4590 (-)\\
-
-- & 0.077 (-) & 0.037 (-) & 0.13 (-) & 4540 (-)\\
+Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4820 (17700)\\
+\textbf{\textbf{Perimeter (m)}} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4820 (17700)\\
+\textbf{\textbf{Shoreline development}} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4820 (17700)\\
+\textbf{\textbf{Elevation (m)}} & 300 (340) & 180 (210) & 400 (460) & 4820 (17700)\\
+\textbf{\textbf{Watershed-lake ratio}} & 7.8 (10) & 3.8 (4.4) & 17 (29) & 4820 (17700)\\
+\textbf{\textbf{Deepest point distance (m)}} & 180 (-) & 110 (-) & 290 (-) & 4820 (-)\\
+\textbf{\textbf{Mean deepest point distance (m)}} & 140 (-) & 87 (-) & 230 (-) & 4820 (-)\\
+\textbf{\textbf{Visual center distance (m)}} & 240 (-) & 160 (-) & 390 (-) & 4820 (-)\\
+Mean inlake slope (m/m) & 0.043 (-) & 0.021 (-) & 0.09 (-) & 4820 (-)\\
+Inlake slope online (m/m) & 0.062 (-) & 0.029 (-) & 0.14 (-) & 4800 (-)\\
+Inlake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4820 (-)\\
+Inlake slopes (m/m) & 0.057 (-) & 0.032 (-) & 0.098 (-) & 4820 (-)\\
+Inlake slopes online (m/m) & 0.07 (-) & 0.034 (-) & 0.15 (-) & 4800 (-)\\
+Nearshore mean slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4820 (-)\\
+Nearshore slope online (m/m) & 0.077 (-) & 0.038 (-) & 0.13 (-) & 4590 (-)\\
+Nearshore slopes online (m/m) & 0.077 (-) & 0.037 (-) & 0.13 (-) & 4540 (-)\\
 \hline

### 2020-11-25 tables/01_predictors.Rmd
fix table S1 variable names

@@
   distinct(lagoslakeid, .keep_all = TRUE) %>%
-  dplyr::select(-inlake_slope_mean)
+  dplyr::select(-matches("inlake_slopes_online"), 
+                -matches("inlake_slope_online"), 
+                -matches("inlake_slope_mean"),
+                -lake_centroidstate.y, 
+                -lake_centroidstate.x)
 dt_bathy <- dt_all %>%
@@
 # setwd("figures")
+# dt_combined$variable
+exclude_vars <- unique(dt_combined$variable)[
+  stringr::str_detect(unique(dt_combined$variable), 
+                     "totalarea|islandperimeter|between|buffer100m|lake_meandepth_m|countyfips|upstream|length_m|width_m|orientation_deg|nws_")]
 slope_distance_vars <- unique(dt_combined$variable)[
@@
                      "dist_|e_slope_|dists_|e_slopes")]
-exclude_vars <- unique(dt_combined$variable)[
-  stringr::str_detect(unique(dt_combined$variable), 
-                     "totalarea|islandperimeter|between|buffer100m|lake_meandepth_m")]
+slope_distance_vars <- slope_distance_vars[!slope_distance_vars %in% exclude_vars]
 
 dt <- dt_combined %>%
-  dplyr::filter(str_detect(variable, 
-                c(paste0(slope_distance_vars, collapse = "|"),
-                  "^lake_|^ws_lake_arearatio"))) %>%
-  dplyr::filter(!str_detect(variable,
-                            paste0(exclude_vars, collapse = "|")))
+  dplyr::filter(variable %in% c(slope_distance_vars, 
+                           "lake_maxdepth_m", "lake_elevation_m",
+                           "lake_perimeter_m", "lake_shorelinedevfactor",
+                           "ws_lake_arearatio"))
+dt$variable
                        # %>%
@@
 taxonomy_key <- read.csv("../data/taxonomy.csv", stringsAsFactors = FALSE)
+
+# dt$variable[!dt$variable %in% taxonomy_key$col_name]
+
+dt[9:16,] <- dplyr::arrange(dt[9:16,], variable)
 
@@
 # clipr::write_clip(
-#   knitr::kable(dt, format = 'latex', align = c("lllll"), escape = FALSE))
+#   gsub("%\\*\\*", "\\\\%\\*\\*",
+#   gsub("_", "\\\\_",
+#   gsub("%\\\\", "\\\\%\\\\",
+#   sub("\\\\midrule", "\\\\hline",
+#   sub("\\\\bottomrule", "\\\\hline",
+#   sub("\\\\toprule", "\\\\hline",
+#       knitr::kable(dt, digits = 2, format = "latex",
+#      booktabs = TRUE, linesep = "", align = c("l"), escape = FALSE)))))))
+#    )
 ```

### 2020-11-25 scripts/99_make_taxonomy.R

@@ -51,30 +51,38 @@ taxonomy_key <- bind_rows(
   c(col_name = "maxdepth_false_false",
     description = "lake maximum depth calculated from slope_mean and dist_viscenter in meters"),
 
-  c(col_name = "inlake_slope",
+  c(col_name = "inlake_slope_pnt",
     description = "slope of the lake bottom assuming a straight line from the shore to the deepest point of the lake in units of depth change in meters per meter distance",
     pretty = "Inlake slope (m/m)"),
 
-  c(col_name = "inlake_slopes",
+  c(col_name = "inlake_slope_pnts",
     description = "slope of the lake bottom assuming a straight line from the shore to the deepest point(s) of the lake in units of depth change in meters per meter distance",
-    pretty = "Mean inlake slope (m/m)"),
+    pretty = "Inlake slopes (m/m)"),
 
   c(col_name = "inlake_slope_mean",
     description = "the mean slope of the lake bottom in units of depth change in meters per meter distance",
-    pretty = "Mean lake slope (m/m)"),
+    pretty = "Mean inlake slope (m/m)"),
 
-  c(col_name = "slope_mean",
+  c(col_name = "inlake_slope_online_mean",
+    description = "average slope of a line extending from the deepest point of a lake to the shoreline",
+    pretty = "Inlake slope online (m/m)"),
+
+  c(col_name = "inlake_slopes_online_mean",
+    description = "average slope of a line (or lines) extending from the deepest point(s) of a lake to the shoreline",
+    pretty = "Inlake slopes online (m/m)"),
+
+  c(col_name = "nearshore_slope_mean",
     description = "average slope of the land surrounding a lake in a 100m buffer",
-    pretty = "Mean slope (m/m)"),
+    pretty = "Nearshore mean slope (m/m)"),
   c(col_name = "buffer_slope",
     description = "average slope of the land surrounding a lake in a 100m buffer",
     pretty = "Nearshore slope (m/m)"),
 
-  c(col_name = "slope_online_mean",
+  c(col_name = "nearshore_slope_online_mean",
     description = "average slope of the land surrounding a lake in a straight-line  extending from the deepest point of a lake and buffered on either side by 20m",
     pretty = "Nearshore slope online (m/m)"),
 
-  c(col_name = "slopes_online_mean",
+  c(col_name = "nearshore_slopes_online_mean",
[…]

### 2020-11-25 manuscript/si_template_2019.tex
fight with table S2 bolding

--- 2020-11-25 manuscript/si_template_2019.tex
@@
 \hline
+Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4820 (17700)\\
 \textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4820 (17700)\\
 \textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4820 (17700)\\
-- & 1500 (1100) & 880 (640) & 2500 (2200) & 4820 (17700)\\
-- & 91 (88) & 41 (40) & 140 (140) & 4820 (17700)\\
-- & 710 (560) & 430 (320) & 1200 (1000) & 4820 (17700)\\
-- & 380 (290) & 230 (170) & 610 (500) & 4820 (17700)\\
-- & 0 (0) & 0 (0) & 58 (25) & 4820 (17700)\\
-- & 0 (0) & 0 (0) & 3 (3) & 4820 (17700)\\
-- & 0 (0) & 0 (0) & 1 (1) & 4820 (17700)\\
-- & 26000 (27000) & 23000 (23000) & 27000 (38000) & 4820 (17700)\\
+\textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4820 (17700)\\
+\textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4820 (17700)\\
+\textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4820 (17700)\\
+\textbf{Watershed-lake ratio} & 7.8 (10) & 3.8 (4.4) & 17 (29) & 4820 (17700)\\
 \textbf{Deepest point distance (m)} & 180 (-) & 110 (-) & 290 (-) & 4820 (-)\\
+Mean deepest point distance (m) & 140 (-) & 87 (-) & 230 (-) & 4820 (-)\\
 \textbf{Visual center distance (m)} & 240 (-) & 160 (-) & 390 (-) & 4820 (-)\\
-- & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4820 (-)\\
-
-- & 0.062 (-) & 0.029 (-) & 0.14 (-) & 4800 (-)\\
-
-- & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4820 (-)\\
-
-- & 0.077 (-) & 0.038 (-) & 0.13 (-) & 4590 (-)\\
-
-- & 0.077 (-) & 0.037 (-) & 0.13 (-) & 4540 (-)\\
+Inlake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4820 (-)\\
+Inlake slope online (m/m) & 0.062 (-) & 0.029 (-) & 0.14 (-) & 4800 (-)\\
+Inlake slopes (m/m) & 0.057 (-) & 0.032 (-) & 0.098 (-) & 4820 (-)\\
+Inlake slopes online (m/m) & 0.07 (-) & 0.034 (-) & 0.15 (-) & 4800 (-)\\
+Mean inlake slope (m/m) & 0.043 (-) & 0.021 (-) & 0.09 (-) & 4820 (-)\\
+Nearshore mean slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4820 (-)\\
+Nearshore slope online (m/m) & 0.077 (-) & 0.038 (-) & 0.13 (-) & 4590 (-)\\
+Nearshore slopes online (m/m) & 0.077 (-) & 0.037 (-) & 0.13 (-) & 4540 (-)\\
 \hline

### 2020-11-25 tables/01_predictors.Rmd
fight with table S2 bolding

@@
                            "lake_perimeter_m", "lake_shorelinedevfactor",
-                           "ws_lake_arearatio"))
-dt$variable
+                           "ws_lake_arearatio", "lake_islandarea_ha", 
+                           "lake_waterarea_ha"))
+# dt$variable
                        # %>%
@@
 
-dt[9:16,] <- dplyr::arrange(dt[9:16,], variable)
-
 dt <- left_join(dt, taxonomy_key, by = c("variable" = "col_name")) %>%
@@
 
+dt[11:18,] <- dplyr::arrange(dt[11:18,], variable)
+
 dt <- dplyr::rename(dt, Variable = variable)
+
+model_predictors <- data.frame(
+  variable = names(read.csv("../data/depth_predictors.csv", 
+                            stringsAsFactors = FALSE)), 
+  stringsAsFactors = FALSE) %>%
+  dplyr::filter(variable != "lake_maxdepth_m") %>%
+  left_join(taxonomy, by = c("variable" = "col_name")) %>%
+  pull(pretty) %>% na.omit() %>% as.character()
 
@@
 dt_md <- dt
-dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)] <- 
+dt_md$Variable[
+  dt_md$Variable %in% model_predictors & !is.na(dt_md$Variable)] <- 
   paste0("**", 
-         dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)],
+dt_md$Variable[
+  dt_md$Variable %in% model_predictors & !is.na(dt_md$Variable)],
          "**")
@@
 # bold Eq 2 predictor variables
-dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)] <- 
+dt$Variable[
+  dt$Variable %in% model_predictors & !is.na(dt$Variable)] <- 
   paste0("\\textbf{", 
-         dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)],
+dt$Variable[
+  dt$Variable %in% model_predictors & !is.na(dt$Variable)],
          "}")
 
-# clipr::write_clip(
-#   gsub("%\\*\\*", "\\\\%\\*\\*",
-#   gsub("_", "\\\\_",
-#   gsub("%\\\\", "\\\\%\\\\",
-#   sub("\\\\midrule", "\\\\hline",
-#   sub("\\\\bottomrule", "\\\\hline",
-#   sub("\\\\toprule", "\\\\hline",
-#       knitr::kable(dt, digits = 2, format = "latex",
-#      booktabs = TRUE, linesep = "", align = c("l"), escape = FALSE)))))))
-#    )
+if(interactive()){
+  clipr::write_clip(
+    gsub("%\\*\\*", "\\\\%\\*\\*",
+    gsub("_", "\\\\_",
+    gsub("%\\\\", "\\\\%\\\\",
+    sub("\\\\midrule", "\\\\hline",
+    sub("\\\\bottomrule", "\\\\hline",
+    sub("\\\\toprule", "\\\\hline",
+        knitr::kable(dt, digits = 2, format = "latex",
+       booktabs = TRUE, linesep = "", align = c("l"), escape = FALSE)))))))
+     )
+}
 ```

### 2020-11-25 manuscript/agujournaltemplate.tex
text edits

@@
 Reported nearshore land slope values are the mean of all points in the buffer.
-In addition to the aforementioned methods of calculating in-lake (and nearshore) slopes and distances, we tried 7 alternative methods which are described in Table S2.
+In addition to the aforementioned techniques of calculating in-lake (and nearshore) slopes and distances, we tried 7 alternate techniques which are described in Table S2.
 
@@
 The purpose of this standardization was to prevent lakes with extremely long proxy distances from having an outsized impact on model evaluation metrics.
-In addition to comparing among model runs using different metric combinations, we compared among sets of model runs where slope and distance measures were calculated using different methods (Table S2).
+In addition to comparing among model runs using different metric combinations, we compared among sets of model runs where slope and distance measures were calculated using different sets of calculation techniques (Table S2).
 
@@
 \end{figure}
+
+Model fit and prediction error differed depending on the technique used to calculate inlake and nearshore geometry metrics (Table S2).
+We found that the best model fit and lowest model error occurred when inlake slope was calculated as the average point-wise slope of all points at maximum lake depth and nearshore slope was calculated as the mean of all points in the lake buffer.
+However, given the small difference in the fit of this model compared to the model where only a single deepest point was used to calculate slope, we limit our discussion to the latter case hereafter.
+The more complex calculations involving all points at maximum depth represent a significant increase in computational load and complexity.
 
@@
 This finding is consistent with \citeA{heathcotePredictingBathymetricFeatures2015} whos results suggest that in-lake slopes are shallower compared to the surrounding land.
-The reason for these shallow in-lake slopes may be related to slope-induced turbidity currents which distribute sediment from shallow high-energy areas of lakes to deep low-energy areas \cite{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}.
+Furthermore, the fact that in-lake slopes were shallower compared to the surrounding land even after controlling for differences in area and depth (Figure S9) is consistent with the idea of topographic scaling (i.e.
+scale invariance) explored in previous work and detailed by \cite{caelVolumeMeanDepth2017}.
+The underlying reason for these shallow in-lake slopes may be related to slope-induced turbidity currents which distribute sediment from shallow high-energy areas of lakes to deep low-energy areas \cite{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}.
 The strength of such sediment focusing is likely greater in "younger" lakes with steeper slopes leading to a smoothing of their bathymetry over time \cite{blaisInfluenceLakeMorphometry1995}.

### 2020-11-26 tables/01_predictors.Rmd
typo fix

@@
   dplyr::filter(variable != "lake_maxdepth_m") %>%
-  left_join(taxonomy, by = c("variable" = "col_name")) %>%
+  left_join(taxonomy_key, by = c("variable" = "col_name")) %>%
   pull(pretty) %>% na.omit() %>% as.character()

### 2020-11-26 figures/01_geometry.Rmd
adjust figure s9 dimensions

@@
 
-```{r slope_by_area_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5.5, fig.height=4.2}
+```{r slope_by_area_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5.7, fig.height=3.3}
 
@@
                method = "lm", se = FALSE) + 
+  xlim(c(0, 0.6)) + ylim(c(0, 0.6)) +
   facet_wrap(~shape_class) +
@@
                method = "lm", se = FALSE) +
+    xlim(c(0, 0.6)) + ylim(c(0, 0.6)) +
     facet_wrap(~shape_class) +

### 2020-11-26 manuscript/si_template_2019.tex
adjust figure s9 dimensions

@@
       \begin{center}
-        \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/slope_by_area_depth-1}
+        \includegraphics[width=0.79\textwidth,keepaspectratio]{../figures/slope_by_area_depth-1}
       \end{center}
-      \caption{Compare the mean lake slopes in concave and convex lakes of the same size and max depth.}\label{figS1}
+      \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth.}\label{figS9}
     \end{figure}

### 2020-11-26 manuscript/agujournaltemplate.tex
text edits

@@
 We compiled bathymetry data on approximately 5,000 lakes in the Northeastern and Midwestern US from nine official state databases (Figure S2).
+These lakes represent a diverse cross section of lakes in terms of their surface areas (4 - 18500 ha) and span a wide geographic extent including glaciated and non-glaciated regions.
 The original data came in a variety of formats including pre-interpolated rasters (Minnesota), contour lines (Nebraska, Michigan, Massachusetts, Kansas, Iowa), contour polygons (New Hampshire, Connecticut), or point depth soundings (Maine).
@@
 Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes whereas proxy nearshore land slope is not representative of true in-lake slope.
-The strong relationship between distance to the center of lakes and distance to the deepest point means that it is possible to compute a "correction factor" to convert between the two measures (Figure \ref{fig2}).
+The strong relationship between distance to the center of lakes and distance to the deepest point means that it is possible to compute a "correction factor" to convert between the two measures (See best-fit equations in Figure \ref{fig2}).
 
@@
 
-Model fit and prediction error differed depending on the technique used to calculate inlake and nearshore geometry metrics (Table S2).
-We found that the best model fit and lowest model error occurred when inlake slope was calculated as the average point-wise slope of all points at maximum lake depth and nearshore slope was calculated as the mean of all points in the lake buffer.
-However, given the small difference in the fit of this model compared to the model where only a single deepest point was used to calculate slope, we limit our discussion to the latter case hereafter.
-The more complex calculations involving all points at maximum depth represent a significant increase in computational load and complexity.
+Model fit and prediction error differed depending on the technique used to calculate in-lake and nearshore geometry metrics (Table S2).
+We found that the best model fit and lowest model error occurred when in-lake slope was calculated as the average point-wise slope of all points at maximum lake depth rather than a single point of maximum depth.
+However, given the small difference in the fit of models using either of these techniques and the significant cost in terms of computational load and complexity, we limit our discussion hereafter to the simpler case involving only a single deepest point.
 

### 2020-11-27 manuscript/figures.Rmd
correct figure/table numbering and placement

@@
 \captionsetup[figure]{labelformat=empty}
-```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuracy metrics (RMSE = root mean square error, R2 = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+```{r depth_model_metrics, fig.cap = "Table 1: Model fit and predictive accuracy metrics (RMSE = root mean square error, R2 = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 # see tables/02_model_metrics.Rmd

### 2020-11-27 tables/02_model_metrics.Rmd
correct figure/table numbering and placement

--- 2020-11-25 tables/02_model_metrics.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 5, fig.height = 6)
 # setwd("tables")
@@
 
+\clearpage
+
 ```{r 02_alternate_metrics, eval=TRUE}
@@
 
-dt_metrics[2,] <- paste0("**", dt_metrics[2,], "**") # highlight row from main text
-clipr::write_clip(
-  gsub("%\\*\\*", "\\\\%\\*\\*",
-  gsub("_", "\\\\_",
-  gsub("%\\\\", "\\\\%\\\\",
-  sub("\\\\midrule", "\\\\hline",
-  sub("\\\\bottomrule", "\\\\hline",
-  sub("\\\\toprule", "\\\\hline",
-      knitr::kable(dt_metrics, digits = 2, format = "latex",
-     booktabs = TRUE, linesep = "", align = c("lllllc"), escape = FALSE)))))))
-   )
+# dt_metrics[2,] <- paste0("**", dt_metrics[2,], "**") # highlight row from main text
+# clipr::write_clip(
+#   gsub("%\\*\\*", "\\\\%\\*\\*",
+#   gsub("_", "\\\\_",
+#   gsub("%\\\\", "\\\\%\\\\",
+#   sub("\\\\midrule", "\\\\hline",
+#   sub("\\\\bottomrule", "\\\\hline",
+#   sub("\\\\toprule", "\\\\hline",
+#       knitr::kable(dt_metrics, digits = 2, format = "latex",
+#      booktabs = TRUE, linesep = "", align = c("lllllc"), escape = FALSE)))))))
+#    )
 ```

### 2020-11-28 manuscript/agujournaltemplate.tex
text edits

@@
 We used our generated bathymetry surfaces to find the location of the deepest point in the lake and we resolved ties by choosing the deepest point that was closest to the center of the lake.
+We used the location of this deepest point to calculate "distance to the deepest point" as the minimum distance to the lake shoreline.
 To account for lakes where the centroid does not intersect lake bathymetry because it is located within an embedded island or peninsula, we calculated the center of the lake not as its centroid but rather by finding the point farthest from the lake shoreline (i.e.
@@
 For our purposes, we set a conservative classification probability threshold of 0.75 to determine whether a lake would be considered a  reservoir.
-We tested the sensitivity of our model output relative to variations of this threshold and found little substantive difference in results.
 Note that our reservoir classification defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water.
 It makes no distinction between different dam types, heights, or uses/purposes.
+% We tested the sensitivity of our model output relative to variations of this threshold and found little substantive difference in results.
 
@@
 One reason that we used the cone model is that, unlike other idealized shapes, it is well suited for modeling maximum depth and does not require any knowledge of lake volume or mean depth.
-Note that Equation \ref{eq1} is a geometric formula and has no intercept or “coefficients” and it produces a perfect depth estimate given true values of slope and distance.
+Note that Equation \ref{eq1} is a geometric formula and has no intercept or coefficients and it produces a perfect depth estimate given true values of slope and distance.
 To use this model to predict the depth of all lakes, there is a necessary assumption that proxy slope and distance measures, which are available for all lakes, are representative of true slope and distance (Figure \ref{fig1}).
@@
 For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S4).
-However,  there were not similar differences among slope and distance metrics for natural lakes versus reservoirs (Figure S4).
+We found evidence that this difference was at leat partly explained by the fact that convex lakes are deeper than concave lakes (Figure S5).
+There were not similar differences among slope and distance metrics for natural lakes versus reservoirs.
 
@@
 
-The most important covariates in offset models were those relating to spatial location, lake area, and perimeter (Figure S5).
-Conversely, watershed metrics and lake elevation had little contribution to random forest model fit (Figure S5).
+The most important covariates in offset models were those relating to spatial location, lake area, and perimeter (Figure S7).
+Conversely, watershed metrics and lake elevation had little contribution to random forest model fit (Figure S7).
 The spatial location (i.e.
 HUC4) covariate was notably less importance in the true slope model compared to the two proxy slope models.
-Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Equation \ref{eq1} (Figure S5).
+Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Equation \ref{eq1} (Figure S7).
 
@@
 Models using non-representative proxies showed increased error and systematic overestimation of depth in concave and reservoir lakes.
-Although our analysis was limited to lakes with available bathymetry data, these lakes did not have characteristics that differed from that of the overall lake population (Figure S6, S8).
+Although our analysis was limited to lakes with available bathymetry data, these lakes did not have characteristics that differed from that of the overall lake population (Figure S8, S10).
 Although there is a possibility that there is some hidden bias not explored for in our analyses, this lack of difference suggests that our results are likely to be broadly applicable to all lakes.
@@
 This finding is consistent with \citeA{heathcotePredictingBathymetricFeatures2015} whos results suggest that in-lake slopes are shallower compared to the surrounding land.
-Furthermore, the fact that in-lake slopes were shallower compared to the surrounding land even after controlling for differences in area and depth (Figure S9) is consistent with the idea of topographic scaling (i.e.
+Furthermore, the fact that in-lake slopes were shallower compared to the surrounding land even after controlling for differences in area (Figure S5) is consistent with the idea of topographic scaling (i.e.
 scale invariance) explored in previous work and detailed by \cite{caelVolumeMeanDepth2017}.
@@
 In addition, convex lakes often had steeper in-lake slopes relative to nearshore land slopes as compared to concave lakes.
-Finally, it was notable that convex lakes were deeper than concave lakes despite having similar distributions of lake surface area (Figure S7).
+Finally, it was notable that convex lakes were deeper than concave lakes despite having similar distributions of lake surface area (Figure S9).
 The underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of lake formation whereby the formation of concave lakes were a result of more intense glacial scouring compared to that of convex lakes \cite{gorhamPhysicalLimnologyNorthern1958}.
@@
 The greater sensitivity of depth predictions to proxy slope measures relative to proxy distance measures may be explained by the fact that proxy slope measures were a more imperfect representation of true in-lake slopes relative to proxy versus true distances.
-In addition, these results help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \cite{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S7).
+We did not find evidence that the outsized sensivity of depth predictions to slope was dependent on variations in how these measures were calculated (Table S2).
+In a general sense, the sensitivity of depth predictions to slope help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \cite{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S9).
 

### 2020-11-28 manuscript/si_template_2019.tex
text edits

@@
       \begin{center}
-        \includegraphics[width=\textwidth,keepaspectratio]{../figures/01_slope_compare-1}
-      \end{center}
-      \caption{Sensitivity analysis regarding different slope calculation methods.}\label{figS1}
-    \end{figure}
-\clearpage
-
-\begin{figure}
-      \begin{center}
         \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/lake_shape}
@@
 \begin{figure}
+      \begin{center}
+        \includegraphics[width=0.79\textwidth,keepaspectratio]{../figures/slope_by_area_depth-1}
+      \end{center}
+      \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth.}\label{figS5}
+    \end{figure}
+\clearpage
+
+\begin{figure}
+      \begin{center}
+        \includegraphics[width=\textwidth,keepaspectratio]{../figures/01_slope_compare-1}
+      \end{center}
+      \caption{Sensitivity analysis regarding different slope calculation methods.}\label{figS6}
+    \end{figure}
+\clearpage
+
+\begin{figure}
       \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
@@
 See Equation 1 for a definition of geometric max depth.
-HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figS5}
+HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figS7}
       \end{figure}
@@
             \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product.
-The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS6}
+The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS8}
       \end{figure}
@@
       \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
-      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS7}
+      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS9}
       \end{figure}
@@
 \begin{figure}
-      \begin{center}
-        \includegraphics[width=0.79\textwidth,keepaspectratio]{../figures/slope_by_area_depth-1}
-      \end{center}
-      \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth.}\label{figS9}
-    \end{figure}
-\clearpage
-
-\begin{figure}
       \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
       \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product.
-For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figS8}
+For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figS10}
       \end{figure}

### 2020-11-28 manuscript/agujournaltemplate.tex
note distance upper bound phenomena

@@
 
-Although proxy distance to lake center was often much larger in magnitude compared to the true distance to the deepest point of lakes’, they were strongly related ($R^2$ = 0.8).
-In contrast, proxy nearshore land slope and true in-lake slope were more weakly related ($R^2$ = 0.17).
+Although proxy distance to lake center was often larger in magnitude compared to the true distance to the deepest point of lakes’, they were strongly related ($R^2$ = 0.8).
+Note that the coefficient of determination for this relationship is not strictly correct given that distance to lake center is an upper bound on distance to the deepest point of lakes.
+In contrast to distance metrics, proxy nearshore land slope and true in-lake slope were more weakly related ($R^2$ = 0.17).
 For slope measures, most lakes had lower magnitude (i.e.

### 2020-11-28 manuscript/si_template_2019.tex
cleanup appendix captions, toc

@@
 
-\item \textbf{Figure S5:} Importance plot for random forest variables
+\item \textbf{Figure S5:} Lake geometry comparison among shape classes by depth and area
 
-\item \textbf{Figure S6:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
+\item \textbf{Figure S6:} Effect of calculation technique on in-lake/nearshore slope comparison.
 
-\item \textbf{Figure S7:} Lake characteristics comparison among shape and reservoir classes
+\item \textbf{Figure S7:} Importance plot for random forest variables
 
-\item \textbf{Figure S8:} Comparison between reported depth and bathymetry depth
+\item \textbf{Figure S8:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
+
+\item \textbf{Figure S9:} Lake characteristics comparison among shape and reservoir classes
+
+\item \textbf{Figure S10:} Comparison between reported depth and bathymetry depth
+
+\item \textbf{Table S1:} Summary of lake characteristics for the present study
+
+\item \textbf{Table S2:} Effect of calculation technique on depth model fit and accuracy.
 
@@
       \end{center}
-      \caption{Sensitivity analysis regarding different slope calculation methods.}\label{figS6}
+      \caption{Comparison between in-lake and nearshore slope using different calculation techniques. See Table S2 for explanation of each.}\label{figS6}
     \end{figure}

### 2020-12-02 manuscript/agujournaltemplate.tex
typo fix, address stray reviewer comment

@@
 Lake depth is an important factor controlling lake physics, chemistry, and biota.
-Deeper lakes generally have higher water clarity and less complete mixing compared to shallow lakes \cite{feeEffectsLakeSize1996, readSimulating2368Temperate2014}.
+For example, deeper lakes generally have higher water clarity and less complete mixing compared to shallow lakes \cite{feeEffectsLakeSize1996, readSimulating2368Temperate2014}.
 These differences are reflected in variation among lakes in terms of biological productivity \cite{qinWaterDepthUnderpins2020} and rates of greenhouse gas production \cite{liSignificantContributionLake2020}.
@@
 The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \cite{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
+Such efforts rely on a strategy of exploiting correlations between nearshore geomorphology and in-lake geometry, which at limited extents can be quite strong, while at larger extents can be dependent on geographic location and lake type \cite{oliver2016prediction, branstrator2009origins}.
 Given the limited prediction accuracy of prior depth prediction efforts ($\pm$ 6-7 m), a major focus has been on improving accuracy using strategies such as employing more diverse covariates \cite{oliver2016prediction}, varying lake buffer sizes \cite{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings (e.g.
 fitting different models for distinct size classes) among lakes \cite{caelVolumeMeanDepth2017, sobekPredictingDepthVolume2011}.
-Unfortunately, the results of these efforts have been mixed from a pure prediction standpoint.
+Unfortunately, the predictive accuracy of these efforts has been limited.
 
@@
 For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S4).
-We found evidence that this difference was at leat partly explained by the fact that convex lakes are deeper than concave lakes (Figure S5).
+We found evidence that this difference was at least partly explained by the fact that convex lakes are deeper than concave lakes (Figure S5).
 There were not similar differences among slope and distance metrics for natural lakes versus reservoirs.

### 2020-12-03 figures/01_geometry.Rmd
bold figs6 panel used in main text

@@
     # geom_abline(slope = 1, intercept = 0) +
-    xlab(stringr::str_replace_all(nearshore_var, c("nearshore_" = "", "_mean" = ""))) +
+    xlab(stringr::str_replace_all(nearshore_var, c("nearshore_" = ""))) +
     ylab(pad_str(stringr::str_replace_all(inlake_var, c("inlake_" = "")), 
@@
     theme_opts +
-    theme(axis.title = element_text(family = "mono"))
+    theme(axis.title = element_text(family = "mono", size = 6))
 
@@
 )
+
+gg_slope_list[[1]] <- gg_slope_list[[1]] +
+  theme(axis.title.y = element_text(face = "bold"))
+
+gg_slope_list[[5]] <- gg_slope_list[[5]] +
+  theme(axis.title.x = element_text(face = "bold"))
 

### 2020-12-03 manuscript/agujournaltemplate.tex
text edits

@@
 \item Using data from thousands of lakes, we show that nearshore land slope is a poor proxy for in-lake slope and increases prediction error.
-\item Prediction errors were systematic such that depth was overpredicted in concave and reservoir lakes.
+\item Systematic overprediction of depth in concave and reservoir lakes highlights the potential for improvement in future depth prediction efforts.
 \end{keypoints}
@@
 \noindent
-One of our models (true slope, proxy distance) was more accurate than even the most accurate of prior studies.
+One of our models (where $depth_{geometric}$ was calculated using true slope and proxy distance) was more accurate than even the most accurate of prior studies.
 However, parameterization of this model requires data on bathymetry which is not available for all lakes.

### 2020-12-03 figures/01_geometry.Rmd
fight with figS6 dimensions

@@
     theme_opts +
-    theme(axis.title = element_text(family = "mono", size = 6))
+    theme(axis.title = element_text(family = "mono", size = 8))
 
@@
                    nrow = 5, byrow = FALSE, 
-                   rel_heights = c(rep(0.6, 4), 1),
+                   rel_heights = c(rep(0.75, 4), 1),
                    rel_widths = c(rep(1, 1), 
-                                  rep(0.37, 2))), 
+                                  rep(0.48, 2))
+                   ), 
 xlab, ncol = 1, rel_heights = c(1, 0.1)), 

### 2020-12-03 manuscript/si_template_2019.tex
add calc techniques to figs6 caption
add takeaway to figs6 caption

@@
       \end{center}
-      \caption{Comparison between in-lake and nearshore slope using different calculation techniques. See Table S2 for explanation of each.}\label{figS6}
+      \caption{Comparison between in-lake and nearshore slope using different calculation techniques.
+The techniques used in the main text analyses are bolded and the combination of these techniques (top-left corner) produces the strongest relationship between the two metrics.
+\texttt{slope\_mean} is the mean slope of all inlake or nearshore buffer points.
+\texttt{slope\_pnts} is the average \texttt{slope\_pnt} of all points at maximum depth.
+\texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope).
+\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except for all inlake points at maximum depth.
+\texttt{dists\_deepest} is the same as \texttt{dist\_deepest} except distance is calculated for all points at maximum depth.}\label{figS6}
     \end{figure}

### 2020-12-04 manuscript/si_template_2019.tex
calculation technique caption grammar and cleanup

@@
 \texttt{slope\_mean} is the mean slope of all inlake or nearshore buffer points.
-\texttt{slope\_pnts} is the average \texttt{slope\_pnt} of all points at maximum depth.
+\texttt{slope\_pnts} is the average slope (i.e.
+\texttt{slope\_pnt}) of all points at maximum depth.
 \texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope).
-\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except for all inlake points at maximum depth.
-\texttt{dists\_deepest} is the same as \texttt{dist\_deepest} except distance is calculated for all points at maximum depth.}\label{figS6}
+\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth.}\label{figS6}
     \end{figure}
@@
 \texttt{slope\_mean} is the mean slope of all inlake or nearshore buffer points.
-\texttt{slope\_pnts} is the average \texttt{slope\_pnt} of all points at maximum depth.
+\texttt{slope\_pnts} is the average slope (i.e.
+\texttt{slope\_pnt}) of all points at maximum depth.
 \texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope).
-\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except for all inlake points at maximum depth.
+\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth.
 \texttt{dists\_deepest} is the same as \texttt{dist\_deepest} except distance is calculated for all points at maximum depth.}  \label{tableS2}

### 2020-12-07 figures/01_geometry.Rmd
add axis superlabels

@@
 cowplot::plot_grid(
+  ggdraw() + 
+    draw_label("True measures", y = 0.37, hjust = 0, size = 12, angle = 90,  fontface = "bold"), 
+cowplot::plot_grid(
+  cowplot::plot_grid(
     gg_dist_base, 
@@
     rel_widths = c(1, 0.9)
-)
+  ), 
+  ggdraw() + 
+    draw_label("Proxy measures", x = 0.45, hjust = 0, size = 12, fontface = "bold"),
+  ncol = 1, rel_heights = c(1, 0.1)),
+ncol = 2, rel_widths = c(0.1, 1))
 

### 2020-12-07 manuscript/agujournaltemplate.tex
stash full paper
pas text edits

@@
 \item Using data from thousands of lakes, we show that nearshore land slope is a poor proxy for in-lake slope and increases prediction error.
-\item Systematic overprediction of depth in concave and reservoir lakes highlights the potential for improvement in future depth prediction efforts.
+\item Systematic overprediction of depth in concave and reservoir lakes highlights potential areas for improved depth prediction.
 \end{keypoints}
@@
 Additionally it is unclear whether model prediction error is related to differences in lake type such those with different cross-section shapes or those classified as reservoirs versus natural lakes.
-Given these knowledge gaps, we asked three research questions: 1) How representative is nearshore land slope of in-lake slope; and how representative is the distance to the center of a lake compared to the distance to the deepest point of a lake?
+To address these knowledge gaps, we asked three research questions: 1) How representative is nearshore land slope of in-lake slope; and how representative is the distance to the center of a lake compared to the distance to the deepest point of a lake?
 2) How does the use of proxies for lake geometry affect lake depth prediction error?
@@
 natural lake vs reservoir) affect depth prediction error?
-To answer these questions, we extracted maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, cross-section shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 lakes from bathymetric map data and supplemented this data with classification estimates of whether lakes are reservoirs or natural lakes.
+To answer these questions, we extracted maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, cross-section shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 lakes from bathymetric map data.
+We supplemented these geometry measures with data classifying waterbodies as reservoirs or natural lakes.
 We used this data to compute geometric depth estimates (Equation \ref{eq1}) and prediction "offsets" to these estimates using the random forest algorithm (Equation \ref{eq3}).
-Covariates used in offset modeling included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table S1).
+Such offsets are model quantities which minimize differences between observed and predicted depth as a function of covariates.
+In our case, covariates included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table S1).
 
@@
 next to the dam) rather than in the center of the reservoir \cite{lanza1985interactions}.
-In a similar fashion, we expect to see overestimation error associated with using a nearshore land slope proxy in lakes with differing cross-section shape such that the depth of U-shaped (i.e.
+In a similar fashion, we expect to see overestimation error associated with using a nearshore land slope proxy in lakes with differing cross-section shape such that the depth of bowl-shaped (i.e.
 concave) lakes will be overpredicted whereas the depth of V-shaped (i.e.
@@
 
-We further classified lakes using the output of a deep convolutional neural network model trained on satellite images labeled according to whether there was visual evidence of a water control structure significantly impacting flow.
+We further classified lakes using the output of a deep convolutional neural network model trained on satellite images labeled according to whether there was visual evidence of a water control structure significantly impacting flow (S.
+Polus, personal communication, 2020).
 This model had an overall validation accuracy of 81\% and produced a probability for each waterbody as to whether it is a reservoir or a natural lake.
@@
 Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes whereas proxy nearshore land slope is not representative of true in-lake slope.
-The strong relationship between distance to the center of lakes and distance to the deepest point means that it is possible to compute a "correction factor" to convert between the two measures (See best-fit equations in Figure \ref{fig2}).
+The strong relationship between distance to the center of lakes and distance to the deepest point means that it is possible to convert between the two measures in subsequent analyses (See best-fit equations in Figure \ref{fig2}).
 
@@
 
-Model fit and prediction error differed depending on the technique used to calculate in-lake and nearshore geometry metrics (Table S2).
+Offset model fit and prediction error differed depending on the technique used to calculate in-lake and nearshore geometry metrics (Table S2).
 We found that the best model fit and lowest model error occurred when in-lake slope was calculated as the average point-wise slope of all points at maximum lake depth rather than a single point of maximum depth.
@@
 
-The most important covariates in offset models were those relating to spatial location, lake area, and perimeter (Figure S7).
-Conversely, watershed metrics and lake elevation had little contribution to random forest model fit (Figure S7).
+The most important covariates in these models were those relating to spatial location, lake area, and perimeter (Figure S7).
+Conversely, watershed metrics and lake elevation had little contribution to random forest model fit.
 The spatial location (i.e.

### 2020-12-07 figures/02_depth_model.Rmd
add xaxis labels to residual density plot

@@
                                rel_widths = c(1, 0.85)), 
+            ggdraw() + draw_label("Residual (m)", size = 10, y = 0.6),
             leg, 
-            nrow = 3, rel_heights = c(0.85, 1, 0.15)), 
+            nrow = 4, rel_heights = c(0.85, 1, 0.1, 0.15)),
+
           ncol = 2, rel_widths = c(0.1, 1)) %>%
@@
                                rel_widths = c(1, 0.85)), 
+            ggdraw() + draw_label("Residual (m)", size = 10, y = 0.6),
             leg, 
-            nrow = 3, rel_heights = c(0.85, 1, 0.15)), 
+            nrow = 4, rel_heights = c(0.85, 1, 0.1, 0.15)), 
           ncol = 2, rel_widths = c(0.1, 1)) %>%

### 2020-12-07 figures/00_maps.Rmd
move map to main text

@@
   dplyr::distinct(lagoslakeid, shape_class) %>%
-  dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::filter(!is.na(shape_class) & shape_class != "neither") %>%
   mutate(lagoslakeid = as.character(lagoslakeid))

### 2020-12-07 manuscript/agujournaltemplate.tex
move map to main text

@@
 \noindent
-We compiled bathymetry data on approximately 5,000 lakes in the Northeastern and Midwestern US from nine official state databases (Figure S2).
+We compiled bathymetry data on approximately 5,000 lakes in the Northeastern and Midwestern US from nine official state databases (Figure 2).
 These lakes represent a diverse cross section of lakes in terms of their surface areas (4 - 18500 ha) and span a wide geographic extent including glaciated and non-glaciated regions.
@@
 The size of the moving window was adjusted iteratively to ensure that each bathymetry raster contained no missing data.
+
+% TODO insert map
+\begin{figure}
+      \begin{center}\includegraphics[width=0.65\textwidth]{../figures/00_map_bathy-1}\end{center}
+      \caption{Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.}\label{fig2}
+      \end{figure}
 
@@
 \noindent
-Lakes belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Figure S2, S3).
+Lakes belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Figure 2, S3).
 For example, concave lakes were nearly absent from Michigan whereas Maine lakes had an overabundance of lakes categorized as neither concave nor convex.
-Lakes in the southern portions of our study area tended to be classified as reservoirs whereas lakes in the northern portions of our study area were a more even mix between reservoirs and natural lakes (Figure S2).
+Lakes in the southern portions of our study area tended to be classified as reservoirs whereas lakes in the northern portions of our study area were a more even mix between reservoirs and natural lakes (Figure 2).
 Approximately 18\%, 80\%, and 2\% of lakes were classified as having a concave, convex, or neither shape respectively whereas approximately 30\% and 70\% of lakes were classified as being a reservoir or a natural lake.
@@
 For slope measures, most lakes had lower magnitude (i.e.
-shallower) nearshore land slope compared to true in-lake slope (Figure \ref{fig2}).
+shallower) nearshore land slope compared to true in-lake slope (Figure \ref{fig3}).
 Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes whereas proxy nearshore land slope is not representative of true in-lake slope.
-The strong relationship between distance to the center of lakes and distance to the deepest point means that it is possible to convert between the two measures in subsequent analyses (See best-fit equations in Figure \ref{fig2}).
+The strong relationship between distance to the center of lakes and distance to the deepest point means that it is possible to convert between the two measures in subsequent analyses (See best-fit equations in Figure \ref{fig3}).
 
@@
   \caption{Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
-A best-fit line and coefficient of determination is shown to illustrate representativeness.}\label{fig2}
+A best-fit line and coefficient of determination is shown to illustrate representativeness.}\label{fig3}
 \end{figure}
@@
 More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.7) and lower prediction error (RMSE = 4.4m, MAPE = 29\%) compared to the proxy slope - true distance model ($R^2$ = 0.32, RMSE = 6.6m, MAPE = 60\%).
-Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure \ref{fig3}).
-We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure \ref{fig3}).
+Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure \ref{fig4}).
+We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure \ref{fig4}).
 
@@
   \end{center}
-  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig3}
+  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig4}
 \end{figure}
@@
 any permanent waterbody that has a water control structure likely to significantly impact flow or pool water) compared to that of conventional classifications that are tied to specific dam types or dam heights.
-Another possible explanation is that conventional reservoir classifications are conceptually biased towards more southern areas with few natural lakes (Figure S2).
+Another possible explanation is that conventional reservoir classifications are conceptually biased towards more southern areas with few natural lakes (Figure 2).
 
@@
 Instead of a concentrated area of concave lakes in formerly glaciated regions, there appears to be a fairly even mix of concave and convex lakes distributed amongst the northern (i.e.
-glaciated) and southern (non-glaciated) portions of our study area (Figure S2).
+glaciated) and southern (non-glaciated) portions of our study area (Figure 2).
 

### 2020-12-07 manuscript/agujournaltemplate.tex
align figure and table numbers

@@
 Reported nearshore land slope values are the mean of all points in the buffer.
-In addition to the aforementioned techniques of calculating in-lake (and nearshore) slopes and distances, we tried 7 alternate techniques which are described in Table S2.
+In addition to the aforementioned techniques of calculating in-lake (and nearshore) slopes and distances, we tried 7 alternate techniques which are described in Figure S5 and Table S2.
 
@@
 For cross-section shape, we categorized lakes as either convex or concave following the method of \citeA{hakansonLakeFormLake1977} by computing normalized lake depth-area relationships (i.e.
-hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S3).
+hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S2).
 
@@
 \noindent
-Lakes belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Figure 2, S3).
+Lakes belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Figure 2).
 For example, concave lakes were nearly absent from Michigan whereas Maine lakes had an overabundance of lakes categorized as neither concave nor convex.
@@
 In addition to overall differences between slope and distance measures, we found differences in these relationships among lake shape classes.
-For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S4).
-We found evidence that this difference was at least partly explained by the fact that convex lakes are deeper than concave lakes (Figure S5).
+For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S3).
+We found evidence that this difference was at least partly explained by the fact that convex lakes are deeper than concave lakes (Figure S8).
 There were not similar differences among slope and distance metrics for natural lakes versus reservoirs.
@@
 Offset model fit and prediction error differed depending on the technique used to calculate in-lake and nearshore geometry metrics (Table S2).
-We found that the best model fit and lowest model error occurred when in-lake slope was calculated as the average point-wise slope of all points at maximum lake depth rather than a single point of maximum depth.
+We found that the best model fit and lowest model error occurred when in-lake slope was calculated as the average point-wise slope of all points at maximum lake depth rather than at single point of maximum depth.
 However, given the small difference in the fit of models using either of these techniques and the significant cost in terms of computational load and complexity, we limit our discussion hereafter to the simpler case involving only a single deepest point.
@@
 
-The most important covariates in these models were those relating to spatial location, lake area, and perimeter (Figure S7).
+The most important covariates in these models were those relating to spatial location, lake area, and perimeter (Figure S6).
 Conversely, watershed metrics and lake elevation had little contribution to random forest model fit.
@@
 HUC4) covariate was notably less importance in the true slope model compared to the two proxy slope models.
-Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Equation \ref{eq1} (Figure S7).
+Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Equation \ref{eq1} (Figure S6).
 
@@
 Models using non-representative proxies showed increased error and systematic overestimation of depth in concave and reservoir lakes.
-Although our analysis was limited to lakes with available bathymetry data, these lakes did not have characteristics that differed from that of the overall lake population (Figure S8, S10).
+Although our analysis was limited to lakes with available bathymetry data, these lakes did not have characteristics that differed from that of the overall lake population (Figure S7, Table S1, Figure S9).
 Although there is a possibility that there is some hidden bias not explored for in our analyses, this lack of difference suggests that our results are likely to be broadly applicable to all lakes.
@@
 This finding is consistent with \citeA{heathcotePredictingBathymetricFeatures2015} whos results suggest that in-lake slopes are shallower compared to the surrounding land.
-Furthermore, the fact that in-lake slopes were shallower compared to the surrounding land even after controlling for differences in area (Figure S5) is consistent with the idea of topographic scaling (i.e.
+Furthermore, the fact that in-lake slopes were shallower compared to the surrounding land even after controlling for differences in area (Figure S4) is consistent with the idea of topographic scaling (i.e.
 scale invariance) explored in previous work and detailed by \cite{caelVolumeMeanDepth2017}.
@@
 In addition, convex lakes often had steeper in-lake slopes relative to nearshore land slopes as compared to concave lakes.
-Finally, it was notable that convex lakes were deeper than concave lakes despite having similar distributions of lake surface area (Figure S9).
+Finally, it was notable that convex lakes were deeper than concave lakes despite having similar distributions of lake surface area (Figure S8).
 The underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of lake formation whereby the formation of concave lakes were a result of more intense glacial scouring compared to that of convex lakes \cite{gorhamPhysicalLimnologyNorthern1958}.
@@
 The greater sensitivity of depth predictions to proxy slope measures relative to proxy distance measures may be explained by the fact that proxy slope measures were a more imperfect representation of true in-lake slopes relative to proxy versus true distances.
-We did not find evidence that the outsized sensivity of depth predictions to slope was dependent on variations in how these measures were calculated (Table S2).
-In a general sense, the sensitivity of depth predictions to slope help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \cite{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S9).
+We did not find evidence that the outsized sensitivity of depth predictions to slope was dependent on variations in how these measures were calculated (Table S2).
+In a general sense, the sensitivity of depth predictions to slope help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \cite{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S8).
 

### 2020-12-07 manuscript/si_template_2019.tex
move map to main text
align figure and table numbers

@@
 
-\item \textbf{Figure S2:} Map of study lakes
+\item \textbf{Figure S2:} Hypsography classification by state
 
-\item \textbf{Figure S3:} Hypsography classification by state
+\item \textbf{Figure S3:} Lake geometry comparison among shape and reservoir classes
 
-\item \textbf{Figure S4:} Lake geometry comparison among shape and reservoir classes
+\item \textbf{Figure S4:} Lake geometry comparison among shape classes by depth and area
 
-\item \textbf{Figure S5:} Lake geometry comparison among shape classes by depth and area
+\item \textbf{Figure S5:} Effect of calculation technique on in-lake/nearshore slope comparison.
 
-\item \textbf{Figure S6:} Effect of calculation technique on in-lake/nearshore slope comparison.
+\item \textbf{Figure S6:} Importance plot for random forest variables
 
-\item \textbf{Figure S7:} Importance plot for random forest variables
+\item \textbf{Figure S7:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
 
-\item \textbf{Figure S8:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
+\item \textbf{Figure S8:} Lake characteristics comparison among shape and reservoir classes
 
-\item \textbf{Figure S9:} Lake characteristics comparison among shape and reservoir classes
-
-\item \textbf{Figure S10:} Comparison between reported depth and bathymetry depth
+\item \textbf{Figure S9:} Comparison between reported depth and bathymetry depth
 
@@
 \clearpage
-
-\begin{figure}
-      \begin{center}\includegraphics[width=0.65\textwidth]{../figures/00_map_bathy-1}\end{center}
-      \caption{Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.}\label{figS2}
-      \end{figure}
 

### 2020-12-07 manuscript/si_template_2019.tex
round geometry metrics to two decimal places

@@
 \textbf{Visual center distance (m)} & 240 (-) & 160 (-) & 390 (-) & 4820 (-)\\
-Inlake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4820 (-)\\
-Inlake slope online (m/m) & 0.062 (-) & 0.029 (-) & 0.14 (-) & 4800 (-)\\
-Inlake slopes (m/m) & 0.057 (-) & 0.032 (-) & 0.098 (-) & 4820 (-)\\
-Inlake slopes online (m/m) & 0.07 (-) & 0.034 (-) & 0.15 (-) & 4800 (-)\\
-Mean inlake slope (m/m) & 0.043 (-) & 0.021 (-) & 0.09 (-) & 4820 (-)\\
-Nearshore mean slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4820 (-)\\
-Nearshore slope online (m/m) & 0.077 (-) & 0.038 (-) & 0.13 (-) & 4590 (-)\\
-Nearshore slopes online (m/m) & 0.077 (-) & 0.037 (-) & 0.13 (-) & 4540 (-)\\
+Inlake slope (m/m) & 0.05 (-) & 0.02 (-) & 0.08 (-) & 4820 (-)\\
+Inlake slope online (m/m) & 0.06 (-) & 0.03 (-) & 0.14 (-) & 4800 (-)\\
+Inlake slopes (m/m) & 0.06 (-) & 0.03 (-) & 0.1 (-) & 4820 (-)\\
+Inlake slopes online (m/m) & 0.07 (-) & 0.03 (-) & 0.15 (-) & 4800 (-)\\
+Mean inlake slope (m/m) & 0.04 (-) & 0.02 (-) & 0.09 (-) & 4820 (-)\\
+Nearshore mean slope (m/m) & 0.08 (-) & 0.05 (-) & 0.11 (-) & 4820 (-)\\
+Nearshore slope online (m/m) & 0.08 (-) & 0.04 (-) & 0.13 (-) & 4590 (-)\\
+Nearshore slopes online (m/m) & 0.08 (-) & 0.04 (-) & 0.13 (-) & 4540 (-)\\
 \hline

### 2020-12-07 tables/01_predictors.Rmd
round geometry metrics to two decimal places

@@
 dt_bathy <- data.frame(dt_bathy)
+# round geometry measures to two decimal points
+clean_geometry <- function(x) as.character(round(as.numeric(x), 2))
+dt_bathy$Median[which(dt_bathy$variable == "inlake_slope_pnt"):nrow(dt_bathy)] <-
+  clean_geometry(
+    dt_bathy$Median[which(dt_bathy$variable == "inlake_slope_pnt"):nrow(dt_bathy)])
+dt_bathy$Q25[which(dt_bathy$variable == "inlake_slope_pnt"):nrow(dt_bathy)] <-
+  clean_geometry(
+    dt_bathy$Q25[which(dt_bathy$variable == "inlake_slope_pnt"):nrow(dt_bathy)])
+dt_bathy$Q75[which(dt_bathy$variable == "inlake_slope_pnt"):nrow(dt_bathy)] <-
+  clean_geometry(
+    dt_bathy$Q75[which(dt_bathy$variable == "inlake_slope_pnt"):nrow(dt_bathy)])
+
 # compute stats for all of lagosus
@@
 dt_all <- dplyr::filter(dt_all, variable %in% dt_bathy$variable)
+
 # interleave

### 2020-12-07 scripts/05_depth_model.R
stash wrr resubmission

@@ -218,6 +218,18 @@ slope_distance_alternatives$proxy_proxy_rsq <- unlist(
   lapply(res, function(x) x$dt_metrics$rsq[4]))
 slope_distance_alternatives$proxy_proxy_mape <- unlist(
   lapply(res, function(x) x$dt_metrics$mape[4]))
+
+more_sensitive_to_slope <- function(x){
+  # x <- res[[1]]$dt_metrics
+  dplyr::filter(x, model == "maxdepth_true_false")$rmse <
+    dplyr::filter(x, model == "maxdepth_false_true")$rmse &
+  dplyr::filter(x, model == "maxdepth_true_false")$rmse <
+    dplyr::filter(x, model == "maxdepth_false_false")$rmse
+}
+slope_distance_alternatives$sensitive_to_slope <- unlist(
+  lapply(res, function(x) more_sensitive_to_slope(x$dt_metrics))
+)
+
 # View(slope_distance_alternatives)
 write.csv(slope_distance_alternatives, "data/01_depth_model/alternatives_metrics.csv",
           row.names = FALSE)

### 2021-02-15 manuscript/si_template_2019.tex
scrub wrr material from supplement, closes #25
more detail to table s1 caption

@@
 
-\title{Supporting Information for "Geometric models overestimate lake depth due to imperfect slope
-measurement"}
+\title{Supporting Information for "Imperfect slope measurements drive overestimation in geometric models of lake depth"}
 %
@@
 Predictor variables for computing random forest offsets (Equation 2) are printed in bold face.
-Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{tableS1}
+Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset.
+The total number lakes is reported as n.} \label{tableS1}
 \centering

### 2021-02-15 manuscript/si_template_2019.tex
report that categories are quantile bins

@@
       \end{center}
-      \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth.}\label{figS5}
+      \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth.
+Categories are quantile bins ($<$ 25\%, 25-50\%, 50-75\%, and 75-100\%).}\label{figS5}
     \end{figure}

### 2021-02-15 scripts/00_get_reservoir_class.R
use correct lagosus field name

@@ -23,7 +23,7 @@ pred_class_raw <- readxl::read_excel(
 dt <- dplyr::bind_rows(manual_class_raw, pred_class_raw) %>%
   arrange(lagoslakeid) %>%
   rename(lagosne_lagoslakeid = lagoslakeid) %>%
-  left_join(dplyr::select(lg$locus$locus_link, lagoslakeid, lagosne_lagoslakeid),
+  left_join(dplyr::select(lg$locus$lake_link, lagoslakeid, lagosne_lagoslakeid),
                           by = "lagosne_lagoslakeid") %>%
   dplyr::filter(!is.na(lagoslakeid)) %>%
   distinct(lagoslakeid, .keep_all = TRUE) %>%

### 2021-08-04 scripts/99_generate_eml.R
name fix

@@ -11,7 +11,7 @@ library(EML)
 
 joe <- eml$creator(
   individualName = eml$individualName(
-    givenName = "Joseph",
+    givenName = "Jemma",
     surName = "Stachelek"),
   electronicMailAddress = "stachel2@msu.edu")
 

### 2021-08-04 figures/01_geometry.Rmd
linting

--- 2020-12-03 figures/01_geometry.Rmd
@@
   summarize(dist_deepest = list(tibble::enframe(
-    quantile(dist_deepest, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+    quantile(dist_deepest, probs = c(0.05, 0.5, 0.95), na.rm = TRUE)))) %>%
   unnest() %>%
@@
   summarize(dist_viscenter = list(tibble::enframe(
-    quantile(dist_viscenter, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+    quantile(dist_viscenter, probs = c(0.05, 0.5, 0.95), na.rm = TRUE)))) %>%
   unnest() %>%
@@
 
-get_pnt_map <- function(x){
+get_pnt_map <- function(x) {
   # x <- 3
@@
 #                    )
-
 ```
@@
 
-gg_scatter_slope <- function(x, y, dt){
+gg_scatter_slope <- function(x, y, dt) {
   ggplot(data = dt) +
@@
   annotate("text", x = 2550, y = 400, label = paste0("R^2 == ", rsq),
-                                parse = TRUE, size = text_size-0.7) +
+    parse = TRUE, size = text_size - 0.7) +
   annotate("label", x = 1800, y = 150,
@@
       lm(dist_viscenter ~ dist_deepest, data = st_drop_geometry(dt))),
-           size = text_size-0.85, fill = "white", label.size = NA) +
+    size = text_size - 0.85, fill = "white", label.size = NA) +
   labs(color = "") +
@@
   annotate("text", x = 0.34, y = 0.05, label = paste0("R^2 == ", rsq),
-                                parse = TRUE, size = text_size-0.7) +
+    parse = TRUE, size = text_size - 0.7) +
   annotate("label", x = 0.27, y = 0.01,
@@
       lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)),
-           size = text_size-0.85, fill = "white", label.size = NA) +
+    size = text_size - 0.85, fill = "white", label.size = NA) +
   # geom_abline(slope = 1, intercept = 0) +
@@
 cowplot::plot_grid(
+  ggdraw() +
+    draw_label("True measures", y = 0.37, hjust = 0, size = 12, angle = 90,  fontface = "bold"),
+  cowplot::plot_grid(
+    cowplot::plot_grid(
       gg_dist_base,
@@
       rel_widths = c(1, 0.9)
-)
-
+    ),
+    ggdraw() +
+      draw_label("Proxy measures", x = 0.45, hjust = 0, size = 12, fontface = "bold"),
+    ncol = 1, rel_heights = c(1, 0.1)),
+  ncol = 2, rel_widths = c(0.1, 1))
 ```
@@
   annotate("text", x = 1800, y = 400, label = paste0("R^2 == ", rsq),
-                                parse = TRUE, size = text_size-0.7) +
+    parse = TRUE, size = text_size - 0.7) +
   annotate("label", x = 1200, y = 150,
@@
       lm(dist_viscenter ~ dists_deepest, data = dt)),
-           size = text_size-0.85, fill = "white", label.size = NA) +
+    size = text_size - 0.85, fill = "white", label.size = NA) +
   labs(color = "") +
@@
 
-gg_slope_alternative <- function(nearshore_var, inlake_var, dt){
+gg_slope_alternative <- function(nearshore_var, inlake_var, dt) {
   # dt <- test
@@
     annotate("label", x = 0.4, y = 0.52, label = paste0("R^2 == ", rsq),
-                                  parse = TRUE, size = text_size-0.7) +
+      parse = TRUE, size = text_size - 0.7) +
     # annotate("label", x = 0.27, y = 0.01,
@@
       #          parse = TRUE) +
-                       theme(#axis.title.x = element_blank(), 
+      theme( # axis.title.x = element_blank(),
         axis.title.y = element_blank()) +
@@
   rel_widths = c(1, 0.87))
-
 ```

### 2021-08-04 figures/01_geometry.Rmd
pheatmap not working, turning off slope-compare for now

@@
 
-```{r slope-compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}
+```{r slope-compare, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide', fig.height=3.5}
 # compare inlake slope, buffer slope, and online slope

### 2021-08-04 figures/01_geometry.Rmd
turn back on slope compare chunk for data processing purposes

--- 2021-08-04 figures/01_geometry.Rmd
@@
 # jsta::jheatmap
-jheatmap(
-  dplyr::select(dt,
-    matches(slope_dist_str))
-  # maxdepth,
-  # meandepth)
-)
+# jheatmap(
+#   dplyr::select(dt,
+#     matches(slope_dist_str))
+#   # maxdepth,
+#   # meandepth)
+# )
 
@@
 ```{r buffer-vs-inlake_slope, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4}
-lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
+lgus_pred <- read.csv(
+  "../data/lagosus_depth_predictors.csv",
+  stringsAsFactors = FALSE) %>%
   mutate(lagoslakeid = as.character(lagoslakeid)) %>%
-  dplyr::select(-matches("inlake_slope"),
-    -matches("dist_|dists_"))
+  dplyr::select(
+    -matches("inlake_slope"),
+    -matches("dist_|dists_")
+  )
 test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>%
   dplyr::filter(!is.na(shape_class)) %>%
+  dplyr::filter(inlake_slope_pnt < 4) %>%
   dplyr::arrange(desc(factor(shape_class,

### 2021-08-04 figures/02_depth_model.Rmd
linting

@@
 knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
-if(interactive()){
+if (interactive()) {
   setwd("figures")
@@
 
-tidy_importance <- function(x){
+tidy_importance <- function(x) {
   # x <- dt_fits[[2]]$fit$fit
   res <- ranger::importance(x) %>%
-  data.frame() %>% cbind(row.names(.), .[,1]) %>%
+    data.frame() %>% cbind(row.names(.), .[, 1]) %>%
     setNames(c("percent_inc_mse", "variable"))
   row.names(res) <- NULL
-  res <- res[,c(2, 1)]
+  res <- res[, c(2, 1)]
   res <- arrange(res, percent_inc_mse)
@@
   "proxy_proxy")
-dt_imp <- bind_rows(lapply(1:3, function(i){
+dt_imp <- bind_rows(lapply(1:3, function(i) {
   mutate(dt_imp[[i]], model = model_names[i])
@@
     yintercept = 1:nlevels(dt_imp$variable)) +
-  theme(text = element_text(family="Times"),
+  theme(text = element_text(family = "Times"),
     panel.background = element_blank(), panel.grid = element_blank(),
@@
       vjust = -0.5, size = 10),
-        axis.text = element_text(family="sans",size = 9, 
+    axis.text = element_text(family = "sans", size = 9,
       color = "black"),
@@
 
-gg_shape <- lapply(unique(dt$maxdepth), function(x){ 
+gg_shape <- lapply(unique(dt$maxdepth), function(x) {
   ggplot(data = dplyr::filter(dt, maxdepth == x)) +
@@
       strip.text = element_blank(),
-        plot.margin = unit(c(1.2,0.2,0.2,0.2), "cm"))  +
+      plot.margin = unit(c(1.2, 0.2, 0.2, 0.2), "cm"))  +
     guides(fill = guide_legend(nrow = 2))
@@
   draw_label("true", x = 0.35, y = 0.9, size = 10) +
-  draw_label("distance", x = 0.55, y = 0.96, size = 10)#  +
+  draw_label("distance", x = 0.55, y = 0.96, size = 10) #  +
 # draw_label("Residual distribution",
@@
 
-gg_reservoir <- lapply(unique(dt$maxdepth), function(x){ 
+gg_reservoir <- lapply(unique(dt$maxdepth), function(x) {
   ggplot(data = dplyr::filter(dt, maxdepth == x)) +
@@
       axis.text = element_text(size = 9.5),
-        plot.margin = unit(c(1.2,0.2,0.2,0.2), "cm"))  +
+      plot.margin = unit(c(1.2, 0.2, 0.2, 0.2), "cm"))  +
     guides(fill = guide_legend(nrow = 2))
@@
   # create some space to the left of the legend
-  gg_noshape + theme(#legend.box.margin = margin(0, 0, 0, 12), 
+  gg_noshape + theme( # legend.box.margin = margin(0, 0, 0, 12),
     legend.direction = "horizontal")
@@
 )
-
 ```

### 2021-08-04 figures/02_depth_model.Rmd
turn off 02_depth_model_fitted chunk for now

@@
 
-```{r 02_depth_model_fitted, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r 02_depth_model_fitted, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
 dt <- readRDS("../data/01_depth_model/depth_model.rds")

### 2021-08-04 figures/02_depth_model.Rmd
also turn off 02_depth_model_resid chunk for now

@@
 
-```{r 02_depth_model_resid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+```{r 02_depth_model_resid, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
 # residual = observed - predicted
@@
 # negative value = overprediction
+dt <- readRDS("../data/01_depth_model/depth_model.rds")
 

### 2021-08-04 figures/01_contrasts.Rmd
linting

@@
 ```{r define-plots, echo=FALSE}
-contrast_plot <- function(..., plot_type){
+contrast_plot <- function(..., plot_type) {
   theme_opts <- cowplot::theme_cowplot()
 
-  if(plot_type == "violin"){
+  if (plot_type == "violin") {
     violin_plot(...) + theme_opts
-  }else{
+  } else {
     pointrange_plot(...) + theme_opts
@@
 
-violin_plot <- function(x, grp, varb, label){
+violin_plot <- function(x, grp, varb, label) {
   ggplot(data = x) +
@@
 
-quantile_collapse <- function(x){
+quantile_collapse <- function(x) {
   list(tibble::enframe(
-    quantile(x, probs = c(0.05,0.5,0.95), na.rm = TRUE)))
+    quantile(x, probs = c(0.05, 0.5, 0.95), na.rm = TRUE)))
 }
 
-pointrange_plot <- function(x, grp, varb, label){
+pointrange_plot <- function(x, grp, varb, label) {
   x %>%
@@
   summarize(lake_maxdepth_m = list(tibble::enframe(
-    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
+    quantile(lake_maxdepth_m, probs = c(0.05, 0.5, 0.95), na.rm = TRUE)))) %>%
   unnest() %>%
@@
   ylab("Depth (m)") + xlab(""))
-
 ```
@@
       1000, 20000, Inf),
-             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
+    labels = ~ paste(sep = "-", thousand_k(.y[1]), thousand_k(.y[2])))
   ) %>%

### 2021-08-04 figures/01_contrasts.Rmd
correct old lagosus depth load command

@@
 # load LAGOSUS DEPTH
-lg_raw <- lagosus_load("depth")$depth$depth
+lg_raw <- lagosus_load("depth")$depth$lake_depth
 lg <- lg_raw %>%

### 2021-08-04 figures/01_contrasts.Rmd
correct old lagos depth column name

--- 2021-08-04 figures/01_contrasts.Rmd
@@
 # load LAGOSUS DEPTH
-lg_raw <- lagosus_load("depth")$depth$depth
+lg_raw <- lagosus_load("depth")$depth$lake_depth
 lg <- lg_raw %>%
   mutate(lagos_effort =
-    forcats::fct_collapse(lg_raw$lagos_effort,
+    forcats::fct_collapse(lg_raw$lake_depth_effort,
       bathymetry = "bathymetry",
-      other = c("LAGOSNE", "manual", "NLA")))
+      other = c("LAGOSNE", "LAGOSUS", "NLA")))
 
@@
 cowplot::plot_grid(
-  gg_quantdot(lg, lagos_effort, "lake_maxdepth_m") +
+  gg_quantdot(lg, lake_depth_effort, "lake_maxdepth_m") +
     ylab("Max depth (m) \n") +
@@
     theme(axis.text.x = element_blank()),
-  gg_quantdot(lg, lagos_effort, "lake_waterarea_ha") +
+  gg_quantdot(lg, lake_depth_effort, "lake_waterarea_ha") +
     scale_y_log10() +
@@
     theme(axis.text.x = element_blank()),
-  gg_quantdot(lg, lagos_effort, "urban_distance") +
+  gg_quantdot(lg, lake_depth_effort, "urban_distance") +
     scale_y_log10() +

### 2021-08-10 figures/02_depth_model.Rmd
compute and arrange complex table 1 showing shape/res sensitivity, #27

@@
 ```{r depth_model_metrics, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=2.6, fig.width=4}
-dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
+dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds")[[1]] %>%
   mutate_if(is.numeric, function(x) as.character(round(x, 2))) %>%

### 2021-08-10 tables/02_model_metrics.Rmd
linting?
linting
compute and arrange complex table 1 showing shape/res sensitivity, #27

@@
 
-dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
+tidy_grid_metrics <- function(x) {
+  res <- x %>%
     mutate_if(is.numeric, function(x) as.character(round(x, 2))) %>%
@@
     dplyr::select(slope, distance, everything(), -.estimator)
+  res
+}
 
-options(knitr.kable.NA = '-')
+dt_metrics <- tidy_grid_metrics(
+  readRDS("../data/01_depth_model/depth_grid_metrics.rds")[[1]])
+
+options(knitr.kable.NA = "-")
 knitr::kable(dt_metrics, digits = 2, format = "markdown")
@@
 # escape = TRUE))
+```
+
+\clearpage
+
+```{r 02_model_sensitivity, eval=TRUE}
+dt_metrics_sensitivity <- lapply(
+  readRDS("../data/01_depth_model/depth_grid_metrics.rds"), function(x)
+    tidy_grid_metrics(x))
+
+labels <- c("All", "Res only", "NL only", "convex only", "concave only")
+dt_metrics_sensitivity <- lapply(
+  seq_len(length(dt_metrics_sensitivity)), function(i) {
+    res <- dplyr::mutate(dt_metrics_sensitivity[[i]], label = labels[i]) %>%
+      dplyr::relocate(label)
+    res$label[2:4] <- NA
+    res
+  })
+
+options(knitr.kable.NA = "-")
+knitr::kable(do.call("rbind", dt_metrics_sensitivity),
+  digits = 2, format = "markdown")
 ```
@@
 
-options(knitr.kable.NA = '-')
+options(knitr.kable.NA = "-")
 knitr::kable(dt_metrics, digits = 2, format = "markdown")

### 2021-08-10 scripts/05_depth_model.R

@@ -225,11 +225,16 @@ res <- lapply(seq_len(nrow(slope_distance_alternatives)),
   })
 
 # add shape_class and reservoir_class sensitivity for Table 1
-test <- res
-test <- c(test, list(fit_alternative(1, nl_res = "Res")))
-test <- c(test, list(fit_alternative(1, nl_res = "NL")))
-test <- c(test, list(fit_alternative(1, concave_convex = "convex")))
-test <- c(test, list(fit_alternative(1, concave_convex = "concave")))
+# test <- res
+shape_reservoir_sensitivity <- c(
+  res[1],
+  list(fit_alternative(1, nl_res = "Res")),
+  list(fit_alternative(1, nl_res = "NL")),
+  list(fit_alternative(1, concave_convex = "convex")),
+  list(fit_alternative(1, concave_convex = "concave"))
+)
+depth_grid_metrics <- lapply(shape_reservoir_sensitivity,
+  function(x) x$dt_metrics)
 
 # lapply res extract proxy_proxy stats and add to slope_distance_alternatives
 slope_distance_alternatives$proxy_proxy_rmse <- unlist(
@@ -258,7 +263,7 @@ write.csv(slope_distance_alternatives,
 saveRDS(res[[1]]$dt_fits, "data/01_depth_model/depth_fits.rds")
 saveRDS(res[[1]]$dt_grid,
   "data/01_depth_model/depth_grid.rds")
-saveRDS(res[[1]]$dt_metrics,
+saveRDS(depth_grid_metrics,
   "data/01_depth_model/depth_grid_metrics.rds")
 saveRDS(res[[17]]$dt_metrics,
   "data/01_depth_model/runner_up_metrics.rds")

### 2021-08-10 figures/00_maps.Rmd
linting

@@
   ggplot() +
-    geom_sf(data = study_bbox[1,], alpha = 1, size = 0.6, 
+  geom_sf(data = study_bbox[1, ], alpha = 1, size = 0.6,
     color = "black", fill = "#1D9DCF") +
@@
 
-us_states <-dplyr::filter(
+us_states <- dplyr::filter(
   dplyr::select(states_all, postal, iso_a2), iso_a2 == "US")

### 2021-08-10 figures/00_maps.Rmd
linting

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
+knitr::opts_chunk$set(echo = TRUE,
+  fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
 # setwd("figures")
@@
 states_focal  <- st_read("../data/gis.gpkg", "states_focal", quiet = TRUE)
-# hu4_focal     <- st_read("../data/gis.gpkg", "hu4s_focal_simple", quiet = TRUE)
+# hu4_focal     <- st_read("../data/gis.gpkg", "hu4s_focal_simple",
+#   quiet = TRUE)
 # hu4_focal     <- st_crop(st_transform(
@@
 dt_bathy <- st_sf(dt_bathy, geometry = dt_bathy$geom)
-dt_bathy <- mutate(dt_bathy, maxdepth_class = droplevels(dt_bathy$maxdepth_class))
+dt_bathy <- mutate(dt_bathy,
+  maxdepth_class = droplevels(dt_bathy$maxdepth_class))
 dt_bathy <- dplyr::filter(dt_bathy, !is.na(maxdepth_class))
@@
     color = "black", fill = "#1D9DCF") +
-  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))), fill = "gray", color = "black", size = 0.3) +
-  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))), fill = "white", color = "black", size = 0.4) +
+  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))),
+    fill = "gray", color = "black", size = 0.3) +
+  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))),
+    fill = "white", color = "black", size = 0.4) +
   geom_sf(data = dt_bathy, aes(color = maxdepth_class),
@@
 ```{r gg_colocated, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
-is_colocated <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
+is_colocated <- read.csv("../data/lagosus_depth.csv",
+  stringsAsFactors = FALSE) %>%
   dplyr::filter(
@@
       !is.na(lake_maxdepth_m) &
-      !(lagoslakeid %in% c(456169))) %>% # remove llids with duplicate manual depth)
+      # remove llids with duplicate manual depth)
+      !(lagoslakeid %in% c(456169))) %>%
   dplyr::select(lagoslakeid, lake_state, lagos_effort, lake_maxdepth_m) %>%
@@
     color = "black", fill = "#1D9DCF") +
-  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))), fill = "gray", color = "black", size = 0.3) +
-  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))), fill = "white", color = "black", size = 0.4) +
+  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))),
+    fill = "gray", color = "black", size = 0.3) +
+  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))),
+    fill = "white", color = "black", size = 0.4) +
   geom_sf(data = dt_bathy, aes(color = colocated),
@@
     color = "black", fill = "#1D9DCF") +
-  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))), fill = "gray", color = "black", size = 0.3) +
-  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))), fill = "white", color = "black", size = 0.4) +
-  geom_sf(data = dplyr::filter(dt, shape_class == "convex"), aes(color = shape_class),
-    size = 0.03, alpha = 0.3) +
-  geom_sf(data = dplyr::filter(dt, shape_class != "convex"), aes(color = shape_class),
-    size = 0.07, alpha = 0.8) +
+  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))),
+    fill = "gray", color = "black", size = 0.3) +
+  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))),
+    fill = "white", color = "black", size = 0.4) +
+  geom_sf(data = dplyr::filter(dt, shape_class == "convex"),
+    aes(color = shape_class), size = 0.03, alpha = 0.3) +
+  geom_sf(data = dplyr::filter(dt, shape_class != "convex"),
+    aes(color = shape_class), size = 0.07, alpha = 0.8) +
   labs(color = "Shape class     ") +
@@
     color = "black", fill = "#1D9DCF") +
-  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))), fill = "gray", color = "black", size = 0.3) +
-  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))), fill = "white", color = "black", size = 0.4) +
+  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))),
+    fill = "gray", color = "black", size = 0.3) +
+  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))),
+    fill = "white", color = "black", size = 0.4) +
   geom_sf(data = dplyr::filter(dt, !is.na(reservoir_class)),

### 2021-08-10 manuscript/figures.Rmd
sync figures Rmd with submitting ms

@@
 knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", dev = c("pdf"))
-knitr::opts_chunk$set(fig.pos = 'H')
+knitr::opts_chunk$set(fig.pos = "H")
 ```
@@
 
-```{r 01_geometry_base, fig.cap = "Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
-A best-fit line and equation is shown to shown to facilitate computation of correction factors for proxy values of lake geometry.
-Coefficients of determination are shown to illustrate representativeness.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
-knitr::include_graphics("../figures/01_geometry_base-1.pdf")
+```{r 00_map_bathy, fig.cap = "Map of study lakes showing A) lake e maximum depth measurements, B) cross-section shape class, and C) reservoir classification.
+The distribution of lake depths from panel A is reported in Figure S4.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/00_map_bathy-1.pdf")
 ```
+
+\vspace{5em}
 
@@
 
+```{r 01_geometry_base, fig.cap = "Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
+A best-fit line and equation is shown to shown to facilitate computation of correction factors for proxy values of lake geometry.
+Coefficients of determination are shown to illustrate representativeness.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+knitr::include_graphics("../figures/01_geometry_base-1.pdf")
+```
+
 ```{r depth_model_grid_resid, fig.cap = "Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}

### 2021-08-12 manuscript/si_template_2019.tex
linting

@@
       % \noindent\textbf{Introduction}
-%Type or paste your text here. The introduction gives a brief overview of the supporting information. You should include information %about as many of the following as possible (when appropriate):
+      %Type or paste your text here.
+The introduction gives a brief overview of the supporting information.
+You should include information %about as many of the following as possible (when appropriate):
       % 1. a general overview of the kind of data files;
@@
             \hline
-Variable & Median & Q25 & Q75 & n\\
+            Variable                            & Median      & Q25         & Q75          & n            \\
             \hline
-Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4820 (17700)\\
-\textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4820 (17700)\\
-\textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4820 (17700)\\
-\textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4820 (17700)\\
-\textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4820 (17700)\\
-\textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4820 (17700)\\
-\textbf{Watershed-lake ratio} & 7.8 (10) & 3.8 (4.4) & 17 (29) & 4820 (17700)\\
-\textbf{Deepest point distance (m)} & 180 (-) & 110 (-) & 290 (-) & 4820 (-)\\
-Mean deepest point distance (m) & 140 (-) & 87 (-) & 230 (-) & 4820 (-)\\
-\textbf{Visual center distance (m)} & 240 (-) & 160 (-) & 390 (-) & 4820 (-)\\
-Inlake slope (m/m) & 0.05 (-) & 0.02 (-) & 0.08 (-) & 4820 (-)\\
-Inlake slope online (m/m) & 0.06 (-) & 0.03 (-) & 0.14 (-) & 4800 (-)\\
-Inlake slopes (m/m) & 0.06 (-) & 0.03 (-) & 0.1 (-) & 4820 (-)\\
-Inlake slopes online (m/m) & 0.07 (-) & 0.03 (-) & 0.15 (-) & 4800 (-)\\
-Mean inlake slope (m/m) & 0.04 (-) & 0.02 (-) & 0.09 (-) & 4820 (-)\\
-Nearshore mean slope (m/m) & 0.08 (-) & 0.05 (-) & 0.11 (-) & 4820 (-)\\
-Nearshore slope online (m/m) & 0.08 (-) & 0.04 (-) & 0.13 (-) & 4590 (-)\\
-Nearshore slopes online (m/m) & 0.08 (-) & 0.04 (-) & 0.13 (-) & 4540 (-)\\
+            Max depth (m)                       & 8.2 (7)     & 4.6 (3.7)   & 14 (12)      & 4820 (17700) \\
+            \textbf{Area (ha)}                  & 55 (33)     & 21 (11)     & 140 (100)    & 4820 (17700) \\
+            \textbf{Island area (ha)}           & 0 (0)       & 0 (0)       & 0.18 (0.076) & 4820 (17700) \\
+            \textbf{Perimeter (m)}              & 4400 (3500) & 2500 (1800) & 8100 (7300)  & 4820 (17700) \\
+            \textbf{Shoreline development}      & 1.7 (1.7)   & 1.4 (1.4)   & 2.1 (2.2)    & 4820 (17700) \\
+            \textbf{Elevation (m)}              & 300 (340)   & 180 (210)   & 400 (460)    & 4820 (17700) \\
+            \textbf{Watershed-lake ratio}       & 7.8 (10)    & 3.8 (4.4)   & 17 (29)      & 4820 (17700) \\
+            \textbf{Deepest point distance (m)} & 180 (-)     & 110 (-)     & 290 (-)      & 4820 (-)     \\
+            Mean deepest point distance (m)     & 140 (-)     & 87 (-)      & 230 (-)      & 4820 (-)     \\
+            \textbf{Visual center distance (m)} & 240 (-)     & 160 (-)     & 390 (-)      & 4820 (-)     \\
+            Inlake slope (m/m)                  & 0.05 (-)    & 0.02 (-)    & 0.08 (-)     & 4820 (-)     \\
+            Inlake slope online (m/m)           & 0.06 (-)    & 0.03 (-)    & 0.14 (-)     & 4800 (-)     \\
+            Inlake slopes (m/m)                 & 0.06 (-)    & 0.03 (-)    & 0.1 (-)      & 4820 (-)     \\
+            Inlake slopes online (m/m)          & 0.07 (-)    & 0.03 (-)    & 0.15 (-)     & 4800 (-)     \\
+            Mean inlake slope (m/m)             & 0.04 (-)    & 0.02 (-)    & 0.09 (-)     & 4820 (-)     \\
+            Nearshore mean slope (m/m)          & 0.08 (-)    & 0.05 (-)    & 0.11 (-)     & 4820 (-)     \\
+            Nearshore slope online (m/m)        & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4590 (-)     \\
+            Nearshore slopes online (m/m)       & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4540 (-)     \\
             \hline
@@
             \hline
-Inlake slope & Nearshore slope & Inlake distance & RMSE & $R^2$ & MAPE\\
+            Inlake slope         & Nearshore slope      & Inlake distance        & RMSE           & $R^2$         & MAPE           \\
             \hline
-slope\_pnts & slope\_mean & dists\_deepest & 6.2 m & 0.38 & 58 \%\\
-\textbf{slope\_pnt} & \textbf{slope\_mean} & \textbf{dist\_deepest} & \textbf{6.4 m} & \textbf{0.35} & \textbf{59 \%}\\
-slope\_pnts & slopes\_online\_mean & dist\_deepest & 6.4 m & 0.32 & 61 \%\\
-slope\_online\_mean & slope\_mean & dists\_deepest & 6.5 m & 0.41 & 63 \%\\
-slope\_pnts & slope\_mean & dist\_deepest & 6.7 m & 0.44 & 58 \%\\
-slope\_online\_mean & slope\_mean & dist\_deepest & 6.7 m & 0.36 & 59 \%\\
-slope\_online\_mean & slopes\_online\_mean & dist\_deepest & 6.7 m & 0.32 & 66 \%\\
-slope\_mean & slope\_mean & dists\_deepest & 6.8 m & 0.36 & 59 \%\\
-slope\_pnt & slopes\_online\_mean & dists\_deepest & 6.8 m & 0.25 & 73 \%\\
-slope\_pnts & slope\_online\_mean & dist\_deepest & 6.9 m & 0.3 & 71 \%\\
-slope\_online\_mean & slope\_online\_mean & dist\_deepest & 6.9 m & 0.32 & 68 \%\\
-slope\_online\_mean & slope\_online\_mean & dists\_deepest & 6.9 m & 0.33 & 65 \%\\
-slope\_mean & slopes\_online\_mean & dists\_deepest & 7 m & 0.24 & 65 \%\\
-slope\_mean & slope\_mean & dist\_deepest & 7.1 m & 0.4 & 64 \%\\
-slopes\_online\_mean & slope\_mean & dist\_deepest & 7.1 m & 0.37 & 56 \%\\
-slope\_mean & slope\_online\_mean & dist\_deepest & 7.1 m & 0.3 & 69 \%\\
-slopes\_online\_mean & slopes\_online\_mean & dists\_deepest & 7.2 m & 0.32 & 63 \%\\
-slopes\_online\_mean & slopes\_online\_mean & dist\_deepest & 7.3 m & 0.25 & 64 \%\\
-slope\_pnt & slope\_mean & dists\_deepest & 7.3 m & 0.35 & 61 \%\\
-slopes\_online\_mean & slope\_mean & dists\_deepest & 7.3 m & 0.36 & 60 \%\\
-slope\_online\_mean & slopes\_online\_mean & dists\_deepest & 7.3 m & 0.29 & 58 \%\\
-slope\_pnts & slope\_online\_mean & dists\_deepest & 7.4 m & 0.27 & 64 \%\\
-slopes\_online\_mean & slope\_online\_mean & dists\_deepest & 7.4 m & 0.33 & 67 \%\\
-slopes\_online\_mean & slope\_online\_mean & dist\_deepest & 7.5 m & 0.26 & 61 \%\\
-slope\_pnt & slopes\_online\_mean & dist\_deepest & 7.5 m & 0.33 & 69 \%\\
-slope\_mean & slopes\_online\_mean & dist\_deepest & 7.6 m & 0.26 & 64 \%\\
-slope\_pnt & slope\_online\_mean & dist\_deepest & 7.7 m & 0.27 & 68 \%\\
-slope\_pnts & slopes\_online\_mean & dists\_deepest & 7.8 m & 0.3 & 65 \%\\
-slope\_pnt & slope\_online\_mean & dists\_deepest & 7.9 m & 0.27 & 67 \%\\
-slope\_mean & slope\_online\_mean & dists\_deepest & 7.9 m & 0.31 & 60 \%\\
+            slope\_pnts          & slope\_mean          & dists\_deepest         & 6.2 m          & 0.38          & 58 \%          \\
+            \textbf{slope\_pnt}  & \textbf{slope\_mean} & \textbf{dist\_deepest} & \textbf{6.4 m} & \textbf{0.35} & \textbf{59 \%} \\
+            slope\_pnts          & slopes\_online\_mean & dist\_deepest          & 6.4 m          & 0.32          & 61 \%          \\
+            slope\_online\_mean  & slope\_mean          & dists\_deepest         & 6.5 m          & 0.41          & 63 \%          \\
+            slope\_pnts          & slope\_mean          & dist\_deepest          & 6.7 m          & 0.44          & 58 \%          \\
+            slope\_online\_mean  & slope\_mean          & dist\_deepest          & 6.7 m          & 0.36          & 59 \%          \\
+            slope\_online\_mean  & slopes\_online\_mean & dist\_deepest          & 6.7 m          & 0.32          & 66 \%          \\
+            slope\_mean          & slope\_mean          & dists\_deepest         & 6.8 m          & 0.36          & 59 \%          \\
+            slope\_pnt           & slopes\_online\_mean & dists\_deepest         & 6.8 m          & 0.25          & 73 \%          \\
+            slope\_pnts          & slope\_online\_mean  & dist\_deepest          & 6.9 m          & 0.3           & 71 \%          \\
+            slope\_online\_mean  & slope\_online\_mean  & dist\_deepest          & 6.9 m          & 0.32          & 68 \%          \\
+            slope\_online\_mean  & slope\_online\_mean  & dists\_deepest         & 6.9 m          & 0.33          & 65 \%          \\
+            slope\_mean          & slopes\_online\_mean & dists\_deepest         & 7 m            & 0.24          & 65 \%          \\
+            slope\_mean          & slope\_mean          & dist\_deepest          & 7.1 m          & 0.4           & 64 \%          \\
+            slopes\_online\_mean & slope\_mean          & dist\_deepest          & 7.1 m          & 0.37          & 56 \%          \\
+            slope\_mean          & slope\_online\_mean  & dist\_deepest          & 7.1 m          & 0.3           & 69 \%          \\
+            slopes\_online\_mean & slopes\_online\_mean & dists\_deepest         & 7.2 m          & 0.32          & 63 \%          \\
+            slopes\_online\_mean & slopes\_online\_mean & dist\_deepest          & 7.3 m          & 0.25          & 64 \%          \\
+            slope\_pnt           & slope\_mean          & dists\_deepest         & 7.3 m          & 0.35          & 61 \%          \\
+            slopes\_online\_mean & slope\_mean          & dists\_deepest         & 7.3 m          & 0.36          & 60 \%          \\
+            slope\_online\_mean  & slopes\_online\_mean & dists\_deepest         & 7.3 m          & 0.29          & 58 \%          \\
+            slope\_pnts          & slope\_online\_mean  & dists\_deepest         & 7.4 m          & 0.27          & 64 \%          \\
+            slopes\_online\_mean & slope\_online\_mean  & dists\_deepest         & 7.4 m          & 0.33          & 67 \%          \\
+            slopes\_online\_mean & slope\_online\_mean  & dist\_deepest          & 7.5 m          & 0.26          & 61 \%          \\
+            slope\_pnt           & slopes\_online\_mean & dist\_deepest          & 7.5 m          & 0.33          & 69 \%          \\
+            slope\_mean          & slopes\_online\_mean & dist\_deepest          & 7.6 m          & 0.26          & 64 \%          \\
+            slope\_pnt           & slope\_online\_mean  & dist\_deepest          & 7.7 m          & 0.27          & 68 \%          \\
+            slope\_pnts          & slopes\_online\_mean & dists\_deepest         & 7.8 m          & 0.3           & 65 \%          \\
+            slope\_pnt           & slope\_online\_mean  & dists\_deepest         & 7.9 m          & 0.27          & 67 \%          \\
+            slope\_mean          & slope\_online\_mean  & dists\_deepest         & 7.9 m          & 0.31          & 60 \%          \\
             \hline

### 2021-08-12 manuscript/figures.Rmd
move table S1 and Figure S6 to the main figures doc, #29

@@
 \captionsetup[figure]{labelformat=empty}
-```{r depth_model_metrics, fig.cap = "Table 1: Model fit and predictive accuracy metrics (RMSE = root mean square error, R2 = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+\captionsetup[figure]{labelformat=empty}
+```{r depth_model_metrics, fig.cap = "Table 1: Summary of lake characteristics for the present study (and for lakes in the contiguous United States).
+Predictor variables for computing random forest offsets (Equation 2) are printed in bold face.
+Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset.
+The total number lakes is reported as n.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
+# see tables/02_model_metrics.Rmd
+knitr::include_graphics("../tables/01_predictors.pdf")
+```
+
+```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuracy metrics (RMSE = root mean square error, R2 = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
 # see tables/02_model_metrics.Rmd
@@
 ```
+
+```{r depth_mode_importance, fig.cap="Importance plot for random forest variables showing increase in mean square error.
+Higher values indicate greater importance to model predictions.
+See Equation 1 for a definition of geometric max depth.
+HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
+knitr::include_graphics("../figures/02_depth_model_importance-1")
+```

### 2021-08-12 manuscript/si_template_2019.tex
move table S1 and Figure S6 to the main figures doc, #29

@@
 
-            \item \textbf{Figure S6:} Importance plot for random forest variables
+            \item \textbf{Figure S6:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
 
-            \item \textbf{Figure S7:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
+            \item \textbf{Figure S7:} Lake characteristics comparison among shape and reservoir classes
 
-            \item \textbf{Figure S8:} Lake characteristics comparison among shape and reservoir classes
+            \item \textbf{Figure S8:} Comparison between reported depth and bathymetry depth
 
-            \item \textbf{Figure S9:} Comparison between reported depth and bathymetry depth
-
-            \item \textbf{Table S1:} Summary of lake characteristics for the present study
-
-            \item \textbf{Table S2:} Effect of calculation technique on depth model fit and accuracy.
+            \item \textbf{Table S1:} Effect of calculation technique on depth model fit and accuracy.
 
@@
 
-\begin{figure}
-      \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
-      \caption{Importance plot for random forest variables showing increase in mean square error.
+% \begin{figure}
+%       \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
+%       \caption{Importance plot for random forest variables showing increase in mean square error.
 Higher values indicate greater importance to model predictions.
@@
 HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figS7}
-\end{figure}
+% \end{figure}
 
@@
 
-% source at tables/01_predictors.Rmd
-\begin{table}
-      \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States).
+% % source at tables/01_predictors.Rmd
+% \begin{table}
+%       \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States).
 Predictor variables for computing random forest offsets (Equation 2) are printed in bold face.
@@
 The total number lakes is reported as n.} \label{tableS1}
-      \centering
-      \setlength\tabcolsep{1.5pt} % default value: 6pt
-      \begin{tabular}{lllll}
-            \hline
-            Variable                            & Median      & Q25         & Q75          & n            \\
-            \hline
-            Max depth (m)                       & 8.2 (7)     & 4.6 (3.7)   & 14 (12)      & 4820 (17700) \\
-            \textbf{Area (ha)}                  & 55 (33)     & 21 (11)     & 140 (100)    & 4820 (17700) \\
-            \textbf{Island area (ha)}           & 0 (0)       & 0 (0)       & 0.18 (0.076) & 4820 (17700) \\
-            \textbf{Perimeter (m)}              & 4400 (3500) & 2500 (1800) & 8100 (7300)  & 4820 (17700) \\
-            \textbf{Shoreline development}      & 1.7 (1.7)   & 1.4 (1.4)   & 2.1 (2.2)    & 4820 (17700) \\
-            \textbf{Elevation (m)}              & 300 (340)   & 180 (210)   & 400 (460)    & 4820 (17700) \\
-            \textbf{Watershed-lake ratio}       & 7.8 (10)    & 3.8 (4.4)   & 17 (29)      & 4820 (17700) \\
-            \textbf{Deepest point distance (m)} & 180 (-)     & 110 (-)     & 290 (-)      & 4820 (-)     \\
-            Mean deepest point distance (m)     & 140 (-)     & 87 (-)      & 230 (-)      & 4820 (-)     \\
-            \textbf{Visual center distance (m)} & 240 (-)     & 160 (-)     & 390 (-)      & 4820 (-)     \\
-            Inlake slope (m/m)                  & 0.05 (-)    & 0.02 (-)    & 0.08 (-)     & 4820 (-)     \\
-            Inlake slope online (m/m)           & 0.06 (-)    & 0.03 (-)    & 0.14 (-)     & 4800 (-)     \\
-            Inlake slopes (m/m)                 & 0.06 (-)    & 0.03 (-)    & 0.1 (-)      & 4820 (-)     \\
-            Inlake slopes online (m/m)          & 0.07 (-)    & 0.03 (-)    & 0.15 (-)     & 4800 (-)     \\
-            Mean inlake slope (m/m)             & 0.04 (-)    & 0.02 (-)    & 0.09 (-)     & 4820 (-)     \\
-            Nearshore mean slope (m/m)          & 0.08 (-)    & 0.05 (-)    & 0.11 (-)     & 4820 (-)     \\
-            Nearshore slope online (m/m)        & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4590 (-)     \\
-            Nearshore slopes online (m/m)       & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4540 (-)     \\
-            \hline
-      \end{tabular}
-\end{table}
+%       \centering
+%       \setlength\tabcolsep{1.5pt} % default value: 6pt
+%       \begin{tabular}{lllll}
+%             \hline
+%             Variable                            & Median      & Q25         & Q75          & n            \\
+%             \hline
+%             Max depth (m)                       & 8.2 (7)     & 4.6 (3.7)   & 14 (12)      & 4820 (17700) \\
+%             \textbf{Area (ha)}                  & 55 (33)     & 21 (11)     & 140 (100)    & 4820 (17700) \\
+%             \textbf{Island area (ha)}           & 0 (0)       & 0 (0)       & 0.18 (0.076) & 4820 (17700) \\
+%             \textbf{Perimeter (m)}              & 4400 (3500) & 2500 (1800) & 8100 (7300)  & 4820 (17700) \\
+%             \textbf{Shoreline development}      & 1.7 (1.7)   & 1.4 (1.4)   & 2.1 (2.2)    & 4820 (17700) \\
+%             \textbf{Elevation (m)}              & 300 (340)   & 180 (210)   & 400 (460)    & 4820 (17700) \\
+%             \textbf{Watershed-lake ratio}       & 7.8 (10)    & 3.8 (4.4)   & 17 (29)      & 4820 (17700) \\
+%             \textbf{Deepest point distance (m)} & 180 (-)     & 110 (-)     & 290 (-)      & 4820 (-)     \\
+%             Mean deepest point distance (m)     & 140 (-)     & 87 (-)      & 230 (-)      & 4820 (-)     \\
+%             \textbf{Visual center distance (m)} & 240 (-)     & 160 (-)     & 390 (-)      & 4820 (-)     \\
+%             Inlake slope (m/m)                  & 0.05 (-)    & 0.02 (-)    & 0.08 (-)     & 4820 (-)     \\
+%             Inlake slope online (m/m)           & 0.06 (-)    & 0.03 (-)    & 0.14 (-)     & 4800 (-)     \\
+%             Inlake slopes (m/m)                 & 0.06 (-)    & 0.03 (-)    & 0.1 (-)      & 4820 (-)     \\
+%             Inlake slopes online (m/m)          & 0.07 (-)    & 0.03 (-)    & 0.15 (-)     & 4800 (-)     \\
+%             Mean inlake slope (m/m)             & 0.04 (-)    & 0.02 (-)    & 0.09 (-)     & 4820 (-)     \\
+%             Nearshore mean slope (m/m)          & 0.08 (-)    & 0.05 (-)    & 0.11 (-)     & 4820 (-)     \\
+%             Nearshore slope online (m/m)        & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4590 (-)     \\
+%             Nearshore slopes online (m/m)       & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4540 (-)     \\
+%             \hline
+%       \end{tabular}
+% \end{table}
 

### 2021-08-12 manuscript/figures.Rmd
typo fixes, #29

@@
 \captionsetup[figure]{labelformat=empty}
-```{r depth_model_metrics, fig.cap = "Table 1: Summary of lake characteristics for the present study (and for lakes in the contiguous United States).
+```{r predictors, fig.cap = "Table 1: Summary of lake characteristics for the present study (and for lakes in the contiguous United States).
 Predictor variables for computing random forest offsets (Equation 2) are printed in bold face.
@@
 HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
-knitr::include_graphics("../figures/02_depth_model_importance-1")
+knitr::include_graphics("../figures/02_depth_model_importance-1.pdf")
 ```

### 2021-08-12 README.Rmd
sync title with ms text
sync title
update title, add reproducibility section to README

@@
 
-**Stachelek et al. In prep**. Bathymetry data from thousands of lakes show that lake depth prediction is confounded by difficulty modeling inlake slope.
+**Stachelek et al. In prep**. Imperfect slope measurements drive overestimation in standard cone model of lake depth
 
@@
 ```
+
+### Reproducibility
+
+```shell
+# local
+make -t data/00_hypso/hypso.csv
+make -t data/00_bathy_depth/bathy_pnts.rds
+make -t data/00_bathy_depth/00_bathy_depth.csv
+make -t data/00_bathy_depth/bathy_geometry.csv
+make -t data/00_hypso/hypso_classes.csv
+make -t data/00_reservoir_classification/reservoir_classes_clean.csv
+make -t data/lagosus_depth_predictors.csv
+make -t data/lagosne_depth_predictors.csv
+make -t data/00_geometry/nearshore.csv
+```
+
+```shell
+# docker
+
+```

### 2021-08-12 manuscript/manuscript.Rmd
stash pre dissertation edit text
name fix
update title, add reproducibility section to README

@@
 ---
-title: Geometric models overestimate lake depth due to imperfect slope measurement
+title: Imperfect slope measurements drive overestimation in standard cone model of lake depth
 
@@
 author:
-  - name: Joseph Stachelek
+  - name: Jemma Stachelek
     affiliation: a
-  - name: Patrick Hanly
+  - name: Patrick J. Hanly
     affiliation: a

### 2021-08-12 tables/01_predictors.Rmd
add word output

@@
 }
+
+# ---- Word output ----
+
+# flextable::save_as_docx(flextable::flextable(dt), path = "../manuscript/inland_waters/table1.docx")
 ```

### 2021-08-12 tables/02_model_metrics.Rmd
add word output

@@
 # escape = TRUE))
+# flextable::save_as_docx(flextable::flextable(dt_metrics), path = "../manuscript/inland_waters/table2.docx")
 ```

### 2021-08-16 tables/02_model_metrics.Rmd
render more complex table 1(2), closes #27

@@
 
-dt_metrics <- tidy_grid_metrics(
-  readRDS("../data/01_depth_model/depth_grid_metrics.rds")[[1]])
+dt_raw <- readRDS("../data/01_depth_model/depth_grid_metrics.rds")
+dt_metrics <- lapply(dt_raw, tidy_grid_metrics)
+dt_metrics <- do.call("rbind", dt_metrics)
+dt_metrics$filter <- ""
+dt_metrics$filter[c(1, 5, 9, 13, 17)] <- c("all", "reservoir", "natural lake", "convex", "concave")
+# dt_metrics$filter <- c(rep("all", 4), rep("reservoir", 4),
+#   rep("natural lake", 4), rep("convex", 4), rep("concave", 4))
+dt_metrics <- dplyr::relocate(dt_metrics, filter)
 

### 2021-08-16 figures/00_maps.Rmd
add supp fig showing map of model residuals

@@
 ```
+
+```{r 00_map_resid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6.5}
+dt_sf <- st_read("../data/gis.gpkg", "dt", quiet = TRUE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid))
+dt   <- readRDS("../data/01_depth_model/depth_grid.rds") %>% # 05_depth_model.R
+  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
+  mutate(resid = case_when(maxdepth == "maxdepth_true_true" ~ NA_real_,
+    TRUE ~ resid)) %>%
+  mutate(maxdepth = factor(maxdepth, levels = c(
+    "maxdepth_true_true", "maxdepth_true_false",
+    "maxdepth_false_true", "maxdepth_false_false"))) %>%
+  left_join(dt_sf, by = "lagoslakeid") %>%
+  st_as_sf()
+dt <- cbind(dt, st_coordinates(dt))
+
+dt <- dt %>% mutate(resid_class =
+  smart_cut(resid, c(-20, -10, 0, 10, 20, max(resid)),
+    brackets = NULL, sep = " - ")
+)
+
+# ----
+
+gg_resid <-
+  ggplot() +
+  geom_sf(data = st_as_sfc(st_bbox(dt)), alpha = 1, size = 0.6,
+    color = "black", fill = "#1D9DCF") +
+  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt))),
+    fill = "gray", color = "black", size = 0.3) +
+  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt))),
+    fill = "white", color = "black", size = 0.4) +
+  geom_sf(data = dplyr::filter(dt, !is.na(resid)),
+    aes(color = resid_class), size = 0.03, alpha = 0.7) +
+  labs(color = "Residual") +
+  coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
+  theme_void()
+
+gg_resid
+```

### 2021-08-16 manuscript/si_template_2019.tex
update title, add reproducibility section to README
add supp fig showing map of model residuals

@@
 
-\title{Supporting Information for "Imperfect slope measurements drive overestimation in geometric models of lake depth"}
+\title{Supporting Information for "Imperfect slope measurements drive overestimation in standard cone model of lake depth"}
 %
@@
 
-            \item \textbf{Figure S6:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
+            \item \textbf{Figure S6:} Spatial distribution of depth model residuals.
 
-            \item \textbf{Figure S7:} Lake characteristics comparison among shape and reservoir classes
+            \item \textbf{Figure S7:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
 
-            \item \textbf{Figure S8:} Comparison between reported depth and bathymetry depth
+            \item \textbf{Figure S8:} Lake characteristics comparison among shape and reservoir classes
+
+            \item \textbf{Figure S9:} Comparison between reported depth and bathymetry depth
 
@@
       \noindent\includegraphics[width=0.85\textwidth]{../figures/01_hypsography-1}
-      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figS3}
+      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figS2}
 \end{figure}
@@
 A dashed 1:1 line is shown for comparison.
-Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figS4}
+Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figS3}
 \end{figure}
@@
       \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth.
-Categories are quantile bins ($<$ 25\%, 25-50\%, 50-75\%, and 75-100\%).}\label{figS5}
+Categories are quantile bins ($<$ 25\%, 25-50\%, 50-75\%, and 75-100\%).}\label{figS4}
 \end{figure}
@@
 \texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope).
-\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth.}\label{figS6}
+\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth.}\label{figS5}
+\end{figure}
+\clearpage
+
+\begin{figure}
+      \begin{center}
+            \includegraphics[width=\textwidth,keepaspectratio]{../figures/00_map_resid-1}
+      \end{center}
+      \caption{Spatial distribution of depth model residuals.}\label{figS6}
 \end{figure}

### 2021-08-16 figures/02_depth_model.Rmd
add supp figure showing proxy-proxy fit, closes #30

@@
 
+```{r 02_depth_model_fitted_allproxy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
+
+dt <- readRDS("../data/01_depth_model/depth_grid.rds") %>%
+  dplyr::filter(maxdepth == "maxdepth_false_false")
+
+rsq <- round(broom::glance(lm(.pred ~ lake_maxdepth_m, data = dt))$r.squared, 2)
+
+ggplot(data = dt, aes(x = lake_maxdepth_m, y = .pred)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  annotate("text", x = 42, y = 3,
+    label = paste0("R^2 == ", rsq),
+    parse = TRUE) +
+  annotate("label", x = 25, y = 3,
+    label = get_formula(
+      lm(.pred ~ lake_maxdepth_m, data = dt)),
+    size = 4, fill = "white", label.size = NA) +
+  ylab("lake_maxdepth_m (predicted)") +
+  # geom_abline() +
+  scale_x_continuous(expand = c(0, 0), limits = c(1.5, 50)) +
+  scale_y_continuous(expand = c(0, 0), limits = c(1.5, NA)) +
+  cowplot::theme_cowplot()
+```
+
 ```{r 02_depth_model_fitted, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}

### 2021-08-16 manuscript/si_template_2019.tex
add supp figure showing proxy-proxy fit, closes #30

@@
 
-            \item \textbf{Figure S9:} Comparison between reported depth and bathymetry depth
+            \item \textbf{Figure S9:} Proxy-proxy model fit showing predicted depth versus measured depth
+
+            \item \textbf{Figure S10:} Comparison between reported depth and bathymetry depth
 
@@
       \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product.
-The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS8}
+The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS7}
 \end{figure}
@@
       \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
-      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS9}
+      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS8}
+\end{figure}
+
+\clearpage
+\begin{figure}
+      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_fitted_allproxy-1}\end{center}
+      \caption{Proxy-proxy model fit showing predicted depth versus measured depth.}\label{figS9}
 \end{figure}

### 2021-08-16 figures/01_geometry.Rmd
show median calculations in fig s5, closes #31

@@
 
-```{r 01_slope_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6.2, fig.height=5.3}
+```{r 01_slope_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6.4, fig.height=5.3}
 # fig 2B (slope) alternatives
@@
 
-inlake_slope_alternatives     <- c("inlake_slope_pnt", "inlake_slope_pnts",
+inlake_slope_alternatives     <- c(
+  "inlake_slope_pnt",
+  "inlake_slope_pnts",
   "inlake_slope_mean",
   "inlake_slope_online_mean",
-  "inlake_slopes_online_mean")
-nearshore_slope_alternatives  <- c("nearshore_slope_mean",
+  "inlake_slopes_online_mean",
+  "inlake_slope_median"
+)
+nearshore_slope_alternatives  <- c(
+  "nearshore_slope_mean",
   "nearshore_slope_online_mean",
-  "nearshore_slopes_online_mean")
+  "nearshore_slopes_online_mean",
+  "nearshore_slope_median"
+)
 slope_alternatives   <- setNames(data.frame(expand.grid(
@@
     theme_opts +
-    theme(axis.title = element_text(family = "mono", size = 8))
+    # theme(axis.title = element_blank()) + ##
+    theme(axis.title = element_text(
+      family = "mono",
+      size = 8))
 
@@
 
-gg_slope_list[[5]] <- gg_slope_list[[5]] +
+gg_slope_list[[6]] <- gg_slope_list[[6]] +
   theme(axis.title.x = element_text(face = "bold"))
 
-gg_slope_list[c(1:5)] <- lapply(
-  gg_slope_list[c(1:5)], function(x) x +
+gg_slope_list[c(1:6)] <- lapply(
+  gg_slope_list[c(1:6)], function(x) x +
     theme(
@@
 
-gg_slope_list[c(1:4, 6:9, 11:14)] <- lapply(
-  gg_slope_list[c(1:4, 6:9, 11:14)], function(x) x +
+gg_slope_list[c(1:5, 7:11, 13:17, 19:23)] <- lapply(
+  gg_slope_list[c(1:5, 7:11, 13:17, 19:23)], function(x) x +
     theme(axis.title.x = element_blank(),
@@
 
-gg_slope_list[c(6:15)] <- lapply(
-  gg_slope_list[c(6:15)], function(x) x +
+gg_slope_list[c(7:24)] <- lapply(
+  gg_slope_list[c(7:24)], function(x) x +
     theme(axis.title.y = element_blank(),
       axis.text.y = element_blank()))
+
+gg_slope_inner <-
+  cowplot::plot_grid(plotlist = gg_slope_list,
+    nrow = 6, byrow = FALSE,
+    rel_heights = c(rep(0.75, 5), 1),
+    rel_widths = c(rep(1, 1),
+      rep(0.41, 3))
+  )
 
@@
   cowplot::plot_grid(
-    cowplot::plot_grid(plotlist = gg_slope_list,
-      nrow = 5, byrow = FALSE,
-      rel_heights = c(rep(0.75, 4), 1),
-      rel_widths = c(rep(1, 1),
-        rep(0.48, 2))
-    ),
+    gg_slope_inner,
     xlab, ncol = 1, rel_heights = c(1, 0.1)),

### 2021-08-16 scripts/06_stats.R
stash stats exploration

@@ -89,3 +89,29 @@ dt <- read.csv("data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>
 
 range(dt$lake_waterarea_ha)
 # 4 - 18500 ha
+
+# --- the percentage of lakes in Maine with a lake shape of neither ----
+dt <- read.csv("data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
+dplyr::filter(lake_state == "ME" & !is.na(shape_class))
+nrow(dplyr::filter(dt, shape_class == "neither")) / 
+nrow(dt) * 100
+
+# --- what are the natural lakes in Southern Iowa?----
+dt         <- readRDS("data/00_bathy_depth/bathy_pnts.rds") %>%
+dplyr::select(llid, pnt_deepest, state)
+hypso_pred <- read.csv("data/00_hypso/hypso_predictors.csv", 
+                       stringsAsFactors = FALSE) %>%
+  mutate(lagoslakeid = as.character(lagoslakeid))
+dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, shape_class), 
+                                  by = c("llid" = "lagoslakeid"))
+dt <- dplyr::filter(dt, state == "IA", reservoir_class == "NL")
+
+mapview::mapview(dt)
+
+
+
+# --- the percentage of glacial/non-glacial lakes ----
+
+
+# --- the percentage of lakes situated in different "landforms" ----
+

### 2021-08-22 manuscript/agujournaltemplate.tex
agu revision requires figures at the same folder level as the tex source
sync title with ms word

@@
 
-\title{Geometric models overestimate lake depth due to imperfect slope
-measurement}
+\title{Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth}
 
@@
 \begin{figure}
-\noindent\includegraphics[width=0.74\textwidth]{../figures/slope_diagram_new}
+      \noindent\includegraphics[width=0.74\textwidth]{slope_diagram_new}
       \caption{Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
@@
 \begin{figure}
-      \begin{center}\includegraphics[width=0.65\textwidth]{../figures/00_map_bathy-1}\end{center}
+      \begin{center}\includegraphics[width=0.65\textwidth]{00_map_bathy-1}\end{center}
       \caption{Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.}\label{fig2}
@@
       \begin{center}
-    \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/01_geometry_base-1}
+            \includegraphics[width=0.75\textwidth,keepaspectratio]{01_geometry_base-1}
       \end{center}
@@
             \hline
-  slope & distance & RMSE & $R^2$ & MAPE\\
+            slope & distance & RMSE  & $R^2$ & MAPE  \\
             \hline
-  true & true & - & - & -\\
-  true & proxy & 4.4 m & 0.70 & 29 \%\\
-  proxy & true & 6.6 m & 0.32 & 60 \%\\
-  proxy & proxy & 6.4 m & 0.35 & 59 \%\\
+            true  & true     & -     & -     & -     \\
+            true  & proxy    & 4.4 m & 0.70  & 29 \% \\
+            proxy & true     & 6.6 m & 0.32  & 60 \% \\
+            proxy & proxy    & 6.4 m & 0.35  & 59 \% \\
             \hline
@@
       \begin{center}
-    \includegraphics[width=\textwidth,keepaspectratio]{../figures/02_depth_model_grid_resid-1}
+            \includegraphics[width=\textwidth,keepaspectratio]{02_depth_model_grid_resid-1}
       \end{center}

### 2021-08-22 manuscript/si_template_2019.tex
title sync
sync title with ms word
add cfl affliation

@@
 
-\title{Supporting Information for "Imperfect slope measurements drive overestimation in standard cone model of lake depth"}
+\title{Supporting Information for "Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth"}
 %
@@
 
-\authors{J. Stachelek\affil{1}, P. J. Hanly\affil{1}, and P. A. Soranno\affil{1}}
+\authors{J. Stachelek\affil{1,2}, P. J. Hanly\affil{1}, and P. A. Soranno\affil{1}}
 
@@
       Wilson Rd., East Lansing, MI 48824, USA}
+\affiliation{2}{Center for Limnology, University of Wisconsin – Madison, Madison, WI, USA}
 

### 2021-11-10 README.Rmd
title sync
sync title with ms word
updates

@@
 
-[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Docker Build](https://img.shields.io/badge/Docker%20Image-jsta/bathymetry-green.svg)](https://cloud.docker.com/repository/docker/jsta/bathymetry)
+[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Data DOI](https://img.shields.io/badge/Data-DOI-blue.svg)](https://doi.org/) [![Docker Build](https://img.shields.io/badge/Docker%20Image-jsta/bathymetry-green.svg)](https://cloud.docker.com/repository/docker/jsta/bathymetry)
 
@@
 
-**Stachelek et al. In prep**. Imperfect slope measurements drive overestimation in standard cone model of lake depth
+**Stachelek J.**, P. J. Hanly, P. A., Soranno, Accepted, Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth. Inland Waters.
 
@@
 
-options(knitr.kable.NA = '-')
+options(knitr.kable.NA = "-")
 knitr::kable(dt)

### 2021-11-10 figures/01_geometry.Rmd
fix label overlap, add 1:1 line

@@
   theme(legend.position = "none",
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 11))
+    axis.text = element_text(size = 9),
+    axis.title = element_text(size = 10))
 text_size <- 3
@@
   geom_smooth(method = "lm", se = FALSE) +
-  annotate("text", x = 2550, y = 400, label = paste0("R^2 == ", rsq),
+  geom_abline(intercept = 0, slope = 1, linetype = 2, color = "darkgreen", size = 1) +
+  annotate("text", x = 2550, y = 430, label = paste0("R^2 == ", rsq), # higher y is up
     parse = TRUE, size = text_size - 0.7) +
-  annotate("label", x = 1800, y = 150,
+  annotate("label", x = 1900, y = 150, # 1800
     label = get_formula(
       lm(dist_viscenter ~ dist_deepest, data = st_drop_geometry(dt))),
-    size = text_size - 0.85, fill = "white", label.size = NA) +
+    size = text_size - 1.1, fill = "white", label.size = NA) +
   labs(color = "") +
@@
   ylab("Distance to \n lake center (m)") +
-  xlim(0, 3000) +
-  theme_opts
+  theme_opts +
+  scale_x_continuous(expand = c(0, 0.01), limits = c(0, 3000)) +
+  scale_y_continuous(expand = c(0, 1), limits = c(0, 3000))
 
@@
   geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt), method = "lm", se = FALSE) +
+  geom_abline(intercept = 0, slope = 1, linetype = 2, color = "darkgreen", size = 1) +
   geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
@@
     parse = TRUE, size = text_size - 0.7) +
-  annotate("label", x = 0.27, y = 0.01,
+  annotate("label", x = 0.22, y = 0.01,
     label = get_formula(
       lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)),
-    size = text_size - 0.85, fill = "white", label.size = NA) +
-  # geom_abline(slope = 1, intercept = 0) +
+    size = text_size - 1.2, fill = "white", label.size = NA, label.padding = unit(0.1, "lines")) +
   xlab("Nearshore slope (mean)") +
@@
   labs(color = "") +
-  theme_opts
+  theme_opts +
+  scale_x_continuous(expand = c(0, 0.01), limits = c(0, NA)) +
+  scale_y_continuous(expand = c(0, 0.01), limits = c(0, NA))
 

### 2021-11-10 manuscript/figures.Rmd
fix label overlap, add 1:1 line

@@
 ```{r 01_geometry_base, fig.cap = "Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
-A best-fit line and equation is shown to shown to facilitate computation of correction factors for proxy values of lake geometry.
+A best-fit line (solid) and equation is shown to shown to facilitate computation of correction factors for proxy values of lake geometry.
+A 1:1 line (dashed) is shown to illustrate bias.
 Coefficients of determination are shown to illustrate representativeness.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}

### 2021-11-10 manuscript/manuscript.Rmd
title sync
sync title with ms word
towards accepted version pdf
prep for accepted ms

@@
 ---
-title: Imperfect slope measurements drive overestimation in standard cone model of lake depth
+title: Imperfect slope measurements drive overestimation in geometric cone model of waterbody depth
 
@@
   - name: Jemma Stachelek
-    affiliation: a
+    affiliation: a,b
   - name: Patrick J. Hanly
@@
     address: Department of Fisheries and Wildlife, Michigan State University, 480 Wilson Rd., East Lansing, MI 48824, USA
+  - code: b
+    address: asdf
 
@@
 # If not specified, defaults to "This version was compiled on \today"
-#date_subtitle: Published in *Journal of Statistical Software*, 2018
-# date_subtitle: > 
-# This is the accepted version of the following article: J.
-Stachelek et al., Ecological Applications.
-In Press, which has been published in final form at DOI:  [DOI](https://doi.org).
+date_subtitle: > 
+  This is the accepted version of the following article: J.
+Stachelek et al., Inland Waters.
+In Press.
 This article may be used for non-commercial purposes in accordance with Wiley Terms and Conditions for Use of Self-Archived Versions.
@@
 First, we examined whether measures of lake geometry (in-lake slope and distance to the deepest point of lakes) were related to geometry proxies (nearshore land slope and distance to lake centers).
-Next, we computed geometric depth estimates (Equation \ref{eq1}) and offsets to these estimates using the random forest algorithm (Equation \ref{eq2}).
+Next, we computed geometric depth estimates (Equation \ref{eq1}) and offsets to these estimates using the random forest algorithm (Equation ).
 Covariates used in offset modeling included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table \ref{table1}).
@@
 its “visual distance to lake center”).
-For these calculations, we used the \texttt{polylabelr} R package \citep{larssonPolylabelrFindPole2019d}, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
+For these calculations, we used the \texttt{polylabelr} R package, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
 We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model (~15x15m grain) accessed using the \texttt{elevatr} R package \citep{hollisterElevatrAccessElevation2017} and the terrain function in the \texttt{raster} R package \citep{hijmansRasterGeographicData2019}.
@@
 \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
-Predictor variables for computing random forest offsets (Eq \ref{eq2}) are printed in bold face.
+Predictor variables for computing random forest offsets (Eq ) are printed in bold face.
 Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
@@
 
-Covariates used in random forest modeling (Table \ref{table1}, Equation \ref{eq2}) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
+Covariates used in random forest modeling (Table \ref{table1}, Equation ) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
 HUC4s), were obtained from the LAGOS-US LOCUS database (Smith et al.

### 2021-11-10 README.Rmd
render
fill out data badge

--- 2020-05-04 README.Rmd
@@
 
-[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Docker Build](https://img.shields.io/badge/Docker%20Image-jsta/bathymetry-green.svg)](https://cloud.docker.com/repository/docker/jsta/bathymetry)
+[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Data DOI](https://img.shields.io/badge/Data-10.6084/m9.figshare.12722246-blue.svg)](https://doi.org/10.6084/m9.figshare.12722246) [![Docker Build](https://img.shields.io/badge/Docker%20Image-jsta/bathymetry-green.svg)](https://cloud.docker.com/repository/docker/jsta/bathymetry)
 
@@
 
-**Stachelek et al. In prep**. Bathymetry data from thousands of lakes show that lake depth prediction is confounded by difficulty modeling inlake slope.
+**Stachelek J.**, P. J. Hanly, P. A., Soranno. Accepted. Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth. Inland Waters.
 
 ### Products
-
-Figures: [manuscript/combined.pdf](manuscript/combined.pdf)
 
@@
 ```
+
+### Reproducibility
+
+```shell
+# local
+make -t data/00_hypso/hypso.csv
+make -t data/00_bathy_depth/bathy_pnts.rds
+make -t data/00_bathy_depth/00_bathy_depth.csv
+make -t data/00_bathy_depth/bathy_geometry.csv
+make -t data/00_hypso/hypso_classes.csv
+make -t data/00_reservoir_classification/reservoir_classes_clean.csv
+make -t data/lagosus_depth_predictors.csv
+make -t data/lagosne_depth_predictors.csv
+make -t data/00_geometry/nearshore.csv
+make all
+```
+
+```shell
+# docker
+
+```

### 2021-11-10 manuscript/manuscript.Rmd
towards accepted ms pdf

@@
   - name: Jemma Stachelek
-    affiliation: a,b
+    affiliation: a,b,c
   - name: Patrick J. Hanly
-    affiliation: a
+    affiliation: c
   - name: Patricia A. Soranno
-    affiliation: a
+    affiliation: c
 
@@
   - code: a
+    address: Earth and Environmental Sciences Division, Los Alamos National Laboratory, Los Alamos, NM, USA
+  - code: b
+    address: Center for Limnology, University of Wisconsin – Madison, Madison, WI, USA
+  - code: c
     address: Department of Fisheries and Wildlife, Michigan State University, 480 Wilson Rd., East Lansing, MI 48824, USA
-  - code: b
-    address: asdf
 
@@
 abstract: |
-  Lake depth is a critical characteristic that influences many important ecological processes in lakes.
-Unfortunately, lake depth measurements are labor-intensive to gather and are only available for a small fraction of lakes globally.
-Therefore, scientists have tried to predict lake depth from characteristics that are easily obtained for all lakes such as lake surface area or the slope of the land surrounding a lake.
-One approach for predicting lake depth simulates lake basins using a geometric model where nearshore land slope is assumed to be a representative proxy for in-lake slope and the distance to the center of the lake is assumed to be a representative proxy for the distance to the deepest point of the lake.
-However, these assumptions have rarely been tested in a broad range of lakes.
-We used bathymetry data from 5,000 lakes and reservoirs to test these assumptions and to examine whether differences in lake type or shape influences depth prediction error.
-We found that nearshore land slope was not a representative proxy of in-lake slope and using it for prediction increases prediction error substantially relative to models using true in-lake slope.
-We also found that models using nearshore land slope as a proxy systematically overpredicts lake depth in concave lakes (i.e.
-bowl-shaped; up to 18% of lakes in the study population) and reservoir lakes (up to 30% of lakes in the study population), suggesting caution in using geometric models for lake prediction in unsampled lakes.
+Lake and reservoir (waterbody) depth is a critical characteristic that influences many important ecological processes.
+Unfortunately, waterbody depth measurements are labor-intensive to gather and are only available for a small fraction of waterbodies globally.
+Therefore, scientists have tried to predict waterbody depth from characteristics that are easily obtained for all waterbodies such as waterbody surface area or the slope of the land surrounding a waterbody.
+One approach for predicting waterbody depth simulates waterbody basins using a geometric cone model where nearshore land slope is assumed to be a representative proxy for in-lake slope and the distance to the center of the waterbody is assumed to be a representative proxy for the distance to the deepest point of the waterbody.
+However, these assumptions have rarely been tested in a broad range of waterbodies.
+We used bathymetry data from approximately 5,000 lakes and reservoirs to test the assumptions of the cone model and to examine whether differences in waterbody type or shape influences depth prediction error.
+We found that nearshore land slope was not representative of in-lake slope and using it for prediction increases error substantially relative to models using true in-lake slope.
+We also found that models using nearshore land slope as a proxy systematically overpredict depth in concave waterbodies (i.e.
+bowl-shaped; up to 18% of waterbodies in the study population) and reservoir waterbodies (up to 30% of waterbodies in the study population).
+Nevertheless, model errors were low (in absolute and relative terms) for concave waterbodies, suggesting that the cone model is an adequate representation of depth for these waterbodies.
 
@@
 acknowledgements: |
-  Funding was provided by the US NSF Macrosystems Biology Program grants, DEB-1638679; DEB-1638550, DEB-1638539, DEB-1638554.
-PAS was also supported by USDA National Institute of Food and Agriculture Hatch Project, Grant Number: 176820.
+This work was supported by the U.S.
+National Science Foundation (NSF) Macrosystems Biology Program (awards EF-1638679, EF-1638554, EF-1638539, and EF-1638550).
+JS was also supported by the NSF Harnessing the Data Revolution Program (OAC-1934633) and Los Alamos National Laboratory (LDRD-20210213ER).
+PAS was also supported by the USDA National Institute of Food and Agriculture, Hatch project 1013544.
 Author contributions: JS conceived of the study, built models, analyzed data, and wrote the paper.
-PH and PAS provided interpretation of results and edited the paper.
+PJH and PAS contributed to the conception of the manuscript, provided interpretation of results, and edited the paper.
+This work benefited from participation in the Global Lake Ecological Observatory Network (GLEON).
+We thank K.S.
+Cheruvelil for a friendly review of an earlier draft.
 
@@
   - lake depth
+  - cone model
+  - slope
   - bathymetry
   - hypsography
-  - geometry
-  - slope
+  - reservoir
 
@@
 
-Lake depth is a major factor controlling lake physics, chemistry, and biota
-\citep{liSignificantContributionLake2020, feeEffectsLakeSize1996, readSimulating2368Temperate2014}.
-However, because measured depth data is only available for a small fraction of all lakes, our ability to understand and predict depth-dependent processes is limited.
-The importance of lake depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all lakes such as lake surface area or the nearshore slope of the land surrounding a lake \citep{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
-The approaches taken in these studies have been varied.
-In addition, they have explored a variety of strategies to maximize prediction accuracy such as employing more diverse covariates \citep{oliver2016prediction}, varying lake buffer sizes \citep{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings among lakes \citep{sobekPredictingDepthVolume2011}.
+Depth is an important factor controlling waterbody physics, chemistry, and biota.
+For example, deeper waterbodies generally have higher water clarity and less complete mixing compared to shallow waterbodies (Fee et al.
+1996, Read et al.
+2014).
+These differences are reflected in variation among waterbodies in terms of biological productivity (Qin et al.
+2020) and rates of greenhouse gas production (Li et al.
+2020).
+However, because measured depth data is only available for a small fraction of waterbodies (about 15% of all waterbodies in the area encompassed by our study), our ability to understand and predict depth-dependent processes is limited.
+The importance of waterbody depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all waterbodies such as waterbody surface area or the nearshore slope of the land surrounding a waterbody (Heathcote et al.
+2015, Oliver et al.
+2016, Sobek et al.
+2011).
+Such efforts rely on a strategy of exploiting correlations between nearshore geomorphology and in-lake geometry, which at limited extents (e.g., within a single North American state or province) can be quite strong, while at larger extents can be dependent on geographic location and waterbody type (Oliver et al.
+2016, Branstrator 2009).
 
-One approach for predicting lake depth involves a geometric model that assumes lake basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \citep{hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
-All such geometric models for lake depth prediction involve implicit assumptions about the terms of geometric formulae.
-In the simplest case, where lakes basins are treated as cones (Eq \ref{eq1}, Figure \ref{fig1}), two assumptions are required to make depth predictions for all lakes: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the lake is a representative proxy for the distance to the deepest point of the lake (Figure \ref{fig1}).
+Given the limited prediction accuracy of prior depth prediction efforts (± 6-7 m, Heathcote et al.
+2015, Oliver et al.
+2016, Sobek et al.
+2011), a major focus has been on improving accuracy using strategies such as employing more diverse covariates (Oliver et al.
+2016), varying waterbody buffer sizes (Heathcote et al.
+2015), or estimating hidden groupings (e.g.
+fitting different models for distinct size classes) among waterbodies (Cael et al.
+2017, Sobek et al.
+2011).
+Unfortunately, the predictive accuracy of these efforts has been limited (± 6-7 m).
+
+One intuitive approach for predicting waterbody depth involves using a geometric model that assumes waterbody basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid (Getirana et al.
+2018, Hollister et al.
+2011, Neumann 1959, Yigzaw et al.
+2018).
+All such geometric models for waterbody depth prediction involve implicit assumptions about the terms of geometric formulae.
+In the simplest case, where waterbody basins are treated as cones (Eq.
+1, Fig.
+1), two assumptions are required to make depth predictions for all waterbodies: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the waterbody is a representative proxy for the distance to the deepest point of the waterbody (Fig.
+1).
 This cone model imposes the following fixed (i.e.
 geometric) relationship between slope and horizontal distance:
+
 \begin{align}
@@
 
-where the product of slope and horizontal distance yields an exact geometric depth estimate ($depth_{geometric}$).
+where the product of slope and horizontal distance yields an exact geometric depth estimate (depthgeometric).
+Cone models of waterbody basins have been used extensively to estimate hypsography in waterbodies where there is no knowledge of volume or mean depth (Read et al.
+2014, Winslow et al.
+2017).
 
-\begin{figure}[H]
-  \begin{center}
-    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/slope_diagram_new}
-  \end{center}
-  \caption{Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
-Geometric depth calculated via Equation \ref{eq1} requires a single distance and slope metric.}\label{fig1}
-\end{figure}
+The assumptions of the cone model (as well as other geometric models) can be tested by comparing proxy measures of waterbody geometry against corresponding “true” (i.e.
+in-lake) values derived from bathymetric maps and by evaluating how waterbody cross-section shapes differ from that of an idealized cone (Johansson et al.
+2007).
+For instance, waterbody cross-section shapes have been shown to vary from narrow "convex" forms to outstretched "concave" forms (Hakanson 1977).
+Because tests of geometric model assumptions require bathymetric map data, which is only available for a small fraction of waterbodies (including about 15% of all waterbodies in our study footprint), existing evidence may not be applicable to all waterbodies.
+The few studies that have tested these assumptions have been limited to individual studies of very large (> 500 ha) waterbodies or studies on small numbers (< 100) of waterbodies (Johansson et al.
+2007).
+Studies focused specifically on reservoirs (as opposed to the more typical case where reservoirs and natural lakes are combined), have been even more restricted to that of large waterbodies > 1000 ha (Lehner et al.
+2011, Messager et al.
+2016).
 
-The assumptions of this model as well as other geometric models can be tested by comparing proxy measures of lake geometry against corresponding “true” (i.e.
-in-lake) values derived from bathymetric maps.
-However, due to the scarcity of bathymetric map data, few studies have tested these assumptions except in individual studies of very large lakes or for small numbers of lakes (< 100) \citep{johanssonNewApproachesModelling2007}.
-As a result, it is unclear what effect proxies have on lake depth predictions using geometric models and whether depth predictions are more sensitive to errors in the horizontal dimension (i.e.
-distance to the deepest point of the lake) or errors in the vertical dimension (i.e.
+As a result of this limited testing, we lack knowledge on both the predictive performance of geometric models, the effect of proxies on depth prediction, and whether depth predictions are more sensitive to measurement errors in the horizontal dimension (i.e.
+distance to the deepest point of the waterbody) or measurement errors in the vertical dimension (i.e.
 in-lake slope).
-Additionally it is unclear whether model prediction error is related to differences in lake type such those with different cross-section shapes or those classified as reservoirs versus natural lakes.
-This information could help direct future research efforts to focus on particular dimensions of lake geometry (i.e.
-horizontal versus vertical) or to stratify model predictions based on specific lake types and cross-section shapes.
+Additionally, it is unclear whether model prediction error is related to differences in waterbody type such those with different cross-section shapes (concave versus convex) or those classified as reservoirs versus natural lakes.
+To address these knowledge gaps, we asked three research questions: (1) How representative is nearshore land slope of in-lake slope; and how representative is the distance to the center of a waterbody compared to the distance to the deepest point of a waterbody?
+(2) How does the use of proxies for waterbody geometry affect waterbody depth prediction error?
+(3) How does waterbody cross-section shape (i.e.
+concave versus convex) and waterbody type (i.e.
+natural lake vs reservoir) affect depth prediction error?
+To answer these questions, we extracted maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, cross-section shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 waterbodies from bathymetric map data.
+We supplemented these geometry measures with data classifying waterbodies as reservoirs or natural lakes.
+We used this data to compute geometric depth estimates (Equation 1) and prediction "offsets" to these estimates using the random forest algorithm (Equation 3).
+Such offsets are model quantities which minimize differences between observed and predicted depth as a function of covariates.
+In our case, covariates included a variety of waterbody, watershed, and hydrologic subbasin measures that are available for all waterbodies (Table 1).
 
-Given the knowledge gaps identified above, we asked three research questions: 1) Is nearshore land slope representative of in-lake slope; and is the distance to the center of a lake representative of the distance to the deepest point of a lake?
-2) How does the use of proxies for lake geometry affect lake depth prediction error?
-3) How does lake shape (i.e.
-concave versus convex) and lake type (i.e.
-natural lake vs reservoir) affect depth prediction error?
+By definition, the distance proxy (distance to the center of the waterbody) must always be greater or equal to the true distance value (distance to the deepest point of the waterbody).
+Therefore, we expect that the use of this proxy will lead to overestimation of waterbody depth (Fig.
+1).
+Furthermore, we expect to see greater overestimation error in reservoirs as compared to natural lakes because many reservoirs are known to be drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
+next to the dam) rather than in the center of the reservoir (Lanza and Silvey 1985).
+In a similar fashion, we expect to see overestimation error associated with using a nearshore land slope proxy in waterbodies with differing cross-section shape such that the depth of bowl-shaped (i.e.
+concave) waterbodies will be overpredicted whereas the depth of V-shaped (i.e.
+convex) waterbodies will be underpredicted (Fig.
+S1).
+Finally, we expect that depth predictions themselves will be strongly related to waterbody area and hydrologic subbasin variables as these measures have been influential in prior studies (Oliver et al.
+2016).
 
-To answer these questions, we extracted the maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 lakes in the Northeastern and Midwestern US from bathymetric map data.
-First, we examined whether measures of lake geometry (in-lake slope and distance to the deepest point of lakes) were related to geometry proxies (nearshore land slope and distance to lake centers).
-Next, we computed geometric depth estimates (Equation \ref{eq1}) and offsets to these estimates using the random forest algorithm (Equation ).
-Covariates used in offset modeling included a variety of lake, watershed, and hydrologic subbasin measures that are available for all lakes (Table \ref{table1}).
-We examined differences in overall prediction error corresponding to different inputs to geometric depth as well as differences in the relative distribution of prediction error in lakes with different characteristics.
-
-Given that, by definition, the distance proxy (distance to the center of the lake) must always be greater or equal to the true distance value (distance to the deepest point of the lake), we expect that the use of this proxy will lead to overestimation of lake depth (Figure \ref{fig1}).
-Furthermore we expect to see greater overestimation error in reservoirs as compared to natural lakes because many reservoirs are known to be drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
-next to the dam) rather than in the center of the reservoir \citep{lanza1985interactions}.
-Similar to distance proxies, we expect that the use of a nearshore land slope proxy will lead to prediction error according to differences in cross-section shape (Figure \ref{fig2}) such that the depth of U-shaped (i.e.
-concave) lakes will be overpredicted whereas the depth of V-shaped (i.e.
-convex) lakes will be underpredicted.
-Finally, we expect that offsets to geometric depth estimates will be most strongly related to lake area and hydrologic subbasin variables as these measures have been influential in prior studies \citep{oliver2016prediction}.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/lake_shape}
-  \end{center}
-  \caption{Diagram showing our expectation that slope-based models of lake depth will under predict true depth in convex lakes (left) and over predict true depth in concave lakes (right).
-Dashed lines represent extrapolated nearshore land slope while solid lines represent the lake bottom.}\label{fig2}
-\end{figure}
+By testing these expectations, we can establish whether barriers to increased depth prediction accuracy lie in lack of correspondence between true and proxy measures of waterbody geometry or in particular characteristics among waterbodies (such as waterbody cross-section shape or reservoir status).
+This information could help direct future research efforts to focus on particular dimensions of waterbody geometry (i.e.
+horizontal versus vertical) or to stratify model predictions based on specific waterbody types and cross-section shapes.
+Ultimately, achieving increased depth prediction accuracy would allow for more precise estimates of depth-dependent biotic and chemical processes across broad spatial extents.
 
@@
 ### Data description
-We compiled bathymetry data on approximately 5,000 lakes in the Northeastern and Midwestern US from nine official state databases (Figure S1).
+We compiled bathymetry data on approximately 5,000 waterbodies in the Northeastern and Midwestern United States from nine official state databases (Fig.
+2).
+These represent approximately 15% of all waterbodies in the States included in our study and are a diverse cross section in terms of their characteristics, surface areas, and span a wide geographic extent including glaciated and non-glaciated regions (Table 1).
+Thus, they can be considered representative of the entire population of lakes in our study extent.
+
 The original data came in a variety of formats including pre-interpolated rasters (Minnesota), contour lines (Nebraska, Michigan, Massachusetts, Kansas, Iowa), contour polygons (New Hampshire, Connecticut), or point depth soundings (Maine).
-For the Minnesota data, we simply clipped the raster for each lake to its outline.
-For data from the remaining states, we processed each lake by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the \texttt{raster} R package \citep{hijmansRasterGeographicData2019}.
+For the Minnesota data, we simply clipped the raster for each waterbody to its outline.
+For data from the remaining States, we processed each waterbody by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the raster R package (Hijmans 2019).
 The size of the moving window was adjusted iteratively to ensure that each bathymetry raster contained no missing data.
+All waterbody bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) National Hydrography Dataset (USGS 2019) waterbodies such that source data and bathymetry surface outputs were clipped to the area of each waterbody polygon.
+We restricted the waterbodies in our study to those with an area of at least 4 ha and a maximum depth of at least 0.3 m (1 ft.).
+The purpose of these restrictions was to ensure that waterbodies had enough contours (or points, or polygons) to generate adequately smooth interpolations with which to calculate in-lake geometry metrics.
 
-All lake bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) NHD \citep{usgsNationalHydrographyDataset2019} waterbodies such that source data and bathymetry surface outputs were clipped to the area of each lake polygon.
-We restricted the lakes in our study to those with an area of at least 4 ha and a maximum depth of at least 0.3 m (1 ft).
-The purpose of these restrictions was to ensure that lakes had enough contours (or points, or polygons) to generate adequately smooth interpolations with which to calculate in-lake geometry metrics.
+We used our generated bathymetry surfaces to find the location of the deepest point in the waterbody and we resolved ties by choosing the deepest point that was closest to the center of the waterbody.
+We used the location of this deepest point to calculate "distance to the deepest point" as the minimum distance to the waterbody shoreline.
+To account for waterbodies where the centroid does not intersect waterbody bathymetry because it is located within an embedded island or peninsula, we calculated the center of the waterbody not as its centroid but rather by finding the point farthest from the waterbody shoreline (i.e.
+its “visual distance to waterbody center”).
+For these calculations, we used the polylabelr R package (Larsson 2019), which interfaces with the Mapbox pole of inaccessibility algorithm (Agafonkin 2019).
+We calculated (maximum) in-lake slope as depth at the deepest point divided by the shortest distance to the deepest point from the shoreline..
+We calculated (mean) nearshore land slope for each waterbody by computing the slope within a 100-m buffer using data from a high-resolution digital elevation model (~ 15x15m grain) accessed using the elevatr R package (Hollister and Shah 2017).
+We explored alternative buffer sizes ranging between 50-1000 m following Sobek et al.
+(2011) and although 100 m provided the lowest model error, we ultimately found little appreciable effect of varying buffers on model performance.
+Slope computations proceeded by passing a 3x3 moving window over the 100-m buffer to calculate the slope at each point using Horn's algorithm via the terrain function in the raster R package (Hijmans 2019).
+Reported nearshore land slope values are the mean of all points in the buffer.
+In addition to the aforementioned techniques of calculating in-lake (and nearshore) slopes and distances, we tried 7 alternate techniques which are described in Fig.
+S5 and Table S2 including measures such as median slope (results not shown).
 
-We used our generated bathymetry surfaces to find the location of the deepest point in the lake and we resolved ties by choosing the deepest point that was closest to the center of the lake.
-We calculated the center of the lake not as its centroid but rather by finding the point farthest from the lake shoreline (i.e.
-its “visual distance to lake center”).
-For these calculations, we used the \texttt{polylabelr} R package, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
-We calculated in-lake slope as maximum lake depth divided by the distance to the deepest point and we calculated nearshore land slope for each lake by computing the slope within a 100-m buffer using data from a high resolution digital elevation model (~15x15m grain) accessed using the \texttt{elevatr} R package \citep{hollisterElevatrAccessElevation2017} and the terrain function in the \texttt{raster} R package \citep{hijmansRasterGeographicData2019}.
+We categorized waterbodies based on their cross-section shape and reservoir class (e.g.
+natural lake, reservoir).
+For cross-section shape, we categorized waterbodies as either convex or concave following the method of Hakanson (1977) by computing normalized waterbody depth-area relationships (i.e.
+hypsographic curves) and assigning class membership based on whether the midpoint of a waterbody’s curve falls above or below that of a simple straight-sided cone (Fig.
+S2).
 
-We categorized lakes based on their cross-section shape and reservoir class.
-For cross-section shape, we categorized lakes as either convex or concave following the method of \citet{hakansonLakeFormLake1977} by computing normalized lake depth-area relationships (i.e.
-hypsographic curves) and assigning class membership based on whether a lake’s curve falls above or below that of a simple straight-sided cone (Figure S4).
-We further classified lakes using data from \citet{polus2020}, which uses the output of a machine learning algorithm to assign a probability to each lake regarding whether it is a reservoir or a natural lake.
-For our purposes, we determined a lake to be a reservoir if the classification probability was 0.75 or greater.
-The \citet{polus2020} data product defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water, beyond simply controlling water level.
-It makes no distinction between different dam types or dam heights.
+We further categorized waterbodies using the output of a deep convolutional neural network model trained on satellite images labelled according to whether there was visual evidence of a water control structure significantly impacting flow (Polus et.
+al 2021).
+This model had an overall validation accuracy of 81% and produced a probability for each waterbody as to whether it is a reservoir or a natural lake.
+For our purposes, we set a conservative classification probability threshold of 0.75 to determine whether a waterbody would be considered a reservoir.
+For example, if the Polus et al.
+(2021) dataset classified a particular waterbody as a reservoir with a probability of 0.74 we categorized it as a natural lake but if the probability was greater than 0.75 we categorized it as a reservoir.
+Note that our reservoir classification defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water.
+It makes no distinction between different dam types, heights, or uses/purposes because the Polus et al.
+(2021) dataset is only based on visual interpretation of waterbody images (via deep convolutional neural network models).
+However, the Polus et al.
+(2021) dataset is unique in that it provides data using a standardized approach at broad spatial extents for waterbodies > 4 ha.
 
-\begin{table}
-\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>).
-Predictor variables for computing random forest offsets (Eq ) are printed in bold face.
-Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.} \label{table1}
-\centering
-\setlength\tabcolsep{1.5pt} % default value: 6pt
-\begin{tabular}{lllll}
-  \hline
-  Variable & Median & Q25 & Q75 & n\\
-  \hline
-  Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4850 (17700)\\
-  \textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4850 (17700)\\
-  \textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4850 (17700)\\
-  \textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4850 (17700)\\
-  \textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4850 (17700)\\
-  \textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4850 (17700)\\
-  \textbf{Watershed-lake ratio} & 7.8 (10) & 3.9 (4.4) & 17 (28) & 4850 (17700)\\
-  Distance to deepest point (m) & 180 (-) & 110 (-) & 290 (-) & 4850 (-)\\
-  Distance to lake center (m) & 240 (-) & 160 (-) & 380 (-) & 4850 (-)\\
-  In-lake slope (m/m) & 0.046 (-) & 0.024 (-) & 0.079 (-) & 4850 (-)\\
-  Nearshore land slope (m/m) & 0.077 (-) & 0.051 (-) & 0.11 (-) & 4850 (-)\\  
-  \bottomrule
-  \end{tabular}  
-\end{table}
-
-Covariates used in random forest modeling (Table \ref{table1}, Equation ) for lake elevation, area, island area, perimeter, shoreline development, watershed to lake area ratio, and hydrologic subbasin (i.e.
-HUC4s), were obtained from the LAGOS-US LOCUS database (Smith et al.
-In Prep).
-Shoreline development is a measure of lake perimeter shape defined as:
+Covariates used in random forest modeling (Table 1, Equation 3, see Random Forests Model sub-section below) for waterbody elevation, area, island area, perimeter, shoreline development, watershed to waterbody area ratio, and hydrologic subbasin (i.e.
+HUC4s), were obtained from the LAGOS-US LOCUS database.
+One such measure, that of shoreline development, is a measure of waterbody perimeter shape defined as:
 
 \begin{align}
-shoreline_{devel} = perimeter / (2 * \sqrt{(\pi * waterarea * 10000)})       
+shoreline_{devel} = perimeter / (2 * \sqrt{(\pi * waterarea * 10000)}) \label{eq2}
 \end{align}
 
-where sinuous lakes have larger values of shoreline development and circular lakes have smaller values of shoreline development.
-Watershed to lake area ratio is an approximation of water residence time and is defined as watershed area divided by lake area.
-All of our data and data processing code are available at [Zenodo DOI].
+where sinuous waterbodies have larger values of shoreline development and circular waterbodies have smaller values of shoreline development.
+Watershed to waterbody area ratio is an approximation of water residence time and is defined as watershed area divided by waterbody area (Timms 2009).
 
 ### Proxy evaluation
-We conducted a qualitative assessment of whether or not proxy measures of lake geometry are representative of their true values by visual inspection (i.e.
-plotting each proxy measure against its corresponding true value) and by computing coefficients of determination ($R^2$).
-We further tested proxy measures by examining their effect on lake depth prediction error.
+We conducted a qualitative assessment of whether or not proxy measures of waterbody geometry (e.g.
+nearshore land slope, distance to the center of the waterbody) are representative of their true values (e.g.
+in-lake slope, distance to the deepest point of the waterbody) by visual inspection (i.e.
+plotting each proxy measure against its corresponding true value) and by computing coefficients of determination (R2).
+We further tested proxy measures by examining their effect on waterbody depth prediction error.
 Our approach involved several steps.
-In the first step, we computed a geometric estimate of lake depth using only geometry information ($depth_{geometric}$, Equation \ref{eq1}).
+In the first step, we computed a geometric estimate of waterbody depth using only geometry information (depthgeometric, Equation 1).
 In the second step, we fit a random forest model to predict observed (i.e.
-true) depth as a function of geometric depth along with several covariates available for all lakes (Table \ref{table1}).
-The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of $depth_{geometric}$ and among different lake types.
-Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table \ref{table2}).
+true) depth as a function of geometric depth along with several covariates available for all waterbodies (Table 1).
+The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of depthgeometric and among different waterbody types.
+Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table 2).
+We conducted additional sensitivity analysis to examine possible interactions between different proxy measures of waterbody geometry and different subsets of the entire dataset where model data was restricted (i.e.
+“filtered”) to include only reservoirs, only natural lakes, only convex waterbodies, or only concave waterbodies (Table 2).
 
@@
 ### Geometric model 
-We used a geometric model of lakes where basins are treated as cones with a fixed relationship between slope and distance (Equation \ref{eq1}).
-Note that Equation \ref{eq1} is a geometric formula and has no intercept or “coefficients” and it produces a perfect depth estimate given true values of slope and distance.
-To use this model to predict the depth of all lakes, there is a necessary assumption that proxy slope and distance measures, which are available for all lakes, are representative of true slope and distance (Figure \ref{fig1}).
+We used a geometric model of waterbodies where basins are treated as cones with a fixed relationship between slope and distance (Equation 1).
+One reason that we used the cone model is that, unlike other idealized shapes, it does not require any knowledge of waterbody volume or mean depth.
+Note that Equation 1 is a geometric formula and has no intercept or coefficients and it produces an exact depth value given true values of slope and distance.
+To use this model to predict the depth of all waterbodies, there is a necessary assumption that proxy slope and distance measures, which are available for all waterbodies, are representative of true slope and distance (Fig.
+1).
 
 ### Random forest models
-Nearly all prior studies predicting lake depth using geometric models include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy.
-For our purposes, this offset modeling enabled us to test our expectations that prediction error would be different among different formulations of $depth_{geometric}$ and among different lake types.
-It also facilitated direct comparison against prior models of lake depth including those that are non-geometric.
-We generated an “offset” to geometric depth \citep[sensu][]{hollisterPredictingMaximumLake2011} using the random forest algorithm and the \texttt{ranger} R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation \ref{eq1}) along with the lake elevation, area, perimeter, and ratio/index measures listed in Table 1:
+Prior studies using geometric models to predict waterbody depth include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy (Hollister et al.
+2011, Yigzaw et al.
+2018).
+This procedure involves fitting a statistical or machine learning model to the residuals of an initial geometric model.
+For our purposes, such offset modeling enabled us to test our expectations that prediction error would be different among different formulations of depthgeometric and among different waterbody types.
+It also facilitated direct comparison against prior models of waterbody depth including those that are non-geometric.
+We generated an “offset” to geometric depth (sensu Hollister et al.
+2011) using the random forest algorithm and the ranger R package (Wright and Ziegler 2017) to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation 1) along with the waterbody elevation, area, perimeter, and ratio/index measures listed in Table 2:
 
 \begin{align}
-  depth_{observed} \sim depth_{geometric} + covariates         
+  depth_{observed} \sim depth_{geometric} + covariates \label{eq3}  
 \end{align}
 
+We evaluated the relative “importance” of individual covariates by comparing model performance between 1) models where a given covariate was left untouched versus 2) models where a given covariate was permuted randomly (Prasad et al.
+2006, Wright and Ziegler 2017).
 Neither cross-section shape nor reservoir class was used as a covariate in any random forest models.
+Random forest training and test data was stratified on shape and reservoir class to match that of the overall waterbody population.
 We used the random forest algorithm because it makes no assumptions about the distribution of model residuals, allows for non-linearity, and is insensitive to interactions (i.e.
-multicollinearity) among covariates \citep{prasadNewerClassificationRegression2006}.
+multicollinearity) among covariates (Prasad et al.
+2006).
 
 ### Model comparisons
-We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
-Prior to entry into Equation \ref{eq1}, we standardized proxy distances to have the same numeric range as their true counterpart.
-The purpose of this standardization was to prevent lakes with extremely long proxy distances from having an outsized impact on model evaluation metrics.
+We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation 1 (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
+Prior to entry into Equation 1, we standardized proxy distances to have the same numeric range as their true counterpart.
+The purpose of this standardization was to prevent waterbodies with extremely long proxy distances from having an outsized impact on model evaluation metrics.
+In addition to comparing among model runs using different metric combinations, we compared among sets of model runs where slope and distance measures were calculated using different sets of calculation techniques (Table S2).
 
 ### Model evaluations
-We evaluated model fit and prediction error using root-mean-square error (RMSE) and coefficient of determination ($R^2$) metrics on a holdout set containing 25\% of all lakes.
-We evaluated the residuals (residual = observed - predicted) of each model relative to lake cross-section shape and reservoir classes to determine whether depth is consistently over or under predicted for some lake types relative to others.
-All of our code for model fitting and evaluation is available at [Zenodo DOI].
+We evaluated model fit and prediction error using root-mean-square error (RMSE), mean absolute percent error (MAPE), and coefficient of determination (R2) metrics on a holdout set (i.e.
+a data subset not used for model training) containing 25% of all waterbodies.
+We evaluated the residuals of each model relative to waterbody cross-section shape and reservoir classes to determine whether depth is consistently over or under predicted for some waterbody types relative to others.
 
 ## RESULTS
-Lakes belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Figure S1).
-For example, concave lakes were nearly absent from Michigan whereas Maine lakes had an overabundance of lakes categorized as neither concave nor convex.
-Lakes in the southern portions of our study area tended to be classified as reservoirs whereas lakes in the northern portions of our study area were a more even mix between reservoirs and natural lakes (Figure S1).
-Approximately 18\%, 80\%, and 2\% of lakes were classified as having a concave, convex, or neither shape respectively whereas approximately 30\% and 70\% of lakes were classified as being a reservoir or a natural lake.
+Waterbodies belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Fig.
+2).
+For example, concave waterbodies were nearly absent from Michigan whereas Maine had an overabundance (~3%) of waterbodies categorized as neither concave nor convex.
+Waterbodies in the southern portions of our study area tended to be classified as reservoirs whereas waterbodies in the northern portions of our study area were a more even mix between reservoirs and natural lakes (Fig.
+2).
+Approximately 18%, 80%, and 2% of waterbodies were classified as having a concave, convex, or neither shape respectively whereas approximately 30% and 70% of waterbodies were classified as being a reservoir or a natural lake.
 
-Although proxy distance to lake center was often much larger in magnitude compared to the true distance to the deepest point of lakes’, they were strongly related ($R^2$ = 0.8).
-In contrast, proxy nearshore land slope and true in-lake slope were more weakly related ($R^2$ = 0.17).
-For slope measures, most lakes had lower magnitude (i.e.
-shallower) nearshore land slope compared to true in-lake slope (Figure \ref{fig3}).
-Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes while proxy nearshore land slope is not representative of true in-lake slope.
-In addition to overall differences between slope and distance measures, we found differences in these relationships among lake shape classes.
-For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S5).
-However,  there were not similar differences among slope and distance metrics for natural lakes versus reservoirs (Figure S5).
+Although proxy distance to waterbody center was often larger in magnitude compared to the true distance to the deepest point of waterbodies’ (rather than being identical), they were strongly related (R2 = 0.8).
+Note that the coefficient of determination for this relationship is not strictly correct given that distance to waterbody center is an upper bound on distance to the deepest point of waterbodies.
+In contrast to distance metrics, proxy nearshore land slope and true in-lake slope were more weakly related (R2 = 0.17).
+For slope measures, most waterbodies had higher magnitude (i.e.
+steeper) nearshore land slope compared to true in-lake slope (Fig.
+3).
+Taken together, these results suggest that proxy distance to the center of waterbodies is representative of true distance to the deepest point of waterbodies whereas proxy nearshore land slope is not representative of true in-lake slope.
+The strong relationship between distance to the center of waterbodies and distance to the deepest point means that it is possible to convert between the two measures in subsequent analyses (See best-fit equations in Fig.
+3).
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/01_geometry_base-1}
-  \end{center}
-  \caption{Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
-A best-fit line and coefficient of determination is shown to illustrate representativeness.}\label{fig3}
-\end{figure}
+In addition to overall differences between slope and distance measures, we found differences in these relationships among waterbody shape classes.
+For example, in-lake slope and distance to the deepest point of the waterbody metrics were consistently larger in magnitude for convex waterbodies as compared to concave waterbodies (Fig.
+S3).
+We found evidence that this difference was at least partly explained by the fact that convex waterbodies are deeper than concave waterbodies (Fig.
+S8).
+Unlike concave and convex waterbodies, there were not clear differences among slope and distance metrics for natural lakes versus reservoirs.
+Offset model fit and prediction error differed depending on the technique used to calculate in-lake and nearshore geometry metrics (Table S2).
+We found that the best model fit and lowest model error occurred when in-lake slope was calculated as the average point-wise slope of all points at maximum waterbody depth rather than at single point of maximum depth.
+However, given the small difference in the fit of models using either of these techniques and the significant cost in terms of computational load and complexity, we limit our discussion hereafter to the simpler case involving only a single deepest point.
 
-The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table \ref{table2}).
-More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model ($R^2$ = 0.26, RMSE = 6.87m).
-Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure \ref{fig4}).
-We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure \ref{fig4}).
+The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to waterbody center (Table 2).
+More specifically, the true slope (in-lake slope) - proxy distance (distance to the center of the waterbody) model had a better fit (R2 = 0.73) and lower prediction error (RMSE = 4.8m, MAPE = 27%) compared to the proxy slope - true distance model (R2 = 0.31, RMSE = 7.3m, MAPE = 64%).
+The fit of the proxy slope -proxy distance model (R2 = 0.36, RMSE = 7.1m, MAPE = 61%) was very similar to the proxy slope – true distance model.
+Predicted depth values for this model were generally underestimates relative to measured depth values (Fig.
+S9).
 
-\begin{table}
-  \caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.} \label{table2}
-  \centering
-  % \setlength\tabcolsep{1.5pt} % default value: 6pt
-\begin{tabular}{llll}
-  \hline
-  slope & distance & RMSE & $R^2$\\
-  \hline
-  true & true & - & -\\
-  true & proxy & 4.2 m & 0.73\\
-  proxy & true & 6.9 m & 0.26\\
-  proxy & proxy & 6.6 m & 0.31\\
-  \bottomrule
-\end{tabular}
-\end{table}
+Furthermore, analysis of model residuals showed overestimation of waterbody depth for concave waterbodies when models included a proxy slope measure (Fig.
+4).
+We observed similar but smaller overestimation depending on if a waterbody was classified as a reservoir rather than a natural lake (Fig.
+4).
+We found that models restricted to consider only concave lakes had lower error (both in absolute and relative terms) compared to models on other data subsets (e.g.
+convex lakes, reservoirs, natural lakes, see Table 2).
+Conversely, we did not observe any notable geographic patterns in model residuals (Fig.
+S6).
 
-\begin{figure*}
-  \begin{center}
-    \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/02_depth_model_grid_resid-1}
-  \end{center}
-  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig4}
-\end{figure*}
-
-The most important covariates in offset models were those relating to spatial location, lake area, and perimeter (Figure S6).
-Conversely, watershed metrics and lake elevation had little contribution to random forest model fit (Figure S6).
+The most important covariates in these models were those relating to spatial location, waterbody area, and perimeter (Fig.
+5).
+Conversely, watershed metrics and waterbody elevation had little contribution to random forest model fit.
 The spatial location (i.e.
-HUC4) covariate was notably less importance in the true slope model compared to the two proxy slope models.
-Model importance calculations indicated that omitting a geometric max depth measure results in a 130\%, 60\%, or 50\% increase in mean square error depending on the formulation of geometric max depth in Eq \ref{eq1} (Figure S6).
+HUC4, hydrologic subbasin) covariate was notably less important in the true slope model compared to the two proxy slope models.
+To evaluate the contribution of our “offset” models relative to the “base” geometric model, we can look at model importance calculations for the geometric max depth input to the random forest model (Fig.
+5).
+These calculations indicate that omitting a geometric max depth term results in a 130%, 60%, or 50% increase in mean square error depending on the formulation of geometric max depth in Eq.
+1.
 
@@
 
-Our tests of geometric lake depth models show that specific proxy measures of lake geometry are not representative of true measures of lake geometry across a broad array of lakes.
-Using a cone model example, we show that nearshore land slope is not representative of in-lake slope.
-Furthermore, our results indicate that use of this metric for prediction results in increased prediction error and systematic overestimation of depth in concave and reservoir lakes.
+Our tests of the geometric cone model of waterbody depth models show that specific proxy measures of waterbody geometry are not representative of true geometry measures across a broad array of waterbodies.
+Models using non-representative proxies showed increased error and systematic overestimation of depth in concave and reservoir waterbodies.
+Although our analysis was limited to waterbodies with available bathymetry data, these waterbodies did not have characteristics that differed from that of the overall waterbody population (apart from the fact that our study waterbodies were somewhat larger in area compared to the overall waterbody population, See Fig.
+S7, Table 1, Fig.
+S10).
+Although there is a possibility that there is some hidden bias not explored for in our analyses, this lack of difference suggests that our results are likely to be broadly applicable to nearly all waterbodies in the study area.
 
-### Representativeness of proxy measures of lake geometry
-In comparing among lake geometry measures, our analysis suggests that proxy distance to lake center is representative of true distance to the deepest point of the lakes but that proxy nearshore land slope is not representative of true in-lake slope.
-A simple indication of this non representativeness is that proxy nearshore land slope was often (in > 74\% of cases) steeper than true in-lake slope.
-This finding is consistent with the suggestion by \citet{heathcotePredictingBathymetricFeatures2015} that in-lake slope matches nearshore land slope at the time of lake formation, but as lakes age, infilling via sedimentation causes them to develop shallower in-lake slopes compared to the surrounding land.
-If true, then these results also lend support to prior work \citep{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}, conducted on a more limited number of lakes, showing that in-lake slopes are controlled by erosion and sediment transport processes.
+### Representativeness of proxy measures of waterbody geometry
+In comparing among waterbody geometry measures, our analysis suggests that proxy distance to waterbody center is representative of true distance to the deepest point of the waterbodies, but that proxy nearshore land slope is not representative of true in-lake slope.
+A simple indication of this non-representativeness is that proxy nearshore land slope was often (in > 74% of cases) steeper than true in-lake slope.
+This finding is consistent with Heathcote et al.
+(2015) whose results suggest that in-lake slopes are shallower compared to the surrounding land.
+Furthermore, the fact that in-lake slopes were shallower compared to the surrounding land even after controlling for differences in area (Fig.
+S4) is consistent with the idea of topographic scaling (i.e.
+scale invariance) explored in previous work and detailed by (Cael et al.
+2017).
+The underlying reason for these shallow in-lake slopes may be related to slope-induced turbidity currents which distribute sediment from shallow high-energy areas of waterbodies to deep low-energy areas (Hakanson 1981, Johansson et al.
+2007).
+The strength of such sediment focusing is likely greater in "younger" waterbodies with steeper slopes leading to a smoothing of their bathymetry over time (Blais and Kalff 1995).
 
-One surprising finding with respect to the relationship between true and proxy geometry measures when examined by lake class was the fact that there was no greater difference between proxy and true distances in reservoirs compared to natural lakes.
+One surprising finding with respect to the relationship between true and proxy geometry measures when examined by waterbody class was the fact that there was no greater difference between proxy and true distances in reservoirs compared to natural lakes.
 This is contrary to the idea that most reservoirs are drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
-next to the dam) rather than in the center of the reservoir \citep{lanza1985interactions}.
-One possible explanation is that the reservoir classification data from \citet{polus2020} uses a more general definition of a reservoir (i.e.
+next to the dam) rather than in the center of the reservoir (Lanza and Silvey 1985).
+One possible explanation is that our reservoir classification data uses a more general definition of a reservoir (i.e.
 any permanent waterbody that has a water control structure likely to significantly impact flow or pool water) compared to that of conventional classifications that are tied to specific dam types or dam heights.
-Another possible explanation is that conventional reservoir classifications are conceptually biased towards more southern areas with few natural lakes (Figure S1).
+Another possible explanation is that conventional reservoir classifications are conceptually biased towards more southern areas with few natural lakes (Figure 2).
+Southern Iowa, for instance is typically considered to have few to no natural lakes.
+In the present study, all of the apparent natural lakes in Southern Iowa were in fact oxbow lakes adjacent to the Missouri River.
 
-We found other differences between convex and concave lakes.
-One finding was that convex lakes, when compared to concave lakes, had longer distances to lake centers relative to corresponding distances to the deepest point of lakes.
-In addition, convex lakes often had steeper in-lake slopes relative to nearshore land slopes as compared to concave lakes.
-These differences are reflected in the class-wise differences in shoreline development (e.g.
-shorelines were more sinuous in convex lakes compared to concave lakes, Figure S3).
-It was notable that convex lakes were deeper than concave lakes despite having similar distributions of lake surface area (Figure S3).
-Given the similarity in lake surface area, the underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of lake formation whereby concave lakes were formed as a result of more intense glacial scouring compared to convex lakes \citep{gorhamPhysicalLimnologyNorthern1958}.
-While there is some evidence in support of this idea, namely that there is a geographic hotspot of concave lakes associated with the glaciated “prairie pothole region” \citep[see][]{hayashiSimpleEquationsRepresent2000}, the overall geographic distribution of lake cross-section shapes does not support this idea.
-Instead of a concentrated area of concave lakes in formerly glaciated regions, there appears to be a fairly even mix of concave and convex lakes distributed amongst the northern (i.e.
-glaciated) and southern (non-glaciated) portions of our study area (Figure S1).
+We found other differences among waterbody geometry measures according to waterbody cross-section shape.
+One finding was that convex waterbodies, when compared to concave waterbodies, had longer distances to waterbody centers relative to corresponding distances to the deepest point of waterbodies.
+In addition, convex waterbodies often had steeper in-lake slopes relative to nearshore land slopes as compared to concave waterbodies.
+Finally, it was notable that convex waterbodies were deeper than concave waterbodies despite having similar distributions of waterbody surface area (Fig.
+S8).
+The underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of waterbody formation whereby the formation of concave waterbodies were a result of more intense glacial scouring compared to that of convex waterbodies (Gorham 1958).
+While our findings provide some evidence in support of this idea, namely that there is a geographic hotspot of concave waterbodies associated with the glaciated “prairie pothole region” (See Hayashi and van der Kamp 2000), the overall geographic distribution of waterbody cross-section shapes does not support this idea.
+Instead of a concentrated area of concave waterbodies in formerly glaciated regions, there appears to be a fairly even mix of concave and convex waterbodies distributed amongst the northern (i.e.
+glaciated) and southern (non-glaciated) portions of our study area (Fig.
+2).
 
-### Effects of proxy measures of lake geometry depth prediction error
-Models using only proxy variables had prediction error rates (RMSE = 6.6m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting lake depth at broad geographic extents \citep{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
-When only a single proxy measure was used there was a difference in model sensitivity depending on if it was a horizontal distance measure or a vertical slope measure.
-In the case of a true slope and proxy distance combination, models were more accurate ($\pm$ 4.2m) than even the most accurate of prior studies \citep{hollisterPredictingMaximumLake2011, oliver2016prediction, messagerEstimatingVolumeAge2016}.
-Conversely, models using a proxy slope and true distance combination had prediction error rates ($\pm$ 6.9m) of a similar magnitude as that of the baseline proxy-proxy model ($\pm$ 6.6m).
+### Effects of proxy measures of waterbody geometry depth prediction error
+Models using only proxy variables (Table 2) had prediction error rates (RMSE = 7.1m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting waterbody depth at broad geographic extents (Hollister et al.
+2011, Oliver et al.
+2016, Messager et al.
+2016).
+When only a single proxy measure was used, there was a difference in model sensitivity depending on if it was a horizontal distance measure or a vertical slope measure.
+In the case of a true slope and proxy distance combination, models were more accurate (± 4.8m, 27%) than even the most accurate of prior studies (Hollister et al.
+2011, Oliver et al.
+2016, Messager et al.
+2016).
+Conversely, models using a proxy slope and true distance combination had prediction error rates (± 7.3m, 64%) of a similar magnitude as that of the baseline proxy-proxy model (± 7.1m, 61%).
 The greater sensitivity of depth predictions to proxy slope measures relative to proxy distance measures may be explained by the fact that proxy slope measures were a more imperfect representation of true in-lake slopes relative to proxy versus true distances.
-In addition, these results help explain the relatively poor predictive performance of prior non-geometric lake depth models given that they rely heavily on lake area as a predictor \citep{messagerEstimatingVolumeAge2016, oliver2016prediction, sobekPredictingDepthVolume2011} and both horizontal distance measures and vertical slope measures appear to be decoupled from lake area (Figure S3).
+We did not find evidence that the sensitivity of depth predictions to slope was dependent on variations in how these measures were calculated (Table S2).
+In a general sense, the sensitivity of depth predictions to slope help explain the relatively poor predictive performance of prior non-geometric waterbody depth models given that they rely heavily on waterbody area as a predictor (Messager et al.
+2016, Oliver et al.
+2016, Sobek et al.
+2011) and both horizontal distance measures and vertical slope measures appear to be decoupled from waterbody area (Fig.
+S8).
 
-### Effects of lake shape and lake type on depth prediction error
-As expected, we found that the maximum depth of concave lakes was systematically overpredicted by a simple geometric model using proxy nearshore land slope.
-However, contrary to our expectation, we did not observe underprediction of depth in convex lakes.
-The reason we did not observe underprediction of the depth of convex lakes is likely because geometric depth itself was always greater than observed maximum depth owing to the fact that proxy distance is constrained to be greater than true distance.
-Although the magnitude of overestimation is likely related to class imbalance in our dataset (i.e.
-there was a greater number of convex lakes), these results suggest that broad scale estimates of lake depth are overestimated particularly when those estimates encompass large numbers of lakes with diverse cross-section shapes.
+### Effects of waterbody shape and waterbody type on depth prediction error
+As expected, we found that the maximum depth of concave waterbodies was systematically overpredicted by a simple geometric model using proxy nearshore land slope (Fig.
+S1).
+However, contrary to our expectation, we did not observe underprediction of depth in convex waterbodies.
+The reason we did not observe underprediction of the depth of convex waterbodies is likely because geometric depth itself was always greater than observed maximum depth owing to the fact that proxy distance is constrained to be greater than true distance.
+Given that models restricted to only concave waterbodies had low error (both in absolute and relative terms), this suggests that despite evidence of overprediction, the cone model is an adequate representation of depth for these waterbodies.
 
 ### Future research
-One of our models (true slope, proxy distance) was more accurate than even the most accurate of prior studies.
-However, parameterization of this model requires data on bathymetry which is not available for all lakes.
-We propose that the error rate of this model ($\pm$ 4.2m) be used as an out-of-sample prediction benchmark for future studies such that they should attempt to match it but not expect to exceed it.
-Because this model requires bathymetry data, this suggests that it may not be possible with current data and models to produce depth predictions for all lakes with error rates below about 6m.
-To approach our benchmark using data available for all lakes, future studies could explore alternative modeling approaches such as ordinal modeling, which would capture whether or not a lake crosses some important depth threshold, but would not seek to predict a specific depth value.
-These studies could also explore emerging data types to boost prediction accuracy such as “topobathymetric” products that integrate both topographic and bathymetric data in a seamless fashion rather than treating them as separate entities.
-This would allow for more robust tests of the representativeness of geometric model inputs.
-Unfortunately, topobathymetric products are rare, have mostly been limited nearshore marine environments, and as such are not widely available for inland waters \citep{danielsonTopobathymetricElevationModel2016}.
+The only model parameterization that was more accurate than the most accurate of prior studies fitting models to waterbody depth data requires data on in-lake slope (true slope, proxy distance) which is not available for all waterbodies thus it is not of practical use for general prediction.
+However, we propose that the error rate of this model (± 4.8m, 27%) be used as an out-of-sample prediction benchmark for future studies such that they should attempt to match it but not expect to exceed it.
 
-Finally, our findings indicate that geometry measures differ according to lake cross-section shape.
+Because this most accurate model requires bathymetry data, this suggests that it may not be possible with current data and models to produce depth predictions for all waterbodies with error rates below about 5m or 30%.
+To achieve high prediction accuracy using data available for all waterbodies, future studies could explore alternative modeling approaches such as ordinal modeling, which would capture whether or not a waterbody crosses some important depth threshold but would not seek to predict a specific depth value, or emerging data types such as “topobathymetric” products that integrate both topographic and bathymetric data in a seamless fashion rather than treating them as separate entities.
+Topobathymetry would allow for more robust tests of the representativeness of geometric model inputs.
+Unfortunately, topobathymetric products are rare, have mostly been limited nearshore marine environments, and as such are not yet widely available for inland waters (Danielson et al.
+2016).
+Other potential explanatory data include information on waterbody origin and development.
+Unfortunately, such “waterbody ontogeny” data is presently available only for select regions and the very largest waterbodies (Sharma and Byrne 2011).
+
+Finally, our findings indicate that geometry measures differ according to waterbody cross-section shape.
 This makes it an attractive target for inclusion in depth prediction models.
-Unfortunately, identifying a lake’s cross-section shape requires bathymetry data which is unavailable for most lakes.
-However, given the conceptual links between cross-section shape, glaciation, and sedimentation \citep{johanssonNewApproachesModelling2007} it may be advantageous for future studies to compile data on sedimentation to determine if this data can be used to predict cross-section shape and boost depth prediction accuracy.
+Unfortunately, identifying a waterbody’s cross-section shape requires bathymetry data which is unavailable for most waterbodies.
+However, given the conceptual links between cross-section shape, glaciation, and sedimentation (Johansson et al.
+2007) it may be advantageous for future studies to compile data on sedimentation to determine if this data can be used to predict cross-section shape and boost depth prediction accuracy.
+We note that such data does not currently exist for large numbers of waterbodies.
+
+## Conclusion
+To our knowledge, the present study is the largest and most comprehensive test to date of the geometric cone model of waterbody depth.
+Using bathymetry data on approximately 5,000 waterbodies, we show that proxy slope measures are not representative of true in-lake slope and this leads to overestimates of depth in concave and reservoir waterbodies.
+Despite these apparent biases, overall prediction accuracy was equivalent to that of prior depth prediction studies (± 6-7m).
+In addition, models restricted to only concave waterbodies had low error (both in absolute and relative terms) suggesting that the cone model is an adequate representation of depth for these waterbodies.
+
+Only our models using a true measure of in-lake slope had greater accuracy than that of prior studies (± 4.8m, 27%).
+Given that this model requires data which is only available for waterbodies with bathymetry data, it is of limited use for general depth prediction.
+Lack of improved prediction accuracy (short of including data that is unavailable for most waterbodies) suggests that improved prediction may require new types of data or novel analysis techniques.
+
+## DATA AVAILABILITY
+
+All data used in the study is available at Stachelek (2021a). All code for data processing, model fitting, and model evaluation is available at [Zenodo DOI].

### 2021-11-10 manuscript/manuscript.Rmd
embed figures

@@
 We also found that models using nearshore land slope as a proxy systematically overpredict depth in concave waterbodies (i.e.
-bowl-shaped; up to 18% of waterbodies in the study population) and reservoir waterbodies (up to 30% of waterbodies in the study population).
+bowl-shaped; up to 18\% of waterbodies in the study population) and reservoir waterbodies (up to 30\% of waterbodies in the study population).
 Nevertheless, model errors were low (in absolute and relative terms) for concave waterbodies, suggesting that the cone model is an adequate representation of depth for these waterbodies.
@@
 
-## INTRODUCTION
+## Introduction
 
 Depth is an important factor controlling waterbody physics, chemistry, and biota.
-For example, deeper waterbodies generally have higher water clarity and less complete mixing compared to shallow waterbodies (Fee et al.
-1996, Read et al.
-2014).
-These differences are reflected in variation among waterbodies in terms of biological productivity (Qin et al.
-2020) and rates of greenhouse gas production (Li et al.
-2020).
+For example, deeper waterbodies generally have higher water clarity and less complete mixing compared to shallow waterbodies \citep{feeEffectsLakeSize1996, readSimulating2368Temperate2014}.
+These differences are reflected in variation among waterbodies in terms of biological productivity \citep{qinWaterDepthUnderpins2020} and rates of greenhouse gas production \citep{liSignificantContributionLake2020}.
 However, because measured depth data is only available for a small fraction of waterbodies (about 15% of all waterbodies in the area encompassed by our study), our ability to understand and predict depth-dependent processes is limited.
-The importance of waterbody depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all waterbodies such as waterbody surface area or the nearshore slope of the land surrounding a waterbody (Heathcote et al.
-2015, Oliver et al.
-2016, Sobek et al.
-2011).
-Such efforts rely on a strategy of exploiting correlations between nearshore geomorphology and in-lake geometry, which at limited extents (e.g., within a single North American state or province) can be quite strong, while at larger extents can be dependent on geographic location and waterbody type (Oliver et al.
-2016, Branstrator 2009).
+The importance of waterbody depth, coupled with its limited availability, has led to numerous attempts to predict depth using measures available for all waterbodies such as waterbody surface area or the nearshore slope of the land surrounding a waterbody \citep{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}.
+Such efforts rely on a strategy of exploiting correlations between nearshore geomorphology and in-lake geometry, which at limited extents (e.g., within a single North American state or province) can be quite strong, while at larger extents can be dependent on geographic location and waterbody type \citep{oliver2016prediction, branstrator2009origins}.
 
@@
 In the simplest case, where waterbody basins are treated as cones (Eq.
-1, Fig.
+\ref{eq1}, Fig.
 1), two assumptions are required to make depth predictions for all waterbodies: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the waterbody is a representative proxy for the distance to the deepest point of the waterbody (Fig.
@@
 2017).
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/slope_diagram_new}
+  \end{center}
+  \caption{Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry.
+Geometric depth calculated via Equation \ref{eq1} requires a single distance and slope metric.}\label{fig1}
+\end{figure}
 
@@
 
-## METHODS
+## Methods
 ### Data description
@@
 
+\begin{figure}[H]
+  \begin{center}
+    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/00_map_bathy-1.pdf}
+  \end{center}
+  \caption{Map of study lakes showing A) lake e maximum depth measurements, B) cross-section shape class, and C) reservoir classification.
+The distribution of lake depths from panel A is reported in Figure S4.}\label{fig2}
+\end{figure}
+
 The original data came in a variety of formats including pre-interpolated rasters (Minnesota), contour lines (Nebraska, Michigan, Massachusetts, Kansas, Iowa), contour polygons (New Hampshire, Connecticut), or point depth soundings (Maine).
@@
 The size of the moving window was adjusted iteratively to ensure that each bathymetry raster contained no missing data.
+
 All waterbody bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) National Hydrography Dataset (USGS 2019) waterbodies such that source data and bathymetry surface outputs were clipped to the area of each waterbody polygon.
@@
 
-## RESULTS
+## Results
 Waterbodies belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Fig.
@@
 
+\begin{figure}[H]
+  \begin{center}
+    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/01_geometry_base-1.pdf}
+  \end{center}
+  \caption{Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope.
+A best-fit line (solid) and equation is shown to shown to facilitate computation of correction factors for proxy values of lake geometry.
+A 1:1 line (dashed) is shown to illustrate bias.
+Coefficients of determination are shown to illustrate representativeness.}\label{fig3}
+\end{figure}
+
 In addition to overall differences between slope and distance measures, we found differences in these relationships among waterbody shape classes.
@@
 Unlike concave and convex waterbodies, there were not clear differences among slope and distance metrics for natural lakes versus reservoirs.
+
 Offset model fit and prediction error differed depending on the technique used to calculate in-lake and nearshore geometry metrics (Table S2).
@@
 
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/02_depth_model_grid_resid-1}
+  \end{center}
+  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig4}
+\end{figure*}
+
 The most important covariates in these models were those relating to spatial location, waterbody area, and perimeter (Fig.
@@
 These calculations indicate that omitting a geometric max depth term results in a 130%, 60%, or 50% increase in mean square error depending on the formulation of geometric max depth in Eq.
-1.
+\ref{eq1}.
 
-## DISCUSSION
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.45\textwidth,keepaspectratio]{../figures/02_depth_model_importance-1.pdf}
+  \end{center}
+  \caption{Importance plot for random forest variables showing increase in mean square error.
+Higher values indicate greater importance to model predictions.
+See Equation 1 for a definition of geometric max depth.
+HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{fig5}
+\end{figure}
+
+## Discussion
 
@@
 
-## DATA AVAILABILITY
+## Data availability
 

### 2021-11-10 manuscript/manuscript.Rmd
stash progress

@@
 
-Given the limited prediction accuracy of prior depth prediction efforts (± 6-7 m, Heathcote et al.
-2015, Oliver et al.
-2016, Sobek et al.
-2011), a major focus has been on improving accuracy using strategies such as employing more diverse covariates (Oliver et al.
-2016), varying waterbody buffer sizes (Heathcote et al.
-2015), or estimating hidden groupings (e.g.
-fitting different models for distinct size classes) among waterbodies (Cael et al.
-2017, Sobek et al.
-2011).
+Given the limited prediction accuracy of prior depth prediction efforts (± 6-7 m, \citealt{heathcotePredictingBathymetricFeatures2015, oliver2016prediction, sobekPredictingDepthVolume2011}), a major focus has been on improving accuracy using strategies such as employing more diverse covariates \citep{oliver2016prediction}, varying waterbody buffer sizes \citep{heathcotePredictingBathymetricFeatures2015}, or estimating hidden groupings (e.g.
+fitting different models for distinct size classes) among waterbodies \citep{caelVolumeMeanDepth2017, sobekPredictingDepthVolume2011}.
 Unfortunately, the predictive accuracy of these efforts has been limited (± 6-7 m).
 
-One intuitive approach for predicting waterbody depth involves using a geometric model that assumes waterbody basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid (Getirana et al.
-2018, Hollister et al.
-2011, Neumann 1959, Yigzaw et al.
-2018).
+One intuitive approach for predicting waterbody depth involves using a geometric model that assumes waterbody basins correspond to an idealized shape such as a cone, bowl, or an elliptic sinusoid \citep{getiranaDerivingThreeDimensional2018, hollisterPredictingMaximumLake2011, neumannMaximumDepthAverage1959, yigzawNewGlobalStorage2018}.
 All such geometric models for waterbody depth prediction involve implicit assumptions about the terms of geometric formulae.
@@
 
-where the product of slope and horizontal distance yields an exact geometric depth estimate (depthgeometric).
+where the product of slope and horizontal distance yields an exact geometric depth estimate ($depth_{geometric}$).
 Cone models of waterbody basins have been used extensively to estimate hypsography in waterbodies where there is no knowledge of volume or mean depth (Read et al.
@@
 The assumptions of the cone model (as well as other geometric models) can be tested by comparing proxy measures of waterbody geometry against corresponding “true” (i.e.
-in-lake) values derived from bathymetric maps and by evaluating how waterbody cross-section shapes differ from that of an idealized cone (Johansson et al.
-2007).
-For instance, waterbody cross-section shapes have been shown to vary from narrow "convex" forms to outstretched "concave" forms (Hakanson 1977).
+in-lake) values derived from bathymetric maps and by evaluating how waterbody cross-section shapes differ from that of an idealized cone \citep{johanssonNewApproachesModelling2007}.
+For instance, waterbody cross-section shapes have been shown to vary from narrow "convex" forms to outstretched "concave" forms \citep{hakansonLakeFormLake1977}.
 Because tests of geometric model assumptions require bathymetric map data, which is only available for a small fraction of waterbodies (including about 15% of all waterbodies in our study footprint), existing evidence may not be applicable to all waterbodies.
-The few studies that have tested these assumptions have been limited to individual studies of very large (> 500 ha) waterbodies or studies on small numbers (< 100) of waterbodies (Johansson et al.
-2007).
-Studies focused specifically on reservoirs (as opposed to the more typical case where reservoirs and natural lakes are combined), have been even more restricted to that of large waterbodies > 1000 ha (Lehner et al.
-2011, Messager et al.
-2016).
+The few studies that have tested these assumptions have been limited to individual studies of very large (> 500 ha) waterbodies or studies on small numbers (< 100) of waterbodies \citep{johanssonNewApproachesModelling2007}.
+Studies focused specifically on reservoirs (as opposed to the more typical case where reservoirs and natural lakes are combined), have been even more restricted to that of large waterbodies > 1000 ha \citep{lehnerHighresolutionMappingWorld2011, messagerEstimatingVolumeAge2016}.
 
@@
 We supplemented these geometry measures with data classifying waterbodies as reservoirs or natural lakes.
-We used this data to compute geometric depth estimates (Equation 1) and prediction "offsets" to these estimates using the random forest algorithm (Equation 3).
+We used this data to compute geometric depth estimates (Equation \ref{eq1}) and prediction "offsets" to these estimates using the random forest algorithm (Equation \ref{eq3}).
 Such offsets are model quantities which minimize differences between observed and predicted depth as a function of covariates.
@@
 S1).
-Finally, we expect that depth predictions themselves will be strongly related to waterbody area and hydrologic subbasin variables as these measures have been influential in prior studies (Oliver et al.
-2016).
+Finally, we expect that depth predictions themselves will be strongly related to waterbody area and hydrologic subbasin variables as these measures have been influential in prior studies \citep{oliver2016prediction}.
 
@@
 We compiled bathymetry data on approximately 5,000 waterbodies in the Northeastern and Midwestern United States from nine official state databases (Fig.
-2).
+\ref{fig2}).
 These represent approximately 15% of all waterbodies in the States included in our study and are a diverse cross section in terms of their characteristics, surface areas, and span a wide geographic extent including glaciated and non-glaciated regions (Table 1).
@@
 For the Minnesota data, we simply clipped the raster for each waterbody to its outline.
-For data from the remaining States, we processed each waterbody by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the raster R package (Hijmans 2019).
+For data from the remaining States, we processed each waterbody by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the raster R package \citep{hijmansRasterGeographicData2019}.
 The size of the moving window was adjusted iteratively to ensure that each bathymetry raster contained no missing data.
 
-All waterbody bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) National Hydrography Dataset (USGS 2019) waterbodies such that source data and bathymetry surface outputs were clipped to the area of each waterbody polygon.
+All waterbody bathymetry was specifically calculated relative to high-resolution (1:24,000 scale) National Hydrography Dataset \citep{usgsNationalHydrographyDataset2019} waterbodies such that source data and bathymetry surface outputs were clipped to the area of each waterbody polygon.
 We restricted the waterbodies in our study to those with an area of at least 4 ha and a maximum depth of at least 0.3 m (1 ft.).
@@
 natural lake, reservoir).
-For cross-section shape, we categorized waterbodies as either convex or concave following the method of Hakanson (1977) by computing normalized waterbody depth-area relationships (i.e.
+For cross-section shape, we categorized waterbodies as either convex or concave following the method of \citet{hakansonLakeFormLake1977} by computing normalized waterbody depth-area relationships (i.e.
 hypsographic curves) and assigning class membership based on whether the midpoint of a waterbody’s curve falls above or below that of a simple straight-sided cone (Fig.
@@
 
-We further categorized waterbodies using the output of a deep convolutional neural network model trained on satellite images labelled according to whether there was visual evidence of a water control structure significantly impacting flow (Polus et.
-al 2021).
+We further categorized waterbodies using the output of a deep convolutional neural network model trained on satellite images labelled according to whether there was visual evidence of a water control structure significantly impacting flow \citep{polus2020}.
 This model had an overall validation accuracy of 81% and produced a probability for each waterbody as to whether it is a reservoir or a natural lake.
 For our purposes, we set a conservative classification probability threshold of 0.75 to determine whether a waterbody would be considered a reservoir.
-For example, if the Polus et al.
-(2021) dataset classified a particular waterbody as a reservoir with a probability of 0.74 we categorized it as a natural lake but if the probability was greater than 0.75 we categorized it as a reservoir.
+For example, if the \citet{polus2020} dataset classified a particular waterbody as a reservoir with a probability of 0.74 we categorized it as a natural lake but if the probability was greater than 0.75 we categorized it as a reservoir.
 Note that our reservoir classification defines reservoirs as any permanent waterbody that has a water control structure likely to significantly impact flow or pool water.
-It makes no distinction between different dam types, heights, or uses/purposes because the Polus et al.
-(2021) dataset is only based on visual interpretation of waterbody images (via deep convolutional neural network models).
-However, the Polus et al.
-(2021) dataset is unique in that it provides data using a standardized approach at broad spatial extents for waterbodies > 4 ha.
+It makes no distinction between different dam types, heights, or uses/purposes because the \citet{polus2020} dataset is only based on visual interpretation of waterbody images (via deep convolutional neural network models).
+However, the \citet{polus2020} dataset is unique in that it provides data using a standardized approach at broad spatial extents for waterbodies > 4 ha.
 
@@
 where sinuous waterbodies have larger values of shoreline development and circular waterbodies have smaller values of shoreline development.
-Watershed to waterbody area ratio is an approximation of water residence time and is defined as watershed area divided by waterbody area (Timms 2009).
+Watershed to waterbody area ratio is an approximation of water residence time and is defined as watershed area divided by waterbody area \citep{timmsGeomorphologyLakeBasins2009}.
 
@@
 in-lake slope, distance to the deepest point of the waterbody) by visual inspection (i.e.
-plotting each proxy measure against its corresponding true value) and by computing coefficients of determination (R2).
+plotting each proxy measure against its corresponding true value) and by computing coefficients of determination ($R^2$).
 We further tested proxy measures by examining their effect on waterbody depth prediction error.
 Our approach involved several steps.
-In the first step, we computed a geometric estimate of waterbody depth using only geometry information (depthgeometric, Equation 1).
+In the first step, we computed a geometric estimate of waterbody depth using only geometry information ($depth_{geometric}$, Equation \ref{eq1}).
 In the second step, we fit a random forest model to predict observed (i.e.
 true) depth as a function of geometric depth along with several covariates available for all waterbodies (Table 1).
-The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of depthgeometric and among different waterbody types.
+The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of $depth_{geometric}$ and among different waterbody types.
 Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table 2).
@@
 ### Geometric model 
-We used a geometric model of waterbodies where basins are treated as cones with a fixed relationship between slope and distance (Equation 1).
+We used a geometric model of waterbodies where basins are treated as cones with a fixed relationship between slope and distance (Equation \ref{eq1}).
 One reason that we used the cone model is that, unlike other idealized shapes, it does not require any knowledge of waterbody volume or mean depth.
-Note that Equation 1 is a geometric formula and has no intercept or coefficients and it produces an exact depth value given true values of slope and distance.
+Note that Equation \ref{eq1} is a geometric formula and has no intercept or coefficients and it produces an exact depth value given true values of slope and distance.
 To use this model to predict the depth of all waterbodies, there is a necessary assumption that proxy slope and distance measures, which are available for all waterbodies, are representative of true slope and distance (Fig.
@@
 ### Random forest models
-Prior studies using geometric models to predict waterbody depth include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy (Hollister et al.
-2011, Yigzaw et al.
-2018).
+Prior studies using geometric models to predict waterbody depth include a statistical or machine learning model “layer” or “offset” to boost predictive accuracy \citep{hollisterPredictingMaximumLake2011, yigzawNewGlobalStorage2018}.
 This procedure involves fitting a statistical or machine learning model to the residuals of an initial geometric model.
-For our purposes, such offset modeling enabled us to test our expectations that prediction error would be different among different formulations of depthgeometric and among different waterbody types.
+For our purposes, such offset modeling enabled us to test our expectations that prediction error would be different among different formulations of $depth_{geometric}$ and among different waterbody types.
 It also facilitated direct comparison against prior models of waterbody depth including those that are non-geometric.
-We generated an “offset” to geometric depth (sensu Hollister et al.
-2011) using the random forest algorithm and the ranger R package (Wright and Ziegler 2017) to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation 1) along with the waterbody elevation, area, perimeter, and ratio/index measures listed in Table 2:
+We generated an “offset” to geometric depth (sensu \citealt{hollisterPredictingMaximumLake2011}) using the random forest algorithm and the ranger R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation \ref{eq1}) along with the waterbody elevation, area, perimeter, and ratio/index measures listed in Table 2:
 
@@
 
-We evaluated the relative “importance” of individual covariates by comparing model performance between 1) models where a given covariate was left untouched versus 2) models where a given covariate was permuted randomly (Prasad et al.
-2006, Wright and Ziegler 2017).
+We evaluated the relative “importance” of individual covariates by comparing model performance between 1) models where a given covariate was left untouched versus 2) models where a given covariate was permuted randomly \citep{prasadNewerClassificationRegression2006, wrightRangerFastImplementation2017}.
 Neither cross-section shape nor reservoir class was used as a covariate in any random forest models.
@@
 We used the random forest algorithm because it makes no assumptions about the distribution of model residuals, allows for non-linearity, and is insensitive to interactions (i.e.
-multicollinearity) among covariates (Prasad et al.
-2006).
+multicollinearity) among covariates \citep{prasadNewerClassificationRegression2006}.
 
 ### Model comparisons
-We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation 1 (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
-Prior to entry into Equation 1, we standardized proxy distances to have the same numeric range as their true counterpart.
+We tested model sensitivity to slope and distance proxies by generating multiple “geometric maximum depth” estimates from 3 different model runs using each of the possible metric combinations for Equation \ref{eq1} (true slope - proxy distance, proxy slope - true distance, proxy slope - proxy distance).
+Prior to entry into Equation \ref{eq1}, we standardized proxy distances to have the same numeric range as their true counterpart.
 The purpose of this standardization was to prevent waterbodies with extremely long proxy distances from having an outsized impact on model evaluation metrics.
@@
 ### Model evaluations
-We evaluated model fit and prediction error using root-mean-square error (RMSE), mean absolute percent error (MAPE), and coefficient of determination (R2) metrics on a holdout set (i.e.
+We evaluated model fit and prediction error using root-mean-square error (RMSE), mean absolute percent error (MAPE), and coefficient of determination ($R^2$) metrics on a holdout set (i.e.
 a data subset not used for model training) containing 25% of all waterbodies.
@@
 
-Although proxy distance to waterbody center was often larger in magnitude compared to the true distance to the deepest point of waterbodies’ (rather than being identical), they were strongly related (R2 = 0.8).
+Although proxy distance to waterbody center was often larger in magnitude compared to the true distance to the deepest point of waterbodies’ (rather than being identical), they were strongly related ($R^2$ = 0.8).
 Note that the coefficient of determination for this relationship is not strictly correct given that distance to waterbody center is an upper bound on distance to the deepest point of waterbodies.
-In contrast to distance metrics, proxy nearshore land slope and true in-lake slope were more weakly related (R2 = 0.17).
+In contrast to distance metrics, proxy nearshore land slope and true in-lake slope were more weakly related ($R^2$ = 0.17).
 For slope measures, most waterbodies had higher magnitude (i.e.
@@
 The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to waterbody center (Table 2).
-More specifically, the true slope (in-lake slope) - proxy distance (distance to the center of the waterbody) model had a better fit (R2 = 0.73) and lower prediction error (RMSE = 4.8m, MAPE = 27%) compared to the proxy slope - true distance model (R2 = 0.31, RMSE = 7.3m, MAPE = 64%).
-The fit of the proxy slope -proxy distance model (R2 = 0.36, RMSE = 7.1m, MAPE = 61%) was very similar to the proxy slope – true distance model.
+More specifically, the true slope (in-lake slope) - proxy distance (distance to the center of the waterbody) model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.8m, MAPE = 27%) compared to the proxy slope - true distance model ($R^2$ = 0.31, RMSE = 7.3m, MAPE = 64%).
+The fit of the proxy slope -proxy distance model ($R^2$ = 0.36, RMSE = 7.1m, MAPE = 61%) was very similar to the proxy slope – true distance model.
 Predicted depth values for this model were generally underestimates relative to measured depth values (Fig.
@@
 Higher values indicate greater importance to model predictions.
-See Equation 1 for a definition of geometric max depth.
+See Equation \ref{eq1} for a definition of geometric max depth.
 HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{fig5}
@@
 A simple indication of this non-representativeness is that proxy nearshore land slope was often (in > 74% of cases) steeper than true in-lake slope.
-This finding is consistent with Heathcote et al.
-(2015) whose results suggest that in-lake slopes are shallower compared to the surrounding land.
+This finding is consistent with \citet{heathcotePredictingBathymetricFeatures2015} whose results suggest that in-lake slopes are shallower compared to the surrounding land.
 Furthermore, the fact that in-lake slopes were shallower compared to the surrounding land even after controlling for differences in area (Fig.
 S4) is consistent with the idea of topographic scaling (i.e.
-scale invariance) explored in previous work and detailed by (Cael et al.
-2017).
-The underlying reason for these shallow in-lake slopes may be related to slope-induced turbidity currents which distribute sediment from shallow high-energy areas of waterbodies to deep low-energy areas (Hakanson 1981, Johansson et al.
-2007).
-The strength of such sediment focusing is likely greater in "younger" waterbodies with steeper slopes leading to a smoothing of their bathymetry over time (Blais and Kalff 1995).
+scale invariance) explored in previous work and detailed by \citep{caelVolumeMeanDepth2017}.
+The underlying reason for these shallow in-lake slopes may be related to slope-induced turbidity currents which distribute sediment from shallow high-energy areas of waterbodies to deep low-energy areas \citep{hakansonLakeBottomDynamics1981, johanssonNewApproachesModelling2007}.
+The strength of such sediment focusing is likely greater in "younger" waterbodies with steeper slopes leading to a smoothing of their bathymetry over time \citep{blaisInfluenceLakeMorphometry1995}.
 
@@
 This is contrary to the idea that most reservoirs are drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
-next to the dam) rather than in the center of the reservoir (Lanza and Silvey 1985).
+next to the dam) rather than in the center of the reservoir \citep{lanza1985interactions}.
 One possible explanation is that our reservoir classification data uses a more general definition of a reservoir (i.e.
@@
 S8).
-The underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of waterbody formation whereby the formation of concave waterbodies were a result of more intense glacial scouring compared to that of convex waterbodies (Gorham 1958).
-While our findings provide some evidence in support of this idea, namely that there is a geographic hotspot of concave waterbodies associated with the glaciated “prairie pothole region” (See Hayashi and van der Kamp 2000), the overall geographic distribution of waterbody cross-section shapes does not support this idea.
+The underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of waterbody formation whereby the formation of concave waterbodies were a result of more intense glacial scouring compared to that of convex waterbodies \citep{gorhamPhysicalLimnologyNorthern1958}.
+While our findings provide some evidence in support of this idea, namely that there is a geographic hotspot of concave waterbodies associated with the glaciated “prairie pothole region” (see \citealt{hayashiSimpleEquationsRepresent2000}), the overall geographic distribution of waterbody cross-section shapes does not support this idea.
 Instead of a concentrated area of concave waterbodies in formerly glaciated regions, there appears to be a fairly even mix of concave and convex waterbodies distributed amongst the northern (i.e.
@@
 ### Effects of proxy measures of waterbody geometry depth prediction error
-Models using only proxy variables (Table 2) had prediction error rates (RMSE = 7.1m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting waterbody depth at broad geographic extents (Hollister et al.
-2011, Oliver et al.
-2016, Messager et al.
-2016).
+Models using only proxy variables (Table 2) had prediction error rates (RMSE = 7.1m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting waterbody depth at broad geographic extents \citep{hollisterPredictingMaximumLake2011, messagerEstimatingVolumeAge2016, oliver2016prediction}.
 When only a single proxy measure was used, there was a difference in model sensitivity depending on if it was a horizontal distance measure or a vertical slope measure.
-In the case of a true slope and proxy distance combination, models were more accurate (± 4.8m, 27%) than even the most accurate of prior studies (Hollister et al.
-2011, Oliver et al.
-2016, Messager et al.
-2016).
+In the case of a true slope and proxy distance combination, models were more accurate (± 4.8m, 27%) than even the most accurate of prior studies \citep{hollisterPredictingMaximumLake2011, messagerEstimatingVolumeAge2016, oliver2016prediction}.
 Conversely, models using a proxy slope and true distance combination had prediction error rates (± 7.3m, 64%) of a similar magnitude as that of the baseline proxy-proxy model (± 7.1m, 61%).
@@
 We did not find evidence that the sensitivity of depth predictions to slope was dependent on variations in how these measures were calculated (Table S2).
-In a general sense, the sensitivity of depth predictions to slope help explain the relatively poor predictive performance of prior non-geometric waterbody depth models given that they rely heavily on waterbody area as a predictor (Messager et al.
-2016, Oliver et al.
-2016, Sobek et al.
-2011) and both horizontal distance measures and vertical slope measures appear to be decoupled from waterbody area (Fig.
+In a general sense, the sensitivity of depth predictions to slope help explain the relatively poor predictive performance of prior non-geometric waterbody depth models given that they rely heavily on waterbody area as a predictor \citep{sobekPredictingDepthVolume2011, messagerEstimatingVolumeAge2016, oliver2016prediction} and both horizontal distance measures and vertical slope measures appear to be decoupled from waterbody area (Fig.
 S8).
@@
 Topobathymetry would allow for more robust tests of the representativeness of geometric model inputs.
-Unfortunately, topobathymetric products are rare, have mostly been limited nearshore marine environments, and as such are not yet widely available for inland waters (Danielson et al.
-2016).
+Unfortunately, topobathymetric products are rare, have mostly been limited nearshore marine environments, and as such are not yet widely available for inland waters \citep{danielsonTopobathymetricElevationModel2016}.
 Other potential explanatory data include information on waterbody origin and development.
-Unfortunately, such “waterbody ontogeny” data is presently available only for select regions and the very largest waterbodies (Sharma and Byrne 2011).
+Unfortunately, such “waterbody ontogeny” data is presently available only for select regions and the very largest waterbodies \citep{sharmaComparisonTitanNorth2011}.
 
@@
 Unfortunately, identifying a waterbody’s cross-section shape requires bathymetry data which is unavailable for most waterbodies.
-However, given the conceptual links between cross-section shape, glaciation, and sedimentation (Johansson et al.
-2007) it may be advantageous for future studies to compile data on sedimentation to determine if this data can be used to predict cross-section shape and boost depth prediction accuracy.
+However, given the conceptual links between cross-section shape, glaciation, and sedimentation \citep{johanssonNewApproachesModelling2007} it may be advantageous for future studies to compile data on sedimentation to determine if this data can be used to predict cross-section shape and boost depth prediction accuracy.
 We note that such data does not currently exist for large numbers of waterbodies.
@@
 
-All data used in the study is available at Stachelek (2021a). All code for data processing, model fitting, and model evaluation is available at [Zenodo DOI].
+All data used in the study is available at \citet{stachelekjemmaBathymetryData0002021}. All code for data processing, model fitting, and model evaluation is available at [Zenodo DOI].

### 2021-11-10 manuscript/manuscript.Rmd
stash progress

@@
 \ref{eq1}, Fig.
-1), two assumptions are required to make depth predictions for all waterbodies: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the waterbody is a representative proxy for the distance to the deepest point of the waterbody (Fig.
-1).
+\ref{fig1}), two assumptions are required to make depth predictions for all waterbodies: 1) that nearshore land slope is a representative proxy for in-lake slope and 2) that the distance to the center of the waterbody is a representative proxy for the distance to the deepest point of the waterbody (Fig.
+\ref{fig1}).
 This cone model imposes the following fixed (i.e.
@@
 where the product of slope and horizontal distance yields an exact geometric depth estimate ($depth_{geometric}$).
-Cone models of waterbody basins have been used extensively to estimate hypsography in waterbodies where there is no knowledge of volume or mean depth (Read et al.
-2014, Winslow et al.
-2017).
+Cone models of waterbody basins have been used extensively to estimate hypsography in waterbodies where there is no knowledge of volume or mean depth \citep{readSimulating2368Temperate2014, winslowLargescaleModeledContemporary2017}.
 
@@
 Therefore, we expect that the use of this proxy will lead to overestimation of waterbody depth (Fig.
-1).
+\ref{fig1}).
 Furthermore, we expect to see greater overestimation error in reservoirs as compared to natural lakes because many reservoirs are known to be drowned river valleys where the deepest point is close to the edge at the end of the reservoir (i.e.
-next to the dam) rather than in the center of the reservoir (Lanza and Silvey 1985).
+next to the dam) rather than in the center of the reservoir \citep{lanza1985interactions}.
 In a similar fashion, we expect to see overestimation error associated with using a nearshore land slope proxy in waterbodies with differing cross-section shape such that the depth of bowl-shaped (i.e.
@@
 its “visual distance to waterbody center”).
-For these calculations, we used the polylabelr R package (Larsson 2019), which interfaces with the Mapbox pole of inaccessibility algorithm (Agafonkin 2019).
+For these calculations, we used the polylabelr R package \citep{larssonPolylabelrFindPole2019}, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
 We calculated (maximum) in-lake slope as depth at the deepest point divided by the shortest distance to the deepest point from the shoreline..
-We calculated (mean) nearshore land slope for each waterbody by computing the slope within a 100-m buffer using data from a high-resolution digital elevation model (~ 15x15m grain) accessed using the elevatr R package (Hollister and Shah 2017).
-We explored alternative buffer sizes ranging between 50-1000 m following Sobek et al.
-(2011) and although 100 m provided the lowest model error, we ultimately found little appreciable effect of varying buffers on model performance.
-Slope computations proceeded by passing a 3x3 moving window over the 100-m buffer to calculate the slope at each point using Horn's algorithm via the terrain function in the raster R package (Hijmans 2019).
+We calculated (mean) nearshore land slope for each waterbody by computing the slope within a 100-m buffer using data from a high-resolution digital elevation model (~ 15x15m grain) accessed using the elevatr R package \citep{hollisterElevatrAccessElevation2017}.
+We explored alternative buffer sizes ranging between 50-1000 m following \citet{sobekPredictingDepthVolume2011} and although 100 m provided the lowest model error, we ultimately found little appreciable effect of varying buffers on model performance.
+Slope computations proceeded by passing a 3x3 moving window over the 100-m buffer to calculate the slope at each point using Horn's algorithm via the terrain function in the raster R package \citep{hijmansRasterGeographicData2019}.
 Reported nearshore land slope values are the mean of all points in the buffer.
@@
 To use this model to predict the depth of all waterbodies, there is a necessary assumption that proxy slope and distance measures, which are available for all waterbodies, are representative of true slope and distance (Fig.
-1).
+\ref{fig1}).
 
@@
 Waterbodies belonging to each cross-section shape and reservoir class were not evenly distributed across our study area (Fig.
-2).
+\ref{fig2}).
 For example, concave waterbodies were nearly absent from Michigan whereas Maine had an overabundance (~3%) of waterbodies categorized as neither concave nor convex.
 Waterbodies in the southern portions of our study area tended to be classified as reservoirs whereas waterbodies in the northern portions of our study area were a more even mix between reservoirs and natural lakes (Fig.
-2).
+\ref{fig2}).
 Approximately 18%, 80%, and 2% of waterbodies were classified as having a concave, convex, or neither shape respectively whereas approximately 30% and 70% of waterbodies were classified as being a reservoir or a natural lake.
@@
 steeper) nearshore land slope compared to true in-lake slope (Fig.
-3).
+\ref{fig3}).
 Taken together, these results suggest that proxy distance to the center of waterbodies is representative of true distance to the deepest point of waterbodies whereas proxy nearshore land slope is not representative of true in-lake slope.
 The strong relationship between distance to the center of waterbodies and distance to the deepest point means that it is possible to convert between the two measures in subsequent analyses (See best-fit equations in Fig.
-3).
+\ref{fig3}).
 
@@
 Furthermore, analysis of model residuals showed overestimation of waterbody depth for concave waterbodies when models included a proxy slope measure (Fig.
-4).
+\ref{fig4}).
 We observed similar but smaller overestimation depending on if a waterbody was classified as a reservoir rather than a natural lake (Fig.
-4).
+\ref{fig4}).
 We found that models restricted to consider only concave lakes had lower error (both in absolute and relative terms) compared to models on other data subsets (e.g.
@@
 The most important covariates in these models were those relating to spatial location, waterbody area, and perimeter (Fig.
-5).
+\ref{fig5}).
 Conversely, watershed metrics and waterbody elevation had little contribution to random forest model fit.
@@
 To evaluate the contribution of our “offset” models relative to the “base” geometric model, we can look at model importance calculations for the geometric max depth input to the random forest model (Fig.
-5).
+\ref{fig5}).
 These calculations indicate that omitting a geometric max depth term results in a 130%, 60%, or 50% increase in mean square error depending on the formulation of geometric max depth in Eq.
@@
 glaciated) and southern (non-glaciated) portions of our study area (Fig.
-2).
+\ref{fig2}).
 

### 2021-11-10 manuscript/si_template_2019.tex
stash progress

@@
 
-\authors{J. Stachelek\affil{1,2}, P. J. Hanly\affil{1}, and P. A. Soranno\affil{1}}
+\authors{J. Stachelek\affil{1,2,3}, P. J. Hanly\affil{3}, and P. A. Soranno\affil{3}}
 
-\affiliation{1}{Department of Fisheries and Wildlife, Michigan State University, 480
+\affiliation{1}{Earth and Environmental Sciences Division, Los Alamos National Laboratory, Los Alamos, NM, USA}
+\affiliation{2}{Center for Limnology, University of Wisconsin – Madison, Madison, WI, USA}
+\affiliation{3}{Department of Fisheries and Wildlife, Michigan State University, 480
       Wilson Rd., East Lansing, MI 48824, USA}
-\affiliation{2}{Center for Limnology, University of Wisconsin – Madison, Madison, WI, USA}
 

### 2021-11-10 manuscript/manuscript.Rmd
stash tables

@@
 Such offsets are model quantities which minimize differences between observed and predicted depth as a function of covariates.
-In our case, covariates included a variety of waterbody, watershed, and hydrologic subbasin measures that are available for all waterbodies (Table 1).
+In our case, covariates included a variety of waterbody, watershed, and hydrologic subbasin measures that are available for all waterbodies (Table \ref{table1}).
+
+\begin{table}
+  \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States).
+Predictor variables for computing random forest offsets (Equation \ref{eq2}) are printed in bold face.
+Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset.
+The total number lakes is reported as n.} \label{table1}
+  \centering
+  \setlength\tabcolsep{1.5pt} % default value: 6pt
+  \begin{tabular}{lllll}
+    \hline
+    Variable & Median & Q25 & Q75 & n\\
+    \hline
+    Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4820 (17700)\\
+    \textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4820 (17700)\\
+    \textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4820 (17700)\\
+    \textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4820 (17700)\\
+    \textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4820 (17700)\\
+    \textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4820 (17700)\\    
+    \textbf{Watershed-lake ratio} & 7.8 (10) & 3.9 (4.4) & 17 (28) & 4820 (17700)\\
+    Deepest point distance (m) & 180 (-) & 110 (-) & 290 (-) & 4820 (-)\\
+    Mean deepest point distance (m) & 140 (-) & 87 (-) & 230 (-) & 4820 (-)\\
+    \textbf{Visual center distance(m)} & 240 (-) & 160 (-) & 390 (-) & 4820 (-)\\
+    Inlake slope (m/m) & 0.05 (-) & 0.02 (-) & 0.08 (-) & 4820 (-)\\
+    Inlake slope online (m/m) & 0.06 (-) & 0.03 (-) & 0.14 (-) & 4800 (-)\\
+    Inlake slopes (m/m) & 0.06 (-) & 0.03 (-) & 0.1 (-) & 4820 (-)\\
+    Inlake slopes online (m/m) & 0.07 (-) & 0.03 (-) & 0.15 (-) & 4800 (-)\\
+    Mean inlake slope (m/m) & 0.04 (-) & 0.02 (-) & 0.09 (-) & 4820 (-)\\
+    Nearshore mean slope (m/m) & 0.08 (-) & 0.05 (-) & 0.11 (-) & 4820 (-)\\
+    Nearshore slope online (m/m) & 0.08 (-) & 0.04 (-) & 0.13 (-) & 4590 (-)\\
+    Nearshore slopes online (m/m) & 0.08 (-) & 0.04 (-) & 0.13 (-) & 4540 (-)\\
+    \bottomrule
+    \end{tabular}  
+  \end{table}
 
@@
 \ref{fig2}).
-These represent approximately 15% of all waterbodies in the States included in our study and are a diverse cross section in terms of their characteristics, surface areas, and span a wide geographic extent including glaciated and non-glaciated regions (Table 1).
+These represent approximately 15% of all waterbodies in the States included in our study and are a diverse cross section in terms of their characteristics, surface areas, and span a wide geographic extent including glaciated and non-glaciated regions (Table \ref{table1}).
 Thus, they can be considered representative of the entire population of lakes in our study extent.
@@
 
-Covariates used in random forest modeling (Table 1, Equation 3, see Random Forests Model sub-section below) for waterbody elevation, area, island area, perimeter, shoreline development, watershed to waterbody area ratio, and hydrologic subbasin (i.e.
+Covariates used in random forest modeling (Table \ref{table1}, Equation 3, see Random Forests Model sub-section below) for waterbody elevation, area, island area, perimeter, shoreline development, watershed to waterbody area ratio, and hydrologic subbasin (i.e.
 HUC4s), were obtained from the LAGOS-US LOCUS database.
@@
 In the second step, we fit a random forest model to predict observed (i.e.
-true) depth as a function of geometric depth along with several covariates available for all waterbodies (Table 1).
+true) depth as a function of geometric depth along with several covariates available for all waterbodies (Table \ref{table1}).
 The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of $depth_{geometric}$ and among different waterbody types.
@@
 “filtered”) to include only reservoirs, only natural lakes, only convex waterbodies, or only concave waterbodies (Table 2).
+
+\begin{table}
+  \caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, R2 = coefficient of determination, MAPE = mean absolute percent error) for all combinations of true (in-lake slope, distance to the deepest point of the waterbody) and proxy (nearshore land slope, distance to waterbody center) metrics.} \label{table2}
+  \centering
+  % \setlength\tabcolsep{1.5pt} % default value: 6pt
+\begin{tabular}{llll}
+  \hline
+  slope & distance & RMSE & $R^2$\\
+  \hline
+  true & true & - & -\\
+  true & proxy & 4.2 m & 0.73\\
+  proxy & true & 6.9 m & 0.26\\
+  proxy & proxy & 6.6 m & 0.31\\
+  \bottomrule
+\end{tabular}
+\end{table}
 
@@
 Although our analysis was limited to waterbodies with available bathymetry data, these waterbodies did not have characteristics that differed from that of the overall waterbody population (apart from the fact that our study waterbodies were somewhat larger in area compared to the overall waterbody population, See Fig.
-S7, Table 1, Fig.
+S7, Table \ref{table1}, Fig.
 S10).

### 2021-11-10 manuscript/manuscript.Rmd
good looking render of accepted ms

@@
 date_subtitle: > 
-  This is the accepted version of the following article: J.
-Stachelek et al., Inland Waters.
-In Press.
-This article may be used for non-commercial purposes in accordance with Wiley Terms and Conditions for Use of Self-Archived Versions.
+  This is an Accepted Manuscript of an article to be published by Taylor & Francis in Inland Waters.
 
@@
 
-where the product of slope and horizontal distance yields an exact geometric depth estimate ($depth_{geometric}$).
+\noindent where the product of slope and horizontal distance yields an exact geometric depth estimate ($depth_{geometric}$).
 Cone models of waterbody basins have been used extensively to estimate hypsography in waterbodies where there is no knowledge of volume or mean depth \citep{readSimulating2368Temperate2014, winslowLargescaleModeledContemporary2017}.
@@
 Additionally, it is unclear whether model prediction error is related to differences in waterbody type such those with different cross-section shapes (concave versus convex) or those classified as reservoirs versus natural lakes.
-To address these knowledge gaps, we asked three research questions: (1) How representative is nearshore land slope of in-lake slope; and how representative is the distance to the center of a waterbody compared to the distance to the deepest point of a waterbody?
-(2) How does the use of proxies for waterbody geometry affect waterbody depth prediction error?
-(3) How does waterbody cross-section shape (i.e.
-concave versus convex) and waterbody type (i.e.
-natural lake vs reservoir) affect depth prediction error?
+To address these knowledge gaps, we asked three research questions:
+
+\begin{enumerate}
+\item How representative is nearshore land slope of in-lake slope; and how representative is the distance to the center of a waterbody compared to the distance to the deepest point of a waterbody?
+\item How does the use of proxies for waterbody geometry affect waterbody depth prediction error?
+\item How does waterbody cross-section shape (i.e. concave versus convex) and waterbody type (i.e. natural lake vs reservoir) affect depth prediction error?
+\end{enumerate}
+
 To answer these questions, we extracted maximum depth (hereafter referred to as “observed maximum depth”), in-lake slope, cross-section shape (i.e., concave versus convex), and distance to the deepest point, of approximately 5,000 waterbodies from bathymetric map data.
@@
 For the Minnesota data, we simply clipped the raster for each waterbody to its outline.
-For data from the remaining States, we processed each waterbody by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the raster R package \citep{hijmansRasterGeographicData2019}.
+For data from the remaining States, we processed each waterbody by converting its original representation to a point layer (if necessary), rasterizing these points, and creating an interpolated bathymetry “surface” using a simple moving window average in the \texttt{raster} R package \citep{hijmansRasterGeographicData2019}.
 The size of the moving window was adjusted iteratively to ensure that each bathymetry raster contained no missing data.
@@
 its “visual distance to waterbody center”).
-For these calculations, we used the polylabelr R package \citep{larssonPolylabelrFindPole2019}, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
-We calculated (maximum) in-lake slope as depth at the deepest point divided by the shortest distance to the deepest point from the shoreline..
-We calculated (mean) nearshore land slope for each waterbody by computing the slope within a 100-m buffer using data from a high-resolution digital elevation model (~ 15x15m grain) accessed using the elevatr R package \citep{hollisterElevatrAccessElevation2017}.
+For these calculations, we used the \texttt{polylabelr} R package \citep{larssonPolylabelrFindPole2019}, which interfaces with the Mapbox pole of inaccessibility algorithm \citep{agafonkinJSLibraryFinding2019}.
+We calculated (maximum) in-lake slope as depth at the deepest point divided by the shortest distance to the deepest point from the shoreline.
+We calculated (mean) nearshore land slope for each waterbody by computing the slope within a 100-m buffer using data from a high-resolution digital elevation model (~ 15x15m grain) accessed using the \texttt{elevatr} R package \citep{hollisterElevatrAccessElevation2017}.
 We explored alternative buffer sizes ranging between 50-1000 m following \citet{sobekPredictingDepthVolume2011} and although 100 m provided the lowest model error, we ultimately found little appreciable effect of varying buffers on model performance.
-Slope computations proceeded by passing a 3x3 moving window over the 100-m buffer to calculate the slope at each point using Horn's algorithm via the terrain function in the raster R package \citep{hijmansRasterGeographicData2019}.
+Slope computations proceeded by passing a 3x3 moving window over the 100-m buffer to calculate the slope at each point using Horn's algorithm via the terrain function in the \texttt{raster} R package \citep{hijmansRasterGeographicData2019}.
 Reported nearshore land slope values are the mean of all points in the buffer.
@@
 
-where sinuous waterbodies have larger values of shoreline development and circular waterbodies have smaller values of shoreline development.
+\noindent where sinuous waterbodies have larger values of shoreline development and circular waterbodies have smaller values of shoreline development.
 Watershed to waterbody area ratio is an approximation of water residence time and is defined as watershed area divided by waterbody area \citep{timmsGeomorphologyLakeBasins2009}.
@@
 The purpose of this random forest “offset” modeling was to more rigorously test our expectations regarding prediction error among different formulations of $depth_{geometric}$ and among different waterbody types.
-Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table 2).
+Each of these steps were executed iteratively for each combination of true and proxy values of slope and distance (Table \ref{table2}).
 We conducted additional sensitivity analysis to examine possible interactions between different proxy measures of waterbody geometry and different subsets of the entire dataset where model data was restricted (i.e.
-“filtered”) to include only reservoirs, only natural lakes, only convex waterbodies, or only concave waterbodies (Table 2).
+“filtered”) to include only reservoirs, only natural lakes, only convex waterbodies, or only concave waterbodies (Table \ref{table2}).
 
@@
   % \setlength\tabcolsep{1.5pt} % default value: 6pt
-\begin{tabular}{llll}
+\begin{tabular}{llllll}
 \hline
-  slope & distance & RMSE & $R^2$\\
+Filter & Slope & Distance & RMSE & $R^2$ & MAPE\\
 \hline
-  true & true & - & -\\
-  true & proxy & 4.2 m & 0.73\\
-  proxy & true & 6.9 m & 0.26\\
-  proxy & proxy & 6.6 m & 0.31\\
+all & true & true &  &  & \\
+ & true & proxy & 4.8 m & 0.73 & 27 \%\\
+ & proxy & true & 7.3 m & 0.31 & 64 \%\\
+ & proxy & proxy & 7.1 m & 0.36 & 61 \%\\
+\hline
+reservoir & true & true &  &  & \\
+ & true & proxy & 5.3 m & 0.66 & 36 \%\\
+ & proxy & true & 7 m & 0.40 & 61 \%\\
+ & proxy & proxy & 7 m & 0.41 & 60 \%\\
+\hline
+natural lake & true & true &  &  & \\
+ & true & proxy & 4.1 m & 0.74 & 22 \%\\
+ & proxy & true & 6.7 m & 0.26 & 68 \%\\
+ & proxy & proxy & 6.6 m & 0.29 & 64 \%\\
+\hline
+convex & true & true &  &  & \\
+ & true & proxy & 4.7 m & 0.74 & 30 \%\\
+ & proxy & true & 7.2 m & 0.34 & 59 \%\\
+ & proxy & proxy & 6.9 m & 0.39 & 58 \%\\
+\hline
+concave & true & true &  &  & \\
+ & true & proxy & 1.6 m & 0.78 & 20 \%\\
+ & proxy & true & 3.1 m & 0.14 & 46 \%\\
+ & proxy & proxy & 3 m & 0.17 & 45 \%\\
 \bottomrule
@@
 It also facilitated direct comparison against prior models of waterbody depth including those that are non-geometric.
-We generated an “offset” to geometric depth (sensu \citealt{hollisterPredictingMaximumLake2011}) using the random forest algorithm and the ranger R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation \ref{eq1}) along with the waterbody elevation, area, perimeter, and ratio/index measures listed in Table 2:
+We generated an “offset” to geometric depth (sensu \citealt{hollisterPredictingMaximumLake2011}) using the random forest algorithm and the \texttt{ranger} R package \citep{wrightRangerFastImplementation2017} to predict observed maximum depth as a function of covariates including geometric maximum depth (from Equation \ref{eq1}) along with the waterbody elevation, area, perimeter, and ratio/index measures listed in Table \ref{table2}:
 
@@
 
-We evaluated the relative “importance” of individual covariates by comparing model performance between 1) models where a given covariate was left untouched versus 2) models where a given covariate was permuted randomly \citep{prasadNewerClassificationRegression2006, wrightRangerFastImplementation2017}.
+\noindent We evaluated the relative “importance” of individual covariates by comparing model performance between 1) models where a given covariate was left untouched versus 2) models where a given covariate was permuted randomly \citep{prasadNewerClassificationRegression2006, wrightRangerFastImplementation2017}.
 Neither cross-section shape nor reservoir class was used as a covariate in any random forest models.
@@
 
-The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to waterbody center (Table 2).
+The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to waterbody center (Table \ref{table2}).
 More specifically, the true slope (in-lake slope) - proxy distance (distance to the center of the waterbody) model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.8m, MAPE = 27%) compared to the proxy slope - true distance model ($R^2$ = 0.31, RMSE = 7.3m, MAPE = 64%).
@@
 We found that models restricted to consider only concave lakes had lower error (both in absolute and relative terms) compared to models on other data subsets (e.g.
-convex lakes, reservoirs, natural lakes, see Table 2).
+convex lakes, reservoirs, natural lakes, see Table \ref{table2}).
 Conversely, we did not observe any notable geographic patterns in model residuals (Fig.
@@
 ### Effects of proxy measures of waterbody geometry depth prediction error
-Models using only proxy variables (Table 2) had prediction error rates (RMSE = 7.1m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting waterbody depth at broad geographic extents \citep{hollisterPredictingMaximumLake2011, messagerEstimatingVolumeAge2016, oliver2016prediction}.
+Models using only proxy variables (Table \ref{table2}) had prediction error rates (RMSE = 7.1m) of a similar magnitude as that of prior studies (RMSE = 6 - 7.3m) predicting waterbody depth at broad geographic extents \citep{hollisterPredictingMaximumLake2011, messagerEstimatingVolumeAge2016, oliver2016prediction}.
 When only a single proxy measure was used, there was a difference in model sensitivity depending on if it was a horizontal distance measure or a vertical slope measure.

### 2021-11-10 figures/00_qa.Rmd
size alignment
rm stale code

@@
   # dplyr::filter(lagoslakeid == 5426)
-  summarize(# lake_maxdepth_std = sd(lake_maxdepth_m), 
+  summarize( # lake_maxdepth_std = sd(lake_maxdepth_m),
     lake_maxdepth_range = range(lake_maxdepth_m)[2] -
@@
 
-```{r lgnemanual-vs-bathy-depth, fig.height=2.5, fig.width=5.5}
+```{r lgnemanual-vs-bathy-depth, fig.height=2.2, fig.width=5}
 dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)
@@
   "LAGOS_NE_All_Lakes_4ha", "lagoslakeid", 61727))
-
 ```
-
-```{r misc, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
-threshold <- 5
-
-dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
-  dplyr::filter(!is.na(lake_maxdepth_m)) %>%
-  group_by(lagoslakeid) %>%
-  add_tally() %>%
-  dplyr::filter(n >= 2) %>%
-  mutate(diff = (max(lake_maxdepth_m) - min(lake_maxdepth_m)) / 
-           max(lake_maxdepth_m)) %>%
-  arrange(desc(diff))
-
-%>%
-  dplyr::filter(!(lagos_effort == "NLA" & diff > threshold)) %>%
-  dplyr::filter(!(lagos_effort == "LAGOSNE" & diff > threshold))
-
-dt %>% View()
-distinct(dt, lagoslakeid, diff) %>%
-  pull(diff) %>% hist(n = 100)
-
-mapview(LAGOSUSgis::query_lake_poly(365518))
-
-r <- raster("../data/nh_bathy/5787.tif")
-
-ggplot() + geom_blank() + 
-  theme(panel.background = element_blank())
-LAGOSNE::lake_info(c(5592, 5787)) %>%
-  dplyr::select(lagoslakeid, lagosname1, maxdepth, gnis_name, 
-                lake_area_ha) %>%
-  t() %>%
-  gridExtra::grid.table()
-
-hist(dt$diff, xlim = c(0, 10), n = 100)
-median(dt$diff, na.rm = TRUE)
-
-res_all <- read.csv("../data/00_bathy_depth/bathy_geometry.csv",
-                    stringsAsFactors = FALSE)
-dt_raw_ne <- read.csv("../data/lagosne_depth_predictors.csv",
-                      stringsAsFactors = FALSE) %>%
-  mutate(llid = as.character(lagoslakeid))
-test      <- left_join(res_all, dt_raw_ne) %>%
-  data.frame()
-
-test$calc_depth_lgne <- calc_depth(test$buffer100m_slope_max, test$dist_deepest,
-                                  grain = 10)
-test$calc_depth_lgus <- calc_depth(test$inlake_slope, test$dist_deepest)
-plot(test$maxdepth, test$calc_depth_lgus)
-# plot(test$lake_maxdepth_m, test$calc_depth_lgus)
-# abline(0, 1)
-#
-# ggplot(data = test) +
-#   geom_point(aes(x = lake_maxdepth_m, y = calc_depth_lgus)) +
-#   geom_abline(aes(slope = 1, intercept = 0)) +
-#   facet_wrap(~state)
-
-# plotly::ggplotly())
-
-#
-# hist(test$inlake_slope)
-# hist(test$dist_deepest)
-# hist(test$calc_depth_lgus)
-#
-# plot(test$calc_depth_lgne, test$calc_depth_lgus)
-# plot(test$lake_maxdepth_m, test$calc_depth_lgne)
-
-#
-# # 2654
-#
-#
-#
-# arrange(res, desc(dist_between)) %>%
-#   View()
-#
-# hist(res$dist_between)
-# plot(res$dist_deepest, res$dist_viscenter)
-# abline(0, 1)
-
-# pnt_surface  <- st_point_on_surface(dt_poly)
-## point furthest from shore
-# test <- as_Spatial(dt_poly)
-# r <- rasterize(SpatialPolygons(test@polygons), dt)
-# r[is.na(r)] <- 0
-# r[r == 1] <- NA
-# test2 <- distance(r)
-# which.max(test2)
-# xy           <- xyFromCell(dt, which.max(test2[]))
-# pnt_furthest  <- st_sfc(st_multipoint(xy), crs = st_crs(dt))
-# pnt_centroid <- st_centroid(dt_poly)
-
-# dt           <- raster("data/mn_bathy/2419.tif")
-# compare_deepest_center(dt)
-# mapview::mapview(dt_poly) +
-#   mapview::mapview(pnt_viscenter) +
-# mapview::mapview(r) +
-#   mapview::mapview(pnt_deepest)
-
-```

### 2021-11-10 scripts/00_get_geometry.R

@@ -15,7 +15,7 @@ lg <- lagosus_load("locus")
 #
 # get_geometry(raster(paste0("data/ct_bathy/", 7324, ".tif")), 7324, ft = 3.281,
 #  dt_poly = LAGOSUSgis::query_gis("LAGOS_US_All_Lakes_1ha", "lagoslakeid", 7324))
-get_geometry <- function(r, llid, deep_positive = TRUE, ft = 1, dt_poly){
+get_geometry <- function(r, llid, deep_positive = TRUE, ft = 1, dt_poly) {
   # llid <- 2645
   # r <- raster(paste0("data/mi_bathy/", llid, ".tif"))
   # deep_positive = TRUE
@@ -24,53 +24,53 @@ get_geometry <- function(r, llid, deep_positive = TRUE, ft = 1, dt_poly){
 
   proj_str_init <- st_crs(r)$proj4string
   proj_str      <- proj_str_init
-  if(nrow(st_coordinates(st_transform(dt_poly, proj_str))) >= 1){
+  if (nrow(st_coordinates(st_transform(dt_poly, proj_str))) >= 1) {
     dt_poly <- st_transform(dt_poly, proj_str)
-  }else{
+  } else {
     proj_str <- st_crs(dt_poly)
     r <- projectRaster(r, crs = proj_str$proj4string)
   }
-  if(is.na(sf::st_is_valid(dt_poly))){
+  if (is.na(sf::st_is_valid(dt_poly))) {
     dt_poly <- sf::st_make_valid(dt_poly)
   }
 
-  if(!st_is_simple(dt_poly) | (
-     st_area(dt_poly) > units::as_units(130000, "m2") &
-     nrow(st_coordinates(dt_poly)) > 32)){
+  if (!st_is_simple(dt_poly) | (
+    st_area(dt_poly) > units::as_units(130000, "m2") &
+      nrow(st_coordinates(dt_poly)) > 32)) {
     dt_poly_raw <- dt_poly
     dt_poly     <- dt_poly_raw %>%
       sf::st_make_valid() %>%
       rmapshaper::ms_simplify(0.1)
-    if(st_area(dt_poly_raw) > units::as_units(2700000, "m2") &
-       nrow(st_coordinates(dt_poly_raw)) > 32){
+    if (st_area(dt_poly_raw) > units::as_units(2700000, "m2") &
+      nrow(st_coordinates(dt_poly_raw)) > 32) {
[…]

### 2021-11-10 scripts/00_get_lagosne_xwalk.R

@@ -26,20 +26,3 @@ res <- lg$locus$locus_link %>%
 
 write.csv(res, "data/00_lagosne/00_lagosne_xwalk.csv",
   row.names = FALSE)
-# res <- read.csv("data/00_lagosne/00_lagosne_xwalk.csv", stringsAsFactors = FALSE)
-
-# plot(res$lake_waterarea_ha, res$lake_area_ha)
-
-# are there lakes that were combined (instead of split)?
-# test <- ungroup(res) %>% group_by(lagoslakeid) %>%
-#   add_tally() %>% arrange(desc(n))
-
-# are there lakes with a 1 to 1 relationship that have a different id?
-# test <- lg$locus$locus_link %>%
-#   distinct(lagoslakeid, .keep_all = TRUE) %>%
-#   dplyr::filter(!is.na(lagosne_lagoslakeid)) %>%
-#   group_by(lagosne_lagoslakeid) %>%
-#   add_tally() %>% arrange(desc(n)) %>%
-#   dplyr::filter(n == 1)
-#
-# which(test$lagoslakeid != test$lagosne_lagoslakeid)

### 2021-11-10 scripts/05_depth_model.R

@@ -281,428 +281,3 @@ saveRDS(depth_grid_metrics,
   "data/01_depth_model/depth_grid_metrics.rds")
 saveRDS(res[[17]]$dt_metrics,
   "data/01_depth_model/runner_up_metrics.rds")
-
-# if(interactive()){
-#
-# # ---- get bootstrap rmse estimates ----
-#
-# library(purrr)
-# fit_model_cv <- function(maxdepth, dt){
-#   # maxdepth <- maxdepth_vec[2]
-#
-#   pred_exclude <- c(maxdepth_vec[!(maxdepth_vec %in% maxdepth)], "inlake_slope", "dist_deepest")
-#   dt_rec       <- recipe(lake_maxdepth_m ~ ., data = dt) %>%
-#     step_normalize(all_numeric(), -all_outcomes(), -maxdepth) %>%
-#     step_log(all_outcomes(), maxdepth) %>%
-#     prep()
-#   dt_jc        <- juice(dt_rec)
-#   preds        <- names(dt_jc)[!(names(dt_jc) %in%
-#                                    c("lagoslakeid", "lake_maxdepth_m",
-#                                      pred_exclude))]
-#   dt_jc        <- dplyr::select(dt_jc, "lake_maxdepth_m",
-#                                 c(all_of(maxdepth), all_of(preds)))
-#
-#   # https://stateofther.github.io/finistR2019/s-tidymodels.html
-#   # lm_mod <-
-#   #   linear_reg(penalty = 0.001) %>%
-#   #   set_engine("glmnet")
-#
-#   lm_mod <- rand_forest(mode = "regression") %>%
-#     set_engine("ranger")
-#
-#   cv_fit <- function(splits, mod, ...){
-#     res_mod <-
-#       fit(mod, lake_maxdepth_m ~ ., data = analysis(splits))
-#     return(res_mod)
-#   }
-#
-#   cv_pred <- function(splits, mod){
-#     holdout     <- assessment(splits)
-#     pred_assess <- bind_cols(truth = holdout$lake_maxdepth_m,
-#                              predict(mod, new_data = holdout))
-#     return(pred_assess)
-#     }
-#
-#   cv_train <- vfold_cv(dt_jc, v = 10, repeats = 5, strata = "lake_maxdepth_m")
-#   res_cv_train <-
-#     cv_train %>%
-#     mutate(res_mod = map(splits, .f = cv_fit, lm_mod),
[…]

### 2021-11-10 scripts/05_hypso_model.R

@@ -22,7 +22,7 @@ dt_sub       <- dt %>%
   # dplyr::filter(lake_centroidstate %in% c("MN", "MI", "CT")) %>%
   dplyr::filter(!is.na(shape_class) & shape_class != "neither") %>%
   mutate(shape_class = droplevels(shape_class)) %>%
-  dplyr::select(# buffer100m_slope_max,
+  dplyr::select( # buffer100m_slope_max,
     shape_class, lake_elevation_m, lake_waterarea_ha, lake_islandarea_ha,
     lake_shorelinedevfactor_nounits, lake_connectivity_class,
     lake_glaciatedlatewisc, ws_area_ha, ws_lake_arearatio,
@@ -45,9 +45,9 @@ learner$train(class_task, row_ids = train_set)
 learner$importance()
 learner$predict(class_task, row_ids = test_set)$confusion
 learner$predict(class_task, row_ids = test_set)$score(measure)
-resampling = rsmp("cv", folds = 10L)
-rr = resample(class_task, learner, resampling, store_models = TRUE)
-pred = rr$prediction()
+resampling <- rsmp("cv", folds = 10L)
+rr <- resample(class_task, learner, resampling, store_models = TRUE)
+pred <- rr$prediction()
 pred$confusion
 pred$score(measure)
 
@@ -64,42 +64,17 @@ learner    <- lrn("classif.ranger")
 learner$train(class_task, row_ids = train_set)
 learner$predict(class_task, row_ids = test_set)$confusion
 learner$predict(class_task, row_ids = test_set)$score(measure)
-resampling = rsmp("cv", folds = 10L)
-rr = resample(class_task, learner, resampling, store_models = TRUE)
-pred = rr$prediction()
+resampling <- rsmp("cv", folds = 10L)
+rr <- resample(class_task, learner, resampling, store_models = TRUE)
+pred <- rr$prediction()
 pred$confusion
 pred$score(measure)
 
 fit_ranger <- list(
-  #importance = learner$importance(),
+  # importance = learner$importance(),
   confusion = pred$confusion,
   score = pred$score(measure))
 
 res[["ranger"]] <- fit_ranger
 
 saveRDS(res, "data/01_hypso_model/hypso_model.rds")
-
-## logistic model yes/no bowl-shaped lake
-# test <- glm(shape_class ~ lake_waterarea_ha + ws_mgbconhull_length_m,
[…]

### 2021-11-10 README.Rmd
hide docker, link ms

@@
 
-[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Data DOI](https://img.shields.io/badge/Data-10.6084/m9.figshare.12722246-blue.svg)](https://doi.org/10.6084/m9.figshare.12722246) [![Docker Build](https://img.shields.io/badge/Docker%20Image-jsta/bathymetry-green.svg)](https://cloud.docker.com/repository/docker/jsta/bathymetry)
+[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Data DOI](https://img.shields.io/badge/Data-10.6084/m9.figshare.12722246-blue.svg)](https://doi.org/10.6084/m9.figshare.12722246)
+
+<!--- [![Docker Build](https://img.shields.io/badge/Docker%20Image-jsta/bathymetry-green.svg)](https://cloud.docker.com/repository/docker/jsta/bathymetry) --->
 
@@
 ### Products
+
+Paper: [manuscript/manuscript.pdf](manuscript/manuscript.pdf)
 
@@
 
-```shell
-# docker
-
-```
+<!--- ```shell --->
+<!--- # TODO: docker --->
+<!--- ``` --->

### 2021-11-11 README.Rmd
mint code doi

@@
 
-[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Data DOI](https://img.shields.io/badge/Data-10.6084/m9.figshare.12722246-blue.svg)](https://doi.org/10.6084/m9.figshare.12722246)
+[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-10.5281/zenodo.5672711-blue.svg)](https://doi.org/10.5281/zenodo.5672711) [![Data DOI](https://img.shields.io/badge/Data-10.6084/m9.figshare.12722246-blue.svg)](https://doi.org/10.6084/m9.figshare.12722246)
 

### 2021-11-11 manuscript/manuscript.Rmd
mint code doi

@@
 
-Covariates used in random forest modeling (Table \ref{table1}, Equation 3, see Random Forests Model sub-section below) for waterbody elevation, area, island area, perimeter, shoreline development, watershed to waterbody area ratio, and hydrologic subbasin (i.e.
+Covariates used in random forest modeling (Table \ref{table1}, Equation \ref{eq3}, see Random Forests Model sub-section below) for waterbody elevation, area, island area, perimeter, shoreline development, watershed to waterbody area ratio, and hydrologic subbasin (i.e.
 HUC4s), were obtained from the LAGOS-US LOCUS database.
@@
 
-All data used in the study is available at \citet{stachelekjemmaBathymetryData0002021}. All code for data processing, model fitting, and model evaluation is available at [Zenodo DOI].
+All data used in the study is available at \citet{stachelekjemmaBathymetryData0002021}.
+All code for data processing, model fitting, and model evaluation is available at \citet{jemstachelekContlimnoBathymetryAccepted2021}.

### 2021-11-12 README.Rmd
add paper doi

--- 2021-11-10 README.Rmd
@@
 
-[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Data DOI](https://img.shields.io/badge/Data-10.6084/m9.figshare.12722246-blue.svg)](https://doi.org/10.6084/m9.figshare.12722246)
+[![Paper DOI](https://img.shields.io/badge/Paper-10.1080/20442041.2021.2006553-blue.svg)](https://doi.org/10.1080/20442041.2021.2006553) [![Code DOI](https://img.shields.io/badge/Code-10.5281/zenodo.5672711-blue.svg)](https://doi.org/10.5281/zenodo.5672711) [![Data DOI](https://img.shields.io/badge/Data-10.6084/m9.figshare.12722246-blue.svg)](https://doi.org/10.6084/m9.figshare.12722246)
 
@@
 
-**Stachelek J.**, P. J. Hanly, P. A., Soranno. Accepted. Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth. Inland Waters.
+**Stachelek J.**, P.
+J.
+Hanly, P.
+A., Soranno.
+Accepted.
+Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth.
+Inland Waters.
+doi:10.1080/20442041.2021.2006553
 

### 2021-12-06 manuscript/manuscript.Rmd
title fix
fight with acknowledgements
typo fix

@@
 ---
-title: Imperfect slope measurements drive overestimation in geometric cone model of waterbody depth
+title: Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth
 
@@
   \end{center}
-  \caption{Map of study lakes showing A) lake e maximum depth measurements, B) cross-section shape class, and C) reservoir classification.
+  \caption{Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.
 The distribution of lake depths from panel A is reported in Figure S4.}\label{fig2}

### 2021-12-06 figures/00_maps.Rmd
towards figure panel labels, #16

@@
   gg_reservoir + theme_opts,
+  labels = c("a", "b", "c")
   ncol = 1)

### 2021-12-06 manuscript/manuscript.Rmd
typo fix
add doi link

--- 2021-11-11 manuscript/manuscript.Rmd
@@
 date_subtitle: > 
-  This is an Accepted Manuscript of an article to be published by Taylor & Francis in Inland Waters.
+  This is an Accepted Manuscript of an article published by Taylor & Francis in Inland Waters, available online: \href{https://doi.org/10.1080/20442041.2021.2006553}{doi:10.1080/20442041.2021.2006553}
 

### 2021-12-06 figures/00_maps.Rmd
cleanup fig 2

--- 2021-08-16 figures/00_maps.Rmd
@@
     smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, max(max_depth)),
-      brackets = NULL, sep = " - ")
+      brackets = NULL, sep = "-")
   ) %>% # unique(dt$maxdepth_class)
@@
   gg_reservoir + theme_opts,
+  labels = c("a", "b", "c"),
+  label_x = 0.16,
   ncol = 1)

### 2021-12-07 manuscript/si_template_2019.tex
reorder appendix

@@
 
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
-
-% \affiliation{=number=}{=Affiliation Address=}
-%(repeat as many times as is necessary)
-
 %% ------------------------------------------------------------------------ %%
@@
 
-            \item \textbf{Figure S2:} Hypsography classification by state
+            \item \textbf{Figure S2:} Effect of calculation technique on in-lake/nearshore slope comparison.
 
-            \item \textbf{Figure S3:} Lake geometry comparison among shape and reservoir classes
+            \item \textbf{Figure S3:} Hypsography classification by state
 
-            \item \textbf{Figure S4:} Lake geometry comparison among shape classes by depth and area
+            \item \textbf{Figure S4:} Lake geometry comparison among shape and reservoir classes
 
-            \item \textbf{Figure S5:} Effect of calculation technique on in-lake/nearshore slope comparison.
+            \item \textbf{Figure S5:} Lake characteristics comparison among shape and reservoir classes
 
-            \item \textbf{Figure S6:} Spatial distribution of depth model residuals.
+            \item \textbf{Figure S6:} Proxy-proxy model fit showing predicted depth versus measured depth
 
-            \item \textbf{Figure S7:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
+            \item \textbf{Figure S7:} Spatial distribution of depth model residuals.
 
-            \item \textbf{Figure S8:} Lake characteristics comparison among shape and reservoir classes
+            \item \textbf{Figure S8:} Characteristics of lakes with bathymetry versus non-bathymetry lakes
 
-            \item \textbf{Figure S9:} Proxy-proxy model fit showing predicted depth versus measured depth
+            \item \textbf{Figure S9:} Comparison between reported depth and bathymetry depth
 
-            \item \textbf{Figure S10:} Comparison between reported depth and bathymetry depth
+            \item \textbf{Figure S10:} Lake geometry comparison among shape classes by depth and area
 
@@
 
-      % \noindent\textbf{Introduction}
-      %Type or paste your text here.
-The introduction gives a brief overview of the supporting information.
-You should include information %about as many of the following as possible (when appropriate):
-      % 1. a general overview of the kind of data files;
-      % 2. information about when and how the data were collected or created;
-      % 3. a general description of processing steps used;
-      % 4. any known imperfections or anomalies in the data.
-
-      %\clearpage
-
-      %Delete all unused file types below. Copy/paste for multiples of each file type as needed.
-      %\noindent\textbf{Text S1.}
-      %Type or paste text here.
-This should be additional explanatory text, such as: extended descriptions of results, full details of models, extended lists of acknowledgements etc.
-It should not be additional discussion, analysis, interpretation or critique.
-It should not be an additional scientific experiment or paper.
-      %
-      %Repeat for any additional Supporting Text
-
-      %%Enter Data Set, Movie, and Audio captions here
-      %%EXAMPLE CAPTIONS
-
-      % \noindent\textbf{Data Set S1.} %Type or paste caption here.
-      %upload your dataset(s) to AGU's journal submission site and select "Supporting Information (SI)" as the file type. Following naming %convention: ds01.
-
-      %Repeat for any additional Supporting data sets
-
-      % \noindent\textbf{Movie S1.} %Type or paste caption here.
-      %upload your movie(s) to AGU's journal submission site and select, "Supporting Information %(SI)" as the file type. Following naming convention: ms01.
-
-      %Repeat any additional Supporting movies
-
-      % \noindent\textbf{Audio S1.} %Type or paste caption here.
-      %upload your audio file(s) to AGU's journal submission site and select "Supporting Information %(SI)" as the file type. Following naming convention: auds01.
-
-      %Repeat for any additional Supporting audio files
-
-      %%% End of body of article:
-      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-      %
-      % Optional Notation section goes here
-      %
-      % Notation -- End each entry with a period.
-      % \begin{notation}
-      % Term & definition.\\
-      % Second term & second definition.\\
-      % \end{notation}
-      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-      %% ------------------------------------------------------------------------ %%
-      %%  REFERENCE LIST AND TEXT CITATIONS
-
-      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-      %
-      %
-      % \bibliography{lagosdepth}
-      %
-      % no need to specify bibliographystyle
-      %
-      % Note that ALL references in this supporting information file must also be referenced in the primary manuscript
-      %
-      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-      % if you get an error about newblock being undefined, uncomment this line:
-      %\newcommand{\newblock}{}
-
-      % \bibliography{ uncomment this line and enter the name of your bibtex file here }
-
-      %Reference citation instructions and examples:
-      %
-      % Please use ONLY \cite and \citeA for reference citations.
-      % \cite for parenthetical references
-      % ...as shown in recent studies (Simpson et al., 2019)
-      % \citeA for in-text citations
-      % ...Simpson et al (2019) have shown...
-      % DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
-      %
-      %
-      %...as shown by \citeA{jskilby}.
-      %...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
-      %...has been shown \cite<e.g.,>{jskilbye}.
-      %...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
-      %...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
-      %
-      % apacite uses < > for prenotes, not [ ]
-      % DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
-      %
-
-      %% ------------------------------------------------------------------------ %%
-      %
-      %  END ARTICLE
-      %
-      %% ------------------------------------------------------------------------ %%
 \end{article}
@@
 \begin{figure}
-      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_hypsography-1}
-      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figS2}
-\end{figure}
-
-\begin{figure}
-      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_geometry_grid-1}
-      \caption{Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.
-A dashed 1:1 line is shown for comparison.
-Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figS3}
-\end{figure}
-
-\begin{figure}
-      \begin{center}
-            \includegraphics[width=0.79\textwidth,keepaspectratio]{../figures/slope_by_area_depth-1}
-      \end{center}
-      \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth.
-Categories are quantile bins ($<$ 25\%, 25-50\%, 50-75\%, and 75-100\%).}\label{figS4}
-\end{figure}
-\clearpage
-
-\begin{figure}
       \begin{center}
@@
 \texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope).
-\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth.}\label{figS5}
+\texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth.}\label{figS2}
 \end{figure}
+
+\clearpage
+
+\begin{figure}
+      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_hypsography-1}
+      \caption{Hypsography classification by state. Numbers on panel labels indicate the percentage of lakes in each state with a convex versus a concave cross-section shape.}\label{figS3}
+\end{figure}
+
+\begin{figure}
+      \noindent\includegraphics[width=0.85\textwidth]{../figures/01_geometry_grid-1}
+      \caption{Comparison among lake shape and reservoir classes for A-B) distance to deepest point versus distance to lake visual center and C-D) nearshore slope versus inlake slope.
+A dashed 1:1 line is shown for comparison.
+Cross-section shape and reservoir class plots are not identical because not all lakes had a reservoir classification exceeding a 0.75 probability confidence level.}\label{figS4}
+\end{figure}
+
+\clearpage
+\begin{figure}
+      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
+      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS5}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_fitted_allproxy-1}\end{center}
+      \caption{Proxy-proxy model fit showing predicted depth versus measured depth.}\label{figS6}
+\end{figure}
+
 \clearpage
@@
       \end{center}
-      \caption{Spatial distribution of depth model residuals.}\label{figS6}
+      \caption{Spatial distribution of depth model residuals.}\label{figS7}
 \end{figure}
 \clearpage
-
-% \begin{figure}
-%       \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
-%       \caption{Importance plot for random forest variables showing increase in mean square error.
-Higher values indicate greater importance to model predictions.
-See Equation 1 for a definition of geometric max depth.
-HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figS7}
-% \end{figure}
 
@@
       \caption{Comparison between characteristics of lakes with bathymetry data against lakes with depth from other sources in the LAGOSUS-Depth product.
-The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS7}
+The distance to urban area metric is calculated using data from the 2018 US Census Urban and Rural Classification.}\label{figS8}
 \end{figure}
@@
 \begin{figure}
-      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/01_contrasts-1}\end{center}
-      \caption{Comparison of lake characteristics according to differences in lake cross-section shape or reservoir status.}\label{figS8}
-\end{figure}
-
-\clearpage
-\begin{figure}
-      \begin{center}\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_fitted_allproxy-1}\end{center}
-      \caption{Proxy-proxy model fit showing predicted depth versus measured depth.}\label{figS9}
+      \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
+      \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product.
+For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figS9}
 \end{figure}
@@
 \begin{figure}
-      \begin{center}\includegraphics[width=0.73\textwidth]{../figures/lgnemanual-vs-bathy-depth-1}\end{center}
-      \caption{Comparison between reported depth and depth extracted from bathymetry surfaces by US State where reported depths come from the LAGOSUS-Depth product.
-For this figure, no reported depth values originated from the same source as its corresponding bathymetry-derived value.}\label{figS10}
+      \begin{center}
+            \includegraphics[width=0.79\textwidth,keepaspectratio]{../figures/slope_by_area_depth-1}
+      \end{center}
+      \caption{Comparison between in-lake and nearshore slopes in concave and convex lakes of the same size and max depth.
+Categories are quantile bins ($<$ 25\%, 25-50\%, 50-75\%, and 75-100\%).}\label{figS10}
 \end{figure}
-
 \clearpage
-
-% % source at tables/01_predictors.Rmd
-% \begin{table}
-%       \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States).
-Predictor variables for computing random forest offsets (Equation 2) are printed in bold face.
-Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset.
-The total number lakes is reported as n.} \label{tableS1}
-%       \centering
-%       \setlength\tabcolsep{1.5pt} % default value: 6pt
-%       \begin{tabular}{lllll}
-%             \hline
-%             Variable                            & Median      & Q25         & Q75          & n            \\
-%             \hline
-%             Max depth (m)                       & 8.2 (7)     & 4.6 (3.7)   & 14 (12)      & 4820 (17700) \\
-%             \textbf{Area (ha)}                  & 55 (33)     & 21 (11)     & 140 (100)    & 4820 (17700) \\
-%             \textbf{Island area (ha)}           & 0 (0)       & 0 (0)       & 0.18 (0.076) & 4820 (17700) \\
-%             \textbf{Perimeter (m)}              & 4400 (3500) & 2500 (1800) & 8100 (7300)  & 4820 (17700) \\
-%             \textbf{Shoreline development}      & 1.7 (1.7)   & 1.4 (1.4)   & 2.1 (2.2)    & 4820 (17700) \\
-%             \textbf{Elevation (m)}              & 300 (340)   & 180 (210)   & 400 (460)    & 4820 (17700) \\
-%             \textbf{Watershed-lake ratio}       & 7.8 (10)    & 3.8 (4.4)   & 17 (29)      & 4820 (17700) \\
-%             \textbf{Deepest point distance (m)} & 180 (-)     & 110 (-)     & 290 (-)      & 4820 (-)     \\
-%             Mean deepest point distance (m)     & 140 (-)     & 87 (-)      & 230 (-)      & 4820 (-)     \\
-%             \textbf{Visual center distance (m)} & 240 (-)     & 160 (-)     & 390 (-)      & 4820 (-)     \\
-%             Inlake slope (m/m)                  & 0.05 (-)    & 0.02 (-)    & 0.08 (-)     & 4820 (-)     \\
-%             Inlake slope online (m/m)           & 0.06 (-)    & 0.03 (-)    & 0.14 (-)     & 4800 (-)     \\
-%             Inlake slopes (m/m)                 & 0.06 (-)    & 0.03 (-)    & 0.1 (-)      & 4820 (-)     \\
-%             Inlake slopes online (m/m)          & 0.07 (-)    & 0.03 (-)    & 0.15 (-)     & 4800 (-)     \\
-%             Mean inlake slope (m/m)             & 0.04 (-)    & 0.02 (-)    & 0.09 (-)     & 4820 (-)     \\
-%             Nearshore mean slope (m/m)          & 0.08 (-)    & 0.05 (-)    & 0.11 (-)     & 4820 (-)     \\
-%             Nearshore slope online (m/m)        & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4590 (-)     \\
-%             Nearshore slopes online (m/m)       & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4540 (-)     \\
-%             \hline
-%       \end{tabular}
-% \end{table}
 

### 2021-12-07 figures/01_geometry.Rmd
format fig 3

@@
   geom_abline(intercept = 0, slope = 1, linetype = 2, color = "darkgreen", size = 1) +
-  annotate("text", x = 2550, y = 430, label = paste0("R^2 == ", rsq), # higher y is up
+  annotate("text", x = 2550, y = 430, label = paste0("italic(R^2) == ", rsq), # higher y is up
     parse = TRUE, size = text_size - 0.7) +
@@
     alpha = 1, fill = "white") +
-  annotate("text", x = 0.34, y = 0.05, label = paste0("R^2 == ", rsq),
+  annotate("text", x = 0.34, y = 0.05, label = paste0("italic(R^2) == ", rsq),
     parse = TRUE, size = text_size - 0.7) +
@@
       gg_slope_base,
-      rel_widths = c(1, 0.9)
+      rel_widths = c(1, 0.9),
+      labels = c("a", "b"),
+      label_x = c(0.08, 0)
     ),

### 2021-12-07 figures/02_depth_model.Rmd
format fig 4

@@
 gg_shape      <- lapply(gg_shape, rm_legend)
-gg_shape[[2]] <- gg_shape[[2]] + theme(axis.text = element_blank(),
+gg_shape[[2]] <- gg_shape[[2]] + theme( # axis.text = element_blank(),
   axis.title = element_blank())
@@
     cowplot::plot_grid(gg_shape[[1]], gg_shape[[2]],
-      rel_widths = c(1, 0.85)),
+      rel_widths = c(0.92, 1)),
     cowplot::plot_grid(gg_shape[[3]], gg_shape[[4]],
@@
     leg,
-    nrow = 4, rel_heights = c(0.85, 1, 0.1, 0.15)),
+    nrow = 4, rel_heights = c(1, 1, 0.1, 0.15)),
 
@@
 gg_reservoir      <- lapply(gg_reservoir, rm_legend)
-gg_reservoir[[2]] <- gg_reservoir[[2]] + theme(axis.text = element_blank(),
+gg_reservoir[[2]] <- gg_reservoir[[2]] + theme( # axis.text = element_blank(),
   axis.title = element_blank())
@@
     cowplot::plot_grid(gg_reservoir[[1]], gg_reservoir[[2]],
-      rel_widths = c(1, 0.85)),
+      rel_widths = c(0.87, 1)),
     cowplot::plot_grid(gg_reservoir[[3]], gg_reservoir[[4]],
@@
     leg,
-    nrow = 4, rel_heights = c(0.85, 1, 0.1, 0.15)),
+    nrow = 4, rel_heights = c(1, 1, 0.1, 0.15)),
   ncol = 2, rel_widths = c(0.1, 1)) %>%
@@
 
-cowplot::plot_grid(gg_shape, gg_reservoir, ncol = 2, rel_widths = c(1, 0.9))
+cowplot::plot_grid(gg_shape, gg_reservoir, ncol = 2, rel_widths = c(1, 0.9), labels = c("a", "b"))
 ```

### 2021-12-07 figures/02_depth_model.Rmd
format fig 5

@@
     legend.key = element_blank()) +
-  ylab("") + xlab("Percent Increase in Mean Square Error")
+  ylab("") + xlab("Percent increase in mean square error")
 ```

### 2021-12-07 manuscript/manuscript.Rmd
fig4 caption

@@
   \end{center}
-  \caption{Map of study lakes showing A) lake e maximum depth measurements, B) cross-section shape class, and C) reservoir classification.
+  \caption{Map of study waterbodies showing A) maximum depth measurements, B) cross-section shape class, and C) reservoir classification.
 The distribution of lake depths from panel A is reported in Figure S4.}\label{fig2}
@@
   \end{center}
-  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig4}
+  \caption{Depth model residuals (residual  = observed - predicted) in meters by (a) cross-section shape and (b) reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig4}
 \end{figure*}

### 2021-12-07 manuscript/si_template_2019.tex
formatting toc for consistency

@@
 
-            \item \textbf{Figure S2:} Effect of calculation technique on in-lake/nearshore slope comparison.
+            \item \textbf{Figure S2:} Effect of calculation technique on in-lake/nearshore slope comparison
 
@@
 
-            \item \textbf{Figure S7:} Spatial distribution of depth model residuals.
+            \item \textbf{Figure S7:} Spatial distribution of depth model residuals
 
@@
 
-            \item \textbf{Table S1:} Effect of calculation technique on depth model fit and accuracy.
+            \item \textbf{Table S1:} Effect of calculation technique on depth model fit and accuracy
 

### 2021-12-07 manuscript/manuscript.Rmd
fix refs to supp material

@@
   \caption{Map of study waterbodies showing A) maximum depth measurements, B) cross-section shape class, and C) reservoir classification.
-The distribution of lake depths from panel A is reported in Figure S4.}\label{fig2}
+The distribution of lake depths from panel A is reported in Figure S10.}\label{fig2}
 \end{figure}
@@
 In addition to the aforementioned techniques of calculating in-lake (and nearshore) slopes and distances, we tried 7 alternate techniques which are described in Fig.
-S5 and Table S2 including measures such as median slope (results not shown).
+S2 and Table S1 including measures such as median slope (results not shown).
 
@@
 hypsographic curves) and assigning class membership based on whether the midpoint of a waterbody’s curve falls above or below that of a simple straight-sided cone (Fig.
-S2).
+S3).
 
@@
 The purpose of this standardization was to prevent waterbodies with extremely long proxy distances from having an outsized impact on model evaluation metrics.
-In addition to comparing among model runs using different metric combinations, we compared among sets of model runs where slope and distance measures were calculated using different sets of calculation techniques (Table S2).
+In addition to comparing among model runs using different metric combinations, we compared among sets of model runs where slope and distance measures were calculated using different sets of calculation techniques (Table S1).
 
@@
 For example, in-lake slope and distance to the deepest point of the waterbody metrics were consistently larger in magnitude for convex waterbodies as compared to concave waterbodies (Fig.
-S3).
+S4).
 We found evidence that this difference was at least partly explained by the fact that convex waterbodies are deeper than concave waterbodies (Fig.
-S8).
+S5).
 Unlike concave and convex waterbodies, there were not clear differences among slope and distance metrics for natural lakes versus reservoirs.
 
-Offset model fit and prediction error differed depending on the technique used to calculate in-lake and nearshore geometry metrics (Table S2).
+Offset model fit and prediction error differed depending on the technique used to calculate in-lake and nearshore geometry metrics (Table S1).
 We found that the best model fit and lowest model error occurred when in-lake slope was calculated as the average point-wise slope of all points at maximum waterbody depth rather than at single point of maximum depth.
@@
 Predicted depth values for this model were generally underestimates relative to measured depth values (Fig.
-S9).
+S6).
 
@@
 Conversely, we did not observe any notable geographic patterns in model residuals (Fig.
-S6).
+S7).
 
@@
 Although our analysis was limited to waterbodies with available bathymetry data, these waterbodies did not have characteristics that differed from that of the overall waterbody population (apart from the fact that our study waterbodies were somewhat larger in area compared to the overall waterbody population, See Fig.
-S7, Table \ref{table1}, Fig.
-S10).
+S8, Table \ref{table1}, Fig.
+S9).
 Although there is a possibility that there is some hidden bias not explored for in our analyses, this lack of difference suggests that our results are likely to be broadly applicable to nearly all waterbodies in the study area.
@@
 Furthermore, the fact that in-lake slopes were shallower compared to the surrounding land even after controlling for differences in area (Fig.
-S4) is consistent with the idea of topographic scaling (i.e.
+S10) is consistent with the idea of topographic scaling (i.e.
 scale invariance) explored in previous work and detailed by \citep{caelVolumeMeanDepth2017}.
@@
 Finally, it was notable that convex waterbodies were deeper than concave waterbodies despite having similar distributions of waterbody surface area (Fig.
-S8).
+S5).
 The underlying cause of these differences is unknown but one possibility is that geometry is tied to the circumstances of waterbody formation whereby the formation of concave waterbodies were a result of more intense glacial scouring compared to that of convex waterbodies \citep{gorhamPhysicalLimnologyNorthern1958}.
@@
 The greater sensitivity of depth predictions to proxy slope measures relative to proxy distance measures may be explained by the fact that proxy slope measures were a more imperfect representation of true in-lake slopes relative to proxy versus true distances.
-We did not find evidence that the sensitivity of depth predictions to slope was dependent on variations in how these measures were calculated (Table S2).
+We did not find evidence that the sensitivity of depth predictions to slope was dependent on variations in how these measures were calculated (Table S1).
 In a general sense, the sensitivity of depth predictions to slope help explain the relatively poor predictive performance of prior non-geometric waterbody depth models given that they rely heavily on waterbody area as a predictor \citep{sobekPredictingDepthVolume2011, messagerEstimatingVolumeAge2016, oliver2016prediction} and both horizontal distance measures and vertical slope measures appear to be decoupled from waterbody area (Fig.
-S8).
+S5).
 

### 2023-04-04 README.Rmd
full citation
hyphen fix
try manual meta tags

@@
 output: github_document
+header-includes:
+  - <meta name="citation_author" content="Stachelek, Jemma">
+  - <meta name="citation_publication_date" content="2022">
+  - <meta name="citation_title" content="Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth">
+  - <meta name="citation_pdf_url" content="https://github.com/cont-limno/bathymetry/raw/master/manuscript/manuscript.pdf">
 ---
@@
 
-**Stachelek J.**, P.
-J.
-Hanly, P.
-A., Soranno.
-Accepted.
-Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth.
-Inland Waters.
-doi:10.1080/20442041.2021.2006553
+**Stachelek, J.**, Hanly, P.J., Soranno, P.A., 2022. Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth. Inland Waters, 12(2), 283-293.
+10.1080/20442041.2021.2006553
 
